### 2022-08-03 text/paper.tex
First commit.

% Page settings
\documentclass[notitlepage,12pt]{article}
\usepackage{fancyvrb, verbatim, listings}                                  % Various for inserting/hosting code
\usepackage{setspace}                                                      % Space between lines (customisable)
\usepackage[margin=2.54cm]{geometry}                                       % Set margins (standard is 2.54cm)
\usepackage[UKenglish,cleanlook]{isodate}                                  % Set default date and date display
\usepackage{fancyhdr} \pagestyle{fancy}                                    % Line the top of the page
\usepackage[activate={true,nocompatibility},final,tracking=true,kerning=true,spacing=true,factor=1100,stretch=10,shrink=10]{microtype}
\microtypecontext{spacing=nonfrench}                                       % Change font and minor spacings to look nicer
\setlength{\parskip}{0.5\baselineskip}                                     % Set default paragraph behaviour: skip a line
%\setlength{\parindent}{0pt}                                                % Set default paragraph behaviour: no indent
%% Packages to use:
% Tables + figures
\usepackage{graphicx}                                                      % control over the import of graphics
\usepackage[table,dvipsnames]{xcolor}                                      % Define tables (with colour control)
\usepackage{float}                                                         % Control over graphics + tables float
\usepackage[justify]{ragged2e}                                             % control over text alignment
\usepackage{booktabs}                                                      % caption graphics + tables (with name in bold)
\usepackage[labelfont=bf]{caption} 
\usepackage{subcaption}
% caption graphics + tables (with name in bold)
\usepackage{tikz}                                                          % Draw graphs inline, guide: https://sites.google.com/site/kochiuyu/Tikz
\usetikzlibrary{shapes,decorations,arrows,calc,arrows.meta,fit,positioning}
\tikzset{
    -Latex,auto,node distance =1 cm and 1 cm,semithick,
    state/.style ={ellipse, draw, minimum width = 0.7 cm},
    point/.style = {circle, draw, inner sep=0.04cm,fill,node contents={}},
    bidirected/.style={Latex-Latex,dashed},
    el/.style = {inner sep=2pt, align=left, sloped}
}
\usepackage{pgfplots} \pgfplotsset{compat=1.16}                            % Automatic graphing, read https://www.overleaf.com/learn/latex/pgfplots_package for examples
\usepackage{subcaption}                                                    % Sub figures.
% Maths + numbers
\usepackage{mathtools}                                                     % Various maths functions
\usepackage{amssymb}                                                       % Various maths functions
\usepackage{amsmath}                                                       % Various maths functions
\usepackage{dsfont}                                                        % Various maths functions
\usepackage{centernot}                                                     % center \not usage
\usepackage{siunitx} \sisetup{round-mode=places, round-precision=3}        % Formalise use of units and numbers among text
\DeclareMathOperator{\eps}{\varepsilon}                                    % epsilon short hand shortcut
\DeclareMathOperator{\st}{\text{ s.t. }}                                   % "such that" short hand shortcut
\DeclareMathOperator{\then}{\text{ then }}                                 % "then" in equation shortcut
\DeclareMathOperator{\ifeq}{\text{ if }}                                   % "if" in equation shortcut
\DeclareMathOperator{\oreq}{\text{ or }}                                   % "or" in equation shortcut
\DeclareMathOperator{\andeq}{\text{ and }}                                 % "and" in equation shortcut
\DeclareMathOperator{\all}{,\; \forall}                                    % all (with spacing) in equation shortcut
\DeclareMathOperator{\N}{\mathbb{N}}                                       % N, natural number, shortcut
\DeclareMathOperator{\R}{\mathbb{R}}                                       % R, real number, shortcut
\DeclareMathOperator{\Ll}{\mathcal{L}}                                     % L, Lagrangian, shortcut
\renewcommand{\vec}[1]{\boldsymbol{\mathit{#1}}}                           % vector notation shortcut
\newcommand{\mat}[1]{\boldsymbol{\mathit{#1}}}                             % matrix notation shortcut
\DeclarePairedDelimiter\abs{\lvert}{\rvert}                                % absolute value notation shortcut
\DeclarePairedDelimiter\norm{\lVert}{\rVert}                               % norm notation shortcut
\newcommand{\Prob}[1]{\Pr\left( #1 \right)}                         % SHortcut for probability notation
\newcommand{\Probgiven}[2]{\Pr\left( #1 \, \middle\vert \, #2 \right)} % SHortcut for probability notation, given
\newcommand{\E}[2][]{\mathbb{E}_{#1} \left[ #2 \right]}                    % Expectation (with optional subscript) shortcut
\newcommand{\Egiven}[3][]{\mathbb{E}_{#1} \left[ #2 \, \middle\vert \, #3 \right]} % Expectation given (with optional subscript) shortcut
\newcommand{\Var}[2][]{\text{Var}_{#1} \left( #2 \right)}                  % Variation (with optional subscript) shortcut
\newcommand{\Cov}[1]{\text{Cov} \left( #1 \right)}                         % Covariance (with optional subscript) shortcut
\newcommand{\median}[1]{\text{median} \left( #1 \right)}                   % Median (with optional subscript) shortcut
\newcommand{\indicator}[1]{\mathds{1}\left\{ #1 \right\}}                  % SHortcut for indicator function
\newcommand{\diff}[2][]{\frac{d#1}{d#2}}                                   % SHortcut for differential fraction as a function
\newcommand{\partialdiff}[2][]{\frac{\partial#1}{\partial#2}}              % SHortcut for partial differential fraction as a function
\newcommand{\converge}[1]{\xrightarrow{ #1 \to\infty}}                     % SHortcut for convergence arrow
\renewcommand{\hat}[1]{\widehat{#1}}                                       % Default estimator notation is widehat
\renewcommand{\bar}[1]{\overline{#1}}                                      % Make over bar look nicer
\renewcommand{\tilde}[1]{\widetilde{#1}}                                   % Make over tilde look better
% Citations
\usepackage[longnamesfirst]{natbib}                                        % Citation package, see https://en.wikibooks.org/wiki/LaTeX/Bibliography_Management#Natbib
\usepackage[backref=page]{hyperref}                                        % Allow for links across the text, with colour options
\hypersetup{colorlinks=true, linkcolor=blue, citecolor=blue, filecolor=magenta, urlcolor=blue}
\def\equationautorefname~#1\null{Equation~(#1)\null}                       % Fix autoref for quations
% IMPORTANT: follow style guide here https://github.com/Wookai/paper-tips-and-tricks

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Title page
% Author
\author{Senan Hogan-Hennessy\thanks{
    Cornell University, Department of Economics.}
    \thanks{I thank Professor Evan Riehl for guidance on this project, Michael Lovenheim, Ronald Ehrenberg, Douglas Miller, and Francine Blau for various comments on the project's direction,
    as well as seminar participants at Cornell University (Spring 2022) for helpful
    discussion.
    This project's Github repository, with materials and (publicly available) data for replication, is available at 
    \url{https://github.com/shoganhennessy/finances-faculty-composition}.
    Any comments or suggestions may be sent to me at \href{mailto:seh325@cornell.edu}{\nolinkurl{seh325@cornell.edu}}, or raised as an issue on the Github project.}}
% Title
\title{Stagnating Public University Finances and Faculty Composition}
\date{\today}
\rhead{\today}
\lhead{Stagnating Public University Finances and Faculty Composition}
% Begin
\begin{document}
\clearpage \maketitle
\thispagestyle{empty}
% Abstract
\begin{abstract}
\noindent
\textbf{This needs re-doing}

When a worker's employer is exposed to demand shocks we may expect the resulting effect on the worker to depend on multiple economic factors, including size of the shock and the worker's outside options.
This paper studies the topic for a unique labour market: that of public university professors in the US.
State spending for public universities has secularly declined the last three decades, yet the isolated effects on professors are not yet clear.
Using an instrumental variables strategy to isolate shocks in public university revenues, I show that universities respond to negative revenue shocks by restricting employment among tenure-track and tenured professors -- possibly due to hiring and retirement channels, respectively.
Results using individual level data regarding every public university professor in Illinois are to follow.

\vspace{0.75cm}
\noindent\textbf{Keywords:} Public Sector Labor Markets, Higher Education, State and Local Budget and Expenditures

\vspace{0.5cm}
\noindent\textbf{JEL Codes:} J45, I23, H72
\end{abstract}
\newpage
\setcounter{page}{1}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Starting the paper
\newpage
\doublespacing

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Plan
\subsection{Plan}
\begin{enumerate}
    \item Document empirical trend about number of professors per students, possibly diverging between private and public unis
    \item Document trend of falling revenues per student in the public sector -- while the private sector is fine
    \item Pull these facts together with shift-share IV approach to attribute the trend to falling state support for PUs.
    \item Document the effect on individuals in Illinois, including local projections for the dynamic effects.
    \item Theory discussion on the goals of a public university: public university charters require a minimum provision of teaching that make teaching a constraint, and other activities (including research).
    Implies exmploying fewer instructors per student, or substitution towards cheaper instructors who focus exclusively on teaching. 
\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Introduction section
\section{Introduction}
\noindent
Public universities educate the majority of higher education students in the US, yet have experienced a secular decline in state funding (per student) the last few decades.
This decline has been shown to lead to worse education and later-life outcomes for students \citep{NBERw23736,NBERw27885}, yet it is not clear what institutional effects or mechanisms driving effects on students.
I show that four-year, degree-granting public universities (PUs) are exposed to tightening financial constraints thanks to falling state financial support (per student), and show this effects the composition of professors at PUs.
Falls in PU revenues, instrumented by state-level finance shocks, lead to a fall in the number of professors per student within a PU, driven by a fall in assistant or tenured professor counts.
US private universities are not exposed to similar financial constraints during this same time period, and do not exhibit any similar changes in the faculty composition, so that stagnating financial support for PUs have implications for the structure of higher education instruction and research in the US.

Universities in the US are widely considered the highest performing in the world,yet there are consequential differences between its universities that operate in the private sector and those established by state governments.
PUs are subject to numerous state-level administrative laws, and rely on their state governments for funding: an average PU received around \$11,600 per enrolled student in 1990, yet only \$8,300 per enrolled student in 2020.\footnote{
    2020 figures are relatively similar to those for year preceding, despite the Covid-19 pnademic and associated financial shocks, since 2020 refers to the academic year 2019-2020 and so the yearly budget was decided in 2019.
    This paper does not directly analyse budget shocks related to the Covid-19 pandemic, or any other events in 2020 and following.
}
This fall is driven by stagnating funding provided by state governments, while total enrollment among PUs rose by 46\% over the same time period.
At the same time, the number of professors per student at private universities stayed relatively stable at 0.045, yet fell from around 0.04 to around 0.35 professors per student at PUs, driven by a fall of over 20\% in the ratio of associate and full (i.e. tenured) professors per student at PUs.\footnote{
    \autoref{fig:fte-perprof} and \autoref{sec:trends} document these trends between private and public universities in full detail.
}

I use a variant of the \cite{NBERw23736,NBERw27885} shift-share instrument for PU revenues from shocks to state appropriations to identify the effect on the faculty composition.
State appropriations for public universities are plausibly endogenous to many outcomes: state governments decide on yearly budgets for their higher education sector, and this process can be influenced by local financial conditions, or even public (and political representatives') perceptions of the state's higher education system.
The shift-share instrument interacts reliance on state funding in a base period with the yearly total appropriations per student in the entire state;
the instrument for yearly state appropriations addresses the endogeneity concerns if state government funding decisions are not associated with any outcomes for any one university in the state \citep{NBERw27885}.
\autoref{sec:approp-shocks} provides further evidence in favour of the exogeneity assumption of the instrument.

Briefly describe results.

Then write about the Illinois data as individual-level data, and empirical approach when it is fully specified.

\cite{NBERw23736} first study the effect of (changes in) PUs' finances on student outcomes by isolating changes in PUs' expenditures on student instruction, addressing endogeneity of expenditures decisions by instrumenting expenditures by an instrument for state appropriations and tuition caps.
\cite{NBERw23736} find that increases in PU spending (via state appropriations) increases enrolment and degree completion among students, but corresponding changes in attendance price (via tuition caps/freezes) do not have an effect.
\cite{chakrabarti2018effect,NBERw27885} use the same instrument for PU revenues at the state level, combined with richer information on students' individual-level outcomes, to show that increases in state appropriations lead to lower time to completion and debt levels among students later in life.
\cite{bound2019public} use another variant of the same approach to show that state-level higher education spending cuts induced PUs to shift toward tuition as their primary source of revenue, and shift institutional resources towards students who pay higher tuition rates.
In a similar vein, \cite{bound2007cohort} document variation in per student state funding for higher education resulting from yearly changes in a state's student-age population; they find that lower per student funding lead to lower higher education completion rate among students.

University professors the core component of this country's educational and research/innovation university system, yet their composition within universities and its relationship with higher education finances has so far not been studied.
\cite{brown2014endowment} presents the closest example by studying how university endowments react to negative financial shocks, exhibiting an association between negative endowment shocks and a fall in tenure-system faculty employed at the university.
\cite{abe2015implications} present a model for university administrator decision-making, and posit that a large component of variation in faculty salaries arises from university politics, and not only economic factors such as outside.
\cite{johnson2009jep,NBERc13879} both empirically document allocation of faculty between research and instruction, and \cite{hemelt2021math} document differences in instructional cost (per student) between departments, primarily thanks to department differences in class size and faculty salaries.
\cite{turner2014impact} document how universities reacted to the Great Recession of 2008, a severe financial shock to state finances, including implementing hiring freezes.
Indeed, Cornell University\footnote{
    Cornell University is, indeed, only partially public, so that this information is not free information, and cannot be presented with aggregate or verified sources of public data on such salaries.
} implemented both hiring freezes and even a nominal salary reduction for professors, in anticipation of oncoming financial shock in early 2020.\footnote{
    The salary cut was not permanent, as the oncoming financial shock turned out to not be as serious as projected, and the salary cut was returned to professors in the year 2021.
}

The number of professors, in absolute terms as well as relative to number of students in the university, may be particularly important for quality of instruction and thus educational outcomes.
\cite{angrist1999using} first causally show that reducing class size induces an increase in test scores for school-age children, yet the magnitude of the effect is not as clear in the university setting.
\cite{bandiera2010heterogeneous} use a fixed effect model, combined with multiple individual observations in a long panel, to find that UK university students perform worse academically in particularly large classes, and the effect is particularly largest for students at the top of the academic distribution.
And for the composition of faculty teaching university classes, 
\cite{bettinger2010does,figlio2015tenure} find US students have better enrollment and learning outcomes from courses taught by contingent or adjunct (i.e. not tenured or on tenure-track) professors (relative to tenured or tenure-track).
While \cite{ehrenberg2005tenured} observe a negative association between utilisation of non-tenured instructors and student graduation rates.
My approach focuses on count of professors per student at universities, so does not directly observe individual class-sizes, yet presents this as a possible mechanism through which the negative effects on student outcomes that falls in PU finance \cite{NBERw23736,NBERw27885} document.

\textbf{Add in an intro conclusion paragraph.}
``Furthermore, because schools that serve students from lower-income backgrounds are most affected by state appropriations cuts, reductions in state support have helped exacerbate inequality and stratification of outcomes in the postsecondary sector.'' \citep{NBERw27885}.

%\begin{table}[H]
%    \centering
%    \caption{Table to Consider Impact of a Given Paper.}
%    \begin{tabular}{l | c c c c c}
%        & concepts & relationships & models & theories & use of methods \\
%        \midrule
%        
%        change:
%        & concepts & relationships & models & theories & use of methods \\
%        
%        challenge:
%        & concepts & relationships & models & theories & use of methods \\
%        
%        fundamentally alter:
%        & concepts & relationships & models & theories & use of methods \\
%    \end{tabular}
%\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Data section
\section{Data and Institutional Context}
\label{sec:data}

\subsection{Data Description}
The data used in this analysis come from two primary sources: \citet[IPEDS]{ipeds} regarding institutional information on finances and enrollment, and \citet[IBHED]{ibhed} for a panel of individual-level of every Professor employed by the Illinois PU system.

IPEDS is a survey of higher educational institutions in the US, and legally requires institutions to participate in order to receive Federal Title IV student aid.\footnote{
    This statement means that IPEDS does not necessarily cover the universe of US higher education institutions, yet thin practice every PU and not for-profit four-year institution is represented.
}
Data are consistent between the years 1990 and 2020,\footnote{
    The years 1987-1989 are represented in these data in an incompletion fashion, so I focus on the years 1990 onwards.
    Year refers to the calendar year of the spring term --- i.e. 1987 refers to the academic year that ran August 1986 to July 1987.
}
and provide information on the total revenues a university received from every source (including state governments), enrollment number, plus faculty count and total expenditures on salaries.\footnote{
    I combine the Urban Institute's 2018 compilation of IPEDS data for the years 1990-2017, and manually combine with raw NCES data regarding the year 2018-2020 for all relevant variables.
    Figures for enrollment, faculty counts and salaries come from the raw NCES version of IPEDS for all years to address inconsistencies in the Urban Institute's data formulation for these variables.
}
I restrict analysis to four-year institutions, as these institutions adhere to a standardised concept of faculty profile -- that is, tenure and title of appointment (lecturer, assistant professor, etc.) are relatively standardised in four-year, degree-granting institutions.
Additionally, for-profit institutions employ and enrol a negligible share of professors and students respectively, while students at two-year institutions by majority intend to eventually enroll at a four-year institution \citep{mountjoy2022}.\footnote{
    Furthermore, \cite{mountjoy2022} documents the effects of expanding two-year higher education access in the US separately for students who would have and those who would not have otherwise have attended college.
    The high rate of eventual enrollment at four-year institutions among two-year students explains the negative impact of increased access to two-year institutions on students who are diverted away from four-year institutions.
}

Importantly, IPEDS reports the count of professors employed by category, as well as total salary expenditures by category.
This gives a resulting panel data-set, where each row represents a university-year, and includes columns for university finances, plus total count and average salary\footnote{
    Real salary is computed by scaling nominal salary to 2021 dollars by the CPI-U.
} for professors by category (lecturer, assistant, tenured, total).
\autoref{tab:ipeds-summary} presents summary statistics for relevant variables in the panel of PU-years.

\begin{table}[h!]
    \onehalfspacing
    \centering
    \caption{IPEDS Summary Statistics, PUs Panel 1987--2020}
    \makebox[\textwidth][c]{\input{tables/ipeds-summary-fte.tex}}
    \label{tab:ipeds-summary}
\end{table}

IPEDS provides information at the university level, yet lacks information on the distribution of salaries and professor count within a university.\footnote{
    IPEDS provides total paid to salary and total faculty employed per university-year, so that I can investigate average professor salary per university-year with IPEDS data.
    Unfortunately, it seems that this measure of professors' salaries is particularly crude in measuring on professors' salaries and other individual-outcomes; summary statistics on IPEDS data do not agree with trends in average professor salary over the sample time period \citep{aau2021survey}.
}
To investigate the distribution, I integregate individual-level data for every PU professor in the state of Illinois between the years 2010-2020.
IBHED host the information;
Public Act 96-0266 (effective 1 January 2010) requires that each PU report to the IBHE the base salary and benefits of the president and all administrators, faculty members, and instructors employed by the college or university.\footnote{
    The universities included are all campuses of the eight Illinois public universities: Chicago State University, Eastern Illinois University, Governors State University, Illinois State University, Northeastern Illinois University, Northern Illinois University, Southern Illinois University  (all five campuses), University of Illinois (all four campuses), Western Illinois University.
}
This publicly available data provides the basis to build a panel of Illinois PU professors 2010-2020; I define a professor as an individual observation by their first plus last name and university pairing, and link this database to IPEDS information on finances for their employing institution.
\begin{table}[h!]
    \onehalfspacing
    \centering
    \caption{IBHED Summary Statistics, Professor Panel 2011--2020.}
    \makebox[\textwidth][c]{\input{tables/illinois-summary.tex}}
    \label{tab:illinois-summary}
\end{table}
The analysis sample focuses on the subset of professors with identified year of hiring between the years 2011 and 2020, representing 1,778 professors in 2011, and 9,099 in 2020.
\autoref{tab:illinois-summary} presents the summary statistics for this sample.\footnote{
    The full data IBHED contain 173,000 observations for every public university professor in Illinois 2010-2020, which representing 16,932 professors in the year 2010 and 15,352 in the year 2020.
    The reasons for restricting to the sample with identified year of hire in the range 2011-2020 is explained in \autoref{sec:iv-model-indiv}.
}

\subsection{Trends in Finances, Enrollment, and Faculty Composition}
\label{sec:trends}
A state government in the US plans an annual budget a couple of years ahead of the fiscal year, with the legislature approving a budget request put forth by the governor's office.\footnote{
    \cite{NBERw23736} present a full discussion of the state appropriation decision-making process, drawing on administrative records originally analysed by \cite{parmley2009state}.}
This process leads to yearly variation in state appropriations not seen in other revenues sources (such as federal government appropriations), since US states differ in their rate of (financial) support for higher education, and are subject to fiscal constraints.
US states governments, by majority, are legally obligated to run a balanced-budget, so that yearly variation in tax revenues (caused by changing economic conditions or otherwise) necessarily affect state expenditures.
State appropriations to higher education are a particularly attractive area of state spending to absorb such shocks to state finances \citep{delaney2011state}.\footnote{
    \cite{delaney2011state} fully describes the financial environment of state expenditures, and what makes spending on higher education an attractive area for state governments to expand funding during years of higher tax revenues, and retract funding in leaner years.
    An analysis of state expenditures for the years 1980-2004 (overlapping with the sample for this analysis) provides solid evidence for these trends.
}
Additionally, there is yearly state variation in the number of higher education students \citep{turner2014impact}, so that per-student state appropriations vary on multiple dimensions.

\autoref{fig:funding} exhibits the trends for the mean of US PUs between the years 1990-2020.
We see a rise in total revenues received by PUs (from all sources), and a notable increase in mean tuition revenues from \$48 million per year-university to \$150 million per year-university. 
Non-institutional revenues refer to revenues coming from external sources, not raised by internal university activities;\footnote{
    US universities raise revenues from many different internal activities, including sales of non-degree related educational courses to publishing deals.
    The non-institutional revenue measure is constructed to consider only revenues resulting from external decisions, including state, local, and federal appropriations and tuition revenues.
    Rates of tuition area considered external since most state governments play a large role in deciding tuition rices for PUs \citep{NBERw23736}.
}
non-institutional revenues rose over this time period almost exclusively thanks to the rise in tuition revenues.
At the same time, total state appropriations stagnated at around \$100 million per year-university for 1990-2008, fell slightly around 2008 and have not recovered.
While PUs experienced a financial stagnation, private universities were not exposed to the same constraints, receiving \$37,000 per student in 1990 and \$49,000 in 2020 with no corresponding decline in any specific component.

\begin{figure}[h!]
    \centering
    \caption{Mean Total Revenues among Public Universities, by Year.}
    \begin{subfigure}[b]{0.495\textwidth}
        \centering
        \caption{Total, millions \$ 2021 CPI-U.}
        \includegraphics[width=\textwidth]{figures/mean-funding-total.png}
        \label{fig:mean-funding-total}
    \end{subfigure}
    \begin{subfigure}[b]{0.495\textwidth}
        \centering
        \caption{Per Enrolled Student, \$ 2021 CPI-U.}
        \includegraphics[width=\textwidth]{figures/mean-funding-fte.png}
        \label{fig:mean-funding-fte}
    \end{subfigure}
    \label{fig:funding}
\end{figure}

At the same time, student enrollment at PUs rose precipitously.
6.2 million students were enrolled in PUs in 1990, and this number rose by 47\% to 9.1 million, with most of the increase occurring over the years 2000-2020.
\begin{figure}[h!]
    \centering
    \caption{Total Student Enrollment, by University Sector, and Year.}
    \begin{subfigure}[b]{0.495\textwidth}
        \centering
        \caption{Total.}
        \includegraphics[width=\textwidth]{figures/enrollment-total.png}
        \label{fig:enrollment-total}
    \end{subfigure}
    \begin{subfigure}[b]{0.495\textwidth}
        \centering
        \caption{Mean, University Level.}
        \includegraphics[width=\textwidth]{figures/enrollment-mean.png}
        \label{fig:enrollment-mean}
    \end{subfigure}
    \label{fig:enrollment}
\end{figure}

\autoref{fig:enrollment} shows that total enrollment at private universities has also risen over the same time period, but not as drastic in either relative or absolute terms; the mean public university also grew over this time period, from 9,800 students in 1990 to 11,800 in 2020.
This means that per student mean revenues per student (seen in \autoref{fig:mean-funding-fte}) have stagnated across all measures, and particularly fallen from \$11,000 per student in 1990 to less than \$8,000 per student in 2020.
\begin{figure}[!h]
    \centering
    \caption{Total Professor Count per Student, by University Sector, Professor Appointment, and Year.}
    \begin{subfigure}[b]{0.495\textwidth}
        \centering
        \caption{Lecturer}
        \includegraphics[width=\textwidth]{figures/lecturer-fte-perprof.png}
        \label{fig:lecturer-fte-perprof}
    \end{subfigure}
    \begin{subfigure}[b]{0.495\textwidth}
        \centering
        \caption{Assistant.}
        \includegraphics[width=\textwidth]{figures/assistant-fte-perprof.png}
        \label{fig:assistant-fte-perprof}
    \end{subfigure}
    \begin{subfigure}[b]{0.495\textwidth}
        \centering
        \caption{Full.}
        \includegraphics[width=\textwidth]{figures/full-fte-perprof.png}
        \label{fig:full-fte-perprof}
    \end{subfigure}
    \begin{subfigure}[b]{0.495\textwidth}
        \centering
        \caption{All.}
        \includegraphics[width=\textwidth]{figures/all-fte-perprof.png}
        \label{fig:all-fte-perprof}
    \end{subfigure}
    \label{fig:fte-perprof}
\end{figure}

\autoref{fig:fte-perprof} documents the divergence in faculty composition (per student) between the average private and public university 1990-2020.
Private universities start with  a higher baseline of around 4.5 professors per 100 students, and exhibits yearly variation of less than 0.5 professor per 100 students over the thirty years.
PUs start with 3.9 professors per student, and this number falls to 3.5 primarily in the 2008-2011 time period.
We see a similar difference in baseline, and fall for the years 2008-2011 for assistant professors.
Private and public universities have similar numbers of associated and full professors before thea year 2000, yet this number has fallen by over 20\% in the next 20 years only for PUs: in 2020 the mean PU has 6 fewer full professors per tudent than the mean private university.
At the same time period, we see the rise in use of non-tenure track instructor positions (referred to as lecturers from here on), who were employed at similar rates in both sectors in 1990 yet have been utilised by PUs at a higher rate since.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Empirical section
\section{Empirical Framework}
\label{sec:empirics}

A professor is employed by a university, so it stands to reason that when PUs experience a secular fall in revenue --- analogous to a demand shock in the private sector --- that their human resources decisions will be affected.
Universities have been shown to freeze faculty hiring and general faculty salaries in response to financial shocks, such as the Great Recession \citep{turner2014impact}.

Na\"ively we can analyse the relationship between professor-outcomes $Y_{i,t}$ at university $i$ in year $t$ as a result of university revenues per student $X_{i,t}$.
Professor outcomes include count of professors per student, and average salary, within each university-year.
This model gives $\beta$ the association with these outcomes and university revenues per student,\footnote{
    Note that dividing by student count also implicitly controls for the size of the university, so that this model accounts for yearly variation in professor count and university revenues arising from growth in a university.
}
including fixed effects to control for effects specific to the university and year.
Log$(.)$ transforming the variables improves the interpretability of $\beta$ as an elasticity for professor count per student with respect to institution revenues.
\begin{equation}
    \label{eqn:naivereg}
    Y_{i,t} = \alpha_i + \gamma_t + \beta X_{i,t} + \epsilon_{i,t}
\end{equation}

And yet a university's finances are not exogenous to decisions made by the university administrator (such as hiring), state outcomes or enrollment.
Instead, the state government and university administration undertake a complex process of alloting resources across multiple different priorities, including instruction, research, or between departments.
Importantly for this analysis, revenues received from a  PU's state government provide opportunity to address this endogeneity.

\subsection{State Appropriation Shocks}
\label{sec:approp-shocks}

\cite{NBERw23736,chakrabarti2018effect,NBERw27885} address endogeneity in PU finances by exploiting a shift-share instrument for changes in state-level funding interacted with university reliance on state funding in a base period.
The system exploits the fact that institutions who rely on state appropriations more will be affected by later shocks to state appropriations.
We start with the shift-share instrument.
\begin{align}
    \label{eqn:public-instrument}
    Z_{i,t} &\coloneqq \log \left[
    \left( \frac{\text{Total State Funding}_{s(i),t}}{\text{Student Population}_{s(i),t}} \right)
    \sum_{\tau = 0}^{3} \frac 14
    \left( \frac{\text{State Funding}_{i,1990 + \tau}}{\text{Total Revenues}_{i,1990 + \tau}} \right) \right]
\end{align}

$Z_{i,t}$ is the instrument for (log) state appropriations for institution $i$ in year $t$, interacting the average funding for university $i$ in state $s(i)$ with reliance on state funding relative to total revenues, averaged across the base years 1990--1993.\footnote{
    1990--1993 are defined as data for PU finance data are most comparable (i.e. without many missing values) beginning in 1990.
    \cite{NBERw23736} use the single year 1990 as the base year, though I use the four years to ameliorate missing values in the single year of 1990.
    Results are similar in either specification.
}
\cite{NBERw27885} notes the tendency for PUs to respond to state funding cuts by increasing reliance on tuition, where \cite{NBERw23736} specifically instruments for tuition revenues with collected information on legislative tuition price controls.
As such, it is important in my set-up to explicitly control for tuition revenues in the first-stage.

The first-stage is then as follows, where $W_{i,t}$ represents (log) tuition revenues, including institution and year fixed effects.
\begin{equation}
    \label{eqn:firststage}
    X_{i,t} = \eta_i + \zeta_t + \delta Z_{i,t} + \kappa W_{i,t} + \epsilon_{i,t}
\end{equation}
We note the conditions for exogeneity in the instrument (following the discussion presented by \citealt{NBERw27885}).
The instrument is exogenous if state policy decisions for funding of, or tuition charges by PUs, are uncorrelated with unobserved changes in the expenditures of any specific college or university in the state.
This assumption is plausible given that the majority of states have multiple (i.e. more than five) public universities, without any single university campus receiving the majority of state funding within any single state.
Secondly, addressing endogeneity in university revenues by the shift-share identification strategy requires exogeneity in either the base-line share or shift component of the instrument.
In this case, we satisfy the second: universities' institutional-level decisions are not correlated with contemporaneous or upcoming shocks to state appropriations.\footnote{
    It would be plausible to consider the case that universities make institutional-level decisions in a consistently different manner to those 
    with differing reliance on state appropriations in 1990, so that exogeneity by the base-line share is not plausible here.
}

\autoref{tab:firststage-reg} presents results of the first-stage regression, separately with and without the tuition control and fixed effects, at the institution level (i.e. using IPEDS data).\footnote{
    Representations for frequentist significance levels (i.e. 10, 5, 1\% etc.) are omitted here, and in all following tables.
}
\begin{table}[!h]
    \onehalfspacing
    \centering
    \caption{First Stage Estimates, for Total University Revenues by Shocks}
    \makebox[\textwidth][c]{\input{tables/firststage-reg-fte.tex}}
    \label{tab:firststage-reg}
    \begin{flushleft}
        \footnotesize
        \textbf{Note}: Standard errors are clustered at the state-year level.
    \end{flushleft}
\end{table}
We note columns (1) and (2) estimate that a shock to state appropriations per student in the state of 10\% is associated with around 2.5\% change in non-institutional revenues per student at the university; the instrument is strong, and we note similarity in estimates with and without inclusion of fixed effects.
Columns (3) and (4) exclude the tuition revenue control (explained above) to exhibit bias in the first-stage if it is not controlled for.
Column (3) shows similar estimates to columns (1), (2) thanks to inclusion of fixed effects, so that the fixed effects was effective in soaking variation in per-student tuition revenues at the institution-year level.
Column (4) exhibits the worst case, and extreme bias in the appropriation shock first-stage without the tuition control.
Column (1) represents the estimates for \autoref{eqn:firststage} with fixed effects, and is the preferred form that I proceed with in the following.

\autoref{fig:firststage-lp} presents local projection estimates for the staying-power of the state appropriation shock to university revenues.
We see the time 0 estimate coincides with that of Column (1) in \autoref{tab:firststage-reg}, the contemporaneous effect of the shock revenues.
We see a decaying, yet real and positive, effect of the shock on revenues for the next 5 years, illustrating the staying power of the effect.

\begin{figure}[h!]
    \centering
    \caption{Local Projection Estimates for First-Stage \autoref{eqn:firststage}.}
    \includegraphics[width=0.6\textwidth]{figures/firststage-lp.png}
    \label{fig:firststage-lp}
\end{figure}

These results, together with the case for exogeneity, show the appropriations shock instrument strongly predicts university revenues in the first-stage estimation.

\subsection{Instrumental Variables Model, University-Level}
\label{sec:iv-model-uni}

The primary empirical model combines the instrumental variable for stata appropriation shocks with the empirical model for the effects of university revenues at the university-level -- i.e. parameter $\beta$ in the following.\footnote{
    It is important to note the treatment effect isolated here; the instrumental variables approach identifies the local average treatment effects (LATE) specific to the instrument.
    So we interpret this treatment effect as the PU's response in employment count and average salaries to revenue changes, changes specific to state appropriation shocks, among the complier group --- i.e. universities with any exposure to shocks, assuming no universities increase employment or salaries in response to revenue shocks.
    \cite{mogstad2021causal} provide a full discussion of the LATE with multiple instruments and thus multiple monotonicity conditions.
}
\begin{eqnarray}
    \label{eqn:secondstage1}
    X_{i,t} &=& \eta_i + \zeta_t + \delta Z_{i,t} + \kappa W_{i,t} + \epsilon_{i,t} \\
    \label{eqn:secondstage2}
    Y_{i,t} &=& \alpha_i + \gamma_t + \beta \widehat X_{i,t} + \lambda W_{i,t} + \varepsilon_{i,t}
\end{eqnarray}
I estimate the system by two stage least squares, including institution and year fixed effects, and investigate outcomes at the university-level to analyse effects on the university as a result of changes in revenues.
Additionally, I estimate the model via local projections \citep{jorda2005,miller2022} to investigate whether the effects of revenues on faculty composition linger for multiple years after the original appropriation shock.
Regarding outcomes, I focus on the composition of the professors employed at the university by analysing (log) count per student, and average salaries paid to, professors employed by the university.

\begin{figure}[H]
    \centering
    \caption{Instrument Variables Model for University Finances and Faculty Composition.}
    \begin{tikzpicture}
        \node[state] (instrument) at (0,0) {$Z$};
        \node at (-2.5,0.5){State};
        \node at (-2.0,0){appropriation};
        \node at (-2.5,-0.5){shock};
        \node[state] (confounder) [above=of instrument] {$W$};
        \node at (-2,2.25) {Tutuion};
        \node at (-2,1.75) {revenues};
        \node[state] (endogenous) [right=of instrument] {$X$};
        \node at (2,1.5) {University};
        \node at (2,1) {revenues};
        \node[state] (outcome) [right=of endogenous] {$Y$};
        \node at (5.5,0.25) {Professor};
        \node at (5.55,-0.25) {per student,};
        \node at (5.5,-0.75) {Average};
        \node at (5.55,-1.25) {Salary};
        % Relevance condition
        \path (instrument) edge (endogenous);
        \path (endogenous) edge (outcome);
        % Confounding relationship
        \path (instrument) edge (confounder);
        \path (confounder) edge (endogenous);
        % Exclusion Restriction
        \path (instrument) edge[bend right=60] (outcome);
        \node[color=red] at (2,-1.2) {X};
    \end{tikzpicture}
    \label{fig:SCM-ivmodel}
\end{figure}

\autoref{fig:SCM-ivmodel} presents the system in graphical format, where tuition revenues $W$ are considered a confounder in the causal effect of state appropriation shocks on total revenues.

\subsection{Instrumental Variables Model, Individual Professor-Level}
\label{sec:iv-model-indiv}

This analysis additionally utilises data regarding individual professors in the Illinois university system, to investigate the effects of changes in university revenues on the individual professors at the universities.
Yet, redefining the level of outcome requires adjustment to the empirical approach to leverage variation in university finances in the years following an individual joins the university.\footnote{
    This formulation follows that presented by \cite{NBERw27885}, where individual student outcomes are analysed via variation in state appropriations after their freshman-year.
    This constrasts with \autoref{sec:iv-model-uni} and \cite{NBERw23736}, where the unit of analysis is the university-year.
}

\autoref{eqn:public-instrument} defines a rolling-share variant of the instrument, $\tilde Z_{j,t}$, where the university's share exposure to state appropriations is based the year a professor joins the university --- and not the base period 1990--1993.
$j$ indexes each professor in year $t$, $\tau(j)$ for the year the professor first joins their Illinois PU.\footnote{
    Identifying $\tau(j)$ is possible for $j$ by restricting to all professors who are hired in the time period 2011-2020 (i.e. in the years after the first in the panel).
    Professors who were hired in the year 2010 are impossible to distinguish from those in year preceding, and so are not considered.
}
\begin{align}
    \label{eqn:rolling-instrument}
    \tilde Z_{j,t} &\coloneqq \log \left[
    \left( \frac{\text{Total State Funding}_{s(j),t}}{\text{Student Population}_{s(j),t}} \right)
    \left( \frac{\text{State Funding}_{\tau(j)}}{\text{Total Revenues}_{i,\tau(j)}} \right) \right]
\end{align}

This approach leverages an insight, made available by level of the data: that an individual professor is affected by changes in university revenues after they have joined the university.\footnote{
    Notice that \autoref{sec:iv-model-uni} considers a the number of professors employed by the university.
    So that whether a professor becomes employed at the university is likely affected by the university revenues.
    The formulation in this section does not consider that outcome, instead taking as given that the professor is employed at the university, and then projecting the effect on the individual given they are employed here.
}
Exogeneity and relevance of the rolling-share instrument, $\tilde Z_{j,t}$, follows the same reasoning as that for the base-share instrument, $Z_{i,t}$, discussed in \autoref{sec:approp-shocks}.
We satisfy the assumptions by noting that none of the Illinois public campuses take the majority of state appropriations, and that the instrument identification strategy relies on exogeneity in changes in state appropriations to individual professor-outcomes, following the year they joined the university.\footnote{
    Additionally, within-institution changes resulting from share reliance on state funding may be correlated with unobserved changes in the outcomes, so that \cite{NBERw27885} note the importance of controlling for the base share and state student population.
    My formulation implicits controls for these factors via the fixed effects; results are relatively similar while including these controls without including fixed effects.
}
\autoref{tab:firststage-illinois} presents results of the first stage estimation, showing that the instrument is strong (after conditioning on tuition revenues) in the same way as that for the university-level outcomes (\autoref{tab:firststage-reg}), with very similar estimates for the association between appropriation shocks and non-institutional revenues.

The full model is defined as follows where $i(j)$ refers to the institution that professor $j$ is employed at, and $Y_{j,t}$ refers to individual-level outcomes total salary, rate of promotion, and propensity to leave the Illinois PU system.
The system includes a fixed effect for the institution and first year of employment.\footnote{
    The instrument varies by institution, based in the year of first employment so that these are the corresponding fixed effects and levels of clustered standard errors.
}
\begin{eqnarray}
    \label{eqn:secondstage1_indiv}
    X_{i(j),t} &=& \theta_{i(j)} + \phi_{\tau(j)} + \delta \tilde Z_{i(j),t} + \kappa W_{i(j),t} + \epsilon_{i(j),t} \\
    \label{eqn:secondstage2_indiv}
    Y_{j,t} &=& \mu_{i(j)} + \nu_{\tau(j)} + \beta \widehat X_{i(j),t} + \lambda W_{i(j),t} + \varepsilon_{i(j),t}
\end{eqnarray}
We then interpret parameter $\beta$ in this formulation as the effect of changes in university revenues, via state appropriation shocks, on an individual professor.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Results section
\section{Results}
\label{sec:results}

\subsection{University-Level}

\autoref{tab:facultycount-shock-reg} presents OLS and IV estimates, where the outcome is (log) count of professors employed at the university, separated by group and combined (columns 7, 8).
Within each group we see a positive correlation between university revenues and professor count per student (the OLS columns), where an increase in university total revenues by 10\% is associated with an increase in professor employment count between 3.5-4.5\%.
Lecturers per student is not shown to be correlated with total univesity revenues.
The 2SLS columns show results after identifying university revenues with the appropriation shock.
All professor count per student increases by 3\% in response to a 10\% rise in total revenues, mediated by revenue shocks.
The effect is driven by increases in the count of assistant and full professors (i.e. tenure-track and tenured) per student, where assistant professor count per student increase by 6.7\%, and full professor 5\%.
The count of lecturers per student decreases by 12\% in response to a (positive) 10\% shock in university revenues.

\begin{table}[!h]
    \onehalfspacing
    \centering
    \caption{OLS and 2SLS Estimates for University Faculty Composition.}
    \makebox[\textwidth][c]{\input{tables/facultycount-shock-reg-fte.tex}}
    \begin{flushleft}
        \footnotesize
        \textbf{Note}: Standard errors are clustered at the state-year level.
    \end{flushleft}
    \label{tab:facultycount-shock-reg}
\end{table}

DISCUSS WHAT THOSE POINT ESTIMATES MEAN, DISCUSSING THE POSSIBILITY THAT LECTURERS ARE INFERIOR GOODS, TENURE TRACK NORMAL GOODS.

This finding is particularly in line with the \cite{turner2014impact} observation regarding hiring freezes particularly for tenure-track hirings, and \cite{brown2014endowment} from shocks to endowments.
Similarly, the effect is exhibited on order of 2.6\% among full professors (those holding tenure), likely through the channel of retirements or replacement hirings (since tenured professors cannot be forcibly fired/not renewed without cause).

Write something about the level of tenure, and that data are available for a subset of years, and shown in Appendix \autoref{tab:tenurecount-shock-reg-fte}.

\autoref{fig:all-count-lp} shows local projection estimates, staying power of the effect on prof count.

\begin{figure}[h!]
    \centering
    \caption{Local Projection Estimates for Professor Count per Student.}
    \includegraphics[width=0.6\textwidth]{figures/all-count-lp.png}
    \label{fig:all-count-lp}
\end{figure}

WRITE HERE ABOUT STAYING POWER.

\subsection{Individual Professor Outcomes}

Estimates for the model where units of analysis are individual professors at Illinois PUs follow.

\begin{table}[!h]
    \onehalfspacing
    \centering
    \caption{2SLS Estimates for Faculty Salaries at Illinois Universities.}
    \makebox[\textwidth][c]{\input{tables/facultysalaries-shock-illinois.tex}}
    \begin{flushleft}
        \footnotesize
        \textbf{Note}: Standard errors are clustered at the institution-year level.
    \end{flushleft}
    \label{tab:facultysalaries-shock-illinois}
\end{table}

\begin{table}[!h]
    \onehalfspacing
    \centering
    \caption{2SLS Estimates for Faculty Salaries, in First-Year, at Illinois Universities.}
    \makebox[\textwidth][c]{\input{tables/newhiresalaries-shock-illinois.tex}}
    \begin{flushleft}
        \footnotesize
        \textbf{Note}: Standard errors are clustered at the institution and first year of employment level.
        Included fixed effects are institution and first-year of employment, since every year-institution interaction over-identifies the system. 
    \end{flushleft}
    \label{tab:newhiresalaries-shock-illinois}
\end{table}

\autoref{tab:promotion-shock-illinois} describes estimates for promotion rate: particularly sharp for the move associate to full.

\begin{table}[!h]
    \onehalfspacing
    \centering
    \caption{2SLS Estimates for Faculty Promotion Rate at Illinois Universities.}
    \makebox[\textwidth][c]{\input{tables/promotion-shock-illinois.tex}}
    \begin{flushleft}
        \footnotesize
        \textbf{Note}: Standard errors are clustered at the institution and first year of employment level.
    \end{flushleft}
    \label{tab:promotion-shock-illinois}
\end{table}

\begin{figure}[h!]
    \centering
    \caption{Local Projection Estimates for Promotion Rate at Illinois Universities.}
    \includegraphics[width=0.6\textwidth]{figures/promoted-illinois-lp.png}
    \label{fig:promoted-illinois-lp}
\end{figure}

\autoref{tab:facultyleaving-shock-illinois} describes estimates for exit rate: not much going on.

\newpage
\subsection{Robustness Checks}

This section will exhibit systematic robustness checks for each of the specifications estimated in \autoref{sec:results}.
In particular, re-estimation of the models scaled by number of students within each university can imply a possible mechanism for the worsening student outcomes in response to negative PU finance shocks shown by \cite{NBERw23736,NBERw27885}.
Similarly to graph the trajectory of finance shocks among PUs: PUs show a secular decline in (per student) revenues from state sources.

Lastly, \cite{blandhol2022tsls} provide reason to doubt the LATE interpretation of two-stage least squares estimates with multiple layers of fixed effects (i.e. both individual and time fixed effects).
Further analysis of their methods, and how it applies to the model presented in \autoref{sec:empirics} is planned.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Conclusion section
\section{Summary and Concluding Remarks}
\label{sec:conclusion}

\textbf{REVISE THIS PARAGRAPH ENTIRELY}

This report has presented the set-up for, and the preliminary results of, a project investigating the relationship between PUs' finances and the professors they employ.
Results show that employment count at the university level is affected by (shocks to) total revenues, concentrated among assistant and full professors while the count of lecturers is unaffected.
Salaries at the university-level are unaffected.

\textbf{ADD A DISCUSSION PARAGRAPH OR TWO TAKING CRITICISM IN ADVANCE}

I present a plan to extend the empirical analysis to individual data for all PU professors in the state of Illinois 2010-2019 --- the first such use of these data in an academic article.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Bibliography section
\bibliographystyle{agsm}
\bibliography{bibliography}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Appendix section
\appendix
\section{Appendix}
\label{appendix}
A number of statistical packages, for the R language \citep{R2022}, made the empirical analysis for this paper possible.
\begin{itemize}
    \item \textit{Tidyverse} \citep{tidyverse} collected tools for data analysis in the R language.
    \item \textit{LFE} \citep{lfe} implemented linear fixed effect models, with instruments, crucial for the empirical estimation in \autoref{sec:empirics}.
    \item \textit{Stargazer} \citep{stargazer} provided a method to efficiently convert empirical results into presentable output in \LaTeX.
    \item \textit{Lpirfs} \citep{lpirfs2019} implemented estimation of the \cite{jorda2005} local projections methods, with instrumental variables, crucial to the local projections estimates presented in this project.
\end{itemize}

\subsection{First Stage Estimates, Individual Outcomes}
\label{appendix:part1}
\begin{table}[H]
    \onehalfspacing
    \centering
    \caption{First Stage Estimates, for Total University Revenues at the Individual-Level}
    \makebox[\textwidth][c]{\input{tables/firststage-illinois.tex}}
    \label{tab:firststage-illinois}
    \begin{flushleft}
        \footnotesize
        \textbf{Note}: Standard errors are clustered at the university-year level.
    \end{flushleft}
\end{table}

\begin{figure}[h!]
    \centering
    \caption{Local Projection Estimates for First-Stage \autoref{eqn:secondstage1_indiv}.}
    \includegraphics[width=0.6\textwidth]{figures/firststage-illinois-lp.png}
    \label{fig:firststage-illinois-lp}
\end{figure}

\subsection{Additional Results, University-Level}

\autoref{tab:facultysalaries-shock-reg} presents OLS and IV estimates where the outcome is (log) mean salary for professors employed at the university, separated by group.
Similarly, \autoref{fig:all-salaries-lp} shows the staying-power via projection methods.
The outcome, mean salary is crude.

\begin{table}[H]
    \onehalfspacing
    \centering
    \caption{OLS and 2SLS Estimates for University Faculty-Tenure Composition.}
    \makebox[\textwidth][c]{\input{tables/tenurecount-shock-reg-fte}}
    \begin{flushleft}
        \footnotesize
        \textbf{Note}: Standard errors are clustered at the university-year level.
    \end{flushleft}
    \label{tab:tenurecount-shock-reg-fte}
\end{table}

\begin{table}[H]
    \onehalfspacing
    \centering
    \caption{OLS and 2SLS Estimates for University Faculty Salaries.}
    \makebox[\textwidth][c]{\input{tables/facultysalaries-shock-reg-fte.tex}}
    \begin{flushleft}
        \footnotesize
        \textbf{Note}: Standard errors are clustered at the state-year level.
    \end{flushleft}
    \label{tab:facultysalaries-shock-reg}
\end{table}

\begin{figure}[h!]
    \centering
    \caption{Local Projection Estimates for Professor Average Salary, within Each University.}
    \includegraphics[width=0.6\textwidth]{figures/all-salaries-lp.png}
    \label{fig:all-salaries-lp}
\end{figure}

\subsection{Additional Results, Individual-Professor Level}

\begin{table}[H]
    \onehalfspacing
    \centering
    \caption{2SLS Estimates for Faculty Exit Rate at Illinois Universities.}
    \makebox[\textwidth][c]{\input{tables/facultyleaving-shock-illinois.tex}}
    \begin{flushleft}
        \footnotesize
        \textbf{Note}: Standard errors are clustered at the institution and first year of employment level.
    \end{flushleft}
    \label{tab:facultyleaving-shock-illinois}
\end{table}

\end{document}

### 2022-08-03 text/tables/tenurecount-shock-reg-fte.tex
First commit.

\begin{tabular}{@{\extracolsep{5pt}}lcccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{8}{c}{Dependent Variable: Employment Count by Tenure Group} \\ 
\cline{2-9} 
\\[-1.8ex] & \multicolumn{2}{c}{Non-tenure} & \multicolumn{2}{c}{Tenure-Track} & \multicolumn{2}{c}{Tenured} & \multicolumn{2}{c}{All} \\ 
 & OLS & 2SLS & OLS & 2SLS & OLS & 2SLS & OLS & 2SLS \\ 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6) & (7) & (8)\\ 
\hline \\[-1.8ex] 
 Appropriations Shock & $-$0.039 & 0.590 & 0.227 & 1.108 & 0.206 & 0.227 & 0.174 & 0.454 \\ 
  & (0.084) & (0.826) & (0.162) & (0.829) & (0.096) & (0.200) & (0.095) & (0.265) \\ 
  Tuition Revenue & 0.069 & $-$0.166 & 0.088 & $-$0.229 & 0.071 & 0.063 & 0.027 & $-$0.074 \\ 
  & (0.109) & (0.370) & (0.109) & (0.355) & (0.038) & (0.106) & (0.054) & (0.158) \\ 
 \hline \\[-1.8ex] 
Observations & 4,825 & 4,825 & 5,094 & 5,094 & 5,130 & 5,130 & 5,181 & 5,181 \\ 
R$^{2}$ & 0.849 & 0.843 & 0.780 & 0.743 & 0.902 & 0.902 & 0.935 & 0.928 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}

### 2022-08-04 text/paper.tex
rough first draft.

@@
 \noindent
-\textbf{This needs re-doing}
-
-When a worker's employer is exposed to demand shocks we may expect the resulting effect on the worker to depend on multiple economic factors, including size of the shock and the worker's outside options.
-This paper studies the topic for a unique labour market: that of public university professors in the US.
-State spending for public universities has secularly declined the last three decades, yet the isolated effects on professors are not yet clear.
-Using an instrumental variables strategy to isolate shocks in public university revenues, I show that universities respond to negative revenue shocks by restricting employment among tenure-track and tenured professors -- possibly due to hiring and retirement channels, respectively.
-Results using individual level data regarding every public university professor in Illinois are to follow.
+State support for public universities has stagnated the last thirty years, while enrollment has secularly increased.
+At the same time, public universities have reduced the number of professors per student on tenure-track or with tenure, while increasing their employment of contingent lecturers.
+This analysis relates changes in the faculty composition at public universities to the changes in per student university revenues from state appropriations, addressing endogeneity in state appropriation decisions by a shift-share instrumental variables approach.
+An increase of university revenues of 10\%, via a change in state appropriations, increases the number of assistant professors per student by 6.6\% and full professors by 5.3\%, yet decreases the number of lecturers per student by 14.3\%.
+Analysis of all the professors hired at Illinois public universities 2011-2021 implies that incumbant professors are not affected by the changes in university revenues, so that changes in faculty composition arises by changes in hiring pratices at public universities.
+These results show the compositional effects of stagnating state suport for higher education, and raise questions about the direction that public education is heading thanks to these financial headwinds.
 
 \vspace{0.75cm}
-\noindent\textbf{Keywords:} Public Sector Labor Markets, Higher Education, State and Local Budget and Expenditures
+\noindent\textbf{Keywords:}
+State and Local Budget and Expenditures,
+Higher Education,
+Public Sector Labor Markets
 
 \vspace{0.5cm}
-\noindent\textbf{JEL Codes:} J45, I23, H72
+\noindent\textbf{JEL Codes:} H72, I23, J45
 \end{abstract}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Plan
-\subsection{Plan}
-\begin{enumerate}
-    \item Document empirical trend about number of professors per students, possibly diverging between private and public unis
-    \item Document trend of falling revenues per student in the public sector -- while the private sector is fine
-    \item Pull these facts together with shift-share IV approach to attribute the trend to falling state support for PUs.
-    \item Document the effect on individuals in Illinois, including local projections for the dynamic effects.
-    \item Theory discussion on the goals of a public university: public university charters require a minimum provision of teaching that make teaching a constraint, and other activities (including research).
-    Implies exmploying fewer instructors per student, or substitution towards cheaper instructors who focus exclusively on teaching. 
-\end{enumerate}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %% Introduction section
-\section{Introduction}
 \noindent
@@
 Universities in the US are widely considered the highest performing in the world,yet there are consequential differences between its universities that operate in the private sector and those established by state governments.
-PUs are subject to numerous state-level administrative laws, and rely on their state governments for funding: an average PU received around \$11,600 per enrolled student in 1990, yet only \$8,300 per enrolled student in 2020.\footnote{
-    2020 figures are relatively similar to those for year preceding, despite the Covid-19 pnademic and associated financial shocks, since 2020 refers to the academic year 2019-2020 and so the yearly budget was decided in 2019.
-    This paper does not directly analyse budget shocks related to the Covid-19 pandemic, or any other events in 2020 and following.
+PUs are subject to numerous state-level administrative laws, and rely on their state governments for funding: an average PU received around \$11,600 per enrolled student in 1990, yet only \$8,300 per enrolled student in 2021.\footnote{
+    2021 figures are relatively similar to those for year preceding, despite the Covid-19 pnademic and associated financial shocks, since 2021 refers to the academic year 2019-2021 and so the yearly budget was decided in 2019.
+    This paper does not directly analyse budget shocks related to the Covid-19 pandemic, or any other events in 2021 and following.
 }
@@
 State appropriations for public universities are plausibly endogenous to many outcomes: state governments decide on yearly budgets for their higher education sector, and this process can be influenced by local financial conditions, or even public (and political representatives') perceptions of the state's higher education system.
-The shift-share instrument interacts reliance on state funding in a base period with the yearly total appropriations per student in the entire state;
-the instrument for yearly state appropriations addresses the endogeneity concerns if state government funding decisions are not associated with any outcomes for any one university in the state \citep{NBERw27885}.
-\autoref{sec:approp-shocks} provides further evidence in favour of the exogeneity assumption of the instrument.
+The shift-share instrument interacts reliance on state funding in a base period with the yearly total appropriations per student in the entire state, to exploit both changes in state support for higher education and how much each university relies on that support.
 
-Briefly describe results.
-
-Then write about the Illinois data as individual-level data, and empirical approach when it is fully specified.
+Findings show that negative changes in university revenues affects faculty composition at public universities, away from tenured and tenure-track professors towards non-tenure track lecturer positions.
+A reduction of 10\% in university revenues, via a shock to state appropriations, leads to a fall in 6.6\% in the number of assistant professors per student within a university, 5.3\% for full (tenured) professors, and a reduction in 14.3\% for lecturers.
+Additionally, I use individual level data on professors who were hired at Illinois public universities 2011-2021 to investigate whether the shocks to university revenues affected incumbant professors at public universities.
+Incumbant professors were not meaningfully affected, in terms of total salary, promotion rate, or rate of leaving the Illinois public university system, which implies that changes in faculty composition came about by changes in the composition of hiring new professors at the university level.
 
@@
     Cornell University is, indeed, only partially public, so that this information is not free information, and cannot be presented with aggregate or verified sources of public data on such salaries.
-} implemented both hiring freezes and even a nominal salary reduction for professors, in anticipation of oncoming financial shock in early 2020.\footnote{
+} implemented both hiring freezes and even a nominal salary reduction for professors, in anticipation of oncoming financial shock in early 2021.\footnote{
     The salary cut was not permanent, as the oncoming financial shock turned out to not be as serious as projected, and the salary cut was returned to professors in the year 2021.
@@
 
-\textbf{Add in an intro conclusion paragraph.}
-``Furthermore, because schools that serve students from lower-income backgrounds are most affected by state appropriations cuts, reductions in state support have helped exacerbate inequality and stratification of outcomes in the postsecondary sector.'' \citep{NBERw27885}.
+This analysis finds that faculty composition is meaningfully affected by the last few decades' stagnation in state support for higher education.
+Analysis of outcomes for incumbant professors in Illinois implies that the explaining mechanism is changes to hiring by public universities in response to the increased financial constraints.
+Previous findings from the literature on the effectiveness of tenured and tenure-track faculty versus contingent faculty in the areas of research and teaching.
+Combined with this analysis, these findings provide a plausible mechanism for the deteriorating student ouctomes shown to be a result of stagnating finances at public universities \citep{NBERw23736,NBERw27885}.
 
@@
 }
-Data are consistent between the years 1990 and 2020,\footnote{
+Data are consistent between the years 1990 and 2021,\footnote{
     The years 1987-1989 are represented in these data in an incompletion fashion, so I focus on the years 1990 onwards.
@@
 and provide information on the total revenues a university received from every source (including state governments), enrollment number, plus faculty count and total expenditures on salaries.\footnote{
-    I combine the Urban Institute's 2018 compilation of IPEDS data for the years 1990-2017, and manually combine with raw NCES data regarding the year 2018-2020 for all relevant variables.
+    I combine the Urban Institute's 2018 compilation of IPEDS data for the years 1990-2017, and manually combine with raw NCES data regarding the year 2018-2021 for all relevant variables.
     Figures for enrollment, faculty counts and salaries come from the raw NCES version of IPEDS for all years to address inconsistencies in the Urban Institute's data formulation for these variables.
@@
     \centering
-    \caption{IPEDS Summary Statistics, PUs Panel 1987--2020}
+    \caption{IPEDS Summary Statistics, PUs Panel 1987--2021}
     \makebox[\textwidth][c]{\input{tables/ipeds-summary-fte.tex}}
@@
 }
-To investigate the distribution, I integregate individual-level data for every PU professor in the state of Illinois between the years 2010-2020.
+To investigate the distribution, I integregate individual-level data for every PU professor in the state of Illinois between the years 2010-2021.
 IBHED host the information;
@@
 }
-This publicly available data provides the basis to build a panel of Illinois PU professors 2010-2020; I define a professor as an individual observation by their first plus last name and university pairing, and link this database to IPEDS information on finances for their employing institution.
+This publicly available data provides the basis to build a panel of Illinois PU professors 2010-2021; I define a professor as an individual observation by their first plus last name and university pairing, and link this database to IPEDS information on finances for their employing institution.
 \begin{table}[h!]
@@
     \centering
-    \caption{IBHED Summary Statistics, Professor Panel 2011--2020.}
-    \makebox[\textwidth][c]{\input{tables/illinois-summary.tex}}
+    \caption{IBHED Summary Statistics, Professor Panel 2011--2021.}
+    \makebox[\textwidth][c]{\input{tables/illinois-summary-rolling.tex}}
     \label{tab:illinois-summary}
 \end{table}
-The analysis sample focuses on the subset of professors with identified year of hiring between the years 2011 and 2020, representing 1,778 professors in 2011, and 9,099 in 2020.
+The analysis sample focuses on the subset of professors with observed year of hiring between the years 2011 and 2021, representing 1,778 professors in 2011, and 9,099 in 2021.
 \autoref{tab:illinois-summary} presents the summary statistics for this sample.\footnote{
-    The full data IBHED contain 173,000 observations for every public university professor in Illinois 2010-2020, which representing 16,932 professors in the year 2010 and 15,352 in the year 2020.
-    The reasons for restricting to the sample with identified year of hire in the range 2011-2020 is explained in \autoref{sec:iv-model-indiv}.
+    The full data IBHED contain 173,000 observations for every public university professor in Illinois 2010-2021, which representing 16,932 professors in the year 2010 and 15,352 in the year 2021.
+    Summary statistics for the entire sample can be see in Appendix \autoref{tab:illinois-summary-entire}.
+    The reasons for restricting to the sample with identified year of hire in the range 2011-2021 is explained in \autoref{sec:iv-model-indiv}.
 }
@@
 
-\autoref{fig:funding} exhibits the trends for the mean of US PUs between the years 1990-2020.
+\autoref{fig:funding} exhibits the trends for the mean of US PUs between the years 1990-2021.
 We see a rise in total revenues received by PUs (from all sources), and a notable increase in mean tuition revenues from \$48 million per year-university to \$150 million per year-university. 
@@
 At the same time, total state appropriations stagnated at around \$100 million per year-university for 1990-2008, fell slightly around 2008 and have not recovered.
-While PUs experienced a financial stagnation, private universities were not exposed to the same constraints, receiving \$37,000 per student in 1990 and \$49,000 in 2020 with no corresponding decline in any specific component.
+While PUs experienced a financial stagnation, private universities were not exposed to the same constraints, receiving \$37,000 per student in 1990 and \$49,000 in 2021 with no corresponding decline in any specific component.
 
@@
 
-At the same time, student enrollment at PUs rose precipitously.
-6.2 million students were enrolled in PUs in 1990, and this number rose by 47\% to 9.1 million, with most of the increase occurring over the years 2000-2020.
 \begin{figure}[h!]
@@
 \end{figure}
-
-\autoref{fig:enrollment} shows that total enrollment at private universities has also risen over the same time period, but not as drastic in either relative or absolute terms; the mean public university also grew over this time period, from 9,800 students in 1990 to 11,800 in 2020.
-This means that per student mean revenues per student (seen in \autoref{fig:mean-funding-fte}) have stagnated across all measures, and particularly fallen from \$11,000 per student in 1990 to less than \$8,000 per student in 2020.
+At the same time, student enrollment at PUs rose precipitously.
+6.2 million students were enrolled in PUs in 1990, and this number rose by 47\% to 9.1 million, with most of the increase occurring over the years 2000-2021.
+\autoref{fig:enrollment} shows that total enrollment at private universities has also risen over the same time period, but not as drastic in either relative or absolute terms; the mean public university also grew over this time period, from 9,800 students in 1990 to 11,800 in 2021.
+This means that per student mean revenues per student (seen in \autoref{fig:mean-funding-fte}) have stagnated across all measures, and particularly fallen from \$11,000 per student in 1990 to less than \$8,000 per student in 2021.
 \begin{figure}[!h]
@@
 
-\autoref{fig:fte-perprof} documents the divergence in faculty composition (per student) between the average private and public university 1990-2020.
+\autoref{fig:fte-perprof} documents the divergence in faculty composition (per student) between the average private and public university 1990-2021.
 Private universities start with  a higher baseline of around 4.5 professors per 100 students, and exhibits yearly variation of less than 0.5 professor per 100 students over the thirty years.
@@
 We see a similar difference in baseline, and fall for the years 2008-2011 for assistant professors.
-Private and public universities have similar numbers of associated and full professors before thea year 2000, yet this number has fallen by over 20\% in the next 20 years only for PUs: in 2020 the mean PU has 6 fewer full professors per tudent than the mean private university.
+Private and public universities have similar numbers of associated and full professors before the year 2000, yet this number has fallen by over 20\% in the next 20 years only for PUs: in 2021 the mean PU has 6 fewer full professors per student than the mean private university.
 At the same time period, we see the rise in use of non-tenure track instructor positions (referred to as lecturers from here on), who were employed at similar rates in both sectors in 1990 yet have been utilised by PUs at a higher rate since.
+
+\subsection{Trends in Illinois}
+\label{sec:trends-illinois}
+
+The state of Illinois particularly exemplifies the stagnation in state support for higher education that the rest of the nation has experienced 1990-2021.
+\autoref{fig:illinois-funding} shows the trends for revenues among the seven Illinois public university campuses, including the falling share of state appropriations and substitute towards tuition revenue.
+
+\begin{figure}[h!]
+    \centering
+    \caption{Mean Revenues among Illinois Public Universities, by Year.}
+    \begin{subfigure}[b]{0.495\textwidth}
+        \centering
+        \caption{Total, millions \$ 2021 CPI-U.}
+        \includegraphics[width=\textwidth]{figures/illinois-funding-total.png}
+        \label{fig:illinois-funding-total}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.495\textwidth}
+        \centering
+        \caption{Per Enrolled Student, \$ 2021 CPI-U.}
+        \includegraphics[width=\textwidth]{figures/illinois-funding-fte.png}
+        \label{fig:illinois-funding-fte}
+    \end{subfigure}
+    \label{fig:illinois-funding}
+\end{figure}
+
+The decade of 2010-2021 has also seen yearly variation in its revenues, particularly around 2016.
+In the calendar year 2015 partisan disagreements between the democratic legislature and republican governor led to the 2016 fiscal year starting with no state budget.
+State agencies, and higher education institutions, employed accounting tehniques to continue operation.\footnote{
+    Fiscal year 2016 refers to June 205 to June 2016, as is the same for the academic year definition in this analysis. 
+    K-12 schools were unaffected thanks to a separate budget agreement allowing non-higher education to operate as before.
+}
+While state institutions continues to operate, there were drastic revenue and spending cuts in response to the budget impasse, as it continued through fiscal year 2017, and ended with a new budget restoring revenues to state agencies and universities for the 2018 fiscal year.
+This variation in public university revenues, stemming entirely from political disagreements --- not from state decisions regarding higher education and its finances \citep{young2020squandered} --- and via the state appropriations channel, mean that Illinois public universities exhibit sizable changes in their state approriations over 2011-2021, of similar order to those for the rest of the country over 1990-2021.
 
@@
 Na\"ively we can analyse the relationship between professor-outcomes $Y_{i,t}$ at university $i$ in year $t$ as a result of university revenues per student $X_{i,t}$.
-Professor outcomes include count of professors per student, and average salary, within each university-year.
+The exact measure of university revenues, $X_{it}$ is non-institutional revenues, the sume of university funding from outside sources -- local, federal, and state appropriations plus tuition revenue.\footnote{
+    This measure is in line with the \cite{NBERw27885} approach of considering revenues from the state, but does not ignore yearly changes in revenues arising from tuition revenues -- which are shown to be heavily related.
+}
+Outcomes include count of professors per student, and average salary, within each university-year.
 This model gives $\beta$ the association with these outcomes and university revenues per student,\footnote{
@@
 including fixed effects to control for effects specific to the university and year.
-Log$(.)$ transforming the variables improves the interpretability of $\beta$ as an elasticity for professor count per student with respect to institution revenues.
+Log$(.)$ transforming the variables improves the interpretability of $\beta$ as an elasticity for professor count per student with respect to university revenues.
 \begin{equation}
@@
 
-\autoref{tab:firststage-reg} presents results of the first-stage regression, separately with and without the tuition control and fixed effects, at the institution level (i.e.
-using IPEDS data).\footnote{
-    Representations for frequentist significance levels (i.e. 10, 5, 1\% etc.) are omitted here, and in all following tables.
-}
 \begin{table}[!h]
@@
 \end{table}
+
+\autoref{tab:firststage-reg} presents results of the first-stage regression, separately with and without the tuition control and fixed effects, at the institution level (i.e.
+using IPEDS data).\footnote{
+    Representations for frequentist significance levels (i.e. 10, 5, 1\% etc.) are omitted here, and in all following tables.
+}
 We note columns (1) and (2) estimate that a shock to state appropriations per student in the state of 10\% is associated with around 2.5\% change in non-institutional revenues per student at the university; the instrument is strong, and we note similarity in estimates with and without inclusion of fixed effects.
@@
 Column (1) represents the estimates for \autoref{eqn:firststage} with fixed effects, and is the preferred form that I proceed with in the following.
-
-\autoref{fig:firststage-lp} presents local projection estimates for the staying-power of the state appropriation shock to university revenues.
-We see the time 0 estimate coincides with that of Column (1) in \autoref{tab:firststage-reg}, the contemporaneous effect of the shock revenues.
-We see a decaying, yet real and positive, effect of the shock on revenues for the next 5 years, illustrating the staying power of the effect.
 
@@
 \end{figure}
+
+\autoref{fig:firststage-lp} presents local projection estimates for the staying-power of the state appropriation shock to university revenues.
+We see the time 0 estimate coincides with that of Column (1) in \autoref{tab:firststage-reg}, the contemporaneous effect of the shock revenues.
+We see a decaying, yet real and positive, effect of the shock on revenues for the next 5 years, illustrating the staying power of the effect.
 
@@
 $j$ indexes each professor in year $t$, $\tau(j)$ for the year the professor first joins their Illinois PU.\footnote{
-    Identifying $\tau(j)$ is possible for $j$ by restricting to all professors who are hired in the time period 2011-2020 (i.e. in the years after the first in the panel).
+    Identifying $\tau(j)$ is possible for $j$ by restricting to all professors who are hired in the time period 2011-2021 (i.e. in the years after the first in the panel).
     Professors who were hired in the year 2010 are impossible to distinguish from those in year preceding, and so are not considered.
@@
 
-DISCUSS WHAT THOSE POINT ESTIMATES MEAN, DISCUSSING THE POSSIBILITY THAT LECTURERS ARE INFERIOR GOODS, TENURE TRACK NORMAL GOODS.
-
-This finding is particularly in line with the \cite{turner2014impact} observation regarding hiring freezes particularly for tenure-track hirings, and \cite{brown2014endowment} from shocks to endowments.
-Similarly, the effect is exhibited on order of 2.6\% among full professors (those holding tenure), likely through the channel of retirements or replacement hirings (since tenured professors cannot be forcibly fired/not renewed without cause).
-
-Write something about the level of tenure, and that data are available for a subset of years, and shown in Appendix \autoref{tab:tenurecount-shock-reg-fte}.
-
-\autoref{fig:all-count-lp} shows local projection estimates, staying power of the effect on prof count.
-
 \begin{figure}[h!]
@@
 \end{figure}
+\autoref{fig:all-count-lp} shows local projection estimates, where the effect on total professor count per student persists at the same magnitude of 0.3 four years after the identified shock.
+These findings are in line with \cite{turner2014impact} observing that universities froze hiring particularly for tenure-track positions in response to negative budget shocks in the late 2000s, and \cite{brown2014endowment} from shocks to university endowments.
+We see a negative coefficient for the number of lecturers per student.
+This effect lines up with two trends we see in \autoref{sec:trends}: that public universities revenues (per student) are decreasing while utilisation of non-tenure track lecturers increased over the same time period.
+And yet, identifying the university revenues by the state appropriations shock increases the magnitude of the effect (Column 2) with respect to the association (Column 1), increasing evidence that the two trends are causally linked.
+Between years 1990-1999 IPEDS provided count of professors by explicit tenure designation (opposed to named position as in \autoref{tab:facultycount-shock-reg}), and Appendix \autoref{tab:tenurecount-shock-reg-fte} presents results of the respective regressions by tenure status.
+Results are relatively similar, yet imprecise from the much smaller number of observations.
 
-WRITE HERE ABOUT STAYING POWER.
+Together these findings imply that public universities increase their count of tenure-track and tenured professors per student in years where revenues are more plentiful, by increasing hiring intensity and efforts to maintain incumbent professors.
+Yet, when revenues are not so bountiful, count of tenure-track and tenured professors per student decreases.
+At the same time, public university enrollment secularly increased over this time period, so that the universities satisfied their teaching obligations with non-tenure track lecturers.
+The mechanism that explains the changes in faculty composition are, however, not clear at this unti of analysis; individual data are needed to delve deeper.
 
@@
 
-Estimates for the model where units of analysis are individual professors at Illinois PUs follow.
+The first outcome is total salary for each professor by their position;
+\autoref{tab:facultysalaries-shock-illinois} presents 2SLS estimates for the elasticity of total salaries with respect to university revenues.
+The results are not distinguishable from zero in any case, which implies that salaries for all years of employment at the university are not causally related to changes in university revenues.
 
@@
     \caption{2SLS Estimates for Faculty Salaries at Illinois Universities.}
-    \makebox[\textwidth][c]{\input{tables/facultysalaries-shock-illinois.tex}}
+    \makebox[\textwidth][c]{\input{tables/facultysalaries-shock-illinois-rolling.tex}}
     \begin{flushleft}
@@
 
+Yet, it is possible that another measure of salary is impacted by the university revenues: salary in first-year of employment.
+\autoref{tab:newhiresalaries-shock-illinois} presents 2SLS for the elasticity of total salary in the first year of a professor' employment  with respect to university revenues.
+The elasticity is non-distinguishable from zero among only lecturers: a 10\% increase in university revenues (via state appropriation shocks) leads to a 9.5\% increase in total salary paid to lecturers.
+
 \begin{table}[!h]
@@
     \caption{2SLS Estimates for Faculty Salaries, in First-Year, at Illinois Universities.}
-    \makebox[\textwidth][c]{\input{tables/newhiresalaries-shock-illinois.tex}}
+    \makebox[\textwidth][c]{\input{tables/newhiresalaries-shock-illinois-rolling.tex}}
     \begin{flushleft}
@@
         \textbf{Note}: Standard errors are clustered at the institution and first year of employment level. 
-        Included fixed effects are institution and first-year of employment, since every year-institution interaction over-identifies the system. 
     \end{flushleft}
@@
 
-\autoref{tab:promotion-shock-illinois} describes estimates for promotion rate: particularly sharp for the move associate to full.
+Note that professor position (assistant, associate, full) is also an outcome in this set-up by the promotion channel.
+\autoref{tab:promotion-shock-illinois} describes estimates where the outcome is rate of promotion within each professor group.
+The lecturer group describes the sample of professors who were ever listed as a non-tenure track lecturer and the binary for promotion equals one in a year they achieved an assistant professor position; assistant the same for the assistant to associate professor promotion; associate same for associate to full promotion.
 
-\begin{table}[!h]
-    \onehalfspacing
-    \centering
-    \caption{2SLS Estimates for Faculty Promotion Rate at Illinois Universities.}
-    \makebox[\textwidth][c]{\input{tables/promotion-shock-illinois.tex}}
-    \begin{flushleft}
-        \footnotesize
-        \textbf{Note}: Standard errors are clustered at the institution and first year of employment level.
-    \end{flushleft}
-    \label{tab:promotion-shock-illinois}
-\end{table}
+There is no discernible relationship between revenues and promotion among lecturers, as only a small group of professors achieves the lecturer to assistant professor promotion in this sample.\footnote{
+    749 in this sample make the transition lecturer to assistant professor, while there are 9,316 distinct lecturers in the data sample.
+}
+There is an imprecise positive relationship between rate of promotion assistant to associate professor, and no clear relationship between promotion rate among associate professors.
+Furthermore, local projection estimates (\autoref{fig:promoted-illinois-lp}) show that the there is no promotion effect, among all professors, for any of the following years.
+Lastly, Appendix \autoref{tab:facultyleaving-shock-illinois} describes estimates for rate of exit from the employing university among Illinois professors, and finds no relationship with university revenues.
+
+\subsection{Discussion}
+
+The results suggest that the recent stagnation in public university revenues per student has affected the composition of faculty within a university, away from professors tenured and on the tenure-track and towards non-tenured lecturers.
+At the same time, there are little discernible effects on individual professors hired in the years 2011-2021 at Illinois public universities, except for the salaries of first year lecturers.
+The rate that faculty leave their university, and their rate of promotion between positions, are unaffected by the university revenues, so this leaves one primary channel to explain changes in faculty composition: hiring.
+
+\cite{turner2014impact} documents the wide-spread practice of hiring freezes at universities in response to budget shocks around the 2008 recession.
+Throughout the last decade, multiple such measures were taken by Illinois public universities in response to their deteriorating finances \citep{furlough2010}.
+The University of Illinois\footnote{
+    The University of Illinois inclves three public university campuses: Urbana-Champaign, Chicago, Springfield.
+} did not receive the allocated state appropriations from the state of Illinois on time, so enacted cost-cutting measures to stay fiscally solvant.
+The university system placed a hold on all hiring for filling state-funded positions and promotions.
+Faculty were furloughed (placed on leave without pay) for a day each month, and university administrators placed on two days per month furlough, or eligible university employees could accept a voluntary, equivalent pay reduction.
+The University of Western Illinois adopted very similar measures in response to the Illinois budget crisis in 2016-2017 \citep{wiu2016}.
+
+These examples show that the Illinois state universities attempted to affect each of the professor outcomes investigated in \autoref{sec:results}, yet the only measurable affect was on count of professor per student (implied via slower hiring), and first-year lecturer salaries.
+To investigate these trends for the entire state, \autoref{fig:newhire-illinois} shows the count of professors hired in the years 2011-2021, and their mean salary, by position.
+In particular, no position had sustained rises in hiring while enrollment within the university was rising (so that new hires per student necessarily fell), and hiring of new administrators jumped once the Illinois budget crisis ended in 2017.
+At the same time, professor new hires' starting salaries remained relatively constant for lecturer and assistant professors, yet rose and fell over the decade for full professors and administrators.
 
@@
     \centering
-    \caption{Local Projection Estimates for Promotion Rate at Illinois Universities.}
-    \includegraphics[width=0.6\textwidth]{figures/promoted-illinois-lp.png}
-    \label{fig:promoted-illinois-lp}
+    \caption{Trends in New Hires at Illinois Public Universities 2011-2021.}
+    \begin{subfigure}[b]{0.495\textwidth}
+        \centering
+        \caption{New Hire Count, Total.}
+        \includegraphics[width=\textwidth]{figures/newhire-count-illinois.png}
+        \label{fig:newhire-count-illinois}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.495\textwidth}
+        \centering
+        \caption{Mean Salary in First Year for New Hires, \$ 2021 CPI-U.}
+        \includegraphics[width=\textwidth]{figures/newhire-salary-illinois.png}
+        \label{fig:newhire-salary-illinois}
+    \end{subfigure}
+    \label{fig:newhire-illinois}
 \end{figure}
 
-\autoref{tab:facultyleaving-shock-illinois} describes estimates for exit rate: not much going on.
+While costs education have been rising in the US, thanks to multiple overlapping trends \citep{ehrenberg2012}, public universities have also had to deal with declining state support.
+It is natural to expect that such headwinds will lead to systematic change at public universities, changes which are not neutral on the professors they employ, or to their goals of research and providing education.
+Furthermore, private universities in this country are not exposed to these same financial headwinds.
+If public universities' hiring of faculty is disrupted by changes in state appropriations, then we can expect that private universities will benefit (in a compeitive sense) by being able to hire better professors from relatively resource constrained public universities.
+While public universities continue to educate the majority of higher education students in the US, and disproptionately students from low-income backgrounds and minorities, we should worry about the effects that restricting their financial support has on both the quality and distribution of education outcomes in this country.
 
-\newpage
-\subsection{Robustness Checks}
+\subsubsection{Threats to Identification}
 
-This section will exhibit systematic robustness checks for each of the specifications estimated in \autoref{sec:results}.
-In particular, re-estimation of the models scaled by number of students within each university can imply a possible mechanism for the worsening student outcomes in response to negative PU finance shocks shown by \cite{NBERw23736,NBERw27885}.
-Similarly to graph the trajectory of finance shocks among PUs: PUs show a secular decline in (per student) revenues from state sources.
-
-Lastly, \cite{blandhol2022tsls} provide reason to doubt the LATE interpretation of two-stage least squares estimates with multiple layers of fixed effects (i.e.
-both individual and time fixed effects).
-Further analysis of their methods, and how it applies to the model presented in \autoref{sec:empirics} is planned.
+\cite{blandhol2022tsls} provide reason to doubt the LATE interpretation of two-stage least squares estimates with multiple layers of fixed effects (i.e. both individual and time fixed effects).
+Needs comment, and to what extent should this question the empirical strategy.
 
@@
 
-\textbf{REVISE THIS PARAGRAPH ENTIRELY}
-
-This report has presented the set-up for, and the preliminary results of, a project investigating the relationship between PUs' finances and the professors they employ.
-Results show that employment count at the university level is affected by (shocks to) total revenues, concentrated among assistant and full professors while the count of lecturers is unaffected.
-Salaries at the university-level are unaffected.
-
-\textbf{ADD A DISCUSSION PARAGRAPH OR TWO TAKING CRITICISM IN ADVANCE}
-
-I present a plan to extend the empirical analysis to individual data for all PU professors in the state of Illinois 2010-2019 --- the first such use of these data in an academic article.
+\textbf{A few paragraphs conclusion go here.}
 
@@
 \label{appendix:part1}
+
+\begin{table}[h!]
+    \onehalfspacing
+    \centering
+    \caption{IBHED Summary Statistics, Entire Professor Panel 2010--2021.}
+    \makebox[\textwidth][c]{\input{tables/illinois-summary.tex}}
+    \label{tab:illinois-summary-entire}
+    \begin{flushleft}
+        \footnotesize
+        \textbf{Note}: \autoref{tab:illinois-summary} presents the summary statistics for professors with an observaed year of first-year at their university, while this table exhibits the data regarding every professor at an Illinoic public university 2010--2021.
+    \end{flushleft}
+\end{table}
+
 \begin{table}[H]
@@
     \centering
-    \caption{First Stage Estimates, for Total University Revenues at the Individual-Level}
-    \makebox[\textwidth][c]{\input{tables/firststage-illinois.tex}}
+    \caption{First Stage Estimates, for Total University Revenues at the Individual-Level.}
+    \makebox[\textwidth][c]{\input{tables/firststage-illinois-rolling.tex}}
     \label{tab:firststage-illinois}
@@
     \caption{Local Projection Estimates for First-Stage \autoref{eqn:secondstage1_indiv}.}
-    \includegraphics[width=0.6\textwidth]{figures/firststage-illinois-lp.png}
+    \includegraphics[width=0.6\textwidth]{figures/firststage-illinois-lp-rolling.png}
     \label{fig:firststage-illinois-lp}
@@
     \caption{OLS and 2SLS Estimates for University Faculty-Tenure Composition.}
-    \makebox[\textwidth][c]{\input{tables/tenurecount-shock-reg-fte}}
+    \makebox[\textwidth][c]{\input{tables/tenurecount-shock-reg-fte.tex}}
     \begin{flushleft}
@@
     \caption{2SLS Estimates for Faculty Exit Rate at Illinois Universities.}
-    \makebox[\textwidth][c]{\input{tables/facultyleaving-shock-illinois.tex}}
+    \makebox[\textwidth][c]{\input{tables/facultyleaving-shock-illinois-rolling.tex}}
     \begin{flushleft}
@@
 
+\begin{table}[!h]
+    \onehalfspacing
+    \centering
+    \caption{2SLS Estimates for Faculty Promotion Rate at Illinois Universities.}
+    \makebox[\textwidth][c]{\input{tables/promotion-shock-illinois-rolling.tex}}
+    \begin{flushleft}
+        \footnotesize
+        \textbf{Note}: Standard errors are clustered at the institution and first year of employment level.
+    \end{flushleft}
+    \label{tab:promotion-shock-illinois}
+\end{table}
+
+\begin{figure}[H]
+    \centering
+    \caption{Local Projection Estimates for Promotion Rate at Illinois Universities.}
+    \includegraphics[width=0.6\textwidth]{figures/promoted-illinois-lp-rolling.png}
+    \label{fig:promoted-illinois-lp}
+\end{figure}
+
 \end{document}

### 2022-08-04 programs/analysis/individual-shock.R

@@ -1,6 +1,6 @@
 #!/usr/bin/R
-## Senan Hogan-Hennessy, 13 June 2022
-## IV for Prof's salaries, using Illinois data.
+## Senan Hogan-Hennessy, 3 August 2022
+## IV for Prof's outcomes, using Illinois data.
 library(tidyverse) ## functions for data manipulation and visualization
 library(lfe)
 # My custom flavour of Stargazer TeX tables:
@@ -63,39 +63,14 @@ reg.data <- reg.data %>%
     mutate(firstyear = min(year, na.rm = TRUE)) %>%
     ungroup()
 
-# Calculate the rolling shock to state appropriations (Lovenheim 2020 p. 13)
-# and link it to the prof's first year
-
-# Generate a Professor's first year
-reg.data <- reg.data %>%
-    arrange(name, year, unitid) %>%
-    group_by(name, unitid) %>%
-    mutate(firstyear = min(year, na.rm = TRUE)) %>%
-    ungroup()
-# Generate the rolling share, for a Professor's first year of employment.
-reg.data <- ipeds.data %>%
-    transmute(unitid = unitid,
-        firstyear = year,
-        staterevenues_rollshare =
-            stateappropriations_real / totalrevenues_real) %>%
-    right_join(reg.data, by = c("unitid", "firstyear")) %>%
-    # Remove the rolling shock if the panel does not include first year for prof
-    mutate(staterevenues_rollshare =
-        ifelse(2010 < firstyear & firstyear < 2021,
-            staterevenues_rollshare, NA)) %>%
-    # Get the shock as a product of rolling share and yearly state approp
-    mutate(appropriationshock_perEnroll_rolling =
-        staterevenues_rollshare * (
-            allstate_stateappropriations_real / stateEnroll_count))
-
 # Remove IPEDS + Illinois data to save memory
 rm(ipeds.data)
 rm(illinois.data)
 gc()
 
-# How many individual professors in 2010 and 2020?
+# How many individual professors in 2010 and 2021?
 reg.data %>% filter(year == 2010) %>% count(unitid, name) %>% nrow() %>% print()
-reg.data %>% filter(year == 2020) %>% count(unitid, name) %>% nrow() %>% print()
+reg.data %>% filter(year == 2021) %>% count(unitid, name) %>% nrow() %>% print()
 
 # Binary for professor position
 reg.data <- reg.data %>%
@@ -118,6 +93,18 @@ reg.data <- reg.data %>%
[…]

### 2022-08-04 programs/analysis/ipeds-public-private.R

@@ -27,7 +27,7 @@ ipeds.data <- ipeds.data %>%
     # Non for-profit unis, 4-year
     filter(fouryear == 1, forprofit == 0) %>%
     # Remove years with incomplete data
-    filter(1990 <= year, year != 2000) %>%
+    filter(1990 <= year, year != 2001, year <= 2021) %>%
     # Restrict to relevant columns and rows.
     select(
         # Identifiers and controls
@@ -76,7 +76,6 @@ enrollment.data <- ipeds.data %>%
     ungroup()
 # Draw the graph for enrollment total
 enrollment_total.plot <- enrollment.data %>%
-    filter(year > 1990) %>%
     ggplot(aes(x = year, y = enrollment_total, colour = factor(public))) +
     geom_point() +
     geom_line() +
@@ -230,13 +229,14 @@ lecturer_fte_perprof.plot <- prof_perfte.data %>%
     scale_x_continuous(name = "Year",
         breaks = seq(1985, 2020, by = 5)) +
     scale_y_continuous(name = "",
-        limits = c(0, 0.004),
-        breaks = seq(0, 0.005, by = 0.001),
+        limits = c(0.003, 0.007),
+        breaks = seq(0, 0.007, by = 0.001),
         labels = scales::comma) +
     theme_bw() +
     theme(plot.title = element_text(hjust = 0.5),
         legend.position = "top") +
     scale_colour_discrete(name = "", labels = c("Private", "Public"))
+lecturer_fte_perprof.plot
 ggsave("../../text/figures/lecturer-fte-perprof.png",
     plot = lecturer_fte_perprof.plot,
     units = "cm", width = fig.width, height = fig.height)

### 2022-08-04 programs/analysis/ipeds-shock.R

@@ -218,6 +218,84 @@ print((
     ) / (enrollment_yearly.data %>% filter(year == 1991) %>% pull(enrollment_reported)
 ))
 
+# Repeat the analysis for Illinois
+# Plot total gov funding by year.
+illinois_funding_fte.data <- reg.data %>%
+    filter(state == "IL") %>%
+    group_by(year) %>%
+    summarise(
+        totalrevenues_real =
+            mean(totalrevenues_real / enrollment_reported, na.rm = TRUE),
+        nonauxrevenues_real =
+            mean(nonauxrevenues_real / enrollment_reported, na.rm = TRUE),
+        stateappropriations_real =
+            mean(stateappropriations_real / enrollment_reported, na.rm = TRUE),
+        tuitionrev_real =
+            mean(tuitionrev_real / enrollment_reported, na.rm = TRUE))
+# Define the plot
+illinois_funding_fte.graph <- illinois_funding_fte.data %>%
+    pivot_longer(!year, names_to = "variable", values_to = "millions") %>%
+    ggplot(aes(x = year, y = millions, colour = variable)) +
+    geom_point() +
+    geom_line() +
+    # Adjust the names and axis
+    scale_x_continuous(name = "Year",
+        breaks = seq(1985, 2020, by = 5)) +
+    scale_y_continuous(name = "",
+        limits = c(0, 52000),
+        breaks = seq(0, 50000, by = 10000),
+        labels = scales::comma) +
+    theme_bw() +
+    theme(plot.title = element_text(hjust = 0.5),
+        legend.position = "top") +
+    scale_colour_discrete(name = "",
+        breaks = c("totalrevenues_real", "nonauxrevenues_real",
+        "stateappropriations_real", "tuitionrev_real"),
+        labels = c("Total", "Non-inst.", "State", "Tution"))
+# Save this plot
+ggsave("../../text/figures/illinois-funding-fte.png",
+    plot = illinois_funding_fte.graph,
+    units = "cm", width = fig.width, height = fig.height)
+
+# Plot total gov funding by year.
+illinois_funding.data <- reg.data %>%
+    filter(state == "IL") %>%
+    group_by(year) %>%
+    summarise(
+        totalrevenues_real =
+            mean(totalrevenues_real, na.rm = TRUE) / (10^6),
+        nonauxrevenues_real =
+            mean(nonauxrevenues_real, na.rm = TRUE) / (10^6),
[…]

### 2022-08-04 programs/data-build/urban-ipeds-build.R

@@ -333,6 +333,10 @@ urban_ipeds.data <- urban_ipeds.data %>%
 urban_ipeds.data <- urban_ipeds.data %>%
     distinct(unitid, year, .keep_all = TRUE)
 
+# Make year refer to the year of spring term, and not autumn of calendar year
+urban_ipeds.data <- urban_ipeds.data %>%
+    mutate(year = year + 1)
+
 
 ################################################################################
 ## Save the constructed data file.

### 2022-08-04 programs/run.sh

@@ -16,12 +16,14 @@ cd analysis
 Rscript ipeds-public-private.R
 # Analysis and regressions at the institution-level.
 Rscript ipeds-shock.R
-# Analysis and regressions at the individual-level.
+# Analysis and regressions at the individual-level, with base-year shock IV.
 Rscript individual-shock.R
+# Analysis and regressions at the individual-level, with rolling shock IV.
+Rscript individual-shock-rolling.R
 cd ..
 
 # Compile paper, with outputs of progams as inputs for TeX files
-cd ..
+cd text
 latexmk -pdf paper.tex
 latexmk -c
 rm paper.bbl

### 2022-08-07 programs/analysis/individual-shock.R
adjusted draft

@@ -192,7 +192,7 @@ reg.data %>%
 
 # Explain Revenues with a shock to (only) state appropriations.
 firststage_approp.reg <- reg.data %>%
-    felm(log(nonauxrevenues_real / enrollment_reported) ~ 1 +
+    felm(log(stateappropriations_real / enrollment_reported) ~ 1 +
         log(appropriationshock_perEnroll_real) |
         unitid + firstyear |
         0 |
@@ -207,7 +207,7 @@ firststage_approp.fstat <-
 # Without the FEs
 # Explain State appropriations with a shock to (only) state appropriations.
 firststage_approp_noFE.reg <- reg.data %>%
-    felm(log(nonauxrevenues_real / enrollment_reported) ~ 1 +
+    felm(log(stateappropriations_real / enrollment_reported) ~ 1 +
         log(appropriationshock_perEnroll_real) |
         0 |
         0 |
@@ -221,7 +221,7 @@ firststage_approp_noFE.fstat <-
 
 # Explain Revenues with a shock to (only) state appropriations.
 firststage_approp_tuit.reg <- reg.data %>%
-    felm(log(nonauxrevenues_real / enrollment_reported) ~ 1 +
+    felm(log(stateappropriations_real / enrollment_reported) ~ 1 +
         log(appropriationshock_perEnroll_real) +
         log(tuitionrev_real / enrollment_reported) |
         unitid + firstyear |
@@ -237,7 +237,7 @@ firststage_approp_tuit.fstat <-
 # Without the FEs
 # Explain State appropriations with a shock to (only) state appropriations.
 firststage_approp_tuit_noFE.reg <- reg.data %>%
-    felm(log(nonauxrevenues_real / enrollment_reported) ~ 1 +
+    felm(log(stateappropriations_real / enrollment_reported) ~ 1 +
         log(appropriationshock_perEnroll_real) +
         log(tuitionrev_real / enrollment_reported) |
         0 |
@@ -282,10 +282,9 @@ lecturer.data <- reg.data %>%
     filter(lecturer == 1)
 # Shift-share IV Regression, explained by state appropriation shock
 shiftshare_lecturer_salaries.reg <- lecturer.data %>%
-    felm(log(salary_real + extra_salary_real) ~ 1 +
-        log(tuitionrev_real / enrollment_reported) |
+    felm(log(salary_real + extra_salary_real) ~ 1 |
         unitid + firstyear |
[…]

### 2022-08-07 programs/analysis/ipeds-shock.R

@@ -72,7 +72,6 @@ reg.data <- reg.data %>%
     filter(1990 <= year,
         !is.na(enrollment_reported), enrollment_reported > 0,
         !is.na(all_prof_count),
-        !is.na(nonauxrevenues_real),
         !is.na(stateappropriations_real),
         !is.na(appropriationshock_perEnroll_real))
 
@@ -88,7 +87,7 @@ reg.data %>%
         enrollment_reported = enrollment_reported,
         stateappropriations_real = stateappropriations_real / (10^6),
         totalrevenues_real = totalrevenues_real / (10^6),
-        nonauxrevenues_real = nonauxrevenues_real / (10^6),
+        #nonauxrevenues_real = nonauxrevenues_real / (10^6),
         lecturer_prof_count = lecturer_prof_count,
         assistant_prof_count = assistant_prof_count,
         full_prof_count = full_prof_count,
@@ -106,7 +105,7 @@ reg.data %>%
             "Enrollment, full-time equivalent",
             "State appropriations (millions 2021 USD)",
             "Total revenues (millions 2021 USD)",
-            "Non-instutional revenues (millions 2021 USD)",
+            #"Non-instutional revenues (millions 2021 USD)",
             "Lecturers count",
             "Assistant professors count",
             "Full professors count",
@@ -301,7 +300,7 @@ ggsave("../../text/figures/illinois-funding-total.png",
 
 # Explain Revenues with a shock to (only) state appropriations.
 firststage_approp.reg <- reg.data %>%
-    felm(log(nonauxrevenues_real / enrollment_reported) ~ 1 +
+    felm(log(stateappropriations_real / enrollment_reported) ~ 1 +
         log(appropriationshock_perEnroll_real) |
         unitid + year |
         0 |
@@ -316,7 +315,7 @@ firststage_approp.fstat <-
 # Without the FEs
 # Explain State appropriations with a shock to (only) state appropriations.
 firststage_approp_noFE.reg <- reg.data %>%
-    felm(log(nonauxrevenues_real / enrollment_reported) ~ 1 +
+    felm(log(stateappropriations_real / enrollment_reported) ~ 1 +
         log(appropriationshock_perEnroll_real) |
         0 |
         0 |
@@ -330,7 +329,7 @@ firststage_approp_noFE.fstat <-
 
 # Explain Revenues with a shock to (only) state appropriations.
[…]

### 2022-08-07 programs/data-build/urban-ipeds-build.R

@@ -29,7 +29,7 @@ tenure.data <- read_csv("../../data/urban-ipeds/raw-data/faculty/ipeds-tenure.cs
 # CPI-U from FREDS (since Urban provided cpi has missing years)
 # Yearly average, seasonally adjusted, base year 1982-1984=100
 # https://fred.stlouisfed.org/series/CPIAUCSL
-cpiu.data <- read_csv("../../data/urban-ipeds/raw-data/freds-cpiu.csv")
+cpiu.data <- read_csv("../../data/inflation-indices/freds-cpiu.csv")
 
 
 # Clean data -----------------------------------------------------------------

### 2022-08-07 programs/analysis/individual-shock-rolling.R
adjusted draft

@@ -1,0 +1,746 @@
+#!/usr/bin/R
+## Senan Hogan-Hennessy, 3 August 2022
+## IV for Prof's outcomes, using Illinois data.
+## Using a rolling-share variant of the instrument.
+library(tidyverse) ## functions for data manipulation and visualization
+library(lfe)
+# My custom flavour of Stargazer TeX tables:
+# devtools::install_github("shoganhennessy/stargazer")
+library(stargazer) ## TeX tables
+set.seed(47)
+# This file follows an adjusted Deming Walters (2017, p.10) approach to
+# appropriations shock instrument.
+
+# Define number of digits in tables and graphs
+digits.no <- 3
+
+# Size for figures
+fig.width <- 10
+fig.height <- fig.width * 0.75
+
+
+# Load data --------------------------------------------------------------------
+
+# Data on Illinois Professors
+illinois.data <- read_csv("../../data/states/illinois-professors.csv")
+
+# Data from IPEDS, with relevant variables
+ipeds.data <- read_csv("../../data/urban-ipeds/urban-clean-publicunis.csv") %>%
+    select(
+        # Identifiers and controls
+        unitid,
+        year,
+        state,
+        # Appropriations to the uni
+        totalrevenues_real,
+        nonauxrevenues_real,
+        stateappropriations_real,
+        tuitionrev_real,
+        # Uni spending.
+        nonauxspending_real,
+        # Uni enrollment
+        enrollment_reported,
+        enrollment_fte,
+        # Appropriation shocks Per student
+        appropriationshock_perEnroll_real,
+        appropriationshock_perFTE_real,
+        appropriationshock_peruni_real,
+        # State-level values
+        allstate_stateappropriations_real,
+        stateEnroll_count)
+
+
+# Clean data -------------------------------------------------------------------
+
+# Join IPEDS data to the individual level
+reg.data <- illinois.data %>%
+    # Join the IPEDS finance data.
+    left_join(ipeds.data, by = c("unitid", "year"))
+
+# Generate a Professor's first year
+reg.data <- reg.data %>%
+    arrange(name, year, unitid) %>%
+    group_by(name, unitid) %>%
+    mutate(firstyear = min(year, na.rm = TRUE)) %>%
+    ungroup()
+
+# Calculate the rolling shock to state appropriations (Lovenheim 2021 p. 13)
[…]

### 2022-08-09 text/paper.tex
adjusted draft
extra detail.

@@
     \thanks{I thank Professor Evan Riehl for guidance on this project, Michael Lovenheim, Ronald Ehrenberg, Douglas Miller, and Francine Blau for various comments on the project's direction,
-    as well as seminar participants at Cornell University (Spring 2022) for helpful
+    as well as seminar participants at Cornell University (Autumn 202) for helpful
     discussion.
@@
 An increase of university revenues of 10\%, via a change in state appropriations, increases the number of assistant professors per student by 6.6\% and full professors by 5.3\%, yet decreases the number of lecturers per student by 14.3\%.
-Analysis of all the professors hired at Illinois public universities 2011-2021 implies that incumbant professors are not affected by the changes in university revenues, so that changes in faculty composition arises by changes in hiring pratices at public universities.
-These results show the compositional effects of stagnating state suport for higher education, and raise questions about the direction that public education is heading thanks to these financial headwinds.
+Analysis of all the professors hired at Illinois public universities 2011-2021 implies that incumbent professors are not affected by the changes in university revenues, so that changes in faculty composition arises by changes in hiring practices at public universities.
+These results show the compositional effects of stagnating state support for higher education, and raise questions about the direction that public education heads as these financial headwinds continue.
 
@@
 \noindent
-Public universities educate the majority of higher education students in the US, yet have experienced a secular decline in state funding (per student) the last few decades.
-This decline has been shown to lead to worse education and later-life outcomes for students \citep{NBERw23736,NBERw27885}, yet it is not clear what institutional effects or mechanisms driving effects on students.
-I show that four-year, degree-granting public universities (PUs) are exposed to tightening financial constraints thanks to falling state financial support (per student), and show this effects the composition of professors at PUs.
-Falls in PU revenues, instrumented by state-level finance shocks, lead to a fall in the number of professors per student within a PU, driven by a fall in assistant or tenured professor counts.
-US private universities are not exposed to similar financial constraints during this same time period, and do not exhibit any similar changes in the faculty composition, so that stagnating financial support for PUs have implications for the structure of higher education instruction and research in the US.
+Public universities educate the majority of higher education students in the US, yet have experienced a secular decline in state funding (per student) the last three decades.
+This decline has been shown to lead to worse education and later-life outcomes for students \citep{NBERw23736,NBERw27885}, yet it is not clear what mechanisms drive these effects or how public universities are affected as institutions.
+I show that four-year, degree-granting public universities have experienced falling state funding, leading to falling revenues per student, effecting the composition of professors within public universities in general.
+Falls in public university revenues from the state, instrumented by state-level finance shocks, lead to a fall in the number of tenure-track or tenured professors per student within a university, an increase in the number of lecturers, and overall fall in the number of faculty per student.
+US private universities were not exposed to such financial constraints during this same time period, and do not exhibit the substitution away from tenure-track and tenured professors, so that the stagnating state support for public universities has implications for the structure of higher education instruction and research in the US.
 
-Universities in the US are widely considered the highest performing in the world,yet there are consequential differences between its universities that operate in the private sector and those established by state governments.
-PUs are subject to numerous state-level administrative laws, and rely on their state governments for funding: an average PU received around \$11,600 per enrolled student in 1990, yet only \$8,300 per enrolled student in 2021.\footnote{
+US Universities are widely considered the highest performing in the world, yet there are consequential differences between its universities that operate in the private sector and those established by state governments.
+public universities are subject to numerous state-level administrative laws, and rely on their state governments for funding: an average PU received around \$11,600 per enrolled student in 1990, yet only \$8,300 per enrolled student in 2021.\footnote{
     2021 figures are relatively similar to those for year preceding, despite the Covid-19 pnademic and associated financial shocks, since 2021 refers to the academic year 2019-2021 and so the yearly budget was decided in 2019.
@@
 }
-This fall is driven by stagnating funding provided by state governments, while total enrollment among PUs rose by 46\% over the same time period.
-At the same time, the number of professors per student at private universities stayed relatively stable at 0.045, yet fell from around 0.04 to around 0.35 professors per student at PUs, driven by a fall of over 20\% in the ratio of associate and full (i.e.
-tenured) professors per student at PUs.\footnote{
+This fall is driven by stagnating funding provided by state governments, while total enrollment among public universities rose by 46\% over the same time period.
+At the same time, the number of professors per student at public universities stayed relatively stable at 0.045, yet fell from around 0.04 to around 0.35 professors per student at public universities, driven by a fall of over 20\% in the ratio of associate and full (i.e.
+tenured) professors per student at public universities.\footnote{
     \autoref{fig:fte-perprof} and \autoref{sec:trends} document these trends between private and public universities in full detail.
@@
 A reduction of 10\% in university revenues, via a shock to state appropriations, leads to a fall in 6.6\% in the number of assistant professors per student within a university, 5.3\% for full (tenured) professors, and a reduction in 14.3\% for lecturers.
-Additionally, I use individual level data on professors who were hired at Illinois public universities 2011-2021 to investigate whether the shocks to university revenues affected incumbant professors at public universities.
-Incumbant professors were not meaningfully affected, in terms of total salary, promotion rate, or rate of leaving the Illinois public university system, which implies that changes in faculty composition came about by changes in the composition of hiring new professors at the university level.
+Additionally, I use individual level data on professors who were hired at Illinois public universities 2011-2021 to investigate whether the shocks to university revenues affected incumbent professors at public universities.
+Incumbent professors were not meaningfully affected, in terms of total salary, promotion rate, or rate of leaving the Illinois public university system, which implies that changes in faculty composition came about by changes in the composition of hiring new professors at the university level.
 
-\cite{NBERw23736} first study the effect of (changes in) PUs' finances on student outcomes by isolating changes in PUs' expenditures on student instruction, addressing endogeneity of expenditures decisions by instrumenting expenditures by an instrument for state appropriations and tuition caps.
+\cite{NBERw23736} first study the effect of (changes in) public universities' finances on student outcomes by isolating changes in public universities' expenditures on student instruction, addressing endogeneity of expenditures decisions by instrumenting expenditures by an instrument for state appropriations and tuition caps.
 \cite{NBERw23736} find that increases in PU spending (via state appropriations) increases enrolment and degree completion among students, but corresponding changes in attendance price (via tuition caps/freezes) do not have an effect.
 \cite{chakrabarti2018effect,NBERw27885} use the same instrument for PU revenues at the state level, combined with richer information on students' individual-level outcomes, to show that increases in state appropriations lead to lower time to completion and debt levels among students later in life.
-\cite{bound2019public} use another variant of the same approach to show that state-level higher education spending cuts induced PUs to shift toward tuition as their primary source of revenue, and shift institutional resources towards students who pay higher tuition rates.
+\cite{bound2019public} use another variant of the same approach to show that state-level higher education spending cuts induced public universities to shift toward tuition as their primary source of revenue, and shift institutional resources towards students who pay higher tuition rates.
 In a similar vein, \cite{bound2007cohort} document variation in per student state funding for higher education resulting from yearly changes in a state's student-age population; they find that lower per student funding lead to lower higher education completion rate among students.
@@
 
-This analysis finds that faculty composition is meaningfully affected by the last few decades' stagnation in state support for higher education.
+This analysis finds that faculty composition is meaningfully affected by recent stagnation in state support for higher education.
 Analysis of outcomes for incumbant professors in Illinois implies that the explaining mechanism is changes to hiring by public universities in response to the increased financial constraints.
@@
 \begin{table}[h!]
-    \onehalfspacing
+    \singlespacing
     \centering
-    \caption{IPEDS Summary Statistics, PUs Panel 1987--2021}
+    \caption{IPEDS Summary Statistics, Public Universities Panel 1987--2021}
     \makebox[\textwidth][c]{\input{tables/ipeds-summary-fte.tex}}
@@
 IBHED host the information;
-Public Act 96-0266 (effective 1 January 2010) requires that each PU report to the IBHE the base salary and benefits of the president and all administrators, faculty members, and instructors employed by the college or university.\footnote{
+Public Act 96-0266 (effective 1 January 2010) requires that each university report base salary and benefits all administrators, faculty members, and instructors employed by the college or university.\footnote{
     The universities included are all campuses of the eight Illinois public universities: Chicago State University, Eastern Illinois University, Governors State University, Illinois State University, Northeastern Illinois University, Northern Illinois University, Southern Illinois University  (all five campuses), University of Illinois (all four campuses), Western Illinois University.
@@
 This publicly available data provides the basis to build a panel of Illinois PU professors 2010-2021; I define a professor as an individual observation by their first plus last name and university pairing, and link this database to IPEDS information on finances for their employing institution.
+
+The analysis sample represents 16,932 professors in the year 2010 and 15,352 in the year 2021, with summary statistics presented in \autoref{tab:illinois-summary}.
+Additionally, further analysis using the first year of a professor's employment focuses on the subset of professors with observed year of hiring between the years 2011 and 2021, representing 1,778 professors in 2011, and 9,099 in 2021.\footnote{
+    The reasons for restricting to the sample with identified year of hire in the range 2011-2021 is explained in \autoref{sec:iv-model-indiv}.
+}
+Summary statistics for this group, which over-represents lecturers compared to assistant and full professors, can be seen in \autoref{tab:illinois-summary-rolling}.
+
 \begin{table}[h!]
-    \onehalfspacing
+    \singlespacing
     \centering
-    \caption{IBHED Summary Statistics, Professor Panel 2011--2021.}
-    \makebox[\textwidth][c]{\input{tables/illinois-summary-rolling.tex}}
+    \caption{IBHED Summary Statistics, Professor Panel 2010--2021.}
+    \makebox[\textwidth][c]{\input{tables/illinois-summary.tex}}
     \label{tab:illinois-summary}
 \end{table}
-The analysis sample focuses on the subset of professors with observed year of hiring between the years 2011 and 2021, representing 1,778 professors in 2011, and 9,099 in 2021.
-\autoref{tab:illinois-summary} presents the summary statistics for this sample.\footnote{
-    The full data IBHED contain 173,000 observations for every public university professor in Illinois 2010-2021, which representing 16,932 professors in the year 2010 and 15,352 in the year 2021.
-    Summary statistics for the entire sample can be see in Appendix \autoref{tab:illinois-summary-entire}.
-    The reasons for restricting to the sample with identified year of hire in the range 2011-2021 is explained in \autoref{sec:iv-model-indiv}.
-}
 
@@
 
-\autoref{fig:funding} exhibits the trends for the mean of US PUs between the years 1990-2021.
-We see a rise in total revenues received by PUs (from all sources), and a notable increase in mean tuition revenues from \$48 million per year-university to \$150 million per year-university. 
+\autoref{fig:funding} exhibits the trends for the mean of US public universities between the years 1990-2021.
+We see a rise in total revenues received by public universities (from all sources), and a notable increase in mean tuition revenues from \$48 million per year-university to \$150 million per year-university.
 Non-institutional revenues refer to revenues coming from external sources, not raised by internal university activities;\footnote{
-    US universities raise revenues from many different internal activities, including sales of non-degree related educational courses to publishing deals.
+    US universities raise revenues from many internal activities, including sales of nondegree related educational courses to publishing deals.
     The non-institutional revenue measure is constructed to consider only revenues resulting from external decisions, including state, local, and federal appropriations and tuition revenues.
-    Rates of tuition area considered external since most state governments play a large role in deciding tuition rices for PUs \citep{NBERw23736}.
+    Rates of tuition are considered external since most state governments play a large role in deciding tuition rices for public universities \citep{NBERw23736}.
 }
@@
 At the same time, total state appropriations stagnated at around \$100 million per year-university for 1990-2008, fell slightly around 2008 and have not recovered.
-While PUs experienced a financial stagnation, private universities were not exposed to the same constraints, receiving \$37,000 per student in 1990 and \$49,000 in 2021 with no corresponding decline in any specific component.
+While public universities experienced a financial stagnation, private universities were not exposed to the same constraints, receiving \$37,000 per student in 1990 and \$49,000 in 2021 with no corresponding decline in any specific component.
 
@@
     \centering
+    \singlespacing
     \caption{Mean Total Revenues among Public Universities, by Year.}
@@
     \centering
+    \singlespacing
     \caption{Total Student Enrollment, by University Sector, and Year.}
@@
 \end{figure}
-At the same time, student enrollment at PUs rose precipitously.
-6.2 million students were enrolled in PUs in 1990, and this number rose by 47\% to 9.1 million, with most of the increase occurring over the years 2000-2021.
-\autoref{fig:enrollment} shows that total enrollment at private universities has also risen over the same time period, but not as drastic in either relative or absolute terms; the mean public university also grew over this time period, from 9,800 students in 1990 to 11,800 in 2021.
+At the same time, student enrollment at public universities rose precipitously.
+6.2 million students were enrolled in public universities in 1990, and this number rose by 47\% to 9.1 million, with most of the increase occurring over the years 2000-2021.
+\autoref{fig:enrollment} shows that total enrollment at public universities has also risen over the same time period, but not as drastic in either relative or absolute terms; the mean public university also grew over this time period, from 9,800 students in 1990 to 11,800 in 2021.
 This means that per student mean revenues per student (seen in \autoref{fig:mean-funding-fte}) have stagnated across all measures, and particularly fallen from \$11,000 per student in 1990 to less than \$8,000 per student in 2021.
@@
     \centering
+    \singlespacing
     \caption{Total Professor Count per Student, by University Sector, Professor Appointment, and Year.}
@@
 \autoref{fig:fte-perprof} documents the divergence in faculty composition (per student) between the average private and public university 1990-2021.
-Private universities start with  a higher baseline of around 4.5 professors per 100 students, and exhibits yearly variation of less than 0.5 professor per 100 students over the thirty years.
-PUs start with 3.9 professors per student, and this number falls to 3.5 primarily in the 2008-2011 time period.
+Public universities start with  a higher baseline of around 4.5 professors per 100 students, and exhibits yearly variation of less than 0.5 professor per 100 students over the thirty years.
+Public universities start with 3.9 professors per student, and this number falls to 3.5 primarily in the 2008-2011 time period.
 We see a similar difference in baseline, and fall for the years 2008-2011 for assistant professors.
-Private and public universities have similar numbers of associated and full professors before the year 2000, yet this number has fallen by over 20\% in the next 20 years only for PUs: in 2021 the mean PU has 6 fewer full professors per student than the mean private university.
-At the same time period, we see the rise in use of non-tenure track instructor positions (referred to as lecturers from here on), who were employed at similar rates in both sectors in 1990 yet have been utilised by PUs at a higher rate since.
+Private and public universities have similar numbers of associated and full professors before the year 2000, yet this number has fallen by over 20\% in the next 20 years only for public universities: in 2021 the mean PU has 6 fewer full professors per student than the mean private university.
+At the same time period, we see the rise in use of non-tenure track instructor positions (referred to as lecturers from here on), who were employed at similar rates in both sectors in 1990 yet have been utilised by public universities at a higher rate since.
 
@@
 The state of Illinois particularly exemplifies the stagnation in state support for higher education that the rest of the nation has experienced 1990-2021.
-\autoref{fig:illinois-funding} shows the trends for revenues among the seven Illinois public university campuses, including the falling share of state appropriations and substitute towards tuition revenue.
+\autoref{fig:illinois-funding} shows the trends for revenues among the seven Illinois public university campuses, including the falling share of state appropriations and substitution towards tuition revenue.
 
@@
     \centering
+    \singlespacing
     \caption{Mean Revenues among Illinois Public Universities, by Year.}
@@
 }
-While state institutions continues to operate, there were drastic revenue and spending cuts in response to the budget impasse, as it continued through fiscal year 2017, and ended with a new budget restoring revenues to state agencies and universities for the 2018 fiscal year.
-This variation in public university revenues, stemming entirely from political disagreements --- not from state decisions regarding higher education and its finances \citep{young2020squandered} --- and via the state appropriations channel, mean that Illinois public universities exhibit sizable changes in their state approriations over 2011-2021, of similar order to those for the rest of the country over 1990-2021.
+While state institutions continued to operate, there were drastic revenue and spending cuts in response to the budget impasse, as it continued through fiscal year 2017, and ended with a new budget restoring revenues to state agencies and universities for the 2018 fiscal year.
+This variation in public university revenues, stemming entirely from political disagreements --- not from state decisions regarding higher education and its finances \citep{young2020squandered} --- via the state appropriations channel, mean that Illinois public universities exhibit sizable changes in their state appropriations over 2011-2021, of similar order to those for the rest of the country over 1990-2021.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Section on possible outcomes
+\subsection{University Responses to Financial Shocks}
+\label{sec:responses}
+Subsection describing how a university may respond to shocks on its sources of revenues.
+Cite other papers on how this went.
+
+Write about shocks in state appropriations primarily here.
+
+Additional paragraph on how the incumbent professors may be affected.
 
@@
 
-A professor is employed by a university, so it stands to reason that when PUs experience a secular fall in revenue --- analogous to a demand shock in the private sector --- that their human resources decisions will be affected.
+A professor is employed by a university, so it stands to reason that when public universities experience a secular fall in revenue --- analogous to a demand shock in the private sector --- that their human resources decisions will be affected.
 Universities have been shown to freeze faculty hiring and general faculty salaries in response to financial shocks, such as the Great Recession \citep{turner2014impact}.
 
+\textbf{Shock to state appropriations only, as shocks to total revenues/spending requires identification for multiple sources.
+Tuition is an outcome of changes in the state appropriations, so discuss whether it is an outcome or a control.
+Leave footnote that results are relatively similar, so leave out this control.}
+
 Na\"ively we can analyse the relationship between professor-outcomes $Y_{i,t}$ at university $i$ in year $t$ as a result of university revenues per student $X_{i,t}$.
-The exact measure of university revenues, $X_{it}$ is non-institutional revenues, the sume of university funding from outside sources -- local, federal, and state appropriations plus tuition revenue.\footnote{
+The exact measure of university revenues, $X_{it}$ is non-institutional revenues, the sum of university funding from outside sources -- local, federal, and state appropriations plus tuition revenue.\footnote{
     This measure is in line with the \cite{NBERw27885} approach of considering revenues from the state, but does not ignore yearly changes in revenues arising from tuition revenues -- which are shown to be heavily related.
@@
 \cite{NBERw23736,chakrabarti2018effect,NBERw27885} address endogeneity in PU finances by exploiting a shift-share instrument for changes in state-level funding interacted with university reliance on state funding in a base period.
-The system exploits the fact that institutions who rely on state appropriations more will be affected by later shocks to state appropriations.
+The system exploits the fact that institutions who rely on state appropriations more will be affected by state appropriation shocks.
 We start with the shift-share instrument.
@@
 }
-\cite{NBERw27885} notes the tendency for PUs to respond to state funding cuts by increasing reliance on tuition, where \cite{NBERw23736} specifically instruments for tuition revenues with collected information on legislative tuition price controls.
+\cite{NBERw27885} notes the tendency for public universities to respond to state funding cuts by increasing reliance on tuition, where \cite{NBERw23736} specifically instruments for tuition revenues with collected information on legislative tuition price controls.
 As such, it is important in my set-up to explicitly control for tuition revenues in the first-stage.
 
-The first-stage is then as follows, where $W_{i,t}$ represents (log) tuition revenues, including institution and year fixed effects.
+The first-stage is then as follows, including institution and year fixed effects.
 \begin{equation}
     \label{eqn:firststage}
-    X_{i,t} = \eta_i + \zeta_t + \delta Z_{i,t} + \kappa W_{i,t} + \epsilon_{i,t}
+    X_{i,t} = \eta_i + \zeta_t + \delta Z_{i,t} + \epsilon_{i,t}
 \end{equation}
 We note the conditions for exogeneity in the instrument (following the discussion presented by \citealt{NBERw27885}).
-The instrument is exogenous if state policy decisions for funding of, or tuition charges by PUs, are uncorrelated with unobserved changes in the expenditures of any specific college or university in the state.
+The instrument is exogenous if state policy decisions for funding of, or tuition charges by public universities, are uncorrelated with unobserved changes in the expenditures of any specific college or university in the state.
 This assumption is plausible given that the majority of states have multiple (i.e.
@@
 \begin{table}[!h]
-    \onehalfspacing
+    \singlespacing
     \centering
-    \caption{First Stage Estimates, for Total University Revenues by Shocks}
+    \caption{First Stage Estimates, for University Revenues by Appropriation Shock.}
     \makebox[\textwidth][c]{\input{tables/firststage-reg-fte.tex}}
@@
 Column (4) exhibits the worst case, and extreme bias in the appropriation shock first-stage without the tuition control.
-Column (1) represents the estimates for \autoref{eqn:firststage} with fixed effects, and is the preferred form that I proceed with in the following.
+Column (3) represents the estimates for \autoref{eqn:firststage} with fixed effects, omitting the tuition revenue control, and is the preferred form that I proceed with in the following.
 
@@
     \centering
+    \singlespacing
     \caption{Local Projection Estimates for First-Stage \autoref{eqn:firststage}.}
@@
 
-The primary empirical model combines the instrumental variable for stata appropriation shocks with the empirical model for the effects of university revenues at the university-level -- i.e.
+The primary empirical model combines the instrumental variable for state appropriation shocks with the empirical model for the effects of university revenues at the university-level --- i.e.
 parameter $\beta$ in the following.\footnote{
     It is important to note the treatment effect isolated here; the instrumental variables approach identifies the local average treatment effects (LATE) specific to the instrument.
-    So we interpret this treatment effect as the PU's response in employment count and average salaries to revenue changes, changes specific to state appropriation shocks, among the complier group --- i.e.
-universities with any exposure to shocks, assuming no universities increase employment or salaries in response to revenue shocks.
-    \cite{mogstad2021causal} provide a full discussion of the LATE with multiple instruments and thus multiple monotonicity conditions.
+    So we interpret this treatment effect as a university's response in employment count and average salaries to revenue changes, changes specific to state appropriation shocks, among the complier group --- i.e.
+universities with any exposure to shocks, assuming no universities increase employment or salaries in response to negative revenue shocks.
+    \cite{mogstad2021causal} provide a full discussion of the LATE with instruments and multiple monotonicity conditions.
 }
@@
     \label{eqn:secondstage1}
-    X_{i,t} &=& \eta_i + \zeta_t + \delta Z_{i,t} + \kappa W_{i,t} + \epsilon_{i,t} \\
+    X_{i,t} &=& \eta_i + \zeta_t + \delta Z_{i,t} + \epsilon_{i,t} \\
     \label{eqn:secondstage2}
-    Y_{i,t} &=& \alpha_i + \gamma_t + \beta \widehat X_{i,t} + \lambda W_{i,t} + \varepsilon_{i,t}
+    Y_{i,t} &=& \alpha_i + \gamma_t + \beta \widehat X_{i,t} + \varepsilon_{i,t}
 \end{eqnarray}
@@
     \centering
+    \singlespacing
     \caption{Instrument Variables Model for University Finances and Faculty Composition.}
@@
         \node[state] (instrument) at (0,0) {$Z$};
-        \node at (-2.5,0.5){State};
-        \node at (-2.0,0){appropriation};
-        \node at (-2.5,-0.5){shock};
-        \node[state] (confounder) [above=of instrument] {$W$};
-        \node at (-2,2.25) {Tutuion};
-        \node at (-2,1.75) {revenues};
+        \node at (-2.6,0.5){State};
+        \node at (-1.9,0){appropriation};
+        \node at (-2.6,-0.5){shock};
         \node[state] (endogenous) [right=of instrument] {$X$};
@@
         \node[state] (outcome) [right=of endogenous] {$Y$};
-        \node at (5.5,0.25) {Professor};
-        \node at (5.55,-0.25) {per student,};
-        \node at (5.5,-0.75) {Average};
-        \node at (5.55,-1.25) {Salary};
+        \node at (5.2,0.25) {Faculty};
+        \node at (5.6,-0.25) {Composition};
         % Relevance condition
@@
         \path (endogenous) edge (outcome);
-        % Confounding relationship
-        \path (instrument) edge (confounder);
-        \path (confounder) edge (endogenous);
         % Exclusion Restriction
@@
 
-\autoref{fig:SCM-ivmodel} presents the system in graphical format, where tuition revenues $W$ are considered a confounder in the causal effect of state appropriation shocks on total revenues.
+\autoref{fig:SCM-ivmodel} presents the system in graphical form, for the causal effect of state appropriation shocks on university revenues and thus faculty composition.
 
@@
 
-This analysis additionally utilises data regarding individual professors in the Illinois university system, to investigate the effects of changes in university revenues on the individual professors at the universities.
-Yet, redefining the level of outcome requires adjustment to the empirical approach to leverage variation in university finances in the years following an individual joins the university.\footnote{
+This analysis additionally uses data regarding individual professors in the Illinois university system, to investigate the effects of changes in university revenues on the individual professors at the universities.
+Redefining the level of outcome requires adjustment to the empirical approach to leverage variation in university finances for the years following when a professor joins the university.\footnote{
     This formulation follows that presented by \cite{NBERw27885}, where individual student outcomes are analysed via variation in state appropriations after their freshman-year.
-    This constrasts with \autoref{sec:iv-model-uni} and \cite{NBERw23736}, where the unit of analysis is the university-year.
+    This contrasts with \autoref{sec:iv-model-uni} and \cite{NBERw23736}, where the unit of analysis is the university-year.
 }
 
-\autoref{eqn:public-instrument} defines a rolling-share variant of the instrument, $\tilde Z_{j,t}$, where the university's share exposure to state appropriations is based the year a professor joins the university --- and not the base period 1990--1993.
+\autoref{eqn:rolling-instrument} defines a rolling-share variant of the instrument, $\tilde Z_{j,t}$, where the university's state appropriations share exposure is based in the year a professor joins the university --- and not the base period 1990--1993.
 $j$ indexes each professor in year $t$, $\tau(j)$ for the year the professor first joins their Illinois PU.\footnote{
-    Identifying $\tau(j)$ is possible for $j$ by restricting to all professors who are hired in the time period 2011-2021 (i.e. in the years after the first in the panel).
-    Professors who were hired in the year 2010 are impossible to distinguish from those in year preceding, and so are not considered.
+    Identifying $\tau(j)$ is possible for $j$ by restricting to all professors hired 2011-2021 --- i.e. in the years after the start of the panel.
+    It is not possible to discern the hiring year for professors who  were hired in the years preceding 2011, and so the entire sample is only possible to analyse using the base-share in years 1990-1993 formulation (\autoref{tab:facultysalaries-shock-illinois}, \ref{tab:promotion-shock-illinois}).
 }
@@
 
-This approach leverages an insight, made available by level of the data: that an individual professor is affected by changes in university revenues after they have joined the university.\footnote{
-    Notice that \autoref{sec:iv-model-uni} considers a the number of professors employed by the university.
-    So that whether a professor becomes employed at the university is likely affected by the university revenues.
-    The formulation in this section does not consider that outcome, instead taking as given that the professor is employed at the university, and then projecting the effect on the individual given they are employed here.
+This approach leverages an insight, made available by level of the data: that an individual professor is effected by changes in university revenues after they have joined the university.\footnote{
+    Notice that \autoref{sec:iv-model-uni} considers the number of professors employed by the university; whether a professor becomes employed at the university is likely affected by that university's finances.
+    The formulation in this section does not consider whether the professor joins the university, instead taking as given that the professor is employed at the university, and then projecting the effect on the individual given they are employed here.
 }
-Exogeneity and relevance of the rolling-share instrument, $\tilde Z_{j,t}$, follows the same reasoning as that for the base-share instrument, $Z_{i,t}$, discussed in \autoref{sec:approp-shocks}.
-We satisfy the assumptions by noting that none of the Illinois public campuses take the majority of state appropriations, and that the instrument identification strategy relies on exogeneity in changes in state appropriations to individual professor-outcomes, following the year they joined the university.\footnote{
+Exogeneity and relevance of the rolling-share instrument, $\tilde Z_{j,t}$, follows the same reasoning as that for the base-share instrument, $Z_{i,t}$, discussed in \autoref{sec:approp-shocks}.\footnote{
+    The base-share instrument is still used for some outcomes with the individual Illinois professors, where appropriate (\autoref{tab:facultysalaries-shock-illinois}, \autoref{tab:promotion-shock-illinois}).
+}
+We satisfy the assumptions for exogeneity by noting that none of the Illinois public campuses take the majority of state appropriations, and that the instrument identification strategy relies on exogeneity in changes in state appropriations to individual professor-outcomes, following the year they joined the university.\footnote{
     Additionally, within-institution changes resulting from share reliance on state funding may be correlated with unobserved changes in the outcomes, so that \cite{NBERw27885} note the importance of controlling for the base share and state student population.
-    My formulation implicits controls for these factors via the fixed effects; results are relatively similar while including these controls without including fixed effects.
+    The formulation here implicitly controls for these factors via the fixed effects; results are relatively similar while including these controls without including fixed effects.
 }
@@
 \begin{table}[!h]
-    \onehalfspacing
+    \singlespacing
     \centering
@@
     \centering
+    \singlespacing
     \caption{Local Projection Estimates for Professor Count per Student.}
@@
 And yet, identifying the university revenues by the state appropriations shock increases the magnitude of the effect (Column 2) with respect to the association (Column 1), increasing evidence that the two trends are causally linked.
-Between years 1990-1999 IPEDS provided count of professors by explicit tenure designation (opposed to named position as in \autoref{tab:facultycount-shock-reg}), and Appendix \autoref{tab:tenurecount-shock-reg-fte} presents results of the respective regressions by tenure status.
+Between years 1990-1999 IPEDS provided count of professors by explicit tenure designation (opposed to named position as in \autoref{tab:facultycount-shock-reg}), and \autoref{tab:tenurecount-shock-reg-fte} presents results of the respective regressions by tenure status.
 Results are relatively similar, yet imprecise from the much smaller number of observations.
@@
 At the same time, public university enrollment secularly increased over this time period, so that the universities satisfied their teaching obligations with non-tenure track lecturers.
-The mechanism that explains the changes in faculty composition are, however, not clear at this unti of analysis; individual data are needed to delve deeper.
+The mechanism that explains the changes in faculty composition are, however, not clear at this unit of analysis; individual data are needed to delve deeper.
 
@@
 
-The first outcome is total salary for each professor by their position;
-\autoref{tab:facultysalaries-shock-illinois} presents 2SLS estimates for the elasticity of total salaries with respect to university revenues.
-The results are not distinguishable from zero in any case, which implies that salaries for all years of employment at the university are not causally related to changes in university revenues.
+I first examine whether a shock to the university's revenues affects the salaries of professors at the university.
+The first outcome is first-year salary for each professor by their position;
+\autoref{tab:newhiresalaries-shock-illinois-rolling} presents 2SLS estimates for the elasticity of total salaries with respect to university revenues, among the sample with a year of starting employment after 2010.
+New hires of assistant, full, or administrative position professors are not realted to revenues, via a shock to appropriations in their first-year of employment.
+Yet, we see the starting salary of lecturers is 3.8\% higher given an increase in their hiring university's state funding.
 
 \begin{table}[!h]
-    \onehalfspacing
+    \singlespacing
+    \centering
+    \caption{2SLS Estimates for Faculty Salaries, in First-Year, at Illinois Universities.}
+    \makebox[\textwidth][c]{\input{tables/newhiresalaries-shock-illinois-rolling.tex}}
+    \begin{flushleft}
+        \footnotesize
+        \textbf{Note}: Standard errors are clustered at the institution and first year of employment level. 
+    \end{flushleft}
+    \label{tab:newhiresalaries-shock-illinois-rolling}
+\end{table}
+
+Regarding total salary for every year of employment, \autoref{tab:facultysalaries-shock-illinois-rolling} presents estimates for the elasticity of total salary with respect to university revenues, for the entire sample of professors hired after 2010.\footnote{Recall that this is a subsample of the entire data, that the rolling shift-share instrument is only defined for the sample with an observed hiring date in the years 2011-2021.}
+The elasticity is non-distinguishable from zero among any group, which implies that long-term salaries are not related to university revenues.
+\autoref{tab:facultysalaries-shock-illinois} presents the same model, for the sample of all professors and using the instrument based in years 1990-1993, similarly finding no relationship.
+
+\begin{table}[!h]
+    \singlespacing
     \centering
@@
     \end{flushleft}
-    \label{tab:facultysalaries-shock-illinois}
+    \label{tab:facultysalaries-shock-illinois-rolling}
 \end{table}
 
-Yet, it is possible that another measure of salary is impacted by the university revenues: salary in first-year of employment.
-\autoref{tab:newhiresalaries-shock-illinois} presents 2SLS for the elasticity of total salary in the first year of a professor' employment  with respect to university revenues.
-The elasticity is non-distinguishable from zero among only lecturers: a 10\% increase in university revenues (via state appropriation shocks) leads to a 9.5\% increase in total salary paid to lecturers.
-
-\begin{table}[!h]
-    \onehalfspacing
-    \centering
-    \caption{2SLS Estimates for Faculty Salaries, in First-Year, at Illinois Universities.}
-    \makebox[\textwidth][c]{\input{tables/newhiresalaries-shock-illinois-rolling.tex}}
-    \begin{flushleft}
-        \footnotesize
-        \textbf{Note}: Standard errors are clustered at the institution and first year of employment level. 
-    \end{flushleft}
-    \label{tab:newhiresalaries-shock-illinois}
-\end{table}
-
-Note that professor position (assistant, associate, full) is also an outcome in this set-up by the promotion channel.
-\autoref{tab:promotion-shock-illinois} describes estimates where the outcome is rate of promotion within each professor group.
+Note that professor position (assistant, associate, full) is also an outcome, thanks to the promotion channel: a university may be less able to promote their professors to higher paying positions when their revenues are constrained.
+\autoref{tab:promotion-shock-illinois-rolling} shows estimates where the outcome is rate of promotion within each professor group.
 The lecturer group describes the sample of professors who were ever listed as a non-tenure track lecturer and the binary for promotion equals one in a year they achieved an assistant professor position; assistant the same for the assistant to associate professor promotion; associate same for associate to full promotion.
-
-There is no discernible relationship between revenues and promotion among lecturers, as only a small group of professors achieves the lecturer to assistant professor promotion in this sample.\footnote{
-    749 in this sample make the transition lecturer to assistant professor, while there are 9,316 distinct lecturers in the data sample.
+There is no discernible relationship between revenues and promotion among any position group.\footnote{
+    \autoref{tab:promotion-shock-illinois} shows estimates using the base-year instrument among the entire sample, finding the same results.
 }
-There is an imprecise positive relationship between rate of promotion assistant to associate professor, and no clear relationship between promotion rate among associate professors.
 Furthermore, local projection estimates (\autoref{fig:promoted-illinois-lp}) show that the there is no promotion effect, among all professors, for any of the following years.
-Lastly, Appendix \autoref{tab:facultyleaving-shock-illinois} describes estimates for rate of exit from the employing university among Illinois professors, and finds no relationship with university revenues.
+Lastly, \autoref{tab:facultyleaving-shock-illinois-rolling}, \ref{tab:facultyleaving-shock-illinois} describes estimates for rate of exit from the employing university among Illinois professors, and finds no relationship with university revenues.
 
@@
 The University of Illinois\footnote{
-    The University of Illinois inclves three public university campuses: Urbana-Champaign, Chicago, Springfield.
-} did not receive the allocated state appropriations from the state of Illinois on time, so enacted cost-cutting measures to stay fiscally solvant.
+    The University of Illinois includes three public university campuses: Urbana-Champaign, Chicago, Springfield.
+} did not receive the allocated state appropriations from the state of Illinois on time, so enacted cost-cutting measures to stay fiscally solvent.
 The university system placed a hold on all hiring for filling state-funded positions and promotions.
@@
     \centering
+    \singlespacing
     \caption{Trends in New Hires at Illinois Public Universities 2011-2021.}
@@
 Furthermore, private universities in this country are not exposed to these same financial headwinds.
-If public universities' hiring of faculty is disrupted by changes in state appropriations, then we can expect that private universities will benefit (in a compeitive sense) by being able to hire better professors from relatively resource constrained public universities.
-While public universities continue to educate the majority of higher education students in the US, and disproptionately students from low-income backgrounds and minorities, we should worry about the effects that restricting their financial support has on both the quality and distribution of education outcomes in this country.
+If public universities' hiring of faculty is disrupted by changes in state appropriations, then we can expect that private universities will benefit (in a competitive sense) by being able to hire better professors from relatively resource constrained public universities.
+While public universities continue to educate the majority of higher education students in the US, and disproportionately students from low-income backgrounds and minorities, we should worry about the effects that restricting their financial support has on the composition of distribution of higher education and research in this country.
 
-\subsubsection{Threats to Identification}
+\begin{figure}[h!]
+    \centering
+    \singlespacing
+    \caption{Faculty Profile at Illinois Public Universities 2011-2021.}
+    \begin{subfigure}[b]{0.495\textwidth}
+        \centering
+        \caption{Faculty Count per Enrolled Student, Total for the State.}
+        \includegraphics[width=\textwidth]{figures/faculty-count-illinois.png}
+        \label{fig:faculty-count-illinois}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.495\textwidth}
+        \centering
+        \caption{Mean Salary for All Faculty, \$ 2021 CPI-U.}
+        \includegraphics[width=\textwidth]{figures/faculty-salary-illinois.png}
+        \label{fig:faculty-salary-illinois}
+    \end{subfigure}
+    \label{fig:faculty-illinois}
+\end{figure}
 
-\cite{blandhol2022tsls} provide reason to doubt the LATE interpretation of two-stage least squares estimates with multiple layers of fixed effects (i.e. both individual and time fixed effects).
-Needs comment, and to what extent should this question the empirical strategy.
+%\subsubsection{Threats to Identification}
+%
+%\cite{blandhol2022tsls} provide reason to doubt the LATE interpretation of two-stage least squares estimates with multiple layers of fixed effects (i.e. both individual and time fixed effects).
+%Needs comment, and to what extent should this question the empirical strategy.
 
@@
 %% Appendix section
+% Set-up the section.
 \appendix
+\setcounter{table}{0}
+\renewcommand{\thetable}{A\arabic{table}}
+\setcounter{figure}{0}
+\renewcommand{\thefigure}{A\arabic{figure}}
+
+% Start appendix
 \section{Appendix}
 \label{appendix}
-A number of statistical packages, for the R language \citep{R2022}, made the empirical analysis for this paper possible.
+A number of statistical packages, for the $R$ language \citep{R2022}, made the empirical analysis for this paper possible.
 \begin{itemize}
@@
 \begin{table}[h!]
-    \onehalfspacing
+    \singlespacing
     \centering
     \caption{IBHED Summary Statistics, Entire Professor Panel 2010--2021.}
-    \makebox[\textwidth][c]{\input{tables/illinois-summary.tex}}
-    \label{tab:illinois-summary-entire}
+    \makebox[\textwidth][c]{\input{tables/illinois-summary-rolling.tex}}
+    \label{tab:illinois-summary-rolling}
     \begin{flushleft}
         \footnotesize
-        \textbf{Note}: \autoref{tab:illinois-summary} presents the summary statistics for professors with an observaed year of first-year at their university, while this table exhibits the data regarding every professor at an Illinoic public university 2010--2021.
+        \textbf{Note}: This table presents the summary statistics for professors with an observed year of first-year at their university.
     \end{flushleft}
@@
 \begin{table}[H]
-    \onehalfspacing
+    \singlespacing
     \centering
@@
     \centering
+    \singlespacing
     \caption{Local Projection Estimates for First-Stage \autoref{eqn:secondstage1_indiv}.}
@@
 
-\begin{figure}[h!]
+\begin{figure}[H]
     \centering
-    \caption{Local Projection Estimates for Professor Average Salary, within Each University.}
+    \singlespacing
+    \caption{Local Projection Estimates for Professor Salary, Mean within Each University.}
     \includegraphics[width=0.6\textwidth]{figures/all-salaries-lp.png}
@@
 \begin{table}[H]
-    \onehalfspacing
+    \singlespacing
     \centering
-    \caption{2SLS Estimates for Faculty Exit Rate at Illinois Universities.}
-    \makebox[\textwidth][c]{\input{tables/facultyleaving-shock-illinois-rolling.tex}}
+    \caption{2SLS Estimates for Faculty Salaries at Illinois Universities.}
+    \makebox[\textwidth][c]{\input{tables/facultysalaries-shock-illinois.tex}}
+    \begin{flushleft}
+        \footnotesize
+        \textbf{Note}: Standard errors are clustered at the institution-year level.
+    \end{flushleft}
+    \label{tab:facultysalaries-shock-illinois}
+\end{table}
+
+\begin{table}[H]
+    \singlespacing
+    \centering
+    \caption{2SLS Estimates for Faculty Promotion Rate at Illinois Universities, using Rolling Instrument.}
+    \makebox[\textwidth][c]{\input{tables/promotion-shock-illinois-rolling.tex}}
     \begin{flushleft}
@@
     \end{flushleft}
-    \label{tab:facultyleaving-shock-illinois}
+    \label{tab:promotion-shock-illinois-rolling}
 \end{table}
 
-\begin{table}[!h]
-    \onehalfspacing
+\begin{table}[H]
+    \singlespacing
     \centering
-    \caption{2SLS Estimates for Faculty Promotion Rate at Illinois Universities.}
-    \makebox[\textwidth][c]{\input{tables/promotion-shock-illinois-rolling.tex}}
+    \caption{2SLS Estimates for Faculty Promotion Rate at Illinois Universities, using Base-Year Instrument.}
+    \makebox[\textwidth][c]{\input{tables/promotion-shock-illinois.tex}}
     \begin{flushleft}
@@
     \centering
+    \singlespacing
     \caption{Local Projection Estimates for Promotion Rate at Illinois Universities.}
@@
 
+\begin{table}[H]
+    \singlespacing
+    \centering
+    \caption{2SLS Estimates for Faculty Exit Rate at Illinois Universities, using Rolling Instrument.}
+    \makebox[\textwidth][c]{\input{tables/facultyleaving-shock-illinois-rolling.tex}}
+    \begin{flushleft}
+        \footnotesize
+        \textbf{Note}: Standard errors are clustered at the institution and first year of employment level.
+    \end{flushleft}
+    \label{tab:facultyleaving-shock-illinois-rolling}
+\end{table}
+
+\begin{table}[H]
+    \singlespacing
+    \centering
+    \caption{2SLS Estimates for Faculty Exit Rate at Illinois Universities, using Base-Year Instrument.}
+    \makebox[\textwidth][c]{\input{tables/facultyleaving-shock-illinois.tex}}
+    \begin{flushleft}
+        \footnotesize
+        \textbf{Note}: Standard errors are clustered at the institution and first year of employment level.
+    \end{flushleft}
+    \label{tab:facultyleaving-shock-illinois}
+\end{table}
+
 \end{document}

### 2022-08-09 programs/analysis/individual-shock-rolling.R

@@ -1,5 +1,6 @@
 #!/usr/bin/R
 ## Senan Hogan-Hennessy, 3 August 2022
+print(c(Sys.time(), Sys.Date()))
 ## IV for Prof's outcomes, using Illinois data.
 ## Using a rolling-share variant of the instrument.
 library(tidyverse) ## functions for data manipulation and visualization
@@ -16,7 +17,7 @@ digits.no <- 3
 
 # Size for figures
 fig.width <- 10
-fig.height <- fig.width * 0.75
+fig.height <- fig.width * 0.85
 
 
 # Load data --------------------------------------------------------------------
@@ -284,7 +285,7 @@ firststage_approp_tuit_noFE.fstat <-
 stargazer(
     firststage_approp_tuit.reg, firststage_approp_tuit_noFE.reg,
     firststage_approp.reg, firststage_approp_noFE.reg,
-    dep.var.caption = "Dependent Variable: Non-institutional Revenues",
+    dep.var.caption = "Dependent Variable: State Funding",
     dep.var.labels.include = FALSE,
     digits = digits.no,
     digits.extra = digits.no,
@@ -381,7 +382,7 @@ stargazer(
     omit = "factor|count|year",
     intercept.bottom = TRUE,
     order = c(2, 1, 3),
-    covariate.labels = c("Non-inst. Revenues", "Tuition Revenue", "Constant"),
+    covariate.labels = c("State Funding", "Tuition Revenue", "Constant"),
     omit.stat = c("LL", "ser", "aic", "wald", "adj.rsq"),
     star.cutoffs = NA,
     header = FALSE, float = FALSE, no.space = TRUE,
@@ -470,7 +471,7 @@ stargazer(
     omit = "factor|count|year",
     intercept.bottom = TRUE,
     order = c(2, 1, 3),
-    covariate.labels = c("Non-inst. Revenues", "Tuition Revenue", "Constant"),
+    covariate.labels = c("State Funding", "Tuition Revenue", "Constant"),
     omit.stat = c("LL", "ser", "aic", "wald", "adj.rsq"),
     star.cutoffs = NA,
     header = FALSE, float = FALSE, no.space = TRUE,
@@ -561,7 +562,7 @@ stargazer(
     omit = "factor|count|year",
     intercept.bottom = TRUE,
     order = c(2, 1, 3),
-    covariate.labels = c("Non-inst. Revenues", "Tuition Revenue", "Constant"),
[…]

### 2022-08-09 programs/analysis/individual-shock.R

@@ -1,5 +1,6 @@
 #!/usr/bin/R
 ## Senan Hogan-Hennessy, 3 August 2022
+print(c(Sys.time(), Sys.Date()))
 ## IV for Prof's outcomes, using Illinois data.
 library(tidyverse) ## functions for data manipulation and visualization
 library(lfe)
@@ -15,7 +16,7 @@ digits.no <- 3
 
 # Size for figures
 fig.width <- 10
-fig.height <- fig.width * 0.75
+fig.height <- fig.width * 0.85
 
 
 # Load data --------------------------------------------------------------------
@@ -254,7 +255,7 @@ firststage_approp_tuit_noFE.fstat <-
 stargazer(
     firststage_approp_tuit.reg, firststage_approp_tuit_noFE.reg,
     firststage_approp.reg, firststage_approp_noFE.reg,
-    dep.var.caption = "Dependent Variable: Non-institutional Revenues",
+    dep.var.caption = "Dependent Variable: State Funding",
     dep.var.labels.include = FALSE,
     digits = digits.no,
     digits.extra = digits.no,
@@ -351,13 +352,82 @@ stargazer(
     omit = "factor|count|year",
     intercept.bottom = TRUE,
     order = c(2, 1, 3),
-    covariate.labels = c("Non-inst. Revenues", "Tuition Revenue", "Constant"),
+    covariate.labels = c("State Funding", "Tuition Revenue", "Constant"),
     omit.stat = c("LL", "ser", "aic", "wald", "adj.rsq"),
     star.cutoffs = NA,
     header = FALSE, float = FALSE, no.space = TRUE,
     omit.table.layout = "n", notes.append = FALSE,
     type = "text",
     out = "../../text/tables/facultysalaries-shock-illinois.tex")
+
+# Investigate count + pay for new professors, by position, in each year Illinois
+faculty_yearly.data <- reg.data %>%
+    mutate(
+        position = ifelse(lecturer == 1, "lecturer",
+            ifelse(assistant == 1, "assistant",
+                ifelse(full == 1, "full",
+                    ifelse(administrator == 1, "administrator", NA))))) %>%
+    # Get the uni-yearly count (enrollment is at the uni leevel, so is averaged)
+    group_by(year, position, unitid) %>%
+    summarise(prof_count = n(),
+        enrollment_reported = mean(enrollment_reported, na.rm = TRUE),
+        totalsalary_real =
[…]

### 2022-08-09 programs/analysis/ipeds-public-private.R

@@ -1,6 +1,7 @@
 #!/usr/bin/R
 ## Senan Hogan-Hennessy, 10 June 2022
 ## Investigate differences between public and private uni prof composition.
+print(c(Sys.time(), Sys.Date()))
 set.seed(47)
 # Functions for data manipulation and visualisation
 library(tidyverse)
@@ -114,21 +115,21 @@ ggsave("../../text/figures/enrollment-mean.png",
 
 # Compare to IPEDS provided table
 # https://nces.ed.gov/ipeds/search/ViewTable?tableId=10102&returnUrl=%2Fipeds%2Fsearch
-ipeds.data %>%
-    group_by(year, public, fouryear) %>%
-    summarise(
-        all_profmeansalary_real = weighted.mean(all_profmeansalary_real,
-            weights = all_prof_count, na.rm = TRUE),
-        full_profmeansalary_real = weighted.mean(full_profmeansalary_real,
-            weights = lecturer_prof_count, na.rm = TRUE),
-        assistant_profmeansalary_real = weighted.mean(assistant_profmeansalary_real,
-            weights = assistant_prof_count, na.rm = TRUE),
-        lecturer_profmeansalary_real = weighted.mean(lecturer_profmeansalary_real,
-            weights = full_prof_count, na.rm = TRUE),
-        full_prof_count = sum(full_prof_count, na.rm = TRUE),
-        all_prof_count =  sum(all_prof_count, na.rm = TRUE)) %>%
-    filter(year %in% 2018:2019) %>%
-    View()
+#ipeds.data %>%
+#    group_by(year, public, fouryear) %>%
+#    summarise(
+#        all_profmeansalary_real = weighted.mean(all_profmeansalary_real,
+#            weights = all_prof_count, na.rm = TRUE),
+#        full_profmeansalary_real = weighted.mean(full_profmeansalary_real,
+#            weights = lecturer_prof_count, na.rm = TRUE),
+#        assistant_profmeansalary_real = weighted.mean(assistant_profmeansalary_real,
+#            weights = assistant_prof_count, na.rm = TRUE),
+#        lecturer_profmeansalary_real = weighted.mean(lecturer_profmeansalary_real,
+#            weights = full_prof_count, na.rm = TRUE),
+#        full_prof_count = sum(full_prof_count, na.rm = TRUE),
+#        all_prof_count =  sum(all_prof_count, na.rm = TRUE)) %>%
+#    filter(year %in% 2018:2019) %>%
+#    View()
[…]

### 2022-08-09 programs/analysis/ipeds-shock.R

@@ -1,6 +1,7 @@
 #!/usr/bin/R
 ## Senan Hogan-Hennessy, 3 August June 2022
 ## IV for Prof's salaries, using IPEDS data.
+print(c(Sys.time(), Sys.Date()))
 library(tidyverse) # Functions for data manipulation and visualization
 library(lfe) # Functions for fast linear models with IV + FEs
 library(plm) # Functions for panel data
@@ -16,7 +17,7 @@ digits.no <- 3
 
 # Size for figures
 fig.width <- 10
-fig.height <- fig.width * 0.75
+fig.height <- fig.width * 0.85
 
 
 # Load data sources ------------------------------------------------------------
@@ -128,8 +129,8 @@ mean_funding_fte.data <- reg.data %>%
     summarise(
         totalrevenues_real =
             mean(totalrevenues_real / enrollment_reported, na.rm = TRUE),
-        nonauxrevenues_real =
-            mean(nonauxrevenues_real / enrollment_reported, na.rm = TRUE),
+        #nonauxrevenues_real =
+        #    mean(nonauxrevenues_real / enrollment_reported, na.rm = TRUE),
         stateappropriations_real =
             mean(stateappropriations_real / enrollment_reported, na.rm = TRUE),
         tuitionrev_real =
@@ -153,7 +154,7 @@ mean_funding_fte.graph <- mean_funding_fte.data %>%
     scale_colour_discrete(name = "",
         breaks = c("totalrevenues_real", "nonauxrevenues_real",
         "stateappropriations_real", "tuitionrev_real"),
-        labels = c("Total", "Non-inst.", "State", "Tution"))
+        labels = c("Total", "Non-aux.", "State", "Tuition"))
 # Save this plot
 ggsave("../../text/figures/mean-funding-fte.png",
     plot = mean_funding_fte.graph,
@@ -165,8 +166,8 @@ mean_funding.data <- reg.data %>%
     summarise(
         totalrevenues_real =
             mean(totalrevenues_real, na.rm = TRUE) / (10^6),
-        nonauxrevenues_real =
-            mean(nonauxrevenues_real, na.rm = TRUE) / (10^6),
+        #nonauxrevenues_real =
+        #    mean(nonauxrevenues_real, na.rm = TRUE) / (10^6),
         stateappropriations_real =
             mean(stateappropriations_real, na.rm = TRUE) / (10^6),
         tuitionrev_real =
@@ -190,23 +191,31 @@ mean_funding.graph <- mean_funding.data %>%
[…]

### 2022-08-09 programs/data-build/illinois-build.R

@@ -1,5 +1,6 @@
 #!/usr/bin/R
 ## Senan Hogan-Hennessy, 12 April 2022
+print(c(Sys.time(), Sys.Date()))
 ## Put raw Illinois data to readable format.
 # functions for data manipulation and visualisation
 library(tidyverse)

### 2022-08-09 programs/data-build/urban-ipeds-build.R

@@ -1,6 +1,7 @@
 #!/usr/bin/R
 ## Senan Hogan-Hennessy, 21 December 2021
 ## Build IPEDS data with the Urban Inst's package
+print(c(Sys.time(), Sys.Date()))
 ## https://educationdata.urban.org/documentation/colleges.html
 ## https://urbaninstitute.github.io/education-data-package-r/
 library(tidyverse) ## functions for data manipulation and visualization
@@ -349,12 +350,3 @@ urban_ipeds.data %>%
 urban_ipeds.data %>%
     filter(fouryear == 1, public == 1, forprofit == 0) %>%
     write_csv("../../data/urban-ipeds/urban-clean-publicunis.csv")
-
-#! WHich years are covered?
-print("all years")
-urban_ipeds.data %>% pull(year) %>% table()
-print("variables")
-urban_ipeds.data %>% filter(!is.na(enrollment_fte)) %>% pull(year) %>% max(na.rm = TRUE)
-urban_ipeds.data %>% filter(!is.na(totalrevenues_real)) %>% pull(year) %>% max(na.rm = TRUE)
-urban_ipeds.data %>% filter(!is.na(nonauxspending_real)) %>% pull(year) %>% max(na.rm = TRUE)
-urban_ipeds.data %>% filter(!is.na(appropriationshock_perEnroll_real)) %>% pull(year) %>% table()

### 2022-08-09 programs/run.sh

@@ -1,29 +1,29 @@
 #!/bin/bash
 ## Senan Hogan-Hennessy, 17 August 2022
 ## Instructions file to run scripts in order, to generate analyse IPEDS data.
+## "R CMD BATCH" generates .Rout files showing console output of each script.
 
 # Build data sets of relevance
 cd data-build
 # Build uni-level analysis data-set from mix of raw IPEDS and Urban Inst. data
-Rscript urban-ipeds-build.R
+R CMD BATCH --no-save urban-ipeds-build.R
 # Build individual-level analysis data-set from raw data on Illinois Profs
-Rscript illinois-build.R
+R CMD BATCH --no-save illinois-build.R
 cd ..
 
 # Statistical analysis code
 cd analysis
 # Generate graphs motivating the introduction.
-Rscript ipeds-public-private.R
+R CMD BATCH --no-save ipeds-public-private.R
 # Analysis and regressions at the institution-level.
-Rscript ipeds-shock.R
+R CMD BATCH --no-save ipeds-shock.R
 # Analysis and regressions at the individual-level, with base-year shock IV.
-Rscript individual-shock.R
+R CMD BATCH --no-save individual-shock.R
 # Analysis and regressions at the individual-level, with rolling shock IV.
-Rscript individual-shock-rolling.R
+R CMD BATCH --no-save individual-shock-rolling.R
 cd ..
 
 # Compile paper, with outputs of progams as inputs for TeX files
 cd text
 latexmk -pdf paper.tex
 latexmk -c
-rm paper.bbl

### 2022-08-12 text/paper.tex
First draft.

@@
 % Title
-\title{Stagnating Public University Finances and Faculty Composition}
+\title{Stagnating State Support for Higher Education and Its Effect on Faculty Composition}
 \date{\today}
 \rhead{\today}
-\lhead{Stagnating Public University Finances and Faculty Composition}
+\lhead{State Support for Higher Education and Faculty Composition}
 % Begin
@@
 \noindent
-State support for public universities has stagnated the last thirty years, while enrollment has secularly increased.
-At the same time, public universities have reduced the number of professors per student on tenure-track or with tenure, while increasing their employment of contingent lecturers.
-This analysis relates changes in the faculty composition at public universities to the changes in per student university revenues from state appropriations, addressing endogeneity in state appropriation decisions by a shift-share instrumental variables approach.
-An increase of university revenues of 10\%, via a change in state appropriations, increases the number of assistant professors per student by 6.6\% and full professors by 5.3\%, yet decreases the number of lecturers per student by 14.3\%.
-Analysis of all the professors hired at Illinois public universities 2011-2021 implies that incumbent professors are not affected by the changes in university revenues, so that changes in faculty composition arises by changes in hiring practices at public universities.
-These results show the compositional effects of stagnating state support for higher education, and raise questions about the direction that public education heads as these financial headwinds continue.
+State support for higher education has stagnated on a per student level the last thirty years, while public universities have reduced the number of tenure-track or with tenured professors per student, and increased their employment of contingent lecturers.
+This analysis relates these changes in the faculty composition at public universities to the fall in per student university revenues from state appropriations.
+A shift-share instrumental variables approaches addresses endogeneity in decisions for state appropriations to higher education, by exploiting the yearly change in state support with an institution's reliance on that support.
+A decrease in state funding of 10\%, via a shock to state appropriations, decreases the number of assistant professors per student at a public university by 1.4\% and full professors by 1.2\%, yet increases the number of lecturers per student by 4.4\%.
+Analysis of all the professors hired at Illinois public universities 2011-2021 shows that incumbent professors are not affected by the changes in state support, implying that these changes in faculty composition arose by lower rate of hiring of tenure-track and tenured professors at public universities.
+These results show one of the long-term effects of stagnating state support for higher education, and raises questions about the direction that public education heads as these financial headwinds show no sign of dissipating.
 
@@
 Public universities educate the majority of higher education students in the US, yet have experienced a secular decline in state funding (per student) the last three decades.
-This decline has been shown to lead to worse education and later-life outcomes for students \citep{NBERw23736,NBERw27885}, yet it is not clear what mechanisms drive these effects or how public universities are affected as institutions.
-I show that four-year, degree-granting public universities have experienced falling state funding, leading to falling revenues per student, effecting the composition of professors within public universities in general.
+This decline has been shown to lead to worse education and later-life outcomes for students \citep{NBERw23736,NBERw27885}, yet it is not clear what mechanisms drive these effects, or how public universities are affected as institutions.
+I show that four-year, degree-granting public universities have experienced falling state funding per student, and this affected their composition of faculty.
 Falls in public university revenues from the state, instrumented by state-level finance shocks, lead to a fall in the number of tenure-track or tenured professors per student within a university, an increase in the number of lecturers, and overall fall in the number of faculty per student.
-US private universities were not exposed to such financial constraints during this same time period, and do not exhibit the substitution away from tenure-track and tenured professors, so that the stagnating state support for public universities has implications for the structure of higher education instruction and research in the US.
+US private universities were not exposed to such financial constraints during this same time period, and do not exhibit the substitution away from tenure-track and tenured professors, so that the stagnating state support for public universities has implications for the wider structure of higher education instruction and research in the US.
 
-US Universities are widely considered the highest performing in the world, yet there are consequential differences between its universities that operate in the private sector and those established by state governments.
-public universities are subject to numerous state-level administrative laws, and rely on their state governments for funding: an average PU received around \$11,600 per enrolled student in 1990, yet only \$8,300 per enrolled student in 2021.\footnote{
-    2021 figures are relatively similar to those for year preceding, despite the Covid-19 pnademic and associated financial shocks, since 2021 refers to the academic year 2019-2021 and so the yearly budget was decided in 2019.
-    This paper does not directly analyse budget shocks related to the Covid-19 pandemic, or any other events in 2021 and following.
-}
-This fall is driven by stagnating funding provided by state governments, while total enrollment among public universities rose by 46\% over the same time period.
-At the same time, the number of professors per student at public universities stayed relatively stable at 0.045, yet fell from around 0.04 to around 0.35 professors per student at public universities, driven by a fall of over 20\% in the ratio of associate and full (i.e.
-tenured) professors per student at public universities.\footnote{
-    \autoref{fig:fte-perprof} and \autoref{sec:trends} document these trends between private and public universities in full detail.
-}
+US universities are widely considered the highest performing in the world, yet there are consequential differences between its universities that operate in the private sector and those established by state governments.
+Public universities are subject to numerous state-level administrative laws, and rely on their state governments for funding: an average public university received around \$11,600 per enrolled student in 1990, yet only \$8,300 per enrolled student in 2021.
+This fall is driven by stagnating funding provided by state governments, while enrolment among public universities rose by 46\% over the same time period.
+At the same time, the number of professors per student at public universities stayed relatively stable at 0.045, yet fell from around 0.04 to around 0.35 professors per student at public universities, driven by a fall of over 20\% in the ratio of associate and full professors per student at public universities.
 
-I use a variant of the \cite{NBERw23736,NBERw27885} shift-share instrument for PU revenues from shocks to state appropriations to identify the effect on the faculty composition.
-State appropriations for public universities are plausibly endogenous to many outcomes: state governments decide on yearly budgets for their higher education sector, and this process can be influenced by local financial conditions, or even public (and political representatives') perceptions of the state's higher education system.
-The shift-share instrument interacts reliance on state funding in a base period with the yearly total appropriations per student in the entire state, to exploit both changes in state support for higher education and how much each university relies on that support.
+I use a variant of the \cite{NBERw23736,NBERw27885} shift-share instrument for state funding from shocks to appropriations for the entire state to identify the effect on the faculty composition.
+State appropriations for public universities are plausibly endogenous to many outcomes: state governments decide on yearly budgets for their higher education sector, and this process can be influenced by local financial conditions, or even public perceptions of the state's higher education system.
+The shift-share instrument interacts reliance on state funding in a base period with the yearly total appropriations per student in the entire state, to exploit both changes in support for higher education and how much each university relies on that support.
 
-Findings show that negative changes in university revenues affects faculty composition at public universities, away from tenured and tenure-track professors towards non-tenure track lecturer positions.
-A reduction of 10\% in university revenues, via a shock to state appropriations, leads to a fall in 6.6\% in the number of assistant professors per student within a university, 5.3\% for full (tenured) professors, and a reduction in 14.3\% for lecturers.
-Additionally, I use individual level data on professors who were hired at Illinois public universities 2011-2021 to investigate whether the shocks to university revenues affected incumbent professors at public universities.
-Incumbent professors were not meaningfully affected, in terms of total salary, promotion rate, or rate of leaving the Illinois public university system, which implies that changes in faculty composition came about by changes in the composition of hiring new professors at the university level.
+Findings show that negative changes in state funding affects faculty composition at public universities, away from tenured and tenure-track professors towards non-tenure track lecturer positions.
+A reduction of 10\% in state funding, via a shock to state appropriations, leads to a fall in 1.4\% in the number of assistant professors per student within a university, 1.2\% for full (tenured) professors, and an increas in 4.4\% the number of lecturers per student.
+Additionally, I use individual level data on professors at Illinois public universities over 2010-2021 to investigate whether the shocks to state funding affected incumbent professors at public universities.
+Incumbent professors were not meaningfully affected, in terms of total salary, promotion rate, or rate of leaving the Illinois public university system, yet lecturers' first year slary decreased by 3.8\% in response to a fall in state funding of 10\%.
+These findings imply that changes in faculty composition came about by changes in the composition of hiring new professors at the university level.
 
-\cite{NBERw23736} first study the effect of (changes in) public universities' finances on student outcomes by isolating changes in public universities' expenditures on student instruction, addressing endogeneity of expenditures decisions by instrumenting expenditures by an instrument for state appropriations and tuition caps.
-\cite{NBERw23736} find that increases in PU spending (via state appropriations) increases enrolment and degree completion among students, but corresponding changes in attendance price (via tuition caps/freezes) do not have an effect.
-\cite{chakrabarti2018effect,NBERw27885} use the same instrument for PU revenues at the state level, combined with richer information on students' individual-level outcomes, to show that increases in state appropriations lead to lower time to completion and debt levels among students later in life.
+\cite{NBERw23736} first study the effect of (changes in) public universities' finances on student outcomes by isolating changes in public universities' expenditures on student instruction, addressing endogeneity of expenditures decisions by instrumenting expenditures with state appropriation shocks and limits on tuition price.
+\cite{NBERw23736} find that increases in public university spending (via appropriation shocks) increases enrolment and degree completion among students, but corresponding changes in attendance price (via limits on tuition price) do not have an effect.
+\cite{chakrabarti2018effect,NBERw27885} use the same instrument for public university revenues at the state level, combined with rich information on students' individual-level outcomes, to show that increases in state appropriations lead to degree lower completion time and later-life student debt.
 \cite{bound2019public} use another variant of the same approach to show that state-level higher education spending cuts induced public universities to shift toward tuition as their primary source of revenue, and shift institutional resources towards students who pay higher tuition rates.
@@
 
-University professors the core component of this country's educational and research/innovation university system, yet their composition within universities and its relationship with higher education finances has so far not been studied.
-\cite{brown2014endowment} presents the closest example by studying how university endowments react to negative financial shocks, exhibiting an association between negative endowment shocks and a fall in tenure-system faculty employed at the university.
-\cite{abe2015implications} present a model for university administrator decision-making, and posit that a large component of variation in faculty salaries arises from university politics, and not only economic factors such as outside.
+Faculty are a core component of this country's educational and research-innovation university system, yet their composition within universities, and its relationship with higher education finances has so far not been studied.
+\cite{brown2014endowment} presents the closest example by studying how university endowments react to negative financial shocks, exhibiting an association between negative endowment shocks and a fall in tenure-system faculty employed at private universities.
+\cite{abe2015implications} present a model for university administrator decision-making, and posit that a large component of variation in faculty salaries arises from university politics, and not only economic factors such as outside options.
 \cite{johnson2009jep,NBERc13879} both empirically document allocation of faculty between research and instruction, and \cite{hemelt2021math} document differences in instructional cost (per student) between departments, primarily thanks to department differences in class size and faculty salaries.
-\cite{turner2014impact} document how universities reacted to the Great Recession of 2008, a severe financial shock to state finances, including implementing hiring freezes.
-Indeed, Cornell University\footnote{
-    Cornell University is, indeed, only partially public, so that this information is not free information, and cannot be presented with aggregate or verified sources of public data on such salaries.
-} implemented both hiring freezes and even a nominal salary reduction for professors, in anticipation of oncoming financial shock in early 2021.\footnote{
-    The salary cut was not permanent, as the oncoming financial shock turned out to not be as serious as projected, and the salary cut was returned to professors in the year 2021.
-}
+\cite{turner2014impact} document how universities reacted to the Great Recession of 2008, a severe financial shock to state finances and support for higher education.
 
-The number of professors, in absolute terms as well as relative to number of students in the university, may be particularly important for quality of instruction and thus educational outcomes.
-\cite{angrist1999using} first causally show that reducing class size induces an increase in test scores for school-age children, yet the magnitude of the effect is not as clear in the university setting.
-\cite{bandiera2010heterogeneous} use a fixed effect model, combined with multiple individual observations in a long panel, to find that UK university students perform worse academically in particularly large classes, and the effect is particularly largest for students at the top of the academic distribution.
-And for the composition of faculty teaching university classes, 
-\cite{bettinger2010does,figlio2015tenure} find US students have better enrollment and learning outcomes from courses taught by contingent or adjunct (i.e.
+The number of professors, both in absolute terms and relative to number of students, may be particularly important for quality of instruction and thus educational outcomes.
+\cite{angrist1999using} first causally show that reducing class size (via religious ruling) induces an increase in test scores for school-age children, yet the magnitude of this effect is not as clear in the higher education setting.
+\cite{bandiera2010heterogeneous} use a fixed effect model, combined with multiple individual observations in a long panel, to find that UK university students perform worse academically in particularly large classes, and the difference is largest for students at the top of the academic distribution.
+For the composition of faculty teaching university classes, \cite{bettinger2010does,figlio2015tenure} find US students have better enrolment and learning outcomes from courses taught by contingent or adjunct (i.e.
 not tenured or on tenure-track) professors (relative to tenured or tenure-track).
-While \cite{ehrenberg2005tenured} observe a negative association between utilisation of non-tenured instructors and student graduation rates.
-My approach focuses on count of professors per student at universities, so does not directly observe individual class-sizes, yet presents this as a possible mechanism through which the negative effects on student outcomes that falls in PU finance \cite{NBERw23736,NBERw27885} document.
-
-This analysis finds that faculty composition is meaningfully affected by recent stagnation in state support for higher education.
-Analysis of outcomes for incumbant professors in Illinois implies that the explaining mechanism is changes to hiring by public universities in response to the increased financial constraints.
-Previous findings from the literature on the effectiveness of tenured and tenure-track faculty versus contingent faculty in the areas of research and teaching.
-Combined with this analysis, these findings provide a plausible mechanism for the deteriorating student ouctomes shown to be a result of stagnating finances at public universities \citep{NBERw23736,NBERw27885}.
+\cite{ehrenberg2005tenured} observe a negative association between utilisation of non-tenured instructors and student graduation rates.
+My approach focuses on count of professors per student at universities, so does not directly observe individual class-sizes, yet presents considers this a possible mechanism through which stagnating state support for higher education negatively affects student outcomes \citep{NBERw23736,NBERw27885}.
 
@@
 \subsection{Data Description}
-The data used in this analysis come from two primary sources: \citet[IPEDS]{ipeds} regarding institutional information on finances and enrollment, and \citet[IBHED]{ibhed} for a panel of individual-level of every Professor employed by the Illinois PU system.
+The data used in this analysis come from two primary sources: \citet[IPEDS]{ipeds} regarding institutional information on finances and enrolment, and \citet[IBHED]{ibhed} for data on every professor in the Illinois public university system.
 
 IPEDS is a survey of higher educational institutions in the US, and legally requires institutions to participate in order to receive Federal Title IV student aid.\footnote{
-    This statement means that IPEDS does not necessarily cover the universe of US higher education institutions, yet thin practice every PU and not for-profit four-year institution is represented.
+    This statement means that IPEDS does not necessarily cover the universe of US higher education institutions, yet in practice every public university and not for-profit four-year institution is represented.
 }
@@
 }
-and provide information on the total revenues a university received from every source (including state governments), enrollment number, plus faculty count and total expenditures on salaries.\footnote{
+and provide information on the total revenues a university received from every source (including state governments), enrolment number, plus faculty count and total expenditures on salaries.\footnote{
     I combine the Urban Institute's 2018 compilation of IPEDS data for the years 1990-2017, and manually combine with raw NCES data regarding the year 2018-2021 for all relevant variables.
-    Figures for enrollment, faculty counts and salaries come from the raw NCES version of IPEDS for all years to address inconsistencies in the Urban Institute's data formulation for these variables.
+    Figures for enrolment, faculty counts and salaries come from the raw NCES version of IPEDS for all years to address inconsistencies in the Urban Institute's data formulation for these variables.
 }
-I restrict analysis to four-year institutions, as these institutions adhere to a standardised concept of faculty profile -- that is, tenure and title of appointment (lecturer, assistant professor, etc.) are relatively standardised in four-year, degree-granting institutions.
-Additionally, for-profit institutions employ and enrol a negligible share of professors and students respectively, while students at two-year institutions by majority intend to eventually enroll at a four-year institution \citep{mountjoy2022}.\footnote{
+I restrict analysis to public, four-year, degree-granting institutions, as these institutions adhere to a standardised concept of faculty profile, where tenure and title of appointment (lecturer, assistant professor, etc.) are relatively standardised.
+For-profit institutions employ and enrol a negligible share of professors and students respectively, while students at two-year institutions by majority intend to eventually enrol at a four-year institution \citep{mountjoy2022},\footnote{
     Furthermore, \cite{mountjoy2022} documents the effects of expanding two-year higher education access in the US separately for students who would have and those who would not have otherwise have attended college.
-    The high rate of eventual enrollment at four-year institutions among two-year students explains the negative impact of increased access to two-year institutions on students who are diverted away from four-year institutions.
+    The high rate of eventual enrolment at four-year institutions among two-year students explains the negative impact of increased access to two-year institutions on students who are diverted away from four-year institutions.
 }
+so that these institutions are not considered.
 
-Importantly, IPEDS reports the count of professors employed by category, as well as total salary expenditures by category.
+IPEDS reports the count of professors employed by position, as well as total salary expenditures by position.
 This gives a resulting panel data-set, where each row represents a university-year, and includes columns for university finances, plus total count and average salary\footnote{
     Real salary is computed by scaling nominal salary to 2021 dollars by the CPI-U.
-} for professors by category (lecturer, assistant, tenured, total).
-\autoref{tab:ipeds-summary} presents summary statistics for relevant variables in the panel of PU-years.
+} for professors by position (lecturer, assistant, tenured, total).
+\autoref{tab:ipeds-summary} presents summary statistics for relevant variables in these data.
 
@@
     IPEDS provides total paid to salary and total faculty employed per university-year, so that I can investigate average professor salary per university-year with IPEDS data.
-    Unfortunately, it seems that this measure of professors' salaries is particularly crude in measuring on professors' salaries and other individual-outcomes; summary statistics on IPEDS data do not agree with trends in average professor salary over the sample time period \citep{aau2021survey}.
+    Unfortunately, this measure of professors' salaries is particularly crude in measuring on professors' salaries and other individual-outcomes; summary statistics on IPEDS data do not agree with trends in average professor salary over the sample time period \citep{aau2021survey}.
+    Similarly, IPEDS reports of the average salary paid to professors at private institutions also disagree with other sources, so are not considered in this analysis.
 }
-To investigate the distribution, I integregate individual-level data for every PU professor in the state of Illinois between the years 2010-2021.
-IBHED host the information;
+To investigate the distribution, I integrate individual-level data for every public university professor in the state of Illinois between the years 2010-2021.
+IBHED hosts the information;
 Public Act 96-0266 (effective 1 January 2010) requires that each university report base salary and benefits all administrators, faculty members, and instructors employed by the college or university.\footnote{
-    The universities included are all campuses of the eight Illinois public universities: Chicago State University, Eastern Illinois University, Governors State University, Illinois State University, Northeastern Illinois University, Northern Illinois University, Southern Illinois University  (all five campuses), University of Illinois (all four campuses), Western Illinois University.
+    The universities included are all campuses of the nine Illinois public universities: Chicago State University, Eastern Illinois University, Governors State University, Illinois State University, Northeastern Illinois University, Northern Illinois University, Southern Illinois University  (all five campuses), University of Illinois (all four campuses), Western Illinois University.
 }
-This publicly available data provides the basis to build a panel of Illinois PU professors 2010-2021; I define a professor as an individual observation by their first plus last name and university pairing, and link this database to IPEDS information on finances for their employing institution.
+This publicly available data provides the basis to build a panel of Illinois public university professors 2010-2021; I define a professor as an individual by their first plus last name and university pairing, and link this database to IPEDS regarding finances for their employing institution.
 
@@
 Additionally, further analysis using the first year of a professor's employment focuses on the subset of professors with observed year of hiring between the years 2011 and 2021, representing 1,778 professors in 2011, and 9,099 in 2021.\footnote{
+    Summary statistics for this group, which over-represents lecturers compared to assistant and full professors, can be seen in \autoref{tab:illinois-summary-rolling}.
     The reasons for restricting to the sample with identified year of hire in the range 2011-2021 is explained in \autoref{sec:iv-model-indiv}.
 }
-Summary statistics for this group, which over-represents lecturers compared to assistant and full professors, can be seen in \autoref{tab:illinois-summary-rolling}.
 
@@
 
-\subsection{Trends in Finances, Enrollment, and Faculty Composition}
+\subsection{Trends in Finances, Enrolment, and Faculty Composition}
 \label{sec:trends}
-A state government in the US plans an annual budget a couple of years ahead of the fiscal year, with the legislature approving a budget request put forth by the governor's office.\footnote{
+A state government plans an annual budget a couple of years ahead of the fiscal year, and the legislature chooses to approve or reject a budget request put forth by the governor's office.\footnote{
     \cite{NBERw23736} present a full discussion of the state appropriation decision-making process, drawing on administrative records originally analysed by \cite{parmley2009state}.}
-This process leads to yearly variation in state appropriations not seen in other revenues sources (such as federal government appropriations), since US states differ in their rate of (financial) support for higher education, and are subject to fiscal constraints.
-US states governments, by majority, are legally obligated to run a balanced-budget, so that yearly variation in tax revenues (caused by changing economic conditions or otherwise) necessarily affect state expenditures.
+US state governments, by majority, are legally obligated to run a balanced-budget, so that yearly variation in tax revenues (caused by changing economic conditions or otherwise) necessarily affect state expenditures.
+This process leads to yearly variation in state appropriations not seen in other revenues sources (such as federal government appropriations), since US states differ in their support for higher education, and are subject to fiscal constraints.
 State appropriations to higher education are a particularly attractive area of state spending to absorb such shocks to state finances \citep{delaney2011state}.\footnote{
@@
 }
-Additionally, there is yearly state variation in the number of higher education students \citep{turner2014impact}, so that per-student state appropriations vary on multiple dimensions.
+Additionally, there is yearly variation in the number of higher education students in each state \citep{turner2014impact}, so that per-student state appropriations vary on multiple dimensions.
 
-\autoref{fig:funding} exhibits the trends for the mean of US public universities between the years 1990-2021.
+\autoref{fig:funding} exhibits the trends for the mean public university for the years 1990-2021.
 We see a rise in total revenues received by public universities (from all sources), and a notable increase in mean tuition revenues from \$48 million per year-university to \$150 million per year-university.
-Non-institutional revenues refer to revenues coming from external sources, not raised by internal university activities;\footnote{
-    US universities raise revenues from many internal activities, including sales of nondegree related educational courses to publishing deals.
-    The non-institutional revenue measure is constructed to consider only revenues resulting from external decisions, including state, local, and federal appropriations and tuition revenues.
-    Rates of tuition are considered external since most state governments play a large role in deciding tuition rices for public universities \citep{NBERw23736}.
-}
-non-institutional revenues rose over this time period almost exclusively thanks to the rise in tuition revenues.
-At the same time, total state appropriations stagnated at around \$100 million per year-university for 1990-2008, fell slightly around 2008 and have not recovered.
-While public universities experienced a financial stagnation, private universities were not exposed to the same constraints, receiving \$37,000 per student in 1990 and \$49,000 in 2021 with no corresponding decline in any specific component.
+At the same time, total state appropriations stagnated at around \$100 million per year-university for 1990-2008, falling around 2008 and have not recovered ever since.
+While public universities experienced a stagnation in state support, private universities were not exposed to the same constraints, receiving \$37,000 per student in 1990 and \$49,000 in 2021, experiencing no corresponding decline in any specific component.
 
@@
     \singlespacing
-    \caption{Total Student Enrollment, by University Sector, and Year.}
+    \caption{Total Student enrolment, by University Sector, and Year.}
     \begin{subfigure}[b]{0.495\textwidth}
@@
     \end{subfigure}
-    \label{fig:enrollment}
+    \label{fig:enrolment}
 \end{figure}
-At the same time, student enrollment at public universities rose precipitously.
+
+At the same time, student enrolment at public universities rose precipitously.
 6.2 million students were enrolled in public universities in 1990, and this number rose by 47\% to 9.1 million, with most of the increase occurring over the years 2000-2021.
-\autoref{fig:enrollment} shows that total enrollment at public universities has also risen over the same time period, but not as drastic in either relative or absolute terms; the mean public university also grew over this time period, from 9,800 students in 1990 to 11,800 in 2021.
-This means that per student mean revenues per student (seen in \autoref{fig:mean-funding-fte}) have stagnated across all measures, and particularly fallen from \$11,000 per student in 1990 to less than \$8,000 per student in 2021.
-\begin{figure}[!h]
+\autoref{fig:enrolment} shows that total enrolment at private universities has also risen over the same time period, but not as drastic in either relative or absolute terms; the mean private university grew from 9,800 students in 1990 to 11,800 in 2021.
+This means that revenues per student have stagnated across all measures (seen in \autoref{fig:mean-funding-fte}), and particularly fallen from a mean of \$11,000 per student in 1990 to less than \$8,000 per student in 2021.
+
+\begin{figure}[h!]
     \centering
@@
 \autoref{fig:fte-perprof} documents the divergence in faculty composition (per student) between the average private and public university 1990-2021.
-Public universities start with  a higher baseline of around 4.5 professors per 100 students, and exhibits yearly variation of less than 0.5 professor per 100 students over the thirty years.
-Public universities start with 3.9 professors per student, and this number falls to 3.5 primarily in the 2008-2011 time period.
+Private universities start with  a higher baseline of around 4.5 professors per 100 students, and exhibited yearly variation of less than 0.5 professor per hundred students over the thirty years.
+Public universities start with 3.9 professors per hundred students, and this number falls to around 3.5 primarily in the 2008-2011 time period.
 We see a similar difference in baseline, and fall for the years 2008-2011 for assistant professors.
-Private and public universities have similar numbers of associated and full professors before the year 2000, yet this number has fallen by over 20\% in the next 20 years only for public universities: in 2021 the mean PU has 6 fewer full professors per student than the mean private university.
-At the same time period, we see the rise in use of non-tenure track instructor positions (referred to as lecturers from here on), who were employed at similar rates in both sectors in 1990 yet have been utilised by public universities at a higher rate since.
+Private and public universities have similar numbers of associated and full professors before the year 2000, yet this number has fallen by over 20\% in the next 20 years only for public universities: in 2021 the mean public university has 6 fewer full professors per hundred students than the mean private university.
+Over the same time period, we see the rise in use of non-tenure track instructor positions (referred to as lecturers from here on), who were employed at similar rates in both sectors in 1990 yet have been utilised by public universities at a higher rate since.
 
@@
 
-The state of Illinois particularly exemplifies the stagnation in state support for higher education that the rest of the nation has experienced 1990-2021.
+The state of Illinois experienced the same stagnation in state support for higher education that the rest of the nation has experienced 1990-2021.
 \autoref{fig:illinois-funding} shows the trends for revenues among the seven Illinois public university campuses, including the falling share of state appropriations and substitution towards tuition revenue.
@@
 In the calendar year 2015 partisan disagreements between the democratic legislature and republican governor led to the 2016 fiscal year starting with no state budget.
-State agencies, and higher education institutions, employed accounting tehniques to continue operation.\footnote{
+State agencies, and higher education institutions, employed accounting techniques to continue operating without any resources provided by the state government.\footnote{
     Fiscal year 2016 refers to June 205 to June 2016, as is the same for the academic year definition in this analysis. 
@@
 }
-While state institutions continued to operate, there were drastic revenue and spending cuts in response to the budget impasse, as it continued through fiscal year 2017, and ended with a new budget restoring revenues to state agencies and universities for the 2018 fiscal year.
-This variation in public university revenues, stemming entirely from political disagreements --- not from state decisions regarding higher education and its finances \citep{young2020squandered} --- via the state appropriations channel, mean that Illinois public universities exhibit sizable changes in their state appropriations over 2011-2021, of similar order to those for the rest of the country over 1990-2021.
+While state institutions were able to stay open, there were drastic revenue and spending cuts in response to the budget impasse, as it continued through fiscal year 2017, and ended with a new budget restoring revenues to state agencies and universities for the 2018 fiscal year.
+This variation in public university revenues via the state appropriations channel, stemming entirely from political disagreements --- not from state decisions regarding higher education and its finances \citep{young2020squandered} ---  mean that Illinois public universities exhibit sizable changes in their state appropriations over 2010-2021, of similar order to those for the rest of the country over 1990-2021.
 
@@
 \label{sec:responses}
-Subsection describing how a university may respond to shocks on its sources of revenues.
-Cite other papers on how this went.
 
-Write about shocks in state appropriations primarily here.
+While state support for public higher education has stagnated at the same time as education costs rose, there are multiple possible ways that a university can respond.
+\cite{NBERw23736} established that corresponding rises in tuition did not offset the falling state support, so that university spending (per student) fell in response to these persistent, negative state funding shocks.
+There are multiple ways that these changes in finances may affect the faculty composition at public universities.
 
-Additional paragraph on how the incumbent professors may be affected.
+Universities every year hire a number of new professors, either to expand their departments or to replace leaving professors.
+The number of professors hired is usually highest among non-tenured adjunct faculty, as these instructors are by majority hired on short-term contingent contracts, so that they are less costly to hire (or to not reinstate) in response to yearly changes in teaching needs.
+Secondly, tenure-track assistant professors are hired in most years at public universities to replace leaving assistant professors, and are most often hired with a four to six year contract and agreement to be considered for tenure at the end of this term.
+Lastly, tenured professors are those that have undergone review and have successfully secured a full-time appointment at their university with no expiration date; tenured professors are sometimes hired from outside a university, though this position has the lowest hiring rate among most universities.
+It is common for universities to restrict their faculty hiring in the face of financial shocks.
+Since the yearly hiring rate is highest among lecturers and assistant professors, financial shocks may lower the number of lecturer and assistant professors at public universities the most, if faculty hiring is affected by falls in state funding.
+
+Additionally, professors' salaries may be disrupted by the stagnation in public university finances.
+When the university has a lower budget for its yearly hiring, it may respond by lowering the salary they offer to new hires.
+This possible effect may not be the same across each position of professor; tenured faculty are often hired away from another university, so that new tenured professor hires could be less likely to accept a lower offer from a public university, and we would see that new tenured hires are not affected salary-wise.\footnote{
+    \cite{blackaby2005} describes the academic outside options among economist faculty, documenting how they differ for academic position and gender.
+}
+Yet salary for all the professors, not just new hires, may also be affected: multiple universities passed a university-wide pay-cut for their faculty in response to state budget cuts around the 2008 recession, for example.\footnote{
+    Indeed, Cornell University implemented both hiring freezes and even a nominal salary reduction for professors, in anticipation of a financial shock in early 2021.
+    The salary cut was not permanent, as the oncoming financial shock turned out to not be as serious as projected, so that the salary cut was returned to professors in the year 2021.
+}
+Faculty are not paid the same by position, so that their are financial consequences for faculty being promoted between positions within a university.
+In response to tightening fiscal restraints, universities may be less likely to grant tenure to assistant professors, or to promote tenured professors from associate to full professor either.
+
+It is not immediately clear which effect will dominate, and which position of professor will be effected the most by changes in their university's finances.
+Yet, there is one empirical fact worth noting: a mean assistant or tenured professor earns more than double that of an adjunct lecturer (\autoref{tab:illinois-summary}), and has more long term salary obligations than a lecturer too.
+If a public university's primary obligation is to teach, and they must fulfil this objective with fewer and fewer resources, then they may substitute away from tenure-track and tenured professor towards contingent lecturers to achieve this obligation.\footnote{
+    That is substituted away from tenure-track and tenured professors according to the salary ratio, and the relative productivity between these two positions.
+    It is not possible to estimate the relative productivity in this paper, so I leave the rate of substitution between the two positions to further research.
+}
 
@@
 
-A professor is employed by a university, so it stands to reason that when public universities experience a secular fall in revenue --- analogous to a demand shock in the private sector --- that their human resources decisions will be affected.
-Universities have been shown to freeze faculty hiring and general faculty salaries in response to financial shocks, such as the Great Recession \citep{turner2014impact}.
+A professor is employed by a university, so it stands to reason that when public universities experience a secular fall in one of their revenue soruces revenue --- analogous to a demand shock in the private sector --- that the composition of their employees will be affected.
 
-\textbf{Shock to state appropriations only, as shocks to total revenues/spending requires identification for multiple sources.
-Tuition is an outcome of changes in the state appropriations, so discuss whether it is an outcome or a control.
-Leave footnote that results are relatively similar, so leave out this control.}
-
-Na\"ively we can analyse the relationship between professor-outcomes $Y_{i,t}$ at university $i$ in year $t$ as a result of university revenues per student $X_{i,t}$.
-The exact measure of university revenues, $X_{it}$ is non-institutional revenues, the sum of university funding from outside sources -- local, federal, and state appropriations plus tuition revenue.\footnote{
-    This measure is in line with the \cite{NBERw27885} approach of considering revenues from the state, but does not ignore yearly changes in revenues arising from tuition revenues -- which are shown to be heavily related.
-}
+Na\"ively we can analyse the relationship between professor-outcomes $Y_{i,t}$ at university $i$ in year $t$ as a result of state funding per student $X_{i,t}$.
 Outcomes include count of professors per student, and average salary, within each university-year.
-This model gives $\beta$ the association with these outcomes and university revenues per student,\footnote{
+This model gives $\beta$ the association with these outcomes and state funding per student,\footnote{
     Note that dividing by student count also implicitly controls for the size of the university, so that this model accounts for yearly variation in professor count and university revenues arising from growth in a university.
@@
 including fixed effects to control for effects specific to the university and year.
-Log$(.)$ transforming the variables improves the interpretability of $\beta$ as an elasticity for professor count per student with respect to university revenues.
+Log$(.)$ transforming the variables improves the interpretability of $\beta$ as an elasticity for professor count per student with respect to state funding per student.
 \begin{equation}
@@
 
-And yet a university's finances are not exogenous to decisions made by the university administrator (such as hiring), state outcomes or enrollment.
-Instead, the state government and university administration undertake a complex process of alloting resources across multiple different priorities, including instruction, research, or between departments.
-Importantly for this analysis, revenues received from a  PU's state government provide opportunity to address this endogeneity.
+Yet a university's finances are not exogenous to state decisions on the yearly level for support of higher education, or exogenous to internal institutional decisions.
+Instead, the state government and university administration undertake a complex process of allotting resources across multiple different priorities, including instruction, research, or between departments.
+Importantly for this analysis, revenues received from an institution's state government provide opportunity to address this endogeneity.
 
@@
 
-\cite{NBERw23736,chakrabarti2018effect,NBERw27885} address endogeneity in PU finances by exploiting a shift-share instrument for changes in state-level funding interacted with university reliance on state funding in a base period.
+\cite{NBERw23736,chakrabarti2018effect,NBERw27885} address endogeneity in public university finances by exploiting a shift-share instrument for changes in state-level funding interacted with university reliance on state funding in a base period.
 The system exploits the fact that institutions who rely on state appropriations more will be affected by state appropriation shocks.
-We start with the shift-share instrument.
 \begin{align}
@@
 $Z_{i,t}$ is the instrument for (log) state appropriations for institution $i$ in year $t$, interacting the average funding for university $i$ in state $s(i)$ with reliance on state funding relative to total revenues, averaged across the base years 1990--1993.\footnote{
-    1990--1993 are defined as data for PU finance data are most comparable (i.e. without many missing values) beginning in 1990.
+    1990--1993 are defined as data for public university finance data are most comparable (i.e. without many missing values) beginning in 1990.
     \cite{NBERw23736} use the single year 1990 as the base year, though I use the four years to ameliorate missing values in the single year of 1990.
@@
 }
-\cite{NBERw27885} notes the tendency for public universities to respond to state funding cuts by increasing reliance on tuition, where \cite{NBERw23736} specifically instruments for tuition revenues with collected information on legislative tuition price controls.
-As such, it is important in my set-up to explicitly control for tuition revenues in the first-stage.
+\cite{NBERw27885} notes the tendency for public universities to respond to state funding cuts by increasing reliance on tuition, where \cite{NBERw23736} specifically instruments for tuition revenues with collected information on legislative tuition price controls.\footnote{
+    \label{foot:control}
+    It may be argued that tuition revenues are confounder between the causal effect of changes in state funding on a university's total revenues, so that this analysis focuses on state support for higher education (no total revenue) as a result.
+    This follows the \cite{NBERw27885} formulation.
+    On the other hand, rises in tuition revenues (per student) may arise as result of tuition hikes thanks declining state support, which would mean controlling for tuition would constitute a bad control.
+    Estimates including tuition revenues (per student) as a control in the second stage of the IV estimates produces results of very similar magnitude and direction, and so are omitted.
+}
+This analysis focuses on one important source of revenues for public universities: state appropriations.
 
@@
 We note the conditions for exogeneity in the instrument (following the discussion presented by \citealt{NBERw27885}).
-The instrument is exogenous if state policy decisions for funding of, or tuition charges by public universities, are uncorrelated with unobserved changes in the expenditures of any specific college or university in the state.
+The instrument is exogenous if state policy decisions for funding of public universities are uncorrelated with unobserved institutional changes of any specific college or university in the state \citep{borusyak2022quasi}.
 This assumption is plausible given that the majority of states have multiple (i.e.
@@
 
-\begin{table}[!h]
+\begin{table}[h!]
     \singlespacing
@@
 
-\autoref{tab:firststage-reg} presents results of the first-stage regression, separately with and without the tuition control and fixed effects, at the institution level (i.e.
-using IPEDS data).\footnote{
+\autoref{tab:firststage-reg} presents results of the first-stage regression, separately with and without a control for per student tuition revenue and year plus institution fixed effects.\footnote{
     Representations for frequentist significance levels (i.e. 10, 5, 1\% etc.) are omitted here, and in all following tables.
 }
-We note columns (1) and (2) estimate that a shock to state appropriations per student in the state of 10\% is associated with around 2.5\% change in non-institutional revenues per student at the university; the instrument is strong, and we note similarity in estimates with and without inclusion of fixed effects.
-Columns (3) and (4) exclude the tuition revenue control (explained above) to exhibit bias in the first-stage if it is not controlled for.
+We note columns (1) and (2) estimate that a shock to state appropriations per student in the state of 10\% is associated with around 9.8\% change in state funding per student at the university; the instrument is strong, and we note similarity in estimates with and without inclusion of fixed effects.
+Columns (3) and (4) include the tuition revenue control (explained in \autoref{foot:control}) to exhibit estimates with the inclusion of this possible collider or bad control.
 Column (3) shows similar estimates to columns (1), (2) thanks to inclusion of fixed effects, so that the fixed effects was effective in soaking variation in per-student tuition revenues at the institution-year level.
-Column (4) exhibits the worst case, and extreme bias in the appropriation shock first-stage without the tuition control.
-Column (3) represents the estimates for \autoref{eqn:firststage} with fixed effects, omitting the tuition revenue control, and is the preferred form that I proceed with in the following.
+Column (1) represents the estimates for \autoref{eqn:firststage} with fixed effects, omitting the tuition revenue control, and is the preferred form that I proceed with in the following.
 
@@
 We see the time 0 estimate coincides with that of Column (1) in \autoref{tab:firststage-reg}, the contemporaneous effect of the shock revenues.
-We see a decaying, yet real and positive, effect of the shock on revenues for the next 5 years, illustrating the staying power of the effect.
+We see a decaying, yet real and positive, effect of the shock on revenues for the next 10 years, illustrating that the state appropriation shock is a strong, if fading, instrument for state funding per student and justifying its use in local projection IV estimation.
 
@@
 
-The primary empirical model combines the instrumental variable for state appropriation shocks with the empirical model for the effects of university revenues at the university-level --- i.e.
+The primary empirical model combines the instrumental variable for state appropriation shocks with the empirical model for the effects of state funding at the university-level --- i.e.
 parameter $\beta$ in the following.\footnote{
     It is important to note the treatment effect isolated here; the instrumental variables approach identifies the local average treatment effects (LATE) specific to the instrument.
-    So we interpret this treatment effect as a university's response in employment count and average salaries to revenue changes, changes specific to state appropriation shocks, among the complier group --- i.e.
+    So we interpret this treatment effect as a university's response in employment count and average salaries to state funding changes, changes specific to state appropriation shocks, among the complier group --- i.e.
 universities with any exposure to shocks, assuming no universities increase employment or salaries in response to negative revenue shocks.
@@
 
-\autoref{fig:SCM-ivmodel} presents the system in graphical form, for the causal effect of state appropriation shocks on university revenues and thus faculty composition.
+\autoref{fig:SCM-ivmodel} presents the system in graphical form, for the causal effect of state appropriation shocks on university revenues and thus faculty composition, making clear the assumption that appropriation shocks affect faculty composition only via affecting university finances.
 
@@
 This analysis additionally uses data regarding individual professors in the Illinois university system, to investigate the effects of changes in university revenues on the individual professors at the universities.
-Redefining the level of outcome requires adjustment to the empirical approach to leverage variation in university finances for the years following when a professor joins the university.\footnote{
+Redefining the level of outcome requires adjustment to the empirical approach, leveraging variation in university finances for the years after a professor joins the university.\footnote{
     This formulation follows that presented by \cite{NBERw27885}, where individual student outcomes are analysed via variation in state appropriations after their freshman-year.
@@
 \autoref{eqn:rolling-instrument} defines a rolling-share variant of the instrument, $\tilde Z_{j,t}$, where the university's state appropriations share exposure is based in the year a professor joins the university --- and not the base period 1990--1993.
-$j$ indexes each professor in year $t$, $\tau(j)$ for the year the professor first joins their Illinois PU.\footnote{
-    Identifying $\tau(j)$ is possible for $j$ by restricting to all professors hired 2011-2021 --- i.e. in the years after the start of the panel.
-    It is not possible to discern the hiring year for professors who  were hired in the years preceding 2011, and so the entire sample is only possible to analyse using the base-share in years 1990-1993 formulation (\autoref{tab:facultysalaries-shock-illinois}, \ref{tab:promotion-shock-illinois}).
+$j$ indexes each professor in year $t$, $\tau(j)$ for the year the professor first joins their institution.\footnote{
+    Identifying $\tau(j)$ is possible for $j$ by restricting to all professors hired 2011-2021 --- i.e. in the years after the start of the full panel.
+    It is not possible to discern the hiring year for professors who  were hired in the years preceding 2011, and so the entire sample is only possible to analyse using the base-share in years 1990-1993 formulation (e.g.
+\autoref{tab:facultysalaries-shock-illinois}, \ref{tab:promotion-shock-illinois}).
 }
@@
 
-This approach leverages an insight, made available by level of the data: that an individual professor is effected by changes in university revenues after they have joined the university.\footnote{
+This approach leverages an insight, made available by level of the data: that an individual professor is affected by changes in university revenues after they have joined the university.\footnote{
     Notice that \autoref{sec:iv-model-uni} considers the number of professors employed by the university; whether a professor becomes employed at the university is likely affected by that university's finances.
-    The formulation in this section does not consider whether the professor joins the university, instead taking as given that the professor is employed at the university, and then projecting the effect on the individual given they are employed here.
+    The formulation in this section does not consider whether the professor joins the university, instead taking as given that the professor is employed at the university, and then projects the effect on the individual.
 }
 Exogeneity and relevance of the rolling-share instrument, $\tilde Z_{j,t}$, follows the same reasoning as that for the base-share instrument, $Z_{i,t}$, discussed in \autoref{sec:approp-shocks}.\footnote{
-    The base-share instrument is still used for some outcomes with the individual Illinois professors, where appropriate (\autoref{tab:facultysalaries-shock-illinois}, \autoref{tab:promotion-shock-illinois}).
+    The base-share instrument is still used for some outcomes with the individual Illinois professors, where appropriate (\autoref{tab:facultysalaries-shock-illinois}, \ref{tab:promotion-shock-illinois}).
 }
@@
     Additionally, within-institution changes resulting from share reliance on state funding may be correlated with unobserved changes in the outcomes, so that \cite{NBERw27885} note the importance of controlling for the base share and state student population.
-    The formulation here implicitly controls for these factors via the fixed effects; results are relatively similar while including these controls without including fixed effects.
+    The formulation here implicitly controls for these factors via the fixed effects; results are relatively similar while including these controls without including fixed effects, and so are omitted.
 }
-\autoref{tab:firststage-illinois} presents results of the first stage estimation, showing that the instrument is strong (after conditioning on tuition revenues) in the same way as that for the university-level outcomes (\autoref{tab:firststage-reg}), with very similar estimates for the association between appropriation shocks and non-institutional revenues.
+\autoref{tab:firststage-illinois} presents results of the first stage estimation, showing that the instrument is strong in the same way as that for the university-level outcomes (\autoref{tab:firststage-reg}), with very similar estimates for the association between appropriation shocks and non-institutional revenues.
 
-The full model is defined as follows where $i(j)$ refers to the institution that professor $j$ is employed at, and $Y_{j,t}$ refers to individual-level outcomes total salary, rate of promotion, and propensity to leave the Illinois PU system.
+The instrumental variables model is then defined as follows where $i(j)$ refers to the institution that professor $j$ is employed at, and $Y_{j,t}$ refers to individual-level outcomes total salary, rate of promotion, and propensity to leave the Illinois public university system.
 The system includes a fixed effect for the institution and first year of employment.\footnote{
-    The instrument varies by institution, based in the year of first employment so that these are the corresponding fixed effects and levels of clustered standard errors.
+    The instrument varies by institution, based in the year of first employment, so that these are the corresponding fixed effects and levels of clustered standard errors.
 }
@@
     \label{eqn:secondstage1_indiv}
-    X_{i(j),t} &=& \theta_{i(j)} + \phi_{\tau(j)} + \delta \tilde Z_{i(j),t} + \kappa W_{i(j),t} + \epsilon_{i(j),t} \\
+    X_{i(j),t} &=& \theta_{i(j)} + \phi_{\tau(j)} + \delta \tilde Z_{i(j),t} + \epsilon_{i(j),t} \\
     \label{eqn:secondstage2_indiv}
-    Y_{j,t} &=& \mu_{i(j)} + \nu_{\tau(j)} + \beta \widehat X_{i(j),t} + \lambda W_{i(j),t} + \varepsilon_{i(j),t}
+    Y_{j,t} &=& \mu_{i(j)} + \nu_{\tau(j)} + \beta \widehat X_{i(j),t} + \varepsilon_{i(j),t}
 \end{eqnarray}
-We then interpret parameter $\beta$ in this formulation as the effect of changes in university revenues, via state appropriation shocks, on an individual professor.
+We then interpret parameter $\beta$ in this formulation as the effect of changes in state funding at an Illinois public university, via state appropriation shocks, on an individual professor.
 
@@
 
-\autoref{tab:facultycount-shock-reg} presents OLS and IV estimates, where the outcome is (log) count of professors employed at the university, separated by group and combined (columns 7, 8).
-Within each group we see a positive correlation between university revenues and professor count per student (the OLS columns), where an increase in university total revenues by 10\% is associated with an increase in professor employment count between 3.5-4.5\%.
-Lecturers per student is not shown to be correlated with total univesity revenues.
-The 2SLS columns show results after identifying university revenues with the appropriation shock.
-All professor count per student increases by 3\% in response to a 10\% rise in total revenues, mediated by revenue shocks.
+\autoref{tab:facultycount-shock-reg} presents OLS and IV estimates, where the outcome is (log) count of professors employed at the university, separated by position, and for all professors (Columns 7, 8).
+For all professors we see a positive correlation between university revenues and professor count per student (the OLS columns), where an increase in university total revenues by 10\% is associated with an increase in professor count per student of 0.83\%, yet negatively correlated with lecturer count of -1.3\%.
+Lecturers per student is not shown to be correlated with total university revenues.
+The 2SLS columns show results after identifying state funding with the appropriation shock.
+All professor count per student increases by 0.53\% in response to a 10\% rise in total revenues.
 The effect is driven by increases in the count of assistant and full professors (i.e.
-tenure-track and tenured) per student, where assistant professor count per student increase by 6.7\%, and full professor 5\%.
-The count of lecturers per student decreases by 12\% in response to a (positive) 10\% shock in university revenues.
+tenure-track and tenured) per student, where assistant professor count per student increase by 1.4\%, and full professor 1.2\%.
+The count of lecturers per student decreases by 4.4\% in response to a (positive) 10\% shock in university revenues.
 
-\begin{table}[!h]
+\begin{table}[h!]
     \singlespacing
@@
 \end{figure}
-\autoref{fig:all-count-lp} shows local projection estimates, where the effect on total professor count per student persists at the same magnitude of 0.3 four years after the identified shock.
+\autoref{fig:all-count-lp} shows local projection estimates, where the effect on total professor count per student persists at the same magnitude of around 0.1 four years after the identified shock.
 These findings are in line with \cite{turner2014impact} observing that universities froze hiring particularly for tenure-track positions in response to negative budget shocks in the late 2000s, and \cite{brown2014endowment} from shocks to university endowments.
@@
 This effect lines up with two trends we see in \autoref{sec:trends}: that public universities revenues (per student) are decreasing while utilisation of non-tenure track lecturers increased over the same time period.
-And yet, identifying the university revenues by the state appropriations shock increases the magnitude of the effect (Column 2) with respect to the association (Column 1), increasing evidence that the two trends are causally linked.
+And yet, identifying the university revenues by the state appropriations shock increases the magnitude of the effect (Column 2) with respect to the association (Column 1), strengthening the case that the two trends are causally linked.
 Between years 1990-1999 IPEDS provided count of professors by explicit tenure designation (opposed to named position as in \autoref{tab:facultycount-shock-reg}), and \autoref{tab:tenurecount-shock-reg-fte} presents results of the respective regressions by tenure status.
@@
 
-Together these findings imply that public universities increase their count of tenure-track and tenured professors per student in years where revenues are more plentiful, by increasing hiring intensity and efforts to maintain incumbent professors.
+Together these findings show that public universities increase their count of tenure-track and tenured professors per student in years where revenues are more plentiful, possibly by increasing hiring intensity or efforts to maintain incumbent professors.
 Yet, when revenues are not so bountiful, count of tenure-track and tenured professors per student decreases.
-At the same time, public university enrollment secularly increased over this time period, so that the universities satisfied their teaching obligations with non-tenure track lecturers.
-The mechanism that explains the changes in faculty composition are, however, not clear at this unit of analysis; individual data are needed to delve deeper.
+The mechanism that explains these changes in faculty composition are, however, not clear at this unit of analysis; individual data are needed to delve deeper.
 
@@
 New hires of assistant, full, or administrative position professors are not realted to revenues, via a shock to appropriations in their first-year of employment.
-Yet, we see the starting salary of lecturers is 3.8\% higher given an increase in their hiring university's state funding.
+Yet, the starting salary of lecturers is 3.8\% higher given an increase in their hiring university's state funding of 10\%.
 
-\begin{table}[!h]
+\begin{table}[h!]
     \singlespacing
@@
 
-Regarding total salary for every year of employment, \autoref{tab:facultysalaries-shock-illinois-rolling} presents estimates for the elasticity of total salary with respect to university revenues, for the entire sample of professors hired after 2010.\footnote{Recall that this is a subsample of the entire data, that the rolling shift-share instrument is only defined for the sample with an observed hiring date in the years 2011-2021.}
-The elasticity is non-distinguishable from zero among any group, which implies that long-term salaries are not related to university revenues.
+Regarding total salary for every year of employment, \autoref{tab:facultysalaries-shock-illinois-rolling} presents estimates for the elasticity of total salary with respect to university revenues, for the entire sample of professors hired after 2010.\footnote{
+    Recall that this is a subsample of the entire data, that the rolling shift-share instrument is only defined for the sample with an observed hiring date in the years 2011-2021.
+}
+The elasticity is non-distinguishable from zero among any position, implying that long-term salaries are not related to state funding.
 \autoref{tab:facultysalaries-shock-illinois} presents the same model, for the sample of all professors and using the instrument based in years 1990-1993, similarly finding no relationship.
 
-\begin{table}[!h]
+\begin{table}[h!]
     \singlespacing
@@
 Note that professor position (assistant, associate, full) is also an outcome, thanks to the promotion channel: a university may be less able to promote their professors to higher paying positions when their revenues are constrained.
-\autoref{tab:promotion-shock-illinois-rolling} shows estimates where the outcome is rate of promotion within each professor group.
-The lecturer group describes the sample of professors who were ever listed as a non-tenure track lecturer and the binary for promotion equals one in a year they achieved an assistant professor position; assistant the same for the assistant to associate professor promotion; associate same for associate to full promotion.
-There is no discernible relationship between revenues and promotion among any position group.\footnote{
+\autoref{tab:promotion-shock-illinois-rolling} shows estimates where the outcome is rate of promotion within each professor position.
+The lecturer position describes the sample of professors who were ever listed as a non-tenure track lecturer and the binary for promotion equals one in a year they achieved an assistant professor position; assistant the same for the assistant to associate professor promotion; associate same for associate to full promotion.
+There is no discernible relationship between revenues and promotion among any position.\footnote{
     \autoref{tab:promotion-shock-illinois} shows estimates using the base-year instrument among the entire sample, finding the same results.
 }
-Furthermore, local projection estimates (\autoref{fig:promoted-illinois-lp}) show that the there is no promotion effect, among all professors, for any of the following years.
+Furthermore, local projection estimates (\autoref{fig:promoted-illinois-lp}) show that the there is no promotion effect, among all professors, for any of the following years either.
 Lastly, \autoref{tab:facultyleaving-shock-illinois-rolling}, \ref{tab:facultyleaving-shock-illinois} describes estimates for rate of exit from the employing university among Illinois professors, and finds no relationship with university revenues.
@@
 
-The results suggest that the recent stagnation in public university revenues per student has affected the composition of faculty within a university, away from professors tenured and on the tenure-track and towards non-tenured lecturers.
+The results show that the recent stagnation in state funding for higher education has affected the composition of faculty within public universities, away from tenure-track and tenured professors towards non-tenured lecturers.
 At the same time, there are little discernible effects on individual professors hired in the years 2011-2021 at Illinois public universities, except for the salaries of first year lecturers.
@@
     The University of Illinois includes three public university campuses: Urbana-Champaign, Chicago, Springfield.
-} did not receive the allocated state appropriations from the state of Illinois on time, so enacted cost-cutting measures to stay fiscally solvent.
+}
+did not receive the allocated state appropriations from the state of Illinois on time, so enacted cost-cutting measures to stay fiscally solvent.
 The university system placed a hold on all hiring for filling state-funded positions and promotions.
@@
 The University of Western Illinois adopted very similar measures in response to the Illinois budget crisis in 2016-2017 \citep{wiu2016}.
-
-These examples show that the Illinois state universities attempted to affect each of the professor outcomes investigated in \autoref{sec:results}, yet the only measurable affect was on count of professor per student (implied via slower hiring), and first-year lecturer salaries.
-To investigate these trends for the entire state, \autoref{fig:newhire-illinois} shows the count of professors hired in the years 2011-2021, and their mean salary, by position.
-In particular, no position had sustained rises in hiring while enrollment within the university was rising (so that new hires per student necessarily fell), and hiring of new administrators jumped once the Illinois budget crisis ended in 2017.
-At the same time, professor new hires' starting salaries remained relatively constant for lecturer and assistant professors, yet rose and fell over the decade for full professors and administrators.
 
@@
 
-While costs education have been rising in the US, thanks to multiple overlapping trends \citep{ehrenberg2012}, public universities have also had to deal with declining state support.
-It is natural to expect that such headwinds will lead to systematic change at public universities, changes which are not neutral on the professors they employ, or to their goals of research and providing education.
-Furthermore, private universities in this country are not exposed to these same financial headwinds.
-If public universities' hiring of faculty is disrupted by changes in state appropriations, then we can expect that private universities will benefit (in a competitive sense) by being able to hire better professors from relatively resource constrained public universities.
-While public universities continue to educate the majority of higher education students in the US, and disproportionately students from low-income backgrounds and minorities, we should worry about the effects that restricting their financial support has on the composition of distribution of higher education and research in this country.
+These examples show that Illinois public universities were effected by the negative shock to state funding for higher education, and instituted policy changes aimed at affecting faculty outcomes investigated in \autoref{sec:results}.
+Yet the only measurable effect was on count of professor per student (implied via slower hiring), and first-year lecturer salaries.
+To investigate these trends for the entire state, \autoref{fig:newhire-illinois} shows the count of professors hired in the years 2011-2021, and their mean salary, by position.
+In particular, no position had sustained rises in hiring while enrolment within the university was rising (so that new hires per student necessarily fell), and hiring of new administrators jumped once the Illinois budget crisis ended in 2017.
+At the same time, professor new hires' starting salaries remained relatively constant for lecturer and assistant professors, yet rose and fell over the decade for full professors and administrators.
 
-\begin{figure}[h!]
-    \centering
-    \singlespacing
-    \caption{Faculty Profile at Illinois Public Universities 2011-2021.}
-    \begin{subfigure}[b]{0.495\textwidth}
-        \centering
-        \caption{Faculty Count per Enrolled Student, Total for the State.}
-        \includegraphics[width=\textwidth]{figures/faculty-count-illinois.png}
-        \label{fig:faculty-count-illinois}
-    \end{subfigure}
-    \begin{subfigure}[b]{0.495\textwidth}
-        \centering
-        \caption{Mean Salary for All Faculty, \$ 2021 CPI-U.}
-        \includegraphics[width=\textwidth]{figures/faculty-salary-illinois.png}
-        \label{fig:faculty-salary-illinois}
-    \end{subfigure}
-    \label{fig:faculty-illinois}
-\end{figure}
+%\begin{figure}[h!]
+%    \centering
+%    \singlespacing
+%    \caption{Faculty Profile at Illinois Public Universities 2011-2021.}
+%    \begin{subfigure}[b]{0.495\textwidth}
+%        \centering
+%        \caption{Faculty Count per Enrolled Student, Total for the State.}
+%        \includegraphics[width=\textwidth]{figures/faculty-count-illinois.png}
+%        \label{fig:faculty-count-illinois}
+%    \end{subfigure}
+%    \begin{subfigure}[b]{0.495\textwidth}
+%        \centering
+%        \caption{Mean Salary for All Faculty, \$ 2021 CPI-U.}
+%        \includegraphics[width=\textwidth]{figures/faculty-salary-illinois.png}
+%        \label{fig:faculty-salary-illinois}
+%    \end{subfigure}
+%    \label{fig:faculty-illinois}
+%\end{figure}
+%
 
@@
 
-\textbf{A few paragraphs conclusion go here.}
+This analysis investigates how the recent stagnation in state support for higher education has affected public universities, and their faculty composition.
+The work contributes to the literature along two primary dimensions.
+Firstly, by exploiting the same empirical approach as \cite{NBERw23736,NBERw27885} to isolate the causal effect of changes in state funding on public universities themselves, the analysis provides a plausible mechanism for the observed negative effects on students outcomes that this stagnation has brought.
+Secondly, this analysis considers individual professors as the unit of analysis using a dataset (IBHED) that has not been considered in the economic literature.
+These data allow for detailed analysis of thousand's of professors' salaries, and are a starting point for further analysis of the determinants of professors wages in the Illinois public university system.
+
+Findings show that public universities have substituted away from tenure-track and tenured professors, towards non-tenured lecturers, in the face of persistent declines in state funding for higher education.
+The effects on incumbent professors in the state of Illinois are non-distinguishable from zero, which implies that the effect is driven by changes in university hiring of new professors.
+Public universities are hiring thousands of non-tenured lecturers to satisfy teaching obligations, and fewer tenure-track and tenured professors that their private university counterparts.
+
+While costs education have been rising in the US, thanks to multiple overlapping trends \citep{ehrenberg2012}, public universities have also dealt with declining state support.
+It is natural to expect that such headwinds will lead to systematic change at public universities, changes which are not neutral to the composition of faculty, or to their goals of research and education.
+Furthermore, private universities were not exposed to financial headwinds of the same magnitude or persistence.
+If public universities' hiring of faculty is disrupted by a fall in state funding, then we can expect that private universities will benefit (in a competitive sense) by being able to hire better professors from relatively resource constrained public universities.
+While public universities continue to educate the majority of higher education students in the US, we should worry about the effects that restricting their financial support has on the composition of faculty, and  higher education itself, in this country.
 
@@
 
-\autoref{tab:facultysalaries-shock-reg} presents OLS and IV estimates where the outcome is (log) mean salary for professors employed at the university, separated by group.
+\autoref{tab:facultysalaries-shock-reg} presents OLS and IV estimates where the outcome is (log) mean salary for professors employed at the university, separated by position.
 Similarly, \autoref{fig:all-salaries-lp} shows the staying-power via projection methods.

### 2022-08-12 programs/analysis/ipeds-shock.R

@@ -369,8 +369,8 @@ firststage_approp_tuit_noFE.fstat <-
 
 # Collate the results to a LaTeX table
 stargazer(
-    firststage_approp_tuit.reg, firststage_approp_tuit_noFE.reg,
     firststage_approp.reg, firststage_approp_noFE.reg,
+    firststage_approp_tuit.reg, firststage_approp_tuit_noFE.reg,
     dep.var.caption = "Dependent Variable: State Funding",
     dep.var.labels.include = FALSE,
     digits = digits.no,

### 2022-08-12 programs/run.sh

@@ -24,6 +24,6 @@ R CMD BATCH --no-save individual-shock-rolling.R
 cd ..
 
 # Compile paper, with outputs of progams as inputs for TeX files
-cd text
+cd ../text
 latexmk -pdf paper.tex
 latexmk -c

### 2022-08-15 text/paper.tex
rename.
adjust script and README.

@@
     This project's Github repository, with materials and (publicly available) data for replication, is available at 
-    \url{https://github.com/shoganhennessy/finances-faculty-composition}.
+    \url{https://github.com/shoganhennessy/state-faculty-composition}.
     Any comments or suggestions may be sent to me at \href{mailto:seh325@cornell.edu}{\nolinkurl{seh325@cornell.edu}}, or raised as an issue on the Github project.}}
@@
 \cite{NBERw23736} first study the effect of (changes in) public universities' finances on student outcomes by isolating changes in public universities' expenditures on student instruction, addressing endogeneity of expenditures decisions by instrumenting expenditures with state appropriation shocks and limits on tuition price.
-\cite{NBERw23736} find that increases in public university spending (via appropriation shocks) increases enrolment and degree completion among students, but corresponding changes in attendance price (via limits on tuition price) do not have an effect.
+\cite{NBERw23736} find that increases in public university spending (via appropriation shocks) increases enrolment and degree completion among students, but corresponding changes in attendance price (via limits on tuition price) do not have an effect.\footnote{
+    \cite{miller2022making} further analyse the effects of falls in university revenues, by finding that reductions in public universities tuition prices (suggestively motivated by falling state funding) leads to reductions in provided financial aid.
+}
 \cite{chakrabarti2018effect,NBERw27885} use the same instrument for public university revenues at the state level, combined with rich information on students' individual-level outcomes, to show that increases in state appropriations lead to degree lower completion time and later-life student debt.

### 2022-08-15 text/paper.tex
adjust script and README.

@@
     Cornell University, Department of Economics.}
-    \thanks{I thank Professor Evan Riehl for guidance on this project, Michael Lovenheim, Ronald Ehrenberg, Douglas Miller, and Francine Blau for various comments on the project's direction,
-    as well as seminar participants at Cornell University (Autumn 202) for helpful
-    discussion.
+    \thanks{I thank Professor Evan Riehl for guidance on this project, Michael Lovenheim, Ronald Ehrenberg, Douglas Miller, and Francine Blau for various comments on the project's direction, as well as seminar participants at Cornell University (Autumn 2022) for helpful discussion.
     This project's Github repository, with materials and (publicly available) data for replication, is available at 
@@
 Findings show that negative changes in state funding affects faculty composition at public universities, away from tenured and tenure-track professors towards non-tenure track lecturer positions.
-A reduction of 10\% in state funding, via a shock to state appropriations, leads to a fall in 1.4\% in the number of assistant professors per student within a university, 1.2\% for full (tenured) professors, and an increas in 4.4\% the number of lecturers per student.
+A reduction of 10\% in state funding, via a shock to state appropriations, leads to a fall in 1.4\% in the number of assistant professors per student within a university, 1.2\% for full (tenured) professors, and an increase in 4.4\% the number of lecturers per student.
 Additionally, I use individual level data on professors at Illinois public universities over 2010-2021 to investigate whether the shocks to state funding affected incumbent professors at public universities.
-Incumbent professors were not meaningfully affected, in terms of total salary, promotion rate, or rate of leaving the Illinois public university system, yet lecturers' first year slary decreased by 3.8\% in response to a fall in state funding of 10\%.
+Incumbent professors were not meaningfully affected, in terms of total salary, promotion rate, or rate of leaving the Illinois public university system, yet lecturers' first year salary decreased by 3.8\% in response to a fall in state funding of 10\%.
 These findings imply that changes in faculty composition came about by changes in the composition of hiring new professors at the university level.
@@
 \cite{brown2014endowment} presents the closest example by studying how university endowments react to negative financial shocks, exhibiting an association between negative endowment shocks and a fall in tenure-system faculty employed at private universities.
-\cite{abe2015implications} present a model for university administrator decision-making, and posit that a large component of variation in faculty salaries arises from university politics, and not only economic factors such as outside options.
+\cite{abe2015implications} present a model for decision-making by university administrators, and posit that a large component of variation in faculty salaries arises from university politics, and not only economic factors such as outside options.
 \cite{johnson2009jep,NBERc13879} both empirically document allocation of faculty between research and instruction, and \cite{hemelt2021math} document differences in instructional cost (per student) between departments, primarily thanks to department differences in class size and faculty salaries.
@@
     Summary statistics for this group, which over-represents lecturers compared to assistant and full professors, can be seen in \autoref{tab:illinois-summary-rolling}.
-    The reasons for restricting to the sample with identified year of hire in the range 2011-2021 is explained in \autoref{sec:iv-model-indiv}.
+    The reasons for considering the sub-sample with identified year of hire in the range 2011-2021 are explained in \autoref{sec:iv-model-indiv}.
 }
@@
 A state government plans an annual budget a couple of years ahead of the fiscal year, and the legislature chooses to approve or reject a budget request put forth by the governor's office.\footnote{
-    \cite{NBERw23736} present a full discussion of the state appropriation decision-making process, drawing on administrative records originally analysed by \cite{parmley2009state}.}
+    \cite{NBERw23736} present a full discussion of the decision-making process for state appropriations, drawing on administrative records originally analysed by \cite{parmley2009state}.
+}
 US state governments, by majority, are legally obligated to run a balanced-budget, so that yearly variation in tax revenues (caused by changing economic conditions or otherwise) necessarily affect state expenditures.
@@
 
-A professor is employed by a university, so it stands to reason that when public universities experience a secular fall in one of their revenue soruces revenue --- analogous to a demand shock in the private sector --- that the composition of their employees will be affected.
+A professor is employed by a university, so it stands to reason that when public universities experience a secular fall in one of their revenue sources revenue --- analogous to a demand shock in the private sector --- that the composition of their employees will be affected.
 
@@
 
-\autoref{tab:firststage-reg} presents results of the first-stage regression, separately with and without a control for per student tuition revenue and year plus institution fixed effects.\footnote{
+\autoref{tab:firststage-reg} presents results of the first-stage regression, separately with and without a control for tuition revenue per student, plus institution and year fixed effects.\footnote{
     Representations for frequentist significance levels (i.e. 10, 5, 1\% etc.) are omitted here, and in all following tables.
 }
-We note columns (1) and (2) estimate that a shock to state appropriations per student in the state of 10\% is associated with around 9.8\% change in state funding per student at the university; the instrument is strong, and we note similarity in estimates with and without inclusion of fixed effects.
+We note columns (1) and (2) estimate that a shock to appropriations (per student in the entire state) of 10\% is associated with 9.8\% change in state funding per student at the university; the instrument is strong, and we note similarity in estimates with and without inclusion of fixed effects.
 Columns (3) and (4) include the tuition revenue control (explained in \autoref{foot:control}) to exhibit estimates with the inclusion of this possible collider or bad control.
@@
     Identifying $\tau(j)$ is possible for $j$ by restricting to all professors hired 2011-2021 --- i.e. in the years after the start of the full panel.
-    It is not possible to discern the hiring year for professors who  were hired in the years preceding 2011, and so the entire sample is only possible to analyse using the base-share in years 1990-1993 formulation (e.g.
-\autoref{tab:facultysalaries-shock-illinois}, \ref{tab:promotion-shock-illinois}).
+    It is not possible to discern the hiring year for professors who  were hired in the years preceding 2011, and so the entire sample is only possible to analyse using the base-share in years 1990-1993 formulation (e.g., \autoref{tab:facultysalaries-shock-illinois}, \ref{tab:promotion-shock-illinois}).
 }
@@
 Exogeneity and relevance of the rolling-share instrument, $\tilde Z_{j,t}$, follows the same reasoning as that for the base-share instrument, $Z_{i,t}$, discussed in \autoref{sec:approp-shocks}.\footnote{
-    The base-share instrument is still used for some outcomes with the individual Illinois professors, where appropriate (\autoref{tab:facultysalaries-shock-illinois}, \ref{tab:promotion-shock-illinois}).
+    The base-share instrument is appropriate for some outcomes with the individual Illinois professors, where appropriate (\autoref{tab:facultysalaries-shock-illinois}, \ref{tab:promotion-shock-illinois}).
 }
@@
 
-Together these findings show that public universities increase their count of tenure-track and tenured professors per student in years where revenues are more plentiful, possibly by increasing hiring intensity or efforts to maintain incumbent professors.
+Together these findings show that public universities increase their count of tenure-track and tenured professors per student in years when revenues are more plentiful, possibly by increasing hiring intensity or efforts to maintain incumbent professors.
 Yet, when revenues are not so bountiful, count of tenure-track and tenured professors per student decreases.
@@
 \autoref{tab:newhiresalaries-shock-illinois-rolling} presents 2SLS estimates for the elasticity of total salaries with respect to university revenues, among the sample with a year of starting employment after 2010.
-New hires of assistant, full, or administrative position professors are not realted to revenues, via a shock to appropriations in their first-year of employment.
+New hires of assistant, full, or administrative position professors are not related to revenues, via a shock to appropriations in their first-year of employment.
 Yet, the starting salary of lecturers is 3.8\% higher given an increase in their hiring university's state funding of 10\%.

### 2022-08-15 programs/analysis/ipeds-shock.R

@@ -103,7 +103,7 @@ reg.data %>%
         digits = 0,
         digits.extra = 0,
         covariate.labels = c(
-            "Enrollment, full-time equivalent",
+            "Enrollment, institution reported",
             "State appropriations (millions 2021 USD)",
             "Total revenues (millions 2021 USD)",
             #"Non-instutional revenues (millions 2021 USD)",

### 2022-11-20 text/paper.tex
Updated draft.

--- 2022-08-12 text/paper.tex
@@
 \usepackage{pgfplots} \pgfplotsset{compat=1.16}                            % Automatic graphing, read https://www.overleaf.com/learn/latex/pgfplots_package for examples
-\usepackage{subcaption}                                                    % Sub figures.
 % Maths + numbers
@@
 \hypersetup{colorlinks=true, linkcolor=blue, citecolor=blue, filecolor=magenta, urlcolor=blue}
-\def\equationautorefname~#1\null{Equation~(#1)\null}                       % Fix autoref for quations
+\def\equationautorefname~#1\null{Equation~(#1)\null}                       % Fix autoref for equations
 % IMPORTANT: follow style guide here https://github.com/Wookai/paper-tips-and-tricks
@@
 \author{Senan Hogan-Hennessy\thanks{
-    Cornell University, Department of Economics.}
-    \thanks{I thank Professor Evan Riehl for guidance on this project, Michael Lovenheim, Ronald Ehrenberg, Douglas Miller, and Francine Blau for various comments on the project's direction,
-    as well as seminar participants at Cornell University (Autumn 202) for helpful
-    discussion.
+    Economics Department, Cornell University.}
+}
+% Title
+\title{Changes in Faculty Composition and Stagnating State Support for Higher Education
+    }
+\date{\today\thanks{
+    I thank Professor Evan Riehl for guidance on this project, Michael Lovenheim, Ronald Ehrenberg, Douglas Miller, Francine Blau, and Ryan Dycus for various comments on the project's direction, as well as seminar participants at Cornell University (Autumn 2022) for helpful discussion.
     This project's Github repository, with materials and (publicly available) data for replication, is available at 
-    \url{https://github.com/shoganhennessy/finances-faculty-composition}.
-    Any comments or suggestions may be sent to me at \href{mailto:seh325@cornell.edu}{\nolinkurl{seh325@cornell.edu}}, or raised as an issue on the Github project.}}
-% Title
-\title{Stagnating State Support for Higher Education and Its Effect on Faculty Composition}
-\date{\today}
+    \url{https://github.com/shoganhennessy/state-faculty-composition}.
+    Any comments or suggestions may be sent to me at \href{mailto:seh325@cornell.edu}{\nolinkurl{seh325@cornell.edu}}, or raised as an issue on the Github project.
+    }
+}
 \rhead{\today}
-\lhead{State Support for Higher Education and Faculty Composition}
+\lhead{Faculty Composition and State Support for Higher Education}
 % Begin
 \begin{document}
-\clearpage \maketitle
+\clearpage
+\maketitle
 \thispagestyle{empty}
@@
 \begin{abstract}
-\noindent
-State support for higher education has stagnated on a per student level the last thirty years, while public universities have reduced the number of tenure-track or with tenured professors per student, and increased their employment of contingent lecturers.
-This analysis relates these changes in the faculty composition at public universities to the fall in per student university revenues from state appropriations.
-A shift-share instrumental variables approaches addresses endogeneity in decisions for state appropriations to higher education, by exploiting the yearly change in state support with an institution's reliance on that support.
-A decrease in state funding of 10\%, via a shock to state appropriations, decreases the number of assistant professors per student at a public university by 1.4\% and full professors by 1.2\%, yet increases the number of lecturers per student by 4.4\%.
-Analysis of all the professors hired at Illinois public universities 2011-2021 shows that incumbent professors are not affected by the changes in state support, implying that these changes in faculty composition arose by lower rate of hiring of tenure-track and tenured professors at public universities.
-These results show one of the long-term effects of stagnating state support for higher education, and raises questions about the direction that public education heads as these financial headwinds show no sign of dissipating.
+    \input{sections/00-abstract.tex}
+\end{abstract}
 
-\vspace{0.75cm}
-\noindent\textbf{Keywords:}
-State and Local Budget and Expenditures,
-Higher Education,
-Public Sector Labor Markets
-
-\vspace{0.5cm}
-\noindent\textbf{JEL Codes:} H72, I23, J45
-\end{abstract}
-\newpage
-\setcounter{page}{1}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newpage
+\setcounter{page}{1}
 \doublespacing
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Introduction section
-\noindent
-Public universities educate the majority of higher education students in the US, yet have experienced a secular decline in state funding (per student) the last three decades.
-This decline has been shown to lead to worse education and later-life outcomes for students \citep{NBERw23736,NBERw27885}, yet it is not clear what mechanisms drive these effects, or how public universities are affected as institutions.
-I show that four-year, degree-granting public universities have experienced falling state funding per student, and this affected their composition of faculty.
-Falls in public university revenues from the state, instrumented by state-level finance shocks, lead to a fall in the number of tenure-track or tenured professors per student within a university, an increase in the number of lecturers, and overall fall in the number of faculty per student.
-US private universities were not exposed to such financial constraints during this same time period, and do not exhibit the substitution away from tenure-track and tenured professors, so that the stagnating state support for public universities has implications for the wider structure of higher education instruction and research in the US.
-
-US universities are widely considered the highest performing in the world, yet there are consequential differences between its universities that operate in the private sector and those established by state governments.
-Public universities are subject to numerous state-level administrative laws, and rely on their state governments for funding: an average public university received around \$11,600 per enrolled student in 1990, yet only \$8,300 per enrolled student in 2021.
-This fall is driven by stagnating funding provided by state governments, while enrolment among public universities rose by 46\% over the same time period.
-At the same time, the number of professors per student at public universities stayed relatively stable at 0.045, yet fell from around 0.04 to around 0.35 professors per student at public universities, driven by a fall of over 20\% in the ratio of associate and full professors per student at public universities.
-
-I use a variant of the \cite{NBERw23736,NBERw27885} shift-share instrument for state funding from shocks to appropriations for the entire state to identify the effect on the faculty composition.
-State appropriations for public universities are plausibly endogenous to many outcomes: state governments decide on yearly budgets for their higher education sector, and this process can be influenced by local financial conditions, or even public perceptions of the state's higher education system.
-The shift-share instrument interacts reliance on state funding in a base period with the yearly total appropriations per student in the entire state, to exploit both changes in support for higher education and how much each university relies on that support.
-
-Findings show that negative changes in state funding affects faculty composition at public universities, away from tenured and tenure-track professors towards non-tenure track lecturer positions.
-A reduction of 10\% in state funding, via a shock to state appropriations, leads to a fall in 1.4\% in the number of assistant professors per student within a university, 1.2\% for full (tenured) professors, and an increas in 4.4\% the number of lecturers per student.
-Additionally, I use individual level data on professors at Illinois public universities over 2010-2021 to investigate whether the shocks to state funding affected incumbent professors at public universities.
-Incumbent professors were not meaningfully affected, in terms of total salary, promotion rate, or rate of leaving the Illinois public university system, yet lecturers' first year slary decreased by 3.8\% in response to a fall in state funding of 10\%.
-These findings imply that changes in faculty composition came about by changes in the composition of hiring new professors at the university level.
-
-\cite{NBERw23736} first study the effect of (changes in) public universities' finances on student outcomes by isolating changes in public universities' expenditures on student instruction, addressing endogeneity of expenditures decisions by instrumenting expenditures with state appropriation shocks and limits on tuition price.
-\cite{NBERw23736} find that increases in public university spending (via appropriation shocks) increases enrolment and degree completion among students, but corresponding changes in attendance price (via limits on tuition price) do not have an effect.
-\cite{chakrabarti2018effect,NBERw27885} use the same instrument for public university revenues at the state level, combined with rich information on students' individual-level outcomes, to show that increases in state appropriations lead to degree lower completion time and later-life student debt.
-\cite{bound2019public} use another variant of the same approach to show that state-level higher education spending cuts induced public universities to shift toward tuition as their primary source of revenue, and shift institutional resources towards students who pay higher tuition rates.
-In a similar vein, \cite{bound2007cohort} document variation in per student state funding for higher education resulting from yearly changes in a state's student-age population; they find that lower per student funding lead to lower higher education completion rate among students.
-
-Faculty are a core component of this country's educational and research-innovation university system, yet their composition within universities, and its relationship with higher education finances has so far not been studied.
-\cite{brown2014endowment} presents the closest example by studying how university endowments react to negative financial shocks, exhibiting an association between negative endowment shocks and a fall in tenure-system faculty employed at private universities.
-\cite{abe2015implications} present a model for university administrator decision-making, and posit that a large component of variation in faculty salaries arises from university politics, and not only economic factors such as outside options.
-\cite{johnson2009jep,NBERc13879} both empirically document allocation of faculty between research and instruction, and \cite{hemelt2021math} document differences in instructional cost (per student) between departments, primarily thanks to department differences in class size and faculty salaries.
-\cite{turner2014impact} document how universities reacted to the Great Recession of 2008, a severe financial shock to state finances and support for higher education.
-
-The number of professors, both in absolute terms and relative to number of students, may be particularly important for quality of instruction and thus educational outcomes.
-\cite{angrist1999using} first causally show that reducing class size (via religious ruling) induces an increase in test scores for school-age children, yet the magnitude of this effect is not as clear in the higher education setting.
-\cite{bandiera2010heterogeneous} use a fixed effect model, combined with multiple individual observations in a long panel, to find that UK university students perform worse academically in particularly large classes, and the difference is largest for students at the top of the academic distribution.
-For the composition of faculty teaching university classes, \cite{bettinger2010does,figlio2015tenure} find US students have better enrolment and learning outcomes from courses taught by contingent or adjunct (i.e.
-not tenured or on tenure-track) professors (relative to tenured or tenure-track).
-\cite{ehrenberg2005tenured} observe a negative association between utilisation of non-tenured instructors and student graduation rates.
-My approach focuses on count of professors per student at universities, so does not directly observe individual class-sizes, yet presents considers this a possible mechanism through which stagnating state support for higher education negatively affects student outcomes \citep{NBERw23736,NBERw27885}.
-
-%\begin{table}[H]
-%    \centering
-%    \caption{Table to Consider Impact of a Given Paper.}
-%    \begin{tabular}{l | c c c c c}
-%        & concepts & relationships & models & theories & use of methods \\
-%        \midrule
-%        
-%        change:
-%        & concepts & relationships & models & theories & use of methods \\
-%        
-%        challenge:
-%        & concepts & relationships & models & theories & use of methods \\
-%        
-%        fundamentally alter:
-%        & concepts & relationships & models & theories & use of methods \\
-%    \end{tabular}
-%\end{table}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Data section
-\section{Data and Institutional Context}
-\label{sec:data}
-
-\subsection{Data Description}
-The data used in this analysis come from two primary sources: \citet[IPEDS]{ipeds} regarding institutional information on finances and enrolment, and \citet[IBHED]{ibhed} for data on every professor in the Illinois public university system.
-
-IPEDS is a survey of higher educational institutions in the US, and legally requires institutions to participate in order to receive Federal Title IV student aid.\footnote{
-    This statement means that IPEDS does not necessarily cover the universe of US higher education institutions, yet in practice every public university and not for-profit four-year institution is represented.
-}
-Data are consistent between the years 1990 and 2021,\footnote{
-    The years 1987-1989 are represented in these data in an incompletion fashion, so I focus on the years 1990 onwards.
-    Year refers to the calendar year of the spring term --- i.e. 1987 refers to the academic year that ran August 1986 to July 1987.
-}
-and provide information on the total revenues a university received from every source (including state governments), enrolment number, plus faculty count and total expenditures on salaries.\footnote{
-    I combine the Urban Institute's 2018 compilation of IPEDS data for the years 1990-2017, and manually combine with raw NCES data regarding the year 2018-2021 for all relevant variables.
-    Figures for enrolment, faculty counts and salaries come from the raw NCES version of IPEDS for all years to address inconsistencies in the Urban Institute's data formulation for these variables.
-}
-I restrict analysis to public, four-year, degree-granting institutions, as these institutions adhere to a standardised concept of faculty profile, where tenure and title of appointment (lecturer, assistant professor, etc.) are relatively standardised.
-For-profit institutions employ and enrol a negligible share of professors and students respectively, while students at two-year institutions by majority intend to eventually enrol at a four-year institution \citep{mountjoy2022},\footnote{
-    Furthermore, \cite{mountjoy2022} documents the effects of expanding two-year higher education access in the US separately for students who would have and those who would not have otherwise have attended college.
-    The high rate of eventual enrolment at four-year institutions among two-year students explains the negative impact of increased access to two-year institutions on students who are diverted away from four-year institutions.
-}
-so that these institutions are not considered.
-
-IPEDS reports the count of professors employed by position, as well as total salary expenditures by position.
-This gives a resulting panel data-set, where each row represents a university-year, and includes columns for university finances, plus total count and average salary\footnote{
-    Real salary is computed by scaling nominal salary to 2021 dollars by the CPI-U.
-} for professors by position (lecturer, assistant, tenured, total).
-\autoref{tab:ipeds-summary} presents summary statistics for relevant variables in these data.
-
-\begin{table}[h!]
-    \singlespacing
-    \centering
-    \caption{IPEDS Summary Statistics, Public Universities Panel 1987--2021}
-    \makebox[\textwidth][c]{\input{tables/ipeds-summary-fte.tex}}
-    \label{tab:ipeds-summary}
-\end{table}
-
-IPEDS provides information at the university level, yet lacks information on the distribution of salaries and professor count within a university.\footnote{
-    IPEDS provides total paid to salary and total faculty employed per university-year, so that I can investigate average professor salary per university-year with IPEDS data.
-    Unfortunately, this measure of professors' salaries is particularly crude in measuring on professors' salaries and other individual-outcomes; summary statistics on IPEDS data do not agree with trends in average professor salary over the sample time period \citep{aau2021survey}.
-    Similarly, IPEDS reports of the average salary paid to professors at private institutions also disagree with other sources, so are not considered in this analysis.
-}
-To investigate the distribution, I integrate individual-level data for every public university professor in the state of Illinois between the years 2010-2021.
-IBHED hosts the information;
-Public Act 96-0266 (effective 1 January 2010) requires that each university report base salary and benefits all administrators, faculty members, and instructors employed by the college or university.\footnote{
-    The universities included are all campuses of the nine Illinois public universities: Chicago State University, Eastern Illinois University, Governors State University, Illinois State University, Northeastern Illinois University, Northern Illinois University, Southern Illinois University  (all five campuses), University of Illinois (all four campuses), Western Illinois University.
-}
-This publicly available data provides the basis to build a panel of Illinois public university professors 2010-2021; I define a professor as an individual by their first plus last name and university pairing, and link this database to IPEDS regarding finances for their employing institution.
-
-The analysis sample represents 16,932 professors in the year 2010 and 15,352 in the year 2021, with summary statistics presented in \autoref{tab:illinois-summary}.
-Additionally, further analysis using the first year of a professor's employment focuses on the subset of professors with observed year of hiring between the years 2011 and 2021, representing 1,778 professors in 2011, and 9,099 in 2021.\footnote{
-    Summary statistics for this group, which over-represents lecturers compared to assistant and full professors, can be seen in \autoref{tab:illinois-summary-rolling}.
-    The reasons for restricting to the sample with identified year of hire in the range 2011-2021 is explained in \autoref{sec:iv-model-indiv}.
-}
-
-\begin{table}[h!]
-    \singlespacing
-    \centering
-    \caption{IBHED Summary Statistics, Professor Panel 2010--2021.}
-    \makebox[\textwidth][c]{\input{tables/illinois-summary.tex}}
-    \label{tab:illinois-summary}
-\end{table}
-
-\subsection{Trends in Finances, Enrolment, and Faculty Composition}
-\label{sec:trends}
-A state government plans an annual budget a couple of years ahead of the fiscal year, and the legislature chooses to approve or reject a budget request put forth by the governor's office.\footnote{
-    \cite{NBERw23736} present a full discussion of the state appropriation decision-making process, drawing on administrative records originally analysed by \cite{parmley2009state}.}
-US state governments, by majority, are legally obligated to run a balanced-budget, so that yearly variation in tax revenues (caused by changing economic conditions or otherwise) necessarily affect state expenditures.
-This process leads to yearly variation in state appropriations not seen in other revenues sources (such as federal government appropriations), since US states differ in their support for higher education, and are subject to fiscal constraints.
-State appropriations to higher education are a particularly attractive area of state spending to absorb such shocks to state finances \citep{delaney2011state}.\footnote{
-    \cite{delaney2011state} fully describes the financial environment of state expenditures, and what makes spending on higher education an attractive area for state governments to expand funding during years of higher tax revenues, and retract funding in leaner years.
-    An analysis of state expenditures for the years 1980-2004 (overlapping with the sample for this analysis) provides solid evidence for these trends.
-}
-Additionally, there is yearly variation in the number of higher education students in each state \citep{turner2014impact}, so that per-student state appropriations vary on multiple dimensions.
-
-\autoref{fig:funding} exhibits the trends for the mean public university for the years 1990-2021.
-We see a rise in total revenues received by public universities (from all sources), and a notable increase in mean tuition revenues from \$48 million per year-university to \$150 million per year-university.
-At the same time, total state appropriations stagnated at around \$100 million per year-university for 1990-2008, falling around 2008 and have not recovered ever since.
-While public universities experienced a stagnation in state support, private universities were not exposed to the same constraints, receiving \$37,000 per student in 1990 and \$49,000 in 2021, experiencing no corresponding decline in any specific component.
-
-\begin{figure}[h!]
-    \centering
-    \singlespacing
-    \caption{Mean Total Revenues among Public Universities, by Year.}
-    \begin{subfigure}[b]{0.495\textwidth}
-        \centering
-        \caption{Total, millions \$ 2021 CPI-U.}
-        \includegraphics[width=\textwidth]{figures/mean-funding-total.png}
-        \label{fig:mean-funding-total}
-    \end{subfigure}
-    \begin{subfigure}[b]{0.495\textwidth}
-        \centering
-        \caption{Per Enrolled Student, \$ 2021 CPI-U.}
-        \includegraphics[width=\textwidth]{figures/mean-funding-fte.png}
-        \label{fig:mean-funding-fte}
-    \end{subfigure}
-    \label{fig:funding}
-\end{figure}
-
-\begin{figure}[h!]
-    \centering
-    \singlespacing
-    \caption{Total Student enrolment, by University Sector, and Year.}
-    \begin{subfigure}[b]{0.495\textwidth}
-        \centering
-        \caption{Total.}
-        \includegraphics[width=\textwidth]{figures/enrollment-total.png}
-        \label{fig:enrollment-total}
-    \end{subfigure}
-    \begin{subfigure}[b]{0.495\textwidth}
-        \centering
-        \caption{Mean, University Level.}
-        \includegraphics[width=\textwidth]{figures/enrollment-mean.png}
-        \label{fig:enrollment-mean}
-    \end{subfigure}
-    \label{fig:enrolment}
-\end{figure}
-
-At the same time, student enrolment at public universities rose precipitously.
-6.2 million students were enrolled in public universities in 1990, and this number rose by 47\% to 9.1 million, with most of the increase occurring over the years 2000-2021.
-\autoref{fig:enrolment} shows that total enrolment at private universities has also risen over the same time period, but not as drastic in either relative or absolute terms; the mean private university grew from 9,800 students in 1990 to 11,800 in 2021.
-This means that revenues per student have stagnated across all measures (seen in \autoref{fig:mean-funding-fte}), and particularly fallen from a mean of \$11,000 per student in 1990 to less than \$8,000 per student in 2021.
-
-\begin{figure}[h!]
-    \centering
-    \singlespacing
-    \caption{Total Professor Count per Student, by University Sector, Professor Appointment, and Year.}
-    \begin{subfigure}[b]{0.495\textwidth}
-        \centering
-        \caption{Lecturer}
-        \includegraphics[width=\textwidth]{figures/lecturer-fte-perprof.png}
-        \label{fig:lecturer-fte-perprof}
-    \end{subfigure}
-    \begin{subfigure}[b]{0.495\textwidth}
-        \centering
-        \caption{Assistant.}
-        \includegraphics[width=\textwidth]{figures/assistant-fte-perprof.png}
-        \label{fig:assistant-fte-perprof}
-    \end{subfigure}
-    \begin{subfigure}[b]{0.495\textwidth}
-        \centering
-        \caption{Full.}
-        \includegraphics[width=\textwidth]{figures/full-fte-perprof.png}
-        \label{fig:full-fte-perprof}
-    \end{subfigure}
-    \begin{subfigure}[b]{0.495\textwidth}
-        \centering
-        \caption{All.}
-        \includegraphics[width=\textwidth]{figures/all-fte-perprof.png}
-        \label{fig:all-fte-perprof}
-    \end{subfigure}
-    \label{fig:fte-perprof}
-\end{figure}
-
-\autoref{fig:fte-perprof} documents the divergence in faculty composition (per student) between the average private and public university 1990-2021.
-Private universities start with  a higher baseline of around 4.5 professors per 100 students, and exhibited yearly variation of less than 0.5 professor per hundred students over the thirty years.
-Public universities start with 3.9 professors per hundred students, and this number falls to around 3.5 primarily in the 2008-2011 time period.
-We see a similar difference in baseline, and fall for the years 2008-2011 for assistant professors.
-Private and public universities have similar numbers of associated and full professors before the year 2000, yet this number has fallen by over 20\% in the next 20 years only for public universities: in 2021 the mean public university has 6 fewer full professors per hundred students than the mean private university.
-Over the same time period, we see the rise in use of non-tenure track instructor positions (referred to as lecturers from here on), who were employed at similar rates in both sectors in 1990 yet have been utilised by public universities at a higher rate since.
-
-\subsection{Trends in Illinois}
-\label{sec:trends-illinois}
-
-The state of Illinois experienced the same stagnation in state support for higher education that the rest of the nation has experienced 1990-2021.
-\autoref{fig:illinois-funding} shows the trends for revenues among the seven Illinois public university campuses, including the falling share of state appropriations and substitution towards tuition revenue.
-
-\begin{figure}[h!]
-    \centering
-    \singlespacing
-    \caption{Mean Revenues among Illinois Public Universities, by Year.}
-    \begin{subfigure}[b]{0.495\textwidth}
-        \centering
-        \caption{Total, millions \$ 2021 CPI-U.}
-        \includegraphics[width=\textwidth]{figures/illinois-funding-total.png}
-        \label{fig:illinois-funding-total}
-    \end{subfigure}
-    \begin{subfigure}[b]{0.495\textwidth}
-        \centering
-        \caption{Per Enrolled Student, \$ 2021 CPI-U.}
-        \includegraphics[width=\textwidth]{figures/illinois-funding-fte.png}
-        \label{fig:illinois-funding-fte}
-    \end{subfigure}
-    \label{fig:illinois-funding}
-\end{figure}
-
-The decade of 2010-2021 has also seen yearly variation in its revenues, particularly around 2016.
-In the calendar year 2015 partisan disagreements between the democratic legislature and republican governor led to the 2016 fiscal year starting with no state budget.
-State agencies, and higher education institutions, employed accounting techniques to continue operating without any resources provided by the state government.\footnote{
-    Fiscal year 2016 refers to June 205 to June 2016, as is the same for the academic year definition in this analysis. 
-    K-12 schools were unaffected thanks to a separate budget agreement allowing non-higher education to operate as before.
-}
-While state institutions were able to stay open, there were drastic revenue and spending cuts in response to the budget impasse, as it continued through fiscal year 2017, and ended with a new budget restoring revenues to state agencies and universities for the 2018 fiscal year.
-This variation in public university revenues via the state appropriations channel, stemming entirely from political disagreements --- not from state decisions regarding higher education and its finances \citep{young2020squandered} ---  mean that Illinois public universities exhibit sizable changes in their state appropriations over 2010-2021, of similar order to those for the rest of the country over 1990-2021.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Section on possible outcomes
-\subsection{University Responses to Financial Shocks}
-\label{sec:responses}
-
-While state support for public higher education has stagnated at the same time as education costs rose, there are multiple possible ways that a university can respond.
-\cite{NBERw23736} established that corresponding rises in tuition did not offset the falling state support, so that university spending (per student) fell in response to these persistent, negative state funding shocks.
-There are multiple ways that these changes in finances may affect the faculty composition at public universities.
-
-Universities every year hire a number of new professors, either to expand their departments or to replace leaving professors.
-The number of professors hired is usually highest among non-tenured adjunct faculty, as these instructors are by majority hired on short-term contingent contracts, so that they are less costly to hire (or to not reinstate) in response to yearly changes in teaching needs.
-Secondly, tenure-track assistant professors are hired in most years at public universities to replace leaving assistant professors, and are most often hired with a four to six year contract and agreement to be considered for tenure at the end of this term.
-Lastly, tenured professors are those that have undergone review and have successfully secured a full-time appointment at their university with no expiration date; tenured professors are sometimes hired from outside a university, though this position has the lowest hiring rate among most universities.
-It is common for universities to restrict their faculty hiring in the face of financial shocks.
-Since the yearly hiring rate is highest among lecturers and assistant professors, financial shocks may lower the number of lecturer and assistant professors at public universities the most, if faculty hiring is affected by falls in state funding.
-
-Additionally, professors' salaries may be disrupted by the stagnation in public university finances.
-When the university has a lower budget for its yearly hiring, it may respond by lowering the salary they offer to new hires.
-This possible effect may not be the same across each position of professor; tenured faculty are often hired away from another university, so that new tenured professor hires could be less likely to accept a lower offer from a public university, and we would see that new tenured hires are not affected salary-wise.\footnote{
-    \cite{blackaby2005} describes the academic outside options among economist faculty, documenting how they differ for academic position and gender.
-}
-Yet salary for all the professors, not just new hires, may also be affected: multiple universities passed a university-wide pay-cut for their faculty in response to state budget cuts around the 2008 recession, for example.\footnote{
-    Indeed, Cornell University implemented both hiring freezes and even a nominal salary reduction for professors, in anticipation of a financial shock in early 2021.
-    The salary cut was not permanent, as the oncoming financial shock turned out to not be as serious as projected, so that the salary cut was returned to professors in the year 2021.
-}
-Faculty are not paid the same by position, so that their are financial consequences for faculty being promoted between positions within a university.
-In response to tightening fiscal restraints, universities may be less likely to grant tenure to assistant professors, or to promote tenured professors from associate to full professor either.
-
-It is not immediately clear which effect will dominate, and which position of professor will be effected the most by changes in their university's finances.
-Yet, there is one empirical fact worth noting: a mean assistant or tenured professor earns more than double that of an adjunct lecturer (\autoref{tab:illinois-summary}), and has more long term salary obligations than a lecturer too.
-If a public university's primary obligation is to teach, and they must fulfil this objective with fewer and fewer resources, then they may substitute away from tenure-track and tenured professor towards contingent lecturers to achieve this obligation.\footnote{
-    That is substituted away from tenure-track and tenured professors according to the salary ratio, and the relative productivity between these two positions.
-    It is not possible to estimate the relative productivity in this paper, so I leave the rate of substitution between the two positions to further research.
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Empirical section
-\section{Empirical Framework}
-\label{sec:empirics}
-
-A professor is employed by a university, so it stands to reason that when public universities experience a secular fall in one of their revenue soruces revenue --- analogous to a demand shock in the private sector --- that the composition of their employees will be affected.
-
-Na\"ively we can analyse the relationship between professor-outcomes $Y_{i,t}$ at university $i$ in year $t$ as a result of state funding per student $X_{i,t}$.
-Outcomes include count of professors per student, and average salary, within each university-year.
-This model gives $\beta$ the association with these outcomes and state funding per student,\footnote{
-    Note that dividing by student count also implicitly controls for the size of the university, so that this model accounts for yearly variation in professor count and university revenues arising from growth in a university.
-}
-including fixed effects to control for effects specific to the university and year.
-Log$(.)$ transforming the variables improves the interpretability of $\beta$ as an elasticity for professor count per student with respect to state funding per student.
-\begin{equation}
-    \label{eqn:naivereg}
-    Y_{i,t} = \alpha_i + \gamma_t + \beta X_{i,t} + \epsilon_{i,t}
-\end{equation}
-
-Yet a university's finances are not exogenous to state decisions on the yearly level for support of higher education, or exogenous to internal institutional decisions.
-Instead, the state government and university administration undertake a complex process of allotting resources across multiple different priorities, including instruction, research, or between departments.
-Importantly for this analysis, revenues received from an institution's state government provide opportunity to address this endogeneity.
-
-\subsection{State Appropriation Shocks}
-\label{sec:approp-shocks}
-
-\cite{NBERw23736,chakrabarti2018effect,NBERw27885} address endogeneity in public university finances by exploiting a shift-share instrument for changes in state-level funding interacted with university reliance on state funding in a base period.
-The system exploits the fact that institutions who rely on state appropriations more will be affected by state appropriation shocks.
-\begin{align}
-    \label{eqn:public-instrument}
-    Z_{i,t} &\coloneqq \log \left[
-    \left( \frac{\text{Total State Funding}_{s(i),t}}{\text{Student Population}_{s(i),t}} \right)
-    \sum_{\tau = 0}^{3} \frac 14
-    \left( \frac{\text{State Funding}_{i,1990 + \tau}}{\text{Total Revenues}_{i,1990 + \tau}} \right) \right]
-\end{align}
-
-$Z_{i,t}$ is the instrument for (log) state appropriations for institution $i$ in year $t$, interacting the average funding for university $i$ in state $s(i)$ with reliance on state funding relative to total revenues, averaged across the base years 1990--1993.\footnote{
-    1990--1993 are defined as data for public university finance data are most comparable (i.e. without many missing values) beginning in 1990.
-    \cite{NBERw23736} use the single year 1990 as the base year, though I use the four years to ameliorate missing values in the single year of 1990.
-    Results are similar in either specification.
-}
-\cite{NBERw27885} notes the tendency for public universities to respond to state funding cuts by increasing reliance on tuition, where \cite{NBERw23736} specifically instruments for tuition revenues with collected information on legislative tuition price controls.\footnote{
-    \label{foot:control}
-    It may be argued that tuition revenues are confounder between the causal effect of changes in state funding on a university's total revenues, so that this analysis focuses on state support for higher education (no total revenue) as a result.
-    This follows the \cite{NBERw27885} formulation.
-    On the other hand, rises in tuition revenues (per student) may arise as result of tuition hikes thanks declining state support, which would mean controlling for tuition would constitute a bad control.
-    Estimates including tuition revenues (per student) as a control in the second stage of the IV estimates produces results of very similar magnitude and direction, and so are omitted.
-}
-This analysis focuses on one important source of revenues for public universities: state appropriations.
-
-The first-stage is then as follows, including institution and year fixed effects.
-\begin{equation}
-    \label{eqn:firststage}
-    X_{i,t} = \eta_i + \zeta_t + \delta Z_{i,t} + \epsilon_{i,t}
-\end{equation}
-We note the conditions for exogeneity in the instrument (following the discussion presented by \citealt{NBERw27885}).
-The instrument is exogenous if state policy decisions for funding of public universities are uncorrelated with unobserved institutional changes of any specific college or university in the state \citep{borusyak2022quasi}.
-This assumption is plausible given that the majority of states have multiple (i.e.
-more than five) public universities, without any single university campus receiving the majority of state funding within any single state.
-Secondly, addressing endogeneity in university revenues by the shift-share identification strategy requires exogeneity in either the base-line share or shift component of the instrument.
-In this case, we satisfy the second: universities' institutional-level decisions are not correlated with contemporaneous or upcoming shocks to state appropriations.\footnote{
-    It would be plausible to consider the case that universities make institutional-level decisions in a consistently different manner to those 
-    with differing reliance on state appropriations in 1990, so that exogeneity by the base-line share is not plausible here.
-}
-
-\begin{table}[h!]
-    \singlespacing
-    \centering
-    \caption{First Stage Estimates, for University Revenues by Appropriation Shock.}
-    \makebox[\textwidth][c]{\input{tables/firststage-reg-fte.tex}}
-    \label{tab:firststage-reg}
-    \begin{flushleft}
-        \footnotesize
-        \textbf{Note}: Standard errors are clustered at the state-year level.
-    \end{flushleft}
-\end{table}
-
-\autoref{tab:firststage-reg} presents results of the first-stage regression, separately with and without a control for per student tuition revenue and year plus institution fixed effects.\footnote{
-    Representations for frequentist significance levels (i.e. 10, 5, 1\% etc.) are omitted here, and in all following tables.
-}
-We note columns (1) and (2) estimate that a shock to state appropriations per student in the state of 10\% is associated with around 9.8\% change in state funding per student at the university; the instrument is strong, and we note similarity in estimates with and without inclusion of fixed effects.
-Columns (3) and (4) include the tuition revenue control (explained in \autoref{foot:control}) to exhibit estimates with the inclusion of this possible collider or bad control.
-Column (3) shows similar estimates to columns (1), (2) thanks to inclusion of fixed effects, so that the fixed effects was effective in soaking variation in per-student tuition revenues at the institution-year level.
-Column (1) represents the estimates for \autoref{eqn:firststage} with fixed effects, omitting the tuition revenue control, and is the preferred form that I proceed with in the following.
-
-\begin{figure}[h!]
-    \centering
-    \singlespacing
-    \caption{Local Projection Estimates for First-Stage \autoref{eqn:firststage}.}
-    \includegraphics[width=0.6\textwidth]{figures/firststage-lp.png}
-    \label{fig:firststage-lp}
-\end{figure}
-
-\autoref{fig:firststage-lp} presents local projection estimates for the staying-power of the state appropriation shock to university revenues.
-We see the time 0 estimate coincides with that of Column (1) in \autoref{tab:firststage-reg}, the contemporaneous effect of the shock revenues.
-We see a decaying, yet real and positive, effect of the shock on revenues for the next 10 years, illustrating that the state appropriation shock is a strong, if fading, instrument for state funding per student and justifying its use in local projection IV estimation.
-
-These results, together with the case for exogeneity, show the appropriations shock instrument strongly predicts university revenues in the first-stage estimation.
-
-\subsection{Instrumental Variables Model, University-Level}
-\label{sec:iv-model-uni}
-
-The primary empirical model combines the instrumental variable for state appropriation shocks with the empirical model for the effects of state funding at the university-level --- i.e.
-parameter $\beta$ in the following.\footnote{
-    It is important to note the treatment effect isolated here; the instrumental variables approach identifies the local average treatment effects (LATE) specific to the instrument.
-    So we interpret this treatment effect as a university's response in employment count and average salaries to state funding changes, changes specific to state appropriation shocks, among the complier group --- i.e.
-universities with any exposure to shocks, assuming no universities increase employment or salaries in response to negative revenue shocks.
-    \cite{mogstad2021causal} provide a full discussion of the LATE with instruments and multiple monotonicity conditions.
-}
-\begin{eqnarray}
-    \label{eqn:secondstage1}
-    X_{i,t} &=& \eta_i + \zeta_t + \delta Z_{i,t} + \epsilon_{i,t} \\
-    \label{eqn:secondstage2}
-    Y_{i,t} &=& \alpha_i + \gamma_t + \beta \widehat X_{i,t} + \varepsilon_{i,t}
-\end{eqnarray}
-I estimate the system by two stage least squares, including institution and year fixed effects, and investigate outcomes at the university-level to analyse effects on the university as a result of changes in revenues.
-Additionally, I estimate the model via local projections \citep{jorda2005,miller2022} to investigate whether the effects of revenues on faculty composition linger for multiple years after the original appropriation shock.
-Regarding outcomes, I focus on the composition of the professors employed at the university by analysing (log) count per student, and average salaries paid to, professors employed by the university.
-
-\begin{figure}[H]
-    \centering
-    \singlespacing
-    \caption{Instrument Variables Model for University Finances and Faculty Composition.}
-    \begin{tikzpicture}
-        \node[state] (instrument) at (0,0) {$Z$};
-        \node at (-2.6,0.5){State};
-        \node at (-1.9,0){appropriation};
-        \node at (-2.6,-0.5){shock};
-        \node[state] (endogenous) [right=of instrument] {$X$};
-        \node at (2,1.5) {University};
-        \node at (2,1) {revenues};
-        \node[state] (outcome) [right=of endogenous] {$Y$};
-        \node at (5.2,0.25) {Faculty};
-        \node at (5.6,-0.25) {Composition};
-        % Relevance condition
-        \path (instrument) edge (endogenous);
-        \path (endogenous) edge (outcome);
-        % Exclusion Restriction
-        \path (instrument) edge[bend right=60] (outcome);
-        \node[color=red] at (2,-1.2) {X};
-    \end{tikzpicture}
-    \label{fig:SCM-ivmodel}
-\end{figure}
-
-\autoref{fig:SCM-ivmodel} presents the system in graphical form, for the causal effect of state appropriation shocks on university revenues and thus faculty composition, making clear the assumption that appropriation shocks affect faculty composition only via affecting university finances.
-
-\subsection{Instrumental Variables Model, Individual Professor-Level}
-\label{sec:iv-model-indiv}
-
-This analysis additionally uses data regarding individual professors in the Illinois university system, to investigate the effects of changes in university revenues on the individual professors at the universities.
-Redefining the level of outcome requires adjustment to the empirical approach, leveraging variation in university finances for the years after a professor joins the university.\footnote{
-    This formulation follows that presented by \cite{NBERw27885}, where individual student outcomes are analysed via variation in state appropriations after their freshman-year.
-    This contrasts with \autoref{sec:iv-model-uni} and \cite{NBERw23736}, where the unit of analysis is the university-year.
-}
-
-\autoref{eqn:rolling-instrument} defines a rolling-share variant of the instrument, $\tilde Z_{j,t}$, where the university's state appropriations share exposure is based in the year a professor joins the university --- and not the base period 1990--1993.
-$j$ indexes each professor in year $t$, $\tau(j)$ for the year the professor first joins their institution.\footnote{
-    Identifying $\tau(j)$ is possible for $j$ by restricting to all professors hired 2011-2021 --- i.e. in the years after the start of the full panel.
-    It is not possible to discern the hiring year for professors who  were hired in the years preceding 2011, and so the entire sample is only possible to analyse using the base-share in years 1990-1993 formulation (e.g.
-\autoref{tab:facultysalaries-shock-illinois}, \ref{tab:promotion-shock-illinois}).
-}
-\begin{align}
-    \label{eqn:rolling-instrument}
-    \tilde Z_{j,t} &\coloneqq \log \left[
-    \left( \frac{\text{Total State Funding}_{s(j),t}}{\text{Student Population}_{s(j),t}} \right)
-    \left( \frac{\text{State Funding}_{\tau(j)}}{\text{Total Revenues}_{i,\tau(j)}} \right) \right]
-\end{align}
-
-This approach leverages an insight, made available by level of the data: that an individual professor is affected by changes in university revenues after they have joined the university.\footnote{
-    Notice that \autoref{sec:iv-model-uni} considers the number of professors employed by the university; whether a professor becomes employed at the university is likely affected by that university's finances.
-    The formulation in this section does not consider whether the professor joins the university, instead taking as given that the professor is employed at the university, and then projects the effect on the individual.
-}
-Exogeneity and relevance of the rolling-share instrument, $\tilde Z_{j,t}$, follows the same reasoning as that for the base-share instrument, $Z_{i,t}$, discussed in \autoref{sec:approp-shocks}.\footnote{
-    The base-share instrument is still used for some outcomes with the individual Illinois professors, where appropriate (\autoref{tab:facultysalaries-shock-illinois}, \ref{tab:promotion-shock-illinois}).
-}
-We satisfy the assumptions for exogeneity by noting that none of the Illinois public campuses take the majority of state appropriations, and that the instrument identification strategy relies on exogeneity in changes in state appropriations to individual professor-outcomes, following the year they joined the university.\footnote{
-    Additionally, within-institution changes resulting from share reliance on state funding may be correlated with unobserved changes in the outcomes, so that \cite{NBERw27885} note the importance of controlling for the base share and state student population.
-    The formulation here implicitly controls for these factors via the fixed effects; results are relatively similar while including these controls without including fixed effects, and so are omitted.
-}
-\autoref{tab:firststage-illinois} presents results of the first stage estimation, showing that the instrument is strong in the same way as that for the university-level outcomes (\autoref{tab:firststage-reg}), with very similar estimates for the association between appropriation shocks and non-institutional revenues.
-
-The instrumental variables model is then defined as follows where $i(j)$ refers to the institution that professor $j$ is employed at, and $Y_{j,t}$ refers to individual-level outcomes total salary, rate of promotion, and propensity to leave the Illinois public university system.
-The system includes a fixed effect for the institution and first year of employment.\footnote{
-    The instrument varies by institution, based in the year of first employment, so that these are the corresponding fixed effects and levels of clustered standard errors.
-}
-\begin{eqnarray}
-    \label{eqn:secondstage1_indiv}
-    X_{i(j),t} &=& \theta_{i(j)} + \phi_{\tau(j)} + \delta \tilde Z_{i(j),t} + \epsilon_{i(j),t} \\
-    \label{eqn:secondstage2_indiv}
-    Y_{j,t} &=& \mu_{i(j)} + \nu_{\tau(j)} + \beta \widehat X_{i(j),t} + \varepsilon_{i(j),t}
-\end{eqnarray}
-We then interpret parameter $\beta$ in this formulation as the effect of changes in state funding at an Illinois public university, via state appropriation shocks, on an individual professor.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Results section
-\section{Results}
-\label{sec:results}
-
-\subsection{University-Level}
-
-\autoref{tab:facultycount-shock-reg} presents OLS and IV estimates, where the outcome is (log) count of professors employed at the university, separated by position, and for all professors (Columns 7, 8).
-For all professors we see a positive correlation between university revenues and professor count per student (the OLS columns), where an increase in university total revenues by 10\% is associated with an increase in professor count per student of 0.83\%, yet negatively correlated with lecturer count of -1.3\%.
-Lecturers per student is not shown to be correlated with total university revenues.
-The 2SLS columns show results after identifying state funding with the appropriation shock.
-All professor count per student increases by 0.53\% in response to a 10\% rise in total revenues.
-The effect is driven by increases in the count of assistant and full professors (i.e.
-tenure-track and tenured) per student, where assistant professor count per student increase by 1.4\%, and full professor 1.2\%.
-The count of lecturers per student decreases by 4.4\% in response to a (positive) 10\% shock in university revenues.
-
-\begin{table}[h!]
-    \singlespacing
-    \centering
-    \caption{OLS and 2SLS Estimates for University Faculty Composition.}
-    \makebox[\textwidth][c]{\input{tables/facultycount-shock-reg-fte.tex}}
-    \begin{flushleft}
-        \footnotesize
-        \textbf{Note}: Standard errors are clustered at the state-year level.
-    \end{flushleft}
-    \label{tab:facultycount-shock-reg}
-\end{table}
-
-\begin{figure}[h!]
-    \centering
-    \singlespacing
-    \caption{Local Projection Estimates for Professor Count per Student.}
-    \includegraphics[width=0.6\textwidth]{figures/all-count-lp.png}
-    \label{fig:all-count-lp}
-\end{figure}
-\autoref{fig:all-count-lp} shows local projection estimates, where the effect on total professor count per student persists at the same magnitude of around 0.1 four years after the identified shock.
-These findings are in line with \cite{turner2014impact} observing that universities froze hiring particularly for tenure-track positions in response to negative budget shocks in the late 2000s, and \cite{brown2014endowment} from shocks to university endowments.
-We see a negative coefficient for the number of lecturers per student.
-This effect lines up with two trends we see in \autoref{sec:trends}: that public universities revenues (per student) are decreasing while utilisation of non-tenure track lecturers increased over the same time period.
-And yet, identifying the university revenues by the state appropriations shock increases the magnitude of the effect (Column 2) with respect to the association (Column 1), strengthening the case that the two trends are causally linked.
-Between years 1990-1999 IPEDS provided count of professors by explicit tenure designation (opposed to named position as in \autoref{tab:facultycount-shock-reg}), and \autoref{tab:tenurecount-shock-reg-fte} presents results of the respective regressions by tenure status.
-Results are relatively similar, yet imprecise from the much smaller number of observations.
-
-Together these findings show that public universities increase their count of tenure-track and tenured professors per student in years where revenues are more plentiful, possibly by increasing hiring intensity or efforts to maintain incumbent professors.
-Yet, when revenues are not so bountiful, count of tenure-track and tenured professors per student decreases.
-The mechanism that explains these changes in faculty composition are, however, not clear at this unit of analysis; individual data are needed to delve deeper.
-
-\subsection{Individual Professor Outcomes}
-
-I first examine whether a shock to the university's revenues affects the salaries of professors at the university.
-The first outcome is first-year salary for each professor by their position;
-\autoref{tab:newhiresalaries-shock-illinois-rolling} presents 2SLS estimates for the elasticity of total salaries with respect to university revenues, among the sample with a year of starting employment after 2010.
-New hires of assistant, full, or administrative position professors are not realted to revenues, via a shock to appropriations in their first-year of employment.
-Yet, the starting salary of lecturers is 3.8\% higher given an increase in their hiring university's state funding of 10\%.
-
-\begin{table}[h!]
-    \singlespacing
-    \centering
-    \caption{2SLS Estimates for Faculty Salaries, in First-Year, at Illinois Universities.}
-    \makebox[\textwidth][c]{\input{tables/newhiresalaries-shock-illinois-rolling.tex}}
-    \begin{flushleft}
-        \footnotesize
-        \textbf{Note}: Standard errors are clustered at the institution and first year of employment level. 
-    \end{flushleft}
-    \label{tab:newhiresalaries-shock-illinois-rolling}
-\end{table}
-
-Regarding total salary for every year of employment, \autoref{tab:facultysalaries-shock-illinois-rolling} presents estimates for the elasticity of total salary with respect to university revenues, for the entire sample of professors hired after 2010.\footnote{
-    Recall that this is a subsample of the entire data, that the rolling shift-share instrument is only defined for the sample with an observed hiring date in the years 2011-2021.
-}
-The elasticity is non-distinguishable from zero among any position, implying that long-term salaries are not related to state funding.
-\autoref{tab:facultysalaries-shock-illinois} presents the same model, for the sample of all professors and using the instrument based in years 1990-1993, similarly finding no relationship.
-
-\begin{table}[h!]
-    \singlespacing
-    \centering
-    \caption{2SLS Estimates for Faculty Salaries at Illinois Universities.}
-    \makebox[\textwidth][c]{\input{tables/facultysalaries-shock-illinois-rolling.tex}}
-    \begin{flushleft}
-        \footnotesize
-        \textbf{Note}: Standard errors are clustered at the institution-year level.
-    \end{flushleft}
-    \label{tab:facultysalaries-shock-illinois-rolling}
-\end{table}
-
-Note that professor position (assistant, associate, full) is also an outcome, thanks to the promotion channel: a university may be less able to promote their professors to higher paying positions when their revenues are constrained.
-\autoref{tab:promotion-shock-illinois-rolling} shows estimates where the outcome is rate of promotion within each professor position.
-The lecturer position describes the sample of professors who were ever listed as a non-tenure track lecturer and the binary for promotion equals one in a year they achieved an assistant professor position; assistant the same for the assistant to associate professor promotion; associate same for associate to full promotion.
-There is no discernible relationship between revenues and promotion among any position.\footnote{
-    \autoref{tab:promotion-shock-illinois} shows estimates using the base-year instrument among the entire sample, finding the same results.
-}
-Furthermore, local projection estimates (\autoref{fig:promoted-illinois-lp}) show that the there is no promotion effect, among all professors, for any of the following years either.
-Lastly, \autoref{tab:facultyleaving-shock-illinois-rolling}, \ref{tab:facultyleaving-shock-illinois} describes estimates for rate of exit from the employing university among Illinois professors, and finds no relationship with university revenues.
-
-\subsection{Discussion}
-
-The results show that the recent stagnation in state funding for higher education has affected the composition of faculty within public universities, away from tenure-track and tenured professors towards non-tenured lecturers.
-At the same time, there are little discernible effects on individual professors hired in the years 2011-2021 at Illinois public universities, except for the salaries of first year lecturers.
-The rate that faculty leave their university, and their rate of promotion between positions, are unaffected by the university revenues, so this leaves one primary channel to explain changes in faculty composition: hiring.
-
-\cite{turner2014impact} documents the wide-spread practice of hiring freezes at universities in response to budget shocks around the 2008 recession.
-Throughout the last decade, multiple such measures were taken by Illinois public universities in response to their deteriorating finances \citep{furlough2010}.
-The University of Illinois\footnote{
-    The University of Illinois includes three public university campuses: Urbana-Champaign, Chicago, Springfield.
-}
-did not receive the allocated state appropriations from the state of Illinois on time, so enacted cost-cutting measures to stay fiscally solvent.
-The university system placed a hold on all hiring for filling state-funded positions and promotions.
-Faculty were furloughed (placed on leave without pay) for a day each month, and university administrators placed on two days per month furlough, or eligible university employees could accept a voluntary, equivalent pay reduction.
-The University of Western Illinois adopted very similar measures in response to the Illinois budget crisis in 2016-2017 \citep{wiu2016}.
-
-\begin{figure}[h!]
-    \centering
-    \singlespacing
-    \caption{Trends in New Hires at Illinois Public Universities 2011-2021.}
-    \begin{subfigure}[b]{0.495\textwidth}
-        \centering
-        \caption{New Hire Count, Total.}
-        \includegraphics[width=\textwidth]{figures/newhire-count-illinois.png}
-        \label{fig:newhire-count-illinois}
-    \end{subfigure}
-    \begin{subfigure}[b]{0.495\textwidth}
-        \centering
-        \caption{Mean Salary in First Year for New Hires, \$ 2021 CPI-U.}
-        \includegraphics[width=\textwidth]{figures/newhire-salary-illinois.png}
-        \label{fig:newhire-salary-illinois}
-    \end{subfigure}
-    \label{fig:newhire-illinois}
-\end{figure}
-
-These examples show that Illinois public universities were effected by the negative shock to state funding for higher education, and instituted policy changes aimed at affecting faculty outcomes investigated in \autoref{sec:results}.
-Yet the only measurable effect was on count of professor per student (implied via slower hiring), and first-year lecturer salaries.
-To investigate these trends for the entire state, \autoref{fig:newhire-illinois} shows the count of professors hired in the years 2011-2021, and their mean salary, by position.
-In particular, no position had sustained rises in hiring while enrolment within the university was rising (so that new hires per student necessarily fell), and hiring of new administrators jumped once the Illinois budget crisis ended in 2017.
-At the same time, professor new hires' starting salaries remained relatively constant for lecturer and assistant professors, yet rose and fell over the decade for full professors and administrators.
-
-%\begin{figure}[h!]
-%    \centering
-%    \singlespacing
-%    \caption{Faculty Profile at Illinois Public Universities 2011-2021.}
-%    \begin{subfigure}[b]{0.495\textwidth}
-%        \centering
-%        \caption{Faculty Count per Enrolled Student, Total for the State.}
-%        \includegraphics[width=\textwidth]{figures/faculty-count-illinois.png}
-%        \label{fig:faculty-count-illinois}
-%    \end{subfigure}
-%    \begin{subfigure}[b]{0.495\textwidth}
-%        \centering
-%        \caption{Mean Salary for All Faculty, \$ 2021 CPI-U.}
-%        \includegraphics[width=\textwidth]{figures/faculty-salary-illinois.png}
-%        \label{fig:faculty-salary-illinois}
-%    \end{subfigure}
-%    \label{fig:faculty-illinois}
-%\end{figure}
-%
-
-%\subsubsection{Threats to Identification}
-%
-%\cite{blandhol2022tsls} provide reason to doubt the LATE interpretation of two-stage least squares estimates with multiple layers of fixed effects (i.e. both individual and time fixed effects).
-%Needs comment, and to what extent should this question the empirical strategy.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Conclusion section
-\section{Summary and Concluding Remarks}
-\label{sec:conclusion}
-
-This analysis investigates how the recent stagnation in state support for higher education has affected public universities, and their faculty composition.
-The work contributes to the literature along two primary dimensions.
-Firstly, by exploiting the same empirical approach as \cite{NBERw23736,NBERw27885} to isolate the causal effect of changes in state funding on public universities themselves, the analysis provides a plausible mechanism for the observed negative effects on students outcomes that this stagnation has brought.
-Secondly, this analysis considers individual professors as the unit of analysis using a dataset (IBHED) that has not been considered in the economic literature.
-These data allow for detailed analysis of thousand's of professors' salaries, and are a starting point for further analysis of the determinants of professors wages in the Illinois public university system.
-
-Findings show that public universities have substituted away from tenure-track and tenured professors, towards non-tenured lecturers, in the face of persistent declines in state funding for higher education.
-The effects on incumbent professors in the state of Illinois are non-distinguishable from zero, which implies that the effect is driven by changes in university hiring of new professors.
-Public universities are hiring thousands of non-tenured lecturers to satisfy teaching obligations, and fewer tenure-track and tenured professors that their private university counterparts.
-
-While costs education have been rising in the US, thanks to multiple overlapping trends \citep{ehrenberg2012}, public universities have also dealt with declining state support.
-It is natural to expect that such headwinds will lead to systematic change at public universities, changes which are not neutral to the composition of faculty, or to their goals of research and education.
-Furthermore, private universities were not exposed to financial headwinds of the same magnitude or persistence.
-If public universities' hiring of faculty is disrupted by a fall in state funding, then we can expect that private universities will benefit (in a competitive sense) by being able to hire better professors from relatively resource constrained public universities.
-While public universities continue to educate the majority of higher education students in the US, we should worry about the effects that restricting their financial support has on the composition of faculty, and  higher education itself, in this country.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Bibliography section
+% Introduction section
+\input{sections/01-introduction.tex}
+% Data section
+\input{sections/02-data.tex}
+% Conceptual Framework
+\input{sections/03-concepts.tex}
+% Empirical methods
+\input{sections/04-empirical.tex}
+% Empirical results
+\input{sections/05-results.tex}
+% Dicussion
+\input{sections/06-discussion.tex}
+% Conclusion
+\input{sections/07-conclusion.tex}
+% Bibliography
 \bibliographystyle{agsm}
-\bibliography{bibliography}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Appendix section
-% Set-up the section.
-\appendix
-\setcounter{table}{0}
-\renewcommand{\thetable}{A\arabic{table}}
-\setcounter{figure}{0}
-\renewcommand{\thefigure}{A\arabic{figure}}
-
-% Start appendix
-\section{Appendix}
-\label{appendix}
-A number of statistical packages, for the $R$ language \citep{R2022}, made the empirical analysis for this paper possible.
-\begin{itemize}
-    \item \textit{Tidyverse} \citep{tidyverse} collected tools for data analysis in the R language.
-    \item \textit{LFE} \citep{lfe} implemented linear fixed effect models, with instruments, crucial for the empirical estimation in \autoref{sec:empirics}.
-    \item \textit{Stargazer} \citep{stargazer} provided a method to efficiently convert empirical results into presentable output in \LaTeX.
-    \item \textit{Lpirfs} \citep{lpirfs2019} implemented estimation of the \cite{jorda2005} local projections methods, with instrumental variables, crucial to the local projections estimates presented in this project.
-\end{itemize}
-
-\subsection{First Stage Estimates, Individual Outcomes}
-\label{appendix:part1}
-
-\begin{table}[h!]
-    \singlespacing
-    \centering
-    \caption{IBHED Summary Statistics, Entire Professor Panel 2010--2021.}
-    \makebox[\textwidth][c]{\input{tables/illinois-summary-rolling.tex}}
-    \label{tab:illinois-summary-rolling}
-    \begin{flushleft}
-        \footnotesize
-        \textbf{Note}: This table presents the summary statistics for professors with an observed year of first-year at their university.
-    \end{flushleft}
-\end{table}
-
-\begin{table}[H]
-    \singlespacing
-    \centering
-    \caption{First Stage Estimates, for Total University Revenues at the Individual-Level.}
-    \makebox[\textwidth][c]{\input{tables/firststage-illinois-rolling.tex}}
-    \label{tab:firststage-illinois}
-    \begin{flushleft}
-        \footnotesize
-        \textbf{Note}: Standard errors are clustered at the university-year level.
-    \end{flushleft}
-\end{table}
-
-\begin{figure}[h!]
-    \centering
-    \singlespacing
-    \caption{Local Projection Estimates for First-Stage \autoref{eqn:secondstage1_indiv}.}
-    \includegraphics[width=0.6\textwidth]{figures/firststage-illinois-lp-rolling.png}
-    \label{fig:firststage-illinois-lp}
-\end{figure}
-
-\subsection{Additional Results, University-Level}
-
-\autoref{tab:facultysalaries-shock-reg} presents OLS and IV estimates where the outcome is (log) mean salary for professors employed at the university, separated by position.
-Similarly, \autoref{fig:all-salaries-lp} shows the staying-power via projection methods.
-The outcome, mean salary is crude.
-
-\begin{table}[H]
-    \singlespacing
-    \centering
-    \caption{OLS and 2SLS Estimates for University Faculty-Tenure Composition.}
-    \makebox[\textwidth][c]{\input{tables/tenurecount-shock-reg-fte.tex}}
-    \begin{flushleft}
-        \footnotesize
-        \textbf{Note}: Standard errors are clustered at the university-year level.
-    \end{flushleft}
-    \label{tab:tenurecount-shock-reg-fte}
-\end{table}
-
-\begin{table}[H]
-    \singlespacing
-    \centering
-    \caption{OLS and 2SLS Estimates for University Faculty Salaries.}
-    \makebox[\textwidth][c]{\input{tables/facultysalaries-shock-reg-fte.tex}}
-    \begin{flushleft}
-        \footnotesize
-        \textbf{Note}: Standard errors are clustered at the state-year level.
-    \end{flushleft}
-    \label{tab:facultysalaries-shock-reg}
-\end{table}
-
-\begin{figure}[H]
-    \centering
-    \singlespacing
-    \caption{Local Projection Estimates for Professor Salary, Mean within Each University.}
-    \includegraphics[width=0.6\textwidth]{figures/all-salaries-lp.png}
-    \label{fig:all-salaries-lp}
-\end{figure}
-
-\subsection{Additional Results, Individual-Professor Level}
-
-\begin{table}[H]
-    \singlespacing
-    \centering
-    \caption{2SLS Estimates for Faculty Salaries at Illinois Universities.}
-    \makebox[\textwidth][c]{\input{tables/facultysalaries-shock-illinois.tex}}
-    \begin{flushleft}
-        \footnotesize
-        \textbf{Note}: Standard errors are clustered at the institution-year level.
-    \end{flushleft}
-    \label{tab:facultysalaries-shock-illinois}
-\end{table}
-
-\begin{table}[H]
-    \singlespacing
-    \centering
-    \caption{2SLS Estimates for Faculty Promotion Rate at Illinois Universities, using Rolling Instrument.}
-    \makebox[\textwidth][c]{\input{tables/promotion-shock-illinois-rolling.tex}}
-    \begin{flushleft}
-        \footnotesize
-        \textbf{Note}: Standard errors are clustered at the institution and first year of employment level.
-    \end{flushleft}
-    \label{tab:promotion-shock-illinois-rolling}
-\end{table}
-
-\begin{table}[H]
-    \singlespacing
-    \centering
-    \caption{2SLS Estimates for Faculty Promotion Rate at Illinois Universities, using Base-Year Instrument.}
-    \makebox[\textwidth][c]{\input{tables/promotion-shock-illinois.tex}}
-    \begin{flushleft}
-        \footnotesize
-        \textbf{Note}: Standard errors are clustered at the institution and first year of employment level.
-    \end{flushleft}
-    \label{tab:promotion-shock-illinois}
-\end{table}
-
-\begin{figure}[H]
-    \centering
-    \singlespacing
-    \caption{Local Projection Estimates for Promotion Rate at Illinois Universities.}
-    \includegraphics[width=0.6\textwidth]{figures/promoted-illinois-lp-rolling.png}
-    \label{fig:promoted-illinois-lp}
-\end{figure}
-
-\begin{table}[H]
-    \singlespacing
-    \centering
-    \caption{2SLS Estimates for Faculty Exit Rate at Illinois Universities, using Rolling Instrument.}
-    \makebox[\textwidth][c]{\input{tables/facultyleaving-shock-illinois-rolling.tex}}
-    \begin{flushleft}
-        \footnotesize
-        \textbf{Note}: Standard errors are clustered at the institution and first year of employment level.
-    \end{flushleft}
-    \label{tab:facultyleaving-shock-illinois-rolling}
-\end{table}
-
-\begin{table}[H]
-    \singlespacing
-    \centering
-    \caption{2SLS Estimates for Faculty Exit Rate at Illinois Universities, using Base-Year Instrument.}
-    \makebox[\textwidth][c]{\input{tables/facultyleaving-shock-illinois.tex}}
-    \begin{flushleft}
-        \footnotesize
-        \textbf{Note}: Standard errors are clustered at the institution and first year of employment level.
-    \end{flushleft}
-    \label{tab:facultyleaving-shock-illinois}
-\end{table}
-
+\bibliography{sections/08-bibliography.bib}
+% Appendix
+\input{sections/09-appendix.tex}
 \end{document}

### 2022-11-20 programs/analysis/individual-shock-rolling.R

@@ -1,13 +1,15 @@
 #!/usr/bin/R
 ## Senan Hogan-Hennessy, 3 August 2022
-print(c(Sys.time(), Sys.Date()))
 ## IV for Prof's outcomes, using Illinois data.
 ## Using a rolling-share variant of the instrument.
-library(tidyverse) ## functions for data manipulation and visualization
-library(lfe)
+print(c(Sys.time(), Sys.Date()))
+library(tidyverse) # Functions for data manipulation and visualization
+library(lfe) # Functions for fast linear models with IV + FEs
+library(plm) # Functions for panel data
+library(car) # Function for F stat regarding IV models
 # My custom flavour of Stargazer TeX tables:
 # devtools::install_github("shoganhennessy/stargazer")
-library(stargazer) ## TeX tables
+library(stargazer) # TeX tables
 set.seed(47)
 # This file follows an adjusted Deming Walters (2017, p.10) approach to
 # appropriations shock instrument.
@@ -43,7 +45,7 @@ ipeds.data <- read_csv("../../data/urban-ipeds/urban-clean-publicunis.csv") %>%
         enrollment_reported,
         enrollment_fte,
         # Appropriation shocks Per student
-        appropriationshock_perEnroll_real,
+        appropriationshock_perEnroll_rolling,
         appropriationshock_perFTE_real,
         appropriationshock_peruni_real,
         # State-level values
@@ -160,7 +162,6 @@ reg.data <- reg.data %>%
     # Only unis + years with measured state appropriations & shocks & Prof count
     filter(!is.na(salary_real), salary_real > 0,
         !is.na(enrollment_reported), enrollment_reported > 0,
-        #!is.na(nonauxrevenues_real),
         !is.na(stateappropriations_real),
         !is.na(appropriationshock_perEnroll_rolling))
 
@@ -195,7 +196,7 @@ reg.data %>%
         extra_salary_real = extra_salary_real) %>%
     as.data.frame() %>%
     stargazer(summary = TRUE,
-        summary.stat = c("mean", "sd", "median", "n"),
+        summary.stat = c("mean", "sd", "n"),
         digits = 0,
         digits.extra = 0,
         covariate.labels = c(
@@ -224,36 +225,43 @@ reg.data %>%
 # Explain Revenues with a shock to (only) state appropriations.
 firststage_approp.reg <- reg.data %>%
[…]

### 2022-11-20 programs/analysis/individual-shock.R

@@ -1,12 +1,15 @@
 #!/usr/bin/R
 ## Senan Hogan-Hennessy, 3 August 2022
-print(c(Sys.time(), Sys.Date()))
 ## IV for Prof's outcomes, using Illinois data.
-library(tidyverse) ## functions for data manipulation and visualization
-library(lfe)
+## Using a base-share version of the instrument.
+print(c(Sys.time(), Sys.Date()))
+library(tidyverse) # Functions for data manipulation and visualization
+library(lfe) # Functions for fast linear models with IV + FEs
+library(plm) # Functions for panel data
+library(car) # Function for F stat regarding IV models
 # My custom flavour of Stargazer TeX tables:
 # devtools::install_github("shoganhennessy/stargazer")
-library(stargazer) ## TeX tables
+library(stargazer) # TeX tables
 set.seed(47)
 # This file follows an adjusted Deming Walters (2017, p.10) approach to
 # appropriations shock instrument.
@@ -130,7 +133,6 @@ reg.data <- reg.data %>%
     # Only unis + years with measured state appropriations & shocks & Prof count
     filter(!is.na(salary_real), salary_real > 0,
         !is.na(enrollment_reported), enrollment_reported > 0,
-        !is.na(nonauxrevenues_real),
         !is.na(stateappropriations_real),
         !is.na(appropriationshock_perEnroll_real))
 
@@ -165,7 +167,7 @@ reg.data %>%
         extra_salary_real = extra_salary_real) %>%
     as.data.frame() %>%
     stargazer(summary = TRUE,
-        summary.stat = c("mean", "sd", "median", "n"),
+        summary.stat = c("mean", "sd", "n"),
         digits = 0,
         digits.extra = 0,
         covariate.labels = c(
@@ -194,14 +196,17 @@ reg.data %>%
 # Explain Revenues with a shock to (only) state appropriations.
 firststage_approp.reg <- reg.data %>%
     felm(log(stateappropriations_real / enrollment_reported) ~ 1 +
-        log(appropriationshock_perEnroll_real) |
+        I(-log(appropriationshock_perEnroll_real)) |
         unitid + firstyear |
         0 |
         unitid + year,
         data = .)
 # Get the F.Stat
 firststage_approp.fstat <-
-    summary(firststage_approp.reg)$P.fstat["F"] %>%
+    linearHypothesis(firststage_approp.reg, test = "F",
[…]

### 2022-11-20 programs/analysis/ipeds-public-private.R

@@ -184,35 +184,35 @@ prof_perfte.data <- ipeds.data %>%
             sum(enrollment_reported, na.rm = TRUE)) %>%
     ungroup()
 
-
-# Draw the graph for all Professors per student
-full_fte_perprof.plot <- prof_perfte.data %>%
-    ggplot(aes(x = year, y = full_fte_perprof, colour = factor(public))) +
+# Draw the graph for lecturers Professors per student
+lecturer_fte_perprof.plot <- prof_perfte.data %>%
+    ggplot(aes(x = year, y = 1 / lecturer_fte_perprof, colour = factor(public))) +
     geom_point() +
     geom_line() +
     scale_x_continuous(name = "Year",
         breaks = seq(1985, 2020, by = 5)) +
     scale_y_continuous(name = "",
-        limits = c(0.015, 0.03),
-        breaks = seq(0, 0.05, by = 0.005),
+        #limits = c(0, 350),
+        breaks = seq(0, 350, by = 25),
         labels = scales::comma) +
     theme_bw() +
     theme(plot.title = element_text(hjust = 0.5),
         legend.position = "top") +
     scale_colour_discrete(name = "", labels = c("Private", "Public"))
-ggsave("../../text/figures/full-fte-perprof.png",
-    plot = full_fte_perprof.plot,
+ggsave("../../text/figures/lecturer-fte-perprof.png",
+    plot = lecturer_fte_perprof.plot,
     units = "cm", width = fig.width, height = fig.height)
+
 # Draw the graph for TT Professors per student
 assistant_fte_perprof.plot <- prof_perfte.data %>%
-    ggplot(aes(x = year, y = assistant_fte_perprof, colour = factor(public))) +
+    ggplot(aes(x = year, y = 1 / assistant_fte_perprof, colour = factor(public))) +
     geom_point() +
     geom_line() +
     scale_x_continuous(name = "Year",
         breaks = seq(1985, 2020, by = 5)) +
     scale_y_continuous(name = "",
-        limits = c(0.005, 0.015),
-        breaks = seq(0, 0.05, by = 0.0025),
+        #limits = c(0, 350),
+        breaks = seq(0, 350, by = 5),
         labels = scales::comma) +
     theme_bw() +
     theme(plot.title = element_text(hjust = 0.5),
@@ -221,35 +221,36 @@ assistant_fte_perprof.plot <- prof_perfte.data %>%
 ggsave("../../text/figures/assistant-fte-perprof.png",
[…]

### 2022-11-20 programs/analysis/ipeds-shock.R

@@ -1,10 +1,11 @@
 #!/usr/bin/R
-## Senan Hogan-Hennessy, 3 August June 2022
+## Senan Hogan-Hennessy, 3 August 2022
 ## IV for Prof's salaries, using IPEDS data.
-print(c(Sys.time(), Sys.Date()))
+print(Sys.time())
 library(tidyverse) # Functions for data manipulation and visualization
 library(lfe) # Functions for fast linear models with IV + FEs
 library(plm) # Functions for panel data
+library(car) # Function for F stat regarding IV models
 # My custom flavour of Stargazer TeX tables:
 # devtools::install_github("shoganhennessy/stargazer")
 library(stargazer) # TeX tables
@@ -66,7 +67,14 @@ reg.data <- ipeds.data %>%
         lecturer_profmeansalary_real,
         all_profmeansalary_real,
         # Total paid on professor salaries
-        all_profoutlays_real)
+        all_profoutlays_real,
+        # Research and instruction expenditures.
+        instructionspending_total_real,
+        instructionspending_salaries_real,
+        researchspending_total_real,
+        researchspending_salaries_real,
+        nonauxspending_real)
+
 
 # Restrict data to unis + years with measured state appropriations & shocks
 reg.data <- reg.data %>%
@@ -88,7 +96,7 @@ reg.data %>%
         enrollment_reported = enrollment_reported,
         stateappropriations_real = stateappropriations_real / (10^6),
         totalrevenues_real = totalrevenues_real / (10^6),
-        #nonauxrevenues_real = nonauxrevenues_real / (10^6),
+        nonauxrevenues_real = nonauxrevenues_real / (10^6),
         lecturer_prof_count = lecturer_prof_count,
         assistant_prof_count = assistant_prof_count,
         full_prof_count = full_prof_count,
@@ -99,14 +107,14 @@ reg.data %>%
         all_profmeansalary_real = all_profmeansalary_real) %>%
     as.data.frame() %>%
     stargazer(summary = TRUE,
-        summary.stat = c("mean", "sd", "median", "n"),
+        summary.stat = c("mean", "sd", "n"),
         digits = 0,
         digits.extra = 0,
         covariate.labels = c(
             "Enrollment, institution reported",
             "State appropriations (millions 2021 USD)",
             "Total revenues (millions 2021 USD)",
[…]

### 2022-11-20 programs/data-build/illinois-build.R

@@ -36,7 +36,7 @@ for (extrayear in c(2011:2021)) {
 # CPI-U from FREDS (since Urban provided cpi has missing years)
 # Yearly average, seasonally adjusted, base year 1982-1984=100
 # https://fred.stlouisfed.org/series/CPIAUCSL
-cpiu.data <- read_csv("../../data/urban-ipeds/raw-data/freds-cpiu.csv")
+cpiu.data <- read_csv("../../data/inflation-indices/freds-cpiu.csv")
 # Clean CPI data, and put in terms of base year 2021
 base.year <- 2021
 base.index <- cpiu.data %>%

### 2022-11-20 programs/data-build/urban-ipeds-build.R

@@ -1,7 +1,7 @@
 #!/usr/bin/R
 ## Senan Hogan-Hennessy, 21 December 2021
 ## Build IPEDS data with the Urban Inst's package
-print(c(Sys.time(), Sys.Date()))
+print(Sys.time())
 ## https://educationdata.urban.org/documentation/colleges.html
 ## https://urbaninstitute.github.io/education-data-package-r/
 library(tidyverse) ## functions for data manipulation and visualization
@@ -79,16 +79,16 @@ finance.data <- finance.data %>%
         tuitionrev_real = replace(rev_tuition_fees_gross,
             rev_tuition_fees_gross <= 0, NA) / cpi2021,
         # Expenditures on instruction
-        instructionspending_total = replace(exp_instruc_total,
+        instructionspending_total_real = replace(exp_instruc_total,
             exp_instruc_total <= 0, NA) / cpi2021,
         # Salary Expenditures on instruction
-        instructionspending_salaries = replace(exp_instruc_salaries,
+        instructionspending_salaries_real = replace(exp_instruc_salaries,
             exp_instruc_salaries <= 0, NA) / cpi2021,
         # Expenditures on research
-        researchspending_total = replace(exp_research_total,
+        researchspending_total_real = replace(exp_research_total,
             exp_research_total <= 0, NA) / cpi2021,
         # Salary Expenditures on research
-        researchspending_salaries = replace(exp_research_salaries,
+        researchspending_salaries_real = replace(exp_research_salaries,
             exp_research_salaries <= 0, NA) / cpi2021,
         # All revenues (including tuition + appropriations)
         totalrevenues_real = replace(rev_total_current,
@@ -302,6 +302,10 @@ urban_ipeds.data <- urban_ipeds.data %>%
         totalrevenues_real,
         nonauxrevenues_real,
         nonauxspending_real,
+        instructionspending_total_real,
+        instructionspending_salaries_real,
+        researchspending_total_real,
+        researchspending_salaries_real,
         fedappropriations_real,
         stateappropriations_real,
         localappropriations_real,

### 2022-11-20 text/sections/00-abstract.tex
Updated draft.

\noindent
State support for higher education has stagnated on a per-student level, while public universities have reduced their employment of tenure-track or tenured professors, and increased their employment of contingent lecturers (per student).
This analysis relates these changes in the faculty composition at public universities to the fall in per-student university revenues from state appropriations.
A shift-share instrumental variables approach addresses endogeneity in decisions for state appropriations to higher education, by exploiting the yearly change in state support with an institution's reliance on that support.
A decrease in state funding of 10\%, via a shock to state appropriations, decreases the number of assistant professors per student at a public university by 1.4\% and full professors by 1.2\%, yet increases the number of lecturers per student by 4.4\%.
Local project estimates show that these effects linger for up to three years after the initial shock.
Analysis of all the professors hired at Illinois public universities 2011-2021 shows that incumbent professors are not affected by the changes in state support, implying that these changes in faculty composition arose by reduced hiring of tenure-track and tenured professors at public universities.
These results exhibit the long-term effects of stagnating state support for higher education, and raises questions about the direction that public education heads as these financial headwinds show no sign of dissipating.

\vspace{0.75cm}
\noindent\textbf{Keywords:}
State and Local Budget and Expenditures,
Higher Education,
Public Sector Labor Markets

\vspace{0.5cm}
\noindent\textbf{JEL Codes:} H72, I23, J45

### 2022-11-20 text/sections/01-introduction.tex
Updated draft.

\noindent
Public universities educate the majority of higher education students in the US, yet have experienced a secular decline in state funding (per student) the last three decades.
This decline has been shown to lead to worse education and later-life outcomes for students \citep{NBERw23736,NBERw27885}, yet it is not clear what mechanisms drive these effects, or how public universities are affected as institutions.
I show that four-year, degree-granting public universities have experienced falling state funding per student, and this affected their composition of faculty.
Falls in public university revenues from the state, instrumented by state-level finance shocks, lead to a fall in the number of tenure-track or tenured professors per student within a university, an increase in the number of lecturers, and overall fall in the number of faculty per student.
US private universities were not exposed to such financial constraints during this same time period, and do not exhibit the substitution away from tenure-track and tenured professors, so that the stagnating state support for public universities has implications for the wider structure of higher education instruction and research in the US.

US universities are widely considered the highest performing in the world, yet there are consequential differences between its universities that operate in the private sector and those established by state governments.
Public universities are subject to numerous state-level administrative laws, and rely on their state governments for funding: an average public university received around \$11,600 per enrolled student in 1990, yet only \$8,300 per enrolled student in 2021.
This fall is driven by stagnating funding provided by state governments, while enrolment among public universities rose by 46\% over the same time period.
At the same time, the number of professors per student at public universities stayed relatively stable at 0.045, yet fell from around 0.04 to around 0.35 professors per student at public universities, driven by a fall of over 20\% in the ratio of associate and full professors per student at public universities.

I use a variant of the \cite{NBERw23736,NBERw27885} shift-share instrument for state funding from shocks to appropriations for the entire state to identify the effect on the faculty composition.
State appropriations for public universities are plausibly endogenous to many outcomes: state governments decide on yearly budgets for their higher education sector, and this process can be influenced by local financial conditions, or even public perceptions of the state's higher education system.
The shift-share instrument interacts reliance on state funding in a base period with the yearly total appropriations per student in the entire state, to exploit both changes in support for higher education and how much each university relies on that support.

Findings show that negative changes in state funding affects faculty composition at public universities, away from tenured and tenure-track professors towards non-tenure track lecturer positions.
A reduction of 10\% in state funding, via a shock to state appropriations, leads to a fall in 1.4\% in the number of assistant professors per student within a university, 1.2\% for full (tenured) professors, and an increase in 4.4\% the number of lecturers per student.
Local projection methods show that these effects linger for a small number of years after the initial revenue shock.
Over this time period, state funding fell by around 35\%, while the count of professors per student fell by 9\%, so that these results show that falls in state funding explain around a third of the observed shift away from tenure-track and tenure professors towards contingent faculty.

Additionally, I use individual level data on professors at Illinois public universities over 2010-2021 to investigate whether the shocks to state funding affected incumbent professors at public universities.
Incumbent professors were not meaningfully affected, in terms of total salary, promotion rate, or rate of leaving the Illinois public university system, yet lecturers' first year salary decreased by 3.8\% in response to a fall in state funding of 10\%.
These findings imply that changes in faculty composition came about by changes in the composition of hiring new professors at the university level.

\cite{NBERw23736} first study the effect of (changes in) public universities' finances on student outcomes by isolating changes in public universities' expenditures on student instruction, addressing endogeneity of expenditures decisions by instrumenting expenditures with state appropriation shocks and limits on tuition price.
\cite{NBERw23736} find that increases in public university spending (via appropriation shocks) increases enrolment and degree completion among students, but corresponding changes in attendance price (via limits on tuition price) do not have an effect.\footnote{
    \cite{miller2022making} further analyse the effects of falls in university revenues, by finding that reductions in public universities tuition prices (suggestively motivated by falling state funding) leads to reductions in provided financial aid.
}
\cite{chakrabarti2018effect,NBERw27885} use the same instrument for public university revenues at the state level, combined with rich information on students' individual-level outcomes, to show that increases in state appropriations lead to degree lower completion time and later-life student debt.
\cite{bound2019public} use another variant of the same approach to show that state-level higher education spending cuts induced public universities to shift toward tuition as their primary source of revenue, and shift institutional resources towards students who pay higher tuition rates.
In a similar vein, \cite{bound2007cohort} document variation in per student state funding for higher education resulting from yearly changes in a state's student-age population; they find that lower per student funding lead to lower higher education completion rate among students.

Faculty are a core component of this country's educational and research-innovation university system, yet their composition within universities, and its relationship with higher education finances has so far not been studied.
\cite{brown2014endowment} presents the closest example by studying how university endowments react to negative financial shocks, exhibiting an association between negative endowment shocks and a fall in tenure-system faculty employed at private universities.
\cite{abe2015implications} present a model for decision-making by university administrators, and posit that a large component of variation in faculty salaries arises from university politics, and not only economic factors such as outside options.
\cite{johnson2009jep,NBERc13879} both empirically document allocation of faculty between research and instruction, and \cite{hemelt2021math} document differences in instructional cost (per student) between departments, primarily thanks to department differences in class size and faculty salaries.
\cite{turner2014impact} document how universities reacted to the Great Recession of 2008, a severe financial shock to state finances and support for higher education.

The number of professors, both in absolute terms and relative to number of students, may be particularly important for quality of instruction and thus educational outcomes.
\cite{angrist1999using} first causally show that reducing class size (via religious ruling) induces an increase in test scores for school-age children, yet the magnitude of this effect is not as clear in the higher education setting.
\cite{bandiera2010heterogeneous} use a fixed effect model, combined with multiple individual observations in a long panel, to find that UK university students perform worse academically in particularly large classes, and the difference is largest for students at the top of the academic distribution.
For the composition of faculty teaching university classes, \cite{bettinger2010does,figlio2015tenure} find US students have better enrolment and learning outcomes from courses taught by contingent or adjunct (i.e. not tenured or on tenure-track) professors (relative to tenured or tenure-track).
\cite{ehrenberg2005tenured} observe a negative association between utilisation of non-tenured instructors and student graduation rates.
My approach focuses on count of professors per student at universities, so does not directly observe individual class-sizes, yet presents considers this a possible mechanism through which stagnating state support for higher education negatively affects student outcomes \citep{NBERw23736,NBERw27885}.

### 2022-11-20 text/sections/02-data.tex
Updated draft.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Data section
\section{Data and Institutional Context}
\label{sec:data}

\subsection{Data Description}
The data used in this analysis come from two primary sources: \citet[IPEDS]{ipeds} regarding institutional information on finances and enrolment, and \citet[IBHED]{ibhed} for data on every professor in the Illinois public university system.

IPEDS is a survey of higher educational institutions in the US, and legally requires institutions to participate in order to receive Federal Title IV student aid.\footnote{
    This statement means that IPEDS does not necessarily cover the universe of US higher education institutions, yet in practice every public university and not for-profit four-year institution is represented.
}
Data are consistent between the years 1990 and 2021,\footnote{
    The years 1987-1989 are represented in these data in an incompletion fashion, so I focus on the years 1990 onwards.
    Year refers to the calendar year of the spring term --- i.e. 1987 refers to the academic year that ran August 1986 to July 1987.
}
and provide information on the total revenues a university received from every source (including state governments), enrolment number, plus faculty count and total expenditures on salaries.\footnote{
    I combine the Urban Institute's 2018 compilation of IPEDS data for the years 1990-2017, and manually combine with raw NCES data regarding the year 2018-2021 for all relevant variables.
    Figures for enrolment, faculty counts and salaries come from the raw NCES version of IPEDS for all years to address inconsistencies in the Urban Institute's data formulation for these variables.
}
I restrict analysis to public, four-year, degree-granting institutions, as these institutions adhere to a standardised concept of faculty profile, where tenure and title of appointment (lecturer, assistant professor, etc.) are relatively standardised.
For-profit institutions employ and enrol a negligible share of professors and students respectively, while students at two-year institutions by majority intend to eventually enrol at a four-year institution \citep{mountjoy2022},\footnote{
    Furthermore, \cite{mountjoy2022} documents the effects of expanding two-year higher education access in the US separately for students who would have and those who would not have otherwise have attended college.
    The high rate of eventual enrolment at four-year institutions among two-year students explains the negative impact of increased access to two-year institutions on students who are diverted away from four-year institutions.
}
so that these institutions are not considered.

IPEDS reports the count of professors employed by position, as well as total salary expenditures by position.
This gives a resulting panel data-set, where each row represents a university-year, and includes columns for university finances, plus total count and average salary\footnote{
    Real salary is computed by scaling nominal salary to 2021 dollars by the CPI-U.
} for professors by position (lecturer, assistant, tenured, total).
\autoref{tab:ipeds-summary} presents summary statistics for relevant variables in these data.

\begin{table}[h!]
    \singlespacing
    \centering
    \caption{IPEDS Summary Statistics, Public Universities Panel 1987--2021}
    \makebox[\textwidth][c]{\input{tables/ipeds-summary-fte.tex}}
    \label{tab:ipeds-summary}
\end{table}

IPEDS provides information at the university level, yet lacks information on the distribution of salaries and professor count within a university.\footnote{
    IPEDS provides total paid to salary and total faculty employed per university-year, so that I can investigate average professor salary per university-year with IPEDS data.
    Unfortunately, this measure of professors' salaries is particularly crude in measuring on professors' salaries and other individual-outcomes; summary statistics on IPEDS data do not agree with trends in average professor salary over the sample time period \citep{aau2021survey}.
    Similarly, IPEDS reports of the average salary paid to professors at private institutions also disagree with other sources, so are not considered in this analysis.
}
To investigate the distribution, I integrate individual-level data for every public university professor in the state of Illinois between the years 2010-2021.
IBHED hosts the information;
Public Act 96-0266 (effective 1 January 2010) requires that each university report base salary and benefits all administrators, faculty members, and instructors employed by the college or university.\footnote{
    The universities included are all campuses of the nine Illinois public universities: Chicago State University, Eastern Illinois University, Governors State University, Illinois State University, Northeastern Illinois University, Northern Illinois University, Southern Illinois University  (all five campuses), University of Illinois (all four campuses), Western Illinois University.
}
This publicly available data provides the basis to build a panel of Illinois public university professors 2010-2021; I define a professor as an individual by their first plus last name and university pairing, and link this database to IPEDS regarding finances for their employing institution.

The analysis sample represents 16,932 professors in the year 2010 and 15,352 in the year 2021, with summary statistics presented in \autoref{tab:illinois-summary}.
Additionally, further analysis using the first year of a professor's employment focuses on the subset of professors with observed year of hiring between the years 2011 and 2021, representing 1,778 professors in 2011, and 9,099 in 2021.\footnote{
    Summary statistics for this group, which over-represents lecturers compared to assistant and full professors, can be seen in \autoref{tab:illinois-summary-rolling}.
    The reasons for considering the sub-sample with identified year of hire in the range 2011-2021 are explained in \autoref{sec:iv-model-indiv}.
}

\begin{table}[h!]
    \singlespacing
    \centering
    \caption{IBHED Summary Statistics, Professor Panel 2010--2021.}
    \makebox[\textwidth][c]{\input{tables/illinois-summary.tex}}
    \label{tab:illinois-summary}
\end{table}

\subsection{Trends in Finances, Enrolment, and Faculty Composition}
\label{sec:trends}
A state government plans an annual budget a couple of years ahead of the fiscal year, and the legislature chooses to approve or reject a budget request put forth by the governor's office.\footnote{
    \cite{NBERw23736} present a full discussion of the decision-making process for state appropriations, drawing on administrative records originally analysed by \cite{parmley2009state}.
}
US state governments, by majority, are legally obligated to run a balanced-budget, so that yearly variation in tax revenues (caused by changing economic conditions or otherwise) necessarily affect state expenditures.
This process leads to yearly variation in state appropriations not seen in other revenues sources (such as federal government appropriations), since US states differ in their support for higher education, and are subject to fiscal constraints.
State appropriations to higher education are a particularly attractive area of state spending to absorb such shocks to state finances \citep{delaney2011state}.\footnote{
    \cite{delaney2011state} fully describes the financial environment of state expenditures, and what makes spending on higher education an attractive area for state governments to expand funding during years of higher tax revenues, and retract funding in leaner years.
    An analysis of state expenditures for the years 1980-2004 (overlapping with the sample for this analysis) provides solid evidence for these trends.
}
Additionally, there is yearly variation in the number of higher education students in each state \citep{turner2014impact}, so that per-student state appropriations vary on multiple dimensions.

\autoref{fig:funding} exhibits the trends for the mean public university for the years 1990-2021.
We see a rise in total revenues received by public universities (from all sources), and a notable increase in mean tuition revenues from \$48 million per year-university to \$150 million per year-university.
At the same time, total state appropriations stagnated at around \$100 million per year-university for 1990-2008, falling around 2008 and have not recovered ever since.
While public universities experienced a stagnation in state support, private universities were not exposed to the same constraints, receiving \$37,000 per student in 1990 and \$49,000 in 2021, experiencing no corresponding decline in any specific component.

\begin{figure}[h!]
    \centering
    \singlespacing
    \caption{Mean Total Revenues among Public Universities, by Year.}
    \begin{subfigure}[b]{0.495\textwidth}
        \centering
        \caption{Total, millions \$ 2021 CPI-U.}
        \includegraphics[width=\textwidth]{figures/mean-funding-total.png}
        \label{fig:mean-funding-total}
    \end{subfigure}
    \begin{subfigure}[b]{0.495\textwidth}
        \centering
        \caption{Per Enrolled Student, \$ 2021 CPI-U.}
        \includegraphics[width=\textwidth]{figures/mean-funding-fte.png}
        \label{fig:mean-funding-fte}
    \end{subfigure}
    \label{fig:funding}
\end{figure}

\begin{figure}[h!]
    \centering
    \singlespacing
    \caption{Total Student enrolment, by University Sector, and Year.}
    \begin{subfigure}[b]{0.495\textwidth}
        \centering
        \caption{Total.}
        \includegraphics[width=\textwidth]{figures/enrollment-total.png}
        \label{fig:enrollment-total}
    \end{subfigure}
    \begin{subfigure}[b]{0.495\textwidth}
        \centering
        \caption{Mean, University Level.}
        \includegraphics[width=\textwidth]{figures/enrollment-mean.png}
        \label{fig:enrollment-mean}
    \end{subfigure}
    \label{fig:enrolment}
\end{figure}

At the same time, student enrolment at public universities rose precipitously.
6.2 million students were enrolled in public universities in 1990, and this number rose by 47\% to 9.1 million, with most of the increase occurring over the years 2000-2021.
\autoref{fig:enrolment} shows that total enrolment at private universities has also risen over the same time period, but not as drastic in either relative or absolute terms; the mean private university grew from 9,800 students in 1990 to 11,800 in 2021.
This means that revenues per student have stagnated across all measures (seen in \autoref{fig:mean-funding-fte}), and particularly fallen from a mean of \$11,000 per student in 1990 to less than \$8,000 per student in 2021.

\begin{figure}[h!]
    \centering
    \singlespacing
    \caption{Student Enrollment per Professor, by University Sector, Professor Appointment, and Year.}
    \begin{subfigure}[b]{0.495\textwidth}
        \centering
        \caption{Lecturer}
        \includegraphics[width=\textwidth]{figures/lecturer-fte-perprof.png}
        \label{fig:lecturer-fte-perprof}
    \end{subfigure}
    \begin{subfigure}[b]{0.495\textwidth}
        \centering
        \caption{Assistant.}
        \includegraphics[width=\textwidth]{figures/assistant-fte-perprof.png}
        \label{fig:assistant-fte-perprof}
    \end{subfigure}
    \begin{subfigure}[b]{0.495\textwidth}
        \centering
        \caption{Full.}
        \includegraphics[width=\textwidth]{figures/full-fte-perprof.png}
        \label{fig:full-fte-perprof}
    \end{subfigure}
    \begin{subfigure}[b]{0.495\textwidth}
        \centering
        \caption{All.}
        \includegraphics[width=\textwidth]{figures/all-fte-perprof.png}
        \label{fig:all-fte-perprof}
    \end{subfigure}
    \label{fig:fte-perprof}
\end{figure}

\autoref{fig:fte-perprof} documents the divergence in faculty composition (per student) between the average private and public university 1990-2021.
Private universities start with  a higher baseline of around 4.5 professors per 100 students, and exhibited yearly variation of less than 0.5 professor per hundred students over the thirty years.
Public universities start with 3.9 professors per hundred students, and this number falls to around 3.5 primarily in the 2008-2011 time period.
We see a similar difference in baseline, and fall for the years 2008-2011 for assistant professors.
Private and public universities have similar numbers of associated and full professors before the year 2000, yet this number has fallen by over 20\% in the next 20 years only for public universities: in 2021 the mean public university has 6 fewer full professors per hundred students than the mean private university.
Over the same time period, we see the rise in use of non-tenure track instructor positions (referred to as lecturers from here on), who were employed at similar rates in both sectors in 1990 yet have been utilised by public universities at a higher rate since.

\subsection{Trends in Illinois}
\label{sec:trends-illinois}

The state of Illinois experienced the same stagnation in state support for higher education that the rest of the nation has experienced 1990-2021.
\autoref{fig:illinois-funding} shows the trends for revenues among the seven Illinois public university campuses, including the falling share of state appropriations and substitution towards tuition revenue.

\begin{figure}[h!]
    \centering
    \singlespacing
    \caption{Mean Revenues among Illinois Public Universities, by Year.}
    \begin{subfigure}[b]{0.495\textwidth}
        \centering
        \caption{Total, millions \$ 2021 CPI-U.}
        \includegraphics[width=\textwidth]{figures/illinois-funding-total.png}
        \label{fig:illinois-funding-total}
    \end{subfigure}
    \begin{subfigure}[b]{0.495\textwidth}
        \centering
        \caption{Per Enrolled Student, \$ 2021 CPI-U.}
        \includegraphics[width=\textwidth]{figures/illinois-funding-fte.png}
        \label{fig:illinois-funding-fte}
    \end{subfigure}
    \label{fig:illinois-funding}
\end{figure}

The decade of 2010-2021 has also seen yearly variation in its revenues, particularly around 2016.
In the calendar year 2015 partisan disagreements between the democratic legislature and republican governor led to the 2016 fiscal year starting with no state budget.
State agencies, and higher education institutions, employed accounting techniques to continue operating without any resources provided by the state government.\footnote{
    Fiscal year 2016 refers to June 205 to June 2016, as is the same for the academic year definition in this analysis. 
    K-12 schools were unaffected thanks to a separate budget agreement allowing non-higher education to operate as before.
}
While state institutions were able to stay open, there were drastic revenue and spending cuts in response to the budget impasse, as it continued through fiscal year 2017, and ended with a new budget restoring revenues to state agencies and universities for the 2018 fiscal year.
This variation in public university revenues via the state appropriations channel, stemming entirely from political disagreements --- not from state decisions regarding higher education and its finances \citep{young2020squandered} --- mean that Illinois public universities exhibit sizable changes in their state appropriations over 2010-2021, of similar order to those for the rest of the country over 1990-2021.

### 2022-11-20 text/sections/03-concepts.tex
Updated draft.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section on possible outcomes
\section{Conceptual Framework}
\label{sec:conceptual}
%
%\textbf{Follow Prof Riehl's comments.
%    Also consider a theory toy-model where reach and teaching are plasuible substitutes, following \cite[Sec.~6.2]{NBERc13879}.
%    Better motivated if research spending is affected.
%}
%The, revenue shock induced, shift away from tenured profs may be mechanically relate to the shift between departments (towards STEM).
%Find a way to mix the “non-replacement of profs” with that for “department composition shift.”
%
%\subsection{University Responses to Financial Shocks}
%\label{sec:responses}

While state support for public higher education has stagnated at the same time as education costs rose, there are multiple possible ways that a university can respond.
\cite{NBERw23736} established that corresponding rises in tuition did not offset the falling state support, so that university spending (per student) fell in response to these persistent, negative state funding shocks.
There are multiple ways that these changes in finances may affect the faculty composition at public universities.

Universities every year hire a number of new professors, either to expand their departments or to replace leaving professors.
The number of professors hired is usually highest among non-tenured adjunct faculty, as these instructors are by majority hired on short-term contingent contracts, so that they are less costly to hire (or to not reinstate) in response to yearly changes in teaching needs.
Secondly, tenure-track assistant professors are hired in most years at public universities to replace leaving assistant professors, and are most often hired with a four to six year contract and agreement to be considered for tenure at the end of this term.
Lastly, tenured professors are those that have undergone review and have successfully secured a full-time appointment at their university with no expiration date; tenured professors may be hired from outside a university, though this position has the lowest hiring rate among most universities.
It is common for universities to restrict their faculty hiring in the face of financial shocks.
Since the yearly hiring rate is highest among lecturers and assistant professors, financial shocks may lower the number of lecturer and assistant professors at public universities the most, if faculty hiring is affected by falls in state funding.

Additionally, professors' salaries may be disrupted by the stagnation in public university finances.
When the university has a lower budget for its yearly hiring, it may respond by lowering the salary they offer to new hires.
This possible effect may not be the same across each position of professor; tenured faculty are often hired away from another university, so that new tenured professor hires could be less likely to accept a lower offer from a public university, and we would see that new tenured hires are not affected salary-wise.\footnote{
    \cite{blackaby2005} describes the academic outside options among economist faculty, documenting how they differ for academic position and gender.
}
Yet salary for all the professors, not just new hires, may also be affected: multiple universities passed a university-wide pay-cut for their faculty in response to state budget cuts around the 2008 recession, for example.\footnote{
    Indeed, Cornell University implemented both hiring freezes and even a nominal salary reduction for professors, in anticipation of a financial shock in early 2021.
    The salary cut was not permanent, as the oncoming financial shock turned out to not be as serious as projected, so that the salary cut was returned to professors in the year 2021.
}
Faculty are not paid the same by position, so that their are financial consequences for faculty being promoted between positions within a university.
In response to tightening fiscal restraints, universities may be less likely to grant tenure to assistant professors, or to promote tenured professors from associate to full professor either.

It is not immediately clear which effect will dominate, and which position of professor will be effected the most by changes in their university's finances.
Yet, there is one empirical fact worth noting: a mean assistant or tenured professor earns more than double that of an adjunct lecturer per year (\autoref{tab:illinois-summary}), and the tenure contract means universities are compelled to pay that higher salary for more years than contingent lecturers.
If a public university's primary obligation is to teach, and they must fulfil this objective with fewer and fewer resources, then they may substitute away from tenure-track and tenured professor towards contingent lecturers to achieve this obligation.\footnote{
    That is substituted away from tenure-track and tenured professors according to the salary ratio, and the relative productivity between these two positions.
    It is not possible to estimate the relative productivity in this paper, so I leave the rate of substitution between the two positions to further research.
}

### 2022-11-20 text/sections/04-empirical.tex
Updated draft.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Empirical section
\section{Empirical Framework}
\label{sec:empirics}

A professor is employed by a university, so it stands to reason that when public universities experience a secular fall in one of their revenue sources revenue --- analogous to a demand shock in the private sector --- that the composition of their employees will be affected.

Na\"ively we can analyse the relationship between professor-outcomes $Y_{i,t}$ at university $i$ in year $t$ as a result of state funding per student $X_{i,t}$.
Outcomes include count of professors per student within each university-year.
This model gives $\beta$ the association with these outcomes and state funding per student,\footnote{
    Note that dividing by student count also implicitly controls for the size of the university, so that this model accounts for yearly variation in professor count and university revenues arising from growth in a university.
}
including fixed effects to control for effects specific to the university and year.
Log$(.)$ transforming the variables improves the interpretability of $\beta$ as an elasticity for professor count per student with respect to state funding per student.
\begin{equation}
    \label{eqn:naivereg}
    Y_{i,t} = \alpha_i + \gamma_t + \beta X_{i,t} + \epsilon_{i,t}
\end{equation}

Yet a university's finances are not exogenous to state decisions for support of higher education, or exogenous to internal institutional decisions.
Instead, the state government and university administration undertake a complex process of allotting resources across multiple different priorities, including instruction, research, or between departments.
Importantly for this analysis, revenues received from an institution's state government provide opportunity to address this endogeneity.

\subsection{State Appropriation Shocks}
\label{sec:approp-shocks}

\cite{NBERw23736,chakrabarti2018effect,NBERw27885} address endogeneity in public university finances by exploiting a shift-share instrument for changes in state-level funding interacted with university reliance on state funding in a base period.
\begin{align}
    \label{eqn:public-instrument}
    Z_{i,t} &\coloneqq - \log \left[
    \left( \frac{\text{Total State Funding}_{s(i),t}}{\text{Student Population}_{s(i),t}} \right)
    \sum_{\tau = 0}^{3} \frac 14
    \left( \frac{\text{State Funding}_{i,1990 + \tau}}{\text{Total Revenues}_{i,1990 + \tau}} \right) \right]
\end{align}
The system exploits the fact that institutions who rely on state appropriations more will be affected by state appropriation shocks.
$Z_{i,t}$ is the instrument for (log) state appropriations for institution $i$ in year $t$, interacting the average funding for universities in state $s(i)$ with reliance on state funding relative to total revenues, averaged across the base years 1990--1993.\footnote{
    1990--1993 are defined as data for public university finance data are most comparable (i.e. without many missing values) beginning in 1990.
    \cite{NBERw23736} use the single year 1990 as the base year, though I use the four years to ameliorate missing values in the single year of 1990.
    Results are similar in either specification.
}
$Z_{i,t}$ is contructed as negative to reflect the fact that the long term trend in, and most of the short-run shocks to, state funding for higher education has been negative.
State funding has been been falling, so that the instrument describes shocks to university revenues, mostly in a negative direction.

\cite{NBERw27885} notes the tendency for public universities to respond to state funding cuts by increasing reliance on tuition, where \cite{NBERw23736} specifically instruments for tuition revenues with collected information on legislative tuition price controls.\footnote{
    \label{foot:control}
    It may be argued that tuition revenues are confounder between the causal effect of changes in state funding on a university's total revenues, so that this analysis focuses on state support for higher education (no total revenue) as a result.
    This follows the \cite{NBERw27885} formulation.
    On the other hand, rises in tuition revenues (per student) may arise as result of tuition hikes thanks declining state support, which would mean controlling for tuition would constitute a bad control.
    Estimates including tuition revenues (per student) as a control in the second stage of the IV estimates produces results of very similar magnitude and direction, and so are omitted.
}
This analysis focuses on one important source of revenues for public universities: state appropriations.

The first-stage is then as follows, including institution and year fixed effects.
\begin{equation}
    \label{eqn:firststage}
    X_{i,t} = \eta_i + \zeta_t + \delta Z_{i,t} + \epsilon_{i,t}
\end{equation}
We note the conditions for exogeneity in the instrument (following the discussion presented by \citealt{NBERw27885}).
The instrument is exogenous if state policy decisions for funding of public universities are uncorrelated with unobserved institutional changes of any specific college or university in the state \citep{borusyak2022quasi}.
This assumption is plausible given that the majority of states have multiple (i.e. more than five) public universities, without any single university campus receiving the majority of state funding within any single state.
Secondly, addressing endogeneity in university revenues by the shift-share identification strategy requires exogeneity in either the base-line share or shift component of the instrument.
In this case, we satisfy the second: universities' institutional-level decisions are not correlated with contemporaneous or upcoming shocks to state appropriations.\footnote{
    It would be plausible to consider the case that universities make institutional-level decisions in a consistently different manner to those 
    with differing reliance on state appropriations in 1990, so that exogeneity by the base-line share is not plausible here.
}

\begin{table}[h!]
    \singlespacing
    \centering
    \caption{First Stage Estimates, for State Funding by Appropriation Shock.}
    \makebox[\textwidth][c]{\input{tables/firststage-reg-fte.tex}}
    \label{tab:firststage-reg}
    \begin{flushleft}
        \footnotesize
        \textbf{Note}: Standard errors are clustered at the state-year level.
    \end{flushleft}
\end{table}

\autoref{tab:firststage-reg} presents results of the first-stage regression, separately with and without a control for tuition revenue per student, plus institution and year fixed effects.\footnote{
    Representations for frequentist significance levels (i.e. 10, 5, 1\% etc.) are omitted here, and in all following tables.
}
We note columns (1) and (2) estimate that a shock to appropriations (per student in the entire state) of 10\% is associated with 9.8\% change in state funding per student at the university; the instrument is strong, and we note similarity in estimates with and without inclusion of fixed effects.
Columns (3) and (4) include the tuition revenue control (explained in \autoref{foot:control}) to exhibit estimates with the inclusion of this possible collider or bad control.
Column (3) shows similar estimates to columns (1), (2) thanks to inclusion of fixed effects, so that the fixed effects was effective in soaking variation in per-student tuition revenues at the institution-year level.
Column (1) represents the estimates for \autoref{eqn:firststage} with fixed effects, omitting the tuition revenue control, and is the preferred form that I proceed with in the following.

\begin{figure}[h!]
    \centering
    \singlespacing
    \caption{Local Projection Estimates for First-Stage \autoref{eqn:firststage}.}
    \includegraphics[width=0.6\textwidth]{figures/firststage-lp.png}
    \label{fig:firststage-lp}
\end{figure}

\autoref{fig:firststage-lp} presents local projection estimates for the staying-power of the state appropriation shock to university revenues, with each year showing the years following effect of a shock to state reveunes in year zero.
We see a decaying, yet real and positive, effect of the shock on revenues for the next 10 years, illustrating that the state appropriation shock is a strong, if fading, instrument for state funding per student and justifying its use in local projection IV estimation.

These results, together with the case for exogeneity, show the appropriations shock instrument strongly predicts university revenues in the first-stage estimation.

\subsection{Instrumental Variables Model, University-Level}
\label{sec:iv-model-uni}

The primary empirical model combines the instrumental variable for state appropriation shocks with the empirical model for the effects of state funding at the university-level --- i.e. parameter $\beta$ in the following.\footnote{
    It is important to note the treatment effect isolated here; the instrumental variables approach identifies the local average treatment effects (LATE) specific to the instrument.
    So we interpret this treatment effect as a university's response in employment count and average salaries to state funding changes, changes specific to state appropriation shocks, among the complier group --- i.e. universities with any exposure to shocks, assuming no universities increase employment or salaries in response to negative revenue shocks.
    \cite{mogstad2021causal} provide a full discussion of the LATE with instruments and multiple monotonicity conditions.
}
\begin{eqnarray}
    \label{eqn:secondstage1}
    X_{i,t} &=& \eta_i + \zeta_t + \delta Z_{i,t} + \epsilon_{i,t} \\
    \label{eqn:secondstage2}
    Y_{i,t} &=& \alpha_i + \gamma_t + \beta \widehat X_{i,t} + \varepsilon_{i,t}
\end{eqnarray}
I estimate the system by two stage least squares, including institution and year fixed effects, and investigate outcomes at the university-level to analyse effects on the university as a result of changes in revenues.
Additionally, I estimate the model via local projections \citep{jorda2005,miller2022} to investigate whether the effects of revenues on faculty composition linger for multiple years after the original appropriation shock.
Regarding outcomes, I focus on the composition of the professors employed at the university by analysing (log) count per student, and average salaries paid to, professors employed by the university.

\begin{figure}[H]
    \centering
    \singlespacing
    \caption{Instrument Variables Model for University Finances and Faculty Composition.}
    \begin{tikzpicture}
        \node[state] (instrument) at (0,0) {$Z$};
        \node at (-2.6,0.5){State};
        \node at (-1.9,0){appropriation};
        \node at (-2.6,-0.5){shock};
        \node[state] (endogenous) [right=of instrument] {$X$};
        \node at (2,1.5) {University};
        \node at (2,1) {revenues};
        \node[state] (outcome) [right=of endogenous] {$Y$};
        \node at (5.2,0.25) {Faculty};
        \node at (5.6,-0.25) {Composition};
        % Relevance condition
        \path (instrument) edge (endogenous);
        \path (endogenous) edge (outcome);
        % Exclusion Restriction
        \path (instrument) edge[bend right=60] (outcome);
        \node[color=red] at (2,-1.2) {X};
    \end{tikzpicture}
    \label{fig:SCM-ivmodel}
\end{figure}

\autoref{fig:SCM-ivmodel} presents the system in graphical form, for the causal effect of state appropriation shocks on university revenues and thus faculty composition, making clear the assumption that appropriation shocks affect faculty composition only via affecting university finances.

\subsection{Instrumental Variables Model, Individual Professor-Level}
\label{sec:iv-model-indiv}

This analysis additionally uses data regarding individual professors in the Illinois university system, to investigate the effects of changes in university revenues on the individual professors at the universities.
Redefining the level of outcome requires adjustment to the empirical approach, leveraging variation in university finances for the years after a professor joins the university.\footnote{
    This formulation follows that presented by \cite{NBERw27885}, where individual student outcomes are analysed via variation in state appropriations after their freshman-year.
    This contrasts with \autoref{sec:iv-model-uni} and \cite{NBERw23736}, where the unit of analysis is the university-year.
}

\autoref{eqn:rolling-instrument} defines a rolling-share variant of the instrument, $\tilde Z_{j,t}$, where the university's state appropriations share exposure is based in the year a professor joins the university --- and not the base period 1990--1993.
$j$ indexes each professor in year $t$, $\tau(j)$ for the year the professor first joins their institution.\footnote{
    Identifying $\tau(j)$ is possible for $j$ by restricting to all professors hired 2011-2021 --- i.e. in the years after the start of the full panel.
    It is not possible to discern the hiring year for professors who  were hired in the years preceding 2011, and so the entire sample is only possible to analyse using the base-share in years 1990-1993 formulation (e.g., \autoref{tab:facultysalaries-shock-illinois}, \ref{tab:promotion-shock-illinois}).
}
\begin{align}
    \label{eqn:rolling-instrument}
    \tilde Z_{j,t} &\coloneqq - \log \left[
    \left( \frac{\text{Total State Funding}_{s(j),t}}{\text{Student Population}_{s(j),t}} \right)
    \left( \frac{\text{State Funding}_{\tau(j)}}{\text{Total Revenues}_{i,\tau(j)}} \right) \right]
\end{align}

This approach leverages an insight, made available by level of the data: that an individual professor is affected by changes in university revenues after they have joined the university.\footnote{
    Notice that \autoref{sec:iv-model-uni} considers the number of professors employed by the university; whether a professor becomes employed at the university is likely affected by that university's finances.
    The formulation in this section does not consider whether the professor joins the university, instead taking as given that the professor is employed at the university, and then projects the effect on the individual.
}
Exogeneity and relevance of the rolling-share instrument, $\tilde Z_{j,t}$, follows the same reasoning as that for the base-share instrument, $Z_{i,t}$, discussed in \autoref{sec:approp-shocks}.\footnote{
    The base-share instrument is appropriate for some outcomes with the individual Illinois professors, where appropriate (\autoref{tab:facultysalaries-shock-illinois}, \ref{tab:promotion-shock-illinois}).
}
We satisfy the assumptions for exogeneity by noting that none of the Illinois public campuses take the majority of state appropriations, and that the instrument identification strategy relies on exogeneity in changes in state appropriations to individual professor-outcomes, following the year they joined the university.\footnote{
    Additionally, within-institution changes resulting from share reliance on state funding may be correlated with unobserved changes in the outcomes, so that \cite{NBERw27885} note the importance of controlling for the base share and state student population.
    The formulation here implicitly controls for these factors via the fixed effects; results are relatively similar while including these controls without including fixed effects, and so are omitted.
}
\autoref{tab:firststage-illinois} presents results of the first stage estimation, showing that the instrument is strong in the same way as that for the university-level outcomes (\autoref{tab:firststage-reg}), with very similar estimates for the association between appropriation shocks and non-institutional revenues.

The instrumental variables model is then defined as follows where $i(j)$ refers to the institution that professor $j$ is employed at, and $Y_{j,t}$ refers to individual-level outcomes total salary, rate of promotion, and propensity to leave the Illinois public university system.
The system includes a fixed effect for the institution and first year of employment.\footnote{
    The instrument varies by institution, based in the year of first employment, so that these are the corresponding fixed effects and levels of clustered standard errors.
}
\begin{eqnarray}
    \label{eqn:secondstage1_indiv}
    X_{i(j),t} &=& \theta_{i(j)} + \phi_{\tau(j)} + \delta \tilde Z_{i(j),t} + \epsilon_{i(j),t} \\
    \label{eqn:secondstage2_indiv}
    Y_{j,t} &=& \mu_{i(j)} + \nu_{\tau(j)} + \beta \widehat X_{i(j),t} + \varepsilon_{j,t}
\end{eqnarray}
We then interpret parameter $\beta$ as the effect of changes in state funding at an Illinois public university, via state appropriation shocks, on an individual professor's outcome $Y_{j,t}$.

### 2022-11-20 text/sections/05-results.tex
Updated draft.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Results section
\section{Results}
\label{sec:results}

\subsection{University-Level}

\autoref{tab:facultycount-shock-reg} presents OLS and IV estimates, where the outcome is (log) count of professors employed at the university, separated by position, and for all professors (Columns 7, 8).
For all professors we see a positive correlation between university revenues and professor count per student (the OLS columns), where an increase in university total revenues by 10\% is associated with an increase in professor count per student of 0.83\%, yet negatively correlated with lecturer count of -1.3\%.
Lecturers per student is not shown to be correlated with total university revenues.
The 2SLS columns show results after identifying state funding with the appropriation shock.
All professor count per student increases by 0.53\% in response to a 10\% rise in total revenues.
The effect is driven by increases in the count of assistant and full professors (i.e. tenure-track and tenured) per student, where assistant professor count per student increase by 1.4\%, and full professor 1.2\%.
The count of lecturers per student decreases by 4.4\% in response to a (positive) 10\% shock in university revenues.

\begin{table}[h!]
    \singlespacing
    \centering
    \caption{OLS and 2SLS Estimates for University Faculty Composition.}
    \makebox[\textwidth][c]{\input{tables/facultycount-shock-reg-fte.tex}}
    \begin{flushleft}
        \footnotesize
        \textbf{Note}: Standard errors are clustered at the state-year level.
    \end{flushleft}
    \label{tab:facultycount-shock-reg}
\end{table}

\autoref{fig:all-count-lp} shows local projection estimates, where the effect on total professor count per student persists at the same magnitude of around 0.1 four years after the identified shock.
These findings are in line with \cite{turner2014impact} observing that universities froze hiring particularly for tenure-track positions in response to negative budget shocks in the late 2000s, and \cite{brown2014endowment} from shocks to university endowments.
We see a negative coefficient for the number of lecturers per student.
This effect lines up with two trends we see in \autoref{sec:trends}: that public universities revenues (per student) are decreasing while utilisation of non-tenure track lecturers increased over the same time period.
And yet, identifying the university revenues by the state appropriations shock increases the magnitude of the effect (Column 2) with respect to the association (Column 1), strengthening the case that the two trends are causally linked.
Between years 1990-1999 IPEDS provided count of professors by explicit tenure designation (opposed to named position as in \autoref{tab:facultycount-shock-reg}), and \autoref{tab:tenurecount-shock-reg-fte} presents results of the respective regressions by tenure status.
Results are relatively similar, yet imprecise from the much smaller number of observations.

Together these findings show that public universities increase (decrease) their count of tenure-track and tenured professors per student in years when revenues are more (less) plentiful, possibly by increasing hiring intensity or efforts to maintain incumbent professors.
In the same vein, when revenues are (not) bountiful, count of tenure-track and tenured professors per student increases (decreases).
%Yet, further analysis is needed to disentangle the possible long-run effects of state revenues on faculty composition among public universities.

\subsection{Long-term Effects on Faculty Composition}

\begin{figure}[h!]
    \centering
    \singlespacing
    \caption{Local Projection Estimates for Professor Count per Student, by Professor Group.}
    \begin{subfigure}[b]{0.495\textwidth}
        \centering
        \caption{Lecturers.}
        \includegraphics[width=\textwidth]{figures/lecturer-count-lp.png}
        \label{fig:lecturer-count-lp}
    \end{subfigure}
    \begin{subfigure}[b]{0.495\textwidth}
        \centering
        \caption{Assistant Professors.}
        \includegraphics[width=\textwidth]{figures/assistant-count-lp.png}
        \label{fig:assistant-count-lp}
    \end{subfigure}
    \begin{subfigure}[b]{0.495\textwidth}
        \centering
        \caption{Full Professors.}
        \includegraphics[width=\textwidth]{figures/full-count-lp.png}
        \label{fig:full-count-lp}
    \end{subfigure}
    \begin{subfigure}[b]{0.495\textwidth}
        \centering
        \caption{All Professors.}
        \includegraphics[width=\textwidth]{figures/all-count-lp.png}
        \label{fig:all-count-lp}
    \end{subfigure}
    \label{fig:count-lp}
\end{figure}

Decreases in university revenues (via state funding shocks) affect faculty composition for at most 4 years after the initial revenue shock.
A 10\% decrease in state funding increases count of lecturers per student by 4\% for the first two years after the initial shock.
While count of assistant professors decreases 10 to 18.5\% three years later;
count of tenured professors decreases 9 to 6\% three years later.
Together, the total count of professors per student decreases around 1\% for three years after the initial funding shock, and the effect is not distinguishable from zero four years after.

\autoref{fig:count-lp} represents these estimates, using the local projection method, as a series of impulse responses for each outcome of professor count.
The mechanism that explains the changes in faculty composition are, however, not clear at this unit of analysis; individual data are needed to delve deeper.

\subsection{Individual Professor Outcomes}

I first examine whether a shock to the university's revenues affects the salaries of professors at the university.
The first outcome is first-year salary for each professor by their position;
\autoref{tab:newhiresalaries-shock-illinois-rolling} presents 2SLS estimates for the elasticity of total salaries with respect to university revenues, among the sample with a year of starting employment after 2010.
New hires of assistant, full, or administrative position professors are not related to revenues, via a shock to appropriations in their first-year of employment.
Yet, the starting salary of lecturers is 3.8\% higher given an increase in their hiring university's state funding of 10\%.

\begin{table}[h!]
    \singlespacing
    \centering
    \caption{2SLS Estimates for Faculty Salaries, in First-Year, at Illinois Universities.}
    \makebox[\textwidth][c]{\input{tables/newhiresalaries-shock-illinois-rolling.tex}}
    \begin{flushleft}
        \footnotesize
        \textbf{Note}: Standard errors are clustered at the institution and first year of employment level. 
    \end{flushleft}
    \label{tab:newhiresalaries-shock-illinois-rolling}
\end{table}

Regarding total salary for every year of employment, \autoref{tab:facultysalaries-shock-illinois-rolling} presents estimates for the elasticity of total salary with respect to university revenues, for the entire sample of professors hired after 2010.\footnote{
    Recall that this is a subsample of the entire data, that the rolling shift-share instrument is only defined for the sample with an observed hiring date in the years 2011-2021.
}
The elasticity is non-distinguishable from zero among any position, implying that long-term salaries are not related to state funding.
\autoref{tab:facultysalaries-shock-illinois} presents the same model, for the sample of all professors and using the instrument based in years 1990-1993, similarly finding no relationship.

\begin{table}[h!]
    \singlespacing
    \centering
    \caption{2SLS Estimates for Faculty Salaries at Illinois Universities.}
    \makebox[\textwidth][c]{\input{tables/facultysalaries-shock-illinois-rolling.tex}}
    \begin{flushleft}
        \footnotesize
        \textbf{Note}: Standard errors are clustered at the institution-year level.
    \end{flushleft}
    \label{tab:facultysalaries-shock-illinois-rolling}
\end{table}

Note that professor position (assistant, associate, full) is also an outcome, thanks to the promotion channel: a university may be less able to promote their professors to higher paying positions when their revenues are constrained.
\autoref{tab:promotion-shock-illinois-rolling} shows estimates where the outcome is rate of promotion within each professor position.
The lecturer position describes the sample of professors who were ever listed as a non-tenure track lecturer and the binary for promotion equals one in a year they achieved an assistant professor position; assistant the same for the assistant to associate professor promotion; associate same for associate to full promotion.
There is no discernible relationship between revenues and promotion among any position.\footnote{
    \autoref{tab:promotion-shock-illinois} shows estimates using the base-year instrument among the entire sample, finding the same results.
}
Furthermore, local projection estimates (\autoref{fig:promoted-illinois-lp}) show that the there is no promotion effect, among all professors, for any of the following years either.
Lastly, \autoref{tab:facultyleaving-shock-illinois-rolling}, \ref{tab:facultyleaving-shock-illinois} describes estimates for rate of exit from the employing university among Illinois professors, and finds no relationship with university revenues.

### 2022-11-20 text/sections/06-discussion.tex
Updated draft.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Discussion section
\section{Discussion}

The results show that the recent stagnation in state funding for higher education has affected the composition of faculty within public universities, away from tenure-track and tenured professors towards non-tenured lecturers.
At the same time, there are little discernible effects on individual professors hired in the years 2011-2021 at Illinois public universities, except for the salaries of first year lecturers.
The rate that faculty leave their university, and their rate of promotion between positions, are unaffected by the university revenues, so this leaves one primary channel to explain changes in faculty composition: hiring.

\cite{turner2014impact} documents the wide-spread practice of hiring freezes at universities in response to budget shocks around the 2008 recession.
Throughout the last decade, multiple such measures were taken by Illinois public universities in response to their deteriorating finances \citep{furlough2010}.
The University of Illinois\footnote{
    The University of Illinois includes three public university campuses: Urbana-Champaign, Chicago, Springfield.
}
did not receive the allocated state appropriations from the state of Illinois on time, so enacted cost-cutting measures to stay fiscally solvent.
The university system placed a hold on all hiring for filling state-funded positions and promotions.
Faculty were furloughed (placed on leave without pay) for a day each month, and university administrators placed on two days per month furlough, or eligible university employees could accept a voluntary, equivalent pay reduction.
The University of Western Illinois adopted very similar measures in response to the Illinois budget crisis in 2016-2017 \citep{wiu2016}.

\begin{figure}[h!]
    \centering
    \singlespacing
    \caption{Trends in New Hires at Illinois Public Universities 2011-2021.}
    \begin{subfigure}[b]{0.495\textwidth}
        \centering
        \caption{New Hire Count, Total.}
        \includegraphics[width=\textwidth]{figures/newhire-count-illinois.png}
        \label{fig:newhire-count-illinois}
    \end{subfigure}
    \begin{subfigure}[b]{0.495\textwidth}
        \centering
        \caption{Mean Salary in First Year for New Hires, \$ 2021 CPI-U.}
        \includegraphics[width=\textwidth]{figures/newhire-salary-illinois.png}
        \label{fig:newhire-salary-illinois}
    \end{subfigure}
    \label{fig:newhire-illinois}
\end{figure}

These examples show that Illinois public universities were effected by the negative shock to state funding for higher education, and instituted policy changes aimed at affecting faculty outcomes investigated in \autoref{sec:results}.
Yet the only measurable effect was on count of professor per student (implied via slower hiring), and first-year lecturer salaries.
To investigate these trends for the entire state, \autoref{fig:newhire-illinois} shows the count of professors hired in the years 2011-2021, and their mean salary, by position.
In particular, no position had sustained rises in hiring while enrolment within the university was rising (so that new hires per student necessarily fell), and hiring of new administrators jumped once the Illinois budget crisis ended in 2017.
At the same time, professor new hires' starting salaries remained relatively constant for lecturer and assistant professors, yet rose and fell over the decade for full professors and administrators.

### 2022-11-20 text/sections/07-conclusion.tex
Updated draft.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Conclusion section
\section{Summary and Concluding Remarks}
\label{sec:conclusion}

This analysis investigates how the recent stagnation in state support for higher education has affected public universities, and their faculty composition.
The work contributes to the literature along two primary dimensions.
Firstly, by exploiting the same empirical approach as \cite{NBERw23736,NBERw27885} to isolate the causal effect of changes in state funding on public universities themselves, the analysis provides a plausible mechanism for the observed negative effects on students outcomes that this stagnation has brought.
Secondly, this analysis considers individual professors as the unit of analysis using a dataset (IBHED) new in the economic literature.
These data allow for detailed analysis of thousand's of professors' salaries, and are a starting point for further analysis of the determinants of professors wages in the Illinois public university system.

Findings show that public universities have substituted away from tenure-track and tenured professors, towards non-tenured lecturers, in the face of persistent declines in state funding for higher education.
The effects on incumbent professors in the state of Illinois are non-distinguishable from zero, which implies that the effect is driven by changes in university hiring of new professors.
Public universities are hiring thousands of non-tenured lecturers to satisfy teaching obligations, and fewer tenure-track and tenured professors that their private counterparts.

While costs education have been rising in the US, thanks to multiple overlapping trends \citep{ehrenberg2012}, public universities have also dealt with declining state support.
It is natural to expect that such headwinds will lead to systematic change at public universities, changes which are not neutral to the composition of faculty, or to their goals of research and education.
Thes results show large changes in faculty composition, and that stagnation in state support explains at least a third of the observed shift away from tenured professors and towards contingent lecturers.
At the same time, private universities were not exposed to financial headwinds of the same magnitude or persistence.
If public universities' hiring of faculty is disrupted by a fall in state funding, then we can expect that private universities will benefit (in a competitive sense) by being able to hire better professors from relatively resource constrained public universities.
While public universities continue to educate the majority of higher education students in the US, we should worry about the effects that restricting their financial support has on the composition of faculty, and  higher education itself, in this country.

### 2022-11-20 text/sections/09-appendix.tex
Updated draft.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Appendix section
% Set-up the section.
\newpage
\appendix
\setcounter{table}{0}
\renewcommand{\thetable}{A\arabic{table}}
\setcounter{figure}{0}
\renewcommand{\thefigure}{A\arabic{figure}}

% Start appendix
\section{Appendix}
\label{appendix}
This project used data which are fully public, and computational tools which are fully open-source.
As such, all code and data involved in this project are available at this project's Github repository, available at \url{https://github.com/shoganhennessy/state-faculty-composition}.
They may be used for replication, or as the basis for further work, as needed.
Any comments or suggestions may be sent to me at \href{mailto:seh325@cornell.edu}{\nolinkurl{seh325@cornell.edu}}, or raised as an issue on the Github project.

A number of statistical packages, for the $R$ language \citep{R2022}, made the empirical analysis for this paper possible.
\begin{itemize}
    \item \textit{Tidyverse} \citep{tidyverse} collected tools for data analysis in the R language.
    \item \textit{LFE} \citep{lfe} implemented linear fixed effect models, with instruments, crucial for the empirical estimation in \autoref{sec:empirics}.
    \item \textit{Stargazer} \citep{stargazer} provided a method to efficiently convert empirical results into presentable output in \LaTeX.
    \item \textit{Lpirfs} \citep{lpirfs2019} implemented estimation of the \cite{jorda2005} local projections methods, with instrumental variables, crucial to the local projections estimates presented in this project.
\end{itemize}

\subsection{First Stage Estimates, Individual Outcomes}
\label{appendix:part1}

\begin{table}[h!]
    \singlespacing
    \centering
    \caption{IBHED Summary Statistics, Entire Professor Panel 2010--2021.}
    \makebox[\textwidth][c]{\input{tables/illinois-summary-rolling.tex}}
    \label{tab:illinois-summary-rolling}
    \begin{flushleft}
        \footnotesize
        \textbf{Note}: This table presents the summary statistics for professors with an observed year of first-year at their university.
    \end{flushleft}
\end{table}

\begin{table}[H]
    \singlespacing
    \centering
    \caption{First Stage Estimates, for Total University Revenues at the Individual-Level.}
    \makebox[\textwidth][c]{\input{tables/firststage-illinois-rolling.tex}}
    \label{tab:firststage-illinois}
    \begin{flushleft}
        \footnotesize
        \textbf{Note}: Standard errors are clustered at the university-year level.
    \end{flushleft}
\end{table}

\begin{figure}[h!]
    \centering
    \singlespacing
    \caption{Local Projection Estimates for First-Stage \autoref{eqn:secondstage1_indiv}.}
    \includegraphics[width=0.6\textwidth]{figures/firststage-illinois-lp-rolling.png}
    \label{fig:firststage-illinois-lp}
\end{figure}

\subsection{Additional Results, University-Level}

\autoref{tab:facultysalaries-shock-reg} presents OLS and IV estimates where the outcome is (log) mean salary for professors employed at the university, separated by position.
Similarly, \autoref{fig:all-salaries-lp} shows the staying-power via projection methods.
The outcome, mean salary is crude.

\begin{table}[H]
    \singlespacing
    \centering
    \caption{OLS and 2SLS Estimates for University Faculty-Tenure Composition.}
    \makebox[\textwidth][c]{\input{tables/tenurecount-shock-reg-fte.tex}}
    \begin{flushleft}
        \footnotesize
        \textbf{Note}: Standard errors are clustered at the university-year level.
    \end{flushleft}
    \label{tab:tenurecount-shock-reg-fte}
\end{table}

\begin{table}[H]
    \singlespacing
    \centering
    \caption{OLS and 2SLS Estimates for University Faculty Salaries.}
    \makebox[\textwidth][c]{\input{tables/facultysalaries-shock-reg-fte.tex}}
    \begin{flushleft}
        \footnotesize
        \textbf{Note}: Standard errors are clustered at the state-year level.
    \end{flushleft}
    \label{tab:facultysalaries-shock-reg}
\end{table}

\begin{figure}[H]
    \centering
    \singlespacing
    \caption{Local Projection Estimates for Professor Salary, Mean within Each University.}
    \includegraphics[width=0.6\textwidth]{figures/all-salaries-lp.png}
    \label{fig:all-salaries-lp}
\end{figure}

\subsection{Additional Results, Individual-Professor Level}

\begin{table}[H]
    \singlespacing
    \centering
    \caption{2SLS Estimates for Faculty Salaries at Illinois Universities.}
    \makebox[\textwidth][c]{\input{tables/facultysalaries-shock-illinois.tex}}
    \begin{flushleft}
        \footnotesize
        \textbf{Note}: Standard errors are clustered at the institution-year level.
    \end{flushleft}
    \label{tab:facultysalaries-shock-illinois}
\end{table}

\begin{table}[H]
    \singlespacing
    \centering
    \caption{2SLS Estimates for Faculty Promotion Rate at Illinois Universities, using Rolling Instrument.}
    \makebox[\textwidth][c]{\input{tables/promotion-shock-illinois-rolling.tex}}
    \begin{flushleft}
        \footnotesize
        \textbf{Note}: Standard errors are clustered at the institution and first year of employment level.
    \end{flushleft}
    \label{tab:promotion-shock-illinois-rolling}
\end{table}

\begin{table}[H]
    \singlespacing
    \centering
    \caption{2SLS Estimates for Faculty Promotion Rate at Illinois Universities, using Base-Year Instrument.}
    \makebox[\textwidth][c]{\input{tables/promotion-shock-illinois.tex}}
    \begin{flushleft}
        \footnotesize
        \textbf{Note}: Standard errors are clustered at the institution and first year of employment level.
    \end{flushleft}
    \label{tab:promotion-shock-illinois}
\end{table}

\begin{figure}[H]
    \centering
    \singlespacing
    \caption{Local Projection Estimates for Promotion Rate at Illinois Universities.}
    \includegraphics[width=0.6\textwidth]{figures/promoted-illinois-lp-rolling.png}
    \label{fig:promoted-illinois-lp}
\end{figure}

\begin{table}[H]
    \singlespacing
    \centering
    \caption{2SLS Estimates for Faculty Exit Rate at Illinois Universities, using Rolling Instrument.}
    \makebox[\textwidth][c]{\input{tables/facultyleaving-shock-illinois-rolling.tex}}
    \begin{flushleft}
        \footnotesize
        \textbf{Note}: Standard errors are clustered at the institution and first year of employment level.
    \end{flushleft}
    \label{tab:facultyleaving-shock-illinois-rolling}
\end{table}

\begin{table}[H]
    \singlespacing
    \centering
    \caption{2SLS Estimates for Faculty Exit Rate at Illinois Universities, using Base-Year Instrument.}
    \makebox[\textwidth][c]{\input{tables/facultyleaving-shock-illinois.tex}}
    \begin{flushleft}
        \footnotesize
        \textbf{Note}: Standard errors are clustered at the institution and first year of employment level.
    \end{flushleft}
    \label{tab:facultyleaving-shock-illinois}
\end{table}

### 2022-12-06 text/paper.tex
Updated draft.
adjust desc.

@@
     }
-\date{\today\thanks{
+\date{Last updated: \today\thanks{
     I thank Professor Evan Riehl for guidance on this project, Michael Lovenheim, Ronald Ehrenberg, Douglas Miller, Francine Blau, and Ryan Dycus for various comments on the project's direction, as well as seminar participants at Cornell University (Autumn 2022) for helpful discussion.
@@
 }
-\rhead{\today}
+\rhead{1 December 2022}
 \lhead{Faculty Composition and State Support for Higher Education}
@@
 \input{sections/07-conclusion.tex}
+
 % Bibliography
+\singlespacing
 \bibliographystyle{agsm}

### 2022-12-07 text/sections/00-abstract.tex
adjust desc.

@@
 A decrease in state funding of 10\%, via a shock to state appropriations, decreases the number of assistant professors per student at a public university by 1.4\% and full professors by 1.2\%, yet increases the number of lecturers per student by 4.4\%.
-Local project estimates show that these effects linger for up to three years after the initial shock.
-Analysis of all the professors hired at Illinois public universities 2011-2021 shows that incumbent professors are not affected by the changes in state support, implying that these changes in faculty composition arose by reduced hiring of tenure-track and tenured professors at public universities.
+Local projection estimates show that these effects linger for up to three years after the initial shock.
+Analysis of all the professors at Illinois public universities 2011-2021 shows that incumbent professors are not affected by the changes in state support, implying that these changes in faculty composition arose by reduced hiring of tenure-track and tenured professors at public universities.
 These results exhibit the long-term effects of stagnating state support for higher education, and raises questions about the direction that public education heads as these financial headwinds show no sign of dissipating.
 
-\vspace{0.75cm}
+\vspace{0.5cm}
 \noindent\textbf{Keywords:}
@@
 Higher Education,
-Public Sector Labor Markets
+Public Sector Labour Markets
 
-\vspace{0.5cm}
+\vspace{0.25cm}
 \noindent\textbf{JEL Codes:} H72, I23, J45

### 2023-01-27 text/paper.tex
adjust desc.
placed a copy of the paper in base directory.

--- 2022-11-20 text/paper.tex
@@
 % Author
-\author{Senan Hogan-Hennessy\thanks{
-    Economics Department, Cornell University.}
+\author{Senan Hogan-Hennessy \\ \vspace{0.1cm}
+    Economics Department, Cornell University
 }
@@
     }
-\date{\today\thanks{
-    I thank Professor Evan Riehl for guidance on this project, Michael Lovenheim, Ronald Ehrenberg, Douglas Miller, Francine Blau, and Ryan Dycus for various comments on the project's direction, as well as seminar participants at Cornell University (Autumn 2022) for helpful discussion.
+\date{9 December 2022%\today
+    \thanks{
+    I thank Professor Evan Riehl for guidance on this project, Michael Lovenheim, Ronald Ehrenberg, Douglas Miller, Francine Blau, Levon Barseghyan, and Ryan Dycus for various comments on the project's direction, as well as seminar participants at Cornell University (Autumn 2022) for helpful discussion.
     This project's Github repository, with materials and (publicly available) data for replication, is available at 
@@
     Any comments or suggestions may be sent to me at \href{mailto:seh325@cornell.edu}{\nolinkurl{seh325@cornell.edu}}, or raised as an issue on the Github project.
+    } \\ \vspace{0.1cm}
+    \href{https://github.com/shoganhennessy/state-faculty-composition/blob/main/text/paper.pdf}{Newest version available here.}
 }
-}
-\rhead{\today}
+\rhead{9 December 2022}%\today}
 \lhead{Faculty Composition and State Support for Higher Education}

### 2023-02-22 text/paper.tex
adjust README

@@
     }
-\date{9 December 2022%\today
-    \thanks{
+\date{\today\thanks{
     I thank Professor Evan Riehl for guidance on this project, Michael Lovenheim, Ronald Ehrenberg, Douglas Miller, Francine Blau, Levon Barseghyan, and Ryan Dycus for various comments on the project's direction, as well as seminar participants at Cornell University (Autumn 2022) for helpful discussion.
@@
 }
-\rhead{9 December 2022}%\today}
-\lhead{Faculty Composition and State Support for Higher Education}
+\rhead{\today}
+\lhead{Faculty Composition and State Support for Higher Education.}
 % Begin

### 2023-02-22 text/sections/04-empirical.tex
adjust README

@@
 }
-$Z_{i,t}$ is contructed as negative to reflect the fact that the long term trend in, and most of the short-run shocks to, state funding for higher education has been negative.
+$Z_{i,t}$ is constructed as negative to reflect the fact that the long term trend in, and most of the short-run shocks to, state funding for higher education has been negative.
 State funding has been been falling, so that the instrument describes shocks to university revenues, mostly in a negative direction.
@@
 The instrument is exogenous if state policy decisions for funding of public universities are uncorrelated with unobserved institutional changes of any specific college or university in the state \citep{borusyak2022quasi}.
-This assumption is plausible given that the majority of states have multiple (i.e.
-more than five) public universities, without any single university campus receiving the majority of state funding within any single state.
+This assumption is plausible given that the majority of states have multiple (i.e., more than five) public universities, without any single university campus receiving the majority of state funding within any single state.
 Secondly, addressing endogeneity in university revenues by the shift-share identification strategy requires exogeneity in either the base-line share or shift component of the instrument.
@@
 Column (3) shows similar estimates to columns (1), (2) thanks to inclusion of fixed effects, so that the fixed effects was effective in soaking variation in per-student tuition revenues at the institution-year level.
-Column (1) represents the estimates for \autoref{eqn:firststage} with fixed effects, omitting the tuition revenue control, and is the preferred form that I proceed with in the following.
+Column (1) represents the estimates for \autoref{eqn:firststage} with fixed effects, omitting the tuition revenue control, and is the preferred form that I proceed with.
 
@@
 
-\autoref{fig:firststage-lp} presents local projection estimates for the staying-power of the state appropriation shock to university revenues, with each year showing the years following effect of a shock to state reveunes in year zero.
+\autoref{fig:firststage-lp} presents local projection estimates for the staying-power of the state appropriation shock to university revenues, with each year showing the years following effect of a shock to state revenues in year zero.
 We see a decaying, yet real and positive, effect of the shock on revenues for the next 10 years, illustrating that the state appropriation shock is a strong, if fading, instrument for state funding per student and justifying its use in local projection IV estimation.
@@
 
-The primary empirical model combines the instrumental variable for state appropriation shocks with the empirical model for the effects of state funding at the university-level --- i.e.
-parameter $\beta$ in the following.\footnote{
+The primary empirical model combines the instrumental variable for state appropriation shocks with the empirical model for the effects of state funding at the university-level --- i.e., parameter $\beta$ in the following.\footnote{
     It is important to note the treatment effect isolated here; the instrumental variables approach identifies the local average treatment effects (LATE) specific to the instrument.
-    So we interpret this treatment effect as a university's response in employment count and average salaries to state funding changes, changes specific to state appropriation shocks, among the complier group --- i.e.
-universities with any exposure to shocks, assuming no universities increase employment or salaries in response to negative revenue shocks.
+    So we interpret this treatment effect as a university's response in employment count and average salaries to state funding changes, changes specific to state appropriation shocks, among the complier group --- i.e., universities with any exposure to shocks, assuming no universities increase employment or salaries in response to negative revenue shocks.
     \cite{mogstad2021causal} provide a full discussion of the LATE with instruments and multiple monotonicity conditions.

### 2023-02-22 text/sections/06-discussion.tex
adjust README

@@
 The rate that faculty leave their university, and their rate of promotion between positions, are unaffected by the university revenues, so this leaves one primary channel to explain changes in faculty composition: hiring.
+
+\subsection{Faculty Hiring at all US Universities}
+
+Write here about the faculty hiring network data, as provided by \cite{wapman2022quantifying}.
+
+Give a visualisation of the correlation between state funding per student (OLS) and faculty hiring.
+Also show the figure of faculty hiring in public vs private: did private unis hire more (in total and per student) over this time period?
+
+Cite Jackson (2021), effects of spending shock.
+
+IV model, where faculty hiring (per student) is the outcome, are presented in \autoref{tab:hiring-shock-reg}.\footnote{
+    Give extra details on this specification.
+}
+
+\begin{table}[h!]
+    \singlespacing
+    \centering
+    \caption{OLS and 2SLS Estimates for University Faculty Hiring, Total for 2011--2020.}
+    \makebox[\textwidth][c]{\input{tables/hiring-shock-reg-fte.tex}}
+    \begin{flushleft}
+        \footnotesize
+        \textbf{Note}: Standard errors are clustered at the state level.
+    \end{flushleft}
+    \label{tab:hiring-shock-reg}
+\end{table}
+
+Note that yearly variation is not observed here, so that only aggregate level, for 180 universities, can be considered.
+Leave it to further research to delve deeper into the role that hiring plays as the primary causal mediator for the effect of state funding shocks on faculty composition.
+
+\subsection{Faculty Hiring at Illinois Public Universities}
 
@@
 
-These examples show that Illinois public universities were effected by the negative shock to state funding for higher education, and instituted policy changes aimed at affecting faculty outcomes investigated in \autoref{sec:results}.
+These examples show that Illinois public universities were affected by the negative shock to state funding for higher education, and instituted policy changes aimed at affecting faculty outcomes investigated in \autoref{sec:results}.
 Yet the only measurable effect was on count of professor per student (implied via slower hiring), and first-year lecturer salaries.

### 2023-02-22 programs/analysis/ipeds-public-private.R

@@ -17,6 +17,7 @@ fig.height <- fig.width * 0.75
 
 
 # Load data sources ------------------------------------------------------------
+
 # Load IPEDS data on all universities
 ipeds.data <- read_csv("../../data/urban-ipeds/urban-clean-allunis.csv")
 

### 2023-02-22 programs/run.sh

@@ -9,6 +9,8 @@ cd data-build
 R CMD BATCH --no-save urban-ipeds-build.R
 # Build individual-level analysis data-set from raw data on Illinois Profs
 R CMD BATCH --no-save illinois-build.R
+# Build uni-level analysis data-set from open-sorce data on hiring networks.
+R CMD BATCH --no-save hiring-build.R
 cd ..
 
 # Statistical analysis code
@@ -21,6 +23,8 @@ R CMD BATCH --no-save ipeds-shock.R
 R CMD BATCH --no-save individual-shock.R
 # Analysis and regressions at the individual-level, with rolling shock IV.
 R CMD BATCH --no-save individual-shock-rolling.R
+# Analysis and regressions for rate of faculty hiring.
+R CMD BATCH --no-save hiring-shock.R
 cd ..
 
 # Compile paper, with outputs of progams as inputs for TeX files

### 2023-03-17 text/sections/00-abstract.tex
Half-way through a major revision.  Needs an intro section on sorting (see Hoxby+ 2008), and revisions of sections 4 (results) onwards.

@@
 \noindent
-State support for higher education has stagnated on a per-student level, while public universities have reduced their employment of tenure-track or tenured professors, and increased their employment of contingent lecturers (per student).
-This analysis relates these changes in the faculty composition at public universities to the fall in per-student university revenues from state appropriations.
-A shift-share instrumental variables approach addresses endogeneity in decisions for state appropriations to higher education, by exploiting the yearly change in state support with an institution's reliance on that support.
-A decrease in state funding of 10\%, via a shock to state appropriations, decreases the number of assistant professors per student at a public university by 1.4\% and full professors by 1.2\%, yet increases the number of lecturers per student by 4.4\%.
-Local projection estimates show that these effects linger for up to three years after the initial shock.
-Analysis of all the professors at Illinois public universities 2011-2021 shows that incumbent professors are not affected by the changes in state support, implying that these changes in faculty composition arose by reduced hiring of tenure-track and tenured professors at public universities.
+State support for higher education per-student has stagnated at the same time as public universities have reduced their employment of tenure-track or tenured professors, and increased their employment of temporary contingent lecturers.
+This analysis relates these changes in the faculty composition at public universities to the fall in per-student revenues from state funding for higher education.
+A shift-share instrumental variables approach addresses endogeneity in decisions for state higher education funding by exploiting (1) yearly change in each state's funding for their public universities and (2) each institution's reliance on that support.
+A decrease in state funding of 10\%, via a shock to state funding, decreases the number of assistant professors per student at a public university by 1.4\% and full professors by 1.2\%, yet increases the number of lecturers per student by 4.4\%.
+Local projection estimates show that these effects linger for up to three years after the initial funding shock.
+Analysis of all the professors at Illinois public universities 2011-2021 shows that incumbent professors are not affected by the changes in state support, while tenured-track hiring counts fall over this time period, showing that these changes in faculty composition arose by reduced hiring of tenure-track and tenured professors at public universities.
 These results exhibit the long-term effects of stagnating state support for higher education, and raises questions about the direction that public education heads as these financial headwinds show no sign of dissipating.
 
-\vspace{0.5cm}
-\noindent\textbf{Keywords:}
+\noindent
+\textbf{Key-words:}
 State and Local Budget and Expenditures,
@@
 
-\vspace{0.25cm}
-\noindent\textbf{JEL Codes:} H72, I23, J45
+\noindent
+\textbf{JEL Codes:} H72, I23, J45

### 2023-03-17 text/sections/01-introduction.tex
Half-way through a major revision.  Needs an intro section on sorting (see Hoxby+ 2008), and revisions of sections 4 (results) onwards.

@@
 I show that four-year, degree-granting public universities have experienced falling state funding per student, and this affected their composition of faculty.
-Falls in public university revenues from the state, instrumented by state-level finance shocks, lead to a fall in the number of tenure-track or tenured professors per student within a university, an increase in the number of lecturers, and overall fall in the number of faculty per student.
+Falls in public university revenues from the state, instrumented by state-level funding shocks, lead to a fall in the number of tenure-track or tenured professors per student within a university, and an increase in the number of lecturers.
 US private universities were not exposed to such financial constraints during this same time period, and do not exhibit the substitution away from tenure-track and tenured professors, so that the stagnating state support for public universities has implications for the wider structure of higher education instruction and research in the US.
@@
 This fall is driven by stagnating funding provided by state governments, while enrolment among public universities rose by 46\% over the same time period.
-At the same time, the number of professors per student at public universities stayed relatively stable at 0.045, yet fell from around 0.04 to around 0.35 professors per student at public universities, driven by a fall of over 20\% in the ratio of associate and full professors per student at public universities.
+At the same time, the number of professors per student at private universities stayed relatively stable at 0.045, yet fell from 0.04 to around 0.35 professors per student at public universities.
+This change was driven by a fall of over 20\% in the count of associate and full professors per student at public universities.
 
-I use a variant of the \cite{NBERw23736,NBERw27885} shift-share instrument for state funding from shocks to appropriations for the entire state to identify the effect on the faculty composition.
+I use a variant of the \cite{NBERw23736,NBERw27885} shift-share instrument for state funding, which exploits yearly shocks to the amount an entire state funds higher education to identify the change in a single public unversity's funding.
 State appropriations for public universities are plausibly endogenous to many outcomes: state governments decide on yearly budgets for their higher education sector, and this process can be influenced by local financial conditions, or even public perceptions of the state's higher education system.
@@
 
-Findings show that negative changes in state funding affects faculty composition at public universities, away from tenured and tenure-track professors towards non-tenure track lecturer positions.
+Negative changes in state funding affects faculty composition at public universities, away from tenured and tenure-track professors towards non-tenure track lecturer positions.
 A reduction of 10\% in state funding, via a shock to state appropriations, leads to a fall in 1.4\% in the number of assistant professors per student within a university, 1.2\% for full (tenured) professors, and an increase in 4.4\% the number of lecturers per student.
-Local projection methods show that these effects linger for a small number of years after the initial revenue shock.
-Over this time period, state funding fell by around 35\%, while the count of professors per student fell by 9\%, so that these results show that falls in state funding explain around a third of the observed shift away from tenure-track and tenure professors towards contingent faculty.
+Local projection estimates show that these effects linger for three to four years after the initial funding shock.
+Over the same time period, state funding fell by around 35\%, while the count of professors per student fell by 9\%, so that these results show that falls in state funding explain around a third of the observed shift away from tenure-track and tenure professors towards contingent faculty.
 
-Additionally, I use individual level data on professors at Illinois public universities over 2010-2021 to investigate whether the shocks to state funding affected incumbent professors at public universities.
-Incumbent professors were not meaningfully affected, in terms of total salary, promotion rate, or rate of leaving the Illinois public university system, yet lecturers' first year salary decreased by 3.8\% in response to a fall in state funding of 10\%.
-These findings imply that changes in faculty composition came about by changes in the composition of hiring new professors at the university level.
+Additionally, I use individual level data on professors at Illinois public universities over 2010-2021 to investigate whether the shocks to state funding affected individual professors.
+Incumbent professors were not meaningfully affected, in terms of total salary, promotion rate, or rate of leaving the Illinois public university system.
+Yet, the hiring rate of tenure-track professors declined over this time period for all public universities.
+This shows that changes in faculty composition came about by hiring reductions of new professors at public universities.
 
-\cite{NBERw23736} first study the effect of (changes in) public universities' finances on student outcomes by isolating changes in public universities' expenditures on student instruction, addressing endogeneity of expenditures decisions by instrumenting expenditures with state appropriation shocks and limits on tuition price.
-\cite{NBERw23736} find that increases in public university spending (via appropriation shocks) increases enrolment and degree completion among students, but corresponding changes in attendance price (via limits on tuition price) do not have an effect.\footnote{
+WRITE HERE ABOUT HOW IT IS NOVEL THAT THESE EFFECTS ARE CAUSAL, AND NOT JUST CONCURRENT TRENDS, IN COMPARISON TO HOXBY 2008 SORTING.
+
+A number of studies have shown that stagnating state funding for higher education negatively impacts student outcomes.
+\cite{NBERw23736} show that increases in public university spending (via appropriation shocks) increases enrolment and degree completion among students,\footnote{
     \cite{miller2022making} further analyse the effects of falls in university revenues, by finding that reductions in public universities tuition prices (suggestively motivated by falling state funding) leads to reductions in provided financial aid.
 }
-\cite{chakrabarti2018effect,NBERw27885} use the same instrument for public university revenues at the state level, combined with rich information on students' individual-level outcomes, to show that increases in state appropriations lead to degree lower completion time and later-life student debt.
-\cite{bound2019public} use another variant of the same approach to show that state-level higher education spending cuts induced public universities to shift toward tuition as their primary source of revenue, and shift institutional resources towards students who pay higher tuition rates.
-In a similar vein, \cite{bound2007cohort} document variation in per student state funding for higher education resulting from yearly changes in a state's student-age population; they find that lower per student funding lead to lower higher education completion rate among students.
+and \cite{bound2019public} that these shocks induce public universities to rley more on tuition revenue.
+\cite{NBERw27885} use the same methodology to show that increases in state appropriations lead to degree lower completion time\footnote{
+    \cite{bound2007cohort} use similar variation in state funding per student to show that lower funding leads to lower higher education completion rate at public universities.
+} and later-life debt for students.
 
-Faculty are a core component of this country's educational and research-innovation university system, yet their composition within universities, and its relationship with higher education finances has so far not been studied.
-\cite{brown2014endowment} presents the closest example by studying how university endowments react to negative financial shocks, exhibiting an association between negative endowment shocks and a fall in tenure-system faculty employed at private universities.
-\cite{abe2015implications} present a model for decision-making by university administrators, and posit that a large component of variation in faculty salaries arises from university politics, and not only economic factors such as outside options.
-\cite{johnson2009jep,NBERc13879} both empirically document allocation of faculty between research and instruction, and \cite{hemelt2021math} document differences in instructional cost (per student) between departments, primarily thanks to department differences in class size and faculty salaries.
-\cite{turner2014impact} document how universities reacted to the Great Recession of 2008, a severe financial shock to state finances and support for higher education.
+Faculty is a core component of the US higher education and research-innovation university system, yet their composition within universities, and its relationship with higher education finances has so far not been studied.
+\cite{brown2014endowment} presents the closest example by studying how university endowments react to negative endowment shocks, and \cite{turner2014impact} for the 2008 recession, while \cite{abe2015implications,johnson2009jep,NBERc13879} model university decision-making and faculty hiring.
+Lastly, it has been shown that instruction costs vary substantially across subject areas, primarily thanks to department differences in class size and faculty salaries \citep{hemelt2021math}.
 
 The number of professors, both in absolute terms and relative to number of students, may be particularly important for quality of instruction and thus educational outcomes.
-\cite{angrist1999using} first causally show that reducing class size (via religious ruling) induces an increase in test scores for school-age children, yet the magnitude of this effect is not as clear in the higher education setting.
-\cite{bandiera2010heterogeneous} use a fixed effect model, combined with multiple individual observations in a long panel, to find that UK university students perform worse academically in particularly large classes, and the difference is largest for students at the top of the academic distribution.
-For the composition of faculty teaching university classes, \cite{bettinger2010does,figlio2015tenure} find US students have better enrolment and learning outcomes from courses taught by contingent or adjunct (i.e.
-not tenured or on tenure-track) professors (relative to tenured or tenure-track).
-\cite{ehrenberg2005tenured} observe a negative association between utilisation of non-tenured instructors and student graduation rates.
-My approach focuses on count of professors per student at universities, so does not directly observe individual class-sizes, yet presents considers this a possible mechanism through which stagnating state support for higher education negatively affects student outcomes \citep{NBERw23736,NBERw27885}.
+\cite{angrist1999using} show that reducing class size (via religious ruling) induces an increase in test scores for school-age children, yet the magnitude of this effect is not as clear in the higher education setting.
+\cite{bandiera2010heterogeneous} show that UK university students perform worse academically in particularly large classes.
+Regarding faculty composition, the literature is not decided on whether contingent lecturers lead to better (or worse) education outcomes, relative to tenured faculty.
+\cite{bettinger2010does,figlio2015tenure} find US students have better enrolment and learning outcomes from courses taught by adjunct professors, while \cite{ehrenberg2005tenured} see that universities who use a greater share of lecturers have lower graduation rates.
+\cite{zhu2021limited} note that adjuncts have lower student outcomes than full-time professors, but this effect disappears among individual lecturers when they gain full-time position. 
+My approach focuses on count of professors per student at universities, and considers this a possible mechanism through which stagnating state support for higher education negatively affects student outcomes \citep{NBERw23736,NBERw27885}.

### 2023-03-17 text/sections/02-data.tex
Half-way through a major revision.  Needs an intro section on sorting (see Hoxby+ 2008), and revisions of sections 4 (results) onwards.

@@
 \subsection{Data Description}
-The data used in this analysis come from two primary sources: \citet[IPEDS]{ipeds} regarding institutional information on finances and enrolment, and \citet[IBHED]{ibhed} for data on every professor in the Illinois public university system.
+The data used in this analysis come from two primary sources: \citet[IPEDS]{ipeds} regarding institutional information including finances and enrolment, and \citet[IBHED]{ibhed} for data on every professor in the Illinois public university system.
 
@@
     The years 1987-1989 are represented in these data in an incompletion fashion, so I focus on the years 1990 onwards.
-    Year refers to the calendar year of the spring term --- i.e. 1987 refers to the academic year that ran August 1986 to July 1987.
+    Year refers to the calendar year of the spring term --- i.e., 1987 refers to the academic year that ran August 1986 to July 1987.
 }
-and provide information on the total revenues a university received from every source (including state governments), enrolment number, plus faculty count and total expenditures on salaries.\footnote{
-    I combine the Urban Institute's 2018 compilation of IPEDS data for the years 1990-2017, and manually combine with raw NCES data regarding the year 2018-2021 for all relevant variables.
-    Figures for enrolment, faculty counts and salaries come from the raw NCES version of IPEDS for all years to address inconsistencies in the Urban Institute's data formulation for these variables.
+and provide information on the total revenues a university received from every source (including state governments), enrolment, plus faculty count and total expenditures on salaries.\footnote{
+    I combine the Urban Institute's 2018 compilation of IPEDS data for the years 1990--2017, and manually combine raw NCES data on 2018-2021 for all relevant variables.
+    Figures for enrolment, faculty counts and salaries come from the raw NCES version of IPEDS for all years, addressing inconsistencies in the Urban Institute's data formulation for these variables.
 }
 I restrict analysis to public, four-year, degree-granting institutions, as these institutions adhere to a standardised concept of faculty profile, where tenure and title of appointment (lecturer, assistant professor, etc.) are relatively standardised.
-For-profit institutions employ and enrol a negligible share of professors and students respectively, while students at two-year institutions by majority intend to eventually enrol at a four-year institution \citep{mountjoy2022},\footnote{
-    Furthermore, \cite{mountjoy2022} documents the effects of expanding two-year higher education access in the US separately for students who would have and those who would not have otherwise have attended college.
-    The high rate of eventual enrolment at four-year institutions among two-year students explains the negative impact of increased access to two-year institutions on students who are diverted away from four-year institutions.
-}
-so that these institutions are not considered.
+For-profit institutions employ and enrol a negligible share of professors and students respectively, while students at two-year institutions by majority intend to eventually enrol at a four-year institution \citep{mountjoy2022}, so that these institutions are not considered.
 
@@
 \label{sec:trends}
+
+The institutional environment for how states appropriate funding allows for wide variation in state funding for higher education, both between states and across years in each state.
 A state government plans an annual budget a couple of years ahead of the fiscal year, and the legislature chooses to approve or reject a budget request put forth by the governor's office.\footnote{
@@
 This process leads to yearly variation in state appropriations not seen in other revenues sources (such as federal government appropriations), since US states differ in their support for higher education, and are subject to fiscal constraints.
-State appropriations to higher education are a particularly attractive area of state spending to absorb such shocks to state finances \citep{delaney2011state}.\footnote{
-    \cite{delaney2011state} fully describes the financial environment of state expenditures, and what makes spending on higher education an attractive area for state governments to expand funding during years of higher tax revenues, and retract funding in leaner years.
-    An analysis of state expenditures for the years 1980-2004 (overlapping with the sample for this analysis) provides solid evidence for these trends.
+State appropriations to higher education are a particularly attractive area of state spending to absorb negative shocks to state finances \citep{delaney2011state}.\footnote{
+    \cite{delaney2011state} fully describe the financial environment of state expenditures, and what makes spending on higher education an attractive area for state governments to expand funding during years of higher tax revenues, and retract funding in leaner years.
+    An analysis of state expenditures for the years 1980-2004 (overlapping with the sample for this analysis) provides solid evidence for these trends, and \autoref{fig:funding} observes these same trends.
 }
@@
 
-\autoref{fig:funding} exhibits the trends for the mean public university for the years 1990-2021.
+Public university revenues on average stagnated across there years 1990--2021.
+\autoref{fig:funding} show the trends in revenues for the mean public university for the years 1990--2021.
 We see a rise in total revenues received by public universities (from all sources), and a notable increase in mean tuition revenues from \$48 million per year-university to \$150 million per year-university.
-At the same time, total state appropriations stagnated at around \$100 million per year-university for 1990-2008, falling around 2008 and have not recovered ever since.
+At the same time, total state appropriations stagnated at around \$100 million per year-university for 1990--2008, falling around 2008 and have not recovered ever since.
 While public universities experienced a stagnation in state support, private universities were not exposed to the same constraints, receiving \$37,000 per student in 1990 and \$49,000 in 2021, experiencing no corresponding decline in any specific component.
@@
 
-\autoref{fig:fte-perprof} documents the divergence in faculty composition (per student) between the average private and public university 1990-2021.
-Private universities start with  a higher baseline of around 4.5 professors per 100 students, and exhibited yearly variation of less than 0.5 professor per hundred students over the thirty years.
-Public universities start with 3.9 professors per hundred students, and this number falls to around 3.5 primarily in the 2008-2011 time period.
+There are large differences in the average number of professors per student between the private and public sector.
+\autoref{fig:fte-perprof} shows that private universities start with  a higher baseline of around 4.5 professors per 100 students, and exhibited yearly variation of less than 0.5 professor per hundred students over the thirty years.
+Public universities start with 3.9 professors per hundred students, and this number falls to around 3., with the largest fall in the 2008-2011 time period.
 We see a similar difference in baseline, and fall for the years 2008-2011 for assistant professors.
 Private and public universities have similar numbers of associated and full professors before the year 2000, yet this number has fallen by over 20\% in the next 20 years only for public universities: in 2021 the mean public university has 6 fewer full professors per hundred students than the mean private university.
-Over the same time period, we see the rise in use of non-tenure track instructor positions (referred to as lecturers from here on), who were employed at similar rates in both sectors in 1990 yet have been utilised by public universities at a higher rate since.
+Over the same time period, we see the rise in use of non-tenure track instructor positions (referred to as lecturers from here on), who were employed at similar rates in both sectors in 1990, yet have been utilised by public universities at a higher rate since.
 
@@
 
-The state of Illinois experienced the same stagnation in state support for higher education that the rest of the nation has experienced 1990-2021.
+The rate that Illinois funds higher education has stagnated to a similar extent as the average for the entire country.
 \autoref{fig:illinois-funding} shows the trends for revenues among the seven Illinois public university campuses, including the falling share of state appropriations and substitution towards tuition revenue.
@@
 
-The decade of 2010-2021 has also seen yearly variation in its revenues, particularly around 2016.
+The decade of 2010--2021 has also seen yearly variation in its revenues, particularly around 2016.
 In the calendar year 2015 partisan disagreements between the democratic legislature and republican governor led to the 2016 fiscal year starting with no state budget.
 State agencies, and higher education institutions, employed accounting techniques to continue operating without any resources provided by the state government.\footnote{
-    Fiscal year 2016 refers to June 205 to June 2016, as is the same for the academic year definition in this analysis. 
-    K-12 schools were unaffected thanks to a separate budget agreement allowing non-higher education to operate as before.
+    Fiscal year 2016 refers to June 205 to June 2016, so is the same as the academic year definition.
 }
-While state institutions were able to stay open, there were drastic revenue and spending cuts in response to the budget impasse, as it continued through fiscal year 2017, and ended with a new budget restoring revenues to state agencies and universities for the 2018 fiscal year.
-This variation in public university revenues via the state appropriations channel, stemming entirely from political disagreements --- not from state decisions regarding higher education and its finances \citep{young2020squandered} --- mean that Illinois public universities exhibit sizable changes in their state appropriations over 2010-2021, of similar order to those for the rest of the country over 1990-2021.
+While public universities were able to stay open, there were drastic revenue and spending cuts in response to the budget impasse, as it continued through fiscal year 2017, and ended with a new budget restoring funding to state institutions for 2018.
+This variation in public university revenues via the state appropriations channel, stemming entirely from political disagreements --- not from state decisions regarding higher education and its finances \citep{young2020squandered} --- mean that Illinois public universities exhibit sizable changes in their state appropriations over 2010-2021, of similar order to those for the rest of the country over 1990--2021.

### 2023-03-17 text/sections/03-concepts.tex
Half-way through a major revision.  Needs an intro section on sorting (see Hoxby+ 2008), and revisions of sections 4 (results) onwards.

@@
 \label{sec:conceptual}
-%
-%\textbf{Follow Prof Riehl's comments.
-%    Also consider a theory toy-model where reach and teaching are plasuible substitutes, following \cite[Sec.~6.2]{NBERc13879}.
-%    Better motivated if research spending is affected.
-%}
-%The, revenue shock induced, shift away from tenured profs may be mechanically relate to the shift between departments (towards STEM).
-%Find a way to mix the “non-replacement of profs” with that for “department composition shift.”
-%
-%\subsection{University Responses to Financial Shocks}
-%\label{sec:responses}
 
@@
 
-Universities every year hire a number of new professors, either to expand their departments or to replace leaving professors.
-The number of professors hired is usually highest among non-tenured adjunct faculty, as these instructors are by majority hired on short-term contingent contracts, so that they are less costly to hire (or to not reinstate) in response to yearly changes in teaching needs.
-Secondly, tenure-track assistant professors are hired in most years at public universities to replace leaving assistant professors, and are most often hired with a four to six year contract and agreement to be considered for tenure at the end of this term.
-Lastly, tenured professors are those that have undergone review and have successfully secured a full-time appointment at their university with no expiration date; tenured professors may be hired from outside a university, though this position has the lowest hiring rate among most universities.
+Most universities hire a number of new professors every year, either to expand their departments or to replace leaving professors.
+The number of professors hired in each year is usually highest among non-tenured adjunct faculty, as these instructors are mostly hired on short-term contingent contracts, so that they are less costly to hire (or to not reinstate) in response to yearly changes in teaching needs.
+Secondly, tenure-track assistant professors are hired to replace leaving assistant professors, and are most often hired with a four to six year contract and agreement for tenure consideration.
+Lastly, tenured professors have undergone review and have successfully secured a full-time appointment at their university with no expiration date; tenured professors may be hired from outside a university, though this position has the lowest yearly hiring rate among most universities.
 It is common for universities to restrict their faculty hiring in the face of financial shocks.
-Since the yearly hiring rate is highest among lecturers and assistant professors, financial shocks may lower the number of lecturer and assistant professors at public universities the most, if faculty hiring is affected by falls in state funding.
+Since the yearly hiring rate is highest among lecturers and assistant professors, financial shocks may lower the number of lecturer and assistant professors at public universities the most, if faculty hiring is affected by falls in funding.
 
-Additionally, professors' salaries may be disrupted by the stagnation in public university finances.
+Additionally, faculty salaries may by affected by changes in public university finances.
 When the university has a lower budget for its yearly hiring, it may respond by lowering the salary they offer to new hires.
-This possible effect may not be the same across each position of professor; tenured faculty are often hired away from another university, so that new tenured professor hires could be less likely to accept a lower offer from a public university, and we would see that new tenured hires are not affected salary-wise.\footnote{
-    \cite{blackaby2005} describes the academic outside options among economist faculty, documenting how they differ for academic position and gender.
+This possible effect may not be the same across each position of professor; tenured faculty are often hired away from another university, so that new tenured professor hires may be less likely to accept a lower salary offer from a public university.\footnote{
+    Faculty often consider their outside options, as in offers for employment and/or promotion from other universities.
+    See \cite{blackaby2005} for an overview of faculty outside options.
 }
 Yet salary for all the professors, not just new hires, may also be affected: multiple universities passed a university-wide pay-cut for their faculty in response to state budget cuts around the 2008 recession, for example.\footnote{
-    Indeed, Cornell University implemented both hiring freezes and even a nominal salary reduction for professors, in anticipation of a financial shock in early 2021.
-    The salary cut was not permanent, as the oncoming financial shock turned out to not be as serious as projected, so that the salary cut was returned to professors in the year 2021.
+    Indeed, Cornell University implemented hiring freezes and a nominal salary reduction for professors, in anticipation of a financial shock in early 2021.
+    The salary cut was not permanent, as the oncoming financial shock turned out to not be as serious as projected, so that the cut was returned to professors in the year 2021.
 }
-Faculty are not paid the same by position, so that their are financial consequences for faculty being promoted between positions within a university.
-In response to tightening fiscal restraints, universities may be less likely to grant tenure to assistant professors, or to promote tenured professors from associate to full professor either.
+Faculty are not paid the same by position, so that there are financial consequences for a university promoting their own faculty.
+In response to tightening fiscal restraints, universities may be less likely to grant tenure to assistant professors (as tenured professors generally receive a salary raise and long term contract), or to promote tenured professors from associate to full professor.
 
-It is not immediately clear which effect will dominate, and which position of professor will be effected the most by changes in their university's finances.
-Yet, there is one empirical fact worth noting: a mean assistant or tenured professor earns more than double that of an adjunct lecturer per year (\autoref{tab:illinois-summary}), and the tenure contract means universities are compelled to pay that higher salary for more years than contingent lecturers.
+It is not immediately clear which effects will dominate, or whether tenured professors vs lecturers will be more affected by changes in their university's finances.
+Yet, there is one empirical fact worth noting: a mean assistant or tenured professor earns more than double that of an adjunct lecturer per year (see \autoref{tab:illinois-summary}), and the tenure contract means universities must pay a higher salary for a guaranteed number years, more so than for contingent lecturers.
 If a public university's primary obligation is to teach, and they must fulfil this objective with fewer and fewer resources, then they may substitute away from tenure-track and tenured professor towards contingent lecturers to achieve this obligation.\footnote{
-    That is substituted away from tenure-track and tenured professors according to the salary ratio, and the relative productivity between these two positions.
-    It is not possible to estimate the relative productivity in this paper, so I leave the rate of substitution between the two positions to further research.
+    That is, substituted away from tenure-track and tenured professors according to the salary ratio, and the relative productivity between these two positions.
 }

### 2023-03-17 text/sections/04-empirical.tex
Half-way through a major revision.  Needs an intro section on sorting (see Hoxby+ 2008), and revisions of sections 4 (results) onwards.

@@
 
-A professor is employed by a university, so it stands to reason that when public universities experience a secular fall in one of their revenue sources revenue --- analogous to a demand shock in the private sector --- that the composition of their employees will be affected.
+Universities employ professors, so it stands to reason that if a university experiences a fall in one of their revenue sources (similar to a demand shock in the private sector) that they may respond by making adjustments with respect to their labour force of professors.
 
 Na\"ively we can analyse the relationship between professor-outcomes $Y_{i,t}$ at university $i$ in year $t$ as a result of state funding per student $X_{i,t}$.
-Outcomes include count of professors per student within each university-year.
-This model gives $\beta$ the association with these outcomes and state funding per student,\footnote{
-    Note that dividing by student count also implicitly controls for the size of the university, so that this model accounts for yearly variation in professor count and university revenues arising from growth in a university.
+This analysis focuses on the outcome of professor count per student within each university-year.
+So that $\beta$ represents the association with these outcomes and state funding per student,\footnote{
+    Note that dividing by student count also implicitly controls for the size of the university, so that this model implicitly accounts for yearly variation in professor count and university revenues arising from growth in a university.
 }
@@
 
-\cite{NBERw23736,chakrabarti2018effect,NBERw27885} address endogeneity in public university finances by exploiting a shift-share instrument for changes in state-level funding interacted with university reliance on state funding in a base period.
+This analysis uses a shift-share instrument to address endogeneity concerns for the amount of funding a state gives to each of its public universities.
+\cite{NBERw23736,NBERw27885} develop the instrument for public university finances by exploiting a shift-share instrument for changes in state-level funding interacted with university reliance on state funding in a base period.
 \begin{align}
@@
 $Z_{i,t}$ is constructed as negative to reflect the fact that the long term trend in, and most of the short-run shocks to, state funding for higher education has been negative.
-State funding has been been falling, so that the instrument describes shocks to university revenues, mostly in a negative direction.
-
-\cite{NBERw27885} notes the tendency for public universities to respond to state funding cuts by increasing reliance on tuition, where \cite{NBERw23736} specifically instruments for tuition revenues with collected information on legislative tuition price controls.\footnote{
+State funding has been falling, so that the instrument describes shocks to university revenues, mostly in a negative direction.\footnote{
     \label{foot:control}
-    It may be argued that tuition revenues are confounder between the causal effect of changes in state funding on a university's total revenues, so that this analysis focuses on state support for higher education (no total revenue) as a result.
-    This follows the \cite{NBERw27885} formulation.
+    \cite{NBERw27885} note the tendency for public universities to respond to state funding cuts by increasing reliance on tuition, where \cite{NBERw23736} specifically instruments for tuition revenues with collected information on legislative tuition price controls.
+    It may be argued that tuition revenues are confounder between the causal effect of changes in state funding on a university's total revenues, so that this analysis focuses on state support for higher education (not total revenue), as do \cite{NBERw27885}.
     On the other hand, rises in tuition revenues (per student) may arise as result of tuition hikes thanks declining state support, which would mean controlling for tuition would constitute a bad control.
-    Estimates including tuition revenues (per student) as a control in the second stage of the IV estimates produces results of very similar magnitude and direction, and so are omitted.
+    Nonetheless, estimates including tuition revenues (per student) as a control in the second stage of the IV estimates produces results of very similar magnitude and direction, and so are omitted.
 }
@@
 This assumption is plausible given that the majority of states have multiple (i.e., more than five) public universities, without any single university campus receiving the majority of state funding within any single state.
-Secondly, addressing endogeneity in university revenues by the shift-share identification strategy requires exogeneity in either the base-line share or shift component of the instrument.
+Secondly, the shift-share identification strategy requires exogeneity in either the base-line share or shift component of the instrument.
 In this case, we satisfy the second: universities' institutional-level decisions are not correlated with contemporaneous or upcoming shocks to state appropriations.\footnote{
@@
 }
+Lastly, the IV model assumes that state funding shocks affect faculty outcomes only via affecting university finances.
 
@@
 
+The shift-share instrument performs excellently as an instrument for universities' yearly state funding.
 \autoref{tab:firststage-reg} presents results of the first-stage regression, separately with and without a control for tuition revenue per student, plus institution and year fixed effects.\footnote{
@@
 The primary empirical model combines the instrumental variable for state appropriation shocks with the empirical model for the effects of state funding at the university-level --- i.e., parameter $\beta$ in the following.\footnote{
-    It is important to note the treatment effect isolated here; the instrumental variables approach identifies the local average treatment effects (LATE) specific to the instrument.
-    So we interpret this treatment effect as a university's response in employment count and average salaries to state funding changes, changes specific to state appropriation shocks, among the complier group --- i.e., universities with any exposure to shocks, assuming no universities increase employment or salaries in response to negative revenue shocks.
-    \cite{mogstad2021causal} provide a full discussion of the LATE with instruments and multiple monotonicity conditions.
+    It is important to note the treatment effect isolated here; the instrumental variables approach identifies the local average treatment effects, one specific to the instrument.
+    So we interpret this treatment effect as a university's response in employment count and average salaries to state funding changes, changes specific to state appropriation shocks, among the complier group --- i.e., universities who respond to funding shocks who would not have made faculty-outcome changes absent the funding shock.
+    Also, we assume that no universities' total revenues increase in response to negative state funding shock (monotonicity).
 }
@@
 Additionally, I estimate the model via local projections \citep{jorda2005,miller2022} to investigate whether the effects of revenues on faculty composition linger for multiple years after the original appropriation shock.
-Regarding outcomes, I focus on the composition of the professors employed at the university by analysing (log) count per student, and average salaries paid to, professors employed by the university.
+Regarding outcomes, I focus on the composition of the professors employed at the university by analysing (log) count per student of, and average salaries paid to, professors employed by the university.
 
-\begin{figure}[H]
-    \centering
-    \singlespacing
-    \caption{Instrument Variables Model for University Finances and Faculty Composition.}
-    \begin{tikzpicture}
-        \node[state] (instrument) at (0,0) {$Z$};
-        \node at (-2.6,0.5){State};
-        \node at (-1.9,0){appropriation};
-        \node at (-2.6,-0.5){shock};
-        \node[state] (endogenous) [right=of instrument] {$X$};
-        \node at (2,1.5) {University};
-        \node at (2,1) {revenues};
-        \node[state] (outcome) [right=of endogenous] {$Y$};
-        \node at (5.2,0.25) {Faculty};
-        \node at (5.6,-0.25) {Composition};
-        % Relevance condition
-        \path (instrument) edge (endogenous);
-        \path (endogenous) edge (outcome);
-        % Exclusion Restriction
-        \path (instrument) edge[bend right=60] (outcome);
-        \node[color=red] at (2,-1.2) {X};
-    \end{tikzpicture}
-    \label{fig:SCM-ivmodel}
-\end{figure}
-
-\autoref{fig:SCM-ivmodel} presents the system in graphical form, for the causal effect of state appropriation shocks on university revenues and thus faculty composition, making clear the assumption that appropriation shocks affect faculty composition only via affecting university finances.
+%\begin{figure}[H]
+%    \centering
+%    \singlespacing
+%    \caption{Instrument Variables Model for University Finances and Faculty Composition.}
+%    \begin{tikzpicture}
+%        \node[state] (instrument) at (0,0) {$Z$};
+%        \node at (-2.6,0.5){State};
+%        \node at (-1.9,0){appropriation};
+%        \node at (-2.6,-0.5){shock};
+%        \node[state] (endogenous) [right=of instrument] {$X$};
+%        \node at (2,1.5) {University};
+%        \node at (2,1) {revenues};
+%        \node[state] (outcome) [right=of endogenous] {$Y$};
+%        \node at (5.2,0.25) {Faculty};
+%        \node at (5.6,-0.25) {Composition};
+%        % Relevance condition
+%        \path (instrument) edge (endogenous);
+%        \path (endogenous) edge (outcome);
+%        % Exclusion Restriction
+%        \path (instrument) edge[bend right=60] (outcome);
+%        \node[color=red] at (2,-1.2) {X};
+%    \end{tikzpicture}
+%    \label{fig:SCM-ivmodel}
+%\end{figure}
 
@@
 
-This analysis additionally uses data regarding individual professors in the Illinois university system, to investigate the effects of changes in university revenues on the individual professors at the universities.
+This analysis additionally uses data on individual professors in the Illinois university system, to investigate the effects of changes in university revenues on the individual professors at the universities.
 Redefining the level of outcome requires adjustment to the empirical approach, leveraging variation in university finances for the years after a professor joins the university.\footnote{
     This formulation follows that presented by \cite{NBERw27885}, where individual student outcomes are analysed via variation in state appropriations after their freshman-year.
-    This contrasts with \autoref{sec:iv-model-uni} and \cite{NBERw23736}, where the unit of analysis is the university-year.
+    This contrasts with \autoref{sec:iv-model-uni} and \cite{NBERw23736}, where the unit of analysis is the university-year, where base year 1990 is more appropriate.
 }
@@
 $j$ indexes each professor in year $t$, $\tau(j)$ for the year the professor first joins their institution.\footnote{
-    Identifying $\tau(j)$ is possible for $j$ by restricting to all professors hired 2011-2021 --- i.e. in the years after the start of the full panel.
+    Identifying $\tau(j)$ is possible for $j$ by restricting to all professors hired 2011-2021 --- i.e., in the years after the start of the full panel.
     It is not possible to discern the hiring year for professors who  were hired in the years preceding 2011, and so the entire sample is only possible to analyse using the base-share in years 1990-1993 formulation (e.g., \autoref{tab:facultysalaries-shock-illinois}, \ref{tab:promotion-shock-illinois}).
@@
 }
-We satisfy the assumptions for exogeneity by noting that none of the Illinois public campuses take the majority of state appropriations, and that the instrument identification strategy relies on exogeneity in changes in state appropriations to individual professor-outcomes, following the year they joined the university.\footnote{
+We satisfy the assumptions for exogeneity by noting that none of the Illinois public campuses take the majority of state appropriations, and that the identification strategy relies on exogeneity in changes in state appropriations to individual professor-outcomes, following the year they joined the university.\footnote{
     Additionally, within-institution changes resulting from share reliance on state funding may be correlated with unobserved changes in the outcomes, so that \cite{NBERw27885} note the importance of controlling for the base share and state student population.
-    The formulation here implicitly controls for these factors via the fixed effects; results are relatively similar while including these controls without including fixed effects, and so are omitted.
+    The formulation here implicitly controls for these factors via the fixed effects; results are relatively similar while including these controls with and without including fixed effects, and so are omitted.
 }
-\autoref{tab:firststage-illinois} presents results of the first stage estimation, showing that the instrument is strong in the same way as that for the university-level outcomes (\autoref{tab:firststage-reg}), with very similar estimates for the association between appropriation shocks and non-institutional revenues.
+\autoref{tab:firststage-illinois} presents results of the first stage estimation, showing that the instrument is strong in the same way as that for the university-level outcomes (\autoref{tab:firststage-reg}), with very similar estimates for the association between appropriation shocks and state funding.
 
-The instrumental variables model is then defined as follows where $i(j)$ refers to the institution that professor $j$ is employed at, and $Y_{j,t}$ refers to individual-level outcomes total salary, rate of promotion, and propensity to leave the Illinois public university system.
-The system includes a fixed effect for the institution and first year of employment.\footnote{
+The instrumental variables model is then defined as follows, where $i(j)$ refers to the institution that professor $j$ is employed at, and $Y_{j,t}$ to individual-level outcomes total salary, rate of promotion, and propensity to leave the Illinois public university system.
+The system includes fixed effects for the institution and first year of employment.\footnote{
     The instrument varies by institution, based in the year of first employment, so that these are the corresponding fixed effects and levels of clustered standard errors.

### 2023-03-17 text/sections/07-conclusion.tex
Half-way through a major revision.  Needs an intro section on sorting (see Hoxby+ 2008), and revisions of sections 4 (results) onwards.

@@
 It is natural to expect that such headwinds will lead to systematic change at public universities, changes which are not neutral to the composition of faculty, or to their goals of research and education.
-Thes results show large changes in faculty composition, and that stagnation in state support explains at least a third of the observed shift away from tenured professors and towards contingent lecturers.
+These results show large changes in faculty composition, and that stagnation in state support explains at least a third of the observed shift away from tenured professors and towards contingent lecturers.
 At the same time, private universities were not exposed to financial headwinds of the same magnitude or persistence.
 If public universities' hiring of faculty is disrupted by a fall in state funding, then we can expect that private universities will benefit (in a competitive sense) by being able to hire better professors from relatively resource constrained public universities.
-While public universities continue to educate the majority of higher education students in the US, we should worry about the effects that restricting their financial support has on the composition of faculty, and  higher education itself, in this country.
+While public universities continue to educate the majority of higher education students in the US, we should worry about the effects that restricting their financial support has on the composition of faculty, and the impact on higher education as a whole.

### 2023-03-19 text/paper.tex
Finished major revision.  Needs more checks of writing style.

@@
     }
-\date{\today\thanks{
+\date{This version: \today\thanks{
     I thank Professor Evan Riehl for guidance on this project, Michael Lovenheim, Ronald Ehrenberg, Douglas Miller, Francine Blau, Levon Barseghyan, and Ryan Dycus for various comments on the project's direction, as well as seminar participants at Cornell University (Autumn 2022) for helpful discussion.
-    This project's Github repository, with materials and (publicly available) data for replication, is available at 
+    This project's Github repository, with materials and data (where publicly available) for replication, is available at 
     \url{https://github.com/shoganhennessy/state-faculty-composition}.
@@
 \begin{document}
-\clearpage
 \maketitle

### 2023-03-19 text/sections/01-introduction.tex
Finished major revision.  Needs more checks of writing style.

@@
 Public universities educate the majority of higher education students in the US, yet have experienced a secular decline in state funding (per student) the last three decades.
-This decline has been shown to lead to worse education and later-life outcomes for students \citep{NBERw23736,NBERw27885}, yet it is not clear what mechanisms drive these effects, or how public universities are affected as institutions.
+This decline has been shown to lead to worse education and later-life outcomes for students, yet it is not clear what mechanisms drive these effects, or how public universities are affected as institutions.
 I show that four-year, degree-granting public universities have experienced falling state funding per student, and this affected their composition of faculty.
@@
 
-I use a variant of the \cite{NBERw23736,NBERw27885} shift-share instrument for state funding, which exploits yearly shocks to the amount an entire state funds higher education to identify the change in a single public unversity's funding.
-State appropriations for public universities are plausibly endogenous to many outcomes: state governments decide on yearly budgets for their higher education sector, and this process can be influenced by local financial conditions, or even public perceptions of the state's higher education system.
-The shift-share instrument interacts reliance on state funding in a base period with the yearly total appropriations per student in the entire state, to exploit both changes in support for higher education and how much each university relies on that support.
+To identify changes in state funding for higher education, I employ a shift-share instrument for state funding of public universities, which exploits reliance on state funding and yearly shocks to the amount an entire state funds higher education \citep{NBERw23736,NBERw27885}.
+State appropriations for public universities are plausibly endogenous to many outcomes: state governments decide on yearly budgets for their higher education sector, and this process can be influenced by local financial conditions, or changing state priorities.
+The shift-share instrument interacts reliance on state funding in a base period with the yearly total appropriations per student in the entire state, to exploit both changes in support for higher education and how much each university historically has relied on that support.
 
@@
 
-WRITE HERE ABOUT HOW IT IS NOVEL THAT THESE EFFECTS ARE CAUSAL, AND NOT JUST CONCURRENT TRENDS, IN COMPARISON TO HOXBY 2008 SORTING.
+Absent this analysis, it is not immediately clear why stagnating state support and changes in public university faculty composition would be causally linked.
+Over the 21st Century there have been drastic changes to college selectivity, where the top universities have become more selective while the average university less selective \citep{hoxby2009changing}.
+While the average public university is becoming less selective, it is possible that their most productive research-focused faculty more often move to more selective and prestigious (and often private) institutions, so that over time public universities (on average) substitute towards contingent lecturers.
+This is one way in which the trends may be concurrent, but not causal.
+This analysis shows that stagnating in state funding for public universities \textit{cause} the changes in faculty composition.
+The state funding channel causes about 40\% of the observed substitution towards contingent lecturers, so that it is possible that other systematic changes in higher education (such as changing selectivity) may explain the rest of the variation.
 
 A number of studies have shown that stagnating state funding for higher education negatively impacts student outcomes.
-\cite{NBERw23736} show that increases in public university spending (via appropriation shocks) increases enrolment and degree completion among students,\footnote{
+\cite{NBERw23736} show that increases in public university spending (via funding shocks) increases enrolment and degree completion among students,\footnote{
     \cite{miller2022making} further analyse the effects of falls in university revenues, by finding that reductions in public universities tuition prices (suggestively motivated by falling state funding) leads to reductions in provided financial aid.
 }
-and \cite{bound2019public} that these shocks induce public universities to rley more on tuition revenue.
+that high school students exposed to similar public school funding shocks are less likely to attend college \citep{jackson2021school}, and that these shocks induce public universities to rely more on tuition as a source of revenue \citep{bound2019public}.
 \cite{NBERw27885} use the same methodology to show that increases in state appropriations lead to degree lower completion time\footnote{
-    \cite{bound2007cohort} use similar variation in state funding per student to show that lower funding leads to lower higher education completion rate at public universities.
+    \cite{bound2007cohort} use similar variation in state funding per student to show that lower funding leads to lower higher education completion rate at public universities, 
 } and later-life debt for students.
 
-Faculty is a core component of the US higher education and research-innovation university system, yet their composition within universities, and its relationship with higher education finances has so far not been studied.
-\cite{brown2014endowment} presents the closest example by studying how university endowments react to negative endowment shocks, and \cite{turner2014impact} for the 2008 recession, while \cite{abe2015implications,johnson2009jep,NBERc13879} model university decision-making and faculty hiring.
-Lastly, it has been shown that instruction costs vary substantially across subject areas, primarily thanks to department differences in class size and faculty salaries \citep{hemelt2021math}.
+Faculty are a core component of the US higher education and research-innovation university system, yet their composition within universities, and its relationship with higher education finances has so far not been studied.
+The closest previous examples study  how university endowments react to negative endowment shocks \citep{brown2014endowment} and the 2008 recession \citep{turner2014impact}, and multiple theoretically model university decision-making and faculty hiring (see e.g., \citealt{abe2015implications,johnson2009jep,NBERc13879}).
+Lastly, instruction costs within individual universities vary substantially across subject areas, primarily thanks to department differences in class size and faculty salaries \citep{hemelt2021math}.
 
 The number of professors, both in absolute terms and relative to number of students, may be particularly important for quality of instruction and thus educational outcomes.
-\cite{angrist1999using} show that reducing class size (via religious ruling) induces an increase in test scores for school-age children, yet the magnitude of this effect is not as clear in the higher education setting.
-\cite{bandiera2010heterogeneous} show that UK university students perform worse academically in particularly large classes.
-Regarding faculty composition, the literature is not decided on whether contingent lecturers lead to better (or worse) education outcomes, relative to tenured faculty.
-\cite{bettinger2010does,figlio2015tenure} find US students have better enrolment and learning outcomes from courses taught by adjunct professors, while \cite{ehrenberg2005tenured} see that universities who use a greater share of lecturers have lower graduation rates.
-\cite{zhu2021limited} note that adjuncts have lower student outcomes than full-time professors, but this effect disappears among individual lecturers when they gain full-time position. 
-My approach focuses on count of professors per student at universities, and considers this a possible mechanism through which stagnating state support for higher education negatively affects student outcomes \citep{NBERw23736,NBERw27885}.
+Small class sizes class size at high school lead to increases in test scores \citep{angrist1999using}, and UK university students perform worse academically in particularly large classes \citep{bandiera2010heterogeneous}.
+Regarding faculty composition, the literature is not decided on whether contingent lecturers lead to better or worse education outcomes relative to tenured faculty.
+Adjunct professors are shown to better undergraduate teachers at highly selective universities \citep{bettinger2010does,figlio2015tenure}, yet universities who use a greater share of lecturers have lower graduation rates \citep{ehrenberg2005tenured}.
+Notably, students taught by part-time adjuncts lecturers have worse student outcomes than full-time professors at a less-selective university \citep{zhu2021limited}, but this difference disappears when individual lecturers gain a full-time teaching position at the university.
+My approach focuses on count of professors per student at universities, and considers faculty count and composition change as possible mechanisms for how stagnating state support for higher education negatively impacts student outcomes \citep{NBERw23736,NBERw27885}.

### 2023-03-19 text/sections/03-concepts.tex
Finished major revision.  Needs more checks of writing style.

@@
 
-It is not immediately clear which effects will dominate, or whether tenured professors vs lecturers will be more affected by changes in their university's finances.
+It is not immediately clear which effects will dominate, or whether tenured professors vs lecturers will be affected more by changes in their university's funding.
 Yet, there is one empirical fact worth noting: a mean assistant or tenured professor earns more than double that of an adjunct lecturer per year (see \autoref{tab:illinois-summary}), and the tenure contract means universities must pay a higher salary for a guaranteed number years, more so than for contingent lecturers.

### 2023-03-19 text/sections/04-empirical.tex
Finished major revision.  Needs more checks of writing style.

@@
 
-\subsection{State Appropriation Shocks}
+\subsection{State Funding Shocks}
 \label{sec:approp-shocks}
@@
 \end{align}
-The system exploits the fact that institutions who rely on state appropriations more will be affected by state appropriation shocks.
+The system exploits the fact that institutions who rely on state funding more will be affected by state appropriation shocks.
 $Z_{i,t}$ is the instrument for (log) state appropriations for institution $i$ in year $t$, interacting the average funding for universities in state $s(i)$ with reliance on state funding relative to total revenues, averaged across the base years 1990--1993.\footnote{
@@
 }
-This analysis focuses on one important source of revenues for public universities: state appropriations.
+The instrument approach relies on the conditional independence assumption, that the instrument is independently assigned to the universities.
+While this assumption is fundamentally untestable, we see that univeresities are by-in-large smaller (in terms of enrolment, total revenues, and professor count) at the top of the state funding shock distribution.
+Though on a per student basis, there is little difference across the distribution (except in the outcomes under consideration).
+Yearly and individual fixed effects are included in regressions throughout to implicitly condition on mean differences between universities, and strengthen the claim of conditional independence of the instrument.
 
-The first-stage is then as follows, including institution and year fixed effects.
+\begin{table}[h!]
+    \singlespacing
+    %\centering
+    \caption{Mean Characteristics for Public Universities, by State Funding Shock Instrument.}
+    \makebox[\textwidth][c]{\input{tables/summary-quantiles.tex}}
+    \label{tab:summary-quantiles}
+    \footnotesize
+    \textbf{Note}:
+    The column labelled ``1st'' refers to the mean for all university-year observations in the first quintile (bottom 20\%) of the funding shock (per student) distribution, and so on.
+\end{table}
+
+The state funding shock is an instrument for the amount of state funding for each university in each year.  
+The first-stage is then as follows, including institution and year fixed effects, where $X_{i,t}$ represents the amount of state funding divided by the number of full-time students attending the university.
 \begin{equation}
@@
 }
-Lastly, the IV model assumes that state funding shocks affect faculty outcomes only via affecting university finances.
+Lastly, the shift-share approach assumes that state funding shocks affect faculty outcomes only via affecting university finances.
 
@@
 Column (1) represents the estimates for \autoref{eqn:firststage} with fixed effects, omitting the tuition revenue control, and is the preferred form that I proceed with.
-
-\begin{figure}[h!]
-    \centering
-    \singlespacing
-    \caption{Local Projection Estimates for First-Stage \autoref{eqn:firststage}.}
-    \includegraphics[width=0.6\textwidth]{figures/firststage-lp.png}
-    \label{fig:firststage-lp}
-\end{figure}
-
-\autoref{fig:firststage-lp} presents local projection estimates for the staying-power of the state appropriation shock to university revenues, with each year showing the years following effect of a shock to state revenues in year zero.
-We see a decaying, yet real and positive, effect of the shock on revenues for the next 10 years, illustrating that the state appropriation shock is a strong, if fading, instrument for state funding per student and justifying its use in local projection IV estimation.
+Lastly, the state funding shock affects university revenues for multiple years after the initial shock (seee \autoref{fig:firststage-lp}).
+We see a decaying, yet real and positive, effect of the shock on revenues for the next 10 years, illustrating that the state appropriation shock is a strong, if fading, instrument for state funding per student and justifying its use in later local projections models.
 
@@
 I estimate the system by two stage least squares, including institution and year fixed effects, and investigate outcomes at the university-level to analyse effects on the university as a result of changes in revenues.
-Additionally, I estimate the model via local projections \citep{jorda2005,miller2022} to investigate whether the effects of revenues on faculty composition linger for multiple years after the original appropriation shock.
+Additionally, I estimate the model via local projections \citep{jorda2005,miller2022} to investigate whether the effects of revenues on faculty composition linger for multiple years after the original appropriation shock.\footnote{
+    The local projections method is an empirical method to estimate dynamic treatment effects when the treatment of interest is not binary, so that time-series confounding (i.e., treatment in time $t$ is correlated with treatment in time $t-1$) is present even for instrumental variable models.
+    Inference for this model is similar to that for an event study design, so I present estimates in graphical form following that example (\autoref
+    {fig:count-lp}).
+}
 Regarding outcomes, I focus on the composition of the professors employed at the university by analysing (log) count per student of, and average salaries paid to, professors employed by the university.

### 2023-03-19 text/sections/05-results.tex
Half-way through a major revision.  Needs an intro section on sorting (see Hoxby+ 2008), and revisions of sections 4 (results) onwards.
Finished major revision.  Needs more checks of writing style.

@@
 
+Changes in state funding have clear effects on the composition of faculty public universities employ.
 \autoref{tab:facultycount-shock-reg} presents OLS and IV estimates, where the outcome is (log) count of professors employed at the university, separated by position, and for all professors (Columns 7, 8).
@@
 
-\autoref{fig:all-count-lp} shows local projection estimates, where the effect on total professor count per student persists at the same magnitude of around 0.1 four years after the identified shock.
-These findings are in line with \cite{turner2014impact} observing that universities froze hiring particularly for tenure-track positions in response to negative budget shocks in the late 2000s, and \cite{brown2014endowment} from shocks to university endowments.
-We see a negative coefficient for the number of lecturers per student.
-This effect lines up with two trends we see in \autoref{sec:trends}: that public universities revenues (per student) are decreasing while utilisation of non-tenure track lecturers increased over the same time period.
-And yet, identifying the university revenues by the state appropriations shock increases the magnitude of the effect (Column 2) with respect to the association (Column 1), strengthening the case that the two trends are causally linked.
-Between years 1990-1999 IPEDS provided count of professors by explicit tenure designation (opposed to named position as in \autoref{tab:facultycount-shock-reg}), and \autoref{tab:tenurecount-shock-reg-fte} presents results of the respective regressions by tenure status.
+Faculty composition is not only impacted in the same years as a state funding shock, but also for a number years following.
+\autoref{fig:all-count-lp} shows local projection estimates, where the effect on total professor count per student persists at a magnitude of around 0.1 four years after the initial funding shock.
+These findings are in line with the observation that universities froze hiring for a couple of yearsin response to 2008 negative budget shocks, particularly for tenure-track positions \citep{turner2014impact}, and similarly from shocks to university endowments \citep{brown2014endowment}.
+For lecturers we see a negative effect, which lines up with two trends noted in \autoref{sec:trends}: public universities revenues (per student) are decreasing at the same time as employment of contingent lecturers increased rapidly.
+And yet, identifying the university revenues by the state appropriations shock increases the magnitude of the effect (Column 2) with respect to the association (Column 1), strengthening the case that the two trends are causally linked.\footnote{
+    Between years 1990-1999 IPEDS provided count of professors by whether they had tenure (opposed to named position as in \autoref{tab:facultycount-shock-reg}), and \autoref{tab:tenurecount-shock-reg-fte} presents results of the respective regressions by tenure status.
     Results are relatively similar, yet imprecise from the much smaller number of observations.
+}
 
-Together these findings show that public universities increase (decrease) their count of tenure-track and tenured professors per student in years when revenues are more (less) plentiful, possibly by increasing hiring intensity or efforts to maintain incumbent professors.
-In the same vein, when revenues are (not) bountiful, count of tenure-track and tenured professors per student increases (decreases).
-%Yet, further analysis is needed to disentangle the possible long-run effects of state revenues on faculty composition among public universities.
-
-\subsection{Long-term Effects on Faculty Composition}
-
+Decreases in university revenues (via state funding shocks) affect faculty composition for at most 4 years after the initial revenue shock.
+A 10\% decrease in state funding increases count of lecturers per student by 4\% for the first two years after the initial shock.
+While count of assistant professors decreases 10 to 18.5\% three years later;
+count of tenured professors decreases 9 to 6\% three years later.
+Together, the total count of professors per student decreases around 1\% for three years after the initial funding shock, and the effect is not distinguishable from zero four years after.
 \begin{figure}[h!]
@@
 \end{figure}
+\autoref{fig:count-lp} represents these estimates, using the local projection method, as a series of impulse responses for each outcome of professor count.
 
-Decreases in university revenues (via state funding shocks) affect faculty composition for at most 4 years after the initial revenue shock.
-A 10\% decrease in state funding increases count of lecturers per student by 4\% for the first two years after the initial shock.
-While count of assistant professors decreases 10 to 18.5\% three years later;
-count of tenured professors decreases 9 to 6\% three years later.
-Together, the total count of professors per student decreases around 1\% for three years after the initial funding shock, and the effect is not distinguishable from zero four years after.
-
-\autoref{fig:count-lp} represents these estimates, using the local projection method, as a series of impulse responses for each outcome of professor count.
+These findings show that public universities increase (decrease) their count of tenure-track and tenured professors per student in years when revenues are more (less) plentiful, possibly by increasing hiring intensity or efforts to maintain incumbent professors.
+In the same vein, when revenues are (not) bountiful, count of tenure-track and tenured professors per student increases (decreases).
 The mechanism that explains the changes in faculty composition are, however, not clear at this unit of analysis; individual data are needed to delve deeper.
@@
 \subsection{Individual Professor Outcomes}
+\label{sec:results-indiv}
 

### 2023-03-19 text/sections/06-discussion.tex
Finished major revision.  Needs more checks of writing style.

--- 2022-11-20 text/sections/06-discussion.tex
@@
 
-The results show that the recent stagnation in state funding for higher education has affected the composition of faculty within public universities, away from tenure-track and tenured professors towards non-tenured lecturers.
+The results show that the recent stagnation in state funding for higher education has affected the composition of faculty within public universities, away from tenure-track and tenured professors towards contingent lecturers.
 At the same time, there are little discernible effects on individual professors hired in the years 2011-2021 at Illinois public universities, except for the salaries of first year lecturers.
@@
 
+Between academic years the number of professors at a university changes: professors leave an institution (retiring, hired elsewhere, etc.) and other join (they are newly hired, end a sabbatical, etc.).
+\autoref{sec:results-indiv} shows that most of these channels are unaffected, implying that falls in hiring for tenured-track and tenured faculty must explain most of the substitution towards lecturers away from tenured faculty.
+Notably, faculty hiring is already know to be heavily impacted by budget shocks.
 \cite{turner2014impact} documents the wide-spread practice of hiring freezes at universities in response to budget shocks around the 2008 recession.
@@
 did not receive the allocated state appropriations from the state of Illinois on time, so enacted cost-cutting measures to stay fiscally solvent.
-The university system placed a hold on all hiring for filling state-funded positions and promotions.
-Faculty were furloughed (placed on leave without pay) for a day each month, and university administrators placed on two days per month furlough, or eligible university employees could accept a voluntary, equivalent pay reduction.
-The University of Western Illinois adopted very similar measures in response to the Illinois budget crisis in 2016-2017 \citep{wiu2016}.
+In response, the university system placed a hold on all hiring for filling state-funded positions and promotions.
 
+The data used so far measure faculty count, so cannot disentangle which of these channels is most affected by changes in state funding at a public university, yet \cite{wapman2022quantifying} provide a measure of the aggregated total faculty hires at universities over the time period 2011--2021.\footnote{
+    \citep{wapman2022quantifying} made an aggregated version of their data open-source, making this sub-analysis possible.
+}
 \begin{figure}[h!]
@@
     \singlespacing
-    \caption{Trends in New Hires at Illinois Public Universities 2011-2021.}
-    \begin{subfigure}[b]{0.495\textwidth}
-        \centering
-        \caption{New Hire Count, Total.}
-        \includegraphics[width=\textwidth]{figures/newhire-count-illinois.png}
-        \label{fig:newhire-count-illinois}
-    \end{subfigure}
-    \begin{subfigure}[b]{0.495\textwidth}
-        \centering
-        \caption{Mean Salary in First Year for New Hires, \$ 2021 CPI-U.}
-        \includegraphics[width=\textwidth]{figures/newhire-salary-illinois.png}
-        \label{fig:newhire-salary-illinois}
-    \end{subfigure}
-    \label{fig:newhire-illinois}
+    \caption{State Funding and Faculty Hired at Public Universities, Total for 2011--2021.}
+    \includegraphics[width=0.95\textwidth]{figures/hiring-correlation.png}
+    \label{fig:hiring-correlation.png}
 \end{figure}
 
-These examples show that Illinois public universities were effected by the negative shock to state funding for higher education, and instituted policy changes aimed at affecting faculty outcomes investigated in \autoref{sec:results}.
-Yet the only measurable effect was on count of professor per student (implied via slower hiring), and first-year lecturer salaries.
-To investigate these trends for the entire state, \autoref{fig:newhire-illinois} shows the count of professors hired in the years 2011-2021, and their mean salary, by position.
-In particular, no position had sustained rises in hiring while enrolment within the university was rising (so that new hires per student necessarily fell), and hiring of new administrators jumped once the Illinois budget crisis ended in 2017.
-At the same time, professor new hires' starting salaries remained relatively constant for lecturer and assistant professors, yet rose and fell over the decade for full professors and administrators.
+Universities with more state funding (for the entire decade) hired more professors (\autoref{fig:hiring-correlation.png}).
+Similarly, the funding shock instrument identifies that an increase in 10\% state funding leads to 13 more faculty hires across the decade 2011--2021, without any clear difference between the genders (\autoref{tab:hiring-shock-reg}).\footnote{
+    These results were produced by integrating the total count of faculty hires for 2010--2021 for the top-ranked 180 US universities with a sum of the funding variables, and then estimating the models specified in \autoref{sec:iv-model-uni}.
+    See \autoref{sec:appendix-hiring} for further details.
+}
+It is clear that public universities' hiring is negatively impacted by falls in state funding, even using these aggregated data for count of hiring. 
+While this analysis strengthens the case negative effects on hiring plays, it is crude for measuring exact magnitudes for the effect on faculty hiring; I leave it to further research to delve deeper into the magnitudes that hiring plays as the primary causal mediator for the effects on faculty composition of falls in state funding for public universities.

### 2023-03-19 text/sections/07-conclusion.tex
Finished major revision.  Needs more checks of writing style.

@@
 This analysis investigates how the recent stagnation in state support for higher education has affected public universities, and their faculty composition.
-The work contributes to the literature along two primary dimensions.
-Firstly, by exploiting the same empirical approach as \cite{NBERw23736,NBERw27885} to isolate the causal effect of changes in state funding on public universities themselves, the analysis provides a plausible mechanism for the observed negative effects on students outcomes that this stagnation has brought.
+This work contributes to the literature along two primary dimensions.
+Firstly, by isolating changes in state funding on public universities via state funding shocks, the analysis provides a plausible mechanism for the observed negative effects on students outcomes that this stagnation has brought \citep{NBERw23736,NBERw27885}.
 Secondly, this analysis considers individual professors as the unit of analysis using a dataset (IBHED) new in the economic literature.
@@
 The effects on incumbent professors in the state of Illinois are non-distinguishable from zero, which implies that the effect is driven by changes in university hiring of new professors.
-Public universities are hiring thousands of non-tenured lecturers to satisfy teaching obligations, and fewer tenure-track and tenured professors that their private counterparts.
+Each year, public universities are using more contingent lecturers to teach their education students, while private universities continue to employ more tenure-track and tenured professors than their public counterparts.
+Private universities have always had more tenured professor per student than public institutions, but each year the gap widens.
 
@@
 If public universities' hiring of faculty is disrupted by a fall in state funding, then we can expect that private universities will benefit (in a competitive sense) by being able to hire better professors from relatively resource constrained public universities.
-While public universities continue to educate the majority of higher education students in the US, we should worry about the effects that restricting their financial support has on the composition of faculty, and the impact on higher education as a whole.
+While public universities continue to educate the majority of higher education students in the US, we should worry about the effects that restricting their funding has on the composition of faculty, and the impact on higher education as a whole.

### 2023-03-19 text/sections/09-appendix.tex
Finished major revision.  Needs more checks of writing style.

@@
 
+\subsection{First Stage Local Projection Estimates}
+
+\begin{figure}[H]
+    \centering
+    \singlespacing
+    \caption{Local Projection Estimates for First-Stage \autoref{eqn:firststage}.}
+    \includegraphics[width=0.6\textwidth]{figures/firststage-lp.png}
+    \label{fig:firststage-lp}
+\end{figure}
+
 \subsection{First Stage Estimates, Individual Outcomes}
@@
 
-\begin{table}[h!]
+\begin{table}[H]
     \singlespacing
@@
 
-\begin{figure}[h!]
+\begin{figure}[H]
     \centering
@@
 \end{figure}
+
+\subsection{Additional Results, Faculty Hiring}
+\label{sec:appendix-hiring}
+
+These results were collects by using the open-source network data by \cite{wapman2022quantifying}, summing the count of hires a university made for 2011--2021 across the network structure of the data, and integrating a sum of all revenue related measures from IPEDS data.
+
+\begin{table}[h!]
+    \singlespacing
+    \centering
+    \caption{OLS and 2SLS Estimates for University Faculty Hiring, Total for 2011--2020.}
+    \makebox[\textwidth][c]{\input{tables/hiring-shock-reg-fte.tex}}
+    \begin{flushleft}
+        \footnotesize
+        \textbf{Note}: Standard errors are clustered at the state level.
+    \end{flushleft}
+    \label{tab:hiring-shock-reg}
+\end{table}
+
+Note that yearly variation is not observed here, so that only the aggregate level, for 180 universities, can be considered.
 

### 2023-03-19 programs/analysis/ipeds-shock.R

@@ -128,6 +128,49 @@ reg.data %>%
         type = "text",
         out = "../../text/tables/ipeds-summary-fte.tex")
 
+# Summary Statistics for the Quantiles of the shift-share instrument.
+quantile.count <- 5
+reg.data %>%
+    mutate(instrument_quantile = round((quantile.count - 1) *
+        ecdf(appropriationshock_perEnroll_real)(
+            appropriationshock_perEnroll_real))) %>%
+    group_by(instrument_quantile) %>%
+    summarise(
+        `State Funding Shock, USD per student` = round(mean(appropriationshock_perEnroll_real, an.rm = TRUE)),
+        `State Funding, per student`           = round(mean(stateappropriations_real / enrollment_reported, na.rm = TRUE)),
+        `Total Full-time Enrolment`            = round(mean(enrollment_reported, na.rme = TRUE)),
+        `State Funding, USD millions`          = round(mean(stateappropriations_real / 10^6, na.rme = TRUE)),
+        `Total Revenues, USD millions`         = round(mean(totalrevenues_real / 10^6, na.rme = TRUE)),
+        `Total Revenues, USD per student`      = round(mean(totalrevenues_real / enrollment_reported, na.rm = TRUE)),
+        `Lecturer Count`                       = round(mean(lecturer_prof_count, na.rm = TRUE)),
+        `Assistant Professor Count`            = round(mean(assistant_prof_count, na.rm = TRUE)),
+        `Full Professor Count`                 = round(mean(full_prof_count, na.rm = TRUE)),
+        `Total Professor Count`                = round(mean(all_prof_count, na.rm = TRUE)),
+        `Lecturers, per student`               = mean(lecturer_prof_count / enrollment_reported, na.rm = TRUE),
+        `Assistant Professors, per student`    = mean(assistant_prof_count / enrollment_reported, na.rm = TRUE),
+        `Full Professors, per student`         = mean(full_prof_count / enrollment_reported, na.rm = TRUE),
+        `Total Professors, per student`        = mean(all_prof_count / enrollment_reported, na.rm = TRUE)
+            ) %>%
+    ungroup() %>%
+    pivot_longer(!instrument_quantile,
+        names_to = "variable", values_to = "value") %>%
+    pivot_wider(names_from = "instrument_quantile",
+        values_from = "value") %>%
+    transmute(`Instrument Quantile:` = variable,
+        `1st` = `0`,
+        `2nd` = `1`,
[…]

### 2023-03-19 programs/run.sh

@@ -31,3 +31,4 @@ cd ..
 cd ../text
 latexmk -pdf paper.tex
 latexmk -c
+cp paper.pdf ../state-faculty-composition-2023.pdf

### 2023-03-19 text/sections/00-abstract.tex
Finished major revision.  Needs more checks of writing style.

@@
 Analysis of all the professors at Illinois public universities 2011-2021 shows that incumbent professors are not affected by the changes in state support, while tenured-track hiring counts fall over this time period, showing that these changes in faculty composition arose by reduced hiring of tenure-track and tenured professors at public universities.
-These results exhibit the long-term effects of stagnating state support for higher education, and raises questions about the direction that public education heads as these financial headwinds show no sign of dissipating.
+These results show the long-term effects of stagnating state support for higher education, and raises questions about the direction that public education heads as these financial headwinds show no sign of dissipating.
 

### 2023-03-20 text/sections/00-abstract.tex
Finished major revision.  Needs more checks of writing style.

@@
 \noindent
-State support for higher education per-student has stagnated at the same time as public universities have reduced their employment of tenure-track or tenured professors, and increased their employment of temporary contingent lecturers.
-This analysis relates these changes in the faculty composition at public universities to the fall in per-student revenues from state funding for higher education.
-A shift-share instrumental variables approach addresses endogeneity in decisions for state higher education funding by exploiting (1) yearly change in each state's funding for their public universities and (2) each institution's reliance on that support.
+State support for higher education per-student has stagnated while universities have substituted towards contingent lecturers and away from tenure-track/tenured professors.
+This analysis relates these changes in the faculty composition at public universities to the fall in state funding per student for higher education, using a shift-share approach to identify changes in a state funding among public universities.
 A decrease in state funding of 10\%, via a shock to state funding, decreases the number of assistant professors per student at a public university by 1.4\% and full professors by 1.2\%, yet increases the number of lecturers per student by 4.4\%.
 Local projection estimates show that these effects linger for up to three years after the initial funding shock.
-Analysis of all the professors at Illinois public universities 2011-2021 shows that incumbent professors are not affected by the changes in state support, while tenured-track hiring counts fall over this time period, showing that these changes in faculty composition arose by reduced hiring of tenure-track and tenured professors at public universities.
-These results show the long-term effects of stagnating state support for higher education, and raises questions about the direction that public education heads as these financial headwinds show no sign of dissipating.
+Analysis of all the professors at Illinois public universities 2011-2021 shows that incumbent professors are not affected by the changes in state funding, and suggestive evidence shows that these changes arose by impacting the hiring of new professors at public universities impacted by the funding shocks.
+These results show the long-term effects of stagnating state support for higher education, and raise questions about the direction that public education heads as these financial headwinds show no sign of dissipating.
 
+\vfill
 \noindent
@@
 
+\vspace{0.1cm}
 \noindent

### 2023-03-20 text/sections/01-introduction.tex
Finished major revision.  Needs more checks of writing style.

@@
 Over the same time period, state funding fell by around 35\%, while the count of professors per student fell by 9\%, so that these results show that falls in state funding explain around a third of the observed shift away from tenure-track and tenure professors towards contingent faculty.
-
 Additionally, I use individual level data on professors at Illinois public universities over 2010-2021 to investigate whether the shocks to state funding affected individual professors.
@@
 Over the 21st Century there have been drastic changes to college selectivity, where the top universities have become more selective while the average university less selective \citep{hoxby2009changing}.
-While the average public university is becoming less selective, it is possible that their most productive research-focused faculty more often move to more selective and prestigious (and often private) institutions, so that over time public universities (on average) substitute towards contingent lecturers.
+While the average public university is becoming less selective, it is possible that their most productive or research-focused faculty more often move to more selective and prestigious (and often private) institutions, so that over time public universities (on average) substitute towards contingent lecturers.
 This is one way in which the trends may be concurrent, but not causal.
-This analysis shows that stagnating in state funding for public universities \textit{cause} the changes in faculty composition.
-The state funding channel causes about 40\% of the observed substitution towards contingent lecturers, so that it is possible that other systematic changes in higher education (such as changing selectivity) may explain the rest of the variation.
+This analysis shows that stagnating in state funding for public universities cause the changes in faculty composition, explaining about 40\% of the observed substitution towards contingent lecturers, so that it is possible that other systematic changes in higher education (such as changing selectivity) may explain the rest of the variation.
 
@@
 The number of professors, both in absolute terms and relative to number of students, may be particularly important for quality of instruction and thus educational outcomes.
-Small class sizes class size at high school lead to increases in test scores \citep{angrist1999using}, and UK university students perform worse academically in particularly large classes \citep{bandiera2010heterogeneous}.
+Small class sizes in secondary schools lead to increases in test scores \citep{angrist1999using}, and UK university students perform worse academically in particularly large classes \citep{bandiera2010heterogeneous}.
 Regarding faculty composition, the literature is not decided on whether contingent lecturers lead to better or worse education outcomes relative to tenured faculty.

### 2023-03-20 text/sections/00-abstract.tex
Finished major revision.  Needs more checks of writing style.

@@
 This analysis relates these changes in the faculty composition at public universities to the fall in state funding per student for higher education, using a shift-share approach to identify changes in a state funding among public universities.
-A decrease in state funding of 10\%, via a shock to state funding, decreases the number of assistant professors per student at a public university by 1.4\% and full professors by 1.2\%, yet increases the number of lecturers per student by 4.4\%.
+A decrease in a university's funding of 10\%, via a shock to state higher education funding, decreases the number of assistant professors per student at a public university by 1.4\% and full professors by 1.2\%, yet increases the number of lecturers per student by 4.4\%.
 Local projection estimates show that these effects linger for up to three years after the initial funding shock.

### 2023-03-20 text/sections/06-discussion.tex
Finished major revision.  Needs more checks of writing style.

@@
     \caption{State Funding and Faculty Hired at Public Universities, Total for 2011--2021.}
-    \includegraphics[width=0.95\textwidth]{figures/hiring-correlation.png}
+    \includegraphics[width=0.8\textwidth]{figures/hiring-correlation.png}
     \label{fig:hiring-correlation.png}
@@
 Universities with more state funding (for the entire decade) hired more professors (\autoref{fig:hiring-correlation.png}).
-Similarly, the funding shock instrument identifies that an increase in 10\% state funding leads to 13 more faculty hires across the decade 2011--2021, without any clear difference between the genders (\autoref{tab:hiring-shock-reg}).\footnote{
+Similarly, the funding shock instrument identifies that an increase in 10\% state funding at an individual university leads to 13 more faculty hires on average across the decade 2011--2021 (see \autoref{tab:hiring-shock-reg}).\footnote{
     These results were produced by integrating the total count of faculty hires for 2010--2021 for the top-ranked 180 US universities with a sum of the funding variables, and then estimating the models specified in \autoref{sec:iv-model-uni}.
+    There were no observable differences in the hiring rate of ale vs female faculty.
     See \autoref{sec:appendix-hiring} for further details.
 }
-It is clear that public universities' hiring is negatively impacted by falls in state funding, even using these aggregated data for count of hiring. 
-While this analysis strengthens the case negative effects on hiring plays, it is crude for measuring exact magnitudes for the effect on faculty hiring; I leave it to further research to delve deeper into the magnitudes that hiring plays as the primary causal mediator for the effects on faculty composition of falls in state funding for public universities.
+It is clear that public universities' hiring is negatively impacted by falls in state funding, even using highly aggregated data for a total count of hiring across an entire decade.
+While this analysis strengthens the case negative effects on hiring plays, it is crude for measuring exact magnitudes for the effect on faculty hiring; I leave it to further research to delve deeper into the magnitudes that hiring plays as the primary causal mediator for the effects of falls in state funding on faculty composition at public universities.

### 2023-03-20 text/sections/00-abstract.tex
Finished major revision.  Needs more checks of writing style.

@@
 Local projection estimates show that these effects linger for up to three years after the initial funding shock.
-Analysis of all the professors at Illinois public universities 2011-2021 shows that incumbent professors are not affected by the changes in state funding, and suggestive evidence shows that these changes arose by impacting the hiring of new professors at public universities impacted by the funding shocks.
+Analysis of all the professors at Illinois public universities 2011-2021 shows that incumbent professors are not affected by the changes in state funding, and suggestive evidence shows that these changes in faculty composition arose by impacting the hiring of new professors at public universities impacted by the funding shocks.
 These results show the long-term effects of stagnating state support for higher education, and raise questions about the direction that public education heads as these financial headwinds show no sign of dissipating.

### 2023-03-30 text/paper.tex
Finished major revision.  Needs more checks of writing style.
adjusted formatting of title page only.

--- 2023-02-22 text/paper.tex
@@
 \author{Senan Hogan-Hennessy \\ \vspace{0.1cm}
-    Economics Department, Cornell University
+    Economics Department, Cornell University\thanks{
+        I thank Professor Evan Riehl for guidance on this project, Michael Lovenheim, Ronald Ehrenberg, Douglas Miller, Francine Blau, Levon Barseghyan, and Ryan Dycus for various comments on the project's direction, as well as seminar participants at Cornell University (Autumn 2022) for helpful discussion.}
 }
@@
     }
-\date{\today\thanks{
-    I thank Professor Evan Riehl for guidance on this project, Michael Lovenheim, Ronald Ehrenberg, Douglas Miller, Francine Blau, Levon Barseghyan, and Ryan Dycus for various comments on the project's direction, as well as seminar participants at Cornell University (Autumn 2022) for helpful discussion.
-    This project's Github repository, with materials and (publicly available) data for replication, is available at 
+\date{This version: \today\thanks{
+    This project's Github repository, with materials and data (where publicly available) for replication, is available at 
     \url{https://github.com/shoganhennessy/state-faculty-composition}.
@@
     } \\ \vspace{0.1cm}
-    \href{https://github.com/shoganhennessy/state-faculty-composition/blob/main/text/paper.pdf}{Newest version available here.}
+    \href{https://github.com/shoganhennessy/state-faculty-composition/blob/main/state-faculty-composition-2023.pdf}{Newest version available here.}
 }

### 2023-05-20 text/paper.tex
Adjusting writing in TeX document.  2nd revision unfinished.

@@
 % Author
-\author{Senan Hogan-Hennessy \\ \vspace{0.1cm}
-    Economics Department, Cornell University\thanks{
-        I thank Professor Evan Riehl for guidance on this project, Michael Lovenheim, Ronald Ehrenberg, Douglas Miller, Francine Blau, Levon Barseghyan, and Ryan Dycus for various comments on the project's direction, as well as seminar participants at Cornell University (Autumn 2022) for helpful discussion.}
+\author{Senan Hogan-Hennessy\thanks{
+        I thank Professor Evan Riehl for his guidance on this project, without which this work would not have possible.
+        Additionally, I thank Michael Lovenheim, Ronald Ehrenberg, Douglas Miller, Francine Blau, Levon Barseghyan, Jake Meyer, and Ryan Dycus for various comments, as well as seminar participants at Cornell University (Autumn 2022) for helpful discussion.} \\
+    \vspace{0.1cm}
+    Economics Department, Cornell University
 }
 % Title
-\title{Changes in Faculty Composition and Stagnating State Support for Higher Education
+\title{Stagnating State Funding for Higher Education and its Effect on Faculty Composition at US Universities
     }
 \date{This version: \today\thanks{
-    This project's Github repository, with materials and data (where publicly available) for replication, is available at 
+    This project's Github repository, with materials and anonymised data for replication, is available at 
     \url{https://github.com/shoganhennessy/state-faculty-composition}.
@@
 \rhead{\today}
-\lhead{Faculty Composition and State Support for Higher Education.}
+\lhead{Stagnating State Funding and its Effect on Faculty Composition.}
 % Begin
@@
 \begin{abstract}
+    \noindent
     \input{sections/00-abstract.tex}
@@
 % Introduction section
+\noindent
 \input{sections/01-introduction.tex}

### 2023-05-20 text/sections/00-abstract.tex
Adjusting writing in TeX document.  2nd revision unfinished.

--- 2023-03-20 text/sections/00-abstract.tex
@@
-\noindent
-State support for higher education per-student has stagnated while universities have substituted towards contingent lecturers and away from tenure-track/tenured professors.
-This analysis relates these changes in the faculty composition at public universities to the fall in state funding per student for higher education, using a shift-share approach to identify changes in a state funding among public universities.
-A decrease in state funding of 10\%, via a shock to state funding, decreases the number of assistant professors per student at a public university by 1.4\% and full professors by 1.2\%, yet increases the number of lecturers per student by 4.4\%.
-Local projection estimates show that these effects linger for up to three years after the initial funding shock.
-Analysis of all the professors at Illinois public universities 2011-2021 shows that incumbent professors are not affected by the changes in state funding, and suggestive evidence shows that these changes arose by impacting the hiring of new professors at public universities impacted by the funding shocks.
-These results show the long-term effects of stagnating state support for higher education, and raise questions about the direction that public education heads as these financial headwinds show no sign of dissipating.
+While states have limited their funding for higher education for the last forty years, US universities have systematically substituted away from tenure-track/tenured professors towards contingent lecturers.
+This analysis relates the changes in the faculty composition at US public universities to the fall in state funding for higher education, using a shift-share approach to isolate the causal effects of changes in state funding.
+A $-10$\% fall in a university's funding, via a state funding shock, decreases the number of assistant professors per student by $-1.4$\% and full professors by $-1.2$\%, yet increases the number of lecturers per student by 4.4\%; long-run estimates show the effects linger for multiple years after the initial funding shock.
+Analysis of all the professors at Illinois public universities 2011-2021 shows that incumbent professors are not affected by the changes in state funding, and suggestive evidence shows that these changes arose by limiting the hiring of new professors at public universities to replace leaving tenure-track/tenured professors.
+These results show the institutional effects of stagnating state funding for higher education, and raise questions about the direction that public education heads as these financial headwinds show no sign of dissipating.
 

### 2023-05-20 text/sections/01-introduction.tex
Adjusting writing in TeX document.  2nd revision unfinished.

@@
-\noindent
 Public universities educate the majority of higher education students in the US, yet have experienced a secular decline in state funding (per student) the last three decades.
-This decline has been shown to lead to worse education and later-life outcomes for students, yet it is not clear what mechanisms drive these effects, or how public universities are affected as institutions.
-I show that four-year, degree-granting public universities have experienced falling state funding per student, and this affected their composition of faculty.
-Falls in public university revenues from the state, instrumented by state-level funding shocks, lead to a fall in the number of tenure-track or tenured professors per student within a university, and an increase in the number of lecturers.
-US private universities were not exposed to such financial constraints during this same time period, and do not exhibit the substitution away from tenure-track and tenured professors, so that the stagnating state support for public universities has implications for the wider structure of higher education instruction and research in the US.
+At the same time, public university student enrollment and employment of contingenet lecturers dramtically increased at public universities.
+I show that four-year, degree-granting public universities have experienced a systematic fall in state funding per student, and this fall led to a substitution away from tenure-track/tenured professors towards contingent lecturers.
+Falls in state funding for public universities, instrumented by state-level shocks to state-wide higher education, led to a fall in the number of tenure-track/tenured professors per student within a university, and an increase in the number of lecturers.
+Private universities were not exposed to similar financial constraints during this same time period, and do not exhibit the substitution away from tenure-track/tenured professors, so that the stagnating state support for public universities has implications for the wider structure of US higher education and research.
 
 US universities are widely considered the highest performing in the world, yet there are consequential differences between its universities that operate in the private sector and those established by state governments.
-Public universities are subject to numerous state-level administrative laws, and rely on their state governments for funding: an average public university received around \$11,600 per enrolled student in 1990, yet only \$8,300 per enrolled student in 2021.
-This fall is driven by stagnating funding provided by state governments, while enrolment among public universities rose by 46\% over the same time period.
-At the same time, the number of professors per student at private universities stayed relatively stable at 0.045, yet fell from 0.04 to around 0.35 professors per student at public universities.
-This change was driven by a fall of over 20\% in the count of associate and full professors per student at public universities.
+Public universities are subject to numerous state-level administrative laws, and rely on their state governments for funding: an average public university received around \$11,600 of state funding per student in 1990, and only \$8,300 in 2021.
+This fall is driven by a stagnation in the absolute level of state funding for higher education in most states, combined with a large increase of 46\% in student enrolment at public universities over the same time period.
+Similarly, the number of lecturers per student over doubled, while the number of tenure-track/tenured professors per students fell by 23\% from 1990 to 2021.
 
-To identify changes in state funding for higher education, I employ a shift-share instrument for state funding of public universities, which exploits reliance on state funding and yearly shocks to the amount an entire state funds higher education \citep{NBERw23736,NBERw27885}.
-State appropriations for public universities are plausibly endogenous to many outcomes: state governments decide on yearly budgets for their higher education sector, and this process can be influenced by local financial conditions, or changing state priorities.
-The shift-share instrument interacts reliance on state funding in a base period with the yearly total appropriations per student in the entire state, to exploit both changes in support for higher education and how much each university historically has relied on that support.
+I employ a shift-share instrument to identify changes in state funding for higher education, which exploits how universities differ in their financial reliance on state funding, combined with yearly shocks to the amount an entire state funds higher education --- following \cite{NBERw23736,NBERw27885}.
+State governments decide how to fund higher education, and everything else, by a complicated process which can be influenced by local economic conditions, changing state priorities, or even lobbying from the state universities themselves.
+As such, it is necessary to employ an identification strategy, such as the shift-share instrument, to account for endogeneity in universities' state funding.
+The shift-share instrument interacts reliance on state funding in a base period with the entire state' total funding per student for each year, to exploit how much each university historically relied on state funding and yearly changes in state funding for higher education.
 
-Negative changes in state funding affects faculty composition at public universities, away from tenured and tenure-track professors towards non-tenure track lecturer positions.
-A reduction of 10\% in state funding, via a shock to state appropriations, leads to a fall in 1.4\% in the number of assistant professors per student within a university, 1.2\% for full (tenured) professors, and an increase in 4.4\% the number of lecturers per student.
-Local projection estimates show that these effects linger for three to four years after the initial funding shock.
-Over the same time period, state funding fell by around 35\%, while the count of professors per student fell by 9\%, so that these results show that falls in state funding explain around a third of the observed shift away from tenure-track and tenure professors towards contingent faculty.
-Additionally, I use individual level data on professors at Illinois public universities over 2010-2021 to investigate whether the shocks to state funding affected individual professors.
+Falls in state funding affects faculty composition at public universities, who substitute away from tenured and tenure-track professors towards contingent lecturers.
+A fall of $-10$\% in state funding leads to a fall of $-1.4$\% in the number of assistant professors per student at a university, a fall of $-1.2$\% for tenured professors, and an increase in 4.4\% the number of lecturers per student.
+Local projection estimates show that these effects linger for three to four years after the initial funding shock, showing that the effect is not isolated to the immediate year of the first funding shock.
+Over the same time period, state funding fell by around $-$35\%, while the count of professors per student fell by $-$9\%, so that these results show that falls in state funding explain around a third of the substitution away from tenure-track and tenure professors towards contingent lecturers.
+Additionally, I use individual level data on all Illinois public university professors in 2010-2021 to investigate whether the shocks to state funding affected individual professors.
 Incumbent professors were not meaningfully affected, in terms of total salary, promotion rate, or rate of leaving the Illinois public university system.
-Yet, the hiring rate of tenure-track professors declined over this time period for all public universities.
-This shows that changes in faculty composition came about by hiring reductions of new professors at public universities.
+Yet, the hiring rate for professors declined over this time period for all public universities, and particularly for universities with the greatest falls in state funding.
+This implies that faculty composition change arose by limiting the hiring of new professors; public universities increasingly hired contingent lecturers, and increasingly did not replace their retiring (or leaving) professors.
 
@@
 that high school students exposed to similar public school funding shocks are less likely to attend college \citep{jackson2021school}, and that these shocks induce public universities to rely more on tuition as a source of revenue \citep{bound2019public}.
-\cite{NBERw27885} use the same methodology to show that increases in state appropriations lead to degree lower completion time\footnote{
-    \cite{bound2007cohort} use similar variation in state funding per student to show that lower funding leads to lower higher education completion rate at public universities, 
-} and later-life debt for students.
+\cite{NBERw27885} use the same methodology to show that increases in state funding lead to degree lower completion time and later-life debt for students.\footnote{
+    \cite{bound2007cohort} use similar variation in state funding per student to show that lower funding leads to lower completion rates at public universities.
+}
 
-Faculty are a core component of the US higher education and research-innovation university system, yet their composition within universities, and its relationship with higher education finances has so far not been studied.
+Faculty composition is an interesting outcome on its own, but its relationship with higher education finances has so far not been studied.
 The closest previous examples study  how university endowments react to negative endowment shocks \citep{brown2014endowment} and the 2008 recession \citep{turner2014impact}, and multiple theoretically model university decision-making and faculty hiring (see e.g., \citealt{abe2015implications,johnson2009jep,NBERc13879}).
-Lastly, instruction costs within individual universities vary substantially across subject areas, primarily thanks to department differences in class size and faculty salaries \citep{hemelt2021math}.
+%Lastly, instruction costs within individual universities vary substantially across subject areas, primarily thanks to department differences in class size and faculty salaries \citep{hemelt2021math}.
+The number of professors per student may be particularly important for quality of instruction and thus educational outcomes.
+Small class sizes in secondary schools lead to increases in test scores \citep{angrist1999using}, and UK university students perform worse academically in particularly large classes \citep{bandiera2010heterogeneous}.
+Though, the economics literature is not decided on whether contingent lecturers lead to better or worse outcomes in higher education.
+Adjunct professors are generally more effective teachers at highly selective universities \citep{bettinger2010does,figlio2015tenure}, yet students taught by part-time lecturers have worse student outcomes \citep{zhu2021limited}, and universities who use a greater share of lecturers have lower graduation rates \citep{ehrenberg2005tenured}.
 
-The number of professors, both in absolute terms and relative to number of students, may be particularly important for quality of instruction and thus educational outcomes.
-Small class sizes in secondary schools lead to increases in test scores \citep{angrist1999using}, and UK university students perform worse academically in particularly large classes \citep{bandiera2010heterogeneous}.
-Regarding faculty composition, the literature is not decided on whether contingent lecturers lead to better or worse education outcomes relative to tenured faculty.
-Adjunct professors are shown to better undergraduate teachers at highly selective universities \citep{bettinger2010does,figlio2015tenure}, yet universities who use a greater share of lecturers have lower graduation rates \citep{ehrenberg2005tenured}.
-Notably, students taught by part-time adjuncts lecturers have worse student outcomes than full-time professors at a less-selective university \citep{zhu2021limited}, but this difference disappears when individual lecturers gain a full-time teaching position at the university.
-My approach focuses on count of professors per student at universities, and considers faculty count and composition change as possible mechanisms for how stagnating state support for higher education negatively impacts student outcomes \citep{NBERw23736,NBERw27885}.
+While it is not immediately clear whether a systematic substitution towards lecturers away from professors reduces teaching effectiveness, it is clear that stagnating state funding has negatively impacted student outcomes \citep{NBERw23736,NBERw27885}.
+This analysis focuses on a university's count of faculty per student, and considers faculty composition change as one possible mechanism for how stagnating state funding for higher education has negatively affected students at public universities for the last 40 years.

### 2023-05-20 text/sections/02-data.tex
Adjusting writing in TeX document.  2nd revision unfinished.

@@
 US state governments, by majority, are legally obligated to run a balanced-budget, so that yearly variation in tax revenues (caused by changing economic conditions or otherwise) necessarily affect state expenditures.
-This process leads to yearly variation in state appropriations not seen in other revenues sources (such as federal government appropriations), since US states differ in their support for higher education, and are subject to fiscal constraints.
-State appropriations to higher education are a particularly attractive area of state spending to absorb negative shocks to state finances \citep{delaney2011state}.\footnote{
+This process leads to yearly variation in state funding not seen in other revenues sources (such as federal government appropriations), since US states differ in their support for higher education, and are subject to fiscal constraints.
+State funding for higher education are a particularly attractive area of state spending to absorb negative shocks to state finances \citep{delaney2011state}.\footnote{
     \cite{delaney2011state} fully describe the financial environment of state expenditures, and what makes spending on higher education an attractive area for state governments to expand funding during years of higher tax revenues, and retract funding in leaner years.
@@
 }
-Additionally, there is yearly variation in the number of higher education students in each state \citep{turner2014impact}, so that per-student state appropriations vary on multiple dimensions.
+Additionally, there is yearly variation in the number of higher education students in each state \citep{turner2014impact}, so that per-student state funding varies on multiple dimensions.
 
@@
 We see a rise in total revenues received by public universities (from all sources), and a notable increase in mean tuition revenues from \$48 million per year-university to \$150 million per year-university.
-At the same time, total state appropriations stagnated at around \$100 million per year-university for 1990--2008, falling around 2008 and have not recovered ever since.
+At the same time, total state funding stagnated at around \$100 million per year-university for 1990--2008, falling around 2008 and have not recovered ever since.
 While public universities experienced a stagnation in state support, private universities were not exposed to the same constraints, receiving \$37,000 per student in 1990 and \$49,000 in 2021, experiencing no corresponding decline in any specific component.
 
-\begin{figure}[h!]
+\begin{figure}[!htbp]
     \centering
@@
         \centering
-        \caption{Total, millions \$ 2021 CPI-U.}
+        \caption{Total, \$ 2021 CPI-U.}
         \includegraphics[width=\textwidth]{figures/mean-funding-total.png}
@@
         \centering
-        \caption{Per Enrolled Student, \$ 2021 CPI-U.}
+        \caption{Per Student, \$ 2021 CPI-U.}
         \includegraphics[width=\textwidth]{figures/mean-funding-fte.png}
@@
 
-\begin{figure}[h!]
+\begin{figure}[!htbp]
     \centering
     \singlespacing
-    \caption{Total Student enrolment, by University Sector, and Year.}
+    \caption{Total Student enrolment, by University Sector and Year.}
     \begin{subfigure}[b]{0.495\textwidth}
@@
 The rate that Illinois funds higher education has stagnated to a similar extent as the average for the entire country.
-\autoref{fig:illinois-funding} shows the trends for revenues among the seven Illinois public university campuses, including the falling share of state appropriations and substitution towards tuition revenue.
+\autoref{fig:illinois-funding} shows the trends for revenues among the seven Illinois public university campuses, including the falling share of state funding and substitution towards tuition revenue.
 
@@
 While public universities were able to stay open, there were drastic revenue and spending cuts in response to the budget impasse, as it continued through fiscal year 2017, and ended with a new budget restoring funding to state institutions for 2018.
-This variation in public university revenues via the state appropriations channel, stemming entirely from political disagreements --- not from state decisions regarding higher education and its finances \citep{young2020squandered} --- mean that Illinois public universities exhibit sizable changes in their state appropriations over 2010-2021, of similar order to those for the rest of the country over 1990--2021.
+This variation in public university revenues via the state funding channel, stemming entirely from political disagreements --- not from state decisions regarding higher education and its finances \citep{young2020squandered} --- mean that Illinois public universities exhibit sizable changes in their state funding over 2010-2021, of similar order to those for the rest of the country over 1990--2021.

### 2023-05-20 programs/analysis/ipeds-shock.R

@@ -17,7 +17,7 @@ set.seed(47)
 digits.no <- 3
 
 # Size for figures
-fig.width <- 10
+fig.width <- 9
 fig.height <- fig.width * 0.85
 
 
@@ -100,29 +100,21 @@ reg.data %>%
         lecturer_prof_count = lecturer_prof_count,
         assistant_prof_count = assistant_prof_count,
         full_prof_count = full_prof_count,
-        all_prof_count = all_prof_count,
-        lecturer_profmeansalary_real = lecturer_profmeansalary_real,
-        assistant_profmeansalary_real = assistant_profmeansalary_real,
-        full_profmeansalary_real = full_profmeansalary_real,
-        all_profmeansalary_real = all_profmeansalary_real) %>%
+        all_prof_count = all_prof_count) %>%
     as.data.frame() %>%
     stargazer(summary = TRUE,
         summary.stat = c("mean", "sd", "n"),
         digits = 0,
         digits.extra = 0,
         covariate.labels = c(
-            "Enrollment, institution reported",
+            "Enrolment",
             "State appropriations (millions 2021 USD)",
             "Total revenues (millions 2021 USD)",
-            "Non-instutional revenues (millions 2021 USD)",
+            "Non-institutional revenues (millions 2021 USD)",
             "Lecturers count",
             "Assistant professors count",
             "Full professors count",
-            "All professors count",
-            "Lecturers mean salary (2021 USD)",
-            "Assistant mean salary (2021 USD)",
-            "Full mean salary (2021 USD)",
-            "All mean salary (2021 USD)"),
+            "All professors count"),
         omit.table.layout = "n",
         header = FALSE, float = FALSE, no.space = TRUE,
         type = "text",
@@ -190,23 +182,26 @@ mean_funding_fte.data <- reg.data %>%
 mean_funding_fte.graph <- mean_funding_fte.data %>%
     select(-totalrevenues_real) %>%
     pivot_longer(!year, names_to = "variable", values_to = "millions") %>%
-    ggplot(aes(x = year, y = millions, colour = variable)) +
+    ggplot(aes(x = year, y = millions / 10^3, colour = variable)) +
     geom_point() +
     geom_line() +
     # Adjust the names and axis
     scale_x_continuous(name = "Year",
         breaks = seq(1985, 2020, by = 5)) +
     scale_y_continuous(name = "",
[…]

### 2023-05-24 text/sections/00-abstract.tex
Second major revision: Concepts needs overhaul.

@@
 While states have limited their funding for higher education for the last forty years, US universities have systematically substituted away from tenure-track/tenured professors towards contingent lecturers.
-This analysis relates the changes in the faculty composition at US public universities to the fall in state funding for higher education, using a shift-share approach to isolate the causal effects of changes in state funding.
+This analysis relates the changes the faculty composition at US public universities to the fall in state funding for higher education, using a shift-share approach to identify causal effects.
 A $-10$\% fall in a university's funding, via a state funding shock, decreases the number of assistant professors per student by $-1.4$\% and full professors by $-1.2$\%, yet increases the number of lecturers per student by 4.4\%; long-run estimates show the effects linger for multiple years after the initial funding shock.
-Analysis of all the professors at Illinois public universities 2011-2021 shows that incumbent professors are not affected by the changes in state funding, and suggestive evidence shows that these changes arose by limiting the hiring of new professors at public universities to replace leaving tenure-track/tenured professors.
+Analysis of all the professors at Illinois public universities shows that incumbent professors are not affected by substantial funding shocks over 2011-2021, and suggestive evidence shows that the faculty changes arose by falling hires of new professors to replace leaving tenure-track/tenured professors.
 These results show the institutional effects of stagnating state funding for higher education, and raise questions about the direction that public education heads as these financial headwinds show no sign of dissipating.

### 2023-05-24 text/sections/01-introduction.tex
Second major revision: Concepts needs overhaul.

@@
 Though, the economics literature is not decided on whether contingent lecturers lead to better or worse outcomes in higher education.
-Adjunct professors are generally more effective teachers at highly selective universities \citep{bettinger2010does,figlio2015tenure}, yet students taught by part-time lecturers have worse student outcomes \citep{zhu2021limited}, and universities who use a greater share of lecturers have lower graduation rates \citep{ehrenberg2005tenured}.
+Adjunct lecturers are generally more effective teachers than research professors at highly selective universities \citep{bettinger2010does,figlio2015tenure}, but students taught by part-time lecturers have worse student outcomes relative to full-time lecturers or faculty \citep{zhu2021limited,ehrenberg2005tenured}.
 
 While it is not immediately clear whether a systematic substitution towards lecturers away from professors reduces teaching effectiveness, it is clear that stagnating state funding has negatively impacted student outcomes \citep{NBERw23736,NBERw27885}.
-This analysis focuses on a university's count of faculty per student, and considers faculty composition change as one possible mechanism for how stagnating state funding for higher education has negatively affected students at public universities for the last 40 years.
+This analysis focuses on a university's count of faculty per student, and considers faculty composition change as one possible mechanism for how stagnating state funding for higher education has negatively affected students at public universities for the last 30 years.

### 2023-05-24 text/sections/02-data.tex
Second major revision: Concepts needs overhaul.

@@
 }
-This publicly available data provides the basis to build a panel of Illinois public university professors 2010-2021; I define a professor as an individual by their first plus last name and university pairing, and link this database to IPEDS regarding finances for their employing institution.
+These publicly available data provides the basis to build a panel of Illinois public university professors 2010-2021; I define a professor as an individual by their first plus last name and university pairing, and link this database to IPEDS regarding finances for their employing institution.
 
@@
     \singlespacing
-    \caption{Mean Total Revenues among Public Universities, by Year.}
+    \caption{Mean Total Funding among Public Universities, by Year.}
     \begin{subfigure}[b]{0.495\textwidth}
@@
         \centering
-        \caption{Total.}
+        \caption{Total Enrolment, Nation-wide.}
         \includegraphics[width=\textwidth]{figures/enrollment-total.png}
@@
         \centering
-        \caption{Mean, University Level.}
+        \caption{Mean Enrolment, per University.}
         \includegraphics[width=\textwidth]{figures/enrollment-mean.png}
@@
     \singlespacing
-    \caption{Student Enrollment per Professor, by University Sector, Professor Appointment, and Year.}
+    \caption{Trends in Mean Student Enrolment per Professor, by University Sector and Faculty level.}
     \begin{subfigure}[b]{0.495\textwidth}
         \centering
-        \caption{Lecturer}
+        \caption{Lecturer.}
         \includegraphics[width=\textwidth]{figures/lecturer-fte-perprof.png}
@@
 The rate that Illinois funds higher education has stagnated to a similar extent as the average for the entire country.
-\autoref{fig:illinois-funding} shows the trends for revenues among the seven Illinois public university campuses, including the falling share of state funding and substitution towards tuition revenue.
+\autoref{fig:illinois-funding} shows the trends for funding sources among the seven Illinois public university campuses, including the falling share of state funding and substitution towards tuition revenue.
 
@@
     \singlespacing
-    \caption{Mean Revenues among Illinois Public Universities, by Year.}
+    \caption{Mean Funding Sources among Illinois Public Universities, by Year.}
     \begin{subfigure}[b]{0.495\textwidth}
         \centering
-        \caption{Total, millions \$ 2021 CPI-U.}
+        \caption{Total, \$ 2021 CPI-U.}
         \includegraphics[width=\textwidth]{figures/illinois-funding-total.png}
@@
         \centering
-        \caption{Per Enrolled Student, \$ 2021 CPI-U.}
+        \caption{Per Student, \$ 2021 CPI-U.}
         \includegraphics[width=\textwidth]{figures/illinois-funding-fte.png}

### 2023-05-24 text/sections/04-empirical.tex
Second major revision: Concepts needs overhaul.

@@
 The instrument approach relies on the conditional independence assumption, that the instrument is independently assigned to the universities.
-While this assumption is fundamentally untestable, we see that univeresities are by-in-large smaller (in terms of enrolment, total revenues, and professor count) at the top of the state funding shock distribution.
+While this assumption is fundamentally untestable, we see that universities are by-in-large smaller (in terms of enrolment, total revenues, and professor count) at the top of the state funding shock distribution.
 Though on a per student basis, there is little difference across the distribution (except in the outcomes under consideration).

### 2023-05-24 text/sections/06-discussion.tex
Second major revision: Concepts needs overhaul.

@@
 Universities with more state funding (for the entire decade) hired more professors (\autoref{fig:hiring-correlation.png}).
-Similarly, the funding shock instrument identifies that an increase in 10\% state funding at an individual university leads to 13 more faculty hires on average across the decade 2011--2021 (see \autoref{tab:hiring-shock-reg}).\footnote{
+Similarly, the funding shock instrument shows that an increase in 10\% state funding leads to 13\% more faculty hires at a university across the decade 2011--2021 (see \autoref{tab:hiring-shock-reg}).\footnote{
     These results were produced by integrating the total count of faculty hires for 2010--2021 for the top-ranked 180 US universities with a sum of the funding variables, and then estimating the models specified in \autoref{sec:iv-model-uni}.
-    There were no observable differences in the hiring rate of ale vs female faculty.
+    There were no observable differences in the hiring rate of male vs female faculty.
     See \autoref{sec:appendix-hiring} for further details.
@@
 It is clear that public universities' hiring is negatively impacted by falls in state funding, even using highly aggregated data for a total count of hiring across an entire decade.
-While this analysis strengthens the case negative effects on hiring plays, it is crude for measuring exact magnitudes for the effect on faculty hiring; I leave it to further research to delve deeper into the magnitudes that hiring plays as the primary causal mediator for the effects of falls in state funding on faculty composition at public universities.
+While this analysis strengthens the case negative effects on hiring plays, it is crude for measuring exact magnitudes for the effect on faculty hiring.
+I leave it to further research to delve deeper into the magnitudes that hiring plays as the primary causal mediator for the effects of falls in state funding on faculty composition at public universities.

### 2023-05-24 programs/analysis/ipeds-public-private.R

@@ -12,8 +12,8 @@ library(plm)
 digits.no <- 3
 
 # Size for figures
-fig.width <- 10
-fig.height <- fig.width * 0.75
+fig.width <- 9
+fig.height <- fig.width * 0.85
 
 
 # Load data sources ------------------------------------------------------------
@@ -78,34 +78,42 @@ enrollment.data <- ipeds.data %>%
     ungroup()
 # Draw the graph for enrollment total
 enrollment_total.plot <- enrollment.data %>%
-    ggplot(aes(x = year, y = enrollment_total, colour = factor(public))) +
+    ggplot(aes(x = year, y = enrollment_total / 10^6, colour = factor(public))) +
     geom_point() +
     geom_line() +
     scale_x_continuous(name = "Year",
         breaks = seq(1985, 2020, by = 5)) +
     scale_y_continuous(name = "",
-        limits = c(0, 10^7),
+        limits = c(0, 10),
+        breaks = seq(0, 10, by = 2),
         labels = scales::comma) +
     theme_bw() +
-    theme(plot.title = element_text(hjust = 0.5),
-        legend.position = "top") +
+    ggtitle("Student Enrolment, millions") +
+    theme(plot.title = element_text(size = rel(1)),
+        plot.margin = unit(c(0.5, 0, 0, 0), "mm"),
+        legend.position = "bottom",
+        legend.margin = margin(t = -10)) +
     scale_colour_discrete(name = "", labels = c("Private", "Public"))
 ggsave("../../text/figures/enrollment-total.png",
     plot = enrollment_total.plot,
     units = "cm", width = fig.width, height = fig.height)
 # Draw the graph for enrollment mean
 enrollment_mean.plot <- enrollment.data %>%
-    ggplot(aes(x = year, y = enrollment_mean, colour = factor(public))) +
+    ggplot(aes(x = year, y = enrollment_mean / 10^3, colour = factor(public))) +
     geom_point() +
     geom_line() +
     scale_x_continuous(name = "Year",
         breaks = seq(1985, 2020, by = 5)) +
     scale_y_continuous(name = "",
-        limits = c(0, 12500),
+        limits = c(0, 12.5),
+        breaks = seq(0, 13, by = 2),
         labels = scales::comma) +
     theme_bw() +
-    theme(plot.title = element_text(hjust = 0.5),
-        legend.position = "top") +
+    ggtitle("Student Enrolment, thousands") +
[…]

### 2023-05-24 programs/analysis/ipeds-shock.R

@@ -9,6 +9,7 @@ library(car) # Function for F stat regarding IV models
 # My custom flavour of Stargazer TeX tables:
 # devtools::install_github("shoganhennessy/stargazer")
 library(stargazer) # TeX tables
+library(xtable)
 set.seed(47)
 # This file follows an adjusted Deming Walters (2017, p.10) approach to
 # appropriations shock instrument.
@@ -128,12 +129,12 @@ reg.data %>%
             appropriationshock_perEnroll_real))) %>%
     group_by(instrument_quantile) %>%
     summarise(
-        `State Funding Shock, USD per student` = round(mean(appropriationshock_perEnroll_real, an.rm = TRUE)),
+        `State Funding Shock, \\$ per student` = round(mean(appropriationshock_perEnroll_real, an.rm = TRUE)),
         `State Funding, per student`           = round(mean(stateappropriations_real / enrollment_reported, na.rm = TRUE)),
         `Total Full-time Enrolment`            = round(mean(enrollment_reported, na.rme = TRUE)),
-        `State Funding, USD millions`          = round(mean(stateappropriations_real / 10^6, na.rme = TRUE)),
-        `Total Revenues, USD millions`         = round(mean(totalrevenues_real / 10^6, na.rme = TRUE)),
-        `Total Revenues, USD per student`      = round(mean(totalrevenues_real / enrollment_reported, na.rm = TRUE)),
+        `State Funding, \\$ millions`          = round(mean(stateappropriations_real / 10^6, na.rme = TRUE)),
+        `Total Revenues, \\$ millions`         = round(mean(totalrevenues_real / 10^6, na.rme = TRUE)),
+        `Total Revenues, \\$ per student`      = round(mean(totalrevenues_real / enrollment_reported, na.rm = TRUE)),
         `Lecturer Count`                       = round(mean(lecturer_prof_count, na.rm = TRUE)),
         `Assistant Professor Count`            = round(mean(assistant_prof_count, na.rm = TRUE)),
         `Full Professor Count`                 = round(mean(full_prof_count, na.rm = TRUE)),
@@ -154,14 +155,16 @@ reg.data %>%
         `3rd` = `2`,
         `4th` = `3`,
         `5th` = `4`) %>%
-    as.data.frame() %>%
-    stargazer(summary = FALSE,
-        rownames = FALSE,
-        digits = 3,
-        omit.table.layout = "n",
-        header = FALSE, float = FALSE, no.space = TRUE,
[…]

### 2023-05-30 text/sections/00-abstract.tex
Finished round of writing revisions.  Only spell-checking remains.

@@
-While states have limited their funding for higher education for the last forty years, US universities have systematically substituted away from tenure-track/tenured professors towards contingent lecturers.
+US universities have systematically substituted away from tenure-track/tenured professors towards contingent lecturers for the last thirty years, while at the same time states have limited how much they fund higher education.
 This analysis relates the changes the faculty composition at US public universities to the fall in state funding for higher education, using a shift-share approach to identify causal effects.
-A $-10$\% fall in a university's funding, via a state funding shock, decreases the number of assistant professors per student by $-1.4$\% and full professors by $-1.2$\%, yet increases the number of lecturers per student by 4.4\%; long-run estimates show the effects linger for multiple years after the initial funding shock.
-Analysis of all the professors at Illinois public universities shows that incumbent professors are not affected by substantial funding shocks over 2011-2021, and suggestive evidence shows that the faculty changes arose by falling hires of new professors to replace leaving tenure-track/tenured professors.
+A fall of $-10$\% in university funding, via a state funding shock, decreases the number of assistant professors per student by $-1.4$\% and full professors by $-1.2$\%, yet increases the number of lecturers per student by 4.4\%; long-run estimates show the effects linger for multiple years after the initial funding shock.
+Analysis of all the professors at Illinois public universities shows that incumbent professors are not affected by substantial funding shocks over 2011-2021, with suggestive evidence showing that the faculty changes arose by limiting the hiring of new professors to replace leaving tenure-track/tenured professors.
 These results show the institutional effects of stagnating state funding for higher education, and raise questions about the direction that public education heads as these financial headwinds show no sign of dissipating.

### 2023-05-30 text/sections/01-introduction.tex
Finished round of writing revisions.  Only spell-checking remains.

@@
-Public universities educate the majority of higher education students in the US, yet have experienced a secular decline in state funding (per student) the last three decades.
-At the same time, public university student enrollment and employment of contingenet lecturers dramtically increased at public universities.
+Public universities educate the majority of higher education students in the US, yet have experienced a secular decline in state funding (per student) over the last three decades.
+At the same time, both public university student enrolment and employment of contingent lecturers dramatically increased at public universities.
 I show that four-year, degree-granting public universities have experienced a systematic fall in state funding per student, and this fall led to a substitution away from tenure-track/tenured professors towards contingent lecturers.
@@
 This fall is driven by a stagnation in the absolute level of state funding for higher education in most states, combined with a large increase of 46\% in student enrolment at public universities over the same time period.
-Similarly, the number of lecturers per student over doubled, while the number of tenure-track/tenured professors per students fell by 23\% from 1990 to 2021.
+Similarly, the number of lecturers per student over doubled, while the number of tenure-track/tenured professors per students fell by $-23$\% from 1990 to 2021.
 
@@
 As such, it is necessary to employ an identification strategy, such as the shift-share instrument, to account for endogeneity in universities' state funding.
-The shift-share instrument interacts reliance on state funding in a base period with the entire state' total funding per student for each year, to exploit how much each university historically relied on state funding and yearly changes in state funding for higher education.
+The shift-share instrument interacts reliance on state funding in a base period with the entire state's total funding per higher education student for each year, to exploit changes in funding and how much each university historically relied on state funding for higher education.
 
-Falls in state funding affects faculty composition at public universities, who substitute away from tenured and tenure-track professors towards contingent lecturers.
+Falls in state funding affects faculty composition at public universities, which substitute away from tenured and tenure-track professors towards contingent lecturers.
 A fall of $-10$\% in state funding leads to a fall of $-1.4$\% in the number of assistant professors per student at a university, a fall of $-1.2$\% for tenured professors, and an increase in 4.4\% the number of lecturers per student.
@@
 Incumbent professors were not meaningfully affected, in terms of total salary, promotion rate, or rate of leaving the Illinois public university system.
-Yet, the hiring rate for professors declined over this time period for all public universities, and particularly for universities with the greatest falls in state funding.
+Yet the hiring rate for new professors at publicuniversities was negatively impacted by the falls in state funding (and by the funding shocks).
 This implies that faculty composition change arose by limiting the hiring of new professors; public universities increasingly hired contingent lecturers, and increasingly did not replace their retiring (or leaving) professors.
-
-Absent this analysis, it is not immediately clear why stagnating state support and changes in public university faculty composition would be causally linked.
-Over the 21st Century there have been drastic changes to college selectivity, where the top universities have become more selective while the average university less selective \citep{hoxby2009changing}.
-While the average public university is becoming less selective, it is possible that their most productive or research-focused faculty more often move to more selective and prestigious (and often private) institutions, so that over time public universities (on average) substitute towards contingent lecturers.
-This is one way in which the trends may be concurrent, but not causal.
-This analysis shows that stagnating in state funding for public universities cause the changes in faculty composition, explaining about 40\% of the observed substitution towards contingent lecturers, so that it is possible that other systematic changes in higher education (such as changing selectivity) may explain the rest of the variation.
 
@@
 
+Over the last half century, there have been drastic changes to the heirarchy of US higher education, where the top universities have become more selective for student applicants, and the average university less selective \citep{hoxby2009changing}.
+Similarly, research faculty at US universities are increasingly concentrated among the graduates of highly selective PhD programmes \citep{wapman2022quantifying}.
 Faculty composition is an interesting outcome on its own, but its relationship with higher education finances has so far not been studied.
@@
 %Lastly, instruction costs within individual universities vary substantially across subject areas, primarily thanks to department differences in class size and faculty salaries \citep{hemelt2021math}.
-The number of professors per student may be particularly important for quality of instruction and thus educational outcomes.
-Small class sizes in secondary schools lead to increases in test scores \citep{angrist1999using}, and UK university students perform worse academically in particularly large classes \citep{bandiera2010heterogeneous}.
+The number of professors per student may be particularly important for quality of instruction and thus educational outcomes;
+small class sizes in secondary schools lead to increases in test scores \citep{angrist1999using}, and UK university students perform worse academically in particularly large classes \citep{bandiera2010heterogeneous}.
 Though, the economics literature is not decided on whether contingent lecturers lead to better or worse outcomes in higher education.
@@
 While it is not immediately clear whether a systematic substitution towards lecturers away from professors reduces teaching effectiveness, it is clear that stagnating state funding has negatively impacted student outcomes \citep{NBERw23736,NBERw27885}.
-This analysis focuses on a university's count of faculty per student, and considers faculty composition change as one possible mechanism for how stagnating state funding for higher education has negatively affected students at public universities for the last 30 years.
+This analysis focuses on a university's count of faculty per student, and considers faculty composition change as one of the mechanisms for how stagnating state funding for higher education transmits to students and has led to worsening education outcomes at public universities for the last thirty years.

### 2023-05-30 text/sections/02-data.tex
Finished round of writing revisions.  Only spell-checking remains.

@@
 The decade of 2010--2021 has also seen yearly variation in its revenues, particularly around 2016.
-In the calendar year 2015 partisan disagreements between the democratic legislature and republican governor led to the 2016 fiscal year starting with no state budget.
+In the calendar year 2015, partisan disagreements between the democratic legislature and republican governor led to the 2016 fiscal year starting with no state budget.
 State agencies, and higher education institutions, employed accounting techniques to continue operating without any resources provided by the state government.\footnote{

### 2023-05-30 text/sections/03-concepts.tex
Finished major revision.  Needs more checks of writing style.
Adjusting writing in TeX document.  2nd revision unfinished.
Second major revision: Concepts needs overhaul.
Finished round of writing revisions.  Only spell-checking remains.

@@
 
-Most universities hire a number of new professors every year, either to expand their departments or to replace leaving professors.
+Universities hire new professors most years, either to expand their departments or to replace leaving professors.
 The number of professors hired in each year is usually highest among non-tenured adjunct faculty, as these instructors are mostly hired on short-term contingent contracts, so that they are less costly to hire (or to not reinstate) in response to yearly changes in teaching needs.
-Secondly, tenure-track assistant professors are hired to replace leaving assistant professors, and are most often hired with a four to six year contract and agreement for tenure consideration.
+Tenure-track assistant professors are hired most often hired with a four to six year contract, and formal agreement for tenure consideration at the end of this term.
 Lastly, tenured professors have undergone review and have successfully secured a full-time appointment at their university with no expiration date; tenured professors may be hired from outside a university, though this position has the lowest yearly hiring rate among most universities.
-It is common for universities to restrict their faculty hiring in the face of financial shocks.
-Since the yearly hiring rate is highest among lecturers and assistant professors, financial shocks may lower the number of lecturer and assistant professors at public universities the most, if faculty hiring is affected by falls in funding.
+It is common for universities to restrict or delay their faculty hiring in the spring term ahead of the next academic year, in response to unforeseen events, or changes in needs.
+So it stands to reason that the hiring rate of new professors could be meaningfully affected by changes in university finances, and thus falls in state funding.
 
-Additionally, faculty salaries may by affected by changes in public university finances.
-When the university has a lower budget for its yearly hiring, it may respond by lowering the salary they offer to new hires.
-This possible effect may not be the same across each position of professor; tenured faculty are often hired away from another university, so that new tenured professor hires may be less likely to accept a lower salary offer from a public university.\footnote{
-    Faculty often consider their outside options, as in offers for employment and/or promotion from other universities.
-    See \cite{blackaby2005} for an overview of faculty outside options.
-}
-Yet salary for all the professors, not just new hires, may also be affected: multiple universities passed a university-wide pay-cut for their faculty in response to state budget cuts around the 2008 recession, for example.\footnote{
-    Indeed, Cornell University implemented hiring freezes and a nominal salary reduction for professors, in anticipation of a financial shock in early 2021.
-    The salary cut was not permanent, as the oncoming financial shock turned out to not be as serious as projected, so that the cut was returned to professors in the year 2021.
-}
-Faculty are not paid the same by position, so that there are financial consequences for a university promoting their own faculty.
-In response to tightening fiscal restraints, universities may be less likely to grant tenure to assistant professors (as tenured professors generally receive a salary raise and long term contract), or to promote tenured professors from associate to full professor.
+%Additionally, faculty salaries may be affected by changes in public university finances.
+%When the university has a lower budget for its yearly hiring, it may respond by lowering the salary they offer to new hires.
+%This possible effect may not be the same across each position of professor; tenured faculty are often hired away from another university, so that new tenured professor hires may be less likely to accept a lower salary offer from a public university.\footnote{
+%    Faculty often consider their outside options, as in offers for employment and/or promotion from other universities.
+%    See \cite{blackaby2005} for an overview of faculty outside options.
+%}
+%Yet salary for all the professors, not just new hires, may also be affected: multiple universities passed a university-wide pay-cut for their faculty in response to state budget cuts around the 2008 recession, for example.\footnote{
+%    Indeed, Cornell University implemented hiring freezes and a nominal salary reduction for professors, in anticipation of a financial shock in early 2021.
+%    The salary cut was not permanent, as the oncoming financial shock turned out to not be as serious as projected, so that the cut was returned to professors in the year 2021.
+%}
+%Faculty are not paid the same by position, so that there are financial consequences for a university promoting their own faculty.
+%In response to tightening fiscal restraints, universities may be less likely to grant tenure to assistant professors (as tenured professors generally receive a salary raise and long term contract), or to promote tenured professors from associate to full professor.
 
-It is not immediately clear which effects will dominate, or whether tenured professors vs lecturers will be affected more by changes in their university's funding.
-Yet, there is one empirical fact worth noting: a mean assistant or tenured professor earns more than double that of an adjunct lecturer per year (see \autoref{tab:illinois-summary}), and the tenure contract means universities must pay a higher salary for a guaranteed number years, more so than for contingent lecturers.
-If a public university's primary obligation is to teach, and they must fulfil this objective with fewer and fewer resources, then they may substitute away from tenure-track and tenured professor towards contingent lecturers to achieve this obligation.\footnote{
-    That is, substituted away from tenure-track and tenured professors according to the salary ratio, and the relative productivity between these two positions.
-}
+Absent this analysis, it is not immediately clear why stagnating state support and changes in faculty composition at public universities would be causally linked.
+Over the 21st Century there have been drastic changes to college selectivity, where the top universities have become more selective while the average university less selective \citep{hoxby2009changing}.
+While the average public university is becoming less selective, it is possible that their most productive or research-focused faculty more often move to more selective and prestigious (and often private) institutions, so that over time public universities (on average) substitute towards contingent lecturers.
+This is one way in which the trends may be concurrent, but not causal.
+This analysis shows that stagnating in state funding for public universities cause the changes in faculty composition, explaining about 40\% of the observed substitution towards contingent lecturers, so that it is possible that other systematic changes in higher education (such as changing selectivity) may explain the rest of the variation.
+
+It is not immediately clear which effects will dominate, or whether tenured professors or lecturers will be affected more by changes in their university's funding.
+Yet there is one empirical fact worth noting: the average assistant or tenured professor earns more than double that of a lecturer (see \autoref{tab:illinois-summary}).
+Additionally, universities must pay the higher salary to tenure-track/tenured professors for a guaranteed number years, which is not true for lecturers who are mostly on shorter term contracts.
+If a public university's primary obligation is to teach, and they must fulfil this objective with fewer and fewer resources, then they may substitute away from tenure-track and tenured professor towards contingent lecturers.

### 2023-05-30 text/sections/04-empirical.tex
Finished round of writing revisions.  Only spell-checking remains.

@@
 
-Universities employ professors, so it stands to reason that if a university experiences a fall in one of their revenue sources (similar to a demand shock in the private sector) that they may respond by making adjustments with respect to their labour force of professors.
+Universities employ professors, so it stands to reason that if a university experiences a fall in one of their revenue sources --- similar to a demand shock in the private sector --- then they may respond by adjusting their employee composition.
 
@@
 Instead, the state government and university administration undertake a complex process of allotting resources across multiple different priorities, including instruction, research, or between departments.
-Importantly for this analysis, revenues received from an institution's state government provide opportunity to address this endogeneity.
+For example, it may be the case that a state government restricts funding for only its lowest performing universities in response to a budget shock, introducing a treatment selection-on-outcome threat to identification.
+Importantly for this analysis, revenues received from an institution's state government provide opportunity to address such endogeneity.
 
@@
 \end{align}
-The system exploits the fact that institutions who rely on state funding more will be affected by state appropriation shocks.
-$Z_{i,t}$ is the instrument for (log) state appropriations for institution $i$ in year $t$, interacting the average funding for universities in state $s(i)$ with reliance on state funding relative to total revenues, averaged across the base years 1990--1993.\footnote{
+The system exploits the fact that institutions who rely on state funding more will be affected by state funding shocks.
+$Z_{i,t}$ is the instrument for (log) state funding for institution $i$ in year $t$, interacting the average funding for universities in state $s(i)$ with reliance on state funding relative to total revenues, averaged across the base years 1990--1993.\footnote{
     1990--1993 are defined as data for public university finance data are most comparable (i.e. without many missing values) beginning in 1990.
@@
 Though on a per student basis, there is little difference across the distribution (except in the outcomes under consideration).
-Yearly and individual fixed effects are included in regressions throughout to implicitly condition on mean differences between universities, and strengthen the claim of conditional independence of the instrument.
+Yearly and individual fixed effects are included in regressions throughout to implicitly condition on mean differences between universities and years, leading to the assumption of conditional independence of the instrument with respect to state funding for each public university.
 
@@
     \singlespacing
-    %\centering
+    \centering
     \caption{Mean Characteristics for Public Universities, by State Funding Shock Instrument.}
@@
     \textbf{Note}:
-    The column labelled ``1st'' refers to the mean for all university-year observations in the first quintile (bottom 20\%) of the funding shock (per student) distribution, and so on.
+    The column labelled ``1st'' refers to the mean for all university-year observations in the first quintile (bottom 20\%) of the funding shock distribution, and so on.
+    The mean of the funding shock within each quintile is shown in the first row of the table.
 \end{table}
@@
 Secondly, the shift-share identification strategy requires exogeneity in either the base-line share or shift component of the instrument.
-In this case, we satisfy the second: universities' institutional-level decisions are not correlated with contemporaneous or upcoming shocks to state appropriations.\footnote{
+In this case, we satisfy the second: universities' institutional-level decisions are not correlated with contemporaneous or upcoming shocks to state funding.\footnote{
     It would be plausible to consider the case that universities make institutional-level decisions in a consistently different manner to those 
-    with differing reliance on state appropriations in 1990, so that exogeneity by the base-line share is not plausible here.
+    with differing reliance on state funding in 1990, so that exogeneity by the base-line share is not plausible here.
 }
@@
     \centering
-    \caption{First Stage Estimates, for State Funding by Appropriation Shock.}
+    \caption{First Stage Estimates, for State Funding by funding shock.}
     \makebox[\textwidth][c]{\input{tables/firststage-reg-fte.tex}}
@@
 }
-We note columns (1) and (2) estimate that a shock to appropriations (per student in the entire state) of 10\% is associated with 9.8\% change in state funding per student at the university; the instrument is strong, and we note similarity in estimates with and without inclusion of fixed effects.
+We note columns (1) and (2) estimate that a funding shock (per student in the entire state) of 10\% is associated with 9.8\% change in state funding per student at the university; the instrument is strong, and we note similarity in estimates with and without inclusion of fixed effects.
 Columns (3) and (4) include the tuition revenue control (explained in \autoref{foot:control}) to exhibit estimates with the inclusion of this possible collider or bad control.
@@
 Column (1) represents the estimates for \autoref{eqn:firststage} with fixed effects, omitting the tuition revenue control, and is the preferred form that I proceed with.
-Lastly, the state funding shock affects university revenues for multiple years after the initial shock (seee \autoref{fig:firststage-lp}).
-We see a decaying, yet real and positive, effect of the shock on revenues for the next 10 years, illustrating that the state appropriation shock is a strong, if fading, instrument for state funding per student and justifying its use in later local projections models.
+Lastly, the state funding shock affects university revenues for multiple years after the initial shock (see \autoref{fig:firststage-lp}).
+We see a decaying, yet real and positive, effect of the shock on revenues for the next 10 years, illustrating that the state funding shock is a strong, if fading, instrument for state funding per student and justifying its use in later local projections models.
 
-These results, together with the case for exogeneity, show the appropriations shock instrument strongly predicts university revenues in the first-stage estimation.
+These results, together with the case for exogeneity, show the funding shock instrument strongly predicts university revenues in the first-stage estimation.
 
@@
 
-The primary empirical model combines the instrumental variable for state appropriation shocks with the empirical model for the effects of state funding at the university-level --- i.e., parameter $\beta$ in the following.\footnote{
+The primary empirical model combines the instrumental variable for state funding shocks with the empirical model for the effects of state funding at the university-level --- i.e., parameter $\beta$ in the following.\footnote{
     It is important to note the treatment effect isolated here; the instrumental variables approach identifies the local average treatment effects, one specific to the instrument.
-    So we interpret this treatment effect as a university's response in employment count and average salaries to state funding changes, changes specific to state appropriation shocks, among the complier group --- i.e., universities who respond to funding shocks who would not have made faculty-outcome changes absent the funding shock.
+    So we interpret this treatment effect as a university's response in employment count and average salaries to state funding changes, changes specific to state funding shocks, among the complier group --- i.e., universities who respond to funding shocks who would not have made faculty-outcome changes absent the funding shock.
     Also, we assume that no universities' total revenues increase in response to negative state funding shock (monotonicity).
@@
 I estimate the system by two stage least squares, including institution and year fixed effects, and investigate outcomes at the university-level to analyse effects on the university as a result of changes in revenues.
-Additionally, I estimate the model via local projections \citep{jorda2005,miller2022} to investigate whether the effects of revenues on faculty composition linger for multiple years after the original appropriation shock.\footnote{
-    The local projections method is an empirical method to estimate dynamic treatment effects when the treatment of interest is not binary, so that time-series confounding (i.e., treatment in time $t$ is correlated with treatment in time $t-1$) is present even for instrumental variable models.
-    Inference for this model is similar to that for an event study design, so I present estimates in graphical form following that example (\autoref
+Additionally, I estimate the model via local projections \citep{jorda2005,miller2022} to investigate whether the effects on faculty composition linger for multiple years after the original funding shock.\footnote{
+    The local projections method is an empirical method to estimate dynamic treatment effects when the treatment is not binary, so that time-series confounding (i.e., treatment in time $t$ is correlated with treatment in time $t-1$) is present, even for instrumental variable models.
+    Inference for this model is similar to that for an event study design, so I present estimates in graphical form (see \autoref
     {fig:count-lp}).
 }
-Regarding outcomes, I focus on the composition of the professors employed at the university by analysing (log) count per student of, and average salaries paid to, professors employed by the university.
+Regarding outcomes, I focus on the composition of the professors employed at the university by analysing (log) count of professors per student by each university.
 
@@
 Redefining the level of outcome requires adjustment to the empirical approach, leveraging variation in university finances for the years after a professor joins the university.\footnote{
-    This formulation follows that presented by \cite{NBERw27885}, where individual student outcomes are analysed via variation in state appropriations after their freshman-year.
+    This formulation follows that presented by \cite{NBERw27885}, where individual student outcomes are analysed via variation in state funding after their freshman-year.
     This contrasts with \autoref{sec:iv-model-uni} and \cite{NBERw23736}, where the unit of analysis is the university-year, where base year 1990 is more appropriate.
@@
 
-\autoref{eqn:rolling-instrument} defines a rolling-share variant of the instrument, $\tilde Z_{j,t}$, where the university's state appropriations share exposure is based in the year a professor joins the university --- and not the base period 1990--1993.
+\autoref{eqn:rolling-instrument} defines a rolling-share variant of the instrument, $\tilde Z_{j,t}$, where the university's state funding share exposure is based in the year a professor joins the university --- and not the base period 1990--1993.
 $j$ indexes each professor in year $t$, $\tau(j)$ for the year the professor first joins their institution.\footnote{
@@
 }
-We satisfy the assumptions for exogeneity by noting that none of the Illinois public campuses take the majority of state appropriations, and that the identification strategy relies on exogeneity in changes in state appropriations to individual professor-outcomes, following the year they joined the university.\footnote{
+We satisfy the assumptions for exogeneity by noting that none of the Illinois public campuses take the majority of state funding, and that the identification strategy relies on exogeneity in changes in state funding to individual professor-outcomes, following the year they joined the university.\footnote{
     Additionally, within-institution changes resulting from share reliance on state funding may be correlated with unobserved changes in the outcomes, so that \cite{NBERw27885} note the importance of controlling for the base share and state student population.
@@
 }
-\autoref{tab:firststage-illinois} presents results of the first stage estimation, showing that the instrument is strong in the same way as that for the university-level outcomes (\autoref{tab:firststage-reg}), with very similar estimates for the association between appropriation shocks and state funding.
+\autoref{tab:firststage-illinois} presents results of the first stage estimation, showing that the instrument is strong in the same way as that for the university-level outcomes (\autoref{tab:firststage-reg}), with very similar estimates for the association between funding shocks and state funding.
 
@@
 \end{eqnarray}
-We then interpret parameter $\beta$ as the effect of changes in state funding at an Illinois public university, via state appropriation shocks, on an individual professor's outcome $Y_{j,t}$.
+We then interpret parameter $\beta$ as the effect of changes in state funding at an Illinois public university, via state funding shocks, on an individual professor's outcome $Y_{j,t}$.

### 2023-05-30 text/sections/05-results.tex
Finished round of writing revisions.  Only spell-checking remains.

@@
 Lecturers per student is not shown to be correlated with total university revenues.
-The 2SLS columns show results after identifying state funding with the appropriation shock.
+The 2SLS columns show results after identifying state funding with the funding shock.
 All professor count per student increases by 0.53\% in response to a 10\% rise in total revenues.
@@
 For lecturers we see a negative effect, which lines up with two trends noted in \autoref{sec:trends}: public universities revenues (per student) are decreasing at the same time as employment of contingent lecturers increased rapidly.
-And yet, identifying the university revenues by the state appropriations shock increases the magnitude of the effect (Column 2) with respect to the association (Column 1), strengthening the case that the two trends are causally linked.\footnote{
+And yet, identifying the university revenues by the state funding shock increases the magnitude of the effect (Column 2) with respect to the association (Column 1), strengthening the case that the two trends are causally linked.\footnote{
     Between years 1990-1999 IPEDS provided count of professors by whether they had tenure (opposed to named position as in \autoref{tab:facultycount-shock-reg}), and \autoref{tab:tenurecount-shock-reg-fte} presents results of the respective regressions by tenure status.
@@
 In the same vein, when revenues are (not) bountiful, count of tenure-track and tenured professors per student increases (decreases).
-The mechanism that explains the changes in faculty composition are, however, not clear at this unit of analysis; individual data are needed to delve deeper.
+The mechanism that explains the changes in faculty composition are, however, not clear at this unit of analysis.
 
@@
 \autoref{tab:newhiresalaries-shock-illinois-rolling} presents 2SLS estimates for the elasticity of total salaries with respect to university revenues, among the sample with a year of starting employment after 2010.
-New hires of assistant, full, or administrative position professors are not related to revenues, via a shock to appropriations in their first-year of employment.
+New hires of assistant, full, or administrative position professors are not related to revenues, via a funding shock, in their first-year of employment.
 Yet, the starting salary of lecturers is 3.8\% higher given an increase in their hiring university's state funding of 10\%.

### 2023-05-30 text/sections/06-discussion.tex
Finished round of writing revisions.  Only spell-checking remains.

@@
 The results show that the recent stagnation in state funding for higher education has affected the composition of faculty within public universities, away from tenure-track and tenured professors towards contingent lecturers.
-At the same time, there are little discernible effects on individual professors hired in the years 2011-2021 at Illinois public universities, except for the salaries of first year lecturers.
+At the same time, there are little discernible effects on individual professors hired in the years 2011-2021 at Illinois public universities.
 The rate that faculty leave their university, and their rate of promotion between positions, are unaffected by the university revenues, so this leaves one primary channel to explain changes in faculty composition: hiring.
 
-Between academic years the number of professors at a university changes: professors leave an institution (retiring, hired elsewhere, etc.) and other join (they are newly hired, end a sabbatical, etc.).
-\autoref{sec:results-indiv} shows that most of these channels are unaffected, implying that falls in hiring for tenured-track and tenured faculty must explain most of the substitution towards lecturers away from tenured faculty.
-Notably, faculty hiring is already know to be heavily impacted by budget shocks.
-\cite{turner2014impact} documents the wide-spread practice of hiring freezes at universities in response to budget shocks around the 2008 recession.
-Throughout the last decade, multiple such measures were taken by Illinois public universities in response to their deteriorating finances \citep{furlough2010}.
-The University of Illinois\footnote{
-    The University of Illinois includes three public university campuses: Urbana-Champaign, Chicago, Springfield.
-}
-did not receive the allocated state appropriations from the state of Illinois on time, so enacted cost-cutting measures to stay fiscally solvent.
-In response, the university system placed a hold on all hiring for filling state-funded positions and promotions.
+The number of professors at a university changes between academic years: professors leave an institution (retiring, hired elsewhere, etc.), while others join (they are newly hired, end a sabbatical, etc.).
+Section~\ref{sec:results-indiv} shows that most of these channels are unaffected, implying that falls in hiring for tenured-track and tenured faculty must explain most of the substitution towards lecturers and away from tenured professors.
+Notably, faculty hiring is suspected to be impacted by changes in university funding; for example, \cite{turner2014impact} documents the wide-spread practice of hiring freezes at universities in response to budget shocks around the 2008 recession.
+Similar measures were taken by Illinois public universities in response to their deteriorating finances in the 2010s \citep{furlough2010}.
+The University of Illinois did not receive funding from the state of Illinois before the next fiscal year started, so enacted cost-cutting measures to stay fiscally solvent in the interim.
+As one such cost-cutting measure, the university system placed a freeze on all hiring for filling state-funded positions and promotions.
 
-The data used so far measure faculty count, so cannot disentangle which of these channels is most affected by changes in state funding at a public university, yet \cite{wapman2022quantifying} provide a measure of the aggregated total faculty hires at universities over the time period 2011--2021.\footnote{
-    \citep{wapman2022quantifying} made an aggregated version of their data open-source, making this sub-analysis possible.
-}
+The data used so far measure faculty count, so cannot disentangle which channels are most affected by changes in state funding at a public university.
+\autoref{fig:hiring-correlation.png} uses data on the total faculty hired at US public universities over 2011--2021, provided by \cite{wapman2022quantifying},\footnote{
+    \cite{wapman2022quantifying} provided an aggregated sample of their data as open-source, making this analysis possible.
+} showing a correlation between professor hiring rate and the total state funding per student.
 \begin{figure}[h!]
@@
 Universities with more state funding (for the entire decade) hired more professors (\autoref{fig:hiring-correlation.png}).
-Similarly, the funding shock instrument shows that an increase in 10\% state funding leads to 13\% more faculty hires at a university across the decade 2011--2021 (see \autoref{tab:hiring-shock-reg}).\footnote{
+Similarly, the funding shock IV model shows that a decrease of $-10$\% in state funding leads to $-13$\% fewer faculty hires per student across the decade 2011--2021 (see \autoref{tab:hiring-shock-reg} for full regression model results).\footnote{
     These results were produced by integrating the total count of faculty hires for 2010--2021 for the top-ranked 180 US universities with a sum of the funding variables, and then estimating the models specified in \autoref{sec:iv-model-uni}.
@@
 }
-It is clear that public universities' hiring is negatively impacted by falls in state funding, even using highly aggregated data for a total count of hiring across an entire decade.
+It is clear that public universities' hiring is negatively impacted by falls in state funding, even using this highly aggregated count of professor hiring across an entire decade.
 While this analysis strengthens the case negative effects on hiring plays, it is crude for measuring exact magnitudes for the effect on faculty hiring.

### 2023-05-30 text/sections/07-conclusion.tex
Finished round of writing revisions.  Only spell-checking remains.

@@
 
-Findings show that public universities have substituted away from tenure-track and tenured professors, towards non-tenured lecturers, in the face of persistent declines in state funding for higher education.
-The effects on incumbent professors in the state of Illinois are non-distinguishable from zero, which implies that the effect is driven by changes in university hiring of new professors.
-Each year, public universities are using more contingent lecturers to teach their education students, while private universities continue to employ more tenure-track and tenured professors than their public counterparts.
-Private universities have always had more tenured professor per student than public institutions, but each year the gap widens.
+Public universities have systematically substituted away from tenure-track and tenured professors, towards non-tenured lecturers, in the face of persistent declines in state funding for higher education.
+The effects on incumbent professors in the state of Illinois are non-distinguishable from zero, while suggestive information shows falls in hiring, which together implies that the changes in faculty composition are driven by reduced hiring of new professors.
+Each year, public universities are using more contingent lecturers to teach their students, while private universities continue to employ more tenure-track and tenured professors than their public counterparts, and each year the gap widens.
 
-While costs education have been rising in the US, thanks to multiple overlapping trends \citep{ehrenberg2012}, public universities have also dealt with declining state support.
-It is natural to expect that such headwinds will lead to systematic change at public universities, changes which are not neutral to the composition of faculty, or to their goals of research and education.
+While costs education have been rising in the US, public universities have also dealt with declining state funding.
+It is natural to expect that such headwinds will lead to systematic change at public universities, changes that affect their faculty, and limit the universities' goals in research and education.
 These results show large changes in faculty composition, and that stagnation in state support explains at least a third of the observed shift away from tenured professors and towards contingent lecturers.
 At the same time, private universities were not exposed to financial headwinds of the same magnitude or persistence.
-If public universities' hiring of faculty is disrupted by a fall in state funding, then we can expect that private universities will benefit (in a competitive sense) by being able to hire better professors from relatively resource constrained public universities.
-While public universities continue to educate the majority of higher education students in the US, we should worry about the effects that restricting their funding has on the composition of faculty, and the impact on higher education as a whole.
+While public universities continue to educate the majority of higher education students in the US, we should worry about the effects of restricting their funding has on faculty composition, research at public universities, and the impact on higher education as a whole.

### 2023-05-31 text/sections/00-abstract.tex
Fixed spelling.

@@
 This analysis relates the changes the faculty composition at US public universities to the fall in state funding for higher education, using a shift-share approach to identify causal effects.
-A fall of $-10$\% in university funding, via a state funding shock, decreases the number of assistant professors per student by $-1.4$\% and full professors by $-1.2$\%, yet increases the number of lecturers per student by 4.4\%; long-run estimates show the effects linger for multiple years after the initial funding shock.
+A fall of $-10$\% in university funding, via a state funding shock, decreases the number of assistant professors per student by $-1.4$\% and full professors by $-1.2$\%, yet increases the number of lecturers per student by 4.4\%; long-run estimates show that the faculty composition is affected both in the same year as the funding shock, and two to three years following.
 Analysis of all the professors at Illinois public universities shows that incumbent professors are not affected by substantial funding shocks over 2011-2021, with suggestive evidence showing that the faculty changes arose by limiting the hiring of new professors to replace leaving tenure-track/tenured professors.

### 2023-05-31 text/sections/04-empirical.tex
Fixed spelling.

@@
     \singlespacing
-    \centering
+    %\centering
     \caption{Mean Characteristics for Public Universities, by State Funding Shock Instrument.}
@@
     \label{tab:summary-quantiles}
+
     \footnotesize
+    %\vspace{1mm}
     \textbf{Note}:
@@
 I estimate the system by two stage least squares, including institution and year fixed effects, and investigate outcomes at the university-level to analyse effects on the university as a result of changes in revenues.
-Additionally, I estimate the model via local projections \citep{jorda2005,miller2022} to investigate whether the effects on faculty composition linger for multiple years after the original funding shock.\footnote{
-    The local projections method is an empirical method to estimate dynamic treatment effects when the treatment is not binary, so that time-series confounding (i.e., treatment in time $t$ is correlated with treatment in time $t-1$) is present, even for instrumental variable models.
-    Inference for this model is similar to that for an event study design, so I present estimates in graphical form (see \autoref
-    {fig:count-lp}).
+Additionally, I estimate the model via local projections \citep{jorda2005,miller2022} to investigate whether faculty composition is affected in the years after the initial funding shock (in addition to the initial year).\footnote{
+    The local projections method is an empirical model used to estimate dynamic treatment effects when the treatment is not binary, so that time-series confounding (i.e., treatment in time $t$ is correlated with treatment in time $t-1$) is present, even for instrumental variable models.
+    Inference in this method is for whether treatment $X_{i,t}$ affects the outcome $Y_{i,t+ k}$ in follow on years $k = 1, \hdots$, in addition to year $t$.
+    I present estimates in graphical form (see \autoref{fig:count-lp}).
 }

### 2023-05-31 text/sections/05-results.tex
Fixed spelling.

@@
 Faculty composition is not only impacted in the same years as a state funding shock, but also for a number years following.
-\autoref{fig:all-count-lp} shows local projection estimates, where the effect on total professor count per student persists at a magnitude of around 0.1 four years after the initial funding shock.
-These findings are in line with the observation that universities froze hiring for a couple of yearsin response to 2008 negative budget shocks, particularly for tenure-track positions \citep{turner2014impact}, and similarly from shocks to university endowments \citep{brown2014endowment}.
-For lecturers we see a negative effect, which lines up with two trends noted in \autoref{sec:trends}: public universities revenues (per student) are decreasing at the same time as employment of contingent lecturers increased rapidly.
+\autoref{fig:all-count-lp} shows local projection estimates, where the effect on total professor count per student is around 0.1 in both the same year year as the initial funding shock, and for the next 2 years
+These findings are in line with the observation that universities froze hiring for a couple of years in response to 2008 negative budget shocks, particularly for tenure-track positions \citep{turner2014impact}, and similarly from shocks to university endowments \citep{brown2014endowment}.
+For lecturers, we see a negative effect which lines up with two trends noted in \autoref{sec:trends}: public universities revenues (per student) are decreasing at the same time as employment of contingent lecturers increased rapidly.
 And yet, identifying the university revenues by the state funding shock increases the magnitude of the effect (Column 2) with respect to the association (Column 1), strengthening the case that the two trends are causally linked.\footnote{

### 2023-05-31 text/sections/07-conclusion.tex
standardised plot axies.

@@
 The effects on incumbent professors in the state of Illinois are non-distinguishable from zero, while suggestive information shows falls in hiring, which together implies that the changes in faculty composition are driven by reduced hiring of new professors.
-Each year, public universities are using more contingent lecturers to teach their students, while private universities continue to employ more tenure-track and tenured professors than their public counterparts, and each year the gap widens.
+Public universities are using more contingent lecturers to teach their students, while private universities continue to employ more tenure-track and tenured professors than their public counterparts, and each year the gap widens.
 

### 2023-06-02 text/sections/00-abstract.tex
Adjusted Table 3 + fixed spelling mistakes.

@@
 US universities have systematically substituted away from tenure-track/tenured professors towards contingent lecturers for the last thirty years, while at the same time states have limited how much they fund higher education.
-This analysis relates the changes the faculty composition at US public universities to the fall in state funding for higher education, using a shift-share approach to identify causal effects.
+This analysis relates the changes the faculty composition at US public universities to the fall in state funding for higher education, using a shift-share approach to instrument for state funding and identify causal effects.
 A fall of $-10$\% in university funding, via a state funding shock, decreases the number of assistant professors per student by $-1.4$\% and full professors by $-1.2$\%, yet increases the number of lecturers per student by 4.4\%; long-run estimates show that the faculty composition is affected both in the same year as the funding shock, and two to three years following.

### 2023-06-02 text/sections/02-data.tex
Fixed spelling.
Adjusted Table 3 + fixed spelling mistakes.

@@
     I combine the Urban Institute's 2018 compilation of IPEDS data for the years 1990--2017, and manually combine raw NCES data on 2018-2021 for all relevant variables.
-    Figures for enrolment, faculty counts and salaries come from the raw NCES version of IPEDS for all years, addressing inconsistencies in the Urban Institute's data formulation for these variables.
+    Figures for enrolment and faculty counts come from the raw NCES version of IPEDS for all years, addressing inconsistencies in the Urban Institute's data formulation for these variables.
 }
@@
     \singlespacing
-    \centering
+    %\centering
     \caption{IPEDS Summary Statistics, Public Universities Panel 1987--2021}
@@
 
-IPEDS provides information at the university level, yet lacks information on the distribution of salaries and professor count within a university.\footnote{
-    IPEDS provides total paid to salary and total faculty employed per university-year, so that I can investigate average professor salary per university-year with IPEDS data.
-    Unfortunately, this measure of professors' salaries is particularly crude in measuring on professors' salaries and other individual-outcomes; summary statistics on IPEDS data do not agree with trends in average professor salary over the sample time period \citep{aau2021survey}.
-    Similarly, IPEDS reports of the average salary paid to professors at private institutions also disagree with other sources, so are not considered in this analysis.
+IPEDS provides information at the university level, but only provides aggregated information within a university.\footnote{
+    IPEDS additionally gives a mean salary measure for faculty by rank (inconsistently) for each university.
+    Unfortunately, this measure of professors' salaries is particularly crude in measuring on professors' salaries and other individual-outcomes; summary statistics on IPEDS data do not agree with trends in average professor salary over the sample time period compared to summary statistics provided in \citealt{aau2021survey}.
+    Similarly, IPEDS reports of the average salary paid to professors at private institutions also disagree with other sources (including those calculated for in Illinois in the IBHED data later in this section) so are not considered in this analysis.
 }

### 2023-06-02 programs/analysis/ipeds-shock.R

@@ -156,6 +156,7 @@ reg.data %>%
         `4th` = `3`,
         `5th` = `4`) %>%
     xtable(type = "latex",
+        digits = digits.no,
         align = "llccccc") %>%
     print(
         sanitize.colnames.function = identity,

### 2023-06-05 text/sections/01-introduction.tex
Adjusted Table 3 + fixed misc spelling.
Fixed spelling in intro + labelled latest substantial revision as 5 June 2023.

@@
 Incumbent professors were not meaningfully affected, in terms of total salary, promotion rate, or rate of leaving the Illinois public university system.
-Yet the hiring rate for new professors at publicuniversities was negatively impacted by the falls in state funding (and by the funding shocks).
+Yet the hiring rate for new professors at public universities was negatively impacted by the falls in state funding (and by the funding shocks).
 This implies that faculty composition change arose by limiting the hiring of new professors; public universities increasingly hired contingent lecturers, and increasingly did not replace their retiring (or leaving) professors.
@@
 While it is not immediately clear whether a systematic substitution towards lecturers away from professors reduces teaching effectiveness, it is clear that stagnating state funding has negatively impacted student outcomes \citep{NBERw23736,NBERw27885}.
-This analysis focuses on a university's count of faculty per student, and considers faculty composition change as one of the mechanisms for how stagnating state funding for higher education transmits to students and has led to worsening education outcomes at public universities for the last thirty years.
+This analysis focuses on a university's count of faculty per student, where faculty composition is a possible mechanism for how the stagnating state funding for higher education transmits to students, and has led to worsening education outcomes at public universities for the last thirty years.

### 2023-06-07 text/paper.tex
Second major revision in progress.

@@
 \author{Senan Hogan-Hennessy\thanks{
-        I thank Professor Evan Riehl for his guidance on this project, without which this work would not have possible.
-        Additionally, I thank Michael Lovenheim, Ronald Ehrenberg, Douglas Miller, Francine Blau, Levon Barseghyan, Jake Meyer, and Ryan Dycus for various comments, as well as seminar participants at Cornell University (Autumn 2022) for helpful discussion.} \\
+        For helpful comments I thank
+        Levon Barseghyan,
+        Francine Blau,
+        Ryan Dycus,
+        Ronald Ehrenberg,
+        Michael Lovenheim,
+        Jake Meyer,
+        Douglas Miller,
+        and Evan Riehl,
+        as well as seminar participants at Cornell University (Autumn 2022) for helpful discussion.} \\
     \vspace{0.1cm}
@@
 % Title
-\title{Stagnating State Funding for Higher Education and its Effect on Faculty Composition at US Universities
+\title{Stagnating State Funding for Higher Education and its Effect on Faculty at US Universities
     }
@@
 \rhead{\today}
-\lhead{Stagnating State Funding and its Effect on Faculty Composition.}
+\lhead{Stagnating State Funding and its Effect on Faculty.}
 % Begin

### 2023-07-11 text/sections/00-abstract.tex
Major revision in progress.

--- 2023-05-30 text/sections/00-abstract.tex
@@
 US universities have systematically substituted away from tenure-track/tenured professors towards contingent lecturers for the last thirty years, while at the same time states have limited how much they fund higher education.
-This analysis relates the changes the faculty composition at US public universities to the fall in state funding for higher education, using a shift-share approach to identify causal effects.
-A fall of $-10$\% in university funding, via a state funding shock, decreases the number of assistant professors per student by $-1.4$\% and full professors by $-1.2$\%, yet increases the number of lecturers per student by 4.4\%; long-run estimates show the effects linger for multiple years after the initial funding shock.
+This paper connects the changes the faculty composition at US public universities to the fall in the states' funding for their public universities, using a shift-share approach to instrument for state funding and identify the causal effects of stagnating funding for higher education.
+A fall of $10$\% in funding for a public university, via a state funding shock, decreases the number of assistant professors per student by $1.4$\% and full professors by $1.2$\%, yet increases the number of lecturers per student by $4.4$\%; long-run estimates show that the faculty composition is affected both in the same year as the funding shock, and two to three years following.
 Analysis of all the professors at Illinois public universities shows that incumbent professors are not affected by substantial funding shocks over 2011-2021, with suggestive evidence showing that the faculty changes arose by limiting the hiring of new professors to replace leaving tenure-track/tenured professors.

### 2023-07-11 text/sections/01-introduction.tex
Major revision in progress.

@@
 The shift-share instrument interacts reliance on state funding in a base period with the entire state's total funding per higher education student for each year, to exploit changes in funding and how much each university historically relied on state funding for higher education.
+
+\textbf{TODO: Write section on the data sources.  See Lovenheimn (2020) on how to do it.}
 
@@
 Though, the economics literature is not decided on whether contingent lecturers lead to better or worse outcomes in higher education.
-Adjunct lecturers are generally more effective teachers than research professors at highly selective universities \citep{bettinger2010does,figlio2015tenure}, but students taught by part-time lecturers have worse student outcomes relative to full-time lecturers or faculty \citep{zhu2021limited,ehrenberg2005tenured}.
+Lecturers are more effective teachers than research professors at a highly selective university \citep{bettinger2010does,figlio2015tenure}, but students taught by part-time lecturers have worse student outcomes relative to full-time lecturers or faculty \citep{zhu2021limited,ehrenberg2005tenured}.
 
 While it is not immediately clear whether a systematic substitution towards lecturers away from professors reduces teaching effectiveness, it is clear that stagnating state funding has negatively impacted student outcomes \citep{NBERw23736,NBERw27885}.
-This analysis focuses on a university's count of faculty per student, where faculty composition is a possible mechanism for how the stagnating state funding for higher education transmits to students, and has led to worsening education outcomes at public universities for the last thirty years.
+\textbf{REWRITE:} where faculty composition is a possible mechanism for how the stagnating state funding for higher education transmits to students, and has led to worsening education outcomes at public universities for the last thirty years.
+
+\textbf{TO-DO:
+    This paper precedes as follows.
+}

### 2023-07-11 text/sections/02-data.tex
Major revision in progress.

@@
 IPEDS is a survey of higher educational institutions in the US, and legally requires institutions to participate in order to receive Federal Title IV student aid.\footnote{
-    This statement means that IPEDS does not necessarily cover the universe of US higher education institutions, yet in practice every public university and not for-profit four-year institution is represented.
+    IPEDS does not necessarily cover the universe of US higher education institutions, yet in practice every public university and not for-profit four-year institution is represented.
 }
@@
 
-\begin{table}[h!]
-    \singlespacing
-    %\centering
-    \caption{IPEDS Summary Statistics, Public Universities Panel 1987--2021}
-    \makebox[\textwidth][c]{\input{tables/ipeds-summary-fte.tex}}
-    \label{tab:ipeds-summary}
-\end{table}
-
 IPEDS provides information at the university level, but only provides aggregated information within a university.\footnote{
@@
 
-The analysis sample represents 16,932 professors in the year 2010 and 15,352 in the year 2021, with summary statistics presented in \autoref{tab:illinois-summary}.
-Additionally, further analysis using the first year of a professor's employment focuses on the subset of professors with observed year of hiring between the years 2011 and 2021, representing 1,778 professors in 2011, and 9,099 in 2021.\footnote{
-    Summary statistics for this group, which over-represents lecturers compared to assistant and full professors, can be seen in \autoref{tab:illinois-summary-rolling}.
-    The reasons for considering the sub-sample with identified year of hire in the range 2011-2021 are explained in \autoref{sec:iv-model-indiv}.
-}
-
-\begin{table}[h!]
-    \singlespacing
-    \centering
-    \caption{IBHED Summary Statistics, Professor Panel 2010--2021.}
-    \makebox[\textwidth][c]{\input{tables/illinois-summary.tex}}
-    \label{tab:illinois-summary}
-\end{table}
+The Illinois sample represents 16,932 professors in the year 2010 and 15,352 in the year 2021, with summary statistics presented in \autoref{tab:illinois-summary}.
+Additionally, further analysis using the first year of a professor's employment focuses on the subset of professors with observed year of hiring between the years 2011 and 2021, representing 1,778 professors in 2011, and 9,099 in 2021.
 
@@
 
-\begin{figure}[!htbp]
-    \centering
-    \singlespacing
-    \caption{Mean Total Funding among Public Universities, by Year.}
-    \begin{subfigure}[b]{0.495\textwidth}
-        \centering
-        \caption{Total, \$ 2021 CPI-U.}
-        \includegraphics[width=\textwidth]{figures/mean-funding-total.png}
-        \label{fig:mean-funding-total}
-    \end{subfigure}
-    \begin{subfigure}[b]{0.495\textwidth}
-        \centering
-        \caption{Per Student, \$ 2021 CPI-U.}
-        \includegraphics[width=\textwidth]{figures/mean-funding-fte.png}
-        \label{fig:mean-funding-fte}
-    \end{subfigure}
-    \label{fig:funding}
-\end{figure}
-
-\begin{figure}[!htbp]
-    \centering
-    \singlespacing
-    \caption{Total Student enrolment, by University Sector and Year.}
-    \begin{subfigure}[b]{0.495\textwidth}
-        \centering
-        \caption{Total Enrolment, Nation-wide.}
-        \includegraphics[width=\textwidth]{figures/enrollment-total.png}
-        \label{fig:enrollment-total}
-    \end{subfigure}
-    \begin{subfigure}[b]{0.495\textwidth}
-        \centering
-        \caption{Mean Enrolment, per University.}
-        \includegraphics[width=\textwidth]{figures/enrollment-mean.png}
-        \label{fig:enrollment-mean}
-    \end{subfigure}
-    \label{fig:enrolment}
-\end{figure}
-
 At the same time, student enrolment at public universities rose precipitously.
@@
 This means that revenues per student have stagnated across all measures (seen in \autoref{fig:mean-funding-fte}), and particularly fallen from a mean of \$11,000 per student in 1990 to less than \$8,000 per student in 2021.
-
-\begin{figure}[h!]
-    \centering
-    \singlespacing
-    \caption{Trends in Mean Student Enrolment per Professor, by University Sector and Faculty level.}
-    \begin{subfigure}[b]{0.495\textwidth}
-        \centering
-        \caption{Lecturer.}
-        \includegraphics[width=\textwidth]{figures/lecturer-fte-perprof.png}
-        \label{fig:lecturer-fte-perprof}
-    \end{subfigure}
-    \begin{subfigure}[b]{0.495\textwidth}
-        \centering
-        \caption{Assistant.}
-        \includegraphics[width=\textwidth]{figures/assistant-fte-perprof.png}
-        \label{fig:assistant-fte-perprof}
-    \end{subfigure}
-    \begin{subfigure}[b]{0.495\textwidth}
-        \centering
-        \caption{Full.}
-        \includegraphics[width=\textwidth]{figures/full-fte-perprof.png}
-        \label{fig:full-fte-perprof}
-    \end{subfigure}
-    \begin{subfigure}[b]{0.495\textwidth}
-        \centering
-        \caption{All.}
-        \includegraphics[width=\textwidth]{figures/all-fte-perprof.png}
-        \label{fig:all-fte-perprof}
-    \end{subfigure}
-    \label{fig:fte-perprof}
-\end{figure}
 
@@
 
-The rate that Illinois funds higher education has stagnated to a similar extent as the average for the entire country.
-\autoref{fig:illinois-funding} shows the trends for funding sources among the seven Illinois public university campuses, including the falling share of state funding and substitution towards tuition revenue.
+Illinois funding for higher education has stagnated, and has experienced serious declines in the decade 2010--2021 --- similar in magnitude to the nation-wide decline over 1990--2021.
+State funding among the seven Illinois public university campuses has fallen by over 50\% for the last decade, on both an absolute and per student basis (see \autoref{fig:illinois-funding}).
+Similar to the figures for the rest of the country, there has been a corresponding substitution towards tuition revenue.
 
-\begin{figure}[h!]
-    \centering
-    \singlespacing
-    \caption{Mean Funding Sources among Illinois Public Universities, by Year.}
-    \begin{subfigure}[b]{0.495\textwidth}
-        \centering
-        \caption{Total, \$ 2021 CPI-U.}
-        \includegraphics[width=\textwidth]{figures/illinois-funding-total.png}
-        \label{fig:illinois-funding-total}
-    \end{subfigure}
-    \begin{subfigure}[b]{0.495\textwidth}
-        \centering
-        \caption{Per Student, \$ 2021 CPI-U.}
-        \includegraphics[width=\textwidth]{figures/illinois-funding-fte.png}
-        \label{fig:illinois-funding-fte}
-    \end{subfigure}
-    \label{fig:illinois-funding}
-\end{figure}
-
-The decade of 2010--2021 has also seen yearly variation in its revenues, particularly around 2016.
+2010--2021 saw not only a decline, but also large yearly funding variation, particularly around 2016.
 In the calendar year 2015, partisan disagreements between the democratic legislature and republican governor led to the 2016 fiscal year starting with no state budget.

### 2023-07-11 text/sections/03-concepts.tex
Major revision in progress.

--- 2023-03-19 text/sections/03-concepts.tex
@@
 
-Most universities hire a number of new professors every year, either to expand their departments or to replace leaving professors.
-The number of professors hired in each year is usually highest among non-tenured adjunct faculty, as these instructors are mostly hired on short-term contingent contracts, so that they are less costly to hire (or to not reinstate) in response to yearly changes in teaching needs.
-Secondly, tenure-track assistant professors are hired to replace leaving assistant professors, and are most often hired with a four to six year contract and agreement for tenure consideration.
-Lastly, tenured professors have undergone review and have successfully secured a full-time appointment at their university with no expiration date; tenured professors may be hired from outside a university, though this position has the lowest yearly hiring rate among most universities.
-It is common for universities to restrict their faculty hiring in the face of financial shocks.
-Since the yearly hiring rate is highest among lecturers and assistant professors, financial shocks may lower the number of lecturer and assistant professors at public universities the most, if faculty hiring is affected by falls in funding.
+It is a priori unclear that stagnating state support and changes in faculty composition at public universities would be causally linked.
+Over the 21st Century there have been drastic changes to college selectivity, where the top universities have become more selective while the average university less selective \citep{hoxby2009changing}.
+While the average public university is becoming less selective, it is possible that their most productive or research-focused faculty more often move to more selective and prestigious (and often private) institutions, so that over time public universities (on average) substitute towards contingent lecturers.
+This is one way in which the trends may be concurrent, but not causal.
 
-Additionally, faculty salaries may by affected by changes in public university finances.
+Employment composition changes likely arise via differences in hiring rate across levels of professor seniority.
+Universities hire new professors most years, either to expand their departments or to replace leaving professors, but not hire at the same rate for each level of seniority.
+The number of professors hired in each year is usually highest among non-tenured adjunct faculty, as these instructors are mostly on short-term or contingent contracts, so  are less costly to hire (or to let go) in response to yearly changes.
+Tenure-track assistant professors are mostly hired with a four to six year contract, and formal agreement for tenure consideration at the end of this term.
+Lastly, tenured professors have successfully secured a full-time appointment at their university with no expiration date; this position has the lowest yearly hiring rate among most universities.
+Similarly, there is highest turn-over (including leaving the university) among lecturers, and lowest among full professors.
+The differences in yearly hiring and turnover rate mean that if hiring is affected by state funding, then there will be faculty composition change in the years following budget cuts.
+
+Faculty salaries may be affected by changes in public university finances.
 When the university has a lower budget for its yearly hiring, it may respond by lowering the salary they offer to new hires.
-This possible effect may not be the same across each position of professor; tenured faculty are often hired away from another university, so that new tenured professor hires may be less likely to accept a lower salary offer from a public university.\footnote{
+This possible effect may not be the same across each position of professor; tenured faculty are often hired away from another university, so that new tenured professor hires may be less likely to accept the lower salary offers from public universities.\footnote{
     Faculty often consider their outside options, as in offers for employment and/or promotion from other universities.
@@
 }
-Faculty are not paid the same by position, so that there are financial consequences for a university promoting their own faculty.
-In response to tightening fiscal restraints, universities may be less likely to grant tenure to assistant professors (as tenured professors generally receive a salary raise and long term contract), or to promote tenured professors from associate to full professor.
+Additionally, it is not immediately clear which faculty (among those already hired) will be most affected by the changes in their university's funding.
+Yet, there is one empirical fact worth noting: the average assistant or tenured professor earns more than double that of a lecturer (see \autoref{tab:illinois-summary}).
+If a public university's primary obligation is to teach, and they must fulfil this objective with less resources, then they may substitute away from tenure-track and tenured professors towards lecturers.
 
-It is not immediately clear which effects will dominate, or whether tenured professors vs lecturers will be affected more by changes in their university's funding.
-Yet, there is one empirical fact worth noting: a mean assistant or tenured professor earns more than double that of an adjunct lecturer per year (see \autoref{tab:illinois-summary}), and the tenure contract means universities must pay a higher salary for a guaranteed number years, more so than for contingent lecturers.
-If a public university's primary obligation is to teach, and they must fulfil this objective with fewer and fewer resources, then they may substitute away from tenure-track and tenured professor towards contingent lecturers to achieve this obligation.\footnote{
-    That is, substituted away from tenure-track and tenured professors according to the salary ratio, and the relative productivity between these two positions.
-}
+Similarly, its is not clear when the effects of funding short-falls will be realised.
+University departments often employ faculty on multiple-year contracts (e.g., the tenure contract), and individual professors' decisions to retire, move university, or leave academia are multiple-year commitments, so that it may take multiple years for university funding to change the faculty composition, or trickle down to faculty.
+For example, university departments coordinate hiring in year-long cycles, so that a budget short-fall in year $t$ may have no effect on the in-progress hiring committee and resulting hiring decisions, yet the hiring cycle in year $t+1$ may be postponed or cancelled in response.
+As such, this paper investigates the dynamics of state funding shocks' effects on faculty, for the years following a university funding shock.

### 2023-07-11 text/sections/04-empirical.tex
Major revision in progress.

--- 2023-05-30 text/sections/04-empirical.tex
@@
 
-Universities employ professors, so it stands to reason that if a university experiences a fall in one of their revenue sources --- similar to a demand shock in the private sector --- then they may respond by adjusting their employee composition.
-
-Na\"ively we can analyse the relationship between professor-outcomes $Y_{i,t}$ at university $i$ in year $t$ as a result of state funding per student $X_{i,t}$.
-This analysis focuses on the outcome of professor count per student within each university-year.
-So that $\beta$ represents the association with these outcomes and state funding per student,\footnote{
-    Note that dividing by student count also implicitly controls for the size of the university, so that this model implicitly accounts for yearly variation in professor count and university revenues arising from growth in a university.
-}
-including fixed effects to control for effects specific to the university and year.
-Log$(.)$ transforming the variables improves the interpretability of $\beta$ as an elasticity for professor count per student with respect to state funding per student.
-\begin{equation}
-    \label{eqn:naivereg}
-    Y_{i,t} = \alpha_i + \gamma_t + \beta X_{i,t} + \epsilon_{i,t}
-\end{equation}
-
-Yet a university's finances are not exogenous to state decisions for support of higher education, or exogenous to internal institutional decisions.
+This paper identifies the effects of changes in state funding on faculty.
+However, a university's state funding is not exogenous to state decisions for support of higher education, or exogenous to internal institutional decisions.
 Instead, the state government and university administration undertake a complex process of allotting resources across multiple different priorities, including instruction, research, or between departments.
@@
 
-This analysis uses a shift-share instrument to address endogeneity concerns for the amount of funding a state gives to each of its public universities.
+This analysis uses a shift-share instrument to address endogeneity concerns for the amount of funding a state allots to each of its public universities.
 \cite{NBERw23736,NBERw27885} develop the instrument for public university finances by exploiting a shift-share instrument for changes in state-level funding interacted with university reliance on state funding in a base period.
@@
     \label{eqn:public-instrument}
-    Z_{i,t} &\coloneqq - \log \left[
+    Z_{i,t} &\coloneqq - \left[
     \left( \frac{\text{Total State Funding}_{s(i),t}}{\text{Student Population}_{s(i),t}} \right)
@@
 The system exploits the fact that institutions who rely on state funding more will be affected by state funding shocks.
-$Z_{i,t}$ is the instrument for (log) state funding for institution $i$ in year $t$, interacting the average funding for universities in state $s(i)$ with reliance on state funding relative to total revenues, averaged across the base years 1990--1993.\footnote{
+$Z_{i,t}$ is the instrument for state funding for institution $i$ in year $t$, interacting the average funding for universities in state $s(i)$ with reliance on state funding relative to total revenues, averaged across the base years 1990--1993.\footnote{
     1990--1993 are defined as data for public university finance data are most comparable (i.e. without many missing values) beginning in 1990.
@@
 The instrument approach relies on the conditional independence assumption, that the instrument is independently assigned to the universities.
-While this assumption is fundamentally untestable, we see that universities are by-in-large smaller (in terms of enrolment, total revenues, and professor count) at the top of the state funding shock distribution.
+While this assumption is fundamentally untestable, we see that universities are by-in-large smaller (in terms of enrolment, total revenues, and professor count) at the top of the state funding shock distribution (see~\autoref{tab:summary-quantiles}).
 Though on a per student basis, there is little difference across the distribution (except in the outcomes under consideration).
+The state funding shock instrument is positively associated with the total enrolment and total amount of state funding for each university, in both \$ and $\log$/percentage change terms, while other the other sources of finances fopr the university are not associated with the funding shock.
+So that the other sources of finances are not clear confounders for the instrumental variables strategy, as they exhibit balance with respect to the funding shock instrument \citep{pei2019poorly}.
 Yearly and individual fixed effects are included in regressions throughout to implicitly condition on mean differences between universities and years, leading to the assumption of conditional independence of the instrument with respect to state funding for each public university.
 
-\begin{table}[h!]
-    \singlespacing
-    \centering
-    \caption{Mean Characteristics for Public Universities, by State Funding Shock Instrument.}
-    \makebox[\textwidth][c]{\input{tables/summary-quantiles.tex}}
-    \label{tab:summary-quantiles}
-    \footnotesize
-    \textbf{Note}:
-    The column labelled ``1st'' refers to the mean for all university-year observations in the first quintile (bottom 20\%) of the funding shock distribution, and so on.
-    The mean of the funding shock within each quintile is shown in the first row of the table.
-\end{table}
-
-The state funding shock is an instrument for the amount of state funding for each university in each year.  
+The state funding shock is an instrument the level of state funding for each university in each year.  
 The first-stage is then as follows, including institution and year fixed effects, where $X_{i,t}$ represents the amount of state funding divided by the number of full-time students attending the university.
@@
 
-\begin{table}[h!]
-    \singlespacing
-    \centering
-    \caption{First Stage Estimates, for State Funding by funding shock.}
-    \makebox[\textwidth][c]{\input{tables/firststage-reg-fte.tex}}
-    \label{tab:firststage-reg}
-    \begin{flushleft}
-        \footnotesize
-        \textbf{Note}: Standard errors are clustered at the state-year level.
-    \end{flushleft}
-\end{table}
-
-The shift-share instrument performs excellently as an instrument for universities' yearly state funding.
-\autoref{tab:firststage-reg} presents results of the first-stage regression, separately with and without a control for tuition revenue per student, plus institution and year fixed effects.\footnote{
-    Representations for frequentist significance levels (i.e. 10, 5, 1\% etc.) are omitted here, and in all following tables.
+\textbf{TO-DO:
+    Add description the dollar-by-dollar, and log terms, of the first stage.
 }
+The shift-share instrument has a strong first stage for universities' yearly state funding.
+\autoref{tab:firststage-reg} presents results of the first-stage regression, separately with and without a control for tuition revenue per student, plus institution and year fixed effects.
 We note columns (1) and (2) estimate that a funding shock (per student in the entire state) of 10\% is associated with 9.8\% change in state funding per student at the university; the instrument is strong, and we note similarity in estimates with and without inclusion of fixed effects.
@@
 Column (1) represents the estimates for \autoref{eqn:firststage} with fixed effects, omitting the tuition revenue control, and is the preferred form that I proceed with.
-Lastly, the state funding shock affects university revenues for multiple years after the initial shock (see \autoref{fig:firststage-lp}).
-We see a decaying, yet real and positive, effect of the shock on revenues for the next 10 years, illustrating that the state funding shock is a strong, if fading, instrument for state funding per student and justifying its use in later local projections models.
 
@@
 
-\subsection{Instrumental Variables Model, University-Level}
+\subsection{Instrumental Variables Model}
 \label{sec:iv-model-uni}
 
-The primary empirical model combines the instrumental variable for state funding shocks with the empirical model for the effects of state funding at the university-level --- i.e., parameter $\beta$ in the following.\footnote{
-    It is important to note the treatment effect isolated here; the instrumental variables approach identifies the local average treatment effects, one specific to the instrument.
-    So we interpret this treatment effect as a university's response in employment count and average salaries to state funding changes, changes specific to state funding shocks, among the complier group --- i.e., universities who respond to funding shocks who would not have made faculty-outcome changes absent the funding shock.
-    Also, we assume that no universities' total revenues increase in response to negative state funding shock (monotonicity).
+I use the instrument defined in \autoref{eqn:public-instrument} to overcome the endogeneity concerns for state funding to each public university, so the primary empirical model is an instrumental variables model.
+\autoref{eqn:firststage} is the first stage for exogenous variation in the state funding for university $i$ in year $t$, and \autoref{eqn:secondstage} the second stage for the effect of state funding on faculty outcomes.\footnote{
+    It is important to note the treatment effect isolated here; the instrumental variables approach identifies the local average treatment effect, one weighted to level of exposure when treatment is continuous as is the case here.
+    So we interpret this treatment effect as a weighted average of effects on faculty at public universities to state funding changes, specific to state funding shocks, among the complier group --- i.e., among universities who respond to funding shocks and would not have made faculty-outcome changes absent the funding shock.
+    Also, we assume that no universities state funding increased in response to the negative state funding shock (monotonicity).
 }
-\begin{eqnarray}
-    \label{eqn:secondstage1}
-    X_{i,t} &=& \eta_i + \zeta_t + \delta Z_{i,t} + \epsilon_{i,t} \\
-    \label{eqn:secondstage2}
-    Y_{i,t} &=& \alpha_i + \gamma_t + \beta \widehat X_{i,t} + \varepsilon_{i,t}
-\end{eqnarray}
-I estimate the system by two stage least squares, including institution and year fixed effects, and investigate outcomes at the university-level to analyse effects on the university as a result of changes in revenues.
-Additionally, I estimate the model via local projections \citep{jorda2005,miller2022} to investigate whether the effects on faculty composition linger for multiple years after the original funding shock.\footnote{
-    The local projections method is an empirical method to estimate dynamic treatment effects when the treatment is not binary, so that time-series confounding (i.e., treatment in time $t$ is correlated with treatment in time $t-1$) is present, even for instrumental variable models.
-    Inference for this model is similar to that for an event study design, so I present estimates in graphical form (see \autoref
-    {fig:count-lp}).
+\begin{equation}
+    \label{eqn:secondstage}
+    Y_{i,t} = \alpha_i + \gamma_t + \beta \widehat X_{i,t} + \varepsilon_{i,t}
+\end{equation}
+I estimate the system by two stage least squares, including institution and year fixed effects.\footnote{
+    Note that dividing by student count also implicitly controls for the size of the university, so that this model implicitly accounts for yearly variation in professor count and university revenues arising from a university's size of growth/decline.
 }
-Regarding outcomes, I focus on the composition of the professors employed at the university by analysing (log) count of professors per student by each university.
+$Y_{i,t}$ represents faculty outcomes for university $i$ in year $t$, $\alpha_i, \gamma_t$ university and year fixed effects, and $\widehat X_{i,t}$ state funding for university $i$ estimated in first stage \eqref{eqn:firststage}.
 
-%\begin{figure}[H]
-%    \centering
-%    \singlespacing
-%    \caption{Instrument Variables Model for University Finances and Faculty Composition.}
-%    \begin{tikzpicture}
-%        \node[state] (instrument) at (0,0) {$Z$};
-%        \node at (-2.6,0.5){State};
-%        \node at (-1.9,0){appropriation};
-%        \node at (-2.6,-0.5){shock};
-%        \node[state] (endogenous) [right=of instrument] {$X$};
-%        \node at (2,1.5) {University};
-%        \node at (2,1) {revenues};
-%        \node[state] (outcome) [right=of endogenous] {$Y$};
-%        \node at (5.2,0.25) {Faculty};
-%        \node at (5.6,-0.25) {Composition};
-%        % Relevance condition
-%        \path (instrument) edge (endogenous);
-%        \path (endogenous) edge (outcome);
-%        % Exclusion Restriction
-%        \path (instrument) edge[bend right=60] (outcome);
-%        \node[color=red] at (2,-1.2) {X};
-%    \end{tikzpicture}
-%    \label{fig:SCM-ivmodel}
-%\end{figure}
-
-\subsection{Instrumental Variables Model, Individual Professor-Level}
-\label{sec:iv-model-indiv}
-
-This analysis additionally uses data on individual professors in the Illinois university system, to investigate the effects of changes in university revenues on the individual professors at the universities.
-Redefining the level of outcome requires adjustment to the empirical approach, leveraging variation in university finances for the years after a professor joins the university.\footnote{
+Additionally, I investigate the effect of changes in state funding among incumbent professors.
+Incumbent professors are faculty who are already employed at the university; state funding changes may affect faculty who are already at the university (beyond affecting whether they get hired), so I base the instrument in the year that the professor was hired, and include fixed effects for the hiring year.\footnote{
     This formulation follows that presented by \cite{NBERw27885}, where individual student outcomes are analysed via variation in state funding after their freshman-year.
     This contrasts with \autoref{sec:iv-model-uni} and \cite{NBERw23736}, where the unit of analysis is the university-year, where base year 1990 is more appropriate.
+    See \autoref{sec:iv-model-indiv} for the instrument and second-stage specification.
 }
+The instrument exogeneity and exclusion follows the same argument as above, where the Illinois legislature did not target any single campus for funding cuts and no single Illinois campus takes the majority of state funding.
 
-\autoref{eqn:rolling-instrument} defines a rolling-share variant of the instrument, $\tilde Z_{j,t}$, where the university's state funding share exposure is based in the year a professor joins the university --- and not the base period 1990--1993.
-$j$ indexes each professor in year $t$, $\tau(j)$ for the year the professor first joins their institution.\footnote{
-    Identifying $\tau(j)$ is possible for $j$ by restricting to all professors hired 2011-2021 --- i.e., in the years after the start of the full panel.
-    It is not possible to discern the hiring year for professors who  were hired in the years preceding 2011, and so the entire sample is only possible to analyse using the base-share in years 1990-1993 formulation (e.g., \autoref{tab:facultysalaries-shock-illinois}, \ref{tab:promotion-shock-illinois}).
+\subsection{Percent Change and Rates of Substitution}
+\label{sec:empirical-substituion}
+
+It is not a priori clear which units are appropriate for this analysis;
+does it make sense to consider state funding in purely dollar amounts per student, or in percent change rate?
+The funding shock instrument is a strong predictor for the average university's level of state funding in either unit; a funding shock of \$-1,000 per student in the entire state leads to around \$-1,176 per student at the university, while a funding shock of -10\% leads to around -9.77\% less state funding per year.
+Yet, the level of state funding (and the outcome variables) vary greatly between states for the unit of analysis.
+For example, the average Illinois public university receives \$10,709 in state funding per student in 1990 and  \$6,713 in 2021 (a fall of over 30\%), while California went from \$19,224 per student to \$12,915 in the same time span (a fall of 37\%).
+While most states are not exactly the same to California and Illinois, this is example is telling for the phenomenon of stagnating state funding:
+states vary in how much the funding higher education in 1990, but most have experienced a decline on order of 30\%, so that the stagnation in state funding has been a percent change trend over this time period.
+As such, I focus on regression specifications where the explanatory and outcome variables are $\log$ transformed (unless otherwise mentioned).
+
+The estimand of a $\log-\log$ specification is a funding elasticity; a 1\% fall in state funding $X_{i,t}$ caused a $\beta$\% fall in outcome $Y_{i,t}$, as in \autoref{eqn:secondstage}.
+If the estimates are negative, then conceptually we consider the outcome to be inferior for a university's budgeting decisions with respect to state funding: if they are funded with more state resources they spend less on outcome $Y_{i,t}$.
+Conversely, outcome $Y_{i,t}$ is normal if the coefficient is positive.
+
+The funding elasticities can be used to recover the rates of substitution between two outcomes.
+For example, write $Y^1$ for the number of lecturers per student at the universities, and $Y^2$ for the number of full professors.
+I use the above approaches to estimate the funding elasticities, where $\% \Delta$ denotes percent change.
+\[ \beta_1 = \frac{\% \Delta Y^1}{\% \Delta X}
+\text{, and }
+\beta_2 = \frac{\% \Delta Y^2}{\% \Delta X} \]
+As such, it is possible to recover the elasticity for substitution between lecturers and full professors by the universities via the respective funding elasticities.
+\[ \frac{\% \Delta Y^1}{\% \Delta Y^2}
+= \frac{\% \Delta Y^1 / \% \Delta X}{\% \Delta Y^2 / \% \Delta X}
+= \frac{\beta_1}{\beta_2} \]
+I present results for the rates of substitution between different levels of faculty by this approach, dividing the relevant coefficient estimates and presenting standard errors calculated by a non-parametric bootstrap.
+% https://stackoverflow.com/questions/63777368/computing-the-standard-error-when-dividing-coefficients-of-different-regressions
+
+\subsection{Effects in Years After the Funding Shock}
+
+The effects of a change in the universities funding may not be immediate, and faculty in may be effected multiple years after a funding shock.
+Yet, the funding shock instrument is significantly auto-correlated year-on-year; a large state-level funding shock in year $t$ is also likely to experience a large funding shock in year $t-1$.
+Similarly, funding shocks to higher education in year are highly correlated with state funding per student in the 5 years before and after the original shock in IPEDS data (see \autoref{fig:lag-firststage}).
+Thus, a linear model correlated state funding in year $t$ with outcomes in year $t$ will suffer from time-series confounding.
+
+I employ a local projections approach to model how faculty outcomes in years $t+k$ are affected by state funding in year $t$, for future years $k > 0$ \citep{jorda2005}.
+The local projections method is an empirical model used to estimate dynamic treatment effects when the treatment is not binary, so that time-series confounding (i.e., treatment in time $t$ is correlated with treatment in time $t-1$) is present, even for instrumental variable models \citep{miller2022}.
+This method estimates the effect of treatment $X_{i,t}$ on outcome $Y_{i, t+ k}$ in follow-on years $k > 0$, while accounting for the auto-correlation between the other years.\footnote{
+    The funding shock has a persistent effect on state funding, multiple years after the initial funding shock, so that the instrument is similarly strong for the local projections method (see \autoref{fig:firststage-lp}).
 }
-\begin{align}
-    \label{eqn:rolling-instrument}
-    \tilde Z_{j,t} &\coloneqq - \log \left[
-    \left( \frac{\text{Total State Funding}_{s(j),t}}{\text{Student Population}_{s(j),t}} \right)
-    \left( \frac{\text{State Funding}_{\tau(j)}}{\text{Total Revenues}_{i,\tau(j)}} \right) \right]
-\end{align}
-
-This approach leverages an insight, made available by level of the data: that an individual professor is affected by changes in university revenues after they have joined the university.\footnote{
-    Notice that \autoref{sec:iv-model-uni} considers the number of professors employed by the university; whether a professor becomes employed at the university is likely affected by that university's finances.
-    The formulation in this section does not consider whether the professor joins the university, instead taking as given that the professor is employed at the university, and then projects the effect on the individual.
+I present estimates in graphical form, where the x-axis represents the year relative to the funding shock (i.e., $x = 0$ represents the year of the state funding shock) and the y-axis represents the estimated effect of the funding shock on that year's outcome.\footnote{
+    See \autoref{fig:count-lp} for the graphical format, and the corresponding note for further explanation.
 }
-Exogeneity and relevance of the rolling-share instrument, $\tilde Z_{j,t}$, follows the same reasoning as that for the base-share instrument, $Z_{i,t}$, discussed in \autoref{sec:approp-shocks}.\footnote{
-    The base-share instrument is appropriate for some outcomes with the individual Illinois professors, where appropriate (\autoref{tab:facultysalaries-shock-illinois}, \ref{tab:promotion-shock-illinois}).
-}
-We satisfy the assumptions for exogeneity by noting that none of the Illinois public campuses take the majority of state funding, and that the identification strategy relies on exogeneity in changes in state funding to individual professor-outcomes, following the year they joined the university.\footnote{
-    Additionally, within-institution changes resulting from share reliance on state funding may be correlated with unobserved changes in the outcomes, so that \cite{NBERw27885} note the importance of controlling for the base share and state student population.
-    The formulation here implicitly controls for these factors via the fixed effects; results are relatively similar while including these controls with and without including fixed effects, and so are omitted.
-}
-\autoref{tab:firststage-illinois} presents results of the first stage estimation, showing that the instrument is strong in the same way as that for the university-level outcomes (\autoref{tab:firststage-reg}), with very similar estimates for the association between funding shocks and state funding.
-
-The instrumental variables model is then defined as follows, where $i(j)$ refers to the institution that professor $j$ is employed at, and $Y_{j,t}$ to individual-level outcomes total salary, rate of promotion, and propensity to leave the Illinois public university system.
-The system includes fixed effects for the institution and first year of employment.\footnote{
-    The instrument varies by institution, based in the year of first employment, so that these are the corresponding fixed effects and levels of clustered standard errors.
-}
-\begin{eqnarray}
-    \label{eqn:secondstage1_indiv}
-    X_{i(j),t} &=& \theta_{i(j)} + \phi_{\tau(j)} + \delta \tilde Z_{i(j),t} + \epsilon_{i(j),t} \\
-    \label{eqn:secondstage2_indiv}
-    Y_{j,t} &=& \mu_{i(j)} + \nu_{\tau(j)} + \beta \widehat X_{i(j),t} + \varepsilon_{j,t}
-\end{eqnarray}
-We then interpret parameter $\beta$ as the effect of changes in state funding at an Illinois public university, via state funding shocks, on an individual professor's outcome $Y_{j,t}$.

### 2023-07-11 text/sections/05-results.tex
Major revision in progress.

--- 2023-05-30 text/sections/05-results.tex
@@
 
-\subsection{University-Level}
+Changes in state funding have clear effects on the composition of faculty public universities employ.
+\autoref{tab:facultycount-shock-reg} presents OLS and IV estimates for the effecet of a change in state funding on the number of faculty at the university, individually by position (and for the total faculty), in both count of professors and in percent terms for faculty per student ratio.
+An extra \$1,000 in state funding per student leads to the average university employing 6 fewer lecturers, with no discernible effect on the count of total professors (\autoref{tab:facultycount-shock-reg}, Panel A).
+In percentage terms, a 10\% increase in state funding lead to a 4.4\% decrease in the number of lecturers per student, an increase of 14\% for both assistant and full professors, leading to a 0.6\% increase in the total number of faculty per student.
+In the state of Illinois, using the count of professors represented in the IBHED databse, state funding is correlated with the number of faculty, but lacks precision to identify the exact effects given the 144 university-year observations in the panel for 2010--2021.
 
-Changes in state funding have clear effects on the composition of faculty public universities employ.
-\autoref{tab:facultycount-shock-reg} presents OLS and IV estimates, where the outcome is (log) count of professors employed at the university, separated by position, and for all professors (Columns 7, 8).
-For all professors we see a positive correlation between university revenues and professor count per student (the OLS columns), where an increase in university total revenues by 10\% is associated with an increase in professor count per student of 0.83\%, yet negatively correlated with lecturer count of -1.3\%.
-Lecturers per student is not shown to be correlated with total university revenues.
-The 2SLS columns show results after identifying state funding with the funding shock.
-All professor count per student increases by 0.53\% in response to a 10\% rise in total revenues.
-The effect is driven by increases in the count of assistant and full professors (i.e.
-tenure-track and tenured) per student, where assistant professor count per student increase by 1.4\%, and full professor 1.2\%.
-The count of lecturers per student decreases by 4.4\% in response to a (positive) 10\% shock in university revenues.
-
-\begin{table}[h!]
-    \singlespacing
-    \centering
-    \caption{OLS and 2SLS Estimates for University Faculty Composition.}
-    \makebox[\textwidth][c]{\input{tables/facultycount-shock-reg-fte.tex}}
-    \begin{flushleft}
-        \footnotesize
-        \textbf{Note}: Standard errors are clustered at the state-year level.
-    \end{flushleft}
-    \label{tab:facultycount-shock-reg}
-\end{table}
+Incumbent professors are relatively unaffected by changes in the state funding for their university in the years following their hire.
+\autoref{tab:faculty-shock-illinois-rolling} show estimates of a 1\% change in state funding per student on faculty salaries, rate of exiting the Illinois public university system, and rate of promotion; there are little effects discernible from zero, so that faculty who are already at the university have little of the effects of funding cuts passed on to them.
 
 Faculty composition is not only impacted in the same years as a state funding shock, but also for a number years following.
-\autoref{fig:all-count-lp} shows local projection estimates, where the effect on total professor count per student persists at a magnitude of around 0.1 four years after the initial funding shock.
-These findings are in line with the observation that universities froze hiring for a couple of yearsin response to 2008 negative budget shocks, particularly for tenure-track positions \citep{turner2014impact}, and similarly from shocks to university endowments \citep{brown2014endowment}.
-For lecturers we see a negative effect, which lines up with two trends noted in \autoref{sec:trends}: public universities revenues (per student) are decreasing at the same time as employment of contingent lecturers increased rapidly.
-And yet, identifying the university revenues by the state funding shock increases the magnitude of the effect (Column 2) with respect to the association (Column 1), strengthening the case that the two trends are causally linked.\footnote{
-    Between years 1990-1999 IPEDS provided count of professors by whether they had tenure (opposed to named position as in \autoref{tab:facultycount-shock-reg}), and \autoref{tab:tenurecount-shock-reg-fte} presents results of the respective regressions by tenure status.
-    Results are relatively similar, yet imprecise from the much smaller number of observations.
-}
+\autoref{fig:all-count-lp} shows local projection estimates, where the elasticity of total professor count per student in year $t+k$ with respect to state funding in year $t$ is around 0.1, for follow-on years $k = 0, \hdots, 3$.
+These findings are in line with the observation that universities froze hiring for a couple of years in response to 2008 negative budget shocks, particularly for tenure-track positions \citep{turner2014impact}, and similarly from shocks to university endowments \citep{brown2014endowment}.
+For lecturers, we see a negative effect which lines up with two trends noted in \autoref{sec:trends}: funding for public universities (per student) is decreasing while relative usage of  lecturers increased rapidly.
 
-Decreases in university revenues (via state funding shocks) affect faculty composition for at most 4 years after the initial revenue shock.
+Decreases in state funding (via funding shocks) affects not only the total counts of professors, but also the composition in the years after thee funding shock.
 A 10\% decrease in state funding increases count of lecturers per student by 4\% for the first two years after the initial shock.
@@
 count of tenured professors decreases 9 to 6\% three years later.
-Together, the total count of professors per student decreases around 1\% for three years after the initial funding shock, and the effect is not distinguishable from zero four years after.
-\begin{figure}[h!]
-    \centering
-    \singlespacing
-    \caption{Local Projection Estimates for Professor Count per Student, by Professor Group.}
-    \begin{subfigure}[b]{0.495\textwidth}
-        \centering
-        \caption{Lecturers.}
-        \includegraphics[width=\textwidth]{figures/lecturer-count-lp.png}
-        \label{fig:lecturer-count-lp}
-    \end{subfigure}
-    \begin{subfigure}[b]{0.495\textwidth}
-        \centering
-        \caption{Assistant Professors.}
-        \includegraphics[width=\textwidth]{figures/assistant-count-lp.png}
-        \label{fig:assistant-count-lp}
-    \end{subfigure}
-    \begin{subfigure}[b]{0.495\textwidth}
-        \centering
-        \caption{Full Professors.}
-        \includegraphics[width=\textwidth]{figures/full-count-lp.png}
-        \label{fig:full-count-lp}
-    \end{subfigure}
-    \begin{subfigure}[b]{0.495\textwidth}
-        \centering
-        \caption{All Professors.}
-        \includegraphics[width=\textwidth]{figures/all-count-lp.png}
-        \label{fig:all-count-lp}
-    \end{subfigure}
-    \label{fig:count-lp}
-\end{figure}
-\autoref{fig:count-lp} represents these estimates, using the local projection method, as a series of impulse responses for each outcome of professor count.
+Together, the total count of professors per student decreases around 1\% for three years after the initial funding shock, and the effect is not distinguishable from zero after this point.
+\autoref{fig:lecturer-count-lp}, \ref{fig:assistant-count-lp}, \ref{fig:full-count-lp} represents these estimates, using the local projection method, as a series of impulse responses for each outcome of professor count.
+These findings show that public universities increase (decrease) their count of tenure-track and tenured professors per student in years when revenues are more (less) plentiful, possibly by increasing hiring intensity or efforts to maintain incumbent professors.
+In the same vein, when revenues are bountiful, count of tenure-track and tenured professors per student increases --- and the opposite in years of funding cuts.
 
-These findings show that public universities increase (decrease) their count of tenure-track and tenured professors per student in years when revenues are more (less) plentiful, possibly by increasing hiring intensity or efforts to maintain incumbent professors.
-In the same vein, when revenues are (not) bountiful, count of tenure-track and tenured professors per student increases (decreases).
-The mechanism that explains the changes in faculty composition are, however, not clear at this unit of analysis.
+\subsection{Rates of Substitution}
+\label{sec:results-substituion}
 
-\subsection{Individual Professor Outcomes}
-\label{sec:results-indiv}
-
-I first examine whether a shock to the university's revenues affects the salaries of professors at the university.
-The first outcome is first-year salary for each professor by their position;
-\autoref{tab:newhiresalaries-shock-illinois-rolling} presents 2SLS estimates for the elasticity of total salaries with respect to university revenues, among the sample with a year of starting employment after 2010.
-New hires of assistant, full, or administrative position professors are not related to revenues, via a funding shock, in their first-year of employment.
-Yet, the starting salary of lecturers is 3.8\% higher given an increase in their hiring university's state funding of 10\%.
-
-\begin{table}[h!]
-    \singlespacing
-    \centering
-    \caption{2SLS Estimates for Faculty Salaries, in First-Year, at Illinois Universities.}
-    \makebox[\textwidth][c]{\input{tables/newhiresalaries-shock-illinois-rolling.tex}}
-    \begin{flushleft}
-        \footnotesize
-        \textbf{Note}: Standard errors are clustered at the institution and first year of employment level. 
-    \end{flushleft}
-    \label{tab:newhiresalaries-shock-illinois-rolling}
-\end{table}
-
-Regarding total salary for every year of employment, \autoref{tab:facultysalaries-shock-illinois-rolling} presents estimates for the elasticity of total salary with respect to university revenues, for the entire sample of professors hired after 2010.\footnote{
-    Recall that this is a subsample of the entire data, that the rolling shift-share instrument is only defined for the sample with an observed hiring date in the years 2011-2021.
-}
-The elasticity is non-distinguishable from zero among any position, implying that long-term salaries are not related to state funding.
-\autoref{tab:facultysalaries-shock-illinois} presents the same model, for the sample of all professors and using the instrument based in years 1990-1993, similarly finding no relationship.
-
-\begin{table}[h!]
-    \singlespacing
-    \centering
-    \caption{2SLS Estimates for Faculty Salaries at Illinois Universities.}
-    \makebox[\textwidth][c]{\input{tables/facultysalaries-shock-illinois-rolling.tex}}
-    \begin{flushleft}
-        \footnotesize
-        \textbf{Note}: Standard errors are clustered at the institution-year level.
-    \end{flushleft}
-    \label{tab:facultysalaries-shock-illinois-rolling}
-\end{table}
-
-Note that professor position (assistant, associate, full) is also an outcome, thanks to the promotion channel: a university may be less able to promote their professors to higher paying positions when their revenues are constrained.
-\autoref{tab:promotion-shock-illinois-rolling} shows estimates where the outcome is rate of promotion within each professor position.
-The lecturer position describes the sample of professors who were ever listed as a non-tenure track lecturer and the binary for promotion equals one in a year they achieved an assistant professor position; assistant the same for the assistant to associate professor promotion; associate same for associate to full promotion.
-There is no discernible relationship between revenues and promotion among any position.\footnote{
-    \autoref{tab:promotion-shock-illinois} shows estimates using the base-year instrument among the entire sample, finding the same results.
-}
-Furthermore, local projection estimates (\autoref{fig:promoted-illinois-lp}) show that the there is no promotion effect, among all professors, for any of the following years either.
-Lastly, \autoref{tab:facultyleaving-shock-illinois-rolling}, \ref{tab:facultyleaving-shock-illinois} describes estimates for rate of exit from the employing university among Illinois professors, and finds no relationship with university revenues.
+Write here about the log results, including a figure for the university funding elasticities of faculty demand.
+Describe here the relevance of the elasticities, and describe the estimates for elasticity of substitution.

### 2023-07-11 text/sections/06-discussion.tex
Major revision in progress.

@@
 \section{Discussion}
+
+Replace focus on hiring channel, describing how the results can be interpretted.
+
+Additionally, put into further context of the trends in university competition \citep{urquiola2020markets}, and XYZ \citep{ehrenberg2012american}.
+
+\textbf{Moved from Concepts section:}
+This paper shows that stagnating in state funding for public universities cause the changes in faculty composition among public universities, explaining about 40\% of the observed substitution towards contingent lecturers, so that it is possible that other systematic changes in higher education (such as changing selectivity) may explain the rest of the variation.
+
+\textbf{TODO: Rewrite this all, putting  a minimal focus on faculty hiring, and focusing on the mechanisms instead (mechanisms for decline in public unis).}
 
@@
 } showing a correlation between professor hiring rate and the total state funding per student.
-\begin{figure}[h!]
-    \centering
-    \singlespacing
-    \caption{State Funding and Faculty Hired at Public Universities, Total for 2011--2021.}
-    \includegraphics[width=0.8\textwidth]{figures/hiring-correlation.png}
-    \label{fig:hiring-correlation.png}
-\end{figure}
 

### 2023-07-11 programs/analysis/individual-shock-rolling.R

@@ -1,764 +1,0 @@
-#!/usr/bin/R
-## Senan Hogan-Hennessy, 3 August 2022
-## IV for Prof's outcomes, using Illinois data.
-## Using a rolling-share variant of the instrument.
-print(c(Sys.time(), Sys.Date()))
-library(tidyverse) # Functions for data manipulation and visualization
-library(lfe) # Functions for fast linear models with IV + FEs
-library(plm) # Functions for panel data
-library(car) # Function for F stat regarding IV models
-# My custom flavour of Stargazer TeX tables:
-# devtools::install_github("shoganhennessy/stargazer")
-library(stargazer) # TeX tables
-set.seed(47)
-# This file follows an adjusted Deming Walters (2017, p.10) approach to
-# appropriations shock instrument.
-
-# Define number of digits in tables and graphs
-digits.no <- 3
-
-# Size for figures
-fig.width <- 10
-fig.height <- fig.width * 0.85
-
-
-# Load data --------------------------------------------------------------------
-
-# Data on Illinois Professors
-illinois.data <- read_csv("../../data/states/illinois-professors-anonymised.csv")
-
-# Data from IPEDS, with relevant variables
-ipeds.data <- read_csv("../../data/urban-ipeds/urban-clean-publicunis.csv") %>%
-    select(
-        # Identifiers and controls
-        unitid,
-        year,
-        state,
-        # Appropriations to the uni
-        totalrevenues_real,
-        nonauxrevenues_real,
-        stateappropriations_real,
-        tuitionrev_real,
-        # Uni spending.
-        nonauxspending_real,
-        # Uni enrollment
-        enrollment_reported,
-        enrollment_fte,
-        # Appropriation shocks Per student
-        appropriationshock_perEnroll_rolling,
-        appropriationshock_perFTE_real,
-        appropriationshock_peruni_real,
-        # State-level values
-        allstate_stateappropriations_real,
-        stateEnroll_count)
-
-
-# Clean data -------------------------------------------------------------------
-
-# Join IPEDS data to the individual level
-reg.data <- illinois.data %>%
-    # Join the IPEDS finance data.
-    left_join(ipeds.data, by = c("unitid", "year"))
-
-# Generate a Professor's first year
-reg.data <- reg.data %>%
-    arrange(name, year, unitid) %>%
[…]

### 2023-07-11 programs/analysis/individual-shock.R

@@ -1,852 +1,0 @@
-#!/usr/bin/R
-## Senan Hogan-Hennessy, 3 August 2022
-## IV for Prof's outcomes, using Illinois data.
-## Using a base-share version of the instrument.
-print(c(Sys.time(), Sys.Date()))
-library(tidyverse) # Functions for data manipulation and visualization
-library(lfe) # Functions for fast linear models with IV + FEs
-library(plm) # Functions for panel data
-library(car) # Function for F stat regarding IV models
-# My custom flavour of Stargazer TeX tables:
-# devtools::install_github("shoganhennessy/stargazer")
-library(stargazer) # TeX tables
-set.seed(47)
-# This file follows an adjusted Deming Walters (2017, p.10) approach to
-# appropriations shock instrument.
-
-# Define number of digits in tables and graphs
-digits.no <- 3
-
-# Size for figures
-fig.width <- 10
-fig.height <- fig.width * 0.85
-
-
-# Load data --------------------------------------------------------------------
-
-# Data on Illinois Professors
-illinois.data <-
-    read_csv("../../data/states/illinois-professors-anonymised.csv")
-
-# Data from IPEDS, with relevant variables
-ipeds.data <- read_csv("../../data/urban-ipeds/urban-clean-publicunis.csv") %>%
-    select(
-        # Identifiers and controls
-        unitid,
-        year,
-        state,
-        # Appropriations to the uni
-        totalrevenues_real,
-        nonauxrevenues_real,
-        stateappropriations_real,
-        tuitionrev_real,
-        # Uni spending.
-        nonauxspending_real,
-        # Uni enrollment
-        enrollment_reported,
-        enrollment_fte,
-        # Appropriation shocks Per student
-        appropriationshock_perEnroll_real,
-        appropriationshock_perFTE_real,
-        appropriationshock_peruni_real,
-        # State-level values
-        allstate_stateappropriations_real,
-        stateEnroll_count)
-
-
-# Clean data -------------------------------------------------------------------
-
-# Join IPEDS data to the individual level
-reg.data <- illinois.data %>%
-    # Join the IPEDS finance data.
-    left_join(ipeds.data, by = c("unitid", "year"))
-
-# Generate a Professor's first year
-reg.data <- reg.data %>%
-    arrange(name, year, unitid) %>%
[…]

### 2023-07-11 programs/analysis/ipeds-shock.R

@@ -18,8 +18,19 @@ set.seed(47)
 digits.no <- 3
 
 # Size for figures
-fig.width <- 9
+fig.width <- 10
 fig.height <- fig.width * 0.85
+
+# Define a function to standardise the dplyr lag function
+# --- negative numbers mean a lag of t time periods, positive a lead of t.
+lag_lead <- function(column, t) {
+    if (t < 0) {
+        lag_column <- dplyr::lag(column, -t)
+    } else if (t >= 0) {
+        lag_column <- dplyr::lead(column, t)
+    }
+    return(lag_column)
+}
 
 
 # Load data sources ------------------------------------------------------------
@@ -83,7 +94,11 @@ reg.data <- reg.data %>%
         !is.na(enrollment_reported), enrollment_reported > 0,
         !is.na(all_prof_count),
         !is.na(stateappropriations_real),
-        !is.na(appropriationshock_perEnroll_real))
+        !is.na(appropriationshock_perEnroll_real)) %>%
+    # Restrict to uni-years with all professor counts.
+    filter(!is.na(lecturer_prof_count), lecturer_prof_count > 0) %>%
+    filter(!is.na(assistant_prof_count), assistant_prof_count > 0) %>%
+    filter(!is.na(full_prof_count), full_prof_count > 0)
 
 
 # Save memory by removing initally loaded files
@@ -109,7 +124,7 @@ reg.data %>%
         digits.extra = 0,
         covariate.labels = c(
             "Enrolment",
-            "State appropriations (millions 2021 USD)",
+            "State Funding (millions 2021 USD)",
             "Total revenues (millions 2021 USD)",
             "Non-institutional revenues (millions 2021 USD)",
             "Lecturers count",
@@ -121,54 +136,27 @@ reg.data %>%
         type = "text",
         out = "../../text/tables/ipeds-summary-fte.tex")
 
-# Summary Statistics for the Quantiles of the shift-share instrument.
-quantile.count <- 5
-reg.data %>%
-    mutate(instrument_quantile = round((quantile.count - 1) *
-        ecdf(appropriationshock_perEnroll_real)(
-            appropriationshock_perEnroll_real))) %>%
-    group_by(instrument_quantile) %>%
-    summarise(
-        `State Funding Shock, \\$ per student` = round(mean(appropriationshock_perEnroll_real, an.rm = TRUE)),
[…]

### 2023-07-11 programs/run.sh

@@ -17,18 +17,22 @@ cd ..
 cd analysis
 # Generate graphs motivating the introduction.
 R CMD BATCH --no-save ipeds-public-private.R
-# Analysis and regressions at the institution-level.
+# Analysis and regressions at the institution-level, IPEDS
 R CMD BATCH --no-save ipeds-shock.R
-# Analysis and regressions at the individual-level, with base-year shock IV.
-R CMD BATCH --no-save individual-shock.R
-# Analysis and regressions at the individual-level, with rolling shock IV.
-R CMD BATCH --no-save individual-shock-rolling.R
+# Analysis and regressions at the institution-level, IPEDS, with count outcomes.
+R CMD BATCH --no-save ipeds-shock-rawcounts.R
+# Analysis and regressions at the individual-level, IBHED, base-year shock IV.
+R CMD BATCH --no-save ibhed-shock.R
+# Analysis and regressions at the individual-level, IBHED, with count outcomes.
+R CMD BATCH --no-save ibhed-shock-rawcounts.R
+# Analysis and regressions at the individual-level, IBHED, rolling shock IV.
+R CMD BATCH --no-save ibhed-shock-rolling.R
 # Analysis and regressions for rate of faculty hiring.
 R CMD BATCH --no-save hiring-shock.R
 cd ..
 
-# Compile paper, with outputs of progams as inputs for TeX files
+# Compile paper, with outputs from R scripts above as inputs for TeX files
 cd ../text
 latexmk -pdf paper.tex
 latexmk -c
-cp paper.pdf ../state-faculty-composition-2023.pdf
+cp paper.pdf ../state-funding-faculty-2023.pdf

### 2023-07-13 text/sections/01-introduction.tex
re-writing

@@
 that high school students exposed to similar public school funding shocks are less likely to attend college \citep{jackson2021school}, and that these shocks induce public universities to rely more on tuition as a source of revenue \citep{bound2019public}.
+\textbf{Faculty outcomes surveyed by \cite{ehrenberg2003studying}.}
 \cite{NBERw27885} use the same methodology to show that increases in state funding lead to degree lower completion time and later-life debt for students.\footnote{
@@
 Though, the economics literature is not decided on whether contingent lecturers lead to better or worse outcomes in higher education.
-Lecturers are more effective teachers than research professors at a highly selective university \citep{bettinger2010does,figlio2015tenure}, but students taught by part-time lecturers have worse student outcomes relative to full-time lecturers or faculty \citep{zhu2021limited,ehrenberg2005tenured}.
+\textbf{Beef up this part to focus more on negative effects on students.}
+Lecturers are more effective teachers than research professors at a highly selective university \citep{bettinger2010does,figlio2015tenure}, but students taught by part-time lecturers have worse student outcomes relative to full-time lecturers or faculty \citep{zhu2021limited,ehrenberg2005tenured,jaeger2011examining}.
+
+\textbf{Tenure research \citep{mcpherson1999tenure,mcpherson1983economics}}
 

### 2023-07-13 text/sections/04-empirical.tex
re-writing

@@
 
+The shift-share instrument has a strong first stage for universities' yearly state funding.
+\autoref{tab:firststage-reg} presents results of the first-stage regression, separately with and without a control for tuition revenue per student, plus institution and year fixed effects;
+Panel A shows estimates of a funding shock of \$1 per student in the state on state funding per student at the university, and Panel B the \% increase effects of a 1\% increase in the shock per student on state funding per student.
 \textbf{TO-DO:
-    Add description the dollar-by-dollar, and log terms, of the first stage.
+    Add description the dollar-by-dollar, of the first stage.
 }
-The shift-share instrument has a strong first stage for universities' yearly state funding.
-\autoref{tab:firststage-reg} presents results of the first-stage regression, separately with and without a control for tuition revenue per student, plus institution and year fixed effects.
 We note columns (1) and (2) estimate that a funding shock (per student in the entire state) of 10\% is associated with 9.8\% change in state funding per student at the university; the instrument is strong, and we note similarity in estimates with and without inclusion of fixed effects.
@@
 
-\subsection{Percent Change and Rates of Substitution}
-\label{sec:empirical-substituion}
-
 It is not a priori clear which units are appropriate for this analysis;
 does it make sense to consider state funding in purely dollar amounts per student, or in percent change rate?
-The funding shock instrument is a strong predictor for the average university's level of state funding in either unit; a funding shock of \$-1,000 per student in the entire state leads to around \$-1,176 per student at the university, while a funding shock of -10\% leads to around -9.77\% less state funding per year.
+The funding shock instrument is a strong predictor for the average university's level of state funding in either unit; a funding shock of \$1,000 per student in the entire state leads to around \$1,176 per student at the university, while a funding shock of -10\% leads to around -9.77\% less state funding per year.
 Yet, the level of state funding (and the outcome variables) vary greatly between states for the unit of analysis.
@@
 states vary in how much the funding higher education in 1990, but most have experienced a decline on order of 30\%, so that the stagnation in state funding has been a percent change trend over this time period.
-As such, I focus on regression specifications where the explanatory and outcome variables are $\log$ transformed (unless otherwise mentioned).
-
-The estimand of a $\log-\log$ specification is a funding elasticity; a 1\% fall in state funding $X_{i,t}$ caused a $\beta$\% fall in outcome $Y_{i,t}$, as in \autoref{eqn:secondstage}.
-If the estimates are negative, then conceptually we consider the outcome to be inferior for a university's budgeting decisions with respect to state funding: if they are funded with more state resources they spend less on outcome $Y_{i,t}$.
-Conversely, outcome $Y_{i,t}$ is normal if the coefficient is positive.
-
-The funding elasticities can be used to recover the rates of substitution between two outcomes.
-For example, write $Y^1$ for the number of lecturers per student at the universities, and $Y^2$ for the number of full professors.
-I use the above approaches to estimate the funding elasticities, where $\% \Delta$ denotes percent change.
-\[ \beta_1 = \frac{\% \Delta Y^1}{\% \Delta X}
-\text{, and }
-\beta_2 = \frac{\% \Delta Y^2}{\% \Delta X} \]
-As such, it is possible to recover the elasticity for substitution between lecturers and full professors by the universities via the respective funding elasticities.
-\[ \frac{\% \Delta Y^1}{\% \Delta Y^2}
-= \frac{\% \Delta Y^1 / \% \Delta X}{\% \Delta Y^2 / \% \Delta X}
-= \frac{\beta_1}{\beta_2} \]
-I present results for the rates of substitution between different levels of faculty by this approach, dividing the relevant coefficient estimates and presenting standard errors calculated by a non-parametric bootstrap.
-% https://stackoverflow.com/questions/63777368/computing-the-standard-error-when-dividing-coefficients-of-different-regressions
+As such, I include regression specifications where the explanatory and outcome variables are $\log$ transformed, and refer to these when stating results in percent change terms.
 
 \subsection{Effects in Years After the Funding Shock}
+\label{sec:empirical-substituion}
 

### 2023-07-13 text/sections/05-results.tex
re-writing

@@
 Changes in state funding have clear effects on the composition of faculty public universities employ.
-\autoref{tab:facultycount-shock-reg} presents OLS and IV estimates for the effecet of a change in state funding on the number of faculty at the university, individually by position (and for the total faculty), in both count of professors and in percent terms for faculty per student ratio.
+\autoref{tab:facultycount-shock-reg} presents OLS and IV estimates for the effect of a change in state funding on the number of faculty at the university, individually by position (and for the total faculty), in both count of professors and in percent terms for faculty per student ratio.
 An extra \$1,000 in state funding per student leads to the average university employing 6 fewer lecturers, with no discernible effect on the count of total professors (\autoref{tab:facultycount-shock-reg}, Panel A).
@@
 
-Incumbent professors are relatively unaffected by changes in the state funding for their university in the years following their hire.
-\autoref{tab:faculty-shock-illinois-rolling} show estimates of a 1\% change in state funding per student on faculty salaries, rate of exiting the Illinois public university system, and rate of promotion; there are little effects discernible from zero, so that faculty who are already at the university have little of the effects of funding cuts passed on to them.
-
 Faculty composition is not only impacted in the same years as a state funding shock, but also for a number years following.
-\autoref{fig:all-count-lp} shows local projection estimates, where the elasticity of total professor count per student in year $t+k$ with respect to state funding in year $t$ is around 0.1, for follow-on years $k = 0, \hdots, 3$.
-These findings are in line with the observation that universities froze hiring for a couple of years in response to 2008 negative budget shocks, particularly for tenure-track positions \citep{turner2014impact}, and similarly from shocks to university endowments \citep{brown2014endowment}.
-For lecturers, we see a negative effect which lines up with two trends noted in \autoref{sec:trends}: funding for public universities (per student) is decreasing while relative usage of  lecturers increased rapidly.
-
-Decreases in state funding (via funding shocks) affects not only the total counts of professors, but also the composition in the years after thee funding shock.
+\autoref{fig:count-lp} shows local projection estimates, as a series of impulse responses, where the estimate is the elasticity for professor count per student in year $t+k$ with respect to state funding in year $t$ for follow-on years $k = 0, \hdots, 10$.
 A 10\% decrease in state funding increases count of lecturers per student by 4\% for the first two years after the initial shock.
@@
 Together, the total count of professors per student decreases around 1\% for three years after the initial funding shock, and the effect is not distinguishable from zero after this point.
-\autoref{fig:lecturer-count-lp}, \ref{fig:assistant-count-lp}, \ref{fig:full-count-lp} represents these estimates, using the local projection method, as a series of impulse responses for each outcome of professor count.
-These findings show that public universities increase (decrease) their count of tenure-track and tenured professors per student in years when revenues are more (less) plentiful, possibly by increasing hiring intensity or efforts to maintain incumbent professors.
-In the same vein, when revenues are bountiful, count of tenure-track and tenured professors per student increases --- and the opposite in years of funding cuts.
 
-\subsection{Rates of Substitution}
-\label{sec:results-substituion}
+Incumbent professors are relatively unaffected by changes in the state funding for their university in the years following their hire.
+\autoref{tab:faculty-shock-illinois-rolling} show estimates of a 1\% change in state funding per student on faculty salaries, rate of exiting the Illinois public university system, and rate of promotion.
+The estimates are not discernible from zero, so that faculty who are already at the university (incumbent faculty) seem to have no effects of state funding cuts passed on to them.
 
-Write here about the log results, including a figure for the university funding elasticities of faculty demand.
-Describe here the relevance of the elasticities, and describe the estimates for elasticity of substitution.
+%These findings show that public universities increase (decrease) their count of tenure-track and tenured professors per student in years when revenues are more (less) plentiful, possibly by increasing hiring intensity or efforts to maintain incumbent professors.
+%In the same vein, when funding increases, count of tenure-track and tenured professors per student increases --- and the opposite in years of funding cuts.
+%These findings are in line with the observation that universities froze hiring for a couple of years in response to 2008 negative budget shocks, particularly for tenure-track positions \citep{turner2014impact}, and similarly from shocks to university endowments \citep{brown2014endowment}.
+%For lecturers, we see a negative effect which lines up with two trends noted in \autoref{sec:trends}: funding for public universities (per student) is decreasing while relative usage of lecturers increased substantially.
+
+The composition of faculty changed at public universities 1990--2017 thanks to stagnating state funding, but there are little discernible effects on incumbent faculty.
+This may imply that the rate that public universities hired faculty was impacted by state funding cuts.
+I take recent data on the total faculty hired at US public universities over 2011--2021 (provided by \citealt{wapman2022quantifying}), to show that universities with lower state funding (per student) also hired fewer professors in the same time period --- shown in \autoref{fig:hiring-correlation}.
+Similarly, the funding shock IV model shows that a $10$\% cut in per student state funding leads to $13$\% fewer faculty hires per student across the decade 2011--2021.\footnote{
+    See \autoref{sec:appendix-hiring} for further details.
+}
+These results provide limited evidence that changes in faculty composition arose by disrupting hiring of new professors at public universities, corresponding to the results that incumbent faculty were not by meaningfully effected by the funding cuts.

### 2023-07-13 text/sections/06-discussion.tex
re-writing

--- 2023-05-24 text/sections/06-discussion.tex
@@
 
-The results show that the recent stagnation in state funding for higher education has affected the composition of faculty within public universities, away from tenure-track and tenured professors towards contingent lecturers.
-At the same time, there are little discernible effects on individual professors hired in the years 2011-2021 at Illinois public universities, except for the salaries of first year lecturers.
-The rate that faculty leave their university, and their rate of promotion between positions, are unaffected by the university revenues, so this leaves one primary channel to explain changes in faculty composition: hiring.
+Universities received less state funding, relative to increasing enrolment, and reacted by increasing their reliance on lecturers relative to assistant and full professors in both the short and medium-run; these results imply that lecturers can be considered substitutes for professors in the tenure-system.\footnote{
+    Additional results in Appendix \autoref{sec:appendix-substitution} formalises the marginal rate of substitution between lecturers and professors on tenure-track, showing how assistant and full professors were substituted for lecturers while there is no substitution between assistant and full professors.
+}
+A public university's state funding per student fell by an average of 36\%, and the number of lecturers per student increased by 113\% (fell by 17\% for assistant and 23\% full professors).
+The estimates of elasticity for faculty count with respect to state funding imply that stagnating state funding explains about 40\% of the observed substitution towards lecturers and away from tenure-track and tenured faculty.
 
-Between academic years the number of professors at a university changes: professors leave an institution (retiring, hired elsewhere, etc.) and other join (they are newly hired, end a sabbatical, etc.).
-\autoref{sec:results-indiv} shows that most of these channels are unaffected, implying that falls in hiring for tenured-track and tenured faculty must explain most of the substitution towards lecturers away from tenured faculty.
-Notably, faculty hiring is already know to be heavily impacted by budget shocks.
-\cite{turner2014impact} documents the wide-spread practice of hiring freezes at universities in response to budget shocks around the 2008 recession.
-Throughout the last decade, multiple such measures were taken by Illinois public universities in response to their deteriorating finances \citep{furlough2010}.
-The University of Illinois\footnote{
-    The University of Illinois includes three public university campuses: Urbana-Champaign, Chicago, Springfield.
+These effects are persistent: public higher education in the US was not exposed to one large instance of funding cuts, but systematic funding short-falls across the last three decades.
+The long-term estimates (i.e., \autoref{fig:count-lp}) show that state funding cuts have effects on universities for multiple years after the initial funding cut, and these effects are compounded by the fact that public universities received persistent funding cuts for so many years.
+The end result is large changes in faculty composition, the long-run trend in stagnating state funding leading to a long-run substitution towards lecturers at public universities.
+
+At the same time, incumbent faculty in the state of Illinois are unaffected by the state funding cuts in terms of salary, promotion rate, and rate of leaving their faculty position.
+These results imply that composition change arose by lower hiring of tenure-track faculty in the public university system.
+The private university system did not grow in any corresponding amount, with the clear implication that there are fewer tenure-track openings at US universities.\footnote{
+    While the number of PhD graduates has continued to rise \citep{aau2021survey}.
 }
-did not receive the allocated state appropriations from the state of Illinois on time, so enacted cost-cutting measures to stay fiscally solvent.
-In response, the university system placed a hold on all hiring for filling state-funded positions and promotions.
+Securing academic employment on the tenure-track is more selective than ever, so that more PhD graduates will end up teaching as lecturers or leaving academia.
+% Note: unclear effects on PhD students choosing to not enter acadmeia, or go towards private unis
+% Would be possible to study with data on PhD students, and whether they are more/less likely to enter academia after budget cuts to their PhD institution.
 
-The data used so far measure faculty count, so cannot disentangle which of these channels is most affected by changes in state funding at a public university, yet \cite{wapman2022quantifying} provide a measure of the aggregated total faculty hires at universities over the time period 2011--2021.\footnote{
-    \citep{wapman2022quantifying} made an aggregated version of their data open-source, making this sub-analysis possible.
-}
-\begin{figure}[h!]
-    \centering
-    \singlespacing
-    \caption{State Funding and Faculty Hired at Public Universities, Total for 2011--2021.}
-    \includegraphics[width=0.8\textwidth]{figures/hiring-correlation.png}
-    \label{fig:hiring-correlation.png}
-\end{figure}
+The results here are squarely in line with theories explaining the increased competition and inequality between US universities.
+\cite{urquiola2020markets} interprets the success of research in US higher education as a result of free market policies, where increased competition in a free market in the early 1900s led to improvements in university research among US private universities.
+These changes have not just affected research, they have affected education, too.
+The rate of selectivity among US universities has become increasingly polarised, where top (and mostly private) universities with the most resources have become more selective at the undergraduate level \citep{hoxby2009changing}, and increasingly dominate academia at the graduate level \citep{wapman2022quantifying}.
+Public universities are by-in-large not selective, so are caught in a relative decline when their selective and private peers are increasingly more selective, and better performing in terms of research and student outcomes.
 
-Universities with more state funding (for the entire decade) hired more professors (\autoref{fig:hiring-correlation.png}).
-Similarly, the funding shock instrument shows that an increase in 10\% state funding leads to 13\% more faculty hires at a university across the decade 2011--2021 (see \autoref{tab:hiring-shock-reg}).\footnote{
-    These results were produced by integrating the total count of faculty hires for 2010--2021 for the top-ranked 180 US universities with a sum of the funding variables, and then estimating the models specified in \autoref{sec:iv-model-uni}.
-    There were no observable differences in the hiring rate of male vs female faculty.
-    See \autoref{sec:appendix-hiring} for further details.
-}
-It is clear that public universities' hiring is negatively impacted by falls in state funding, even using highly aggregated data for a total count of hiring across an entire decade.
-While this analysis strengthens the case negative effects on hiring plays, it is crude for measuring exact magnitudes for the effect on faculty hiring.
-I leave it to further research to delve deeper into the magnitudes that hiring plays as the primary causal mediator for the effects of falls in state funding on faculty composition at public universities.
+While market forces have been effective in the successes of US higher education research, we should worry about the effects on education.
+Enrolment in higher education has increased drastically since 1990, and public universities educate more than twice as many undergraduates in 2021 as their private counterparts.
+%At the same time returns to higher education have increased drastically \citep{berman1998implications}.
+Increasingly these students are being taught by lecturers, and not tenured faculty.
+Lecturers are often employed on short-term or part-time contracts (adjunct), with limited job stability.
+While we should worry for faculty, and their working conditions, there is credible evidence that relying on adjunct lecturers leads to worse student outcomes, relative to full-time lecturers and professors \citep{zhu2021limited}.
+As tenure has increasingly become a private sector phenomenon, we should worry about the long-term effects on higher education, and the effect on the average student who attends a public university in the US.
+
+%Additionally, put into further context of the trends in university competition \citep{urquiola2020markets}, and XYZ \citep{ehrenberg2012american}.

### 2023-07-13 text/sections/07-conclusion.tex
re-writing

@@
 
-This analysis investigates how the recent stagnation in state support for higher education has affected public universities, and their faculty composition.
+This analysis investigates how the recent stagnation in state support for higher education has affected faculty, and their composition, at US public universities.
 This work contributes to the literature along two primary dimensions.
-Firstly, by isolating changes in state funding on public universities via state funding shocks, the analysis provides a plausible mechanism for the observed negative effects on students outcomes that this stagnation has brought \citep{NBERw23736,NBERw27885}.
-Secondly, this analysis considers individual professors as the unit of analysis using a dataset (IBHED) new in the economic literature.
-These data allow for detailed analysis of thousand's of professors' salaries, and are a starting point for further analysis of the determinants of professors wages in the Illinois public university system.
+Firstly, by isolating changes in state funding on public universities via state funding shocks, this work provides an explanation for the increased relaiance on lecturers and away from tenure-track and tenured professors at US public universities.
+This approach used multiple methods to estimate the short- and medium-run effects of state funding cuts on faculty outcomes.
+Secondly, this work asks investigates individual faculty are affected by changes in state funding for their university using a dataset new in the economic literature (IBHED).
+These data allow for detailed analysis of thousands of faculty salaries, and employment outcomes in the Illinois university system.
 
-Public universities have systematically substituted away from tenure-track and tenured professors, towards non-tenured lecturers, in the face of persistent declines in state funding for higher education.
+I found that public universities have systematically substituted away from tenure-track and tenured professors, towards lecturers, in the face of persistent declines in state funding for higher education.
 The effects on incumbent professors in the state of Illinois are non-distinguishable from zero, while suggestive information shows falls in hiring, which together implies that the changes in faculty composition are driven by reduced hiring of new professors.
@@
 While costs education have been rising in the US, public universities have also dealt with declining state funding.
-It is natural to expect that such headwinds will lead to systematic change at public universities, changes that affect their faculty, and limit the universities' goals in research and education.
+I show that these headwinds led to systematic change at public universities, changes that affect their faculty, likely limiting public universities in their goals in research and education.
 These results show large changes in faculty composition, and that stagnation in state support explains at least a third of the observed shift away from tenured professors and towards contingent lecturers.

### 2023-07-13 programs/analysis/ipeds-shock.R

@@ -873,11 +873,10 @@ fixest::coefplot(list(
 # https://cran.r-project.org/web/packages/jtools/vignettes/summ.html#plot_summs()_and_plot_coefs()
 
 # Calculate the elasticities of susbtituion, with SEs bootstrapped.
-# TODO: CALCULATE WITH A BOOTSTRAP FROM THE FOLLOWING:
-# TODO https://stackoverflow.com/questions/63777368/computing-the-standard-error-when-dividing-coefficients-of-different-regressions
+boot.count <- 10000
 
 ## write a function to calculate the lecturer <-> assistant prof elasticity
-assistant.fun <- function(data, inds){
+lecturer_assistant.fun <- function(data, inds){
     substitution.reg <- fixest::feols(c(
             log(lecturer_prof_count / enrollment_reported),
             log(assistant_prof_count / enrollment_reported)
@@ -896,12 +895,26 @@ assistant.fun <- function(data, inds){
     return(substitution.est)
 }
 # bootstrap the function
-boot.calc <- boot(reg.data, statistic = assistant.fun, R = 10000)
-# calculate confidence intervals
-boot.ci(boot.calc, type = c("perc", "bca"))
+boot.calc <- boot::boot(reg.data,
+    statistic = lecturer_assistant.fun, R = boot.count, parallel = "multicore")
+# calculate the standard error
+summary(boot.calc)
+boot.mean <- mean(boot.calc$t)
+boot.se <- sd(boot.calc$t)
+boot.ci <- quantile(boot.calc$t, probs = c(0.025, 0.975))
+# Check for normality -> result roughly normal
+boot.calc$t %>%
+    as.data.frame() %>%
+    ggplot(aes(V1)) +
+    geom_histogram(aes(y = after_stat(density)),
+        fill = "lightgray", col = "black") +
+    stat_function(fun = dnorm, args = list(mean = boot.mean, sd = boot.se))
+print(c("Calculated point est for substitution between lecturers + ast profs",
+    boot.mean, "with SEs", boot.se,
+    "and 95 % CI", boot.ci))
 
 ## write a function to calculate the lecturer <-> full prof elasticity
-full.fun <- function(data, inds){
+lecturer_full.fun <- function(data, inds){
     substitution.reg <- fixest::feols(c(
             log(lecturer_prof_count / enrollment_reported),
             log(full_prof_count / enrollment_reported)
@@ -920,9 +933,61 @@ full.fun <- function(data, inds){
[…]

### 2023-07-13 text/paper.tex
Major revision in progress.
re-writing

@@
     This project's Github repository, with materials and anonymised data for replication, is available at 
-    \url{https://github.com/shoganhennessy/state-faculty-composition}.
+    \url{https://github.com/shoganhennessy/https://github.com/shoganhennessy/state-funding-faculty}.
     Any comments or suggestions may be sent to me at \href{mailto:seh325@cornell.edu}{\nolinkurl{seh325@cornell.edu}}, or raised as an issue on the Github project.
     } \\ \vspace{0.1cm}
-    \href{https://github.com/shoganhennessy/state-faculty-composition/blob/main/state-faculty-composition-2023.pdf}{Newest version available here.}
+    \href{https://github.com/shoganhennessy/https://github.com/shoganhennessy/state-funding-faculty/blob/main/https://github.com/shoganhennessy/state-funding-faculty-2023.pdf}{Newest version available here.}
 }
 \rhead{\today}
-\lhead{Stagnating State Funding and its Effect on Faculty.}
+\lhead{Stagnating State Funding for Higher Education and its Effect on Faculty.}
 % Begin
@@
 \bibliography{sections/08-bibliography.bib}
+
+% Separated Figures \& Tables
+\newpage
+\input{sections/09-figures-tables.tex}
+
 % Appendix
-\input{sections/09-appendix.tex}
+\input{sections/10-appendix.tex}
 \end{document}

### 2023-07-13 text/sections/01-introduction.tex
re-writing

@@
-Public universities educate the majority of higher education students in the US, yet have experienced a secular decline in state funding (per student) over the last three decades.
-At the same time, both public university student enrolment and employment of contingent lecturers dramatically increased at public universities.
-I show that four-year, degree-granting public universities have experienced a systematic fall in state funding per student, and this fall led to a substitution away from tenure-track/tenured professors towards contingent lecturers.
-Falls in state funding for public universities, instrumented by state-level shocks to state-wide higher education, led to a fall in the number of tenure-track/tenured professors per student within a university, and an increase in the number of lecturers.
+Public universities educate the majority of higher education students in the US, yet have experienced a stagnation in state funding over the last three decades .
+At the same time, public university employment of lecturers dramatically increased at public universities, while enrolment rose dramatically.
+Previous research has examined the effects of the decline in state funding on student outcomes \citep{NBERw23736,NBERw27885}, but there is little evidence of the effects of thid funding decline on faculty.
+I show that four-year, degree-granting public universities have experienced a systematic decline in state funding per student, and use a shift-share instrument to show that this fall led to a substitution away from tenure-track/tenured professors towards contingent lecturers.
+Analsysis of every public university faculty member in the state of Illinois shows that incumbent faculty were relatively unaffected, implying that changes in faculty composition arose by disrupting public universities' ability to hire or replace their tenure-track/tenured faculty.
 Private universities were not exposed to similar financial constraints during this same time period, and do not exhibit the substitution away from tenure-track/tenured professors, so that the stagnating state support for public universities has implications for the wider structure of US higher education and research.
+
+There are various margins through which state fundings cuts can impact a university's faculty.
+A university's level of resources affect the ability that departments can hire new professors, retain their current fauclty, or provide retirement incentives.
+So that it is natural to expect that funding cuts may affect the composition of professors at a university.
+Similarly, the wages a university pays their faculty is affected by funding for the university, including the starting salary it offers to new professors and the amount of yearly raises (if any) for incumbent faculty.
+So that it is natural to expect that systematic funding cuts will lead to changes in faculty composition, and possibly affect incumbent professors, too.
+Yet, it is a priori unclear whether/how much each of these margins adjusts to funding cuts, as these decision depend on how the university administrations view the productivity of different inputs, plus the degree of wage and employment stickiness over time.
 
@@
 
+I use data from the Integrated Postsecondary Education Data System (IPEDS) to study the impact of changes in state funding for US public univeristies 1990--2017.
+I combine these data with panel data for every public university faculty member in the state of Illinois for 2010--2021 (Illinois Board of Hdigher Education Database, IBHED), to measure how faculty are affected by changes in state funding.
+These data allow me to answer the following research question: if a US public university receives an extra \$1,000 for each of their students (or a 10\% increase), then do they adjust the number of and composition of faculty they employ, and how does this affect individual faculty already at the university?
+
 I employ a shift-share instrument to identify changes in state funding for higher education, which exploits how universities differ in their financial reliance on state funding, combined with yearly shocks to the amount an entire state funds higher education --- following \cite{NBERw23736,NBERw27885}.
 State governments decide how to fund higher education, and everything else, by a complicated process which can be influenced by local economic conditions, changing state priorities, or even lobbying from the state universities themselves.
-As such, it is necessary to employ an identification strategy, such as the shift-share instrument, to account for endogeneity in universities' state funding.
+As such, state funding is not randomly assigned to universities and so it is necessary to employ an identification strategy, such as the shift-share instrument, to account for this.
 The shift-share instrument interacts reliance on state funding in a base period with the entire state's total funding per higher education student for each year, to exploit changes in funding and how much each university historically relied on state funding for higher education.
 
-\textbf{TODO: Write section on the data sources.  See Lovenheimn (2020) on how to do it.}
-
+\textbf{Adjust this to have dollar figures, and generall re-do desc of results.}
 Falls in state funding affects faculty composition at public universities, which substitute away from tenured and tenure-track professors towards contingent lecturers.
@@
 
+\textit{Re-do the lit review to focus on mechanisms of student quality.}
 A number of studies have shown that stagnating state funding for higher education negatively impacts student outcomes.
@@
 
-While it is not immediately clear whether a systematic substitution towards lecturers away from professors reduces teaching effectiveness, it is clear that stagnating state funding has negatively impacted student outcomes \citep{NBERw23736,NBERw27885}.
-\textbf{REWRITE:} where faculty composition is a possible mechanism for how the stagnating state funding for higher education transmits to students, and has led to worsening education outcomes at public universities for the last thirty years.
+Mine is the frst paper to provide causal evidence of the impact of state funding for higher education on faculty composition and wages.
+The closest related research has examined (\textbf{theoretical papers here}) $\hdots$
 
-\textbf{TO-DO:
-    This paper precedes as follows.
-}
+My results provide suggestive evidence on the mechanisms for previous research on the negative effects of state funding cuts on student outcomes at US public universities \citep{zhu2021limited,ehrenberg2012american}.
+
+Lastly, my results providde evidence that the long term trends in higher education funding and of non-tenure track faculty are causally related.
+\textbf{Sentence on how they will continue.}
+
+This paper proceeds as follows.
+\autoref{sec:data} describes the data for university finances and faculty in Illinois, and trends in public university funding for the last three decades.
+\autoref{sec:conceptual} gives the conceptual framework for how state funding may affect faculty.
+\autoref{sec:empirics} draws the empirical framework for isolating the causal effects of state funding on faculty composition and individual faculty, and \autoref{sec:results} presents the empirical results.
+\autoref{sec:discussion} discusses the context and implications for the findings.
+\autoref{sec:conclusion} concludes.

### 2023-07-13 text/sections/04-empirical.tex
re-writing

@@
 Panel A shows estimates of a funding shock of \$1 per student in the state on state funding per student at the university, and Panel B the \% increase effects of a 1\% increase in the shock per student on state funding per student.
-\textbf{TO-DO:
-    Add description the dollar-by-dollar, of the first stage.
-}
-We note columns (1) and (2) estimate that a funding shock (per student in the entire state) of 10\% is associated with 9.8\% change in state funding per student at the university; the instrument is strong, and we note similarity in estimates with and without inclusion of fixed effects.
-Columns (3) and (4) include the tuition revenue control (explained in \autoref{foot:control}) to exhibit estimates with the inclusion of this possible collider or bad control.
-Column (3) shows similar estimates to columns (1), (2) thanks to inclusion of fixed effects, so that the fixed effects was effective in soaking variation in per-student tuition revenues at the institution-year level.
+Columns (1) shows that a funding shock of \$10 (per student in the entire state) is associated with \$11.76 less state funding per student at the univeristy, with the corresponding -10\% funding shock leads to -9.77\%
+less funding (coumn 1, panel B), with similar estimates with and without including fixed effects.
+Column (2) shows estimates of the first-stage without including fixed effects, and gives less precise estimates for the funding shock, likely thanks to systematic differences in universities unaccounted for without fixed effects.
+Columns (3) and (4) include the tuition revenue control (explained in \autoref{foot:control}) to exhibit estimates with the inclusion of this possible collider (or bad control).
+Column (3) shows similar estimates to column (1) in both Panels A and B thanks to inclusion of fixed effects, so that the uncertainty in including tuition revenue as a possible bad control for the level of state funding does not matter thanks to the inclusion of fixed effects.
 Column (1) represents the estimates for \autoref{eqn:firststage} with fixed effects, omitting the tuition revenue control, and is the preferred form that I proceed with.

### 2023-07-13 text/paper.tex
re-writing

@@
     This project's Github repository, with materials and anonymised data for replication, is available at 
-    \url{https://github.com/shoganhennessy/https://github.com/shoganhennessy/state-funding-faculty}.
+    \url{https://github.com/shoganhennessy/state-funding-faculty}.
     Any comments or suggestions may be sent to me at \href{mailto:seh325@cornell.edu}{\nolinkurl{seh325@cornell.edu}}, or raised as an issue on the Github project.
     } \\ \vspace{0.1cm}
-    \href{https://github.com/shoganhennessy/https://github.com/shoganhennessy/state-funding-faculty/blob/main/https://github.com/shoganhennessy/state-funding-faculty-2023.pdf}{Newest version available here.}
+    \href{https://github.com/shoganhennessy/state-funding-faculty/blob/main/https://github.com/shoganhennessy/state-funding-faculty-2023.pdf}{Newest version available here.}
 }

### 2023-07-16 text/sections/00-abstract.tex
writing revision 2 mostly finished.

@@
-US universities have systematically substituted away from tenure-track/tenured professors towards contingent lecturers for the last thirty years, while at the same time states have limited how much they fund higher education.
+US universities have systematically substituted away from tenure-track/tenured professors towards contingent lecturers for the last thirty years, while states have limited how much they fund higher education.
 This paper connects the changes the faculty composition at US public universities to the fall in the states' funding for their public universities, using a shift-share approach to instrument for state funding and identify the causal effects of stagnating funding for higher education.
-A fall of $10$\% in funding for a public university, via a state funding shock, decreases the number of assistant professors per student by $1.4$\% and full professors by $1.2$\%, yet increases the number of lecturers per student by $4.4$\%; long-run estimates show that the faculty composition is affected both in the same year as the funding shock, and two to three years following.
-Analysis of all the professors at Illinois public universities shows that incumbent professors are not affected by substantial funding shocks over 2011-2021, with suggestive evidence showing that the faculty changes arose by limiting the hiring of new professors to replace leaving tenure-track/tenured professors.
+A fall in funding of \$1,000 per student leads to univeristies employing 6 more lecturers on average; relative to enrolment, a fall of $10$\% in funding per student decreases the number of assistant professors per student by $1.4$\% and full professors by $1.2$\%, yet increases the number of lecturers per student by $4.4$\%.
+Dynamic estimates show that state funding cuts affect composition of faculty for the next 3 years, showing that funding cuts have lingering effects.
+Analysis of all the professors at Illinois public universities shows that incumbent professors' salaries, promotion rate, and quit rate are all are not affected by funding cuts over 2011-2021, implying that the composition change arose by limiting the hiring of new professors to replace leaving tenure-track/tenured professors.
 These results show the institutional effects of stagnating state funding for higher education, and raise questions about the direction that public education heads as these financial headwinds show no sign of dissipating.

### 2023-07-16 text/sections/01-introduction.tex
writing revision 2 mostly finished.

@@
 At the same time, public university employment of lecturers dramatically increased at public universities, while enrolment rose dramatically.
-Previous research has examined the effects of the decline in state funding on student outcomes \citep{NBERw23736,NBERw27885}, but there is little evidence of the effects of thid funding decline on faculty.
+Previous research has examined the effects of the decline in state funding on student outcomes \citep{NBERw23736,NBERw27885}, but there is little evidence of the effects of this funding decline on faculty.
 I show that four-year, degree-granting public universities have experienced a systematic decline in state funding per student, and use a shift-share instrument to show that this fall led to a substitution away from tenure-track/tenured professors towards contingent lecturers.
-Analsysis of every public university faculty member in the state of Illinois shows that incumbent faculty were relatively unaffected, implying that changes in faculty composition arose by disrupting public universities' ability to hire or replace their tenure-track/tenured faculty.
-Private universities were not exposed to similar financial constraints during this same time period, and do not exhibit the substitution away from tenure-track/tenured professors, so that the stagnating state support for public universities has implications for the wider structure of US higher education and research.
+Analysis of all public university faculty in the state of Illinois shows that incumbent faculty were relatively unaffected, in terms of wages, promotion and quit rates, implying that changes in faculty composition arose by disrupting public universities' ability to hire or replace their tenure-track/tenured faculty.
+Private universities were not exposed to similar financial constraints during this same time period, and do not exhibit the substitution away from tenure-track/tenured professors, so that the stagnating state funding for public universities has implications for the wider structure of US higher education and research.
 
-There are various margins through which state fundings cuts can impact a university's faculty.
-A university's level of resources affect the ability that departments can hire new professors, retain their current fauclty, or provide retirement incentives.
+There are various margins through which funding cuts can impact a university's faculty.
+A university's level of resources affect the ability that departments can hire new professors, retain their current faculty, or provide retirement incentives.
 So that it is natural to expect that funding cuts may affect the composition of professors at a university.
@@
 I employ a shift-share instrument to identify changes in state funding for higher education, which exploits how universities differ in their financial reliance on state funding, combined with yearly shocks to the amount an entire state funds higher education --- following \cite{NBERw23736,NBERw27885}.
-State governments decide how to fund higher education, and everything else, by a complicated process which can be influenced by local economic conditions, changing state priorities, or even lobbying from the state universities themselves.
+State governments decide how to fund higher education by a complicated process which can be influenced by local economic conditions, changing state priorities, or even lobbying from the state universities themselves.
 As such, state funding is not randomly assigned to universities and so it is necessary to employ an identification strategy, such as the shift-share instrument, to account for this.
-The shift-share instrument interacts reliance on state funding in a base period with the entire state's total funding per higher education student for each year, to exploit changes in funding and how much each university historically relied on state funding for higher education.
+The shift-share instrument interacts how much a university relies on state funding as a percent of revenues in a base period with the entire state's total funding per higher education student for each year, exploiting changes in funding and how much each university historically relied on state funding for higher education.
 
-\textbf{Adjust this to have dollar figures, and generall re-do desc of results.}
-Falls in state funding affects faculty composition at public universities, which substitute away from tenured and tenure-track professors towards contingent lecturers.
-A fall of $-10$\% in state funding leads to a fall of $-1.4$\% in the number of assistant professors per student at a university, a fall of $-1.2$\% for tenured professors, and an increase in 4.4\% the number of lecturers per student.
-Local projection estimates show that these effects linger for three to four years after the initial funding shock, showing that the effect is not isolated to the immediate year of the first funding shock.
-Over the same time period, state funding fell by around $-$35\%, while the count of professors per student fell by $-$9\%, so that these results show that falls in state funding explain around a third of the substitution away from tenure-track and tenure professors towards contingent lecturers.
+Falls in state funding affect faculty composition at public universities, which substitute away from tenured and tenure-track professors towards contingent lecturers.
+A funding cut of \$1,000 per student leads to an average university employing 6 more lecturers;
+in percentage terms, a (funding per student) cut of 10\% leads to a fall of 1.4\% in the number of assistant professors per student at a university, a fall of 1.2\% for tenured professors, and an increase in 4.4\% the number of lecturers per student.
+Local projection estimates show that these effects linger for three to four years after the initial funding shock, showing that the effect is not isolated to the year of the funding shock. 
+Over the same time period, state funding per student fell by around 35\%, count of (tenure-tack/tenured) professors per student fell by 9\%, and lecturers per student increased by 99\%, so that these results show that falls in state funding explain around 53\% of the fall in professors per student
+% i.e. -(35 * 0.137) / -9
+and 15.5\% of the rise in lecturers per student.
+% ie., (35 * 0.437) / 99
 Additionally, I use individual level data on all Illinois public university professors in 2010-2021 to investigate whether the shocks to state funding affected individual professors.
@@
 
-\textit{Re-do the lit review to focus on mechanisms of student quality.}
-A number of studies have shown that stagnating state funding for higher education negatively impacts student outcomes.
-\cite{NBERw23736} show that increases in public university spending (via funding shocks) increases enrolment and degree completion among students,\footnote{
-    \cite{miller2022making} further analyse the effects of falls in university revenues, by finding that reductions in public universities tuition prices (suggestively motivated by falling state funding) leads to reductions in provided financial aid.
-}
-that high school students exposed to similar public school funding shocks are less likely to attend college \citep{jackson2021school}, and that these shocks induce public universities to rely more on tuition as a source of revenue \citep{bound2019public}.
-\textbf{Faculty outcomes surveyed by \cite{ehrenberg2003studying}.}
-\cite{NBERw27885} use the same methodology to show that increases in state funding lead to degree lower completion time and later-life debt for students.\footnote{
-    \cite{bound2007cohort} use similar variation in state funding per student to show that lower funding leads to lower completion rates at public universities.
-}
+Mine is the first paper to provide causal evidence on the impact of state funding on faculty composition, and wages.
+The closest related research on higher education funding has examined how university spending affected graduation rates and levels of student debt \citep{NBERw23736,NBERw27885}, and university finances \citep{miller2022making,bound2019public,brown2014endowment}.
+For faculty outcomes, multiple theoretically model university decision-making and faculty hiring (see e.g., \citealt{abe2015implications,johnson2009jep,NBERc13879}), while others measured how faculty at universities were affected by endowment shocks and the 2008 recession \citep{brown2014endowment,turner2014impact}.
+These papers measure effects of funding and changes in university revenues, but not measure effects on faculty, or other outcomes related to university instruction.
 
-Over the last half century, there have been drastic changes to the heirarchy of US higher education, where the top universities have become more selective for student applicants, and the average university less selective \citep{hoxby2009changing}.
-Similarly, research faculty at US universities are increasingly concentrated among the graduates of highly selective PhD programmes \citep{wapman2022quantifying}.
-Faculty composition is an interesting outcome on its own, but its relationship with higher education finances has so far not been studied.
-The closest previous examples study  how university endowments react to negative endowment shocks \citep{brown2014endowment} and the 2008 recession \citep{turner2014impact}, and multiple theoretically model university decision-making and faculty hiring (see e.g., \citealt{abe2015implications,johnson2009jep,NBERc13879}).
-%Lastly, instruction costs within individual universities vary substantially across subject areas, primarily thanks to department differences in class size and faculty salaries \citep{hemelt2021math}.
-The number of professors per student may be particularly important for quality of instruction and thus educational outcomes;
-small class sizes in secondary schools lead to increases in test scores \citep{angrist1999using}, and UK university students perform worse academically in particularly large classes \citep{bandiera2010heterogeneous}.
-Though, the economics literature is not decided on whether contingent lecturers lead to better or worse outcomes in higher education.
-\textbf{Beef up this part to focus more on negative effects on students.}
-Lecturers are more effective teachers than research professors at a highly selective university \citep{bettinger2010does,figlio2015tenure}, but students taught by part-time lecturers have worse student outcomes relative to full-time lecturers or faculty \citep{zhu2021limited,ehrenberg2005tenured,jaeger2011examining}.
-
-\textbf{Tenure research \citep{mcpherson1999tenure,mcpherson1983economics}}
-
-Mine is the frst paper to provide causal evidence of the impact of state funding for higher education on faculty composition and wages.
-The closest related research has examined (\textbf{theoretical papers here}) $\hdots$
-
-My results provide suggestive evidence on the mechanisms for previous research on the negative effects of state funding cuts on student outcomes at US public universities \citep{zhu2021limited,ehrenberg2012american}.
-
-Lastly, my results providde evidence that the long term trends in higher education funding and of non-tenure track faculty are causally related.
-\textbf{Sentence on how they will continue.}
+My results provide suggestive evidence on mechanisms for the previous research connecting state funding cuts for higher education and worsening student outcomes.
+Substitution towards lecturers is likely a cost-cutting measure for funding constrained universities, though relying on these faculty --- who are often over-worked and not granted long-term employment protections --- may lead to worse student outcomes \citep{ehrenberg2005tenured,zhu2021limited,jaeger2011examining}.
+Lastly, my results provide evidence that the long term trends in higher education funding are causally related, contributing to the literature in the long run trends in faculty outcomes \citep{ehrenberg2003studying} and trends in US higher education and funding \citep{hoxby2009changing}.
 

### 2023-07-16 text/sections/04-empirical.tex
writing revision 2 mostly finished.

@@
 does it make sense to consider state funding in purely dollar amounts per student, or in percent change rate?
-The funding shock instrument is a strong predictor for the average university's level of state funding in either unit; a funding shock of \$1,000 per student in the entire state leads to around \$1,176 per student at the university, while a funding shock of -10\% leads to around -9.77\% less state funding per year.
+The funding shock instrument is a strong predictor for the average university's level of state funding in either unit; a funding shock of \$1,000 per student in the entire state leads to around \$1,176 per student at the university, while a funding shock of $-10$\% leads to around 9.77\% less state funding per year.
 Yet, the level of state funding (and the outcome variables) vary greatly between states for the unit of analysis.

### 2023-07-18 text/sections/09-figures-tables.tex
Revision 2 finished (though needs check over of writing).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section to host Figures
\section{Figures}
\label{sec:figures}

\begin{figure}[H]
    \centering
    \singlespacing
    \caption{Mean Total Funding among Public Universities, by Year.}
    \begin{subfigure}[b]{0.495\textwidth}
        \centering
        \caption{Total, \$ 2021 CPI-U.}
        \includegraphics[width=\textwidth]{figures/mean-funding-total.png}
        \label{fig:mean-funding-total}
    \end{subfigure}
    \begin{subfigure}[b]{0.495\textwidth}
        \centering
        \caption{Per Student, \$ 2021 CPI-U.}
        \includegraphics[width=\textwidth]{figures/mean-funding-fte.png}
        \label{fig:mean-funding-fte}
    \end{subfigure}
    \label{fig:funding}
    \justify
    \footnotesize
    \textbf{Note}:
    This figure shows the mean funding for US public universities as a total (in figure a.) and divided by student enrolment (figure b.).
    The numbers are adjusted to 2021 figures by CPI-U.
    Non-institutional revenues refers to the sum of federal, state, and local funding plus tuition revenues; these sum to the majority of university funding, but exclude numbers such as university income from capital projects.
    These figures are calculated in IPEDS data.
\end{figure}

\begin{figure}[H]
    \centering
    \singlespacing
    \caption{Total Student enrolment, by University Sector and Year.}
    \begin{subfigure}[b]{0.495\textwidth}
        \centering
        \caption{Total Enrolment, Nation-wide.}
        \includegraphics[width=\textwidth]{figures/enrollment-total.png}
        \label{fig:enrollment-total}
    \end{subfigure}
    \begin{subfigure}[b]{0.495\textwidth}
        \centering
        \caption{Mean Enrolment, per University.}
        \includegraphics[width=\textwidth]{figures/enrollment-mean.png}
        \label{fig:enrollment-mean}
    \end{subfigure}
    \label{fig:enrolment}
    \justify
    \footnotesize
    \textbf{Note}:
    This figure shows the total and mean enrolment for US universities, comparing public and private universities.
    Most of the higher education enrolment increase for the last 30 years was in public universities, who continue to enrol the vast majority of higher education students in the US.
    These figures are calculated in IPEDS data.
\end{figure}

\newpage
\begin{figure}[H]
    \centering
    \singlespacing
    \caption{Trends in Mean Student Enrolment per Professor, by University Sector and Faculty level.}
    \begin{subfigure}[b]{0.495\textwidth}
        \centering
        \caption{Lecturer.}
        \includegraphics[width=\textwidth]{figures/lecturer-fte-perprof.png}
        \label{fig:lecturer-fte-perprof}
    \end{subfigure}
    \begin{subfigure}[b]{0.495\textwidth}
        \centering
        \caption{Assistant.}
        \includegraphics[width=\textwidth]{figures/assistant-fte-perprof.png}
        \label{fig:assistant-fte-perprof}
    \end{subfigure}
    \begin{subfigure}[b]{0.495\textwidth}
        \centering
        \caption{Full.}
        \includegraphics[width=\textwidth]{figures/full-fte-perprof.png}
        \label{fig:full-fte-perprof}
    \end{subfigure}
    \begin{subfigure}[b]{0.495\textwidth}
        \centering
        \caption{All.}
        \includegraphics[width=\textwidth]{figures/all-fte-perprof.png}
        \label{fig:all-fte-perprof}
    \end{subfigure}
    \label{fig:fte-perprof}

    \justify
    \footnotesize
    \textbf{Note}:
    This figure shows the average number of students per faculty member, by different faculty position, at US public universities.
    E.g., panel A calculates the mean of (lecturer count) / (student enrolment) at US public universities, for each year of 1990--2017, to show the average trend in faculty composition.
    These figures are calculated in IPEDS data.
\end{figure}

\newpage
\begin{figure}[H]
    \centering
    \singlespacing
    \caption{Mean Funding Sources among Illinois Public Universities, by Year.}
    \begin{subfigure}[b]{0.495\textwidth}
        \centering
        \caption{Total, \$ 2021 CPI-U.}
        \includegraphics[width=\textwidth]{figures/illinois-funding-total.png}
        \label{fig:illinois-funding-total}
    \end{subfigure}
    \begin{subfigure}[b]{0.495\textwidth}
        \centering
        \caption{Per Student, \$ 2021 CPI-U.}
        \includegraphics[width=\textwidth]{figures/illinois-funding-fte.png}
        \label{fig:illinois-funding-fte}
    \end{subfigure}
    \label{fig:illinois-funding}

    \justify
    \footnotesize
    \textbf{Note}:       
    This figure shows the mean funding for Illinois public universities as a total (in figure a.) and divided by student enrolment (figure b.).
    The numbers are adjusted to 2021 figures by CPI-U.
    Non-institutional revenues refers to the sum of federal, state, and local funding plus tuition revenues; these sum to the majority of university funding, but exclude numbers such as university income from capital projects.
    These figures are calculated in IPEDS data.
\end{figure}

\newpage
\begin{figure}[H]
    \centering
    \singlespacing
    \caption{Local Projection Estimates for Professor Count per Student, by Professor Group.}
    \begin{subfigure}[b]{0.495\textwidth}
        \centering
        \caption{Lecturers.}
        \includegraphics[width=\textwidth]{figures/lecturer-count-lp.png}
        \label{fig:lecturer-count-lp}
    \end{subfigure}
    \begin{subfigure}[b]{0.495\textwidth}
        \centering
        \caption{Assistant Professors.}
        \includegraphics[width=\textwidth]{figures/assistant-count-lp.png}
        \label{fig:assistant-count-lp}
    \end{subfigure}
    \begin{subfigure}[b]{0.495\textwidth}
        \centering
        \caption{Full Professors.}
        \includegraphics[width=\textwidth]{figures/full-count-lp.png}
        \label{fig:full-count-lp}
    \end{subfigure}
    \begin{subfigure}[b]{0.495\textwidth}
        \centering
        \caption{All Professors.}
        \includegraphics[width=\textwidth]{figures/all-count-lp.png}
        \label{fig:all-count-lp}
    \end{subfigure}
    \label{fig:count-lp}
    \justify
    \footnotesize
    \textbf{Note}:
    These figures show the local projections estimates of regression specification \eqref{eqn:secondstage}, with the funding shock as an instrument for state funding, using IPEDS data.
    The coefficient estimate is effect of state funding ($X_{i,t}$) on faculty count per student ($Y_{i,t}$), while accounting for auto-correlation between different time periods --- i.e., between $X_{i,t}, X_{i,t-1}$ and $Y_{i,t}, Y_{i,t-1}$.
    These results use a $\log-\log$ specification, so the estimates are for the elasticity of professor count per student in a year $t+k$ with respect to state funding in year $t$, where years $k = 0, \hdots, 10$ are on the x-axis. 
    Standard errors are clustered at the state-year level.
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section to host Tables
\newpage
\section{Tables}
\label{sec:tables}

\begin{table}[H]
    \singlespacing
    %\centering
    \caption{IPEDS Summary Statistics, Public Universities Panel 1990--2021}
    \makebox[\textwidth][c]{\input{tables/ipeds-summary-fte.tex}}
    \label{tab:ipeds-summary}
    \justify
    \footnotesize
    \textbf{Note}:
    This table shows the summary statistics for every public university--year observation in IPEDS data.
    The numbers are adjusted to 2021 figures by CPI-U.
    Non-institutional revenues refers to the sum of federal, state, and local funding plus tuition revenues; these sum to the majority of university funding, but exclude numbers such as university income from capital projects.
\end{table}

\begin{table}[H]
    \singlespacing
    \centering
    \caption{IBHED Summary Statistics, Professor Panel 2010--2021.}
    \makebox[\textwidth][c]{\input{tables/illinois-summary.tex}}
    \label{tab:illinois-summary}

    \justify
    \footnotesize
    \textbf{Note}:
    This table shows the summary statistics for every faculty--year observation in the IBHED data, which represents every faculty member in the Illinois public university system over years 2010--2021.
    The numbers are adjusted to 2021 figures by CPI-U.
    Lecturer is a binary for whether the faculty member is designated as a lecturer in the databse, and similarly for the assistant professors, full professors, and administrative faculty;
    salary refers the sum of base salary and benefits.
    All salary and benefits refers to summary statistics on the salary and benefits, respectively, of all faculty (regardless of position).
\end{table}

\newpage
\begin{table}[h!]
    \singlespacing
    \centering
    \caption{First Stage Estimates, for State Funding by funding shock.}
    \textbf{Panel A: units in \$ per student}

    \makebox[\textwidth][c]{\input{tables/firststage-reg-rawcount.tex}}

    \textbf{Panel B: units in log \$ per student}

    \makebox[\textwidth][c]{\input{tables/firststage-reg-fte.tex}}

    \label{tab:firststage-reg}
    \justify
    \footnotesize
    \textbf{Note}:
    These tables show the first stage OLS estimates of regression specification \eqref{eqn:firststage}, showing the effect of the funding shock on state funding to gauge performance as an instrument.
    Each observation is a public univeristy-year, in the IPEDS data.
    Panel A shows the effect of an funding shock of \$-1 per student in the state on the number of \$'s of state funding per student at the university --- i.e.,
    \$-1 funding shock per student in the state leads to \$1.176 less state funding per student at the university according to preferred specification column 1.
    Panel B shows the effect of a $-10$\% change funding shock per student in the state on $10$\% change in state funding per student at the university --- i.e.,
    $-10$\% funding shock per student in the state leads to $-9.77$\% less state funding per student at the university according to prefferred specification column 1.        
    Standard errors are clustered at the state-year level, and univeristy $+$ year fixed effects are included where noted.
\end{table}

\newpage
\begin{table}[H]
    \singlespacing
    \centering
    \caption{Shift-Share Instrument Balance Test, in IPEDS 1990--2017.}

    \textbf{Panel A: units in \$}

    \makebox[\textwidth][c]{\input{tables/firststage-balance-count.tex}}

    \textbf{Panel B: units in log \$ per student}

    \makebox[\textwidth][c]{\input{tables/firststage-balance-log.tex}}

    \justify
    \footnotesize
    \textbf{Note}:
    These tables show the OLS estimates of regression specification \eqref{eqn:firststage}, showing the effect of the funding shock on other outcomes to gauge balance of the instrument with respect to university characteristics
    ENrolment is measured in thousands of students.
    Each observation is a public univeristy-year, in the IPEDS data.
    Panel A shows the effect of an funding shock of \$-1,000 per student in the state on the counts of the outcome --- i.e.,
    \$-1,000 funding shock per student in the state is associated with 517 more students (as enrolment is measured in thousands) in column 1.
    Panel B shows the effect of a $-10$\% change funding shock per student in the state on $10$\% change in the outcomes --- i.e.,
    $-10$\% funding shock per student in the state is associated with an increase in 1.07\% more students at the unviersity according to prefferred column 1.        
    Standard errors are clustered at the state-year level, and univeristy $+$ year fixed effects are included where noted.
    \label{tab:facultycount-shock-reg}
\end{table}

\newpage
\begin{table}[H]
    \singlespacing
    \centering
    \caption{Effects of Changes in State Funding on University Faculty Composition, IPEDS 1990--2017, OLS and 2SLS Estimates.}

    \textbf{Panel A: units in \$ per student}

    \makebox[\textwidth][c]{\input{tables/facultycount-shock-reg-rawcount.tex}}

    \textbf{Panel B: units in log \$ per student}

    \makebox[\textwidth][c]{\input{tables/facultycount-shock-reg-fte.tex}}

    \label{tab:facultycount-shock-reg}
    \justify
    \footnotesize
    \textbf{Note}:
    These tables show the second stage OLS and 2SLS estimates of regression specification \eqref{eqn:secondstage}, showing the effect of state funding changes on faculty outcomes, using the funding shock to instrument for state funding in the columns labelled 2SLS.
    Each observation is a public university-year, in the IPEDS data.
    Panel A shows the effect of a fall in state funding \$-1,000 per student in the state on the number of professors --- i.e.,
    an extra \$1,000 per student leads to 6 fewer lecturers according to column 2.
    Panel B shows the effect of a $10$\% change in state funding per student at the university on the 10\% change in the number of professors per students --- i.e.,
    an extra 10\% of state funding per student leads to 4.37\% fewer lecturers per student according to column 2.
    Outcome-mean is the mean of the outcome, for Panel A the number of professors per student, for Panel B the number of faculty per student.
    Panel B uses $\log$ faculty count per student as the outcome, though the outcome mean is count of faculty per student (not in $\log$ terms).
    Standard errors are clustered at the state-year level, and university $+$ year fixed effects are included through--out.
\end{table}

\newpage
\begin{table}[H]
    \singlespacing
    \centering
    \caption{Effects of Changes in State Funding on University Faculty Composition, in Illinois 2010--2021, OLS and 2SLS Estimates.}

    \textbf{Panel A: units in \$ per student}

    \makebox[\textwidth][c]{\input{tables/facultycount-illinois-reg-rawcount.tex}}

    \textbf{Panel B: units in log \$ per student}

    \makebox[\textwidth][c]{\input{tables/facultycount-illinois-reg-fte.tex}}

    \label{tab:facultycount-illinois-reg}
    \justify
    \footnotesize
    \textbf{Note}:
    These tables show the second stage OLS and 2SLS estimates of regression specification \eqref{eqn:secondstage}, showing the effect of state funding changes on number of faculty per student in Illinois universities, using the funding shock to instrument for state funding in the columns labelled 2SLS.
    Each observation is a public university-year in the state of Illinois, where funding data come from IPEDS and faculty count come from IBHED data.
    Panel A shows the effect of a fall in state funding \$-1,000 per student in the state on the number of professors.
    Panel B shows the effect of a $10$\% change in state funding per student at the university on the 10\% change in the number of professors per students.
    Outcome-mean is the mean of the outcome, for Panel A the number of professors per student, for Panel B the number of faculty per student.
    Panel B uses $\log$ faculty count per student as the outcome, though the outcome mean is count of faculty per student (not in $\log$ terms).
    Standard errors are clustered at the university-year level, and university $+$ year fixed effects are included through--out.
\end{table}

\newpage
\begin{table}[H]
    \singlespacing
    \centering
    \caption{
        Effects of Changes in State Funding on Faculty Salaries and Exit Rate, in Illinois 2010--2021, 2SLS Estimates.}
    \makebox[\textwidth][c]{\input{tables/facultysalaries-shock-illinois-rolling.tex}}

    \makebox[\textwidth][c]{\input{tables/facultyleaving-shock-illinois-rolling.tex}}

    \makebox[\textwidth][c]{\input{tables/promotion-shock-illinois-rolling.tex}}

    \justify
    \footnotesize
    \textbf{Note}: 
    These tables show the second stage 2SLS estimates of regression specification \eqref{eqn:secondstage}, showing the effect of state funding changes on faculty outcomes at Illinois universities, using the funding shock to instrument for state funding.
    The shift-share instrument is based in the year the professor was hired, following definition in \autoref{sec:iv-model-indiv}.
    Each observation is a faculty member--year at an Illinois public university, where funding data come from IPEDS and faculty outcomes from IBHED data.
    The panels shows the effect of a $1$\% change in state funding per student at the university on faculties' salaries (base salary $+$ benefits), promotion rate (e.g., assistant professor to associate professor) , and rate of leaving their university (i.e., by quitting, retiring, or moving to another university).
    Standard errors are clustered at the university-year of hire level, and university $+$ year of hire fixed effects are included through--out.
    \label{tab:faculty-shock-illinois-rolling}
\end{table}

### 2023-07-18 text/sections/10-appendix.tex
Revision 2 finished (though needs check over of writing).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Appendix section
% Set-up the section.
\newpage
\appendix
\setcounter{table}{0}
\renewcommand{\thetable}{A\arabic{table}}
\setcounter{figure}{0}
\renewcommand{\thefigure}{A\arabic{figure}}

% Start appendix
\section{Appendix}
\label{appendix}
This project used data which are fully public, and computational tools which are fully open-source.
As such, all code and data involved in this project are available at this project's Github repository, available at \url{https://github.com/shoganhennessy/state-faculty-composition}.
They may be used for replication, or as the basis for further work, as needed.
Any comments or suggestions may be sent to me at \href{mailto:seh325@cornell.edu}{\nolinkurl{seh325@cornell.edu}}, or raised as an issue on the Github project.

A number of statistical packages, for the $R$ language \citep{R2022}, made the empirical analysis for this paper possible.
\begin{itemize}
    \item \textit{Tidyverse} \citep{tidyverse} collected tools for data analysis in the R language.
    \item \textit{LFE} \citep{lfe} implemented linear fixed effect models, with instruments, crucial for the empirical estimation in \autoref{sec:empirics}.
    \item \textit{Stargazer} \citep{stargazer} provided methods to efficiently convert empirical results into presentable output in \LaTeX.
    \item \textit{Lpirfs} \citep{lpirfs2019} implemented estimation of the \cite{jorda2005} local projections methods, with instrumental variables, crucial to the local projections estimates presented in this project.
\end{itemize}

\subsection{IPEDS First Stage}

\begin{table}[H]
    \singlespacing
    %\centering
    \caption{Mean Characteristics for Public Universities, by State Funding Shock Instrument.}
    \makebox[\textwidth][c]{\input{tables/summary-quantiles.tex}}
    \label{tab:summary-quantiles}
    \justify
    \footnotesize
    \textbf{Note}:
    This table shows the summary statistics for every public university--year observation in IPEDS data, for each of the 5 quantiles of the funding shock instrument.
    The column labelled ``1st'' refers to the mean for all university-year observations in the first quintile (bottom 20\%) of the funding shock distribution, and so on.
    The numbers are adjusted to 2021 figures by CPI-U.
    Non-institutional revenues refers to the sum of federal, state, and local funding plus tuition revenues; these sum to the majority of university funding, but exclude numbers such as university income from capital projects.
\end{table}

\begin{figure}[H]
    \centering
    \singlespacing
    \caption{Correlation Between State Funding Shock and Public University State Funding in Surrounding Years.}
    \includegraphics[width=\textwidth]{figures/lag-firststage.png}
    \label{fig:lag-firststage}
    \justify
    \footnotesize
    \textbf{Note}:
    This figure shows the correlation between state funding in year $t+k$ with the funding shock in year $t$ for a university, where $k = -7, \hdots, 5$ are the years on the x-axis.
    This shows that state funding and the funding shock have dynamic correaltion, so that dynamic effects must be estimated by localc projects -- and not simple OLS or 2SLS.
    The estimates are of \eqref{eqn:firststage}, calculated with IPEDS data, separatly for each year relative to initial shock, using the $\log-\log$ specification, including fixed effects for university $+$ year, and clustering standard errors by university $+$ year.
\end{figure}

\begin{figure}[H]
    \centering
    \singlespacing
    \caption{Local Projection Estimates for Funding Shock on State Funding, in IPEDS Data.}
    \includegraphics[width=0.6\textwidth]{figures/firststage-lp.png}
    \label{fig:firststage-lp}
    \justify
    \footnotesize
    \textbf{Note}:
    These figures show the local projections estimates of regression specification \eqref{eqn:firststage}, with the funding shock as an instrument for state funding, using IPEDS data.
    The coefficient estimate is effect of funding shock ($Z_{i,t}$) on state funding ($X_{i,t}$), while accounting for auto-correlation between different time periods --- i.e., between $Z_{i,t}, Z_{i,t-1}$ and $X_{i,t}, X_{i,t-1}$.
    These results use a $\log-\log$ specification, so the estimates are for the elasticity of state funding per student in a year $t+k$ with respect to funding shock in year $t$, where years $k = 0, \hdots, 10$ are on the x-axis. 
    Standard errors are clustered at the state-year level.
\end{figure}

\newpage
\subsection{Illinois IBHED First Stage}
\label{sec:iv-model-indiv}

This paper uses data on individual professors in the Illinois university system, to investigate the effects of changes in university revenues on the individual professors at the universities.
The outcomes here now refer to individual professors (e.g., their salary and promotion rate), so requires adjustment to the empirical approach, leveraging variation in university funding for the years after a professor joins the university.

\autoref{eqn:rolling-instrument} defines a rolling-share variant of the instrument, $\tilde Z_{j,t}$, where the university's state funding share exposure is based in the year a professor joins the university --- and not the base period 1990--1993.
$j$ indexes each professor in year $t$, $\tau(j)$ for the year the professor first joins their institution.
Identifying $\tau(j)$ is possible for $j$ by restricting to all professors hired 2011-2021 --- i.e., in the years after the start of the full panel.
It is not possible to discern the hiring year for professors who  were hired in the years preceding 2011, and so the entire sample is only possible to analyse using the base-share in years 1990-1993 formulation.

\begin{align}
    \label{eqn:rolling-instrument}
    \tilde Z_{j,t} &\coloneqq - \left[
    \left( \frac{\text{Total State Funding}_{s(j),t}}{\text{Student Population}_{s(j),t}} \right)
    \left( \frac{\text{State Funding}_{\tau(j)}}{\text{Total Revenues}_{i,\tau(j)}} \right) \right]
\end{align}

This approach leverages an insight, made available by level of the data: that an individual professor is affected by changes in university revenues after they have joined the university.
\autoref{sec:iv-model-uni} considers the number of professors employed by the university; whether a professor becomes employed at the university is likely affected by that university's finances.
The formulation here takes as given that the professor is already employed at the university, and then projects the effect of changes in state funding on these \textit{incumbent} professors following the state funding shock.

Exogeneity and relevance of the rolling-share instrument, $\tilde Z_{j,t}$, follows the same reasoning as that for the base-share instrument, $Z_{i,t}$, discussed in \autoref{sec:approp-shocks}.
The base-share instrument is appropriate for some outcomes with the individual Illinois professors, where appropriate.
We satisfy the assumptions for exogeneity by noting that none of the Illinois public campuses take the majority of state funding, and that the identification strategy relies on exogeneity in changes in state funding to individual professor-outcomes, following the year they joined the university.
Additionally, within-institution changes resulting from share reliance on state funding may be correlated with unobserved changes in the outcomes, so that \cite{NBERw27885} note the importance of controlling for the base share and state student population.
The formulation here implicitly controls for these factors via the fixed effects; results are relatively similar while including these controls with and without including fixed effects, and so are omitted.

The instrumental variables model is then defined as follows, where $i(j)$ refers to the institution that professor $j$ is employed at, and $Y_{j,t}$ for salary, rate of promotion, and propensity to leave the Illinois public university system.
The system includes fixed effects for the institution and first year of employment.
The instrument varies by institution, based in the year of first employment, so that these are the corresponding fixed effects and level of clustered standard errors.
\begin{eqnarray}
    \label{eqn:secondstage1_indiv}
    X_{i(j),t} &=& \theta_{i(j)} + \phi_{\tau(j)} + \delta \tilde Z_{i(j),t} + \epsilon_{i(j),t} \\
    \label{eqn:secondstage2_indiv}
    Y_{j,t} &=& \mu_{i(j)} + \nu_{\tau(j)} + \beta \widehat X_{i(j),t} + \varepsilon_{j,t}
\end{eqnarray}
We then interpret parameter $\beta$ as the effect of changes in state funding at an Illinois public university, via state funding shocks, on an individual professor's outcome $Y_{j,t}$.

\newpage
\subsection{Additional Results, Faculty Hiring}
\label{sec:appendix-hiring}

These results were produced by integrating the total count of faculty hires for 2010--2021 for the top-ranked 180 US universities with a sum of the funding variables, and then estimating the models specified in \autoref{sec:iv-model-uni}.
There were no observable differences in the hiring rate of male vs female faculty.

\begin{figure}[h!]
    \centering
    \singlespacing
    \caption{State Funding and Faculty Hired at Public Universities, Total for 2011--2021.}
    \includegraphics[width=0.8\textwidth]{figures/hiring-correlation.png}
    \label{fig:hiring-correlation}
    \justify
    \footnotesize
    \textbf{Note}:
    This figure shows the correlation between state funding and the number of total number of faculty hired at public universities between 2011--2021, and the line is a line of best-fit.
    Data for the number of faculty hired at public universities are provided by \cite{wapman2022quantifying}, and funding data from IPEDS.
\end{figure}

\newpage
\begin{table}[h!]
    \singlespacing
    \centering
    \caption{OLS and 2SLS Estimates for University Faculty Hires, in Illinois 2011--2021.}

    \textbf{Panel A: units in \$ per student}

    \makebox[\textwidth][c]{\input{tables/facultyhires-illinois-reg-rawcount.tex}}

    \textbf{Panel B: units in log \$ per student}

    \makebox[\textwidth][c]{\input{tables/facultyhires-illinois-reg-fte.tex}}
    \label{tab:facultyhires-illinois-reg}
    \justify
    \footnotesize
    \textbf{Note}:
    These tables show the second stage OLS and 2SLS estimates of regression specification \eqref{eqn:secondstage}, showing the effect of state funding changes on number of faculty hires at Illinois universities, using the funding shock to instrument for state funding in the columns labelled 2SLS.
    Each observation is a public university-year in the state of Illinois, where funding data come from IPEDS and faculty count come from IBHED data.
    Panel A shows the effect of a fall in state funding \$-1,000 per student in the state on the number of new faculty hires by position.
    Panel B shows the effect of a $10$\% change in state funding per student at the university on the 10\% change in the number of faculty hires per students.
    Outcome-mean is the mean of the outcome, for Panel A the number of faculty hires, for Panel B the number of faculty hires per student.
    Panel B uses $\log$ new faculty hires per student as the outcome, though the outcome mean is count of new faculty hires  per student (not in $\log$ terms).
    Standard errors are clustered at the university-year level, and university $+$ year fixed effects are included through--out.
\end{table}

Yearly variation is not observed here, so that only the aggregate level, for 180 universities, can be considered.
\begin{table}[H]
    \singlespacing
    \centering
    \caption{OLS and 2SLS Estimates for University Faculty Hiring, Total for 2011--2020.}
    \makebox[\textwidth][c]{\input{tables/hiring-shock-reg-fte.tex}}
    \label{tab:hiring-shock-reg}
    \justify
    \footnotesize
    \textbf{Note}: 
    This table show the second stage 2SLS estimates of regression specification \eqref{eqn:secondstage}, showing the effect of state funding changes on the number of faculty hires (per student) total for 2011--2021 at US public universities, using the funding shock to instrument for state funding.
    Each observation is a university--year at a public university, where funding data come from IPEDS and faculty count total for 2011--2021 from \citep{wapman2022quantifying}.
    The panels shows the effect of a $1$\% change in state funding per student at the university (total for 2011--2021) on the number of new faculty hires by gender (and all).
    Standard errors are clustered at the university-year level, and university $+$ year fixed effects are included through--out.
\end{table}

\newpage
\subsection{Additional Results, Rates of Substitution}
\label{sec:appendix-substitution}

The funding elasticities can be used to recover the marginal rate of substitution between two outcomes.
For example, write $Y^1$ for the number of lecturers per student at a university, and $Y^2$ for the number of full professors.
I use the above approaches to estimate the funding elasticities, where $\% \Delta$ denotes percent change.
\[ \beta_1 = \frac{\% \Delta Y^1}{\% \Delta X}
\text{, and }
\beta_2 = \frac{\% \Delta Y^2}{\% \Delta X} \]
As such, it is possible to recover the elasticity for substitution between lecturers and full professors by the universities via the respective funding elasticities.
\[ \frac{\% \Delta Y^1}{\% \Delta Y^2}
= \frac{\% \Delta Y^1 / \% \Delta X}{\% \Delta Y^2 / \% \Delta X}
= \frac{\beta_1}{\beta_2} \]
I present results for the rates of substitution between different levels of faculty by this approach, dividing the relevant coefficient estimates and presenting standard errors calculated by a non-parametric bootstrap.
In practice, this corresponds to division of the estimates of the elasticity for employment of professors (by rank) with respect to state funding, presented in Panel B \autoref{tab:facultycount-shock-reg}, and bootstrapping the results to generate standard errors and confidence intervals.
% https://stackoverflow.com/questions/63777368/computing-the-standard-error-when-dividing-coefficients-of-different-regressions

The implied marginal rate of substitution between lecturers and assistant professors is estimated as -3.26 (standard error 0.50), based on based on 10,000 bootstrap samples.
This means that public universities increased their number lecturers per student by 3.26\% when they decreased their count of assistant professors, on average and subject to the changes in state funding they experienced 1990--2017.
Between lecturers and full professors the rate of subsitution is -3.19 (0.34), which implies that universities subsitute between lecturers and full professors in the same way.
Between assistant and full professors the rate of substitution is 0.99 (0.11), which intuitively implies that universities treated assistant and full professors (i.e., those before and after tenure in the tenure system) as complements.

%"Calculated point est for substitution between lecturers + ast profs" 
%                                                                      
%                                                  "-3.26372574229461" 
%                                                                      
%                                                           "with SEs" 
%                                                                      
%                                                  "0.501721750806993" 
%                                                                      
%                                                        "and 95 % CI" 
%                                                                 2.5% 
%                                                  "-4.38380704259523" 
%                                                                97.5% 
%                                                  "-2.40372605889863"

% "Calculated point est for substitution between lecturers + full profs" 
%                                                                       
%                                                    "-3.1876596143219" 
%                                                                       
%                                                            "with SEs" 
%                                                                       
%                                                   "0.340087353132307" 
%                                                                       
%                                                         "and 95 % CI" 
%                                                                  2.5% 
%                                                   "-3.88637781601822" 
%                                                                 97.5% 
%                                                   "-2.56722681973736"

% "Calculated point est for substitution between ast + full profs" 
%                                                                  
%                                              "0.990674987591129" 
%                                                                  
%                                                       "with SEs" 
%                                                                  
%                                              "0.116874891784044" 
%                                                                  
%                                                    "and 95 % CI" 
%                                                             2.5% 
%                                              "0.771569155168642" 
%                                                            97.5% 
%                                               "1.23151772759759"

### 2023-07-18 text/tables/facultycount-illinois-reg-rawcount.tex
Revision 2 finished (though needs check over of writing).

\begin{tabular}{@{\extracolsep{5pt}}lcccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{8}{c}{Dependent Variable: Employment Count by Professor Group} \\ 
\cline{2-9} 
\\[-1.8ex] & \multicolumn{2}{c}{Lecturer} & \multicolumn{2}{c}{Assistant} & \multicolumn{2}{c}{Full} & \multicolumn{2}{c}{All} \\ 
 & OLS & 2SLS & OLS & 2SLS & OLS & 2SLS & OLS & 2SLS \\ 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6) & (7) & (8)\\ 
\hline \\[-1.8ex] 
 State Funding & $-$3.729 & 0.376 & 0.415 & 3.208 & $-$1.269 & 0.117 & $-$4.084 & 3.890 \\ 
  & (1.721) & (1.725) & (1.728) & (2.493) & (1.548) & (1.200) & (4.352) & (5.668) \\ 
 \hline \\[-1.8ex] 
Outcome Mean & 351.306 & 351.306 & 273.757 & 273.757 & 477.597 & 477.597 & 1303.014 & 1303.014 \\ 
Observations & 144 & 144 & 144 & 144 & 144 & 144 & 144 & 144 \\ 
R$^{2}$ & 0.886 & 0.882 & 0.970 & 0.969 & 0.991 & 0.991 & 0.981 & 0.980 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}

### 2023-07-18 text/tables/facultycount-shock-reg-rawcount.tex
Revision 2 finished (though needs check over of writing).

\begin{tabular}{@{\extracolsep{5pt}}lcccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{8}{c}{Dependent Variable: Faculty Count per 100 Students, by Professor Group} \\ 
\cline{2-9} 
\\[-1.8ex] & \multicolumn{2}{c}{Lecturer} & \multicolumn{2}{c}{Assistant} & \multicolumn{2}{c}{Full} & \multicolumn{2}{c}{All} \\ 
 & OLS & 2SLS & OLS & 2SLS & OLS & 2SLS & OLS & 2SLS \\ 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6) & (7) & (8)\\ 
\hline \\[-1.8ex] 
 State Funding & $-$0.451 & $-$5.957 & $-$0.479 & 1.031 & $-$0.104 & 2.288 & $-$1.198 & $-$3.333 \\ 
  & (0.177) & (1.725) & (0.211) & (1.232) & (0.275) & (2.910) & (0.612) & (4.259) \\ 
 \hline \\[-1.8ex] 
Outcome Mean & 59.253 & 59.253 & 116.121 & 116.121 & 269.103 & 269.103 & 452.507 & 452.507 \\ 
Observations & 17,012 & 17,012 & 17,012 & 17,012 & 17,012 & 17,012 & 17,012 & 17,012 \\ 
R$^{2}$ & 0.742 & 0.595 & 0.887 & 0.881 & 0.973 & 0.971 & 0.954 & 0.954 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}

### 2023-07-18 text/tables/firststage-balance-count.tex
Revision 2 finished (though needs check over of writing).

\begin{tabular}{@{\extracolsep{5pt}}lcccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{8}{c}{Dependent Variables: University Characteristics} \\ 
\cline{2-9}
\\[-1.8ex] & Enrolment & State & Total & Non-inst. & Lecturers & Assistant & Full & All \\ 
\\[-1.8ex] &           & Funding & Revenues & Revenues & Lecturers & Professors & Professors & Professors \\
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6) & (7) & (8)\\ 
\hline \\[-1.8ex] 
 Funding & 0.517   & $-$9,217 & $-$7,633 & 886 & 0.007 & $-$0.001 & $-$0.003 & 0.004 \\ 
 Shock   & (0.199) & (2,452) & (12,771) & (4,093) & (0.002) & (0.001) & (0.003) & (0.005) \\ 
 \hline \\[-1.8ex] 
Uni. + Year \\
fixed effects? & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\
Observations & 17,012 & 17,012 & 17,012 & 17,012 & 17,012 & 17,012 & 17,012 & 17,012 \\ 
R$^{2}$ & 0.960 & 0.950 & 0.881 & 0.932 & 0.745 & 0.886 & 0.973 & 0.954 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}

### 2023-07-18 text/tables/firststage-balance-log.tex
Revision 2 finished (though needs check over of writing).

--- 2023-07-18 text/tables/firststage-balance-count.tex
@@
 \hline \\[-1.8ex] 
- Funding & 0.517   & $-$9,217 & $-$7,633 & 886 & 0.007 & $-$0.001 & $-$0.003 & 0.004 \\ 
- Shock   & (0.199) & (2,452) & (12,771) & (4,093) & (0.002) & (0.001) & (0.003) & (0.005) \\ 
+ Funding & 0.107 & $-$0.870 & 0.031 & $-$0.091 & 0.534 & $-$0.024 & $-$0.027 & 0.043 \\ 
+ Shock   & (0.037) & (0.079) & (0.082) & (0.063) & (0.122) & (0.066) & (0.034) & (0.038) \\ 
 \hline \\[-1.8ex] 
@@
 Observations & 17,012 & 17,012 & 17,012 & 17,012 & 17,012 & 17,012 & 17,012 & 17,012 \\ 
-R$^{2}$ & 0.960 & 0.950 & 0.881 & 0.932 & 0.745 & 0.886 & 0.973 & 0.954 \\ 
+R$^{2}$ & 0.976 & 0.914 & 0.979 & 0.975 & 0.782 & 0.902 & 0.963 & 0.965 \\ 
 \hline 

### 2023-07-18 programs/analysis/hiring-shock.R

@@ -1,0 +1,235 @@
+#!/usr/bin/R
+## Senan Hogan-Hennessy, 3 February 2023
+## IV for hiring of faculty, using IPEDS + hiring networks data.
+print(Sys.time())
+set.seed(47)
+# Functions for data manipulation and visualisation
+library(tidyverse)
+# Functions for fast linear models with IV + FEs
+library(lfe)
+# TeX tables
+library(stargazer)
+# Define number of digits in tables and graphs
+digits.no <- 3
+# Size for figures
+fig.width <- 15
+fig.height <- fig.width * 1 / 2
+
+
+# Load data sources ------------------------------------------------------------
+
+# Load IPEDS data on all universities
+ipeds.data <- read_csv("../../data/urban-ipeds/urban-clean-allunis.csv")
+
+# Load hiring data, which is a sum for years 2010--2021
+hiring.data <- read_csv("../../data/states/hiring-count.csv")
+
+
+# Clean data -------------------------------------------------------------------
+
+# Select IPEDS variables, and define the panel at the university level
+ipeds.data <- ipeds.data %>%
+    # Non for-profit unis, 4-year
+    filter(fouryear == 1, forprofit == 0) %>%
+    # Remove years with incomplete data
+    filter(1990 <= year, year != 2001, year <= 2021) %>%
+    # Restrict to relevant columns and rows.
+    select(
+        # Identifiers and controls
+        unitid,
+        inst_name,
+        year,
+        state,
+        public,
+        # Appropriations to the uni
+        totalrevenues_real,
+        nonauxrevenues_real,
+        stateappropriations_real,
+        tuitionrev_real,
+        # Uni spending.
+        nonauxspending_real,
+        # Uni enrollment
+        enrollment_reported,
+        enrollment_fte,
+        # Appropriation + tuition shocks Per uni
+        appropriationshock_perEnroll_real,
+        appropriationshock_perFTE_real,
+        appropriationshock_peruni_real)
+
+# Sum data for 2011-2020, which the hiring data pertain to.
+ipeds.data <- ipeds.data %>%
+    filter(2011 <= year, year <= 2021) %>%
+    # Overate the sum.
+    group_by(unitid, state, public) %>%
+    summarise(
+        # Appropriations to the uni
+        totalrevenues_real = mean(totalrevenues_real, na.rm = TRUE),
[…]

### 2023-07-18 programs/analysis/ibhed-shock-rawcounts.R

@@ -1,0 +1,479 @@
+#!/usr/bin/R
+## Senan Hogan-Hennessy, 3 August 2022
+## IV for Prof's outcomes, using Illinois data.
+## Using a base-share version of the instrument.
+print(c(Sys.time(), Sys.Date()))
+library(tidyverse) # Functions for data manipulation and visualization
+library(lfe) # Functions for fast linear models with IV + FEs
+library(plm) # Functions for panel data
+library(car) # Function for F stat regarding IV models
+# My custom flavour of Stargazer TeX tables:
+# devtools::install_github("shoganhennessy/stargazer")
+library(stargazer) # TeX tables
+set.seed(47)
+# This file follows an adjusted Deming Walters (2017, p.10) approach to
+# appropriations shock instrument.
+
+# Define number of digits in tables and graphs
+digits.no <- 3
+
+# Size for figures
+fig.width <- 10
+fig.height <- fig.width * 0.85
+
+
+# Load data --------------------------------------------------------------------
+
+# Data on Illinois Professors
+illinois.data <-
+    read_csv("../../data/states/illinois-professors-anonymised.csv")
+
+# Data from IPEDS, with relevant variables
+ipeds.data <- read_csv("../../data/urban-ipeds/urban-clean-publicunis.csv") %>%
+    select(
+        # Identifiers and controls
+        unitid,
+        year,
+        state,
+        # Appropriations to the uni
+        totalrevenues_real,
+        nonauxrevenues_real,
+        stateappropriations_real,
+        tuitionrev_real,
+        # Uni spending.
+        nonauxspending_real,
+        # Uni enrollment
+        enrollment_reported,
+        enrollment_fte,
+        # Appropriation shocks Per student
+        appropriationshock_perEnroll_real,
+        appropriationshock_perFTE_real,
+        appropriationshock_peruni_real,
+        # State-level values
+        allstate_stateappropriations_real,
+        stateEnroll_count)
+
+
+# Clean data -------------------------------------------------------------------
+
+# Join IPEDS data to the individual level
+reg.data <- illinois.data %>%
+    # Join the IPEDS finance data.
+    left_join(ipeds.data, by = c("unitid", "year"))
+
+# Remove IPEDS + Illinois data to save memory
+rm(ipeds.data)
+rm(illinois.data)
+gc()
+
+# Calculate the first and final years in the panel of professors
[…]

### 2023-07-18 programs/analysis/ibhed-shock-rolling.R

@@ -1,0 +1,762 @@
+#!/usr/bin/R
+## Senan Hogan-Hennessy, 3 August 2022
+## IV for Prof's outcomes, using Illinois data.
+## Using a rolling-share variant of the instrument.
+print(c(Sys.time(), Sys.Date()))
+library(tidyverse) # Functions for data manipulation and visualization
+library(lfe) # Functions for fast linear models with IV + FEs
+library(plm) # Functions for panel data
+library(car) # Function for F stat regarding IV models
+# My custom flavour of Stargazer TeX tables:
+# devtools::install_github("shoganhennessy/stargazer")
+library(stargazer) # TeX tables
+set.seed(47)
+# This file follows an adjusted Deming Walters (2017, p.10) approach to
+# appropriations shock instrument.
+
+# Define number of digits in tables and graphs
+digits.no <- 3
+
+# Size for figures
+fig.width <- 10
+fig.height <- fig.width * 0.85
+
+
+# Load data --------------------------------------------------------------------
+
+# Data on Illinois Professors
+illinois.data <- read_csv("../../data/states/illinois-professors-anonymised.csv")
+
+# Data from IPEDS, with relevant variables
+ipeds.data <- read_csv("../../data/urban-ipeds/urban-clean-publicunis.csv") %>%
+    select(
+        # Identifiers and controls
+        unitid,
+        year,
+        state,
+        # Appropriations to the uni
+        totalrevenues_real,
+        nonauxrevenues_real,
+        stateappropriations_real,
+        tuitionrev_real,
+        # Uni spending.
+        nonauxspending_real,
+        # Uni enrollment
+        enrollment_reported,
+        enrollment_fte,
+        # Appropriation shocks Per student
+        appropriationshock_perFTE_real,
+        appropriationshock_peruni_real,
+        # State-level values
+        allstate_stateappropriations_real,
+        stateEnroll_count)
+
+
+# Clean data -------------------------------------------------------------------
+
+# Join IPEDS data to the individual level
+reg.data <- illinois.data %>%
+    # Join the IPEDS finance data.
+    left_join(ipeds.data, by = c("unitid", "year"))
+
+# Generate a Professor's first year
+reg.data <- reg.data %>%
+    arrange(name, year, unitid) %>%
+    group_by(name, unitid) %>%
[…]

### 2023-07-18 programs/analysis/ibhed-shock.R

@@ -1,0 +1,1205 @@
+#!/usr/bin/R
+## Senan Hogan-Hennessy, 3 August 2022
+## IV for Prof's outcomes, using Illinois data.
+## Using a base-share version of the instrument.
+print(c(Sys.time(), Sys.Date()))
+library(tidyverse) # Functions for data manipulation and visualization
+library(lfe) # Functions for fast linear models with IV + FEs
+library(plm) # Functions for panel data
+library(car) # Function for F stat regarding IV models
+# My custom flavour of Stargazer TeX tables:
+# devtools::install_github("shoganhennessy/stargazer")
+library(stargazer) # TeX tables
+set.seed(47)
+# This file follows an adjusted Deming Walters (2017, p.10) approach to
+# appropriations shock instrument.
+
+# Define number of digits in tables and graphs
+digits.no <- 3
+
+# Size for figures
+fig.width <- 10
+fig.height <- fig.width * 0.85
+
+
+# Load data --------------------------------------------------------------------
+
+# Data on Illinois Professors
+illinois.data <-
+    read_csv("../../data/states/illinois-professors-anonymised.csv")
+
+# Data from IPEDS, with relevant variables
+ipeds.data <- read_csv("../../data/urban-ipeds/urban-clean-publicunis.csv") %>%
+    select(
+        # Identifiers and controls
+        unitid,
+        year,
+        state,
+        # Appropriations to the uni
+        totalrevenues_real,
+        nonauxrevenues_real,
+        stateappropriations_real,
+        tuitionrev_real,
+        # Uni spending.
+        nonauxspending_real,
+        # Uni enrollment
+        enrollment_reported,
+        enrollment_fte,
+        # Appropriation shocks Per student
+        appropriationshock_perEnroll_real,
+        appropriationshock_perFTE_real,
+        appropriationshock_peruni_real,
+        # State-level values
+        allstate_stateappropriations_real,
+        stateEnroll_count)
+
+
+# Clean data -------------------------------------------------------------------
+
+# Join IPEDS data to the individual level
+reg.data <- illinois.data %>%
+    # Join the IPEDS finance data.
+    left_join(ipeds.data, by = c("unitid", "year"))
+
+# Remove IPEDS + Illinois data to save memory
+rm(ipeds.data)
+rm(illinois.data)
+gc()
+
+# Calculate the first and final years in the panel of professors
[…]

### 2023-07-18 programs/analysis/ipeds-shock-rawcounts.R

@@ -1,0 +1,331 @@
+#!/usr/bin/R
+## Senan Hogan-Hennessy, 5 July 2023
+## IV for Prof's salaries, using IPEDS data.
+print(Sys.time())
+library(tidyverse) # Functions for data manipulation and visualization
+library(lfe) # Functions for fast linear models with IV + FEs
+library(plm) # Functions for panel data
+library(car) # Function for F stat regarding IV models
+# My custom flavour of Stargazer TeX tables:
+# devtools::install_github("shoganhennessy/stargazer")
+# TeX tables:
+library(stargazer)
+library(xtable)
+set.seed(47)
+# This file follows an adjusted Deming Walters (2017, p.10) approach to
+# Funding Shock instrument.
+
+# Define number of digits in tables and graphs
+digits.no <- 3
+
+# Size for figures
+fig.width <- 10
+fig.height <- fig.width * 0.85
+
+# Define a function to standardise the dplyr lag function
+# --- negative numbers mean a lag of t time periods, positive a lead of t.
+lag_lead <- function(column, t){
+    if (t < 0){
+        lag_column <- dplyr::lag(column, -t)
+    }
+    else if (t >= 0){
+        lag_column <- dplyr::lead(column, t)
+    }
+    return(lag_column)
+}
+
+
+# Load data sources ------------------------------------------------------------
+
+# Load IPEDS data
+ipeds.data <- read_csv("../../data/urban-ipeds/urban-clean-publicunis.csv")
+
+
+# Clean data -------------------------------------------------------------------
+
+# Select IPEDS variables, and define the panel at the university level
+reg.data <- ipeds.data %>%
+    # Create independent columns and rows.
+    select(
+        # Identifiers and controls
+        unitid,
+        year,
+        state,
+        # Appropriations to the uni
+        totalrevenues_real,
+        nonauxrevenues_real,
+        stateappropriations_real,
+        tuitionrev_real,
+        # Uni spending.
+        nonauxspending_real,
+        # Uni enrollment
+        enrollment_reported,
+        enrollment_fte,
+        # Appropriation shocks Per student
+        appropriationshock_perEnroll_real,
+        appropriationshock_perFTE_real,
+        appropriationshock_peruni_real,
[…]

### 2023-07-18 programs/data-build/hiring-build.R

@@ -1,0 +1,106 @@
+#!/usr/bin/R
+## Senan Hogan-Hennessy, 3 February 2023
+print(Sys.time())
+set.seed(47)
+## Take faculty hiring network data from Scince (2022) article,
+# As a measure for uni hiring for the time period 2010--2021
+# functions for data manipulation and visualisation
+library(tidyverse)
+# Fuzzy matching
+library(fuzzyjoin)
+
+
+################################################################################
+## Load raw data, open source from Scieence article
+#! PUT CITATION HERE.
+
+# Load the academic hiring network (edges) data.
+hiring_edges.data <-
+    read_csv("../../data/us-faculty-hiring-networks/data/edge_lists.csv")
+
+
+################################################################################
+## Clean data
+
+hiring_edges.data %>% head(100)
+hiring_edges.data %>% names()
+# Count the sum of hires, by each uni 2011--2020, from the edge data
+hiring.data <- hiring_edges.data %>%
+    transmute(
+        # This inst_id is not the same as the one for IPEDS
+        inst_id = InstitutionId,
+        inst_name = InstitutionName,
+        men_hired = Men,
+        women_hired = Women,
+        total_hired = Total) %>%
+    group_by(inst_id, inst_name) %>%
+    summarise(
+        men_hired = sum(men_hired, na.rm = TRUE),
+        women_hired = sum(women_hired, na.rm = TRUE),
+        total_hired = sum(total_hired, na.rm = TRUE)) %>%
+    ungroup()
+
+
+################################################################################
+## Connect to IPEDS finance data.
+
+# Load constructed IPEDS data, 2011 onwards
+ipeds.data <- read_csv("../../data/urban-ipeds/urban-clean-allunis.csv") %>%
+    filter(year >= 2011)
+gc()
+
+# Get the connecting variables for IPEDs: unitid, college name
+ipeds.key <- ipeds.data %>%
+    #mutate(inst_name = str_sub(inst_name, end = 30)) %>%
+    group_by(unitid, inst_name) %>%
+    summarise(count = n()) %>%
+    ungroup() %>%
+    select(unitid, inst_name)
+
+ipeds.key %>% arrange(unitid, inst_name)
+hiring.data %>% arrange(inst_id, inst_name)
+
+
+# Get the connecting variable for hiring data: college name
+hiring.key <- hiring.data %>%
+    #mutate(inst_name = str_sub(inst_name, end = 30)) %>%
[…]

### 2023-07-18 text/paper.tex
Revision 2 finished (though needs check over of writing).

@@
     } \\ \vspace{0.1cm}
-    \href{https://github.com/shoganhennessy/state-funding-faculty/blob/main/https://github.com/shoganhennessy/state-funding-faculty-2023.pdf}{Newest version available here.}
+    \href{https://github.com/shoganhennessy/state-funding-faculty/blob/main/state-funding-faculty-2023.pdf}{Newest version available here.}
 }

### 2023-07-18 text/sections/01-introduction.tex
Revision 2 finished (though needs check over of writing).

@@
 For faculty outcomes, multiple theoretically model university decision-making and faculty hiring (see e.g., \citealt{abe2015implications,johnson2009jep,NBERc13879}), while others measured how faculty at universities were affected by endowment shocks and the 2008 recession \citep{brown2014endowment,turner2014impact}.
-These papers measure effects of funding and changes in university revenues, but not measure effects on faculty, or other outcomes related to university instruction.
+These papers measure effects of funding and changes in university revenues, but do not measure effects on faculty, or other outcomes related to university instruction.
 

### 2023-07-21 text/sections/05-results.tex
add LP estimates in IBHED data (no effects worth mentioning).

@@
 \autoref{tab:faculty-shock-illinois-rolling} show estimates of a 1\% change in state funding per student on faculty salaries, rate of exiting the Illinois public university system, and rate of promotion.
-The estimates are not discernible from zero, so that faculty who are already at the university (incumbent faculty) seem to have no effects of state funding cuts passed on to them.
+The estimates are not discernible from zero, so that faculty who are already at the university (incumbent faculty) seem to have no effects of state funding cuts passed on to them, in the year of the funding cut.
+In the years following, Illinois professors' salaries are not affected by state funding cuts.
+Lecturers on the other hand, see a 0.75\% increase in salary in the second and third year after a 10\% increase in state funding (\autoref{fig:salaries-illinois-lp-rolling}), reflecting how lecturers are again more affected changes in public university finances.
+Similarly, promotion rate and exit rate among incumbent faculty  (from the Illinois public university system) are unaffected in the years after an initial funding shock (\autoref{fig:promotion-illinois-lp-rolling}, \ref{fig:exit-illinois-lp-rolling}).
 

### 2023-07-21 text/sections/09-figures-tables.tex
add LP estimates in IBHED data (no effects worth mentioning).

@@
     \singlespacing
-    \caption{Local Projection Estimates for Professor Count per Student, by Professor Group.}
+    \caption{Local Projection Estimates for Effect of State Funding on Faculty Count per Student at Universities, by Professor Group.}
     \begin{subfigure}[b]{0.495\textwidth}
@@
     \textbf{Note}:
-    These figures show the local projections estimates of regression specification \eqref{eqn:secondstage}, with the funding shock as an instrument for state funding, using IPEDS data.
+    These figures show the local projections estimates of regression specification \eqref{eqn:secondstage}, with the funding shock as an instrument for state funding
+    The unit of analysis is the university, and uses IPEDS data.
     The coefficient estimate is effect of state funding ($X_{i,t}$) on faculty count per student ($Y_{i,t}$), while accounting for auto-correlation between different time periods --- i.e., between $X_{i,t}, X_{i,t-1}$ and $Y_{i,t}, Y_{i,t-1}$.
@@
     Standard errors are clustered at the state-year level.
+\end{figure}
+
+\newpage
+\begin{figure}[H]
+    \centering
+    \singlespacing
+    \caption{Local Projection Estimates for Effect of State Funding on Faculty Salaries per Student at Illinois Public Universities, by Professor Group.}
+    \begin{subfigure}[b]{0.495\textwidth}
+        \centering
+        \caption{Lecturers.}
+        \includegraphics[width=\textwidth]{figures/salaries-lecturer-illinois-lp-rolling.png}
+        \label{fig:salaries-lecturer-illinois-lp-rolling}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.495\textwidth}
+        \centering
+        \caption{Assistant Professors.}
+        \includegraphics[width=\textwidth]{figures/salaries-assistant-illinois-lp-rolling.png}
+        \label{fig:salaries-assistant-illinois-lp-rolling}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.495\textwidth}
+        \centering
+        \caption{Full Professors.}
+        \includegraphics[width=\textwidth]{figures/salaries-full-illinois-lp-rolling.png}
+        \label{fig:salaries-full-illinois-lp-rolling}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.495\textwidth}
+        \centering
+        \caption{Administrator Professors.}
+        \includegraphics[width=\textwidth]{figures/salaries-administrator-illinois-lp-rolling.png}
+        \label{fig:salaries-administrator-illinois-lp-rolling}
+    \end{subfigure}
+    \label{fig:salaries-illinois-lp-rolling}
+    \justify
+    \footnotesize
+    \textbf{Note}:
+    These figures show the local projections estimates of regression specification \eqref{eqn:secondstage}, with the funding shock as an instrument for state funding.
+    The unit of analysis is an individual faculty member (at an Illinois public university); funding data come from IPEDS, and faculty salaries from IBHED.
+    The coefficient estimate is effect of state funding ($X_{i(j),t}$) on faculty salaries ($Y_{j,t}$), while accounting for auto-correlation between different time periods --- i.e., between $X_{i(j),t}, X_{i(j),t-1}$ and $Y_{i(j),t}, Y_{i(j),t-1}$.
+    These results use a $\log-\log$ specification, so the estimates are for the elasticity of faculty salaries student in a year $t+k$ with respect to state funding in year $t$, where years $k = 0, \hdots, 4$ are on the x-axis.
+    Standard errors are clustered at the university-year level.
 \end{figure}

### 2023-07-21 text/sections/10-appendix.tex
add LP estimates in IBHED data (no effects worth mentioning).

@@
 We then interpret parameter $\beta$ as the effect of changes in state funding at an Illinois public university, via state funding shocks, on an individual professor's outcome $Y_{j,t}$.
+
+\newpage
+\begin{figure}[H]
+    \centering
+    \singlespacing
+    \caption{Local Projection Estimates for Effect of State Funding on Faculty Promotion Rate at Illinois Public Universities, by Professor Group.}
+    \begin{subfigure}[b]{0.495\textwidth}
+        \centering
+        \caption{Lecturers.}
+        \includegraphics[width=\textwidth]{figures/promotion-lecturer-illinois-lp-rolling.png}
+        \label{fig:promotion-lecturer-illinois-lp-rolling}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.495\textwidth}
+        \centering
+        \caption{Assistant Professors.}
+        \includegraphics[width=\textwidth]{figures/promotion-assistant-illinois-lp-rolling.png}
+        \label{fig:promotion-assistant-illinois-lp-rolling}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.495\textwidth}
+        \centering
+        \caption{Full Professors.}
+        \includegraphics[width=\textwidth]{figures/promotion-full-illinois-lp-rolling.png}
+        \label{fig:promotion-full-illinois-lp-rolling}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.495\textwidth}
+        \centering
+        \caption{Administrator Professors.}
+        \includegraphics[width=\textwidth]{figures/promotion-administrator-illinois-lp-rolling.png}
+        \label{fig:promotion-administrator-illinois-lp-rolling}
+    \end{subfigure}
+    \label{fig:promotion-illinois-lp-rolling}
+    \justify
+    \footnotesize
+    \textbf{Note}:
+    These figures show the local projections estimates of regression specification \eqref{eqn:secondstage}, with the funding shock as an instrument for state funding.
+    The unit of analysis is an individual faculty member (at an Illinois public university); funding data come from IPEDS, and faculty promotion rate from IBHED.
+    The coefficient estimate is effect of state funding ($X_{i(j),t}$) on faculty promotion rate ($Y_{j,t}$), while accounting for auto-correlation between different time periods --- i.e., between $X_{i(j),t}, X_{i(j),t-1}$ and $Y_{i(j),t}, Y_{i(j),t-1}$.
+    These results use a rate$-\log$ specification, so the estimates are for the rate of promotion in a year $t+k$ affected by a 1\% change in state funding in year $t$, where years $k = 0, \hdots, 4$ are on the x-axis.
+    Standard errors are clustered at the university-year level.
+\end{figure}
+
+\newpage
+\begin{figure}[H]
+    \centering
+    \singlespacing
+    \caption{Local Projection Estimates for Effect of State Funding on Faculty Exit Rate at Illinois Public Universities, by Professor Group.}
+    \begin{subfigure}[b]{0.495\textwidth}
+        \centering
+        \caption{Lecturers.}
+        \includegraphics[width=\textwidth]{figures/exit-lecturer-illinois-lp-rolling.png}
+        \label{fig:exit-lecturer-illinois-lp-rolling}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.495\textwidth}
+        \centering
+        \caption{Assistant Professors.}
+        \includegraphics[width=\textwidth]{figures/exit-assistant-illinois-lp-rolling.png}
+        \label{fig:exit-assistant-illinois-lp-rolling}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.495\textwidth}
+        \centering
+        \caption{Full Professors.}
+        \includegraphics[width=\textwidth]{figures/exit-full-illinois-lp-rolling.png}
+        \label{fig:exit-full-illinois-lp-rolling}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.495\textwidth}
+        \centering
+        \caption{Administrator Professors.}
+        \includegraphics[width=\textwidth]{figures/exit-administrator-illinois-lp-rolling.png}
+        \label{fig:exit-administrator-illinois-lp-rolling}
+    \end{subfigure}
+    \label{fig:exit-illinois-lp-rolling}
+    \justify
+    \footnotesize
+    \textbf{Note}:
+    These figures show the local projections estimates of regression specification \eqref{eqn:secondstage}, with the funding shock as an instrument for state funding.
+    The unit of analysis is an individual faculty member (at an Illinois public university); funding data come from IPEDS, and faculty promotion rate from IBHED.
+    The coefficient estimate is effect of state funding ($X_{i(j),t}$) on faculty promotion rate ($Y_{j,t}$), while accounting for auto-correlation between different time periods --- i.e., between $X_{i(j),t}, X_{i(j),t-1}$ and $Y_{i(j),t}, Y_{i(j),t-1}$.
+    These results use a rate$-\log$ specification, so the estimates are for the rate of promotion in a year $t+k$ affected by a 1\% change in state funding in year $t$, where years $k = 0, \hdots, 4$ are on the x-axis.
+    Standard errors are clustered at the university-year level.
+\end{figure}
 

### 2023-07-21 programs/analysis/ibhed-shock-longterm.R

@@ -1,0 +1,468 @@
+#!/usr/bin/R
+## Senan Hogan-Hennessy, 21 July 2023
+## IV for Prof's outcomes, using Illinois data.
+## Long term outcomes by local projections method.
+## Using a rolling-share variant of the instrument.
+print(Sys.time())
+library(tidyverse) # Functions for data manipulation and visualisation
+library(lfe) # Functions for fast linear models with IV + FEs
+library(plm) # Functions for panel data
+library(car) # Function for F stat regarding IV models
+# https://cran.r-project.org/web/packages/lpirfs/lpirfs.pdf
+library(lpirfs) # Function for LP regression.
+# My custom flavour of Stargazer TeX tables:
+# devtools::install_github("shoganhennessy/stargazer")
+library(stargazer) # TeX tables
+set.seed(47)
+# This file follows an adjusted Deming Walters (2017, p.10) approach to
+# appropriations shock instrument.
+
+# Define number of digits in tables and graphs
+digits.no <- 3
+
+# Size for figures
+fig.width <- 10
+fig.height <- fig.width * 0.85
+
+
+# Load data --------------------------------------------------------------------
+
+# Data on Illinois Professors
+illinois.data <- read_csv("../../data/states/illinois-professors-anonymised.csv")
+
+# Data from IPEDS, with relevant variables
+ipeds.data <- read_csv("../../data/urban-ipeds/urban-clean-publicunis.csv") %>%
+    select(
+        # Identifiers and controls
+        unitid,
+        year,
+        state,
+        # Appropriations to the uni
+        totalrevenues_real,
+        nonauxrevenues_real,
+        stateappropriations_real,
+        tuitionrev_real,
+        # Uni spending.
+        nonauxspending_real,
+        # Uni enrollment
+        enrollment_reported,
+        enrollment_fte,
+        # Appropriation shocks Per student
+        appropriationshock_perFTE_real,
+        appropriationshock_peruni_real,
+        # State-level values
+        allstate_stateappropriations_real,
+        stateEnroll_count)
+
+
+# Clean data -------------------------------------------------------------------
+
+# Join IPEDS data to the individual level
+reg.data <- illinois.data %>%
+    # Join the IPEDS finance data.
+    left_join(ipeds.data, by = c("unitid", "year"))
+
+# Generate a Professor's first year
+reg.data <- reg.data %>%
[…]

### 2023-07-21 programs/analysis/ibhed-shock-rolling.R

@@ -174,6 +174,7 @@ reg.data <- reg.data %>%
     mutate(unitid = factor(unitid),
         name = factor(name))
 
+
 # Summary Table ----------------------------------------------------------------
 
 # Generate the TeX summary table.

### 2023-07-21 programs/analysis/ipeds-shock-longterm.R

@@ -1,0 +1,272 @@
+#!/usr/bin/R
+## Senan Hogan-Hennessy, 3 August 2022
+## IV for Prof's salaries, using IPEDS data.
+print(Sys.time())
+library(tidyverse) # Functions for data manipulation and visualization
+library(lfe) # Functions for fast linear models with IV + FEs
+library(plm) # Functions for panel data
+# My custom flavour of Stargazer TeX tables:
+# devtools::install_github("shoganhennessy/stargazer")
+library(stargazer) # TeX tables
+set.seed(47)
+# This file follows an adjusted Deming Walters (2017, p.10) approach to
+# appropriations shock instrument.
+
+# Define number of digits in tables and graphs
+digits.no <- 3
+
+# Size for figures
+fig.width <- 10
+fig.height <- fig.width * 0.85
+
+
+# Load data sources ------------------------------------------------------------
+
+# Load IPEDS data
+ipeds.data <- read_csv("../../data/urban-ipeds/urban-clean-publicunis.csv")
+
+
+# Clean data -------------------------------------------------------------------
+
+# Select IPEDS variables, and define the panel at the university level
+reg.data <- ipeds.data %>%
+    # Create independent columns and rows.
+    select(
+        # Identifiers and controls
+        unitid,
+        year,
+        state,
+        # Appropriations to the uni
+        totalrevenues_real,
+        nonauxrevenues_real,
+        stateappropriations_real,
+        tuitionrev_real,
+        # Uni spending.
+        nonauxspending_real,
+        # Uni enrollment
+        enrollment_reported,
+        enrollment_fte,
+        # Appropriation shocks Per student
+        appropriationshock_perEnroll_real,
+        appropriationshock_perFTE_real,
+        appropriationshock_peruni_real,
+        # Count of full-time professors (tenured or not)
+        lecturer_prof_count,
+        assistant_prof_count,
+        full_prof_count,
+        all_prof_count,
+        # Count of professors by tenure-rate (only 1990-)
+        nontenured_tenure_count,
+        tenuretrack_tenure_count,
+        tenured_tenure_count,
+        all_tenure_count,
+        # Average faculty salary
+        full_profmeansalary_real,
+        assistant_profmeansalary_real,
+        lecturer_profmeansalary_real,
+        all_profmeansalary_real,
+        # Total paid on professor salaries
[…]

### 2023-07-21 programs/run.sh

@@ -21,12 +21,16 @@ R CMD BATCH --no-save ipeds-public-private.R
 R CMD BATCH --no-save ipeds-shock.R
 # Analysis and regressions at the institution-level, IPEDS, with count outcomes.
 R CMD BATCH --no-save ipeds-shock-rawcounts.R
+# Analysis and regressions at the institution-level, IPEDS, local porjections.
+R CMD BATCH --no-save ipeds-shock-longterm.R
 # Analysis and regressions at the individual-level, IBHED, base-year shock IV.
 R CMD BATCH --no-save ibhed-shock.R
-# Analysis and regressions at the individual-level, IBHED, with count outcomes.
+# Analysis and regressions at the using IBHED count outcomes, instituion level.
 R CMD BATCH --no-save ibhed-shock-rawcounts.R
 # Analysis and regressions at the individual-level, IBHED, rolling shock IV.
 R CMD BATCH --no-save ibhed-shock-rolling.R
+# Analysis and regressions at the institutionindividual-level, IBHED, local porjections.
+R CMD BATCH --no-save ibhed-shock-longterm.R
 # Analysis and regressions for rate of faculty hiring.
 R CMD BATCH --no-save hiring-shock.R
 cd ..

### 2023-07-21 text/sections/01-introduction.tex
added LP estimates in IBHED data (no effects worth mentioning).

@@
 The shift-share instrument interacts how much a university relies on state funding as a percent of revenues in a base period with the entire state's total funding per higher education student for each year, exploiting changes in funding and how much each university historically relied on state funding for higher education.
+Additionally, I estimate effect of state funding on faculty multiple years after the initial funding cuts, using the local projections method thanks to the presence of time-series confounding between the funding shock and later years' level of state funding.\footnote{
+    To the best of my knowledge, this is also the first paper to use the local projections method for a shift-share instrument.
+    %Previous work ignored the potential time-series confoduning, and estimated linear instrumental variables.
+}
 

### 2023-07-21 text/sections/05-results.tex
added LP estimates in IBHED data (no effects worth mentioning).

@@
 Lecturers on the other hand, see a 0.75\% increase in salary in the second and third year after a 10\% increase in state funding (\autoref{fig:salaries-illinois-lp-rolling}), reflecting how lecturers are again more affected changes in public university finances.
-Similarly, promotion rate and exit rate among incumbent faculty  (from the Illinois public university system) are unaffected in the years after an initial funding shock (\autoref{fig:promotion-illinois-lp-rolling}, \ref{fig:exit-illinois-lp-rolling}).
+Similarly, promotion rate and exit rate among incumbent faculty  (from the Illinois public university system) are unaffected in the years after an initial funding shock (\autoref{fig:promoted-illinois-lp-rolling}, \ref{fig:exit-illinois-lp-rolling}).
 

### 2023-07-21 text/sections/10-appendix.tex
added LP estimates in IBHED data (no effects worth mentioning).

@@
         \caption{Lecturers.}
-        \includegraphics[width=\textwidth]{figures/promotion-lecturer-illinois-lp-rolling.png}
-        \label{fig:promotion-lecturer-illinois-lp-rolling}
+        \includegraphics[width=\textwidth]{figures/promoted-lecturer-illinois-lp-rolling.png}
+        \label{fig:promoted-lecturer-illinois-lp-rolling}
     \end{subfigure}
@@
         \caption{Assistant Professors.}
-        \includegraphics[width=\textwidth]{figures/promotion-assistant-illinois-lp-rolling.png}
-        \label{fig:promotion-assistant-illinois-lp-rolling}
+        \includegraphics[width=\textwidth]{figures/promoted-assistant-illinois-lp-rolling.png}
+        \label{fig:promoted-assistant-illinois-lp-rolling}
     \end{subfigure}
@@
         \caption{Full Professors.}
-        \includegraphics[width=\textwidth]{figures/promotion-full-illinois-lp-rolling.png}
-        \label{fig:promotion-full-illinois-lp-rolling}
+        \includegraphics[width=\textwidth]{figures/promoted-full-illinois-lp-rolling.png}
+        \label{fig:promoted-full-illinois-lp-rolling}
     \end{subfigure}
-    \begin{subfigure}[b]{0.495\textwidth}
-        \centering
-        \caption{Administrator Professors.}
-        \includegraphics[width=\textwidth]{figures/promotion-administrator-illinois-lp-rolling.png}
-        \label{fig:promotion-administrator-illinois-lp-rolling}
-    \end{subfigure}
-    \label{fig:promotion-illinois-lp-rolling}
+    \label{fig:promoted-illinois-lp-rolling}
     \justify

### 2023-07-21 programs/analysis/ibhed-shock-longterm.R

@@ -191,6 +191,7 @@ lp.data <- reg.data %>%
         academicyear = factor(year),
         lecturer = lecturer,
         assistant = assistant,
+        associate = associate,
         full = full,
         administrator = administrator,
         salary_real = log(salary_real + extra_salary_real),
@@ -228,9 +229,7 @@ lpreg.plot <- function(model.lpreg) {
         geom_line(aes(y = conf.high), linetype = "dashed") +
         scale_x_continuous(name = "Years, Relative to Initital Shock",
             breaks = model.lpdata$t, expand = c(0.01, 0.01)) +
-        scale_y_continuous(name = "",
-            limits = c(-0.05, 0.1),
-            breaks = seq(-2, 1, by = 0.025)) +
+        scale_y_continuous(name = "") +
         theme_bw() +
         ggtitle("Estimate") +
         theme(plot.title = element_text(size = rel(1)),
@@ -266,7 +265,7 @@ ggsave("../../text/figures/firststage-illinois-lp-rolling.png",
 
 # LP estimation for salaries of lecturers.
 lecturer_salaries.lpreg <-
-    lp_lin_panel(data_set = filter(lp.data, lecturer == 1),
+    lp_lin_panel(data_set = lp.data %>% filter(lecturer == 1),
         # Outcome variable
         endog_data = "salary_real",
         # Predictor variable
@@ -284,12 +283,15 @@ lecturer_salaries.lpreg <-
         hor = time.horizon)
 # Save this plot
 ggsave("../../text/figures/salaries-lecturer-illinois-lp-rolling.png",
-    plot = lpreg.plot(lecturer_salaries.lpreg),
+    plot = (lpreg.plot(lecturer_salaries.lpreg) +
+        scale_y_continuous(name = "",
+            limits = c(-0.05, 0.1),
+            breaks = seq(-2, 1, by = 0.025))),
     units = "cm", width = fig.width, height = fig.height)
 
 # LP estimation for salaries of APs.
 assistant_salaries.lpreg <-
-    lp_lin_panel(data_set = filter(lp.data, assistant == 1),
+    lp_lin_panel(data_set = lp.data %>% filter(assistant == 1),
         # Outcome variable
         endog_data = "salary_real",
         # Predictor variable
@@ -307,12 +309,15 @@ assistant_salaries.lpreg <-
         hor = time.horizon)
 # Save this plot
 ggsave("../../text/figures/salaries-assistant-illinois-lp-rolling.png",
[…]

### 2023-08-10 text/sections/01-introduction.tex
adjusted writing

@@
 There are various margins through which funding cuts can impact a university's faculty.
-A university's level of resources affect the ability that departments can hire new professors, retain their current faculty, or provide retirement incentives.
-So that it is natural to expect that funding cuts may affect the composition of professors at a university.
+A university's level of resources affect the ability that departments can hire new professors, retain their current faculty, or provide retirement incentives, so it is natural to expect that funding cuts may affect the composition of professors at a university.
 Similarly, the wages a university pays their faculty is affected by funding for the university, including the starting salary it offers to new professors and the amount of yearly raises (if any) for incumbent faculty.
@@
 
-I use data from the Integrated Postsecondary Education Data System (IPEDS) to study the impact of changes in state funding for US public univeristies 1990--2017.
-I combine these data with panel data for every public university faculty member in the state of Illinois for 2010--2021 (Illinois Board of Hdigher Education Database, IBHED), to measure how faculty are affected by changes in state funding.
-These data allow me to answer the following research question: if a US public university receives an extra \$1,000 for each of their students (or a 10\% increase), then do they adjust the number of and composition of faculty they employ, and how does this affect individual faculty already at the university?
+I use data from the \cite{ipeds} to study the impact of changes in state funding for US public universities 1990--2017.
+I combine these data with panel data for every public university faculty member in the state of Illinois for 2010--2021 \citep{ibhed}, to measure how faculty are affected by changes in state funding.
+These data allow me to answer the following research question: if a US public university receives an extra \$1,000 for each of their students (or a 10\% increase), do they change the composition of faculty they employ?
+If so, which positions do universities substitute away from, or towards, and do these funding cuts affect the faculty members themselves?
 
@@
 A funding cut of \$1,000 per student leads to an average university employing 6 more lecturers;
-in percentage terms, a (funding per student) cut of 10\% leads to a fall of 1.4\% in the number of assistant professors per student at a university, a fall of 1.2\% for tenured professors, and an increase in 4.4\% the number of lecturers per student.
+in percentage terms, a funding cut of 10\% per student leads to a fall of 1.4\% in the number of assistant professors per student at a university, a fall of 1.2\% for tenured professors, and an increase in 4.4\% the number of lecturers per student.
 Local projection estimates show that these effects linger for three to four years after the initial funding shock, showing that the effect is not isolated to the year of the funding shock. 

### 2023-08-10 text/sections/03-concepts.tex
adjusted writing

@@
 Employment composition changes likely arise via differences in hiring rate across levels of professor seniority.
-Universities hire new professors most years, either to expand their departments or to replace leaving professors, but not hire at the same rate for each level of seniority.
-The number of professors hired in each year is usually highest among non-tenured adjunct faculty, as these instructors are mostly on short-term or contingent contracts, so  are less costly to hire (or to let go) in response to yearly changes.
+Universities hire new professors most years, either to expand their departments or to replace leaving professors, but do not hire at the same rate for each level of seniority.
+The yearly hiring is usually highest among non-tenured adjunct faculty, where universities hire instructors often on short-term or contingent contracts, so that lecturers are less costly to hire (or to let go) in response to yearly changes.
 Tenure-track assistant professors are mostly hired with a four to six year contract, and formal agreement for tenure consideration at the end of this term.
@@
 University departments often employ faculty on multiple-year contracts (e.g., the tenure contract), and individual professors' decisions to retire, move university, or leave academia are multiple-year commitments, so that it may take multiple years for university funding to change the faculty composition, or trickle down to faculty.
-For example, university departments coordinate hiring in year-long cycles, so that a budget short-fall in year $t$ may have no effect on the in-progress hiring committee and resulting hiring decisions, yet the hiring cycle in year $t+1$ may be postponed or cancelled in response.
+For example, university departments coordinate hiring in annual cycles, so that a budget short-fall in year $t$ may have no effect on the in-progress hiring committee and resulting hiring decisions, yet the hiring cycle in year $t+1$ may be postponed or cancelled in response.
 As such, this paper investigates the dynamics of state funding shocks' effects on faculty, for the years following a university funding shock.

### 2023-08-10 text/sections/04-empirical.tex
adjusted writing

@@
 The shift-share instrument has a strong first stage for universities' yearly state funding.
-\autoref{tab:firststage-reg} presents results of the first-stage regression, separately with and without a control for tuition revenue per student, plus institution and year fixed effects;
-Panel A shows estimates of a funding shock of \$1 per student in the state on state funding per student at the university, and Panel B the \% increase effects of a 1\% increase in the shock per student on state funding per student.
-Columns (1) shows that a funding shock of \$10 (per student in the entire state) is associated with \$11.76 less state funding per student at the univeristy, with the corresponding -10\% funding shock leads to -9.77\%
-less funding (coumn 1, panel B), with similar estimates with and without including fixed effects.
+\autoref{tab:firststage-reg} presents results of the first-stage regression, separately with and without a control for tuition revenue per student, plus institution and year fixed effects.
+\autoref{tab:firststage-reg} Panel A shows estimates of a funding shock of \$1 per student in the state on state funding per student at the university, and Panel B the \% increase effects of a 1\% increase in the shock per student on state funding per student.
+Column (1) shows that a funding shock of \$10 (per student in the entire state) is associated with \$11.76 less state funding per student at the university, with the corresponding -10\% funding shock leads to -9.77\%
+less funding (column 1, panel B), with similar estimates with and without including fixed effects.
 Column (2) shows estimates of the first-stage without including fixed effects, and gives less precise estimates for the funding shock, likely thanks to systematic differences in universities unaccounted for without fixed effects.

### 2023-08-10 text/sections/10-appendix.tex
adjusted writing

@@
     This figure shows the correlation between state funding in year $t+k$ with the funding shock in year $t$ for a university, where $k = -7, \hdots, 5$ are the years on the x-axis.
-    This shows that state funding and the funding shock have dynamic correaltion, so that dynamic effects must be estimated by localc projects -- and not simple OLS or 2SLS.
-    The estimates are of \eqref{eqn:firststage}, calculated with IPEDS data, separatly for each year relative to initial shock, using the $\log-\log$ specification, including fixed effects for university $+$ year, and clustering standard errors by university $+$ year.
+    This shows that state funding and the funding shock have dynamic correlation, so that dynamic effects must be estimated by local projections -- and not simple OLS or 2SLS.
+    The estimates are of \eqref{eqn:firststage}, calculated with IPEDS data, separately for each year relative to initial shock, using the $\log-\log$ specification, including fixed effects for university $+$ year, and clustering standard errors by university $+$ year.
 \end{figure}
@@
     Each observation is a university--year at a public university, where funding data come from IPEDS and faculty count total for 2011--2021 from \citep{wapman2022quantifying}.
-    The panels shows the effect of a $1$\% change in state funding per student at the university (total for 2011--2021) on the number of new faculty hires by gender (and all).
+    The panels show the effect of a $1$\% change in state funding per student at the university (total for 2011--2021) on the number of new faculty hires by gender (and all).
     Standard errors are clustered at the university-year level, and university $+$ year fixed effects are included through--out.
@@
 
-The implied marginal rate of substitution between lecturers and assistant professors is estimated as -3.26 (standard error 0.50), based on based on 10,000 bootstrap samples.
+The implied marginal rate of substitution between lecturers and assistant professors is estimated as -3.26 (standard error 0.50), based on 10,000 bootstrap samples.
 This means that public universities increased their number lecturers per student by 3.26\% when they decreased their count of assistant professors, on average and subject to the changes in state funding they experienced 1990--2017.
-Between lecturers and full professors the rate of subsitution is -3.19 (0.34), which implies that universities subsitute between lecturers and full professors in the same way.
+Between lecturers and full professors the rate of substitution is -3.19 (0.34), which implies that universities substitute between lecturers and full professors in the same way.
 Between assistant and full professors the rate of substitution is 0.99 (0.11), which intuitively implies that universities treated assistant and full professors (i.e., those before and after tenure in the tenure system) as complements.

### 2023-08-10 text/sections/04-empirical.tex
adjusted writing

@@
 For example, it may be the case that a state government restricts funding for only its lowest performing universities in response to a budget shock, introducing a treatment selection-on-outcome threat to identification.
-Importantly for this analysis, revenues received from an institution's state government provide opportunity to address such endogeneity.
+Importantly, funding from a university's state government provides opportunity to address such endogeneity.
 
@@
 
-This analysis uses a shift-share instrument to address endogeneity concerns for the amount of funding a state allots to each of its public universities.
+I use a shift-share instrument to address endogeneity concerns for the amount of funding a state allots to each of its public universities.
 \cite{NBERw23736,NBERw27885} develop the instrument for public university finances by exploiting a shift-share instrument for changes in state-level funding interacted with university reliance on state funding in a base period.
@@
 }
-$Z_{i,t}$ is constructed as negative to reflect the fact that the long term trend in, and most of the short-run shocks to, state funding for higher education has been negative.
+$Z_{i,t}$ is constructed as negative to reflect the fact that the long term trend in, and most of the short-run shocks to, state funding for higher education have been negative.\footnote{
+    When used in log terms, the instrument is the negative of the logged shock --- i.e., $- \log \left( -Z_{i,t} \right)$ as opposed to log of the negative shock $\log Z_{i,t}$ directly.
+}
 State funding has been falling, so that the instrument describes shocks to university revenues, mostly in a negative direction.\footnote{

### 2023-08-11 text/sections/00-abstract.tex
adjusted writing

@@
-US universities have systematically substituted away from tenure-track/tenured professors towards contingent lecturers for the last thirty years, while states have limited how much they fund higher education.
+US public universities universities have systematically substituted away from tenure-track/tenured professors towards contingent lecturers for the last thirty years, while their states have systematically reduced the funding to these universities relative to enrolment.
 This paper connects the changes the faculty composition at US public universities to the fall in the states' funding for their public universities, using a shift-share approach to instrument for state funding and identify the causal effects of stagnating funding for higher education.
 A fall in funding of \$1,000 per student leads to univeristies employing 6 more lecturers on average; relative to enrolment, a fall of $10$\% in funding per student decreases the number of assistant professors per student by $1.4$\% and full professors by $1.2$\%, yet increases the number of lecturers per student by $4.4$\%.
-Dynamic estimates show that state funding cuts affect composition of faculty for the next 3 years, showing that funding cuts have lingering effects.
-Analysis of all the professors at Illinois public universities shows that incumbent professors' salaries, promotion rate, and quit rate are all are not affected by funding cuts over 2011-2021, implying that the composition change arose by limiting the hiring of new professors to replace leaving tenure-track/tenured professors.
-These results show the institutional effects of stagnating state funding for higher education, and raise questions about the direction that public education heads as these financial headwinds show no sign of dissipating.
+Dynamic estimates show that these effects linger multiple years after the initial funding shock.
+Analysis of all faculty at Illinois public universities shows that incumbent professors' salaries, promotion rate, and quit rate are all are not affected by funding cuts over 2011-2021, implying that the composition change arose by limiting the hiring of new tenure-track/tenured professors to replace leaving tenure-track/tenured professors.
+These results show the institutional effects of stagnating state funding for higher education, and exhibit one mechanism for worsening student outcomes at US public universities over this same time period.
 

### 2023-08-11 text/sections/01-introduction.tex
adjusted writing

@@
 I use data from the \cite{ipeds} to study the impact of changes in state funding for US public universities 1990--2017.
-I combine these data with panel data for every public university faculty member in the state of Illinois for 2010--2021 \citep{ibhed}, to measure how faculty are affected by changes in state funding.
+I combine these data with panel data for every faculty member at Illinois public universities for 2010--2021 \citep{ibhed}, to measure how incumbent faculty (i.e., those already employed at the university) are affected by changes in state funding.
 These data allow me to answer the following research question: if a US public university receives an extra \$1,000 for each of their students (or a 10\% increase), do they change the composition of faculty they employ?

### 2023-08-14 text/sections/04-empirical.tex
adjusted writing

@@
 I employ a local projections approach to model how faculty outcomes in years $t+k$ are affected by state funding in year $t$, for future years $k > 0$ \citep{jorda2005}.
-The local projections method is an empirical model used to estimate dynamic treatment effects when the treatment is not binary, so that time-series confounding (i.e., treatment in time $t$ is correlated with treatment in time $t-1$) is present, even for instrumental variable models \citep{miller2022}.
+The local projections method is an empirical model used to estimate dynamic treatment effects when the treatment is not binary, so that time-series confounding (i.e., treatment in time $t$ is correlated with treatment in time $t-1$) is present, even for instrumental variable models \citep{montiel2021local}.
 This method estimates the effect of treatment $X_{i,t}$ on outcome $Y_{i, t+ k}$ in follow-on years $k > 0$, while accounting for the auto-correlation between the other years.\footnote{
@@
 }
+Additionally, the approach accomodates use of an instrument for the regressor \citep{olea2021inference}, so that I use the shift-share instument as part of the local projections estimation.
 I present estimates in graphical form, where the x-axis represents the year relative to the funding shock (i.e., $x = 0$ represents the year of the state funding shock) and the y-axis represents the estimated effect of the funding shock on that year's outcome.\footnote{

### 2023-08-15 text/sections/00-abstract.tex
adjusted writing

@@
 Analysis of all faculty at Illinois public universities shows that incumbent professors' salaries, promotion rate, and quit rate are all are not affected by funding cuts over 2011-2021, implying that the composition change arose by limiting the hiring of new tenure-track/tenured professors to replace leaving tenure-track/tenured professors.
-These results show the institutional effects of stagnating state funding for higher education, and exhibit one mechanism for worsening student outcomes at US public universities over this same time period.
+These results show the institutional effects of stagnating state funding for higher education, and exhibit one possible mechanism for deteriorating student outcomes at US public universities over this same time period.
 

### 2023-08-15 text/sections/05-results.tex
adjusted writing

--- 2023-07-13 text/sections/05-results.tex
@@
 The estimates are not discernible from zero, so that faculty who are already at the university (incumbent faculty) seem to have no effects of state funding cuts passed on to them.
+Additionally, local projection estimates show that professors' salaries are also not affected on average in yeard following a funding shock  (\autoref{fig:salaries-illinois-lp-rolling}).
+Lecturers on the other hand, see a 0.75\% increase in salary in the second and third year after a 10\% increase in state funding (\autoref{fig:salaries-lecturer-illinois-lp-rolling}), reflecting how lecturers are again more affected by changes in public university funding.
+Similarly, promotion rate and exit rate among incumbent faculty  (from the Illinois public university system) are unaffected in the years after an initial funding shock (\autoref{fig:promoted-illinois-lp-rolling}, \ref{fig:exit-illinois-lp-rolling}).
 
-%These findings show that public universities increase (decrease) their count of tenure-track and tenured professors per student in years when revenues are more (less) plentiful, possibly by increasing hiring intensity or efforts to maintain incumbent professors.
-%In the same vein, when funding increases, count of tenure-track and tenured professors per student increases --- and the opposite in years of funding cuts.
-%These findings are in line with the observation that universities froze hiring for a couple of years in response to 2008 negative budget shocks, particularly for tenure-track positions \citep{turner2014impact}, and similarly from shocks to university endowments \citep{brown2014endowment}.
-%For lecturers, we see a negative effect which lines up with two trends noted in \autoref{sec:trends}: funding for public universities (per student) is decreasing while relative usage of lecturers increased substantially.
+These findings show that public universities increase (decrease) their count of tenure-track and tenured professors per student in years when revenues are more (less) plentiful.
+In the same vein, when funding increased, count of tenure-track and tenured professors per student increased --- and the opposite in years of funding cuts.
+These findings are in line with the observation that universities froze hiring for a couple of years in response to 2008 negative budget shocks, particularly for tenure-track positions \citep{turner2014impact}, and similarly from shocks to university endowments \citep{brown2014endowment}.
+For lecturers, we see a negative effect which lines up with two trends noted in \autoref{sec:trends}: funding for public universities (per student) decreased while relative usage of lecturers increased substantially.
 
 The composition of faculty changed at public universities 1990--2017 thanks to stagnating state funding, but there are little discernible effects on incumbent faculty.
-This may imply that the rate that public universities hired faculty was impacted by state funding cuts.
+This implied that changes in faculty compoisition arose by state funding cuts impacting the rate that public universioties hired new professors thanks.
 I take recent data on the total faculty hired at US public universities over 2011--2021 (provided by \citealt{wapman2022quantifying}), to show that universities with lower state funding (per student) also hired fewer professors in the same time period --- shown in \autoref{fig:hiring-correlation}.
 Similarly, the funding shock IV model shows that a $10$\% cut in per student state funding leads to $13$\% fewer faculty hires per student across the decade 2011--2021.\footnote{
+    This limited sample of data are purel;y a cross-section for count of professor hires 2011--2021, so that local projection estimates are not possible here.
     See \autoref{sec:appendix-hiring} for further details.
 }
-These results provide limited evidence that changes in faculty composition arose by disrupting hiring of new professors at public universities, corresponding to the results that incumbent faculty were not by meaningfully effected by the funding cuts.
+These results provide limited evidence that changes in faculty composition arose by disrupting hiring of new professors at public universities, corresponding to the concurrent results that faculty composition changed while incumbent faculty were not by meaningfully affected by the funding cuts.

### 2023-08-15 text/sections/06-discussion.tex
re-writing
add LP estimates in IBHED data (no effects worth mentioning).
added LP estimates in IBHED data (no effects worth mentioning).
adjusted writing

@@
 \section{Discussion}
+\label{sec:discussion}
 
+%\textbf{implication: public unis are increasing sheer count of lecturers, in response to enrolment, and not their tenured faculty.
+%SO prof per student is where the movemenet is most stark.}
 Universities received less state funding, relative to increasing enrolment, and reacted by increasing their reliance on lecturers relative to assistant and full professors in both the short and medium-run; these results imply that lecturers can be considered substitutes for professors in the tenure-system.\footnote{
@@
 Lecturers are often employed on short-term or part-time contracts (adjunct), with limited job stability.
-While we should worry for faculty, and their working conditions, there is credible evidence that relying on adjunct lecturers leads to worse student outcomes, relative to full-time lecturers and professors \citep{zhu2021limited}.
-As tenure has increasingly become a private sector phenomenon, we should worry about the long-term effects on higher education, and the effect on the average student who attends a public university in the US.
+While we should worry about the effect on faculty working conditions, there is credible evidence that relying on adjunct lecturers leads to worse education and student outcomes, relative to full-time lecturers and professors \citep{zhu2021limited}.
+As tenure has increasingly become a private sector phenomenon, and the majority of US higher education students attend public institutions, we should worry about the long-term effects of stagnating funding for higher education in the US.
 

### 2023-08-15 text/sections/07-conclusion.tex
adjusted writing

@@
 
-This analysis investigates how the recent stagnation in state support for higher education has affected faculty, and their composition, at US public universities.
+This paper investigates how the recent stagnation in state support for higher education has affected faculty, and their composition, at US public universities.
 This work contributes to the literature along two primary dimensions.
@@
 
-I found that public universities have systematically substituted away from tenure-track and tenured professors, towards lecturers, in the face of persistent declines in state funding for higher education.
-The effects on incumbent professors in the state of Illinois are non-distinguishable from zero, while suggestive information shows falls in hiring, which together implies that the changes in faculty composition are driven by reduced hiring of new professors.
+Public universities have systematically substituted away from tenure-track and tenured professors, towards lecturers, in the face of persistent declines in state funding for higher education.
+The effects on incumbent professors in the state of Illinois are non-distinguishable from zero, with suggestive evidence for falls in faculty hiring, implying that the changes in faculty composition are driven by reduced hiring of new professors.
 Public universities are using more contingent lecturers to teach their students, while private universities continue to employ more tenure-track and tenured professors than their public counterparts, and each year the gap widens.
 
-While costs education have been rising in the US, public universities have also dealt with declining state funding.
-I show that these headwinds led to systematic change at public universities, changes that affect their faculty, likely limiting public universities in their goals in research and education.
-These results show large changes in faculty composition, and that stagnation in state support explains at least a third of the observed shift away from tenured professors and towards contingent lecturers.
+At the same time as higher education costs have been rising in the US, public universities have also dealt with declining state funding.
+I show that these headwinds led to systematic change at public universities, changes that affect their faculty, and likely limiting public universities in their goals in research and education.
+These results show large changes in faculty composition, and that stagnation in state support explains at least a third of the observed shift away from tenured professors and towards lecturers.
 At the same time, private universities were not exposed to financial headwinds of the same magnitude or persistence.
-While public universities continue to educate the majority of higher education students in the US, we should worry about the effects of restricting their funding has on faculty composition, research at public universities, and the impact on higher education as a whole.
+While public universities continue to educate the majority of higher education students in the US, we should worry about the effects of restricting their funding has on faculty, and the wider impact on higher education as a whole.

### 2023-08-15 text/sections/10-appendix.tex
adjusted writing

@@
 This project used data which are fully public, and computational tools which are fully open-source.
-As such, all code and data involved in this project are available at this project's Github repository, available at \url{https://github.com/shoganhennessy/state-faculty-composition}.
+As such, all code and data (anonymised versions where necessary) involved in this project are available at this project's Github repository, available at \url{https://github.com/shoganhennessy/state-faculty-composition}.
 They may be used for replication, or as the basis for further work, as needed.

### 2023-08-15 text/sections/01-introduction.tex
adjusted writing.

@@
 Additionally, I estimate effect of state funding on faculty multiple years after the initial funding cuts, using the local projections method thanks to the presence of time-series confounding between the funding shock and later years' level of state funding.\footnote{
-    To the best of my knowledge, this is also the first paper to use the local projections method for a shift-share instrument.
+    To the best of my knowledge, this is also the first paper to use the local projections method for a shift-share instrument, and to acknowledge the time-series confounding for shift-share IV models when the outcome is non-contemporaneous.
     %Previous work ignored the potential time-series confoduning, and estimated linear instrumental variables.

### 2023-08-15 text/sections/04-empirical.tex
adjusted writing.

@@
 Though on a per student basis, there is little difference across the distribution (except in the outcomes under consideration).
-The state funding shock instrument is positively associated with the total enrolment and total amount of state funding for each university, in both \$ and $\log$/percentage change terms, while other the other sources of finances fopr the university are not associated with the funding shock.
+The state funding shock instrument is positively associated with the total enrolment and total amount of state funding for each university, in both \$ and $\log$/percentage change terms, while other the other sources of university finances are not associated with the funding shock.
 So that the other sources of finances are not clear confounders for the instrumental variables strategy, as they exhibit balance with respect to the funding shock instrument \citep{pei2019poorly}.
@@
 While most states are not exactly the same to California and Illinois, this is example is telling for the phenomenon of stagnating state funding:
-states vary in how much the funding higher education in 1990, but most have experienced a decline on order of 30\%, so that the stagnation in state funding has been a percent change trend over this time period.
+states vary in their absolute funding for higher education in 1990, but most have experienced a decline of similar order to 30\%, so that the stagnation in state funding has been a percent change trend over this time period.
 As such, I include regression specifications where the explanatory and outcome variables are $\log$ transformed, and refer to these when stating results in percent change terms.
 
-\subsection{Effects in Years After the Funding Shock}
-\label{sec:empirical-substituion}
-
-The effects of a change in the universities funding may not be immediate, and faculty in may be effected multiple years after a funding shock.
+\subsection{Effects in Years After the Initial Funding Shock}
+\label{sec:local-projections}
+The effects of a change in the universities funding may not be immediate, and faculty may be affected multiple years after a funding shock.
 Yet, the funding shock instrument is significantly auto-correlated year-on-year; a large state-level funding shock in year $t$ is also likely to experience a large funding shock in year $t-1$.
-Similarly, funding shocks to higher education in year are highly correlated with state funding per student in the 5 years before and after the original shock in IPEDS data (see \autoref{fig:lag-firststage}).
+Similarly, funding shocks to higher education in a year are highly correlated with state funding per student in the 5 years before and after the initial shock in IPEDS data (see \autoref{fig:lag-firststage}).
 Thus, a linear model correlated state funding in year $t$ with outcomes in year $t$ will suffer from time-series confounding.
@@
 I employ a local projections approach to model how faculty outcomes in years $t+k$ are affected by state funding in year $t$, for future years $k > 0$ \citep{jorda2005}.
-The local projections method is an empirical model used to estimate dynamic treatment effects when the treatment is not binary, so that time-series confounding (i.e., treatment in time $t$ is correlated with treatment in time $t-1$) is present, even for instrumental variable models \citep{montiel2021local}.
+The local projections method is an empirical model used to estimate dynamic treatment effects when the treatment is not binary, so that time-series confounding (i.e., treatment in time $t$ is correlated with treatment in time $t-1$) is present \citep{montiel2021local}.
 This method estimates the effect of treatment $X_{i,t}$ on outcome $Y_{i, t+ k}$ in follow-on years $k > 0$, while accounting for the auto-correlation between the other years.\footnote{
@@
 }
-Additionally, the approach accomodates use of an instrument for the regressor \citep{olea2021inference}, so that I use the shift-share instument as part of the local projections estimation.
+Additionally, the approach accomodates use of an instrument for state funding \citep{olea2021inference}, so that I use the shift-share instument in local projections estimates as in the rest of the empirical analysis.
 I present estimates in graphical form, where the x-axis represents the year relative to the funding shock (i.e., $x = 0$ represents the year of the state funding shock) and the y-axis represents the estimated effect of the funding shock on that year's outcome.\footnote{

### 2023-08-15 text/sections/05-results.tex
adjusted writing.

@@
 This implied that changes in faculty compoisition arose by state funding cuts impacting the rate that public universioties hired new professors thanks.
-I take recent data on the total faculty hired at US public universities over 2011--2021 (provided by \citealt{wapman2022quantifying}), to show that universities with lower state funding (per student) also hired fewer professors in the same time period --- shown in \autoref{fig:hiring-correlation}.
+I take recent data on the total faculty hired at US public universities over 2011--2021 (provided by \citealt{wapman2022quantifying,wapman2022zenodo}), to show that universities with lower state funding (per student) also hired fewer professors in the same time period --- shown in \autoref{fig:hiring-correlation}.
 Similarly, the funding shock IV model shows that a $10$\% cut in per student state funding leads to $13$\% fewer faculty hires per student across the decade 2011--2021.\footnote{

### 2023-08-15 text/sections/09-figures-tables.tex
adjusted writing.

@@
     Non-institutional revenues refers to the sum of federal, state, and local funding plus tuition revenues; these sum to the majority of university funding, but exclude numbers such as university income from capital projects.
-    These figures are calculated in IPEDS data.
+    These figures are calculated with IPEDS data.
 \end{figure}
@@
     Most of the higher education enrolment increase for the last 30 years was in public universities, who continue to enrol the vast majority of higher education students in the US.
-    These figures are calculated in IPEDS data.
+    These figures are calculated with IPEDS data.
 \end{figure}
@@
     This figure shows the average number of students per faculty member, by different faculty position, at US public universities.
-    E.g., panel A calculates the mean of (lecturer count) / (student enrolment) at US public universities, for each year of 1990--2017, to show the average trend in faculty composition.
-    These figures are calculated in IPEDS data.
+    E.g., panel A calculates the mean of (lecturer count) / (student enrolment) at US public universities, for each year of 1990--2017, to show the average trend in faculty composition compared to student enrolment.
+    These figures are calculated with IPEDS data.
 \end{figure}
@@
     The unit of analysis is the university, and uses IPEDS data.
-    The coefficient estimate is effect of state funding ($X_{i,t}$) on faculty count per student ($Y_{i,t}$), while accounting for auto-correlation between different time periods --- i.e., between $X_{i,t}, X_{i,t-1}$ and $Y_{i,t}, Y_{i,t-1}$.
+    The coefficient estimate is effect of state funding ($X_{i,t}$) on faculty count per student ($Y_{i,t}$), using the funding shock instrument ($Z_{i,t}$), while accounting for auto-correlation between different time periods --- i.e., between $X_{i,t}, X_{i,t-1}$ and $Y_{i,t}, Y_{i,t-1}$.
     These results use a $\log-\log$ specification, so the estimates are for the elasticity of professor count per student in a year $t+k$ with respect to state funding in year $t$, where years $k = 0, \hdots, 10$ are on the x-axis.
@@
     The unit of analysis is an individual faculty member (at an Illinois public university); funding data come from IPEDS, and faculty salaries from IBHED.
-    The coefficient estimate is effect of state funding ($X_{i(j),t}$) on faculty salaries ($Y_{j,t}$), while accounting for auto-correlation between different time periods --- i.e., between $X_{i(j),t}, X_{i(j),t-1}$ and $Y_{i(j),t}, Y_{i(j),t-1}$.
+    The coefficient estimate is effect of state funding ($X_{i(j),t}$) on faculty salaries ($Y_{j,t}$), using the funding shock instrument ($Z_{i(j),t}$), while accounting for auto-correlation between different time periods --- i.e., between $X_{i(j),t}, X_{i(j),t-1}$ and $Y_{i(j),t}, Y_{i(j),t-1}$.
     These results use a $\log-\log$ specification, so the estimates are for the elasticity of faculty salaries student in a year $t+k$ with respect to state funding in year $t$, where years $k = 0, \hdots, 4$ are on the x-axis.
-    Standard errors are clustered at the university-year level.
+    Standard errors are clustered at the university-year level, and \autoref{sec:iv-model-indiv} fully describes the differences in empirical specification when unit of analysis is an individual faculty member.
 \end{figure}
@@
     These tables show the OLS estimates of regression specification \eqref{eqn:firststage}, showing the effect of the funding shock on other outcomes to gauge balance of the instrument with respect to university characteristics
-    ENrolment is measured in thousands of students.
+    Enrolment is measured in thousands of students.
     Each observation is a public univeristy-year, in the IPEDS data.
@@
     Standard errors are clustered at the state-year level, and univeristy $+$ year fixed effects are included where noted.
-    \label{tab:facultycount-shock-reg}
+    \label{tab:firststage-balance}
 \end{table}

### 2023-08-15 text/sections/10-appendix.tex
adjusted writing.

@@
     This figure shows the correlation between state funding in year $t+k$ with the funding shock in year $t$ for a university, where $k = -7, \hdots, 5$ are the years on the x-axis.
-    This shows that state funding and the funding shock have dynamic correlation, so that dynamic effects must be estimated by local projections -- and not simple OLS or 2SLS.
+    This shows that state funding and the funding shock have dynamic correlation, so that dynamic effects must be estimated by local projections --- and not simple OLS or 2SLS.
     The estimates are of \eqref{eqn:firststage}, calculated with IPEDS data, separately for each year relative to initial shock, using the $\log-\log$ specification, including fixed effects for university $+$ year, and clustering standard errors by university $+$ year.
@@
 
-\autoref{eqn:rolling-instrument} defines a rolling-share variant of the instrument, $\tilde Z_{j,t}$, where the university's state funding share exposure is based in the year a professor joins the university --- and not the base period 1990--1993.
+\autoref{eqn:rolling-instrument} defines a rolling-share variant of the instrument, $\tilde Z_{i(j),t}$, where the university's state funding share exposure is based in the year a professor joins the university --- and not the base period 1990--1993.
 $j$ indexes each professor in year $t$, $\tau(j)$ for the year the professor first joins their institution.
@@
     \label{eqn:rolling-instrument}
-    \tilde Z_{j,t} &\coloneqq - \left[
+    \tilde Z_{i(j),t} &\coloneqq - \left[
     \left( \frac{\text{Total State Funding}_{s(j),t}}{\text{Student Population}_{s(j),t}} \right)
@@
 
-Exogeneity and relevance of the rolling-share instrument, $\tilde Z_{j,t}$, follows the same reasoning as that for the base-share instrument, $Z_{i,t}$, discussed in \autoref{sec:approp-shocks}.
+Exogeneity and relevance of the rolling-share instrument, $\tilde Z_{i(j),t}$, follows the same reasoning as that for the base-share instrument, $Z_{i,t}$, discussed in \autoref{sec:approp-shocks}.
 The base-share instrument is appropriate for some outcomes with the individual Illinois professors, where appropriate.
@@
     The unit of analysis is an individual faculty member (at an Illinois public university); funding data come from IPEDS, and faculty promotion rate from IBHED.
-    The coefficient estimate is effect of state funding ($X_{i(j),t}$) on faculty promotion rate ($Y_{j,t}$), while accounting for auto-correlation between different time periods --- i.e., between $X_{i(j),t}, X_{i(j),t-1}$ and $Y_{i(j),t}, Y_{i(j),t-1}$.
-    These results use a rate$-\log$ specification, so the estimates are for the rate of promotion in a year $t+k$ affected by a 1\% change in state funding in year $t$, where years $k = 0, \hdots, 4$ are on the x-axis.
-    Standard errors are clustered at the university-year level.
+    The coefficient estimate is effect of state funding ($X_{i(j),t}$) on faculty promotion rate ($Y_{j,t}$), using the funding shock instrument ($Z_{i(j),t}$), while accounting for auto-correlation between different time periods --- i.e., between $X_{i(j),t}, X_{i(j),t-1}$ and $Y_{i(j),t}, Y_{i(j),t-1}$.
+    These results use a $\log-\log$ specification, so the estimates are for the rate of promotion in a year $t+k$ affected by a 1\% change in state funding in year $t$, where years $k = 0, \hdots, 4$ are on the x-axis.
+    Standard errors are clustered at the university-year level, and \autoref{sec:iv-model-indiv} fully describes the differences in empirical specification when unit of analysis is an individual faculty member.
 \end{figure}
@@
     The unit of analysis is an individual faculty member (at an Illinois public university); funding data come from IPEDS, and faculty promotion rate from IBHED.
-    The coefficient estimate is effect of state funding ($X_{i(j),t}$) on faculty promotion rate ($Y_{j,t}$), while accounting for auto-correlation between different time periods --- i.e., between $X_{i(j),t}, X_{i(j),t-1}$ and $Y_{i(j),t}, Y_{i(j),t-1}$.
+    The coefficient estimate is effect of state funding ($X_{i(j),t}$) on faculty promotion rate ($Y_{j,t}$), using the funding shock instrument ($Z_{i(j),t}$), while accounting for auto-correlation between different time periods --- i.e., between $X_{i(j),t}, X_{i(j),t-1}$ and $Y_{i(j),t}, Y_{i(j),t-1}$.
     These results use a rate$-\log$ specification, so the estimates are for the rate of promotion in a year $t+k$ affected by a 1\% change in state funding in year $t$, where years $k = 0, \hdots, 4$ are on the x-axis.
-    Standard errors are clustered at the university-year level.
+    Standard errors are clustered at the university-year level, and \autoref{sec:iv-model-indiv} fully describes the differences in empirical specification when unit of analysis is an individual faculty member.
 \end{figure}
@@
     This figure shows the correlation between state funding and the number of total number of faculty hired at public universities between 2011--2021, and the line is a line of best-fit.
-    Data for the number of faculty hired at public universities are provided by \cite{wapman2022quantifying}, and funding data from IPEDS.
+    Data for the number of faculty hired at public universities are provided by \cite{wapman2022quantifying,wapman2022zenodo}, and funding data from IPEDS.
 \end{figure}
@@
     This table show the second stage 2SLS estimates of regression specification \eqref{eqn:secondstage}, showing the effect of state funding changes on the number of faculty hires (per student) total for 2011--2021 at US public universities, using the funding shock to instrument for state funding.
-    Each observation is a university--year at a public university, where funding data come from IPEDS and faculty count total for 2011--2021 from \citep{wapman2022quantifying}.
+    Each observation is a university--year at a public university, where funding data come from IPEDS and faculty count total for 2011--2021 from \citep{wapman2022quantifying,wapman2022zenodo}.
     The panels show the effect of a $1$\% change in state funding per student at the university (total for 2011--2021) on the number of new faculty hires by gender (and all).

### 2023-08-17 text/sections/02-data.tex
adjusted writing
adjusted writing.

@@
 These publicly available data provides the basis to build a panel of Illinois public university professors 2010-2021; I define a professor as an individual by their first plus last name and university pairing, and link this database to IPEDS regarding finances for their employing institution.
-
 The Illinois sample represents 16,932 professors in the year 2010 and 15,352 in the year 2021, with summary statistics presented in \autoref{tab:illinois-summary}.
@@
 State agencies, and higher education institutions, employed accounting techniques to continue operating without any resources provided by the state government.\footnote{
-    Fiscal year 2016 refers to June 205 to June 2016, so is the same as the academic year definition.
+    Fiscal year 2016 refers to June 2015 to June 2016, so is the same as the academic year definition.
 }

### 2023-08-17 text/sections/04-empirical.tex
adjusted writing.

@@
 Similarly, funding shocks to higher education in a year are highly correlated with state funding per student in the 5 years before and after the initial shock in IPEDS data (see \autoref{fig:lag-firststage}).
-Thus, a linear model correlated state funding in year $t$ with outcomes in year $t$ will suffer from time-series confounding.
+Thus, a linear model correlated state funding in year $t$ with outcomes in year $t+k$ will suffer from time-series confounding.
 
@@
 Additionally, the approach accomodates use of an instrument for state funding \citep{olea2021inference}, so that I use the shift-share instument in local projections estimates as in the rest of the empirical analysis.
-I present estimates in graphical form, where the x-axis represents the year relative to the funding shock (i.e., $x = 0$ represents the year of the state funding shock) and the y-axis represents the estimated effect of the funding shock on that year's outcome.\footnote{
+I present estimates in graphical form, where the $x$-axis represents the year relative to the funding shock (i.e., $x = 0$ represents the year of the state funding shock) and the $y$-axis represents the estimated effect of the funding shock on that year's outcome.\footnote{
     See \autoref{fig:count-lp} for the graphical format, and the corresponding note for further explanation.

### 2023-08-17 text/sections/09-figures-tables.tex
adjusted writing.

@@
     The coefficient estimate is effect of state funding ($X_{i,t}$) on faculty count per student ($Y_{i,t}$), using the funding shock instrument ($Z_{i,t}$), while accounting for auto-correlation between different time periods --- i.e., between $X_{i,t}, X_{i,t-1}$ and $Y_{i,t}, Y_{i,t-1}$.
-    These results use a $\log-\log$ specification, so the estimates are for the elasticity of professor count per student in a year $t+k$ with respect to state funding in year $t$, where years $k = 0, \hdots, 10$ are on the x-axis.
+    These results use a $\log-\log$ specification, so the estimates are for the elasticity of professor count per student in a year $t+k$ with respect to state funding in year $t$, where years $k = 0, \hdots, 10$ are on the $x$-axis.
     Standard errors are clustered at the state-year level.
@@
     The coefficient estimate is effect of state funding ($X_{i(j),t}$) on faculty salaries ($Y_{j,t}$), using the funding shock instrument ($Z_{i(j),t}$), while accounting for auto-correlation between different time periods --- i.e., between $X_{i(j),t}, X_{i(j),t-1}$ and $Y_{i(j),t}, Y_{i(j),t-1}$.
-    These results use a $\log-\log$ specification, so the estimates are for the elasticity of faculty salaries student in a year $t+k$ with respect to state funding in year $t$, where years $k = 0, \hdots, 4$ are on the x-axis.
+    These results use a $\log-\log$ specification, so the estimates are for the elasticity of faculty salaries student in a year $t+k$ with respect to state funding in year $t$, where years $k = 0, \hdots, 4$ are on the $x$-axis.
     Standard errors are clustered at the university-year level, and \autoref{sec:iv-model-indiv} fully describes the differences in empirical specification when unit of analysis is an individual faculty member.
@@
     Each observation is a public univeristy-year, in the IPEDS data.
-    Panel A shows the effect of an funding shock of \$-1,000 per student in the state on the counts of the outcome --- i.e.,
-    \$-1,000 funding shock per student in the state is associated with 517 more students (as enrolment is measured in thousands) in column 1.
+    Panel A shows the effect of an funding shock of \$$-1,000$ per student in the state on the counts of the outcome --- i.e.,
+    \$$-1,000$ funding shock per student in the state is associated with 517 more students (as enrolment is measured in thousands) in column 1.
     Panel B shows the effect of a $-10$\% change funding shock per student in the state on $10$\% change in the outcomes --- i.e.,
@@
     Each observation is a public university-year, in the IPEDS data.
-    Panel A shows the effect of a fall in state funding \$-1,000 per student in the state on the number of professors --- i.e.,
-    an extra \$1,000 per student leads to 6 fewer lecturers according to column 2.
+    Panel A shows the effect of a fall in state funding \$$-1,000$ per student in the state on the number of professors --- i.e.,
+    an extra \$$1,000$ per student leads to 6 fewer lecturers according to column 2.
     Panel B shows the effect of a $10$\% change in state funding per student at the university on the 10\% change in the number of professors per students --- i.e.,
@@
     Each observation is a public university-year in the state of Illinois, where funding data come from IPEDS and faculty count come from IBHED data.
-    Panel A shows the effect of a fall in state funding \$-1,000 per student in the state on the number of professors.
+    Panel A shows the effect of a fall in state funding \$$-1,000$ per student in the state on the number of professors.
     Panel B shows the effect of a $10$\% change in state funding per student at the university on the 10\% change in the number of professors per students.
@@
     \centering
-    \caption{
-        Effects of Changes in State Funding on Faculty Salaries and Exit Rate, in Illinois 2010--2021, 2SLS Estimates.}
+    \caption{Effects of Changes in State Funding on Faculty Salaries and Exit Rate, in Illinois 2010--2021, 2SLS Estimates.}
+    \label{tab:faculty-shock-illinois-rolling}
     \makebox[\textwidth][c]{\input{tables/facultysalaries-shock-illinois-rolling.tex}}
@@
     Each observation is a faculty member--year at an Illinois public university, where funding data come from IPEDS and faculty outcomes from IBHED data.
-    The panels shows the effect of a $1$\% change in state funding per student at the university on faculties' salaries (base salary $+$ benefits), promotion rate (e.g., assistant professor to associate professor) , and rate of leaving their university (i.e., by quitting, retiring, or moving to another university).
+    The panels shows the effect of a $1$\% change in state funding per student at the university on faculties' salaries (base salary $+$ benefits), promotion rate (e.g., assistant professor to associate professor), and rate of leaving their university (i.e., by quitting, retiring, or moving to another university).
     Standard errors are clustered at the university-year of hire level, and university $+$ year of hire fixed effects are included through--out.
-    \label{tab:faculty-shock-illinois-rolling}
 \end{table}

### 2023-08-17 text/sections/10-appendix.tex
adjusted writing.

@@
 \begin{itemize}
-    \item \textit{Tidyverse} \citep{tidyverse} collected tools for data analysis in the R language.
+    \item \textit{Tidyverse} \citep{tidyverse} collected tools for data analysis in the $R$ language.
     \item \textit{LFE} \citep{lfe} implemented linear fixed effect models, with instruments, crucial for the empirical estimation in \autoref{sec:empirics}.
@@
     \textbf{Note}:
-    This figure shows the correlation between state funding in year $t+k$ with the funding shock in year $t$ for a university, where $k = -7, \hdots, 5$ are the years on the x-axis.
+    This figure shows the correlation between state funding in year $t+k$ with the funding shock in year $t$ for a university, where $k = -7, \hdots, 5$ are the years on the $x$-axis.
     This shows that state funding and the funding shock have dynamic correlation, so that dynamic effects must be estimated by local projections --- and not simple OLS or 2SLS.
@@
     The coefficient estimate is effect of funding shock ($Z_{i,t}$) on state funding ($X_{i,t}$), while accounting for auto-correlation between different time periods --- i.e., between $Z_{i,t}, Z_{i,t-1}$ and $X_{i,t}, X_{i,t-1}$.
-    These results use a $\log-\log$ specification, so the estimates are for the elasticity of state funding per student in a year $t+k$ with respect to funding shock in year $t$, where years $k = 0, \hdots, 10$ are on the x-axis.
+    These results use a $\log-\log$ specification, so the estimates are for the elasticity of state funding per student in a year $t+k$ with respect to funding shock in year $t$, where years $k = 0, \hdots, 10$ are on the $x$-axis.
     Standard errors are clustered at the state-year level.
@@
     The coefficient estimate is effect of state funding ($X_{i(j),t}$) on faculty promotion rate ($Y_{j,t}$), using the funding shock instrument ($Z_{i(j),t}$), while accounting for auto-correlation between different time periods --- i.e., between $X_{i(j),t}, X_{i(j),t-1}$ and $Y_{i(j),t}, Y_{i(j),t-1}$.
-    These results use a $\log-\log$ specification, so the estimates are for the rate of promotion in a year $t+k$ affected by a 1\% change in state funding in year $t$, where years $k = 0, \hdots, 4$ are on the x-axis.
+    These results use a $\log-\log$ specification, so the estimates are for the rate of promotion in a year $t+k$ affected by a 1\% change in state funding in year $t$, where years $k = 0, \hdots, 4$ are on the $x$-axis.
     Standard errors are clustered at the university-year level, and \autoref{sec:iv-model-indiv} fully describes the differences in empirical specification when unit of analysis is an individual faculty member.
@@
     The coefficient estimate is effect of state funding ($X_{i(j),t}$) on faculty promotion rate ($Y_{j,t}$), using the funding shock instrument ($Z_{i(j),t}$), while accounting for auto-correlation between different time periods --- i.e., between $X_{i(j),t}, X_{i(j),t-1}$ and $Y_{i(j),t}, Y_{i(j),t-1}$.
-    These results use a rate$-\log$ specification, so the estimates are for the rate of promotion in a year $t+k$ affected by a 1\% change in state funding in year $t$, where years $k = 0, \hdots, 4$ are on the x-axis.
+    These results use a rate$-\log$ specification, so the estimates are for the rate of promotion in a year $t+k$ affected by a 1\% change in state funding in year $t$, where years $k = 0, \hdots, 4$ are on the $x$-axis.
     Standard errors are clustered at the university-year level, and \autoref{sec:iv-model-indiv} fully describes the differences in empirical specification when unit of analysis is an individual faculty member.
@@
     Each observation is a public university-year in the state of Illinois, where funding data come from IPEDS and faculty count come from IBHED data.
-    Panel A shows the effect of a fall in state funding \$-1,000 per student in the state on the number of new faculty hires by position.
+    Panel A shows the effect of a fall in state funding \$$-1,000$ per student in the state on the number of new faculty hires by position.
     Panel B shows the effect of a $10$\% change in state funding per student at the university on the 10\% change in the number of faculty hires per students.
     Outcome-mean is the mean of the outcome, for Panel A the number of faculty hires, for Panel B the number of faculty hires per student.
-    Panel B uses $\log$ new faculty hires per student as the outcome, though the outcome mean is count of new faculty hires  per student (not in $\log$ terms).
+    Panel B uses new faculty hires per student as the outcome (in $\log$ terms), though the outcome mean is count of new faculty hires per student (not in $\log$ terms).
     Standard errors are clustered at the university-year level, and university $+$ year fixed effects are included through--out.

### 2023-08-17 text/sections/09-figures-tables.tex
adjusted writing.

@@
     Enrolment is measured in thousands of students.
-    Each observation is a public univeristy-year, in the IPEDS data.
-    Panel A shows the effect of an funding shock of \$$-1,000$ per student in the state on the counts of the outcome --- i.e.,
+    Each observation is a public universit--year, in the IPEDS data.
+    Panel A shows the effect of a funding shock of \$$-1,000$ (per student in the state) on the outcome --- i.e.,
     \$$-1,000$ funding shock per student in the state is associated with 517 more students (as enrolment is measured in thousands) in column 1.
     Panel B shows the effect of a $-10$\% change funding shock per student in the state on $10$\% change in the outcomes --- i.e.,
-    $-10$\% funding shock per student in the state is associated with an increase in 1.07\% more students at the unviersity according to prefferred column 1.        
-    Standard errors are clustered at the state-year level, and univeristy $+$ year fixed effects are included where noted.
+    $-10$\% funding shock per student in the state is associated with an increase in 1.07\% more students at the university according to preferred column 1.
+    Standard errors are clustered at the state-year level, and university $+$ year fixed effects are included where noted.
     \label{tab:firststage-balance}

### 2023-08-18 text/sections/00-abstract.tex
adjusted writing.

@@
 Dynamic estimates show that these effects linger multiple years after the initial funding shock.
-Analysis of all faculty at Illinois public universities shows that incumbent professors' salaries, promotion rate, and quit rate are all are not affected by funding cuts over 2011-2021, implying that the composition change arose by limiting the hiring of new tenure-track/tenured professors to replace leaving tenure-track/tenured professors.
+Analysis of all faculty at Illinois public universities shows that incumbent professors' salaries, promotion rate, and quit rate are all are not affected by funding cuts over 2011-2021, implying that the composition change arose by limiting the hiring of new tenure-track/tenured professors.
 These results show the institutional effects of stagnating state funding for higher education, and exhibit one possible mechanism for deteriorating student outcomes at US public universities over this same time period.
@@
 
-\vspace{0.1cm}
+\vspace{0.05cm}
 \noindent

### 2023-08-18 text/sections/10-appendix.tex
adjusted writing.

@@
 
-Yearly variation is not observed here, so that only the aggregate level, for 180 universities, can be considered.
-\begin{table}[H]
+\newpage
+\begin{table}[h!]
     \singlespacing
@@
     This table show the second stage 2SLS estimates of regression specification \eqref{eqn:secondstage}, showing the effect of state funding changes on the number of faculty hires (per student) total for 2011--2021 at US public universities, using the funding shock to instrument for state funding.
-    Each observation is a university--year at a public university, where funding data come from IPEDS and faculty count total for 2011--2021 from \citep{wapman2022quantifying,wapman2022zenodo}.
+    Yearly variation in total hires is not observed here, so only the total hires across 2011--2021 for 157 universities, can be considered.
+    Each observation is a public university across the years 2011--2021, where data on total funding across 2011--2021 come from IPEDS and faculty count total from \citep{wapman2022quantifying,wapman2022zenodo}.
     The panels show the effect of a $1$\% change in state funding per student at the university (total for 2011--2021) on the number of new faculty hires by gender (and all).
-    Standard errors are clustered at the university-year level, and university $+$ year fixed effects are included through--out.
+    Standard errors are clustered at the state level, and state fixed effects are included through--out.
 \end{table}

### 2023-08-21 text/sections/00-abstract.tex
adjusted writing.

@@
-US public universities universities have systematically substituted away from tenure-track/tenured professors towards contingent lecturers for the last thirty years, while their states have systematically reduced the funding to these universities relative to enrolment.
+US public universities have systematically substituted away from tenure-track/tenured professors towards contingent lecturers for the last thirty years, while their states have systematically reduced the funding to these universities relative to enrolment.
 This paper connects the changes the faculty composition at US public universities to the fall in the states' funding for their public universities, using a shift-share approach to instrument for state funding and identify the causal effects of stagnating funding for higher education.

### 2023-08-21 text/sections/01-introduction.tex
adjusted writing.

@@
-Public universities educate the majority of higher education students in the US, yet have experienced a stagnation in state funding over the last three decades .
-At the same time, public university employment of lecturers dramatically increased at public universities, while enrolment rose dramatically.
+Public universities educate the majority of higher education students in the US, yet have experienced a stagnation in state funding over the last three decades.
+At the same time, use of lecturers dramatically increased, while enrolment rose dramatically.
 Previous research has examined the effects of the decline in state funding on student outcomes \citep{NBERw23736,NBERw27885}, but there is little evidence of the effects of this funding decline on faculty.

### 2023-10-27 text/sections/00-abstract.tex
Newest revision.

@@
-US public universities have systematically substituted away from tenure-track/tenured professors towards contingent lecturers for the last thirty years, while their states have systematically reduced the funding to these universities relative to enrolment.
-This paper connects the changes the faculty composition at US public universities to the fall in the states' funding for their public universities, using a shift-share approach to instrument for state funding and identify the causal effects of stagnating funding for higher education.
-A fall in funding of \$1,000 per student leads to univeristies employing 6 more lecturers on average; relative to enrolment, a fall of $10$\% in funding per student decreases the number of assistant professors per student by $1.4$\% and full professors by $1.2$\%, yet increases the number of lecturers per student by $4.4$\%.
-Dynamic estimates show that these effects linger multiple years after the initial funding shock.
-Analysis of all faculty at Illinois public universities shows that incumbent professors' salaries, promotion rate, and quit rate are all are not affected by funding cuts over 2011-2021, implying that the composition change arose by limiting the hiring of new tenure-track/tenured professors.
+US public universities have systematically substituted away from tenure-track/tenured professors towards contingent lecturers for the last thirty years, while their states have systematically cut higher education funding despite rising enrolment.
+This paper asks whether the changes the faculty composition at US public universities are caused by the fall in state funding, using a shift-share approach to instrument for state funding.
+I find that universities employ $4.4$\% more lecturers per student in response to a fall of $10$\% per student funding, and correspondingly decrease their reliance on tenure-track faculty by employing $1.4$\% fewer assistant professors and $1.2$\% fewer full professors per student.
+%Dynamic estimates show that these effects linger multiple years after the initial funding cut.
+In a secondary analysis, using data on all faculty at Illinois public universities, I find that incumbent professors' salaries, promotion rate, and quit rate are all not affected by funding cuts, implying that faculty composition change arose by limiting the hiring of new tenure-track/tenured professors at public universities.
 These results show the institutional effects of stagnating state funding for higher education, and exhibit one possible mechanism for deteriorating student outcomes at US public universities over this same time period.

### 2023-10-27 text/sections/01-introduction.tex
Newest revision.

@@
-Public universities educate the majority of higher education students in the US, yet have experienced a stagnation in state funding over the last three decades.
-At the same time, use of lecturers dramatically increased, while enrolment rose dramatically.
-Previous research has examined the effects of the decline in state funding on student outcomes \citep{NBERw23736,NBERw27885}, but there is little evidence of the effects of this funding decline on faculty.
-I show that four-year, degree-granting public universities have experienced a systematic decline in state funding per student, and use a shift-share instrument to show that this fall led to a substitution away from tenure-track/tenured professors towards contingent lecturers.
-Analysis of all public university faculty in the state of Illinois shows that incumbent faculty were relatively unaffected, in terms of wages, promotion and quit rates, implying that changes in faculty composition arose by disrupting public universities' ability to hire or replace their tenure-track/tenured faculty.
-Private universities were not exposed to similar financial constraints during this same time period, and do not exhibit the substitution away from tenure-track/tenured professors, so that the stagnating state funding for public universities has implications for the wider structure of US higher education and research.
+Public universities have experienced a stagnation in state funding over the last three decades, despite a dramatic increase in enrolments.
+At the same time, public unviersities have increasingly subsituted away from tenure-track faculty towards lecturers, who often work on contingenet and lower-salaried contracts.
+Previous research has examined the effects of the decline in state funding on student outcomes \citep{NBERw23736,NBERw27885}, but there is little evidence on whether or how much this funding decline has impacted faculty.
+In this paper, I show there is a causal relation between the trends of declining state funding for higher education and increasing reliance on lecturers.
+In a secondary analysis of all public university faculty in the state of Illinois, I show  shows that incumbent faculty were relatively unaffected, in terms of wages, promotion and quit rates.
+This implies that changes in faculty composition arose by disrupting public universities' ability to hire or replace their tenure-track/tenured faculty. 
+Private universities were not exposed to similar financial constraints during this same time period, and do not exhibit the substitution away from tenure-track/tenured professors.
+Thus stagnating state funding for public universities has implications for the wider structure of US higher education and research.
 
-There are various margins through which funding cuts can impact a university's faculty.
+It is a priori unclear how faculty may be affected, as there are various margins for university to respond to funding cuts.
 A university's level of resources affect the ability that departments can hire new professors, retain their current faculty, or provide retirement incentives, so it is natural to expect that funding cuts may affect the composition of professors at a university.
@@
 So that it is natural to expect that systematic funding cuts will lead to changes in faculty composition, and possibly affect incumbent professors, too.
-Yet, it is a priori unclear whether/how much each of these margins adjusts to funding cuts, as these decision depend on how the university administrations view the productivity of different inputs, plus the degree of wage and employment stickiness over time.
+However, over the same time period, higher education has undergone dramatic changes in the rate of selectivity thanks to greater competition \citep{hoxby2009changing}.
+These changes may have lead public universities to resist relying on lecturers, and focus more on raising revenue from other sources, such as federal or private funds.
 
-US universities are widely considered the highest performing in the world, yet there are consequential differences between its universities that operate in the private sector and those established by state governments.
+I beging the paper by documenting the dramatic decline in state funding for public universities over recent decades, and drastic changes in faculty composition over the same time frame.
+While US universities are widely considered the highest performing in the world, there are consequential differences between its universities that operate in the private sector and those established by state governments.
 Public universities are subject to numerous state-level administrative laws, and rely on their state governments for funding: an average public university received around \$11,600 of state funding per student in 1990, and only \$8,300 in 2021.
@@
 
-I use data from the \cite{ipeds} to study the impact of changes in state funding for US public universities 1990--2017.
-I combine these data with panel data for every faculty member at Illinois public universities for 2010--2021 \citep{ibhed}, to measure how incumbent faculty (i.e., those already employed at the university) are affected by changes in state funding.
+I use two different analyses to explore the causal effects of state funding on faculty composition and salaries.
+My first analysis uses data from the \cite{ipeds} to study the impact of changes in state funding for all US public universities 1990--2017.
+My second analysis uses panel data for every faculty member at Illinois public universities for 2010--2021 \citep{ibhed}, to measure how incumbent faculty (i.e., those already employed at the university) are affected by changes in state funding.
 These data allow me to answer the following research question: if a US public university receives an extra \$1,000 for each of their students (or a 10\% increase), do they change the composition of faculty they employ?
@@
 
-I employ a shift-share instrument to identify changes in state funding for higher education, which exploits how universities differ in their financial reliance on state funding, combined with yearly shocks to the amount an entire state funds higher education --- following \cite{NBERw23736,NBERw27885}.
+In both analyses, I employ a shift-share instrument to identify changes in state funding for higher education, which exploits how universities differ in their financial reliance on state funding, combined with yearly shocks to the amount an entire state funds higher education --- following \cite{NBERw23736,NBERw27885}.
 State governments decide how to fund higher education by a complicated process which can be influenced by local economic conditions, changing state priorities, or even lobbying from the state universities themselves.
@@
 
-Falls in state funding affect faculty composition at public universities, which substitute away from tenured and tenure-track professors towards contingent lecturers.
+In my national-level analysis, I find that state funding cuts cause public universities to substitute away from tenured and tenure-track professors towards contingent lecturers.
 A funding cut of \$1,000 per student leads to an average university employing 6 more lecturers;
@@
 % ie., (35 * 0.437) / 99
-Additionally, I use individual level data on all Illinois public university professors in 2010-2021 to investigate whether the shocks to state funding affected individual professors.
-Incumbent professors were not meaningfully affected, in terms of total salary, promotion rate, or rate of leaving the Illinois public university system.
-Yet the hiring rate for new professors at public universities was negatively impacted by the falls in state funding (and by the funding shocks).
+
+In my analysis of all Illinois faculty, I find that incumbent professors were not meaningfully affected, in terms of total salary, promotion rate, or rate of leaving the Illinois public university system.
+Yet the hiring rate for new professors at public universities was negatively impacted by the state funding cuts.
 This implies that faculty composition change arose by limiting the hiring of new professors; public universities increasingly hired contingent lecturers, and increasingly did not replace their retiring (or leaving) professors.
 
-Mine is the first paper to provide causal evidence on the impact of state funding on faculty composition, and wages.
+Mine is the first paper to provide causal evidence on the impact of state funding on faculty composition and wages.
 The closest related research on higher education funding has examined how university spending affected graduation rates and levels of student debt \citep{NBERw23736,NBERw27885}, and university finances \citep{miller2022making,bound2019public,brown2014endowment}.
-For faculty outcomes, multiple theoretically model university decision-making and faculty hiring (see e.g., \citealt{abe2015implications,johnson2009jep,NBERc13879}), while others measured how faculty at universities were affected by endowment shocks and the 2008 recession \citep{brown2014endowment,turner2014impact}.
-These papers measure effects of funding and changes in university revenues, but do not measure effects on faculty, or other outcomes related to university instruction.
+For faculty outcomes, there is applied theory work that models university decision-making and faculty hiring (see e.g., \citealt{abe2015implications,johnson2009jep,NBERc13879}).
+In addition, several empirical papers measure how universities were affected by endowment shocks \citep{brown2014endowment}, and the academic job market following the 2008 recession \citep{turner2014impact}.
+These papers measure effects of state funding and changes in university revenues, but do not measure effects on faculty composition or salaries, or other outcomes related to university instruction.
 
-My results provide suggestive evidence on mechanisms for the previous research connecting state funding cuts for higher education and worsening student outcomes.
+Another contribution of my paper is to provide evidence on possible  mechanisms for the previous research connecting state funding cuts for higher education and worsening student outcomes.
 Substitution towards lecturers is likely a cost-cutting measure for funding constrained universities, though relying on these faculty --- who are often over-worked and not granted long-term employment protections --- may lead to worse student outcomes \citep{ehrenberg2005tenured,zhu2021limited,jaeger2011examining}.

### 2023-10-27 text/sections/04-empirical.tex
Newest revision.

@@
 
-The shift-share instrument has a strong first stage for universities' yearly state funding.
-\autoref{tab:firststage-reg} presents results of the first-stage regression, separately with and without a control for tuition revenue per student, plus institution and year fixed effects.
-\autoref{tab:firststage-reg} Panel A shows estimates of a funding shock of \$1 per student in the state on state funding per student at the university, and Panel B the \% increase effects of a 1\% increase in the shock per student on state funding per student.
-Column (1) shows that a funding shock of \$10 (per student in the entire state) is associated with \$11.76 less state funding per student at the university, with the corresponding -10\% funding shock leads to -9.77\%
-less funding (column 1, panel B), with similar estimates with and without including fixed effects.
-Column (2) shows estimates of the first-stage without including fixed effects, and gives less precise estimates for the funding shock, likely thanks to systematic differences in universities unaccounted for without fixed effects.
-Columns (3) and (4) include the tuition revenue control (explained in \autoref{foot:control}) to exhibit estimates with the inclusion of this possible collider (or bad control).
-Column (3) shows similar estimates to column (1) in both Panels A and B thanks to inclusion of fixed effects, so that the uncertainty in including tuition revenue as a possible bad control for the level of state funding does not matter thanks to the inclusion of fixed effects.
-Column (1) represents the estimates for \autoref{eqn:firststage} with fixed effects, omitting the tuition revenue control, and is the preferred form that I proceed with.
-
-These results, together with the case for exogeneity, show the funding shock instrument strongly predicts university revenues in the first-stage estimation.
-
 \subsection{Instrumental Variables Model}
@@
 
-Additionally, I investigate the effect of changes in state funding among incumbent professors.
+Additionally, I investigate the effect of changes in state funding among incumbent professors using the Illinois data (IBHED).
 Incumbent professors are faculty who are already employed at the university; state funding changes may affect faculty who are already at the university (beyond affecting whether they get hired), so I base the instrument in the year that the professor was hired, and include fixed effects for the hiring year.\footnote{
@@
 }
-Additionally, the approach accomodates use of an instrument for state funding \citep{olea2021inference}, so that I use the shift-share instument in local projections estimates as in the rest of the empirical analysis.
+Additionally, the approach accommodates use of an instrument for state funding \citep{olea2021inference}, so that I use the shift-share instument in local projections estimates as in the rest of the empirical analysis.
 I present estimates in graphical form, where the $x$-axis represents the year relative to the funding shock (i.e., $x = 0$ represents the year of the state funding shock) and the $y$-axis represents the estimated effect of the funding shock on that year's outcome.\footnote{

### 2023-10-27 text/sections/05-results.tex
Newest revision.

@@
 
+\subsection{First Stage Estimates}
+The shift-share instrument has a strong first stage for universities' yearly state funding.
+\autoref{tab:firststage-reg} presents results of the first-stage regression among IPEDS data, separately with and without a control for tuition revenue per student, plus institution and year fixed effects.
+I use the same empirical strategy in the analysis of Illinois faculty, so  \autoref{tab:firststage-reg} presents similar first-stage estimates among the IBHED data.
+\autoref{tab:firststage-reg} Panel A shows estimates of a funding shock of \$1 per student in the state on state funding per student at the university, and Panel B the \% increase effects of a 1\% increase in the shock per student on state funding per student.
+Column (1) shows that a funding shock of \$10 (per student in the entire state) is associated with \$11.76 less state funding per student at the university, with the corresponding -10\% funding shock leads to -9.77\%
+less funding (column 1, panel B), with similar estimates with and without including fixed effects.
+Column (2) shows estimates of the first-stage without including fixed effects, and gives less precise estimates for the funding shock, likely thanks to systematic differences in universities unaccounted for without fixed effects.
+Columns (3) and (4) include the tuition revenue control (explained in \autoref{foot:control}) to exhibit estimates with the inclusion of this possible collider (or bad control).
+Column (3) shows similar estimates to column (1) in both Panels A and B thanks to inclusion of fixed effects, so that the uncertainty in including tuition revenue as a possible bad control for the level of state funding does not matter thanks to the inclusion of fixed effects.
+Column (1) represents the estimates for \autoref{eqn:firststage} with fixed effects, omitting the tuition revenue control, and is the preferred form that I proceed with.
+
+The above shows that the first stage is strong among both IPEDS and IBHED data, and the argument in \autoref{sec:empirics} for exogenity comes from independence of the state funding mechanism to any indivdiaul public university campus.
+\autoref{tab:firststage-balance} provides a balance test, showing that the shift-share instrument is correlated with the professor outcomes, but not with other sources of university funding.
+These results, together with the case for exogeneity, show the funding shock instrument
+strongly predicts university revenues in the first-stage estimation.
+
+\subsection{National Level, IPEDS}
 Changes in state funding have clear effects on the composition of faculty public universities employ.
 \autoref{tab:facultycount-shock-reg} presents OLS and IV estimates for the effect of a change in state funding on the number of faculty at the university, individually by position (and for the total faculty), in both count of professors and in percent terms for faculty per student ratio.
-An extra \$1,000 in state funding per student leads to the average university employing 6 fewer lecturers, with no discernible effect on the count of total professors (\autoref{tab:facultycount-shock-reg}, Panel A).
-In percentage terms, a 10\% increase in state funding lead to a 4.4\% decrease in the number of lecturers per student, an increase of 14\% for both assistant and full professors, leading to a 0.6\% increase in the total number of faculty per student.
-In the state of Illinois, using the count of professors represented in the IBHED databse, state funding is correlated with the number of faculty, but lacks precision to identify the exact effects given the 144 university-year observations in the panel for 2010--2021.
+An extra \$1,000 in state funding per student leads to the average university employing 6 fewer lecturers, while effects for the other positions are not discernible from zero in absolute measure (\autoref{tab:facultycount-shock-reg}, Panel A).
+However, Panel A \autoref{tab:facultycount-shock-reg} measures professor count in percentage terms for professors per student.
+A 10\% increase in state funding lead to a 4.4\% decrease in the number of lecturers per student, an increase of 14\% for both assistant and full professors, leading to a 0.6\% increase in the total number of faculty per student.
+% In the state of Illinois, using the count of professors represented in the IBHED databse, state funding is correlated with the number of faculty, but lacks precision to identify the exact effects given the 144 university-year observations in the panel for 2010--2021.
 
-Faculty composition is not only impacted in the same years as a state funding shock, but also for a number years following.
+Faculty composition is not only impacted in the same year as a state funding shock, but also for a number years following.
 \autoref{fig:count-lp} shows local projection estimates, as a series of impulse responses, where the estimate is the elasticity for professor count per student in year $t+k$ with respect to state funding in year $t$ for follow-on years $k = 0, \hdots, 10$.
@@
 
+\subsection{Illinois Faculty, IBHED}
+This section refers to results using data on every professor at a public university in the state of Illinois.
 Incumbent professors are relatively unaffected by changes in the state funding for their university in the years following their hire.
@@
 The composition of faculty changed at public universities 1990--2017 thanks to stagnating state funding, but there are little discernible effects on incumbent faculty.
-This implied that changes in faculty compoisition arose by state funding cuts impacting the rate that public universioties hired new professors thanks.
+This implies that changes in faculty composition arose by state funding cuts impacting the rate that public universities hired new professors.
 I take recent data on the total faculty hired at US public universities over 2011--2021 (provided by \citealt{wapman2022quantifying,wapman2022zenodo}), to show that universities with lower state funding (per student) also hired fewer professors in the same time period --- shown in \autoref{fig:hiring-correlation}.
 Similarly, the funding shock IV model shows that a $10$\% cut in per student state funding leads to $13$\% fewer faculty hires per student across the decade 2011--2021.\footnote{
-    This limited sample of data are purel;y a cross-section for count of professor hires 2011--2021, so that local projection estimates are not possible here.
+    This limited sample of data are a cross-section for count of professor hires 2011--2021, so that local projection estimates are not possible here.
     See \autoref{sec:appendix-hiring} for further details.

### 2023-10-27 text/sections/09-figures-tables.tex
Newest revision.

@@
     \centering
-    \caption{First Stage Estimates, for State Funding by funding shock.}
+    \caption{First Stage Estimates, for State Funding by Funding Shock in IPEDS Data.}
     \textbf{Panel A: units in \$ per student}
@@
     \centering
-    \caption{Effects of Changes in State Funding on University Faculty Composition, in Illinois 2010--2021, OLS and 2SLS Estimates.}
-
-    \textbf{Panel A: units in \$ per student}
-
-    \makebox[\textwidth][c]{\input{tables/facultycount-illinois-reg-rawcount.tex}}
-
-    \textbf{Panel B: units in log \$ per student}
-
-    \makebox[\textwidth][c]{\input{tables/facultycount-illinois-reg-fte.tex}}
-
-    \label{tab:facultycount-illinois-reg}
-    \justify
-    \footnotesize
-    \textbf{Note}:
-    These tables show the second stage OLS and 2SLS estimates of regression specification \eqref{eqn:secondstage}, showing the effect of state funding changes on number of faculty per student in Illinois universities, using the funding shock to instrument for state funding in the columns labelled 2SLS.
-    Each observation is a public university-year in the state of Illinois, where funding data come from IPEDS and faculty count come from IBHED data.
-    Panel A shows the effect of a fall in state funding \$$-1,000$ per student in the state on the number of professors.
-    Panel B shows the effect of a $10$\% change in state funding per student at the university on the 10\% change in the number of professors per students.
-    Outcome-mean is the mean of the outcome, for Panel A the number of professors per student, for Panel B the number of faculty per student.
-    Panel B uses $\log$ faculty count per student as the outcome, though the outcome mean is count of faculty per student (not in $\log$ terms).
-    Standard errors are clustered at the university-year level, and university $+$ year fixed effects are included through--out.
-\end{table}
-
-\newpage
-\begin{table}[H]
-    \singlespacing
-    \centering
     \caption{Effects of Changes in State Funding on Faculty Salaries and Exit Rate, in Illinois 2010--2021, 2SLS Estimates.}

### 2023-10-27 text/sections/10-appendix.tex
Newest revision.

@@
 The formulation here takes as given that the professor is already employed at the university, and then projects the effect of changes in state funding on these \textit{incumbent} professors following the state funding shock.
+\autoref{tab:firststage-illinois} presents the first-stage results in Illinois data .
+
+\begin{table}[h!]
+    \singlespacing
+    \centering
+    \caption{First Stage Estimates, for State Funding by Funding Shock in IBHED Data.}
+    %\textbf{Panel A: units in \$ per student}
+    %
+    %\makebox[\textwidth][c]{\input{tables/firststage-reg-rawcount.tex}}
+
+    \textbf{Panel A: units in log \$ per student}
+
+    \makebox[\textwidth][c]{\input{tables/firststage-illinois.tex}}
+
+    \label{tab:firststage-illinois}
+    \justify
+    \footnotesize
+    \textbf{Note}:
+    These tables show the first stage OLS estimates of regression specification \eqref{eqn:secondstage1_indiv}, showing the effect of the funding shock on state funding to gauge performance as an instrument.
+    Each observation is a professor-year, in the IBHED data, and funding data are merged from IPEDS.
+    %Panel A shows the effect of an funding shock of \$-1 per student in the state on the number of \$'s of state funding per student at the university --- i.e.,
+    %\$-1 funding shock per student in the state leads to \$1.176 less state funding per student at the university according to preferred specification column 1.
+    Panel A shows the effect of a $-10$\% change funding shock per student in the state on $10$\% change in state funding per student at the university --- i.e.,
+    $-10$\% funding shock per student in the state leads to $-9.77$\% less state funding per student at the university according to prefferred specification column 1.        
+    Standard errors are clustered at the institution-year level, and institution $+$ year fixed effects are included where noted.
+\end{table}
 
@@
 \end{figure}
+
+\newpage
+\begin{table}[H]
+    \singlespacing
+    \centering
+    \caption{Effects of Changes in State Funding on University Faculty Composition, in Illinois 2010--2021, OLS and 2SLS Estimates.}
+
+    \textbf{Panel A: units in \$ per student}
+
+    \makebox[\textwidth][c]{\input{tables/facultycount-illinois-reg-rawcount.tex}}
+
+    \textbf{Panel B: units in log \$ per student}
+
+    \makebox[\textwidth][c]{\input{tables/facultycount-illinois-reg-fte.tex}}
+
+    \label{tab:facultycount-illinois-reg}
+    \justify
+    \footnotesize
+    \textbf{Note}:
+    These tables show the second stage OLS and 2SLS estimates of regression specification \eqref{eqn:secondstage}, showing the effect of state funding changes on number of faculty per student in Illinois universities, using the funding shock to instrument for state funding in the columns labelled 2SLS.
+    Each observation is a public university-year in the state of Illinois, where funding data come from IPEDS and faculty count come from IBHED data.
+    Panel A shows the effect of a fall in state funding \$$-1,000$ per student in the state on the number of professors.
+    Panel B shows the effect of a $10$\% change in state funding per student at the university on the 10\% change in the number of professors per students.
+    Outcome-mean is the mean of the outcome, for Panel A the number of professors per student, for Panel B the number of faculty per student.
+    Panel B uses $\log$ faculty count per student as the outcome, though the outcome mean is count of faculty per student (not in $\log$ terms).
+    Standard errors are clustered at the university-year level, and university $+$ year fixed effects are included through--out.
+\end{table}
 

### 2023-10-27 programs/run.sh

@@ -29,7 +29,7 @@ R CMD BATCH --no-save ibhed-shock.R
 R CMD BATCH --no-save ibhed-shock-rawcounts.R
 # Analysis and regressions at the individual-level, IBHED, rolling shock IV.
 R CMD BATCH --no-save ibhed-shock-rolling.R
-# Analysis and regressions at the institutionindividual-level, IBHED, local porjections.
+# Analysis and regressions at the institution-individual-level, IBHED, local porjections.
 R CMD BATCH --no-save ibhed-shock-longterm.R
 # Analysis and regressions for rate of faculty hiring.
 R CMD BATCH --no-save hiring-shock.R

### 2023-11-13 text/sections/00-abstract.tex
Adjustments.

@@
-US public universities have systematically substituted away from tenure-track/tenured professors towards contingent lecturers for the last thirty years, while their states have systematically cut higher education funding despite rising enrolment.
-This paper asks whether the changes the faculty composition at US public universities are caused by the fall in state funding, using a shift-share approach to instrument for state funding.
-I find that universities employ $4.4$\% more lecturers per student in response to a fall of $10$\% per student funding, and correspondingly decrease their reliance on tenure-track faculty by employing $1.4$\% fewer assistant professors and $1.2$\% fewer full professors per student.
-%Dynamic estimates show that these effects linger multiple years after the initial funding cut.
-In a secondary analysis, using data on all faculty at Illinois public universities, I find that incumbent professors' salaries, promotion rate, and quit rate are all not affected by funding cuts, implying that faculty composition change arose by limiting the hiring of new tenure-track/tenured professors at public universities.
-These results show the institutional effects of stagnating state funding for higher education, and exhibit one possible mechanism for deteriorating student outcomes at US public universities over this same time period.
+US public universities have systematically substituted away from tenure-track/tenured professors towards lecturers over the last thirty years, while states have systematically cut public funding for higher education.
+This paper shows that falling state funding caused changes in faculty composition, using a shift-share approach to instrument for state funding cuts.
+Universities employ $4.4$\% more lecturers per student following a $10$\% funding cut, and correspondingly rely less on tenure-track faculty by employing $1.4$\% fewer assistant professors and $1.2$\% fewer full professors per student.
+In a secondary analysis, using data on all faculty at Illinois public universities, incumbent professors' salaries, promotion rate, and quit rate are not affected by funding cuts, implying that composition change arose by limiting hiring of new tenure-track/tenured professors.
+These results show institutional effects of stagnating public funding for education, exhibiting one possible mechanism for deteriorating student outcomes over this same time period.
 
@@
 \noindent
-\textbf{JEL Codes:} H72, I23, J45
+\textbf{JEL Codes:} H72, I22, J45

### 2023-11-13 text/sections/01-introduction.tex
Adjustments.

@@
 At the same time, public unviersities have increasingly subsituted away from tenure-track faculty towards lecturers, who often work on contingenet and lower-salaried contracts.
-Previous research has examined the effects of the decline in state funding on student outcomes \citep{NBERw23736,NBERw27885}, but there is little evidence on whether or how much this funding decline has impacted faculty.
+Previous research has examined the effects of the decline in state funding on student outcomes \citep{NBERw23736,NBERw27885}, but there is little evidence on whether this funding decline has impacted faculty.
 In this paper, I show there is a causal relation between the trends of declining state funding for higher education and increasing reliance on lecturers.
@@
 Private universities were not exposed to similar financial constraints during this same time period, and do not exhibit the substitution away from tenure-track/tenured professors.
-Thus stagnating state funding for public universities has implications for the wider structure of US higher education and research.
+Thus stagnating state funding for public universities has implications for the wider landscape of US higher education and research.
 
-It is a priori unclear how faculty may be affected, as there are various margins for university to respond to funding cuts.
-A university's level of resources affect the ability that departments can hire new professors, retain their current faculty, or provide retirement incentives, so it is natural to expect that funding cuts may affect the composition of professors at a university.
-Similarly, the wages a university pays their faculty is affected by funding for the university, including the starting salary it offers to new professors and the amount of yearly raises (if any) for incumbent faculty.
-So that it is natural to expect that systematic funding cuts will lead to changes in faculty composition, and possibly affect incumbent professors, too.
-However, over the same time period, higher education has undergone dramatic changes in the rate of selectivity thanks to greater competition \citep{hoxby2009changing}.
-These changes may have lead public universities to resist relying on lecturers, and focus more on raising revenue from other sources, such as federal or private funds.
-
-I beging the paper by documenting the dramatic decline in state funding for public universities over recent decades, and drastic changes in faculty composition over the same time frame.
-While US universities are widely considered the highest performing in the world, there are consequential differences between its universities that operate in the private sector and those established by state governments.
-Public universities are subject to numerous state-level administrative laws, and rely on their state governments for funding: an average public university received around \$11,600 of state funding per student in 1990, and only \$8,300 in 2021.
+State funding for public universities has dramatically declined over recent decades, and the composition of faculty at these universities changed drastically over the same time frame.
 This fall is driven by a stagnation in the absolute level of state funding for higher education in most states, combined with a large increase of 46\% in student enrolment at public universities over the same time period.
 Similarly, the number of lecturers per student over doubled, while the number of tenure-track/tenured professors per students fell by $-23$\% from 1990 to 2021.
+%While US universities are widely considered the highest performing in the world, there are consequential differences between its universities that operate in the private sector and those established by state governments.
+Public universities rely on their state governments for funding, and receive over a quarter of their total revenue from state governments.
+An average public university received around \$11,600 of state funding per student in 1990, and only \$8,300 in 2021.
+
+It is a priori unclear how faculty may be affected, as there are various margins that universities may to respond to funding cuts.
+University administrations may directly cut salaries or lay-off incumbent staff; in 2015, the University of Illinois furloughed faculty around a budget crisis \citep{furlough2010}.
+Similarly, faculty hiring committees are often cancelled or disrupted in response to funding cuts, meaning there are fewer faculty entering a department to replace those leaving.
+On the other hand, some faculty positions may grow at universities in response to funding cuts.
+Public universities do not restrict enrolment in response to funding cuts, so someone must teach the next year of students.
+Lecturers can fill in teaching duties, at a lower cost for the universities than tenure-track professors, with less employment protections (e.g., no tenure).
+However, over the same time period, higher education has undergone dramatic changes in the rate of selectivity thanks to greater competition \citep{hoxby2009changing}.
+This selectivity pressure may temper universities from relying on lecturers, and instead lead them to fill in the funding gap by raising revenue from other sources, such as private funds or raising tuition.
 
@@
     To the best of my knowledge, this is also the first paper to use the local projections method for a shift-share instrument, and to acknowledge the time-series confounding for shift-share IV models when the outcome is non-contemporaneous.
-    %Previous work ignored the potential time-series confoduning, and estimated linear instrumental variables.
 }
@@
 The closest related research on higher education funding has examined how university spending affected graduation rates and levels of student debt \citep{NBERw23736,NBERw27885}, and university finances \citep{miller2022making,bound2019public,brown2014endowment}.
-For faculty outcomes, there is applied theory work that models university decision-making and faculty hiring (see e.g., \citealt{abe2015implications,johnson2009jep,NBERc13879}).
-In addition, several empirical papers measure how universities were affected by endowment shocks \citep{brown2014endowment}, and the academic job market following the 2008 recession \citep{turner2014impact}.
-These papers measure effects of state funding and changes in university revenues, but do not measure effects on faculty composition or salaries, or other outcomes related to university instruction.
+For faculty outcomes, previous applied theory work models university decision-making and faculty hiring (see e.g., \citealt{abe2015implications,johnson2009jep,NBERc13879}).
+In addition, several empirical papers measure how universities were effected by endowment shocks \citep{brown2014endowment}, and trends in the academic job market following the 2008 recession \citep{turner2014impact}.
+These papers measure changes in state funding and university revenues, but do not measure effects on faculty composition or salaries, or other outcomes related to university instruction.
 

### 2023-11-13 text/sections/02-data.tex
Newest revision.
Adjustments.

@@
 \subsection{Data Description}
-The data used in this analysis come from two primary sources: \citet[IPEDS]{ipeds} regarding institutional information including finances and enrolment, and \citet[IBHED]{ibhed} for data on every professor in the Illinois public university system.
+The data used in this paper come from two primary sources: \citet[IPEDS]{ipeds} regarding institutional information including finances and enrolment, and \citet[IBHED]{ibhed} for data on every professor in the Illinois public university system.
 
@@
     The years 1987-1989 are represented in these data in an incompletion fashion, so I focus on the years 1990 onwards.
-    Year refers to the calendar year of the spring term --- i.e., 1987 refers to the academic year that ran August 1986 to July 1987.
+    Year refers to the calendar year of the spring term --- i.e., 1990 refers to the academic year that ran August 1989 to July 1990.
 }

### 2023-11-13 text/sections/06-discussion.tex
Adjustments.

@@
 These effects are persistent: public higher education in the US was not exposed to one large instance of funding cuts, but systematic funding short-falls across the last three decades.
-The long-term estimates (i.e., \autoref{fig:count-lp}) show that state funding cuts have effects on universities for multiple years after the initial funding cut, and these effects are compounded by the fact that public universities received persistent funding cuts for so many years.
+The long-term estimates (\autoref{fig:count-lp}) show that state funding cuts have effects on universities for multiple years after the initial funding cut, and these effects are compounded by the fact that public universities received persistent funding cuts for so many years.
 The end result is large changes in faculty composition, the long-run trend in stagnating state funding leading to a long-run substitution towards lecturers at public universities.

### 2023-11-13 text/sections/09-figures-tables.tex
Adjustments.

@@
     \textbf{Note}:
-    These figures show the local projections estimates of regression specification \eqref{eqn:secondstage}, with the funding shock as an instrument for state funding
+    These figures show the local projections estimates of regression specification \eqref{eqn:secondstage}, with the funding shock as an instrument for state funding.
     The unit of analysis is the university, and uses IPEDS data.
@@
 \newpage
-\begin{table}[h!]
+\begin{table}[H]
     \singlespacing
@@
     Standard errors are clustered at the state-year level, and univeristy $+$ year fixed effects are included where noted.
-\end{table}
-
-\newpage
-\begin{table}[H]
-    \singlespacing
-    \centering
-    \caption{Shift-Share Instrument Balance Test, in IPEDS 1990--2017.}
-
-    \textbf{Panel A: units in \$}
-
-    \makebox[\textwidth][c]{\input{tables/firststage-balance-count.tex}}
-
-    \textbf{Panel B: units in log \$ per student}
-
-    \makebox[\textwidth][c]{\input{tables/firststage-balance-log.tex}}
-
-    \justify
-    \footnotesize
-    \textbf{Note}:
-    These tables show the OLS estimates of regression specification \eqref{eqn:firststage}, showing the effect of the funding shock on other outcomes to gauge balance of the instrument with respect to university characteristics
-    Enrolment is measured in thousands of students.
-    Each observation is a public universit--year, in the IPEDS data.
-    Panel A shows the effect of a funding shock of \$$-1,000$ (per student in the state) on the outcome --- i.e.,
-    \$$-1,000$ funding shock per student in the state is associated with 517 more students (as enrolment is measured in thousands) in column 1.
-    Panel B shows the effect of a $-10$\% change funding shock per student in the state on $10$\% change in the outcomes --- i.e.,
-    $-10$\% funding shock per student in the state is associated with an increase in 1.07\% more students at the university according to preferred column 1.
-    Standard errors are clustered at the state-year level, and university $+$ year fixed effects are included where noted.
-    \label{tab:firststage-balance}
 \end{table}

### 2023-11-13 text/sections/10-appendix.tex
Adjustments.

@@
 \subsection{IPEDS First Stage}
+\begin{table}[H]
+    \singlespacing
+    \centering
+    \caption{Shift-Share Instrument Balance Test, in IPEDS 1990--2017.}
+
+    \textbf{Panel A: units in \$}
+
+    \makebox[\textwidth][c]{\input{tables/firststage-balance-count.tex}}
+
+    \textbf{Panel B: units in log \$ per student}
+
+    \makebox[\textwidth][c]{\input{tables/firststage-balance-log.tex}}
+
+    \justify
+    \footnotesize
+    \textbf{Note}:
+    These tables show the OLS estimates of regression specification \autoref{eqn:firststage}, showing the effect of the funding shock on other outcomes to gauge balance of the instrument with respect to university characteristics
+    Enrolment is measured in thousands of students.
+    Each observation is a public universit--year, in the IPEDS data.
+    Panel A shows the effect of a funding shock of \$$-1,000$ (per student in the state) on the outcome --- i.e.,
+    \$$-1,000$ funding shock per student in the state is associated with 517 more students (as enrolment is measured in thousands) in column 1.
+    Panel B shows the effect of a $-10$\% change funding shock per student in the state on $10$\% change in the outcomes --- i.e.,
+    $-10$\% funding shock per student in the state is associated with an increase in 1.07\% more students at the university according to preferred column 1.
+    Standard errors are clustered at the state-year level, and university $+$ year fixed effects are included where noted.
+    \label{tab:firststage-balance}
+\end{table}
 
@@
 \newpage
-\subsection{Additional Results, Faculty Hiring}
+\subsection{Faculty Hiring}
 \label{sec:appendix-hiring}
@@
 \newpage
-\subsection{Additional Results, Rates of Substitution}
+\subsection{Rates of Substitution}
 \label{sec:appendix-substitution}

### 2023-11-13 programs/analysis/ibhed-shock-longterm.R

@@ -229,7 +229,9 @@ lpreg.plot <- function(model.lpreg) {
         geom_line(aes(y = conf.high), linetype = "dashed") +
         scale_x_continuous(name = "Years, Relative to Initital Shock",
             breaks = model.lpdata$t, expand = c(0.01, 0.01)) +
-        scale_y_continuous(name = "") +
+        scale_y_continuous(name = "",
+            limits = c(-0.2, 0.2),
+            breaks = seq(-1, , by = 0.05)) +
         theme_bw() +
         ggtitle("Estimate") +
         theme(plot.title = element_text(size = rel(1)),
@@ -275,6 +277,9 @@ lecturer_salaries.lpreg <-
         # Option to use IV for predictor endogeneity
         iv_reg = TRUE,
         instrum = "appropriationshock_perEnroll_rolling",
+        # LP options
+        diff_shock = FALSE,
+        cumul_mult = FALSE,
         # Add clustered SEs in the panel
         panel_model = "pooling",
         robust_cov = "vcovHC",
@@ -283,10 +288,7 @@ lecturer_salaries.lpreg <-
         hor = time.horizon)
 # Save this plot
 ggsave("../../text/figures/salaries-lecturer-illinois-lp-rolling.png",
-    plot = (lpreg.plot(lecturer_salaries.lpreg) +
-        scale_y_continuous(name = "",
-            limits = c(-0.05, 0.1),
-            breaks = seq(-2, 1, by = 0.025))),
+    plot = lpreg.plot(lecturer_salaries.lpreg),
     units = "cm", width = fig.width, height = fig.height)
 
 # LP estimation for salaries of APs.
@@ -301,6 +303,9 @@ assistant_salaries.lpreg <-
         # Option to use IV for predictor endogeneity
         iv_reg = TRUE,
         instrum = "appropriationshock_perEnroll_rolling",
+        # LP options
+        diff_shock = FALSE,
+        cumul_mult = FALSE,
         # Add clustered SEs in the panel
         panel_model = "pooling",
         robust_cov = "vcovHC",
@@ -309,10 +314,7 @@ assistant_salaries.lpreg <-
         hor = time.horizon)
 # Save this plot
 ggsave("../../text/figures/salaries-assistant-illinois-lp-rolling.png",
-    plot = (lpreg.plot(assistant_salaries.lpreg) +
-        scale_y_continuous(name = "",
-            limits = c(-0.05, 0.1),
[…]

### 2023-11-13 programs/analysis/ipeds-shock-longterm.R

@@ -98,8 +98,7 @@ reg.data <- reg.data %>%
         stateappropriations_real =
             log(stateappropriations_real / enrollment_reported),
         appropriationshock_perEnroll_real =
-            - log(appropriationshock_perEnroll_real),
-        tuitionrev_real = log(tuitionrev_real / enrollment_reported)) %>%
+            - log(appropriationshock_perEnroll_real)) %>%
     pdata.frame(index = c("unitid", "year"), drop.index = FALSE)
 
 # Save memory by removing initally loaded files
@@ -119,9 +118,10 @@ gc()
 library(plm)
 library(lpirfs)
 time.horizon <- 11
+limits.vector <- c(-0.1, 0.6)
 
 ## Define a function to nicely plot the LP results.
-lpreg.plot <- function(model.lpreg) {
+lpreg.plot <- function(model.lpreg, limits.vector) {
     ## Input a LP-reg (estimates an elesticity \in [-1,1])
     ## And the time.hirozon for how far to plot LP model
     # Get the estimates
@@ -143,8 +143,8 @@ lpreg.plot <- function(model.lpreg) {
         scale_x_continuous(name = "Years, Relative to Initital Shock",
             breaks = model.lpdata$t, expand = c(0.01, 0.01)) +
         scale_y_continuous(name = "",
-            limits = c(-0.2, 0.75),
-            breaks = seq(-2, 1, by = 0.1)) +
+            limits = limits.vector,
+            breaks = seq(-5, 5, by = 0.1)) +
         theme_bw() +
         ggtitle("Estimate") +
         theme(plot.title = element_text(size = rel(1)),
@@ -163,6 +163,9 @@ firststage.lpreg <-
         shock = "appropriationshock_perEnroll_real",
         # Option to use IV for predictor endogeneity (not used here)
         iv_reg = FALSE,
+        # LP options
+        diff_shock = FALSE,
+        cumul_mult = FALSE,
         # Add fixed effects + clsutered SEs in the panel
         panel_model = "within",
         panel_effect = "twoways",
@@ -170,11 +173,10 @@ firststage.lpreg <-
         robust_cluster = c("group", "time"),
         confint = 1.96,
         hor = time.horizon)
+plot(firststage.lpreg)
 # Save this plot
 ggsave("../../text/figures/firststage-lp.png",
-    plot = (lpreg.plot(firststage.lpreg) +
-        scale_y_continuous(name = "",
[…]

### 2023-12-19 text/paper.tex
adjusted writing
Adjustments.
Wiritng updates, new title.

@@
 \hypersetup{colorlinks=true, linkcolor=blue, citecolor=blue, filecolor=magenta, urlcolor=blue}
+\def\sectionautorefname~#1\null{Section~#1\null}                       % Fix autoref for sections
 \def\equationautorefname~#1\null{Equation~(#1)\null}                       % Fix autoref for equations
@@
         and Evan Riehl,
-        as well as seminar participants at Cornell University (Autumn 2022) for helpful discussion.} \\
+        as well as seminar participants at Cornell University for helpful discussion.} \\
     \vspace{0.1cm}
@@
 % Title
-\title{Stagnating State Funding for Higher Education and its Effect on Faculty at US Universities
-    }
+\title{Less Funding, More Lecturers, and Fewer Professors: \\ \vspace{0.1cm}
+    \Large{Stagnating State Funding for Higher Education and its Effect on Faculty at Public Universities}}
 \date{This version: \today\thanks{
@@
 \rhead{\today}
-\lhead{Stagnating State Funding for Higher Education and its Effect on Faculty.}
+\lhead{Less Funding, More Lecturers, and Fewer Professors.}
 % Begin

### 2023-12-19 text/sections/00-abstract.tex
Wiritng updates, new title.

@@
-US public universities have systematically substituted away from tenure-track/tenured professors towards lecturers over the last thirty years, while states have systematically cut public funding for higher education.
-This paper shows that falling state funding caused changes in faculty composition, using a shift-share approach to instrument for state funding cuts.
-Universities employ $4.4$\% more lecturers per student following a $10$\% funding cut, and correspondingly rely less on tenure-track faculty by employing $1.4$\% fewer assistant professors and $1.2$\% fewer full professors per student.
-In a secondary analysis, using data on all faculty at Illinois public universities, incumbent professors' salaries, promotion rate, and quit rate are not affected by funding cuts, implying that composition change arose by limiting hiring of new tenure-track/tenured professors.
-These results show institutional effects of stagnating public funding for education, exhibiting one possible mechanism for deteriorating student outcomes over this same time period.
+Public universities employ more lecturers than ever, while state funding for higher education has stagnated for over thirty years.
+This paper shows falling state funding caused a substitution away from professors towards lecturers at US public universities, using a shift-share approach to instrument for state funding cuts relative to enrolment.
+Universities employ $4.4$\% more lecturers per student following a $10$\% funding cut, relying less on professors by employing $1.4$\% fewer assistant professors and $1.2$\% fewer full professors per student.
+Incumbent professors' salaries, promotion rate, and quit rate at Illinois universities are not affected by funding cuts, so that the substitution arose by limiting hiring of new tenure-track/tenured professors.
+Stagnating public funding affects public universities as institutions, and likely contributed to deteriorating student outcomes at public universities since the 1990s.
 

### 2023-12-19 text/sections/01-introduction.tex
Wiritng updates, new title.

@@
-Public universities have experienced a stagnation in state funding over the last three decades, despite a dramatic increase in enrolments.
-At the same time, public unviersities have increasingly subsituted away from tenure-track faculty towards lecturers, who often work on contingenet and lower-salaried contracts.
-Previous research has examined the effects of the decline in state funding on student outcomes \citep{NBERw23736,NBERw27885}, but there is little evidence on whether this funding decline has impacted faculty.
-In this paper, I show there is a causal relation between the trends of declining state funding for higher education and increasing reliance on lecturers.
-In a secondary analysis of all public university faculty in the state of Illinois, I show  shows that incumbent faculty were relatively unaffected, in terms of wages, promotion and quit rates.
-This implies that changes in faculty composition arose by disrupting public universities' ability to hire or replace their tenure-track/tenured faculty. 
-Private universities were not exposed to similar financial constraints during this same time period, and do not exhibit the substitution away from tenure-track/tenured professors.
-Thus stagnating state funding for public universities has implications for the wider landscape of US higher education and research.
+States fund their public universities with roughly the same amount today as they did in the 1990s, despite their enrolment rising by over 50\% on average.
+At the same time, public universities have systematically substituted away from professors towards lecturers, who teach often work on contingent and lower-salaried contracts.
+Previous research shows declining in state funding led to worsening student outcomes \citep{NBERw23736,NBERw27885}, but there is little evidence on whether this funding decline has impacted faculty.
+In this paper, I show that falling state funding caused a substitution away from professors toward lecturers at US public universities.
 
-State funding for public universities has dramatically declined over recent decades, and the composition of faculty at these universities changed drastically over the same time frame.
-This fall is driven by a stagnation in the absolute level of state funding for higher education in most states, combined with a large increase of 46\% in student enrolment at public universities over the same time period.
-Similarly, the number of lecturers per student over doubled, while the number of tenure-track/tenured professors per students fell by $-23$\% from 1990 to 2021.
-%While US universities are widely considered the highest performing in the world, there are consequential differences between its universities that operate in the private sector and those established by state governments.
-Public universities rely on their state governments for funding, and receive over a quarter of their total revenue from state governments.
-An average public university received around \$11,600 of state funding per student in 1990, and only \$8,300 in 2021.
+Per student state funding for higher education has stagnated the last thirty years.
+The average public university received around \$11,600 of state funding per student in 1990 and \$8,300 in 2021, while the number of lecturers per student increased by 109\% and professors per student decreased by $23$\% over 1990--2021.
+I use a shift-share instrumental variables approach to estimate how faculty are affected by these funding cuts.
+Universities employ $4.4$\% more lecturers per student following a $10$\% funding cut, and correspondingly rely less on tenure-track faculty by employing $1.4$\% fewer assistant professors and $1.2$\% fewer full professors per student.
+In a secondary analysis of all public university faculty in the state of Illinois, I show that incumbent faculty were relatively unaffected, in terms of wages, promotion and quit rates.
+This implies that the faculty substitution arose by disrupting hiring of new tenure-track/tenured professors.
+% Private universities were not exposed to similar financial constraints during this same time period, and do not exhibit the substitution away from tenure-track/tenured professors.
+This paper shows the institutional effects of stagnating state funding for higher education, and offers one possible mechanism for how funding cuts affected student outcomes.
+
+I use two different analyses to explore the causal effects of state funding on faculty composition and salaries.
+My first analysis uses data from the \cite{ipeds} to study the impact of changes in state funding for all US public universities 1990--2021.
+My second analysis uses panel data for every faculty member at Illinois public universities for 2010--2021 \citep{ibhed}, to measure how incumbent faculty (i.e., those already employed at the university) are affected by changes in state funding.
+These data allow me to answer the following research question: if a US public university receives an extra \$1,000 per student (or a 10\% increase), do they change how many faculty they employ?
+If so, which positions do universities substitute away from, or towards, and do these funding cuts affect the faculty themselves?
 
 It is a priori unclear how faculty may be affected, as there are various margins that universities may to respond to funding cuts.
-University administrations may directly cut salaries or lay-off incumbent staff; in 2015, the University of Illinois furloughed faculty around a budget crisis \citep{furlough2010}.
+For example, university administrations may directly cut salaries or lay-off incumbent staff.
+This happened in 2015 when the state of Illinois shut-down without setting a state budget, and the University of Illinois furloughed faculty thanks to the funding shortfall \citep{furlough2010}.
 Similarly, faculty hiring committees are often cancelled or disrupted in response to funding cuts, meaning there are fewer faculty entering a department to replace those leaving.
@@
 Lecturers can fill in teaching duties, at a lower cost for the universities than tenure-track professors, with less employment protections (e.g., no tenure).
-However, over the same time period, higher education has undergone dramatic changes in the rate of selectivity thanks to greater competition \citep{hoxby2009changing}.
-This selectivity pressure may temper universities from relying on lecturers, and instead lead them to fill in the funding gap by raising revenue from other sources, such as private funds or raising tuition.
+But it is not certain that universities would hire more lecturers in the face of budget cuts; they may not hire any more instructional staff thanks to lack of resources.
+Similarly, the universities may be limited from hiring more lecturers by competitive pressures from peer institutions \citep{hoxby2009changing,urquiola2020markets}, and instead focus on filling the funding gap by raising revenue from other sources, such as private funds or raising tuition.
 
-I use two different analyses to explore the causal effects of state funding on faculty composition and salaries.
-My first analysis uses data from the \cite{ipeds} to study the impact of changes in state funding for all US public universities 1990--2017.
-My second analysis uses panel data for every faculty member at Illinois public universities for 2010--2021 \citep{ibhed}, to measure how incumbent faculty (i.e., those already employed at the university) are affected by changes in state funding.
-These data allow me to answer the following research question: if a US public university receives an extra \$1,000 for each of their students (or a 10\% increase), do they change the composition of faculty they employ?
-If so, which positions do universities substitute away from, or towards, and do these funding cuts affect the faculty members themselves?
-
-In both analyses, I employ a shift-share instrument to identify changes in state funding for higher education, which exploits how universities differ in their financial reliance on state funding, combined with yearly shocks to the amount an entire state funds higher education --- following \cite{NBERw23736,NBERw27885}.
-State governments decide how to fund higher education by a complicated process which can be influenced by local economic conditions, changing state priorities, or even lobbying from the state universities themselves.
-As such, state funding is not randomly assigned to universities and so it is necessary to employ an identification strategy, such as the shift-share instrument, to account for this.
-The shift-share instrument interacts how much a university relies on state funding as a percent of revenues in a base period with the entire state's total funding per higher education student for each year, exploiting changes in funding and how much each university historically relied on state funding for higher education.
-Additionally, I estimate effect of state funding on faculty multiple years after the initial funding cuts, using the local projections method thanks to the presence of time-series confounding between the funding shock and later years' level of state funding.\footnote{
-    To the best of my knowledge, this is also the first paper to use the local projections method for a shift-share instrument, and to acknowledge the time-series confounding for shift-share IV models when the outcome is non-contemporaneous.
+In both analyses, I employ a shift-share instrument to identify changes in state funding for higher education.
+State governments decide how to fund higher education by a complicated process which can be influenced by local economic conditions, changing state priorities, or informal lobbying from universities.
+As such, state funding is not randomly assigned to universities and so it is necessary to employ an identification strategy to account for this.
+The shift-share instrument identifies exogenous changes in state funding by exploiting differences in how much each university relied on state funding, interacted with state-wide changes in higher education funding --- following \cite{NBERw23736,NBERw27885}.
+The approach relies on the insight that yearly state government budgets are decided without targetting individual universities for funding cuts, and exploits the fact that more reliant universities will be more affected by state-wide changes in higher education funding.
+Additionally, I estimate effect of state funding on faculty multiple years after the initial funding cuts, using the local projections method thanks to time-series confounding between the funding shock and later years' level of state funding.\footnote{
+    To the best of my knowledge, this is also the first paper to use the local projections method for a shift-share instrument, and to acknowledge the time-series confounding for shift-share IV models when the outcome is in future time periods.
 }
 
-In my national-level analysis, I find that state funding cuts cause public universities to substitute away from tenured and tenure-track professors towards contingent lecturers.
+In my national-level analysis (IPEDS), I find that state funding cuts cause public universities to substitute away from tenured and tenure-track professors towards lecturers.
 A funding cut of \$1,000 per student leads to an average university employing 6 more lecturers;
 in percentage terms, a funding cut of 10\% per student leads to a fall of 1.4\% in the number of assistant professors per student at a university, a fall of 1.2\% for tenured professors, and an increase in 4.4\% the number of lecturers per student.
-Local projection estimates show that these effects linger for three to four years after the initial funding shock, showing that the effect is not isolated to the year of the funding shock. 
-Over the same time period, state funding per student fell by around 35\%, count of (tenure-tack/tenured) professors per student fell by 9\%, and lecturers per student increased by 99\%, so that these results show that falls in state funding explain around 53\% of the fall in professors per student
+Local projection estimates show that these effects linger for many years after the initial funding shock, showing that the effect is not isolated to the year of the funding cut. 
+Over the same time period, state funding per student fell by around 35\%, professors per student fell by 9\%, and lecturers per student increased by 99\%.
+These results show that falls in state funding explain around 53\% of the fall in professors per student,
 % i.e. -(35 * 0.137) / -9
@@
 % ie., (35 * 0.437) / 99
-
-In my analysis of all Illinois faculty, I find that incumbent professors were not meaningfully affected, in terms of total salary, promotion rate, or rate of leaving the Illinois public university system.
+In my analysis of all Illinois faculty (IBHED), I find that incumbent professors were not meaningfully affected, in terms of total salary, promotion rate, or rate of leaving the Illinois public university system.
 Yet the hiring rate for new professors at public universities was negatively impacted by the state funding cuts.
-This implies that faculty composition change arose by limiting the hiring of new professors; public universities increasingly hired contingent lecturers, and increasingly did not replace their retiring (or leaving) professors.
+This implies that faculty composition change arose by limiting the hiring of new professors; public universities increasingly hired contingent lecturers, and increasingly did not replace their retiring (or otherwise leaving) professors.
+Suggestive evidence using data on cumulative hiring counts supports this hypothesis.
 
 Mine is the first paper to provide causal evidence on the impact of state funding on faculty composition and wages.
-The closest related research on higher education funding has examined how university spending affected graduation rates and levels of student debt \citep{NBERw23736,NBERw27885}, and university finances \citep{miller2022making,bound2019public,brown2014endowment}.
-For faculty outcomes, previous applied theory work models university decision-making and faculty hiring (see e.g., \citealt{abe2015implications,johnson2009jep,NBERc13879}).
-In addition, several empirical papers measure how universities were effected by endowment shocks \citep{brown2014endowment}, and trends in the academic job market following the 2008 recession \citep{turner2014impact}.
+The closest related research on higher education funding examined how university spending affected graduation rates and levels of student debt \citep{NBERw23736,NBERw27885}, and university finances \citep{miller2022making,bound2019public,brown2014endowment}.
+For faculty outcomes, applied theory work models university decision-making and faculty hiring (see e.g., \citealt{abe2015implications,johnson2009jep,NBERc13879}).
+In addition, empirical papers measure how universities responded to endowment shocks \citep{brown2014endowment}, and trends in the academic job market following the 2008 recession \citep{turner2014impact}.
 These papers measure changes in state funding and university revenues, but do not measure effects on faculty composition or salaries, or other outcomes related to university instruction.
 
-Another contribution of my paper is to provide evidence on possible mechanisms for the previous research connecting state funding cuts for higher education and worsening student outcomes.
+Another contribution of my paper is to provide evidence on possible mechanisms connecting state funding cuts for higher education and worsening student outcomes.
 Substitution towards lecturers is likely a cost-cutting measure for funding constrained universities, though relying on these faculty --- who are often over-worked and not granted long-term employment protections --- may lead to worse student outcomes \citep{ehrenberg2005tenured,zhu2021limited,jaeger2011examining}.
-Lastly, my results provide evidence that the long term trends in higher education funding are causally related, contributing to the literature in the long run trends in faculty outcomes \citep{ehrenberg2003studying} and trends in US higher education and funding \citep{hoxby2009changing}.
+Lastly, my results provide evidence that the long term trends in higher education funding are causally related, contributing to the literature on faculty outcomes \citep{ehrenberg2003studying} and trends in US higher education and funding \citep{hoxby2009changing,ehrenberg2012american}.
 
@@
 \autoref{sec:conceptual} gives the conceptual framework for how state funding may affect faculty.
-\autoref{sec:empirics} draws the empirical framework for isolating the causal effects of state funding on faculty composition and individual faculty, and \autoref{sec:results} presents the empirical results.
+\autoref{sec:empirics} draws the empirical framework for isolating the causal effects of state funding on faculty composition and individual faculty, with \autoref{sec:results} presenting the empirical results.
 \autoref{sec:discussion} discusses the context and implications for the findings.

### 2023-12-19 text/sections/02-data.tex
Wiritng updates, new title.

@@
     Unfortunately, this measure of professors' salaries is particularly crude in measuring on professors' salaries and other individual-outcomes; summary statistics on IPEDS data do not agree with trends in average professor salary over the sample time period compared to summary statistics provided in \citealt{aau2021survey}.
-    Similarly, IPEDS reports of the average salary paid to professors at private institutions also disagree with other sources (including those calculated for in Illinois in the IBHED data later in this section) so are not considered in this analysis.
+    Similarly, IPEDS reports of the average salary paid to professors at private institutions also disagree with other sources (including those calculated for Illinois in IBHED data later in this section) so are not considered in this analysis.
 }
@@
 
-The institutional environment for how states appropriate funding allows for wide variation in state funding for higher education, both between states and across years in each state.
-A state government plans an annual budget a couple of years ahead of the fiscal year, and the legislature chooses to approve or reject a budget request put forth by the governor's office.\footnote{
+States vary largely in how much they fund their public university systems, thanks in part to the state-wide budgeting process.
+Planning for an annual budget begins two years ahead of the fiscal year, and the legislature votes to approve or reject the governor office's budget request a number of months before the fiscal year begins.\footnote{
     \cite{NBERw23736} present a full discussion of the decision-making process for state appropriations, drawing on administrative records originally analysed by \cite{parmley2009state}.
@@
 US state governments, by majority, are legally obligated to run a balanced-budget, so that yearly variation in tax revenues (caused by changing economic conditions or otherwise) necessarily affect state expenditures.
-This process leads to yearly variation in state funding not seen in other revenues sources (such as federal government appropriations), since US states differ in their support for higher education, and are subject to fiscal constraints.
-State funding for higher education are a particularly attractive area of state spending to absorb negative shocks to state finances \citep{delaney2011state}.\footnote{
+Public universities have relatively lower lobbying power than other funding beneificaries, so often bear the brunt of state government funding cuts \citep{delaney2011state}.\footnote{
     \cite{delaney2011state} fully describe the financial environment of state expenditures, and what makes spending on higher education an attractive area for state governments to expand funding during years of higher tax revenues, and retract funding in leaner years.
@@
 }
-Additionally, there is yearly variation in the number of higher education students in each state \citep{turner2014impact}, so that per-student state funding varies on multiple dimensions.
+These features lead to yearly variation in state funding not seen in other revenues sources (such as federal funding), since US states differ in their support for higher education, and are subject to fiscal constraints.
+Additionally, the number of higher education students in each state varies thanks to the size of each birth cohort.
+For example, the birth cohort of 1971 was larger than for neighbouring years, leading to more student demand for limited public university places for students turning 18 in 1989 \citep{bound2007cohort}.
+This leads to variation in the amount of funding per student across years, showing that per student state funding varies on multiple dimensions.
 
-Public university revenues on average stagnated across there years 1990--2021.
+Public university revenues on average stagnated over 1990--2021.
 \autoref{fig:funding} show the trends in revenues for the mean public university for the years 1990--2021.
@@
 There are large differences in the average number of professors per student between the private and public sector.
-\autoref{fig:fte-perprof} shows that private universities start with  a higher baseline of around 4.5 professors per 100 students, and exhibited yearly variation of less than 0.5 professor per hundred students over the thirty years.
+\autoref{fig:fte-perprof} shows that private universities start with a higher baseline of around 4.5 professors per hundred students, and exhibited yearly variation of less than 0.5 professor per hundred students over the thirty years.
 Public universities start with 3.9 professors per hundred students, and this number falls to around 3., with the largest fall in the 2008-2011 time period.
@@
 
-2010--2021 saw not only a decline, but also large yearly funding variation, particularly around 2016.
+There was not only a stagnation in state funding in this time period, but also large annual rises and falls, particularly around 2016.
 In the calendar year 2015, partisan disagreements between the democratic legislature and republican governor led to the 2016 fiscal year starting with no state budget.
@@
 }
-While public universities were able to stay open, there were drastic revenue and spending cuts in response to the budget impasse, as it continued through fiscal year 2017, and ended with a new budget restoring funding to state institutions for 2018.
-This variation in public university revenues via the state funding channel, stemming entirely from political disagreements --- not from state decisions regarding higher education and its finances \citep{young2020squandered} --- mean that Illinois public universities exhibit sizable changes in their state funding over 2010-2021, of similar order to those for the rest of the country over 1990--2021.
+While most public universities were able to stay open, there were drastic revenue and spending cuts in response to the budget impasse, as it continued through fiscal year 2017, and ended with a new budget restoring funding to state institutions for 2018.
+This means that Illinois public universities exhibit sizable changes in their state funding over 2010--2021, of similar order to those for the rest of the country over 1990--2021.
+Additionally, the 2016 episode stemmed entirely from political disagreements, and not from state decisions regarding higher education and its finances \citep{young2020squandered}, exhibiting how state-level changes in funding affect public universities thanks to unrelated issues.

### 2023-12-19 text/sections/03-concepts.tex
Wiritng updates, new title.

@@
 
-While state support for public higher education has stagnated at the same time as education costs rose, there are multiple possible ways that a university can respond.
-\cite{NBERw23736} established that corresponding rises in tuition did not offset the falling state support, so that university spending (per student) fell in response to these persistent, negative state funding shocks.
-There are multiple ways that these changes in finances may affect the faculty composition at public universities.
+Universities can respond multiple ways to the falls in per student state funding.
+\cite{NBERw23736} established that public universities who experienced state funding cuts did not fully off-set by raising tution, and cut total spending.
+But it is not clear how these spending cuts went on to affect faculty --- or whether they did at all.
 
-It is a priori unclear that stagnating state support and changes in faculty composition at public universities would be causally linked.
-Over the 21st Century there have been drastic changes to college selectivity, where the top universities have become more selective while the average university less selective \citep{hoxby2009changing}.
-While the average public university is becoming less selective, it is possible that their most productive or research-focused faculty more often move to more selective and prestigious (and often private) institutions, so that over time public universities (on average) substitute towards contingent lecturers.
+Wider changes in the American higher education sector may concurrently explain the trends in state funding and faculty substitition.
+College selectivity changed drastically over the 21st century, where the top universities have become more selective while the average university less selective \citep{hoxby2009changing}.
+While the average public university is becoming less selective, it is possible that their most productive or research-focused faculty more often move to more selective and prestigious (and often private) institutions, leaving public universities substituting toward lecturers over time.
 This is one way in which the trends may be concurrent, but not causal.
+I address this issue by using ssift-share instrument approach to isolate funding cuts that affected more funding-intensive universities, measuring the rate of substitution between lecturers and professors in response to yearly changes in state-wide funding.
 
-Employment composition changes likely arise via differences in hiring rate across levels of professor seniority.
-Universities hire new professors most years, either to expand their departments or to replace leaving professors, but do not hire at the same rate for each level of seniority.
-The yearly hiring is usually highest among non-tenured adjunct faculty, where universities hire instructors often on short-term or contingent contracts, so that lecturers are less costly to hire (or to let go) in response to yearly changes.
-Tenure-track assistant professors are mostly hired with a four to six year contract, and formal agreement for tenure consideration at the end of this term.
-Lastly, tenured professors have successfully secured a full-time appointment at their university with no expiration date; this position has the lowest yearly hiring rate among most universities.
-Similarly, there is highest turn-over (including leaving the university) among lecturers, and lowest among full professors.
-The differences in yearly hiring and turnover rate mean that if hiring is affected by state funding, then there will be faculty composition change in the years following budget cuts.
+Universities could respond to funding cuts by cutting faculty salaries, leading to high-performing faculty leaving the university.
+Multiple universities passed a university-wide pay-cut for their faculty in response to state budget cuts around the 2008 recession, and Cornell University implemented a professor salary cut in 2020.\footnote{
+    The Cornell University administration expected large fiscal squeezes in mid-2020, so imposed a faculty salary cut, and then returned the amount cut later in 2020 when the financial squeeze did not materialise.}
+Funding cuts could also limit the amount public universities can offer to new faculty hires, leading to fewer or less accomplished faculty accepting these offers.\footnote{
+    Faculty often consider their outside options in job decisions, considering offers for employment and/or promotion from multiple universities at the same time (see \citealt{blackaby2005} for an overview of faculty outside options).
+    If public universities can only make low salary offers thanks to funding cuts, then they are less likely to land professors with multiple offers.
+}
+This effect may not be the same across faculty seniority; most universities hire more junior faculty each year, so that yearly hiring disruptions may disrupt the looser market for established full professors more than tighter market for assistant professors or lecturers.
+While these factors make it unclear who will be affected most by funding cuts, there is one empirical fact worth noting: the average assistant or tenured professor earns more than double that of a lecturer (see \autoref{tab:illinois-summary}).
+Secondly, lecturers are (by definition) granted shorter contracts than professors; assistant professors mostly have 5--6 year long contracts, full professors indefinite contracts (tenure), lecturers yearly or term-wise contracts.
+If a public university's primary obligation is to teach, and they must fulfil this objective with fewer resources, then substitution away from tenure-track and tenured professors towards lecturers is rationalisable.
 
-Faculty salaries may be affected by changes in public university finances.
-When the university has a lower budget for its yearly hiring, it may respond by lowering the salary they offer to new hires.
-This possible effect may not be the same across each position of professor; tenured faculty are often hired away from another university, so that new tenured professor hires may be less likely to accept the lower salary offers from public universities.\footnote{
-    Faculty often consider their outside options, as in offers for employment and/or promotion from other universities.
-    See \cite{blackaby2005} for an overview of faculty outside options.
-}
-Yet salary for all the professors, not just new hires, may also be affected: multiple universities passed a university-wide pay-cut for their faculty in response to state budget cuts around the 2008 recession, for example.\footnote{
-    Indeed, Cornell University implemented hiring freezes and a nominal salary reduction for professors, in anticipation of a financial shock in early 2021.
-    The salary cut was not permanent, as the oncoming financial shock turned out to not be as serious as projected, so that the cut was returned to professors in the year 2021.
-}
-Additionally, it is not immediately clear which faculty (among those already hired) will be most affected by the changes in their university's funding.
-Yet, there is one empirical fact worth noting: the average assistant or tenured professor earns more than double that of a lecturer (see \autoref{tab:illinois-summary}).
-If a public university's primary obligation is to teach, and they must fulfil this objective with less resources, then they may substitute away from tenure-track and tenured professors towards lecturers.
-
-Similarly, its is not clear when the effects of funding short-falls will be realised.
+It is not clear when the effects of funding short-falls will be realised.
 University departments often employ faculty on multiple-year contracts (e.g., the tenure contract), and individual professors' decisions to retire, move university, or leave academia are multiple-year commitments, so that it may take multiple years for university funding to change the faculty composition, or trickle down to faculty.
 For example, university departments coordinate hiring in annual cycles, so that a budget short-fall in year $t$ may have no effect on the in-progress hiring committee and resulting hiring decisions, yet the hiring cycle in year $t+1$ may be postponed or cancelled in response.
-As such, this paper investigates the dynamics of state funding shocks' effects on faculty, for the years following a university funding shock.
+As such, this paper investigates the dynamics of state funding shocks' effects on faculty, for the years following a state funding cut.

### 2023-12-19 programs/analysis/ipeds-shock.R

@@ -863,6 +863,61 @@ stargazer(
     type = "text",
     out = "../../text/tables/facultycount-shock-reg-fte.tex")
 
+# Plot the results.
+library(jtools)
+# Estimate with separate names
+plot_lecturer_count.reg <- reg.data %>%
+    mutate(lect_funding = log(stateappropriations_real / enrollment_reported)) %>%
+    felm(log(lecturer_prof_count / enrollment_reported) ~ 1 |
+        unitid + year |
+        (lect_funding ~ I(-log(appropriationshock_perEnroll_real))) |
+        state + year,
+        data = .)
+plot_assistant_count.reg <- reg.data %>%
+    mutate(asst_funding = log(stateappropriations_real / enrollment_reported)) %>%
+    felm(log(assistant_prof_count / enrollment_reported) ~ 1 |
+        unitid + year |
+        (asst_funding ~ I(-log(appropriationshock_perEnroll_real))) |
+        state + year,
+        data = .)
+plot_full_count.reg <- reg.data %>%
+    mutate(full_funding = log(stateappropriations_real / enrollment_reported)) %>%
+    felm(log(full_prof_count / enrollment_reported) ~ 1 |
+        unitid + year |
+        (full_funding ~ I(-log(appropriationshock_perEnroll_real))) |
+        state + year,
+        data = .)
+plot_all_count.reg <- reg.data %>%
+    mutate(all_funding = log(stateappropriations_real / enrollment_reported)) %>%
+    felm(log(all_prof_count / enrollment_reported) ~ 1 |
+        unitid + year |
+        (all_funding ~ I(-log(appropriationshock_perEnroll_real))) |
+        state + year,
+        data = .)
+elasticity.plot <- plot_summs(
+    plot_lecturer_count.reg,
+    plot_assistant_count.reg,
+    plot_full_count.reg,
+    plot_all_count.reg,
+    coefs = c(
+        "Lecturers" = "`lect_funding(fit)`",
+        "Assistant Professors" = "`asst_funding(fit)`",
+        "Full Professors" = "`full_funding(fit)`",
+        "All Professors" = "`all_funding(fit)`"),
+    legend.title = "",
+    plot.distributions = TRUE,
+    inner_ci_level = 0.95) +
+    theme_bw() +
+    theme(
+        #plot.title = element_text(size = rel(1)),
+        plot.margin = unit(c(0.5, 0, 0, 0), "mm"),
+        legend.position = "none",
[…]

### 2023-12-19 text/paper.tex
Writing updates, Table A1 needs re-do.

@@
         Douglas Miller,
-        and Evan Riehl,
-        as well as seminar participants at Cornell University for helpful discussion.} \\
+        and Evan Riehl.
+        I thank seminar participants at Cornell University (2022, 2023) for helpful discussion.} \\
     \vspace{0.1cm}

### 2023-12-19 text/sections/00-abstract.tex
Writing updates, Table A1 needs re-do.

@@
-Public universities employ more lecturers than ever, while state funding for higher education has stagnated for over thirty years.
+Public universities employ more lecturers and fewer professors than at any other point in the last thirty years, relative to student enrolment.
+At the same time, state funding for higher education has stagnated.
 This paper shows falling state funding caused a substitution away from professors towards lecturers at US public universities, using a shift-share approach to instrument for state funding cuts relative to enrolment.
@@
 Incumbent professors' salaries, promotion rate, and quit rate at Illinois universities are not affected by funding cuts, so that the substitution arose by limiting hiring of new tenure-track/tenured professors.
-Stagnating public funding affects public universities as institutions, and likely contributed to deteriorating student outcomes at public universities since the 1990s.
+Stagnating public funding affects faculty and public universities as institutions, likely contributing to worsening student outcomes at public universities since the 1990s.
 

### 2023-12-19 text/sections/01-introduction.tex
Writing updates, Table A1 needs re-do.

@@
 For example, university administrations may directly cut salaries or lay-off incumbent staff.
-This happened in 2015 when the state of Illinois shut-down without setting a state budget, and the University of Illinois furloughed faculty thanks to the funding shortfall \citep{furlough2010}.
+This happened in 2015 when the state of Illinois shut-down without setting a state budget, and the University of Illinois furloughed faculty unable to pay salaries \citep{furlough2010}.
 Similarly, faculty hiring committees are often cancelled or disrupted in response to funding cuts, meaning there are fewer faculty entering a department to replace those leaving.
@@
 Lecturers can fill in teaching duties, at a lower cost for the universities than tenure-track professors, with less employment protections (e.g., no tenure).
-But it is not certain that universities would hire more lecturers in the face of budget cuts; they may not hire any more instructional staff thanks to lack of resources.
+But it is not certain that universities would hire more lecturers in the face of budget cuts; they may not hire any more instructional staff at all.
 Similarly, the universities may be limited from hiring more lecturers by competitive pressures from peer institutions \citep{hoxby2009changing,urquiola2020markets}, and instead focus on filling the funding gap by raising revenue from other sources, such as private funds or raising tuition.
@@
 }
+This approach allows me to test whether universities recover in years following the fundings cuts e.g., by resuming faculty hiring in later years.
 
@@
 In my analysis of all Illinois faculty (IBHED), I find that incumbent professors were not meaningfully affected, in terms of total salary, promotion rate, or rate of leaving the Illinois public university system.
-Yet the hiring rate for new professors at public universities was negatively impacted by the state funding cuts.
-This implies that faculty composition change arose by limiting the hiring of new professors; public universities increasingly hired contingent lecturers, and increasingly did not replace their retiring (or otherwise leaving) professors.
-Suggestive evidence using data on cumulative hiring counts supports this hypothesis.
+Yet the hiring rate for new professors at public universities was negatively impacted by state funding cuts.
+This implies that faculty substitution arose by limiting the hiring of new professors, which is supported by suggestive evidence on cumulative faculty hiring counts.
+The substitution is persistent for up to ten years, meaning that universities did not resume professor hiring in years following the funding cut, making the substitution towards lecturers long-run.
 

### 2023-12-19 text/sections/03-concepts.tex
Writing updates, Table A1 needs re-do.

@@
 Universities can respond multiple ways to the falls in per student state funding.
-\cite{NBERw23736} established that public universities who experienced state funding cuts did not fully off-set by raising tution, and cut total spending.
-But it is not clear how these spending cuts went on to affect faculty --- or whether they did at all.
+\cite{NBERw23736} established that public universities who experienced state funding cuts did not fully off-set by raising tuition, and cutting total spending.
+But it is not clear how these spending cuts went on to affect faculty --- or whether they affected faculty at all.
 
-Wider changes in the American higher education sector may concurrently explain the trends in state funding and faculty substitition.
+Wider changes in the American higher education sector may concurrently explain the trends in state funding and faculty substitution.
 College selectivity changed drastically over the 21st century, where the top universities have become more selective while the average university less selective \citep{hoxby2009changing}.
@@
 This is one way in which the trends may be concurrent, but not causal.
-I address this issue by using ssift-share instrument approach to isolate funding cuts that affected more funding-intensive universities, measuring the rate of substitution between lecturers and professors in response to yearly changes in state-wide funding.
+I address this issue by using shift-share instrument approach to isolate funding cuts that affected more funding-intensive universities, measuring the rate of substitution between lecturers and professors in response to yearly changes in state-wide funding.
 
-Universities could respond to funding cuts by cutting faculty salaries, leading to high-performing faculty leaving the university.
+Universities could respond to funding cuts by cutting faculty salaries, leading to more faculty leaving for jobs at other universities.
 Multiple universities passed a university-wide pay-cut for their faculty in response to state budget cuts around the 2008 recession, and Cornell University implemented a professor salary cut in 2020.\footnote{
@@
 This effect may not be the same across faculty seniority; most universities hire more junior faculty each year, so that yearly hiring disruptions may disrupt the looser market for established full professors more than tighter market for assistant professors or lecturers.
-While these factors make it unclear who will be affected most by funding cuts, there is one empirical fact worth noting: the average assistant or tenured professor earns more than double that of a lecturer (see \autoref{tab:illinois-summary}).
-Secondly, lecturers are (by definition) granted shorter contracts than professors; assistant professors mostly have 5--6 year long contracts, full professors indefinite contracts (tenure), lecturers yearly or term-wise contracts.
-If a public university's primary obligation is to teach, and they must fulfil this objective with fewer resources, then substitution away from tenure-track and tenured professors towards lecturers is rationalisable.
+While these factors make it unclear who will be affected most by funding cuts, there is one empirical fact worth noting: the average Illinois assistant professor is paid $\$76,897$, tenured professor $\$109,283$, while lecturers earn $\$31,449$ more than double that of a lecturer (annual salary for 2010--2021 in 2021 USD, see \autoref{tab:illinois-summary}).
+Secondly, lecturers are (by definition) granted shorter contracts than professors; assistant professors mostly work on 5--6 year long contracts, full professors on indefinite contracts (tenure), lecturers on yearly or term-wise contracts.
+If a public university's primary obligation is to teach, and they must fulfil this objective with fewer resources, substituting away from tenure-track and tenured professors towards lecturers is rationalisable.
 
-It is not clear when the effects of funding short-falls will be realised.
+It is not clear when the effects of funding cuts will be realised.
 University departments often employ faculty on multiple-year contracts (e.g., the tenure contract), and individual professors' decisions to retire, move university, or leave academia are multiple-year commitments, so that it may take multiple years for university funding to change the faculty composition, or trickle down to faculty.
-For example, university departments coordinate hiring in annual cycles, so that a budget short-fall in year $t$ may have no effect on the in-progress hiring committee and resulting hiring decisions, yet the hiring cycle in year $t+1$ may be postponed or cancelled in response.
+For example, university departments coordinate hiring in annual cycles, so that a budget short-fall in year $t$ may have no effect on the in-progress hiring committee and resulting hiring decisions, yet the hiring cycle in year $t+1$ may be postponed or cancelled.
 As such, this paper investigates the dynamics of state funding shocks' effects on faculty, for the years following a state funding cut.

### 2023-12-19 text/sections/04-empirical.tex
Writing updates, Table A1 needs re-do.

@@
 Instead, the state government and university administration undertake a complex process of allotting resources across multiple different priorities, including instruction, research, or between departments.
-For example, it may be the case that a state government restricts funding for only its lowest performing universities in response to a budget shock, introducing a treatment selection-on-outcome threat to identification.
+For example, it may be the case that a state government restricts funding for only its lowest performing universities in response to a budget shock, introducing a selection-on-outcomes threat to identification.
 Importantly, funding from a university's state government provides opportunity to address such endogeneity.
@@
 The local projections method is an empirical model used to estimate dynamic treatment effects when the treatment is not binary, so that time-series confounding (i.e., treatment in time $t$ is correlated with treatment in time $t-1$) is present \citep{montiel2021local}.
-This method estimates the effect of treatment $X_{i,t}$ on outcome $Y_{i, t+ k}$ in follow-on years $k > 0$, while accounting for the auto-correlation between the other years.\footnote{
+This method estimates the effect of state funding $X_{i,t}$ on faculty outcomes $Y_{i, t+ k}$ in follow-on years $k > 0$, while accounting for the auto-correlation between state funding in surrounding years.\footnote{
     The funding shock has a persistent effect on state funding, multiple years after the initial funding shock, so that the instrument is similarly strong for the local projections method (see \autoref{fig:firststage-lp}).
 }
-Additionally, the approach accommodates use of an instrument for state funding \citep{olea2021inference}, so that I use the shift-share instument in local projections estimates as in the rest of the empirical analysis.
-I present estimates in graphical form, where the $x$-axis represents the year relative to the funding shock (i.e., $x = 0$ represents the year of the state funding shock) and the $y$-axis represents the estimated effect of the funding shock on that year's outcome.\footnote{
-    See \autoref{fig:count-lp} for the graphical format, and the corresponding note for further explanation.
-}
+Additionally, the approach accommodates the shift-share instrument for state funding \citep{olea2021inference}, so that I present instrumental variables estimates for this method as in the rest of the empirical analysis.

### 2023-12-19 text/sections/09-figures-tables.tex
Writing updates, Table A1 needs re-do.

@@
     \textbf{Note}:
-    These figures show the local projections estimates of regression specification \eqref{eqn:secondstage}, with the funding shock as an instrument for state funding.
+    These figures show the local projections estimates of regression specification \eqref{eqn:secondstage}, with the funding shock shift-share as an instrument for state funding.
     The unit of analysis is the university, and uses IPEDS data.
+    The $x$-axis represents the year relative to the funding shock (i.e., $x = 0$ represents the year of the state funding cut) and the $y$-axis represents the estimated effect in years relative to initial funding shock.
     The coefficient estimate is effect of state funding ($X_{i,t}$) on faculty count per student ($Y_{i,t}$), using the funding shock instrument ($Z_{i,t}$), while accounting for auto-correlation between different time periods --- i.e., between $X_{i,t}, X_{i,t-1}$ and $Y_{i,t}, Y_{i,t-1}$.
@@
     \textbf{Note}:
-    These figures show the local projections estimates of regression specification \eqref{eqn:secondstage}, with the funding shock as an instrument for state funding.
+    These figures show the local projections estimates of regression specification \eqref{eqn:secondstage}, with the funding shock shift-share as an instrument for state funding.
     The unit of analysis is an individual faculty member (at an Illinois public university); funding data come from IPEDS, and faculty salaries from IBHED.
+    The $x$-axis represents the year relative to the funding shock (i.e., $x = 0$ represents the year of the state funding cut) and the $y$-axis represents the estimated effect in years relative to initial funding shock.
     The coefficient estimate is effect of state funding ($X_{i(j),t}$) on faculty salaries ($Y_{j,t}$), using the funding shock instrument ($Z_{i(j),t}$), while accounting for auto-correlation between different time periods --- i.e., between $X_{i(j),t}, X_{i(j),t-1}$ and $Y_{i(j),t}, Y_{i(j),t-1}$.

### 2023-12-19 programs/analysis/ibhed-shock-longterm.R

@@ -227,7 +227,7 @@ lpreg.plot <- function(model.lpreg) {
         geom_line(aes(y = estimate), colour = "blue") +
         geom_line(aes(y = conf.low), linetype = "dashed") +
         geom_line(aes(y = conf.high), linetype = "dashed") +
-        scale_x_continuous(name = "Years, Relative to Initital Shock",
+        scale_x_continuous(name = "Years, Relative to Funding Cut",
             breaks = model.lpdata$t, expand = c(0.01, 0.01)) +
         scale_y_continuous(name = "",
             limits = c(-0.2, 0.2),

### 2023-12-19 programs/analysis/ipeds-shock-longterm.R

@@ -140,7 +140,7 @@ lpreg.plot <- function(model.lpreg, limits.vector) {
         geom_line(aes(y = estimate), colour = "blue") +
         geom_line(aes(y = conf.low), linetype = "dashed") +
         geom_line(aes(y = conf.high), linetype = "dashed") +
-        scale_x_continuous(name = "Years, Relative to Initital Shock",
+        scale_x_continuous(name = "Years, Relative to Funding Cut",
             breaks = model.lpdata$t, expand = c(0.01, 0.01)) +
         scale_y_continuous(name = "",
             limits = limits.vector,

### 2023-12-19 text/sections/01-introduction.tex
Writing updates, Table A1 needs re-do.

@@
 On the other hand, some faculty positions may grow at universities in response to funding cuts.
-Public universities do not restrict enrolment in response to funding cuts, so someone must teach the next year of students.
-Lecturers can fill in teaching duties, at a lower cost for the universities than tenure-track professors, with less employment protections (e.g., no tenure).
-But it is not certain that universities would hire more lecturers in the face of budget cuts; they may not hire any more instructional staff at all.
-Similarly, the universities may be limited from hiring more lecturers by competitive pressures from peer institutions \citep{hoxby2009changing,urquiola2020markets}, and instead focus on filling the funding gap by raising revenue from other sources, such as private funds or raising tuition.
+Lecturers can fill in teaching duties when professors are unavailable, and may do so at a lower cost  on on short-term and contracts with less employment protections.
+But it is not certain that universities would hire more lecturers in the face of budget cuts; they may not hire any more lewcturers or professors at all.
+Similarly, the universities may be limited from hiring more lecturers by competitive pressures from peer institutions \citep{hoxby2009changing,urquiola2020markets}, and instead focus on filling the funding gap by courting private funding or raising tuition prices.
 
-In both analyses, I employ a shift-share instrument to identify changes in state funding for higher education.
-State governments decide how to fund higher education by a complicated process which can be influenced by local economic conditions, changing state priorities, or informal lobbying from universities.
-As such, state funding is not randomly assigned to universities and so it is necessary to employ an identification strategy to account for this.
-The shift-share instrument identifies exogenous changes in state funding by exploiting differences in how much each university relied on state funding, interacted with state-wide changes in higher education funding --- following \cite{NBERw23736,NBERw27885}.
-The approach relies on the insight that yearly state government budgets are decided without targetting individual universities for funding cuts, and exploits the fact that more reliant universities will be more affected by state-wide changes in higher education funding.
+State governments decide how to fund higher education by a complicated process, which could be influenced by lobbying, changing state priorities, or local economic conditions.
+As such, it would be na\"ive to think of state higher education funding as independent of public university faculty, so I employ a shift-share instrumental variable approach to identify causal effects of state funding cuts.
+The shift-share instrument identifies exogenous changes in state funding by exploiting differences in how much universities relied on that funding, interacted with state-wide changes in higher education funding --- following \cite{NBERw23736,NBERw27885}.
+Yearly state government budgets are decided without targetting individual universities for funding cuts, so more reliant universities will be more affected by state-wide changes in higher education funding, allowing the shift-share approach to estimate causal effects among more reliant unversities.
 Additionally, I estimate effect of state funding on faculty multiple years after the initial funding cuts, using the local projections method thanks to time-series confounding between the funding shock and later years' level of state funding.\footnote{

### 2023-12-20 text/sections/01-introduction.tex
Writing updates, Table A1 needs re-do.

@@
 Previous research shows declining in state funding led to worsening student outcomes \citep{NBERw23736,NBERw27885}, but there is little evidence on whether this funding decline has impacted faculty.
-In this paper, I show that falling state funding caused a substitution away from professors toward lecturers at US public universities.
+In this paper, I show that falling state funding caused a change in the composition of faculty at US public universities, substituting professors for more lecturers.
 
@@
 Universities employ $4.4$\% more lecturers per student following a $10$\% funding cut, and correspondingly rely less on tenure-track faculty by employing $1.4$\% fewer assistant professors and $1.2$\% fewer full professors per student.
-In a secondary analysis of all public university faculty in the state of Illinois, I show that incumbent faculty were relatively unaffected, in terms of wages, promotion and quit rates.
+In a secondary analysis of all public university faculty in the state of Illinois, I show that incumbent faculty were unaffected, in terms of wages, promotion and quit rates.
 This implies that the faculty substitution arose by disrupting hiring of new tenure-track/tenured professors.
@@
 Similarly, faculty hiring committees are often cancelled or disrupted in response to funding cuts, meaning there are fewer faculty entering a department to replace those leaving.
-On the other hand, some faculty positions may grow at universities in response to funding cuts.
-Lecturers can fill in teaching duties when professors are unavailable, and may do so at a lower cost  on on short-term and contracts with less employment protections.
-But it is not certain that universities would hire more lecturers in the face of budget cuts; they may not hire any more lewcturers or professors at all.
+On the other hand, lecturers could be hired to fill in teaching duties, on shorter and cheaper contracts with fewer employment protections.
+But it is not certain that universities would hire more lecturers in the face of budget cuts; they may not hire any more faculty at all.
 Similarly, the universities may be limited from hiring more lecturers by competitive pressures from peer institutions \citep{hoxby2009changing,urquiola2020markets}, and instead focus on filling the funding gap by courting private funding or raising tuition prices.
@@
 }
-This approach allows me to test whether universities recover in years following the fundings cuts e.g., by resuming faculty hiring in later years.
+This approach allows me to test whether universities recover in years following the funding cuts e.g., by resuming faculty hiring in later years.
 
@@
 Substitution towards lecturers is likely a cost-cutting measure for funding constrained universities, though relying on these faculty --- who are often over-worked and not granted long-term employment protections --- may lead to worse student outcomes \citep{ehrenberg2005tenured,zhu2021limited,jaeger2011examining}.
-Lastly, my results provide evidence that the long term trends in higher education funding are causally related, contributing to the literature on faculty outcomes \citep{ehrenberg2003studying} and trends in US higher education and funding \citep{hoxby2009changing,ehrenberg2012american}.
+My results provide evidence that the long term trends in higher education funding are causally related, contributing to the literature on faculty outcomes \citep{ehrenberg2003studying} and trends in US higher education and funding \citep{hoxby2009changing,ehrenberg2012american}.
 

### 2023-12-20 text/sections/03-concepts.tex
Writing updates, Table A1 needs re-do.

@@
 While these factors make it unclear who will be affected most by funding cuts, there is one empirical fact worth noting: the average Illinois assistant professor is paid $\$76,897$, tenured professor $\$109,283$, while lecturers earn $\$31,449$ more than double that of a lecturer (annual salary for 2010--2021 in 2021 USD, see \autoref{tab:illinois-summary}).
-Secondly, lecturers are (by definition) granted shorter contracts than professors; assistant professors mostly work on 5--6 year long contracts, full professors on indefinite contracts (tenure), lecturers on yearly or term-wise contracts.
+Secondly, lecturers are (by definition) granted shorter contracts than professors; assistant professors mostly work on 5--6 year long contracts, full professors on indefinite contracts (tenure), lecturers on yearly or termly contracts.
 If a public university's primary obligation is to teach, and they must fulfil this objective with fewer resources, substituting away from tenure-track and tenured professors towards lecturers is rationalisable.

### 2023-12-20 text/sections/04-empirical.tex
Writing updates, Table A1 needs re-do.

@@
 \cite{NBERw23736,NBERw27885} develop the instrument for public university finances by exploiting a shift-share instrument for changes in state-level funding interacted with university reliance on state funding in a base period.
+
 \begin{align}
@@
 \end{align}
+
 The system exploits the fact that institutions who rely on state funding more will be affected by state funding shocks.

### 2023-12-20 text/sections/05-results.tex
Writing updates, Table A1 needs re-do.

@@
 \autoref{tab:firststage-reg} presents results of the first-stage regression among IPEDS data, separately with and without a control for tuition revenue per student, plus institution and year fixed effects.
-I use the same empirical strategy in the analysis of Illinois faculty, so  \autoref{tab:firststage-reg} presents similar first-stage estimates among the IBHED data.
+I use the same empirical strategy in the analysis of Illinois faculty, so  \autoref{tab:firststage-illinois} presents first-stage estimates among the IBHED data (which are very similar to those calculated in IPEDS data).
 \autoref{tab:firststage-reg} Panel A shows estimates of a funding shock of \$1 per student in the state on state funding per student at the university, and Panel B the \% increase effects of a 1\% increase in the shock per student on state funding per student.
@@
 
-The composition of faculty changed at public universities 1990--2017 thanks to stagnating state funding, but there are little discernible effects on incumbent faculty.
-This implies that changes in faculty composition arose by state funding cuts impacting the rate that public universities hired new professors.
+\subsection{Reduced Hiring of New Professors}
+Public universities substituted professors for more lecturers thanks to stagnating state funding, but there are little discernible effects on incumbent professors or lecturers.
+This implies that the substitution arose by impacting the hiring of new professors to replace those retiring or leaving academia.
+% \textbf{Expand this paragraph}.
 I take recent data on the total faculty hired at US public universities over 2011--2021 (provided by \citealt{wapman2022quantifying,wapman2022zenodo}), to show that universities with lower state funding (per student) also hired fewer professors in the same time period --- shown in \autoref{fig:hiring-correlation}.
-Similarly, the funding shock IV model shows that a $10$\% cut in per student state funding leads to $13$\% fewer faculty hires per student across the decade 2011--2021.\footnote{
+The shift-share IV model shows that a $10$\% cut in per student state funding leads to $13$\% fewer faculty hires per student across the decade 2011--2021.\footnote{
     This limited sample of data are a cross-section for count of professor hires 2011--2021, so that local projection estimates are not possible here.
@@
 }
-These results provide limited evidence that changes in faculty composition arose by disrupting hiring of new professors at public universities, corresponding to the concurrent results that faculty composition changed while incumbent faculty were not by meaningfully affected by the funding cuts.
+These results provide suggestive evidence that funding cuts caused faculty substitution by disrupting hiring of new professors at public universities.
+This provides a coherent mechanism for how incumbent faculty (those who are already hired at the university) were not meaningfully impacted by funding cuts, yet their wider composition nonetheless changed when their universities experienced funding cuts.
+
+%\subsection{Robustness checks}
+% Describe robustness checks here.

### 2023-12-20 text/sections/07-conclusion.tex
Writing updates, Table A1 needs re-do.

@@
 
-This paper investigates how the recent stagnation in state support for higher education has affected faculty, and their composition, at US public universities.
+This paper studies how stagnating state funding for higher education affected faculty, and their composition, at US public universities.
 This work contributes to the literature along two primary dimensions.
-Firstly, by isolating changes in state funding on public universities via state funding shocks, this work provides an explanation for the increased relaiance on lecturers and away from tenure-track and tenured professors at US public universities.
+First, by isolating changes in state funding on public universities via state funding shocks, this work provides an explanation for the increased reliance on lecturers and away from tenure-track and tenured professors at US public universities.
 This approach used multiple methods to estimate the short- and medium-run effects of state funding cuts on faculty outcomes.
-Secondly, this work asks investigates individual faculty are affected by changes in state funding for their university using a dataset new in the economic literature (IBHED).
+Second, this work investigates how individual faculty are affected by changes in state funding for their university using a dataset new in the economic literature (IBHED).
 These data allow for detailed analysis of thousands of faculty salaries, and employment outcomes in the Illinois university system.
 
-Public universities have systematically substituted away from tenure-track and tenured professors, towards lecturers, in the face of persistent declines in state funding for higher education.
-The effects on incumbent professors in the state of Illinois are non-distinguishable from zero, with suggestive evidence for falls in faculty hiring, implying that the changes in faculty composition are driven by reduced hiring of new professors.
+Public universities have systematically substituted away from tenure-track and tenured professors, toward lecturers, in the face of persistent declines in state funding for higher education.
+The effects on incumbent professors in Illinois are non-distinguishable from zero, with suggestive evidence for reduced hiring of new professors, implying that the faculty substitution is driven by reduced hiring of new professors.
 Public universities are using more contingent lecturers to teach their students, while private universities continue to employ more tenure-track and tenured professors than their public counterparts, and each year the gap widens.

### 2023-12-20 text/paper.tex
Writing updates, Table A1 needs re-do.
Updates.

@@
         and Evan Riehl.
-        I thank seminar participants at Cornell University (2022, 2023) for helpful discussion.} \\
+        I thank seminar participants at Cornell University (2022, 2023) for helpful discussion.
+        This project's repository, with all materials and anonymised data needed for replication, is available at 
+        \url{https://github.com/shoganhennessy/state-funding-faculty}.
+        Any comments or suggestions may be sent to me at \href{mailto:seh325@cornell.edu}{\nolinkurl{seh325@cornell.edu}}, or raised as an issue on the Github project.
+    } \\
     \vspace{0.1cm}
-    Economics Department, Cornell University
+    Economics Department, Cornell University\footnote{
+        Address: Uris Hall \#447, Economics Department, Cornell University NY 14853 USA.
+    }
 }
@@
     \Large{Stagnating State Funding for Higher Education and its Effect on Faculty at Public Universities}}
-\date{This version: \today\thanks{
-    This project's Github repository, with materials and anonymised data for replication, is available at 
-    \url{https://github.com/shoganhennessy/state-funding-faculty}.
-    Any comments or suggestions may be sent to me at \href{mailto:seh325@cornell.edu}{\nolinkurl{seh325@cornell.edu}}, or raised as an issue on the Github project.
-    } \\ \vspace{0.1cm}
+\date{This version: \today \\ \vspace{0.1cm}
     \href{https://github.com/shoganhennessy/state-funding-faculty/blob/main/state-funding-faculty-2023.pdf}{Newest version available here.}
@@
 \input{sections/07-conclusion.tex}
+% Rev of Econ Ed Statements.
+\subsubsection*{Credit Author Statement}
+As the only author on this paper, I am responsible for all contributions in this paper.
+\subsubsection*{Declaration of Competing Interest}
+This paper's author has no competing interests to declare.
 

### 2023-12-20 text/sections/02-data.tex
Updates.

@@
 US state governments, by majority, are legally obligated to run a balanced-budget, so that yearly variation in tax revenues (caused by changing economic conditions or otherwise) necessarily affect state expenditures.
-Public universities have relatively lower lobbying power than other funding beneificaries, so often bear the brunt of state government funding cuts \citep{delaney2011state}.\footnote{
+Public universities have relatively lower lobbying power than other funding beneficiaries, so often bear the brunt of state government funding cuts \citep{delaney2011state}.\footnote{
     \cite{delaney2011state} fully describe the financial environment of state expenditures, and what makes spending on higher education an attractive area for state governments to expand funding during years of higher tax revenues, and retract funding in leaner years.
@@
 There are large differences in the average number of professors per student between the private and public sector.
-\autoref{fig:fte-perprof} shows that private universities start with a higher baseline of around 4.5 professors per hundred students, and exhibited yearly variation of less than 0.5 professor per hundred students over the thirty years.
-Public universities start with 3.9 professors per hundred students, and this number falls to around 3., with the largest fall in the 2008-2011 time period.
-We see a similar difference in baseline, and fall for the years 2008-2011 for assistant professors.
+\autoref{fig:fte-perprof} shows that private universities start with an average of 40 students for every one full professor, and little average trend until 2021.
+Public universities start with 40 students per professor, and by 2023 there are 52.4 students professors for every full professor --- with the largest rise coming in the 2008-2011 time period.
+The trend, and divergence between public and private universities, is similar for the number of assistant professors.
 Private and public universities have similar numbers of associated and full professors before the year 2000, yet this number has fallen by over 20\% in the next 20 years only for public universities: in 2021 the mean public university has 6 fewer full professors per hundred students than the mean private university.
-Over the same time period, we see the rise in use of non-tenure track instructor positions (referred to as lecturers from here on), who were employed at similar rates in both sectors in 1990, yet have been utilised by public universities at a higher rate since.
+Over the same time period, we see the rise in use of non-tenure track instructor positions (collectively referred to as  lecturers), who were employed at similar rates in both sectors in 1990, yet have been utilised by public universities at a higher rate since.
 
@@
 Illinois funding for higher education has stagnated, and has experienced serious declines in the decade 2010--2021 --- similar in magnitude to the nation-wide decline over 1990--2021.
-State funding among the seven Illinois public university campuses has fallen by over 50\% for the last decade, on both an absolute and per student basis (see \autoref{fig:illinois-funding}).
+State funding among the seven Illinois public university campuses has fallen by over 50\% for the last decade, on both an absolute and per-student basis (see \autoref{fig:illinois-funding}).
 Similar to the figures for the rest of the country, there has been a corresponding substitution towards tuition revenue.

### 2023-12-20 text/sections/05-results.tex
Updates.

@@
 \autoref{tab:firststage-reg} presents results of the first-stage regression among IPEDS data, separately with and without a control for tuition revenue per student, plus institution and year fixed effects.
-I use the same empirical strategy in the analysis of Illinois faculty, so  \autoref{tab:firststage-illinois} presents first-stage estimates among the IBHED data (which are very similar to those calculated in IPEDS data).
-\autoref{tab:firststage-reg} Panel A shows estimates of a funding shock of \$1 per student in the state on state funding per student at the university, and Panel B the \% increase effects of a 1\% increase in the shock per student on state funding per student.
-Column (1) shows that a funding shock of \$10 (per student in the entire state) is associated with \$11.76 less state funding per student at the university, with the corresponding -10\% funding shock leads to -9.77\%
-less funding (column 1, panel B), with similar estimates with and without including fixed effects.
-Column (2) shows estimates of the first-stage without including fixed effects, and gives less precise estimates for the funding shock, likely thanks to systematic differences in universities unaccounted for without fixed effects.
-Columns (3) and (4) include the tuition revenue control (explained in \autoref{foot:control}) to exhibit estimates with the inclusion of this possible collider (or bad control).
-Column (3) shows similar estimates to column (1) in both Panels A and B thanks to inclusion of fixed effects, so that the uncertainty in including tuition revenue as a possible bad control for the level of state funding does not matter thanks to the inclusion of fixed effects.
-Column (1) represents the estimates for \autoref{eqn:firststage} with fixed effects, omitting the tuition revenue control, and is the preferred form that I proceed with.
+I use the same empirical strategy in the analysis of Illinois faculty, and \autoref{tab:firststage-illinois} presents first-stage estimates among the IBHED data ; results are similar to those calculated in IPEDS data.
+\autoref{tab:firststage-reg} Panel A shows estimates of a state-wide funding shock of \$1 per student on state funding per student at the university, and Panel B the effects of a 1\% increase in the state-wide funding shock on state funding per student.
+Column (1) shows that a state-wide funding shock of \$10 per student is associated with \$11.76 less state funding per student at the university; -10\% state-wide funding shock leads to -9.77\% state funding per student at the average university.
+Estimates with and without including fixed effects.
+Column (2) shows estimates of the first-stage without including fixed effects, and gives less precise estimates for the funding shock --- this may be due to systematic differences in universities, which are not accounted for without fixed effects.
+Columns (3) and (4) include the tuition revenue as a control (explained in \autoref{foot:control}) to exhibit estimates with or without this control variable.
+Column (3) shows similar estimates to column (1) in both Panels A and B thanks to inclusion of fixed effects, so that 
+whether tuition revenue is a bad control or not does not affect empirical estimates.
+Column (1) represents the estimates for \autoref{eqn:firststage} with fixed effects (not including tuition revenue as a control variable), and is the preferred form that I proceed with.
 
-The above shows that the first stage is strong among both IPEDS and IBHED data, and the argument in \autoref{sec:empirics} for exogenity comes from independence of the state funding mechanism to any indivdiaul public university campus.
+The first stage is strong in both IPEDS and IBHED data, and the argument in \autoref{sec:empirics} for exogeneity comes from independence of the state funding mechanism to any individual public university campus.
 \autoref{tab:firststage-balance} provides a balance test, showing that the shift-share instrument is correlated with the professor outcomes, but not with other sources of university funding.
@@
 This implies that the substitution arose by impacting the hiring of new professors to replace those retiring or leaving academia.
-% \textbf{Expand this paragraph}.
+% \textbf{Expand this paragraph, referencing the figure, and the OLS tables for the appendix}.
 I take recent data on the total faculty hired at US public universities over 2011--2021 (provided by \citealt{wapman2022quantifying,wapman2022zenodo}), to show that universities with lower state funding (per student) also hired fewer professors in the same time period --- shown in \autoref{fig:hiring-correlation}.
-The shift-share IV model shows that a $10$\% cut in per student state funding leads to $13$\% fewer faculty hires per student across the decade 2011--2021.\footnote{
+The shift-share IV model shows that a $10$\% per-student state funding cut leads to $13$\% fewer faculty hires per student across the decade 2011--2021.\footnote{
     This limited sample of data are a cross-section for count of professor hires 2011--2021, so that local projection estimates are not possible here.
@@
 }
+
 These results provide suggestive evidence that funding cuts caused faculty substitution by disrupting hiring of new professors at public universities.

### 2023-12-20 text/sections/09-figures-tables.tex
Updates.

@@
 \end{table}
+
+\newpage
+\begin{figure}[H]
+    \centering
+    \singlespacing
+    \caption{State Funding and Total Professor Hiring at Public Universities, Total for 2011--2021.}
+    \includegraphics[width=0.8\textwidth]{figures/hiring-correlation.png}
+    \label{fig:hiring-correlation}
+    \justify
+    \footnotesize
+    \textbf{Note}:
+    This figure shows the correlation between state funding and the number of total number of professors hired at public universities between 2011--2021, and the line is an OLS line of best-fit.
+    Data for the number of faculty hired at public universities are provided by \cite{wapman2022quantifying,wapman2022zenodo}, and funding data from IPEDS.
+\end{figure}

### 2023-12-20 text/sections/10-appendix.tex
Updates.

@@
 
-These results were produced by integrating the total count of faculty hires for 2010--2021 for the top-ranked 180 US universities with a sum of the funding variables, and then estimating the models specified in \autoref{sec:iv-model-uni}.
+These results were produced by integrating the total count of professor hires for 2010--2021 for the top-ranked 180 US universities with a sum of the funding variables, and then estimating the models specified in \autoref{sec:iv-model-uni}.
 There were no observable differences in the hiring rate of male vs female faculty.
 
-\begin{figure}[h!]
-    \centering
-    \singlespacing
-    \caption{State Funding and Faculty Hired at Public Universities, Total for 2011--2021.}
-    \includegraphics[width=0.8\textwidth]{figures/hiring-correlation.png}
-    \label{fig:hiring-correlation}
-    \justify
-    \footnotesize
-    \textbf{Note}:
-    This figure shows the correlation between state funding and the number of total number of faculty hired at public universities between 2011--2021, and the line is a line of best-fit.
-    Data for the number of faculty hired at public universities are provided by \cite{wapman2022quantifying,wapman2022zenodo}, and funding data from IPEDS.
-\end{figure}
-
-\newpage
 \begin{table}[h!]

### 2023-12-23 text/paper.tex
Updates.

@@
 % Rev of Econ Ed Statements.
+\subsubsection*{Data Statement}
+Every source of data, and the code to fully replicate the entire analysis is freely available at the following link:
+\url{https://github.com/shoganhennessy/state-funding-faculty}.
+IBHED data, while public knowledge, include personally identifiable names, so the data included in the replication package have names replaced with non-identifiable placeholders.
 \subsubsection*{Credit Author Statement}

### 2023-12-23 text/sections/09-figures-tables.tex
Updates.

@@
 
+\newpage
+\begin{figure}[H]
+    \centering
+    \singlespacing
+    \caption{State Funding and Total Professor Hiring at Public Universities, Total for 2011--2021.}
+    \includegraphics[width=0.8\textwidth]{figures/hiring-correlation.png}
+    \label{fig:hiring-correlation}
+    \justify
+    \footnotesize
+    \textbf{Note}:
+    This figure shows the correlation between state funding and total number of professors hired at each public university between 2011--2021; the line is an OLS line of best-fit.
+    Data for the number of faculty hired at public universities are provided by \cite{wapman2022quantifying,wapman2022zenodo}, and funding data from IPEDS.
+\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \end{table}
-
-\newpage
-\begin{figure}[H]
-    \centering
-    \singlespacing
-    \caption{State Funding and Total Professor Hiring at Public Universities, Total for 2011--2021.}
-    \includegraphics[width=0.8\textwidth]{figures/hiring-correlation.png}
-    \label{fig:hiring-correlation}
-    \justify
-    \footnotesize
-    \textbf{Note}:
-    This figure shows the correlation between state funding and the number of total number of professors hired at public universities between 2011--2021, and the line is an OLS line of best-fit.
-    Data for the number of faculty hired at public universities are provided by \cite{wapman2022quantifying,wapman2022zenodo}, and funding data from IPEDS.
-\end{figure}

### 2024-01-02 text/paper.tex
writing revision (working on the final revision).

@@
 \title{Less Funding, More Lecturers, and Fewer Professors: \\ \vspace{0.1cm}
-    \Large{Stagnating State Funding for Higher Education and its Effect on Faculty at Public Universities}}
+    Stagnating State Funding for Higher Education and its Effect on Faculty}
 \date{This version: \today \\ \vspace{0.1cm}
@@
 \rhead{\today}
-\lhead{Less Funding, More Lecturers, and Fewer Professors.}
+\lhead{Stagnating State Funding for Higher Education and its Effect on Faculty.}
 % Begin

### 2024-01-03 text/sections/00-abstract.tex
Updates.
writing revision (working on the final revision).

@@
 At the same time, state funding for higher education has stagnated.
-This paper shows falling state funding caused a substitution away from professors towards lecturers at US public universities, using a shift-share approach to instrument for state funding cuts relative to enrolment.
-Universities employ $4.4$\% more lecturers per student following a $10$\% funding cut, relying less on professors by employing $1.4$\% fewer assistant professors and $1.2$\% fewer full professors per student.
-Incumbent professors' salaries, promotion rate, and quit rate at Illinois universities are not affected by funding cuts, so that the substitution arose by limiting hiring of new tenure-track/tenured professors.
-Stagnating public funding affects faculty and public universities as institutions, likely contributing to worsening student outcomes at public universities since the 1990s.
+This paper shows that the decline in state funding led to a substitution away from professors toward lecturers at US public universities.
+Using a shift-share approach to instrument for state funding, I find that universities employ 4.4\% more lecturers per student following a 10\% funding cut.
+This shift is accompanied by a reduction in assistant professors and full professors per student by 1.4\% and 1.2\%, respectively.
+Incumbent professors' salaries, promotion rates, and quit rates at Illinois universities remain unaffected by funding cuts, indicating that the substitution arose from limiting the hiring of new tenure-track/tenured professors.
+Stagnating state funding impacts public universities and faculty, likely contributing to the deterioration of student outcomes at public universities since the 1990s.
 
@@
 \textbf{Key-words:}
-State and Local Budget and Expenditures,
+Faculty,
 Higher Education,
-Public Sector Labour Markets
+State Funding
 

### 2024-01-03 text/sections/01-introduction.tex
Updates.
writing revision (working on the final revision).

@@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Introduction section
+\section{Introduction}
+\label{sec:intro}
+
 States fund their public universities with roughly the same amount today as they did in the 1990s, despite their enrolment rising by over 50\% on average.
@@
 My second analysis uses panel data for every faculty member at Illinois public universities for 2010--2021 \citep{ibhed}, to measure how incumbent faculty (i.e., those already employed at the university) are affected by changes in state funding.
-These data allow me to answer the following research question: if a US public university receives an extra \$1,000 per student (or a 10\% increase), do they change how many faculty they employ?
-If so, which positions do universities substitute away from, or towards, and do these funding cuts affect the faculty themselves?
+These data allow me to answer the following research question: if a US public university receives an extra \$1,000 per student (or a 10\% increase), do they change how many faculty members they employ?
+If so, which positions do universities substitute away from, and do these funding cuts affect the faculty themselves?
 
@@
 
-In my national-level analysis (IPEDS), I find that state funding cuts cause public universities to substitute away from tenured and tenure-track professors towards lecturers.
+In my university-level analysis (IPEDS), I find that state funding cuts cause public universities to substitute away from tenured and tenure-track professors towards lecturers.
 A funding cut of \$1,000 per student leads to an average university employing 6 more lecturers;
@@
 Yet the hiring rate for new professors at public universities was negatively impacted by state funding cuts.
-This implies that faculty substitution arose by limiting the hiring of new professors, which is supported by suggestive evidence on cumulative faculty hiring counts.
+This implies that faculty substitution arose by limiting the hiring of new professors, and is supported by suggestive evidence on professor hiring counts.
 The substitution is persistent for up to ten years, meaning that universities did not resume professor hiring in years following the funding cut, making the substitution towards lecturers long-run.
@@
 These papers measure changes in state funding and university revenues, but do not measure effects on faculty composition or salaries, or other outcomes related to university instruction.
+My results provide evidence that the long term trends in higher education funding and faculty outcomes are causally related, contributing to the literature on faculty outcomes \citep{ehrenberg2003studying} and trends in US higher education and funding \citep{hoxby2009changing,ehrenberg2012american}.
 
 Another contribution of my paper is to provide evidence on possible mechanisms connecting state funding cuts for higher education and worsening student outcomes.
-Substitution towards lecturers is likely a cost-cutting measure for funding constrained universities, though relying on these faculty --- who are often over-worked and not granted long-term employment protections --- may lead to worse student outcomes \citep{ehrenberg2005tenured,zhu2021limited,jaeger2011examining}.
-My results provide evidence that the long term trends in higher education funding are causally related, contributing to the literature on faculty outcomes \citep{ehrenberg2003studying} and trends in US higher education and funding \citep{hoxby2009changing,ehrenberg2012american}.
+Substitution towards lecturers is likely a cost-cutting measure for funding constrained universities, but universities that rely on more lecturers have persistently lower student retention graduation rates \citep{jaeger2011examining,ehrenberg2005tenured}.\footnote{
+    \cite{bettinger2010does} provide an alternative perspective, that adjunct lecturers improve subject retention for undergraduates, but do not consider graduation rate or any other student outcomes.
+    \cite{figlio2015tenure} have similar results at a highly-selective private university, a different sector to that studied in this paper.
+}
+According to survey evidence, around 50\% of lecturers are employed full-time, with a guaranteed annual salary and university benefits, and the remaining 50\% are employed part-time on adjunct contracts, without university benefits and paid by the class \citep{aau2014Characteristics}.
+Adjunct lecturers consistently lead to worse students outcomes, compared to part-time lecturers, largely thanks to the worse working conditions for adjunct lecturers \citep{zhu2021limited}.
+Combining the evidence that adjunct lecturers lead to worse student outcomes with the trends in education finances and performance \citep{NBERw23736,NBERw27885}, implies that substituting towards lecturers contributed to deteriorating student outcomes at public universities \citep{NBERw23736,NBERw27885}.
 
@@
 \autoref{sec:conceptual} gives the conceptual framework for how state funding may affect faculty.
-\autoref{sec:empirics} draws the empirical framework for isolating the causal effects of state funding on faculty composition and individual faculty, with \autoref{sec:results} presenting the empirical results.
+\autoref{sec:empirics} draws the empirical framework for isolating the causal effects of state funding on faculty, with \autoref{sec:results} presenting the empirical results.
 \autoref{sec:discussion} discusses the context and implications for the findings.

### 2024-01-03 text/sections/02-data.tex
writing revision (working on the final revision).

@@
 \subsection{Data Description}
-The data used in this paper come from two primary sources: \citet[IPEDS]{ipeds} regarding institutional information including finances and enrolment, and \citet[IBHED]{ibhed} for data on every professor in the Illinois public university system.
+The data used in this paper come from two primary sources: \citet[IPEDS]{ipeds} for data on university funding, finances, and enrolment, and \citet[IBHED]{ibhed} for data on every professor in the Illinois public university system.
 
@@
 }
-Data are consistent between the years 1990 and 2021,\footnote{
+Data are consistent for 1990--2021,\footnote{
     The years 1987-1989 are represented in these data in an incompletion fashion, so I focus on the years 1990 onwards.
@@
 }
-and provide information on the total revenues a university received from every source (including state governments), enrolment, plus faculty count and total expenditures on salaries.\footnote{
-    I combine the Urban Institute's 2018 compilation of IPEDS data for the years 1990--2017, and manually combine raw NCES data on 2018-2021 for all relevant variables.
+and provide information on university funding, enrolment, and numerous other characteristics.\footnote{
+    I combine the Urban Institute's 2018 compilation of IPEDS data for the years 1990--2017, and manually combine raw National Center for Education Statistics (NCES) data on 2018--2021 for all relevant variables.
     Figures for enrolment and faculty counts come from the raw NCES version of IPEDS for all years, addressing inconsistencies in the Urban Institute's data formulation for these variables.
 }
-I restrict analysis to public, four-year, degree-granting institutions, as these institutions adhere to a standardised concept of faculty profile, where tenure and title of appointment (lecturer, assistant professor, etc.) are relatively standardised.
+I restrict analysis to public, four-year, degree-granting institutions, as these institutions have the largest majority of faculty employed and students enrolled.
 For-profit institutions employ and enrol a negligible share of professors and students respectively, while students at two-year institutions by majority intend to eventually enrol at a four-year institution \citep{mountjoy2022}, so that these institutions are not considered.
+IPEDS reports the count of professors employed by position.\footnote{
+    IPEDS gives a mean salary measure for faculty by rank, but these figures have many missing values and disagree with calculated values from other reputable sources.
+    Over 40\% of university--year observations are missing in the IPEDS panel data-set, representing an average of 55\% undergraduate enrolment in each year.
+    Yearly averages of faculty salaries by the non-missing values do not agree with trends in average professor salary over the sample time period compared to summary statistics provided in \cite{aau2021survey}.
+    IPEDS values consistently report that public university professors are paid on average 20\% more than private sector faculty, which disagrees with other sources on the salary gap direction, and IPEDS figures are consistently higher than averages calculated from individual-level IBHED data.
+    For these reasons I do not further analyse IPEDS salary data, and use salary data from IBHED to analyse faculty salaries.
+    Notably, IPEDS data on faculty counts do not suffer from these problems.
+}
+This gives a resulting panel data-set, where each row represents a university-year, and includes columns for university funding and tuition revenue, plus total count\footnote{
+    Real salary is computed by scaling nominal salary to 2021 dollars by the CPI-U.
+} for each faculty position (lecturer, assistant professor, tenured professor, and total faculty), and other university characteristics.
+\autoref{tab:ipeds-summary} presents summary statistics for these variables in IPEDS data.
 
-IPEDS reports the count of professors employed by position, as well as total salary expenditures by position.
-This gives a resulting panel data-set, where each row represents a university-year, and includes columns for university finances, plus total count and average salary\footnote{
-    Real salary is computed by scaling nominal salary to 2021 dollars by the CPI-U.
-} for professors by position (lecturer, assistant, tenured, total).
-\autoref{tab:ipeds-summary} presents summary statistics for relevant variables in these data.
-
-IPEDS provides information at the university level, but only provides aggregated information within a university.\footnote{
-    IPEDS additionally gives a mean salary measure for faculty by rank (inconsistently) for each university.
-    Unfortunately, this measure of professors' salaries is particularly crude in measuring on professors' salaries and other individual-outcomes; summary statistics on IPEDS data do not agree with trends in average professor salary over the sample time period compared to summary statistics provided in \citealt{aau2021survey}.
-    Similarly, IPEDS reports of the average salary paid to professors at private institutions also disagree with other sources (including those calculated for Illinois in IBHED data later in this section) so are not considered in this analysis.
+IPEDS provides information at the university level, but only provides aggregated information within a university.
+To investigate outcomes for individual faculty members, I integrate individual-level data for all faculty members at all Illinois public universities 2010--2021.
+These data allow me to investigate how individual faculty members are affected by funding cuts to their university, investigating outcomes are more granular details than total faculty counts.
+IBHED freely provides this information, as the state of Illinois is required to publicly report base salary and benefits for all administrators, faculty members, and instructors employed by each public college or university.\footnote{
+    Public Act 96-0266, effective 1 January 2010, is the relevant Illinois law that requires publicly publishing salary data for all public university faculty salary and benefits \citep{illinois-public-act}.
+    A pdf copy of Public Act 096-0266 is included in the supplmentary data of this paper.
+    This law applies to all nine Illinois public universities: Chicago State University, Eastern Illinois University, Governors State University, Illinois State University, Northeastern Illinois University, Northern Illinois University, Southern Illinois University (all five campuses), University of Illinois (all four campuses), Western Illinois University.
 }
-To investigate the distribution, I integrate individual-level data for every public university professor in the state of Illinois between the years 2010-2021.
-IBHED hosts the information;
-Public Act 96-0266 (effective 1 January 2010) requires that each university report base salary and benefits all administrators, faculty members, and instructors employed by the college or university.\footnote{
-    The universities included are all campuses of the nine Illinois public universities: Chicago State University, Eastern Illinois University, Governors State University, Illinois State University, Northeastern Illinois University, Northern Illinois University, Southern Illinois University  (all five campuses), University of Illinois (all four campuses), Western Illinois University.
+These data provide the basis to build a panel of Illinois public university faculty across years 2010--2021; I define a faculty member as an individual by their first plus last name and university pairing,\footnote{
+    Some faculty are listed multiple times in these panel data.
+    For example, Professor Alberto Agustin Lopez-Scala of the University of Illinois Chicago has two appointments, one as an adjunct faculty and one as a department director.
+    For this observation I take the highest paid position (department director) as the primary appointment and drop the secondary appointment (adjunct faculty).
+    In analysis of faculty exit rate, I collapse the appointments into one and only consider faculty exits as having a faculty appointment versus no appointment.
 }
-These publicly available data provides the basis to build a panel of Illinois public university professors 2010-2021; I define a professor as an individual by their first plus last name and university pairing, and link this database to IPEDS regarding finances for their employing institution.
+and link this database to IPEDS for data on their university employer.
 The Illinois sample represents 16,932 professors in the year 2010 and 15,352 in the year 2021, with summary statistics presented in \autoref{tab:illinois-summary}.
-Additionally, further analysis using the first year of a professor's employment focuses on the subset of professors with observed year of hiring between the years 2011 and 2021, representing 1,778 professors in 2011, and 9,099 in 2021.
+Analysis of professors' in their first year on the job focuses on the subset of professors with observed year of hiring 2011--2021 (i.e., after the panel's first year), representing 1,778 professors in 2011, and 9,099 in 2021.
 
@@
 
-States vary largely in how much they fund their public university systems, thanks in part to the state-wide budgeting process.
+States vary largely in how much they fund their public university systems, thanks in part to the state-wide budgeting process \citep{NBERw23736}.
 Planning for an annual budget begins two years ahead of the fiscal year, and the legislature votes to approve or reject the governor office's budget request a number of months before the fiscal year begins.\footnote{
-    \cite{NBERw23736} present a full discussion of the decision-making process for state appropriations, drawing on administrative records originally analysed by \cite{parmley2009state}.
+    \cite{NBERw23736} present a full discussion of the decision-making process for state funding, drawing on administrative records originally analysed by \cite{parmley2009state}.
 }
-US state governments, by majority, are legally obligated to run a balanced-budget, so that yearly variation in tax revenues (caused by changing economic conditions or otherwise) necessarily affect state expenditures.
-Public universities have relatively lower lobbying power than other funding beneficiaries, so often bear the brunt of state government funding cuts \citep{delaney2011state}.\footnote{
+US state governments, by majority, are legally obligated to run a balanced-budget, so that yearly variation in tax revenues (e.g., caused by changing economic conditions) necessarily affect state expenditures.
+Public universities have lower lobbying power than other state institutions, so often bear the brunt of state government funding cuts \citep{delaney2011state}.\footnote{
     \cite{delaney2011state} fully describe the financial environment of state expenditures, and what makes spending on higher education an attractive area for state governments to expand funding during years of higher tax revenues, and retract funding in leaner years.
@@
 Additionally, the number of higher education students in each state varies thanks to the size of each birth cohort.
-For example, the birth cohort of 1971 was larger than for neighbouring years, leading to more student demand for limited public university places for students turning 18 in 1989 \citep{bound2007cohort}.
-This leads to variation in the amount of funding per student across years, showing that per student state funding varies on multiple dimensions.
+For example, the birth cohort of 1971 was larger than that of 1970 or 1972, leading to more student demand for limited public university places for students turning 18 in 1989 \citep{bound2007cohort}.
+This leads to yearly variation in the amount of funding per student across years, showing that per student state funding varies on multiple dimensions.
 
@@
 Illinois funding for higher education has stagnated, and has experienced serious declines in the decade 2010--2021 --- similar in magnitude to the nation-wide decline over 1990--2021.
-State funding among the seven Illinois public university campuses has fallen by over 50\% for the last decade, on both an absolute and per-student basis (see \autoref{fig:illinois-funding}).
-Similar to the figures for the rest of the country, there has been a corresponding substitution towards tuition revenue.
+State funding Illinois public universities fell by over 50\% over 2010--2021, in both absolute and per-student terms (see \autoref{fig:illinois-funding}).
+%Similar to the figures for the rest of the country, there has been a corresponding substitution towards tuition revenue.
 

### 2024-01-03 text/sections/03-concepts.tex
writing revision (working on the final revision).

@@
 Universities could respond to funding cuts by cutting faculty salaries, leading to more faculty leaving for jobs at other universities.
-Multiple universities passed a university-wide pay-cut for their faculty in response to state budget cuts around the 2008 recession, and Cornell University implemented a professor salary cut in 2020.\footnote{
+Multiple universities passed a university-wide pay-cut for their faculty in response to state budget cuts around the 2008 recession, and Cornell University implemented a professor salary cut in 2020 at the onset of the Covid-19 pandemic.\footnote{
     The Cornell University administration expected large fiscal squeezes in mid-2020, so imposed a faculty salary cut, and then returned the amount cut later in 2020 when the financial squeeze did not materialise.}
-Funding cuts could also limit the amount public universities can offer to new faculty hires, leading to fewer or less accomplished faculty accepting these offers.\footnote{
-    Faculty often consider their outside options in job decisions, considering offers for employment and/or promotion from multiple universities at the same time (see \citealt{blackaby2005} for an overview of faculty outside options).
+Funding cuts could also limit the salaries public universities can offer to new faculty hires, leading to fewer or less accomplished faculty accepting offers at public universities.\footnote{
+    Faculty often consider their outside options in job decisions, considering offers for employment and/or promotion from multiple universities at the same time --- see \cite{blackaby2005} for an overview of faculty outside options.
     If public universities can only make low salary offers thanks to funding cuts, then they are less likely to land professors with multiple offers.
@@
 This effect may not be the same across faculty seniority; most universities hire more junior faculty each year, so that yearly hiring disruptions may disrupt the looser market for established full professors more than tighter market for assistant professors or lecturers.
-While these factors make it unclear who will be affected most by funding cuts, there is one empirical fact worth noting: the average Illinois assistant professor is paid $\$76,897$, tenured professor $\$109,283$, while lecturers earn $\$31,449$ more than double that of a lecturer (annual salary for 2010--2021 in 2021 USD, see \autoref{tab:illinois-summary}).
-Secondly, lecturers are (by definition) granted shorter contracts than professors; assistant professors mostly work on 5--6 year long contracts, full professors on indefinite contracts (tenure), lecturers on yearly or termly contracts.
-If a public university's primary obligation is to teach, and they must fulfil this objective with fewer resources, substituting away from tenure-track and tenured professors towards lecturers is rationalisable.
+While these factors make it unclear who will be affected most by funding cuts, there is one empirical fact worth noting: lecturers have substantially lower annual salaries than professors.
+The average Illinois assistant professor's annual salary is $\$76,897$, tenured professors' is $\$109,283$, while lecturers' is $\$31,449$ (annual salary for 2010--2021 in 2021 USD, see \autoref{tab:illinois-summary}).
+Secondly, hiring a new professor is a long-term costly investment, when compared to hiring a new lecturer.
+Lecturers have shorter employment contracts than professors; assistant professors mostly work on 5--6 year long contracts, full professors on indefinite contracts (tenure), lecturers on yearly or termly contracts.
+If a public university's primary obligation is to teach, and they must fulfil this objective with less funding, then substituting away from tenure-track and tenured professors towards lecturers is rationalisable purely on this basis.
 
-It is not clear when the effects of funding cuts will be realised.
-University departments often employ faculty on multiple-year contracts (e.g., the tenure contract), and individual professors' decisions to retire, move university, or leave academia are multiple-year commitments, so that it may take multiple years for university funding to change the faculty composition, or trickle down to faculty.
-For example, university departments coordinate hiring in annual cycles, so that a budget short-fall in year $t$ may have no effect on the in-progress hiring committee and resulting hiring decisions, yet the hiring cycle in year $t+1$ may be postponed or cancelled.
-As such, this paper investigates the dynamics of state funding shocks' effects on faculty, for the years following a state funding cut.
+It is not clear when the effects of state funding cuts will be realised.
+Universities employ professors on multiple-year contracts (e.g., the tenure contract), and decisions to retire, move university, or leave academia are multiple-year commitments, so that it may take multiple years for effects to trickle down to faculty.
+On the other hand, university departments hire in annual cycles, so that a funding cut in one year may immediately impact faculty hiring in the same year, and thus faculty counts in the next academic year.
+Yet, it could be the case that faculty hiring resumes the very next year, or even hires more to compensate for last year's cancelled hiring cycle.
+This would mean that state funding cuts leads to only transitory changes in faculty counts.
+As such, this paper investigates the dynamics of state funding cuts and their effects on faculty, by employing local projections methods to measure the persistence of effects for multiple years following a state funding cut.

### 2024-01-03 text/sections/04-empirical.tex
Updates.
writing revision (working on the final revision).

@@
 This paper identifies the effects of changes in state funding on faculty.
-However, a university's state funding is not exogenous to state decisions for support of higher education, or exogenous to internal institutional decisions.
-Instead, the state government and university administration undertake a complex process of allotting resources across multiple different priorities, including instruction, research, or between departments.
-For example, it may be the case that a state government restricts funding for only its lowest performing universities in response to a budget shock, introducing a selection-on-outcomes threat to identification.
-Importantly, funding from a university's state government provides opportunity to address such endogeneity.
+However, a university's state funding is not exogenous to state decisions for support of higher education.
+Instead, the state government undertakes a complex process of budgeting funding across many priorities, while higher education and faculty outcomes may not be independent in this process.
+For example, it may be the case that a state government cuts funding for higher education as the demographics of the state change, and the average age grows older (such as in Maine) 
+At the same time, it may be harder to attract younger faculty to work at universities in a state with an older population, so that faculty outcomes are correlated with state funding thanks to changes in state demographics.
+This means that faculty outcomes would not be conditionally independent of faculty outcomes, and correlations between the two are not causal relationships.
+Shifts in state for higher education, however, provide opportunity to address such endogeneity.
 
-\subsection{State Funding Shocks}
+\subsection{State Funding Shift-Share}
 \label{sec:approp-shocks}
@@
 \cite{NBERw23736,NBERw27885} develop the instrument for public university finances by exploiting a shift-share instrument for changes in state-level funding interacted with university reliance on state funding in a base period.
-
 \begin{align}
@@
 
-The system exploits the fact that institutions who rely on state funding more will be affected by state funding shocks.
+The system exploits the fact that institutions who rely on state funding more will be affected by shift-share instruments.
 $Z_{i,t}$ is the instrument for state funding for institution $i$ in year $t$, interacting the average funding for universities in state $s(i)$ with reliance on state funding relative to total revenues, averaged across the base years 1990--1993.\footnote{
-    1990--1993 are defined as data for public university finance data are most comparable (i.e. without many missing values) beginning in 1990.
+    1990--1993 are defined as data for public university finance data are most comparable (i.e., without many missing values) beginning in 1990.
     \cite{NBERw23736} use the single year 1990 as the base year, though I use the four years to ameliorate missing values in the single year of 1990.
@@
 }
-$Z_{i,t}$ is constructed as negative to reflect the fact that the long term trend in, and most of the short-run shocks to, state funding for higher education have been negative.\footnote{
+$Z_{i,t}$ is negative to reflect the fact that the long term trend in, and most of the short-run shifts in, state funding for higher education have been negative.\footnote{
     When used in log terms, the instrument is the negative of the logged shock --- i.e., $- \log \left( -Z_{i,t} \right)$ as opposed to log of the negative shock $\log Z_{i,t}$ directly.
@@
     On the other hand, rises in tuition revenues (per student) may arise as result of tuition hikes thanks declining state support, which would mean controlling for tuition would constitute a bad control.
-    Nonetheless, estimates including tuition revenues (per student) as a control in the second stage of the IV estimates produces results of very similar magnitude and direction, and so are omitted.
+    Nonetheless, estimates including tuition revenues (per student) as a control in the second stage of the instrumental variables estimates produces results of very similar magnitude and direction, and so are omitted.
 }
-The instrument approach relies on the conditional independence assumption, that the instrument is independently assigned to the universities.
-While this assumption is fundamentally untestable, we see that universities are by-in-large smaller (in terms of enrolment, total revenues, and professor count) at the top of the state funding shock distribution (see~\autoref{tab:summary-quantiles}).
-Though on a per student basis, there is little difference across the distribution (except in the outcomes under consideration).
-The state funding shock instrument is positively associated with the total enrolment and total amount of state funding for each university, in both \$ and $\log$/percentage change terms, while other the other sources of university finances are not associated with the funding shock.
-So that the other sources of finances are not clear confounders for the instrumental variables strategy, as they exhibit balance with respect to the funding shock instrument \citep{pei2019poorly}.
-Yearly and individual fixed effects are included in regressions throughout to implicitly condition on mean differences between universities and years, leading to the assumption of conditional independence of the instrument with respect to state funding for each public university.
 
-The state funding shock is an instrument the level of state funding for each university in each year.  
+The shift-share instrument is an instrument for state funding for each university in each year.  
 The first-stage is then as follows, including institution and year fixed effects, where $X_{i,t}$ represents the amount of state funding divided by the number of full-time students attending the university.
@@
 \end{equation}
-We note the conditions for exogeneity in the instrument (following the discussion presented by \citealt{NBERw27885}).
-The instrument is exogenous if state policy decisions for funding of public universities are uncorrelated with unobserved institutional changes of any specific college or university in the state \citep{borusyak2022quasi}.
+The instrument must be conditionally independent of faculty outcomes for the instrumental variables results to be valid causal eestimates.
+The instrument is conditionally independent if state policy decisions for funding of public universities are uncorrelated with unobserved institutional changes of any specific college or university in the state.
 This assumption is plausible given that the majority of states have multiple (i.e., more than five) public universities, without any single university campus receiving the majority of state funding within any single state.
-Secondly, the shift-share identification strategy requires exogeneity in either the base-line share or shift component of the instrument.
-In this case, we satisfy the second: universities' institutional-level decisions are not correlated with contemporaneous or upcoming shocks to state funding.\footnote{
-    It would be plausible to consider the case that universities make institutional-level decisions in a consistently different manner to those 
-    with differing reliance on state funding in 1990, so that exogeneity by the base-line share is not plausible here.
+Secondly, the shift-share identification strategy requires exogeneity in either the base-line share or shift component of the instrument \citep{borusyak2022quasi}.
+We satisfy the second: faculty or universitiy institutional-level decisions are not correlated with contemporaneous (or upcoming shocks) to state funding, so that the shift-share approach yields causal estimates by instrumental variables \citep{NBERw27885}.\footnote{
+    It would be plausible to consider the case that universities make institutional-level decisions in a consistently different manner to those with differing reliance on state funding in 1990, so that exogeneity by the base-line share is not plausible here.
 }
-Lastly, the shift-share approach assumes that state funding shocks affect faculty outcomes only via affecting university finances.
+Additionally, the shift-share approach assumes that shift-share instruments affect faculty outcomes only via affecting university funding (exclusion restriction).
+
+The conditional independence assumption for the shift-share instrument makes sense in theory for state funding, but is fundamentally untestable with data.
+However, the assumption is bolstered by the fact that universities with higher values of the shift-share instrument differ little in observed characteristics compared to those with lower values \citep{pei2019poorly}.
+Universities are by-in-large smaller (in terms of enrolment, total revenues, and professor count) at the top of the shift-share instrument distribution (see~\autoref{tab:summary-quantiles}).
+Though on a per student basis, there is little difference across the distribution (except in the outcomes under consideration).
+The shift-share instrument is positively associated with the total enrolment and total amount of state funding for each university, in both \$ and $\log$/percentage change terms, while other the other sources of university finances are not associated with the funding shock.
+So that the other sources of finances are not clear confounders for the instrumental variables strategy, as they exhibit balance with respect to the funding shock instrument \citep{pei2019poorly}.
+Yearly and individual fixed effects are included in regressions throughout to implicitly condition on mean differences between universities and years.
 
@@
     It is important to note the treatment effect isolated here; the instrumental variables approach identifies the local average treatment effect, one weighted to level of exposure when treatment is continuous as is the case here.
-    So we interpret this treatment effect as a weighted average of effects on faculty at public universities to state funding changes, specific to state funding shocks, among the complier group --- i.e., among universities who respond to funding shocks and would not have made faculty-outcome changes absent the funding shock.
-    Also, we assume that no universities state funding increased in response to the negative state funding shock (monotonicity).
+    So we interpret this treatment effect as a weighted average of effects on faculty at public universities to state funding changes, specific to shift-share instruments, among the complier group --- i.e., among universities who respond to funding shocks and would not have made faculty-outcome changes absent the funding shock.
+    Also, we assume that no universities state funding increased in response to the negative shift-share instrument (monotonicity).
 }

### 2024-01-03 text/sections/06-discussion.tex
writing revision (working on the final revision).

@@
 }
-A public university's state funding per student fell by an average of 36\%, and the number of lecturers per student increased by 113\% (fell by 17\% for assistant and 23\% full professors).
-The estimates of elasticity for faculty count with respect to state funding imply that stagnating state funding explains about 40\% of the observed substitution towards lecturers and away from tenure-track and tenured faculty.
+The average public university's state funding fell by 36\% per student, whilethe number of lecturers per student increased by 113\%, and assistant and full professors  fell by 17\% and 23\%, respectively.
+Stagnating state funding explains about 40\% of the observed substitution towards lecturers away from tenure-track and tenured professors.
 
-These effects are persistent: public higher education in the US was not exposed to one large instance of funding cuts, but systematic funding short-falls across the last three decades.
-The long-term estimates (\autoref{fig:count-lp}) show that state funding cuts have effects on universities for multiple years after the initial funding cut, and these effects are compounded by the fact that public universities received persistent funding cuts for so many years.
-The end result is large changes in faculty composition, the long-run trend in stagnating state funding leading to a long-run substitution towards lecturers at public universities.
+These effects are persistent: public higher education in the US was not exposed to one large funding cut, but systematic funding short-falls since the 1990s.
+The results in this paper show that state funding cuts had persistent effects on universities and faculty for multiple years after the initial funding cut, which were compounded by many years of further funding cuts.
+These long-run changes resulted in large changes in the composition of faculty, with more lecturers and fewer professors at public universities.
 
@@
 These results imply that composition change arose by lower hiring of tenure-track faculty in the public university system.
-The private university system did not grow in any corresponding amount, with the clear implication that there are fewer tenure-track openings at US universities.\footnote{
-    While the number of PhD graduates has continued to rise \citep{aau2021survey}.
+The private university system did not grow in any corresponding amount, implying that there are fewer tenure-track openings at US universities.\footnote{
+    At the same time as these changes occured, the number of PhD graduates in the US has continued to rise \citep{aau2021survey,wapman2022quantifying}.
 }
 Securing academic employment on the tenure-track is more selective than ever, so that more PhD graduates will end up teaching as lecturers or leaving academia.
-% Note: unclear effects on PhD students choosing to not enter acadmeia, or go towards private unis
-% Would be possible to study with data on PhD students, and whether they are more/less likely to enter academia after budget cuts to their PhD institution.
 
-The results here are squarely in line with theories explaining the increased competition and inequality between US universities.
+The results here are in line with theories explaining the increased competition and inequality between US universities.
 \cite{urquiola2020markets} interprets the success of research in US higher education as a result of free market policies, where increased competition in a free market in the early 1900s led to improvements in university research among US private universities.
-These changes have not just affected research, they have affected education, too.
 The rate of selectivity among US universities has become increasingly polarised, where top (and mostly private) universities with the most resources have become more selective at the undergraduate level \citep{hoxby2009changing}, and increasingly dominate academia at the graduate level \citep{wapman2022quantifying}.
-Public universities are by-in-large not selective, so are caught in a relative decline when their selective and private peers are increasingly more selective, and better performing in terms of research and student outcomes.
+% These changes have not just affected research, they have affected teaching, too.
+Public universities are by-in-large not selective, so are caught in a relative decline when their selective and private competitors are more selective, and have research and student outcomes.
 
@@
 %At the same time returns to higher education have increased drastically \citep{berman1998implications}.
-Increasingly these students are being taught by lecturers, and not tenured faculty.
+Increasingly these students are being taught by lecturers, and not professors.
 Lecturers are often employed on short-term or part-time contracts (adjunct), with limited job stability.
 While we should worry about the effect on faculty working conditions, there is credible evidence that relying on adjunct lecturers leads to worse education and student outcomes, relative to full-time lecturers and professors \citep{zhu2021limited}.
-As tenure has increasingly become a private sector phenomenon, and the majority of US higher education students attend public institutions, we should worry about the long-term effects of stagnating funding for higher education in the US.
-
-%Additionally, put into further context of the trends in university competition \citep{urquiola2020markets}, and XYZ \citep{ehrenberg2012american}.
+As tenure has increasingly becomes a private sector phenomenon, and most US undergraduate attend public universities, we should worry about the long-term effects of stagnating funding for higher education.

### 2024-01-03 text/sections/09-figures-tables.tex
writing revision (working on the final revision).

--- 2023-12-19 text/sections/09-figures-tables.tex
@@
     \label{fig:funding}
+    \vspace{-1cm}
     \justify
@@
     \singlespacing
-    \caption{Total Student enrolment, by University Sector and Year.}
+    \caption{Total Student Enrolment, by University Sector and Year.}
     \begin{subfigure}[b]{0.495\textwidth}
@@
     \label{fig:enrolment}
+    \vspace{-1cm}
     \justify
@@
 
-\newpage
 \begin{figure}[H]
@@
     \label{fig:fte-perprof}
-
+    \vspace{-1cm}
     \justify
@@
     \label{fig:illinois-funding}
-
+    \vspace{-1cm}
     \justify
@@
     \label{fig:count-lp}
+    \vspace{-1cm}
     \justify
@@
     \textbf{Note}:
-    These figures show the local projections estimates of regression specification \eqref{eqn:secondstage}, with the funding shock shift-share as an instrument for state funding.
+    These figures show the local projections estimates of regression specification \eqref{eqn:secondstage}, with the state funding shift-share as an instrument for state funding.
     The unit of analysis is the university, and uses IPEDS data.
-    The $x$-axis represents the year relative to the funding shock (i.e., $x = 0$ represents the year of the state funding cut) and the $y$-axis represents the estimated effect in years relative to initial funding shock.
-    The coefficient estimate is effect of state funding ($X_{i,t}$) on faculty count per student ($Y_{i,t}$), using the funding shock instrument ($Z_{i,t}$), while accounting for auto-correlation between different time periods --- i.e., between $X_{i,t}, X_{i,t-1}$ and $Y_{i,t}, Y_{i,t-1}$.
+    The $x$-axis represents the year relative to the state funding shift-share (i.e., $x = 0$ represents the year of the state funding cut) and the $y$-axis represents the estimated effect in years relative to initial state funding shift-share.
+    The coefficient estimate is effect of state funding ($X_{i,t}$) on faculty count per student ($Y_{i,t}$), using the state funding shift-share instrument ($Z_{i,t}$), while accounting for auto-correlation between different time periods --- i.e., between $X_{i,t}, X_{i,t-1}$ and $Y_{i,t}, Y_{i,t-1}$.
     These results use a $\log-\log$ specification, so the estimates are for the elasticity of professor count per student in a year $t+k$ with respect to state funding in year $t$, where years $k = 0, \hdots, 10$ are on the $x$-axis.
@@
     \singlespacing
-    \caption{Local Projection Estimates for Effect of State Funding on Faculty Salaries per Student at Illinois Public Universities, by Professor Group.}
+    \caption{Local Projection Estimates for Effect of State Funding on Faculty Salaries Illinois Public Universities, by Professor Group.}
     \begin{subfigure}[b]{0.495\textwidth}
@@
     \label{fig:salaries-illinois-lp-rolling}
+    \vspace{-1cm}
     \justify
@@
     \textbf{Note}:
-    These figures show the local projections estimates of regression specification \eqref{eqn:secondstage}, with the funding shock shift-share as an instrument for state funding.
+    These figures show the local projections estimates of regression specification \eqref{eqn:secondstage}, with the state funding shift-share as an instrument for state funding.
     The unit of analysis is an individual faculty member (at an Illinois public university); funding data come from IPEDS, and faculty salaries from IBHED.
-    The $x$-axis represents the year relative to the funding shock (i.e., $x = 0$ represents the year of the state funding cut) and the $y$-axis represents the estimated effect in years relative to initial funding shock.
-    The coefficient estimate is effect of state funding ($X_{i(j),t}$) on faculty salaries ($Y_{j,t}$), using the funding shock instrument ($Z_{i(j),t}$), while accounting for auto-correlation between different time periods --- i.e., between $X_{i(j),t}, X_{i(j),t-1}$ and $Y_{i(j),t}, Y_{i(j),t-1}$.
+    The $x$-axis represents the year relative to the state funding shift-share (i.e., $x = 0$ represents the year of the state funding cut) and the $y$-axis represents the estimated effect in years relative to initial state funding shift-share.
+    The coefficient estimate is effect of state funding ($X_{i(j),t}$) on faculty salaries ($Y_{j,t}$), using the state funding shift-share instrument ($Z_{i(j),t}$), while accounting for auto-correlation between different time periods --- i.e., between $X_{i(j),t}, X_{i(j),t-1}$ and $Y_{i(j),t}, Y_{i(j),t-1}$.
     These results use a $\log-\log$ specification, so the estimates are for the elasticity of faculty salaries student in a year $t+k$ with respect to state funding in year $t$, where years $k = 0, \hdots, 4$ are on the $x$-axis.
@@
 
+\begin{figure}[H]
+    \centering
+    \singlespacing
+    \caption{State Funding and Total Professor Hiring at Public Universities, Total for 2011--2021.}
+    \includegraphics[width=0.8\textwidth]{figures/hiring-correlation.png}
+    \label{fig:hiring-correlation}
+    \justify
+    \footnotesize
+    \textbf{Note}:
+    This figure shows the correlation between state funding and total number of professors hired at each public university between 2011--2021.
+    Each dot is a US public university, with the $y$-axis the total number of professors (not lecturers) they hired 2011--2021, the $x$-axis the total state funding per student 2011--2021; the line is an OLS line of best-fit, with shaded region for 95\% confidence region of the correlation.
+    Data for the number of professors hired at public universities are provided by \cite{wapman2022quantifying,wapman2022zenodo}, and funding data from IPEDS.
+\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %% Section to host Tables
-\newpage
+
 \section{Tables}
@@
     \label{tab:illinois-summary}
-
     \justify
@@
 
-\newpage
 \begin{table}[H]
@@
     \centering
-    \caption{First Stage Estimates, for State Funding by Funding Shock in IPEDS Data.}
+    \caption{First Stage Estimates, for State Funding by state funding shift-share in IPEDS Data.}
     \textbf{Panel A: units in \$ per student}
@@
     \textbf{Note}:
-    These tables show the first stage OLS estimates of regression specification \eqref{eqn:firststage}, showing the effect of the funding shock on state funding to gauge performance as an instrument.
-    Each observation is a public univeristy-year, in the IPEDS data.
-    Panel A shows the effect of an funding shock of \$-1 per student in the state on the number of \$'s of state funding per student at the university --- i.e.,
-    \$-1 funding shock per student in the state leads to \$1.176 less state funding per student at the university according to preferred specification column 1.
-    Panel B shows the effect of a $-10$\% change funding shock per student in the state on $10$\% change in state funding per student at the university --- i.e.,
-    $-10$\% funding shock per student in the state leads to $-9.77$\% less state funding per student at the university according to prefferred specification column 1.        
-    Standard errors are clustered at the state-year level, and univeristy $+$ year fixed effects are included where noted.
+    These tables show the first stage OLS estimates of regression specification \eqref{eqn:firststage}, showing the effect of the state funding shift-share on state funding to gauge performance as an instrument.
+    Each observation is a public university-year, in the IPEDS data.
+    Panel A shows the effect of a state funding shift-share of \$-1 per student in the state on the number of \$'s of state funding per student at the university --- i.e.,
+    \$-1 state funding shift-share per student in the state leads to \$1.176 less state funding per student at the university according to preferred specification column 1.
+    Panel B shows the effect of a $-10$\% change state funding shift-share per student in the state on $10$\% change in state funding per student at the university --- i.e.,
+    $-10$\% state funding shift-share per student in the state leads to $-9.77$\% less state funding per student at the university according to prefferred specification column 1.        
+    Standard errors are clustered at the state-year level, and university $+$ year fixed effects are included where noted.
 \end{table}
 
-\newpage
 \begin{table}[H]
@@
     \textbf{Note}:
-    These tables show the second stage OLS and 2SLS estimates of regression specification \eqref{eqn:secondstage}, showing the effect of state funding changes on faculty outcomes, using the funding shock to instrument for state funding in the columns labelled 2SLS.
+    These tables show the second stage OLS and 2SLS estimates of regression specification \eqref{eqn:secondstage}, showing the effect of state funding changes on faculty outcomes, using the state funding shift-share to instrument for state funding in the columns labelled 2SLS.
     Each observation is a public university-year, in the IPEDS data.
@@
     \textbf{Note}: 
-    These tables show the second stage 2SLS estimates of regression specification \eqref{eqn:secondstage}, showing the effect of state funding changes on faculty outcomes at Illinois universities, using the funding shock to instrument for state funding.
+    These tables show the second stage 2SLS estimates of regression specification \eqref{eqn:secondstage}, showing the effect of state funding changes on faculty outcomes at Illinois universities, using the state funding shift-share to instrument for state funding.
     The shift-share instrument is based in the year the professor was hired, following definition in \autoref{sec:iv-model-indiv}.

### 2024-01-03 programs/run.sh

@@ -39,4 +39,4 @@ cd ..
 cd ../text
 latexmk -pdf paper.tex
 latexmk -c
-cp paper.pdf ../state-funding-faculty-2023.pdf
+cp paper.pdf ../state-funding-faculty-2024.pdf

### 2024-01-03 text/paper.tex
writing revision (working on the final revision).

@@
 \date{This version: \today \\ \vspace{0.1cm}
-    \href{https://github.com/shoganhennessy/state-funding-faculty/blob/main/state-funding-faculty-2023.pdf}{Newest version available here.}
+    \href{https://github.com/shoganhennessy/state-funding-faculty/blob/main/state-funding-faculty-2024.pdf}{Newest version available here.}
 }
@@
 \subsubsection*{Declaration of Competing Interest}
-This paper's author has no competing interests to declare.
-
+I have no competing interests to declare.
+% Separated Figures \& Tables
+\newpage
+\input{sections/09-figures-tables.tex}
 % Bibliography
@@
 \bibliography{sections/08-bibliography.bib}
-
-% Separated Figures \& Tables
-\newpage
-\input{sections/09-figures-tables.tex}
-
 % Appendix

### 2024-01-04 text/sections/05-results.tex
writing revision (working on the final revision).

@@
 
-\subsection{National Level, IPEDS}
-Changes in state funding have clear effects on the composition of faculty public universities employ.
-\autoref{tab:facultycount-shock-reg} presents OLS and IV estimates for the effect of a change in state funding on the number of faculty at the university, individually by position (and for the total faculty), in both count of professors and in percent terms for faculty per student ratio.
+\subsection{University Level, IPEDS}
+Changes in state funding have clear effects on the composition of faculty at US public universities.
+\autoref{tab:facultycount-shock-reg} presents OLS and IV estimates for the effect of a change in state funding on the number of faculty at the university, for each faculty position (lecturer, assistant and full professors, total).
+The results are presented in two different units: Panel A shows how a \$1 increase in state funding changed the number of faculty per students, while Panel B shows how a 1\% increase in state funding led to a percent change in the number of faculty per students.
 An extra \$1,000 in state funding per student leads to the average university employing 6 fewer lecturers, while effects for the other positions are not discernible from zero in absolute measure (\autoref{tab:facultycount-shock-reg}, Panel A).
-However, Panel A \autoref{tab:facultycount-shock-reg} measures professor count in percentage terms for professors per student.
 A 10\% increase in state funding lead to a 4.4\% decrease in the number of lecturers per student, an increase of 14\% for both assistant and full professors, leading to a 0.6\% increase in the total number of faculty per student.
+
 % In the state of Illinois, using the count of professors represented in the IBHED databse, state funding is correlated with the number of faculty, but lacks precision to identify the exact effects given the 144 university-year observations in the panel for 2010--2021.

### 2024-02-28 text/paper.tex
writing edits made.

@@
 \DeclareMathOperator{\eps}{\varepsilon}                                    % epsilon short hand shortcut
-\DeclareMathOperator{\st}{\text{ s.t. }}                                   % "such that" short hand shortcut
-\DeclareMathOperator{\then}{\text{ then }}                                 % "then" in equation shortcut
-\DeclareMathOperator{\ifeq}{\text{ if }}                                   % "if" in equation shortcut
-\DeclareMathOperator{\oreq}{\text{ or }}                                   % "or" in equation shortcut
-\DeclareMathOperator{\andeq}{\text{ and }}                                 % "and" in equation shortcut
-\DeclareMathOperator{\all}{,\; \forall}                                    % all (with spacing) in equation shortcut
-\DeclareMathOperator{\N}{\mathbb{N}}                                       % N, natural number, shortcut
-\DeclareMathOperator{\R}{\mathbb{R}}                                       % R, real number, shortcut
-\DeclareMathOperator{\Ll}{\mathcal{L}}                                     % L, Lagrangian, shortcut
 \renewcommand{\vec}[1]{\boldsymbol{\mathit{#1}}}                           % vector notation shortcut
 \newcommand{\mat}[1]{\boldsymbol{\mathit{#1}}}                             % matrix notation shortcut
-\DeclarePairedDelimiter\abs{\lvert}{\rvert}                                % absolute value notation shortcut
-\DeclarePairedDelimiter\norm{\lVert}{\rVert}                               % norm notation shortcut
-\newcommand{\Prob}[1]{\Pr\left( #1 \right)}                         % SHortcut for probability notation
-\newcommand{\Probgiven}[2]{\Pr\left( #1 \, \middle\vert \, #2 \right)} % SHortcut for probability notation, given
-\newcommand{\E}[2][]{\mathbb{E}_{#1} \left[ #2 \right]}                    % Expectation (with optional subscript) shortcut
-\newcommand{\Egiven}[3][]{\mathbb{E}_{#1} \left[ #2 \, \middle\vert \, #3 \right]} % Expectation given (with optional subscript) shortcut
-\newcommand{\Var}[2][]{\text{Var}_{#1} \left( #2 \right)}                  % Variation (with optional subscript) shortcut
-\newcommand{\Cov}[1]{\text{Cov} \left( #1 \right)}                         % Covariance (with optional subscript) shortcut
-\newcommand{\median}[1]{\text{median} \left( #1 \right)}                   % Median (with optional subscript) shortcut
-\newcommand{\indicator}[1]{\mathds{1}\left\{ #1 \right\}}                  % SHortcut for indicator function
-\newcommand{\diff}[2][]{\frac{d#1}{d#2}}                                   % SHortcut for differential fraction as a function
-\newcommand{\partialdiff}[2][]{\frac{\partial#1}{\partial#2}}              % SHortcut for partial differential fraction as a function
-\newcommand{\converge}[1]{\xrightarrow{ #1 \to\infty}}                     % SHortcut for convergence arrow
 \renewcommand{\hat}[1]{\widehat{#1}}                                       % Default estimator notation is widehat
@@
 \def\sectionautorefname~#1\null{Section~#1\null}                           % Fix autoref for sections
+\def\subsectionautorefname~#1\null{Subsection~#1\null}                     % Fix autoref for sections
 \def\equationautorefname~#1\null{Equation~(#1)\null}                       % Fix autoref for equations
@@
         Francine Blau,
+        Romain Campbell,
         Ryan Dycus,
@@
         and Evan Riehl.
-        I thank seminar participants at Cornell University (2022, 2023) for helpful discussion.
+        I thank seminar participants at Cornell University (2023, 2024), the Eastern Economic Association Annual Meetings (2024) for helpful discussion.
         This project's repository, with all materials and anonymised data needed for replication, is available at 
@@
 % Separated Figures \& Tables
-\newpage
 \input{sections/09-figures-tables.tex}

### 2024-02-28 text/sections/01-introduction.tex
writing edits made.

@@
 \label{sec:intro}
-
-States fund their public universities with roughly the same amount today as they did in the 1990s, despite their enrolment rising by over 50\% on average.
+States fund their public universities with roughly the same amount today as they did in the 1990s, despite enrolment rising by over 50\%.
 At the same time, public universities have systematically substituted away from professors towards lecturers, who teach often work on contingent and lower-salaried contracts.
-Previous research shows declining in state funding led to worsening student outcomes \citep{NBERw23736,NBERw27885}, but there is little evidence on whether this funding decline has impacted faculty.
+Previous research shows state funding cuts led to worsening student outcomes \citep{NBERw23736,NBERw27885}, but there is little evidence on whether these funding cuts impacted faculty.
 In this paper, I show that falling state funding caused a change in the composition of faculty at US public universities, substituting professors for more lecturers.
@@
 I use a shift-share instrumental variables approach to estimate how faculty are affected by these funding cuts.
-Universities employ $4.4$\% more lecturers per student following a $10$\% funding cut, and correspondingly rely less on tenure-track faculty by employing $1.4$\% fewer assistant professors and $1.2$\% fewer full professors per student.
+Universities employ $4.4$\% more lecturers per student following a $10$\% funding cut, and correspondingly rely less on tenure-track faculty by employing $1.4$\% fewer tenure-track professors and $1.2$\% fewer tenured professors per student.
 In a secondary analysis of all public university faculty in the state of Illinois, I show that incumbent faculty were unaffected, in terms of wages, promotion and quit rates.
-This implies that the faculty substitution arose by disrupting hiring of new tenure-track/tenured professors.
+This implies that the faculty substitution arose by disrupting hiring of new tenure-track/tenured professors, which is supported supplementary data on professor hiring at US public universities over 2010--2021.
 % Private universities were not exposed to similar financial constraints during this same time period, and do not exhibit the substitution away from tenure-track/tenured professors.
@@
 I use two different analyses to explore the causal effects of state funding on faculty composition and salaries.
-My first analysis uses data from the \cite{ipeds} to study the impact of changes in state funding for all US public universities 1990--2021.
-My second analysis uses panel data for every faculty member at Illinois public universities for 2010--2021 \citep{ibhed}, to measure how incumbent faculty (i.e., those already employed at the university) are affected by changes in state funding.
+My first analysis uses data from the \citet[IPEDS]{ipeds} to study the impact of changes in state funding for all US public universities 1990--2021.
+My second analysis uses panel data for every faculty member at Illinois public universities for 2010--2021 \citet[IBHED]{ibhed}, to measure how incumbent faculty (i.e., those already employed at the university) are affected by changes in state funding.
 These data allow me to answer the following research question: if a US public university receives an extra \$1,000 per student (or a 10\% increase), do they change how many faculty members they employ?
@@
 It is a priori unclear how faculty may be affected, as there are various margins that universities may to respond to funding cuts.
-For example, university administrations may directly cut salaries or lay-off incumbent staff.
-This happened in 2015 when the state of Illinois shut-down without setting a state budget, and the University of Illinois furloughed faculty unable to pay salaries \citep{furlough2010}.
+On one hand, faculty may be negatively affected by universities cutting salaries or laying-off incumbent faculty.
+This happened in 2010 when the University of Illinois System furloughed university staff following budget cuts after the Great Recession \citep{furlough2010}.
 Similarly, faculty hiring committees are often cancelled or disrupted in response to funding cuts, meaning there are fewer faculty entering a department to replace those leaving.
@@
 But it is not certain that universities would hire more lecturers in the face of budget cuts; they may not hire any more faculty at all.
-Similarly, the universities may be limited from hiring more lecturers by competitive pressures from peer institutions \citep{hoxby2009changing,urquiola2020markets}, and instead focus on filling the funding gap by courting private funding or raising tuition prices.
+It is possible that universities are be limited from hiring more lecturers by competitive pressures from peer institutions \citep{hoxby2009changing,urquiola2020markets}, leading university administrations to court private funding or raise tuition prices to fill funding gaps, leaving faculty unaffected.
 
+State funding for higher education is not decided randomly, so correlations between state funding and faculty outcomes do not necessarily represent causal relationships.
 State governments decide how to fund higher education by a complicated process, which could be influenced by lobbying, changing state priorities, or local economic conditions.
-As such, it would be na\"ive to think of state higher education funding as independent of public university faculty, so I employ a shift-share instrumental variable approach to identify causal effects of state funding cuts.
-The shift-share instrument identifies exogenous changes in state funding by exploiting differences in how much universities relied on that funding, interacted with state-wide changes in higher education funding --- following \cite{NBERw23736,NBERw27885}.
-Yearly state government budgets are decided without targetting individual universities for funding cuts, so more reliant universities will be more affected by state-wide changes in higher education funding, allowing the shift-share approach to estimate causal effects among more reliant unversities.
-Additionally, I estimate effect of state funding on faculty multiple years after the initial funding cuts, using the local projections method thanks to time-series confounding between the funding shock and later years' level of state funding.\footnote{
+As such, it would be na\"ive to think of state higher education funding as independent of public university faculty, so I employ a shift-share instrumental variables (IV) approach to identify causal effects of state funding cuts.
+The shift-share instrument identifies exogenous changes in state funding by exploiting differences in how much universities relied on state funding, interacted with state-wide shifts in higher education funding --- following the methods of \cite{NBERw23736,NBERw27885}.
+Yearly state government budgets are decided without targetting individual universities for funding cuts, so more reliant universities will be more affected by state-wide changes in higher education funding, allowing the shift-share approach to estimate causal effects among more reliant universities.
+Additionally, I estimate effect of state funding on faculty multiple years after the initial funding cuts, using the local projections method.\footnote{
     To the best of my knowledge, this is also the first paper to use the local projections method for a shift-share instrument, and to acknowledge the time-series confounding for shift-share IV models when the outcome is in future time periods.
 }
-This approach allows me to test whether universities recover in years following the funding cuts e.g., by resuming faculty hiring in later years.
+This approach allows me to test whether the effects of funding cuts are persistent or universities recover quickly e.g., by quickly resuming faculty hiring in years following funding cuts.
 
 In my university-level analysis (IPEDS), I find that state funding cuts cause public universities to substitute away from tenured and tenure-track professors towards lecturers.
-A funding cut of \$1,000 per student leads to an average university employing 6 more lecturers;
-in percentage terms, a funding cut of 10\% per student leads to a fall of 1.4\% in the number of assistant professors per student at a university, a fall of 1.2\% for tenured professors, and an increase in 4.4\% the number of lecturers per student.
-Local projection estimates show that these effects linger for many years after the initial funding shock, showing that the effect is not isolated to the year of the funding cut. 
+A funding cut of 10\% per student leads to a fall of 1.4\% in the number of assistant professors per student at a university, a fall of 1.2\% for tenured professors, and an increase in 4.4\% the number of lecturers per student.
+Local projection estimates show that these effects linger for up to ten years after the initial funding cut, showing that effects are persistent and not transitory.
 Over the same time period, state funding per student fell by around 35\%, professors per student fell by 9\%, and lecturers per student increased by 99\%.
@@
 In my analysis of all Illinois faculty (IBHED), I find that incumbent professors were not meaningfully affected, in terms of total salary, promotion rate, or rate of leaving the Illinois public university system.
-Yet the hiring rate for new professors at public universities was negatively impacted by state funding cuts.
 This implies that faculty substitution arose by limiting the hiring of new professors, and is supported by suggestive evidence on professor hiring counts.
-The substitution is persistent for up to ten years, meaning that universities did not resume professor hiring in years following the funding cut, making the substitution towards lecturers long-run.
 
@@
 For faculty outcomes, applied theory work models university decision-making and faculty hiring (see e.g., \citealt{abe2015implications,johnson2009jep,NBERc13879}).
-In addition, empirical papers measure how universities responded to endowment shocks \citep{brown2014endowment}, and trends in the academic job market following the 2008 recession \citep{turner2014impact}.
+Other papers measure university decision--making after endowment shocks \citep{brown2014endowment}, and trends in the academic job market following the 2008 recession \citep{turner2014impact}.
 These papers measure changes in state funding and university revenues, but do not measure effects on faculty composition or salaries, or other outcomes related to university instruction.
@@
 Another contribution of my paper is to provide evidence on possible mechanisms connecting state funding cuts for higher education and worsening student outcomes.
-Substitution towards lecturers is likely a cost-cutting measure for funding constrained universities, but universities that rely on more lecturers have persistently lower student retention graduation rates \citep{jaeger2011examining,ehrenberg2005tenured}.\footnote{
-    \cite{bettinger2010does} provide an alternative perspective, that adjunct lecturers improve subject retention for undergraduates, but do not consider graduation rate or any other student outcomes.
-    \cite{figlio2015tenure} have similar results at a highly-selective private university, a different sector to that studied in this paper.
+In 2021, around 65\% of lecturers at US public universities are employed part-time (IPEDS); this arrangement is often arranged as an adjunct contract, where faculty are paid based on how many courses they teach, without employment benefits \citep{aau2014Characteristics}.
+While funding cuts force universities to substitute towards lecturers, there is consistent evidence that students are worse off being taught by part-time lecturers than full-time faculty.
+Universities that rely on more lecturers have persistently lower student retention and graduation rates \citep{jaeger2011examining,ehrenberg2005tenured}
+Similarly, students at a large public university have consistently worse education outcomes if taught by adjunct lecturers than those taught by full-time faculty, largely thanks to the worse working conditions adjunct lecturers face \citep{zhu2021limited}.\footnote{
+    \cite{bettinger2010does,figlio2015tenure} study lecturers at a highly selective universities, finding that adjunct lecturers improve subject retention for undergraduates compared to professors.
+    This paper considers public universities, which are by-majority not selective, and where lecturers are more often employed part-time with less institutional support.
 }
-According to survey evidence, around 50\% of lecturers are employed full-time, with a guaranteed annual salary and university benefits, and the remaining 50\% are employed part-time on adjunct contracts, without university benefits and paid by the class \citep{aau2014Characteristics}.
-Adjunct lecturers consistently lead to worse students outcomes, compared to part-time lecturers, largely thanks to the worse working conditions for adjunct lecturers \citep{zhu2021limited}.
-Combining the evidence that adjunct lecturers lead to worse student outcomes with the trends in education finances and performance \citep{NBERw23736,NBERw27885}, implies that substituting towards lecturers contributed to deteriorating student outcomes at public universities \citep{NBERw23736,NBERw27885}.
+Combining the evidence that adjunct lecturers lead to worse student outcomes with the trends in education finances and performance \citep{NBERw23736,NBERw27885}, implies that substituting towards lecturers contributed to deteriorating student outcomes at public universities.
 
 This paper proceeds as follows.
-\autoref{sec:data} describes the data for university finances and faculty in Illinois, and trends in public university funding for the last three decades.
+\autoref{sec:data} describes the data for university funding, data on faculty in Illinois, and recent trends in higher education funding.
 \autoref{sec:conceptual} gives the conceptual framework for how state funding may affect faculty.

### 2024-02-28 text/sections/02-data.tex
writing edits made.

@@
 \subsection{Data Description}
-The data used in this paper come from two primary sources: \citet[IPEDS]{ipeds} for data on university funding, finances, and enrolment, and \citet[IBHED]{ibhed} for data on every professor in the Illinois public university system.
+The data used in this paper come from two primary sources: \citet[IPEDS]{ipeds} for data on university funding, finances, and enrolment, and \citet[IBHED]{ibhed} for data on every faculty member in the Illinois public university system.
 
@@
 
-\subsection{Trends in Finances, Enrolment, and Faculty Composition}
+\subsection{Trends in Funding, Enrolment, and Faculty Counts}
 \label{sec:trends}
@@
 }
-These features lead to yearly variation in state funding not seen in other revenues sources (such as federal funding), since US states differ in their support for higher education, and are subject to fiscal constraints.
 Additionally, the number of higher education students in each state varies thanks to the size of each birth cohort.
 For example, the birth cohort of 1971 was larger than that of 1970 or 1972, leading to more student demand for limited public university places for students turning 18 in 1989 \citep{bound2007cohort}.
-This leads to yearly variation in the amount of funding per student across years, showing that per student state funding varies on multiple dimensions.
+These features lead to yearly variation in state funding not seen in other revenues sources, such as federal funding.
 
-Public university revenues on average stagnated over 1990--2021.
+State funding for public universities stagnated over 1990--2021.
 \autoref{fig:funding} show the trends in revenues for the mean public university for the years 1990--2021.
@@
 At the same time, student enrolment at public universities rose precipitously.
-6.2 million students were enrolled in public universities in 1990, and this number rose by 47\% to 9.1 million, with most of the increase occurring over the years 2000-2021.
+6.2 million students were enrolled in public universities in 1990, and this number rose by 47\% to 9.1 million in 2021, with most of the increase occurring after the year 2000.
 \autoref{fig:enrolment} shows that total enrolment at private universities has also risen over the same time period, but not as drastic in either relative or absolute terms; the mean private university grew from 9,800 students in 1990 to 11,800 in 2021.
-This means that revenues per student have stagnated across all measures (seen in \autoref{fig:mean-funding-fte}), and particularly fallen from a mean of \$11,000 per student in 1990 to less than \$8,000 per student in 2021.
+This means that funding per student has stagnated for all sources (seen in \autoref{fig:mean-funding-fte}), falling from \$11,000 per student on average in 1990 to less than \$8,000 per student in 2021.
 
 There are large differences in the average number of professors per student between the private and public sector.
-\autoref{fig:fte-perprof} shows that private universities start with an average of 40 students for every one full professor, and little average trend until 2021.
-Public universities start with 40 students per professor, and by 2023 there are 52.4 students professors for every full professor --- with the largest rise coming in the 2008-2011 time period.
-The trend, and divergence between public and private universities, is similar for the number of assistant professors.
-Private and public universities have similar numbers of associated and full professors before the year 2000, yet this number has fallen by over 20\% in the next 20 years only for public universities: in 2021 the mean public university has 6 fewer full professors per hundred students than the mean private university.
-Over the same time period, we see the rise in use of non-tenure track instructor positions (collectively referred to as  lecturers), who were employed at similar rates in both sectors in 1990, yet have been utilised by public universities at a higher rate since.
+\autoref{fig:fte-perprof} shows that private universities start with an average of 40 students for every one full professor, and little change thereafter.
+Public universities start with 40 students per professor, and by 2021 there are 52.4 students professors for every full professor --- with the largest rise coming in the 2008--2011 time period.
+The general trend is similar for the number of assistant professors at both public and private universities.
+Private and public universities have similar numbers of associate and full professors before the year 2000, yet this number fell by over 20\% afterwards only for public universities: in 2021 the average public university has 6 fewer full professors per hundred students than the average private university.
+Over the same time period, we see the rise in use of lecturers; lecturers were were employed at similar rates in private and public sectors in 1990, but have public universities increased their reliance on lecturers more than private universities ever since.
 
@@
 \label{sec:trends-illinois}
-
 Illinois funding for higher education has stagnated, and has experienced serious declines in the decade 2010--2021 --- similar in magnitude to the nation-wide decline over 1990--2021.

### 2024-02-28 text/sections/03-concepts.tex
writing edits made.

@@
 
-Universities can respond multiple ways to the falls in per student state funding.
-\cite{NBERw23736} established that public universities who experienced state funding cuts did not fully off-set by raising tuition, and cutting total spending.
-But it is not clear how these spending cuts went on to affect faculty --- or whether they affected faculty at all.
+Universities can respond multiple ways to state funding cuts.
+\cite{NBERw23736} established that public universities experiencing state funding cuts did not fully off-set by raising tuition, and instead cut spending.
+It is not clear how these funding or spending cuts went on to affect faculty --- or whether they affected faculty at all.
 
-Wider changes in the American higher education sector may concurrently explain the trends in state funding and faculty substitution.
+Wider changes in the American higher education sector could concurrently explain the trends in state funding and faculty substitution.
 College selectivity changed drastically over the 21st century, where the top universities have become more selective while the average university less selective \citep{hoxby2009changing}.
 While the average public university is becoming less selective, it is possible that their most productive or research-focused faculty more often move to more selective and prestigious (and often private) institutions, leaving public universities substituting toward lecturers over time.
-This is one way in which the trends may be concurrent, but not causal.
-I address this issue by using shift-share instrument approach to isolate funding cuts that affected more funding-intensive universities, measuring the rate of substitution between lecturers and professors in response to yearly changes in state-wide funding.
+This is one way in which the trends may be concurrent, but not causally related.
+I address this issue by using shift-share IV approach to isolate funding cuts that affected more funding-intensive universities, measuring the rate of substitution between lecturers and professors in response to yearly changes in state-wide funding.
 
@@
 The average Illinois assistant professor's annual salary is $\$76,897$, tenured professors' is $\$109,283$, while lecturers' is $\$31,449$ (annual salary for 2010--2021 in 2021 USD, see \autoref{tab:illinois-summary}).
-Secondly, hiring a new professor is a long-term costly investment, when compared to hiring a new lecturer.
-Lecturers have shorter employment contracts than professors; assistant professors mostly work on 5--6 year long contracts, full professors on indefinite contracts (tenure), lecturers on yearly or termly contracts.
 If a public university's primary obligation is to teach, and they must fulfil this objective with less funding, then substituting away from tenure-track and tenured professors towards lecturers is rationalisable purely on this basis.
 
-It is not clear when the effects of state funding cuts will be realised.
+If faculty are affected by state funding cuts, then it is still not clear whether effects are persistent or merely transitory.
 Universities employ professors on multiple-year contracts (e.g., the tenure contract), and decisions to retire, move university, or leave academia are multiple-year commitments, so that it may take multiple years for effects to trickle down to faculty.
 On the other hand, university departments hire in annual cycles, so that a funding cut in one year may immediately impact faculty hiring in the same year, and thus faculty counts in the next academic year.
-Yet, it could be the case that faculty hiring resumes the very next year, or even hires more to compensate for last year's cancelled hiring cycle.
-This would mean that state funding cuts leads to only transitory changes in faculty counts.
+% Yet, it could be the case that faculty hiring resumes the very next year, or even hires more to compensate for last year's cancelled hiring cycle.
+%This means that state funding cuts could effect faculty in only transitory changes in faculty counts.
 As such, this paper investigates the dynamics of state funding cuts and their effects on faculty, by employing local projections methods to measure the persistence of effects for multiple years following a state funding cut.

### 2024-02-28 text/sections/04-empirical.tex
writing edits made.

@@
 
-The system exploits the fact that institutions who rely on state funding more will be affected by shift-share instruments.
-$Z_{i,t}$ is the instrument for state funding for institution $i$ in year $t$, interacting the average funding for universities in state $s(i)$ with reliance on state funding relative to total revenues, averaged across the base years 1990--1993.\footnote{
+The system exploits the fact that institutions who rely more on state funding will be affected by state-wide funding cuts.
+$Z_{i,t}$ is the shift-share instrument for institution $i$ in year $t$.
+It is contructed by interacting the average funding for universities in state $s(i)$ with reliance on state funding relative to total revenues, averaged across the base years 1990--1993.\footnote{
     1990--1993 are defined as data for public university finance data are most comparable (i.e., without many missing values) beginning in 1990.
@@
     On the other hand, rises in tuition revenues (per student) may arise as result of tuition hikes thanks declining state support, which would mean controlling for tuition would constitute a bad control.
-    Nonetheless, estimates including tuition revenues (per student) as a control in the second stage of the instrumental variables estimates produces results of very similar magnitude and direction, and so are omitted.
+    Nonetheless, estimates including tuition revenues (per student) as a control in the second stage of the IV estimates produce results of very similar magnitude and direction, and so are omitted.
 }
 
-The shift-share instrument is an instrument for state funding for each university in each year.  
 The first-stage is then as follows, including institution and year fixed effects, where $X_{i,t}$ represents the amount of state funding divided by the number of full-time students attending the university.
@@
 \end{equation}
-The instrument must be conditionally independent of faculty outcomes for the instrumental variables results to be valid causal eestimates.
+The instrument must be conditionally independent of faculty outcomes for the IV results to be valid causal estimates.
 The instrument is conditionally independent if state policy decisions for funding of public universities are uncorrelated with unobserved institutional changes of any specific college or university in the state.
@@
 Secondly, the shift-share identification strategy requires exogeneity in either the base-line share or shift component of the instrument \citep{borusyak2022quasi}.
-We satisfy the second: faculty or universitiy institutional-level decisions are not correlated with contemporaneous (or upcoming shocks) to state funding, so that the shift-share approach yields causal estimates by instrumental variables \citep{NBERw27885}.\footnote{
+We satisfy the second: faculty or university institutional-level decisions are not correlated with contemporaneous (or upcoming shocks) to state funding, so that the shift-share approach yields causal estimates by IV \citep{NBERw27885}.\footnote{
     It would be plausible to consider the case that universities make institutional-level decisions in a consistently different manner to those with differing reliance on state funding in 1990, so that exogeneity by the base-line share is not plausible here.
@@
 Additionally, the shift-share approach assumes that shift-share instruments affect faculty outcomes only via affecting university funding (exclusion restriction).
+This means that state-wide changes in higher education funding are assumed to affect faculty at universities only by changing how much their university is funded.
 
@@
 Though on a per student basis, there is little difference across the distribution (except in the outcomes under consideration).
-The shift-share instrument is positively associated with the total enrolment and total amount of state funding for each university, in both \$ and $\log$/percentage change terms, while other the other sources of university finances are not associated with the funding shock.
-So that the other sources of finances are not clear confounders for the instrumental variables strategy, as they exhibit balance with respect to the funding shock instrument \citep{pei2019poorly}.
+The shift-share instrument is positively associated with the total enrolment and total amount of state funding for each university, in both \$ and $\log$/percentage change terms, while other the other sources of university finances are not associated with the funding shift-share.
+So that the other sources of finances are not clear confounders for the IV strategy, as they exhibit balance with respect to the funding shift-share \citep{pei2019poorly}.
 Yearly and individual fixed effects are included in regressions throughout to implicitly condition on mean differences between universities and years.
 
-\subsection{Instrumental Variables Model}
+\subsection{Instrumental Variables (IV) Model}
 \label{sec:iv-model-uni}
 
-I use the instrument defined in \autoref{eqn:public-instrument} to overcome the endogeneity concerns for state funding to each public university, so the primary empirical model is an instrumental variables model.
+I use the instrument defined in \autoref{eqn:public-instrument} to overcome the endogeneity concerns for state funding to each public university, for an IV approach.
 \autoref{eqn:firststage} is the first stage for exogenous variation in the state funding for university $i$ in year $t$, and \autoref{eqn:secondstage} the second stage for the effect of state funding on faculty outcomes.\footnote{
-    It is important to note the treatment effect isolated here; the instrumental variables approach identifies the local average treatment effect, one weighted to level of exposure when treatment is continuous as is the case here.
-    So we interpret this treatment effect as a weighted average of effects on faculty at public universities to state funding changes, specific to shift-share instruments, among the complier group --- i.e., among universities who respond to funding shocks and would not have made faculty-outcome changes absent the funding shock.
+    The IV approach identifies the local average treatment effect, weighted towards larger values of the instrument in the continuous case.
+    So we interpret this treatment effect as a weighted average of causal effects on faculty at public universities to state funding changes, specific to shift-share instruments, among compliers --- i.e., among universities who respond to funding cut and would not have made faculty-outcome changes absent the funding cut.
     Also, we assume that no universities state funding increased in response to the negative shift-share instrument (monotonicity).
@@
 does it make sense to consider state funding in purely dollar amounts per student, or in percent change rate?
-The funding shock instrument is a strong predictor for the average university's level of state funding in either unit; a funding shock of \$1,000 per student in the entire state leads to around \$1,176 per student at the university, while a funding shock of $-10$\% leads to around 9.77\% less state funding per year.
+The funding shift-share is a strong predictor for the average university's level of state funding in either unit; a funding shift-share of \$1,000 per student in the entire state leads to around \$1,176 per student at the university, while a funding shift-share of $-10$\% leads to around 9.77\% less state funding per year.
 Yet, the level of state funding (and the outcome variables) vary greatly between states for the unit of analysis.
 For example, the average Illinois public university receives \$10,709 in state funding per student in 1990 and  \$6,713 in 2021 (a fall of over 30\%), while California went from \$19,224 per student to \$12,915 in the same time span (a fall of 37\%).
-While most states are not exactly the same to California and Illinois, this is example is telling for the phenomenon of stagnating state funding:
-states vary in their absolute funding for higher education in 1990, but most have experienced a decline of similar order to 30\%, so that the stagnation in state funding has been a percent change trend over this time period.
+While most states are not exactly the same as California or Illinois, this is example is telling for the phenomenon of stagnating state funding:
+states vary in their absolute funding for higher education in 1990, but most have experienced a decline of around 30\% in per student funding, so that the trend is a common phenomenon in percentage terms across the US.
 As such, I include regression specifications where the explanatory and outcome variables are $\log$ transformed, and refer to these when stating results in percent change terms.
 
-\subsection{Effects in Years After the Initial Funding Shock}
+\subsection{Effects in Years After the Initial Funding Cut}
 \label{sec:local-projections}
-The effects of a change in the universities funding may not be immediate, and faculty may be affected multiple years after a funding shock.
-Yet, the funding shock instrument is significantly auto-correlated year-on-year; a large state-level funding shock in year $t$ is also likely to experience a large funding shock in year $t-1$.
-Similarly, funding shocks to higher education in a year are highly correlated with state funding per student in the 5 years before and after the initial shock in IPEDS data (see \autoref{fig:lag-firststage}).
-Thus, a linear model correlated state funding in year $t$ with outcomes in year $t+k$ will suffer from time-series confounding.
+The effects of a change in the universities funding may not be immediate, and faculty may be affected multiple years after a funding cut.
+Yet, the funding shift-share instrument is significantly auto-correlated year-on-year; a large state-level funding cut in year $t$ is also likely to experience a large funding cut in year $t-1$.
+\autoref{fig:lag-firststage} shows that the shift-share instrument is correlated with state funding five years into the past and future.
+This level of time-series correlation would confound IV estimates of the effect of state funding in year $t$ on faculty outcomes in later years $t+k$, providing inconsistent estimates for the causal relationships.
 
 I employ a local projections approach to model how faculty outcomes in years $t+k$ are affected by state funding in year $t$, for future years $k > 0$ \citep{jorda2005}.
-The local projections method is an empirical model used to estimate dynamic treatment effects when the treatment is not binary, so that time-series confounding (i.e., treatment in time $t$ is correlated with treatment in time $t-1$) is present \citep{montiel2021local}.
-This method estimates the effect of state funding $X_{i,t}$ on faculty outcomes $Y_{i, t+ k}$ in follow-on years $k > 0$, while accounting for the auto-correlation between state funding in surrounding years.\footnote{
-    The funding shock has a persistent effect on state funding, multiple years after the initial funding shock, so that the instrument is similarly strong for the local projections method (see \autoref{fig:firststage-lp}).
+The local projections method is an empirical model used to estimate dynamic treatment effects when the treatment is not binary, when time-series confounding is present \citep{montiel2021local}.
+Faculty outcomes could be affected in only a transitory manner, where effects in later years are close to zero, or in a persistent manner, where the effects of state funding cuts linger for multiple years.
+This method allows me to estimate the effect of state funding $X_{i,t}$ on faculty outcomes $Y_{i, t+ k}$ in follow-on years $k > 0$, credibly testing whether effects are transitory or long-lasting.
+The approach accommodates the shift-share instrument for state funding \citep{olea2021inference}, so that I present results using the shift-share to instrument for state funding (as in the rest of the empirical analysis).\footnote{
+    The funding shift-share has a persistent effect on state funding, multiple years after the initial funding shift-share, so that the instrument is similarly strong for the local projections method --- see \autoref{sec:appendix-ipeds-firststage} further evidence on the strength of the first stage in local projections estimation.
 }
-Additionally, the approach accommodates the shift-share instrument for state funding \citep{olea2021inference}, so that I present instrumental variables estimates for this method as in the rest of the empirical analysis.

### 2024-02-28 text/sections/05-results.tex
writing edits made.

@@
 \autoref{tab:firststage-reg} presents results of the first-stage regression among IPEDS data, separately with and without a control for tuition revenue per student, plus institution and year fixed effects.
-I use the same empirical strategy in the analysis of Illinois faculty, and \autoref{tab:firststage-illinois} presents first-stage estimates among the IBHED data ; results are similar to those calculated in IPEDS data.
-\autoref{tab:firststage-reg} Panel A shows estimates of a state-wide funding shock of \$1 per student on state funding per student at the university, and Panel B the effects of a 1\% increase in the state-wide funding shock on state funding per student.
-Column (1) shows that a state-wide funding shock of \$10 per student is associated with \$11.76 less state funding per student at the university; -10\% state-wide funding shock leads to -9.77\% state funding per student at the average university.
-Estimates with and without including fixed effects.
-Column (2) shows estimates of the first-stage without including fixed effects, and gives less precise estimates for the funding shock --- this may be due to systematic differences in universities, which are not accounted for without fixed effects.
+\autoref{tab:firststage-reg} Panel A shows estimates of a state funding shift-share of \$1 per student on state funding per student at the university, and Panel B the effects of a 1\% increase in the state funding shift-share on state funding per student.
+Column (1) shows that a state funding shift-share of \$10 per student is associated with \$11.76 less state funding per student at the university; -10\% state funding shift-share leads to -9.77\% state funding per student at the average university.
+Estimates are presented for specifications with and without fixed effects.
+Column (2) shows estimates of the first-stage without including fixed effects, and gives less precise estimates for the funding shift-share --- this may be due to systematic differences in universities, which are not accounted for without fixed effects.
 Columns (3) and (4) include the tuition revenue as a control (explained in \autoref{foot:control}) to exhibit estimates with or without this control variable.
 Column (3) shows similar estimates to column (1) in both Panels A and B thanks to inclusion of fixed effects, so that 
-whether tuition revenue is a bad control or not does not affect empirical estimates.
+whether tuition revenue is a bad control does not affect empirical estimates.
 Column (1) represents the estimates for \autoref{eqn:firststage} with fixed effects (not including tuition revenue as a control variable), and is the preferred form that I proceed with.
+I use the same empirical strategy in the analysis of Illinois faculty, and \autoref{tab:firststage-illinois} presents first-stage estimates among the IBHED data; results are similar to those calculated in IPEDS data.
+These results, show the funding shift-share instrument strongly predicts university revenues in the first-stage estimation.
 
 The first stage is strong in both IPEDS and IBHED data, and the argument in \autoref{sec:empirics} for exogeneity comes from independence of the state funding mechanism to any individual public university campus.
-\autoref{tab:firststage-balance} provides a balance test, showing that the shift-share instrument is correlated with the professor outcomes, but not with other sources of university funding.
-These results, together with the case for exogeneity, show the funding shock instrument
-strongly predicts university revenues in the first-stage estimation.
+The IV approach also assumes that shifts in state-wide funding for higher education (the instrument) affect outcomes only via affecting state funding for the university (exclusion restriction).
+This assumption is fundamentally untestable, but is bolstered by the lack of relationship with university characteristics.
+\autoref{tab:summary-quantiles} shows the mean of various university variables, across the instrument distribution.
+The university--years in the lowest 20\% of the instrument distribution have more research spending than those in the next 20\%; this is to be expected, as university spending is heavily related to state--wide changes in funding.
+On the other hand, the measures of selectivity (acceptance rate, 6--year graduation rate) are not substantially different across the distribution of the instrument.
+This allays concerns that state-wide shifts in state funding are targetted more towards less selective universities (a possible confounder), and supports use of the shift-share IV empirical design.
 
 \subsection{University Level, IPEDS}
-Changes in state funding have clear effects on the composition of faculty at US public universities.
+Funding cuts lead to measurable changes in the number of faculty at US public universities.
 \autoref{tab:facultycount-shock-reg} presents OLS and IV estimates for the effect of a change in state funding on the number of faculty at the university, for each faculty position (lecturer, assistant and full professors, total).
 The results are presented in two different units: Panel A shows how a \$1 increase in state funding changed the number of faculty per students, while Panel B shows how a 1\% increase in state funding led to a percent change in the number of faculty per students.
-An extra \$1,000 in state funding per student leads to the average university employing 6 fewer lecturers, while effects for the other positions are not discernible from zero in absolute measure (\autoref{tab:facultycount-shock-reg}, Panel A).
-A 10\% increase in state funding lead to a 4.4\% decrease in the number of lecturers per student, an increase of 14\% for both assistant and full professors, leading to a 0.6\% increase in the total number of faculty per student.
 
-% In the state of Illinois, using the count of professors represented in the IBHED databse, state funding is correlated with the number of faculty, but lacks precision to identify the exact effects given the 144 university-year observations in the panel for 2010--2021.
+An extra \$1,000 in state funding per student leads to the average university employing 6 fewer lecturers, while effects for the other positions are not discernible from zero (\autoref{tab:facultycount-shock-reg}, Panel A).
+A 10\% state funding cut lead to a 4.4\% increase in the number of lecturers per student, and a decrease of 1.35\% and 1.37\% for assistant and full professors, respectively.
+A state funding cut of 10\% leads to a 0.65\% decrease in the total number of faculty per student (\autoref{tab:facultycount-shock-reg}, Panel B).
+These are large substitution effects, and explain roughly 40\% of the trend in subsitutional towards lecturers over this same time period.
+The average public university has roughly 36\% less state funding per student in 2021 than they did in 1990, while employing over double the number of lecturers and 20\% fewer professors (relative to enrolment, \autoref{fig:mean-funding-fte}, \ref{fig:fte-perprof}).
 
-Faculty composition is not only impacted in the same year as a state funding shock, but also for a number years following.
-\autoref{fig:count-lp} shows local projection estimates, as a series of impulse responses, where the estimate is the elasticity for professor count per student in year $t+k$ with respect to state funding in year $t$ for follow-on years $k = 0, \hdots, 10$.
-A 10\% decrease in state funding increases count of lecturers per student by 4\% for the first two years after the initial shock.
-While count of assistant professors decreases 10 to 18.5\% three years later;
-count of tenured professors decreases 9 to 6\% three years later.
-Together, the total count of professors per student decreases around 1\% for three years after the initial funding shock, and the effect is not distinguishable from zero after this point.
+Once the funding cut has passed, universities do not re-adjust to hiring more professors to make up for last year's substitution towards lecturers.
+\autoref{fig:count-lp} shows local projection estimates, where the estimate is the elasticity for professor count per student in year $t+k$ with respect to state funding in year $t$ for later years $k = 0, \hdots, 10$.
+These estimates correlespond to the elasticity estimates in \autoref{tab:facultycount-shock-reg} Panel B, where the outcome is faculty count per students in later years.
+State funding cuts affectthe number of lecturers and professors at universities up to 9 years after the initial funding cut.
+A 10\% state funding cut leads to 4.2\% (2.9--5.5 in a 95 percent confident interval) more lecturers per student the following year, and around 3.1\% (0.6--5.5 ``'') more lecturers nine years later.
+A 10\% state funding cut leads to 1.5\% (0.6--2.2 in a 95 percent confident interval) fewer assistant professors per student the following year, and around 1.1\% (0.2--2.0 ``'') fewer nine years later;
+the same cut leads to 1.2\% (0.6--1.8 in a 95 percent confident interval) fewer full professors per student the following year, and around 0.9\% (0.1--1.7 ``'') fewer nine years later.
+The effects are a substitution with the composition of faculty, as the total number of faculty per students are relatively unaffected in the follow-on years.
 
 \subsection{Illinois Faculty, IBHED}
-This section refers to results using data on every professor at a public university in the state of Illinois.
-Incumbent professors are relatively unaffected by changes in the state funding for their university in the years following their hire.
-\autoref{tab:faculty-shock-illinois-rolling} show estimates of a 1\% change in state funding per student on faculty salaries, rate of exiting the Illinois public university system, and rate of promotion.
-The estimates are not discernible from zero, so that faculty who are already at the university (incumbent faculty) seem to have no effects of state funding cuts passed on to them.
-Additionally, local projection estimates show that professors' salaries are also not affected on average in yeard following a funding shock  (\autoref{fig:salaries-illinois-lp-rolling}).
-Lecturers on the other hand, see a 0.75\% increase in salary in the second and third year after a 10\% increase in state funding (\autoref{fig:salaries-lecturer-illinois-lp-rolling}), reflecting how lecturers are again more affected by changes in public university funding.
-Similarly, promotion rate and exit rate among incumbent faculty  (from the Illinois public university system) are unaffected in the years after an initial funding shock (\autoref{fig:promoted-illinois-lp-rolling}, \ref{fig:exit-illinois-lp-rolling}).
+Incumbent professors at Illinois universities are relatively unaffected by state funding cuts for their university.
+\autoref{tab:faculty-shock-illinois-rolling} show estimates of a 10\% change in state funding per student on faculty salaries, rate of exiting the Illinois public university system, and rate of promotion.
+The estimates are not discernible from zero, so that faculty who are already at the university (incumbent faculty) seem to have little or no state funding cuts passed on to them for these outcomes.
+Additionally, local projection estimates show that professors' salaries are also not affected in the years after a state funding cut (\autoref{fig:salaries-illinois-lp-rolling}).
+Similarly, promotion rate and exit rate among incumbent faculty (from the Illinois public university system) are unaffected in the years after an initial funding shift-share (\autoref{fig:promoted-illinois-lp-rolling}, \ref{fig:exit-illinois-lp-rolling}).
+There is evidence that lecturers see a 0.75\% increase in salary in the second and third year after a 10\% increase in state funding (\autoref{fig:salaries-lecturer-illinois-lp-rolling}), though this is the only evidence that incumbent faculty are affected by state funding cuts in the Illinois public university system.
 
@@
 In the same vein, when funding increased, count of tenure-track and tenured professors per student increased --- and the opposite in years of funding cuts.
-These findings are in line with the observation that universities froze hiring for a couple of years in response to 2008 negative budget shocks, particularly for tenure-track positions \citep{turner2014impact}, and similarly from shocks to university endowments \citep{brown2014endowment}.
-For lecturers, we see a negative effect which lines up with two trends noted in \autoref{sec:trends}: funding for public universities (per student) decreased while relative usage of lecturers increased substantially.
+These findings are in line with the observation that universities froze hiring for a couple of years after the 2008 recession and funding cuts \citep{turner2014impact}.
+For lecturers, we see a negative effect which lines up with two trends noted in \autoref{sec:trends}: state funding for public universities (per student) stagnated while usage of lecturers increased substantially.
+
+%\subsection{Robustness checks}
+%\textbf{Describe robustness checks in one or two paragraphs here; see how Lovenheim$+$ did this section.}
 
 \subsection{Reduced Hiring of New Professors}
-Public universities substituted professors for more lecturers thanks to stagnating state funding, but there are little discernible effects on incumbent professors or lecturers.
-This implies that the substitution arose by impacting the hiring of new professors to replace those retiring or leaving academia.
-% \textbf{Expand this paragraph, referencing the figure, and the OLS tables for the appendix}.
-I take recent data on the total faculty hired at US public universities over 2011--2021 (provided by \citealt{wapman2022quantifying,wapman2022zenodo}), to show that universities with lower state funding (per student) also hired fewer professors in the same time period --- shown in \autoref{fig:hiring-correlation}.
-The shift-share IV model shows that a $10$\% per-student state funding cut leads to $13$\% fewer faculty hires per student across the decade 2011--2021.\footnote{
-    This limited sample of data are a cross-section for count of professor hires 2011--2021, so that local projection estimates are not possible here.
-    See \autoref{sec:appendix-hiring} for further details.
+The results above present a condundrum: there are large substitution effects among faculty, but faculty themselves are unaffected by funding cuts.
+The analysis of Illinois faculty (IBHED) studies the effect of state funding cuts on faculty already at the university --- i.e., incumbent faculty.
+The findings above imply that subsitution towards lecturers came about by limiting hiring of new professors, to replace leaving or retiring professors, so that they are replaced by hiring lecturers.
+
+To further study this issue, I take recent data on the total faculty hired at US public universities over 2011--2021 provided by \cite{wapman2022quantifying,wapman2022zenodo}.
+\cite{wapman2022quantifying} study the composition of professor hiring in the years 2011--2021, and publiccally released data on the number of professors (not lecturers) hired at US universities.\footnote{
+    These data are not panel data, instead only give the total professors hired over the ten year period 2011--2021.
 }
+I combine this measure, with total state funding for public universities acorss the same time period from IPEDS, to validate the claim that state funding cuts led to faculty subtitution via limiting hiring of new professors.
+
+There is a strong correlation between the number of professors hired (relative to student enrolment), where an extra \$1,000 of state funding per student is associated with the university hiring 0.12 more professors per hundred students.
+I apply the shift-share model to these data, summing the state funding and funding shifts for the decade of 2012--2021.
+I find that a state funding cut of 10\% leads to 13\% fewer professor hires (relative to enrolment).
+See \autoref{sec:appendix-hiring} for further details on this additional analysis of professor hiring.
 
@@
 This provides a coherent mechanism for how incumbent faculty (those who are already hired at the university) were not meaningfully impacted by funding cuts, yet their wider composition nonetheless changed when their universities experienced funding cuts.
-
-%\subsection{Robustness checks}
-% Describe robustness checks here.

### 2024-02-28 text/sections/06-discussion.tex
writing edits made.

@@
 }
-The average public university's state funding fell by 36\% per student, whilethe number of lecturers per student increased by 113\%, and assistant and full professors  fell by 17\% and 23\%, respectively.
+The average public university's state funding fell by 36\% per student, while the number of lecturers per student increased by 113\%, and assistant and full professors  fell by 17\% and 23\%, respectively.
 Stagnating state funding explains about 40\% of the observed substitution towards lecturers away from tenure-track and tenured professors.
@@
 The private university system did not grow in any corresponding amount, implying that there are fewer tenure-track openings at US universities.\footnote{
-    At the same time as these changes occured, the number of PhD graduates in the US has continued to rise \citep{aau2021survey,wapman2022quantifying}.
+    At the same time as these changes occurred, the number of PhD graduates in the US has continued to rise \citep{aau2021survey,wapman2022quantifying}.
 }

### 2024-02-28 text/sections/09-figures-tables.tex
writing edits made.

@@
         \centering
-        \caption{Lecturer.}
+        \caption{Lecturers.}
         \includegraphics[width=\textwidth]{figures/lecturer-fte-perprof.png}
@@
         \centering
-        \caption{Assistant.}
+        \caption{Assistant Professors.}
         \includegraphics[width=\textwidth]{figures/assistant-fte-perprof.png}
@@
         \centering
-        \caption{Full.}
+        \caption{Full Professors.}
         \includegraphics[width=\textwidth]{figures/full-fte-perprof.png}
@@
         \centering
-        \caption{All.}
+        \caption{All Faculty.}
         \includegraphics[width=\textwidth]{figures/all-fte-perprof.png}
@@
     This figure shows the average number of students per faculty member, by different faculty position, at US public universities.
-    E.g., panel A calculates the mean of (lecturer count) / (student enrolment) at US public universities, for each year of 1990--2017, to show the average trend in faculty composition compared to student enrolment.
+    E.g., panel A calculates the mean of (student enrolment) / (lecturer count) at US public universities, for each year of 1990--2017, to show the average trend in faculty composition compared to student enrolment.
     These figures are calculated with IPEDS data.
@@
     \centering
-    \caption{IBHED Summary Statistics, Professor Panel 2010--2021.}
+    \caption{IBHED Summary Statistics, Faculty Panel 2010--2021.}
     \makebox[\textwidth][c]{\input{tables/illinois-summary.tex}}
@@
     \centering
-    \caption{First Stage Estimates, for State Funding by state funding shift-share in IPEDS Data.}
+    \caption{First Stage Estimates, Predicting State Funding by Shift-Share shift-share in IPEDS Data.}
     \textbf{Panel A: units in \$ per student}
@@
     \centering
-    \caption{Effects of Changes in State Funding on University Faculty Composition, IPEDS 1990--2017, OLS and 2SLS Estimates.}
+    \caption{Effects of Changes in State Funding on University Faculty Composition, IPEDS 1990--2021, OLS and 2SLS Estimates.}
 
@@
     Each observation is a public university-year, in the IPEDS data.
-    Panel A shows the effect of a fall in state funding \$$-1,000$ per student in the state on the number of professors --- i.e.,
+    Panel A shows the effect of a fall in state funding \$$-1,000$ per student on the number of professors --- i.e.,
     an extra \$$1,000$ per student leads to 6 fewer lecturers according to column 2.

### 2024-02-28 text/sections/10-appendix.tex
Updates.
writing edits made.

--- 2023-10-27 text/sections/10-appendix.tex
@@
 \subsection{IPEDS First Stage}
+\label{sec:appendix-ipeds-firststage}
 
@@
     %\centering
-    \caption{Mean Characteristics for Public Universities, by State Funding Shock Instrument.}
+    \caption{Shift-Share Instrument Balance, Mean Characteristics Across Instrument Distribution.}
     \makebox[\textwidth][c]{\input{tables/summary-quantiles.tex}}
@@
     \textbf{Note}:
-    This table shows the summary statistics for every public university--year observation in IPEDS data, for each of the 5 quantiles of the funding shock instrument.
-    The column labelled ``1st'' refers to the mean for all university-year observations in the first quintile (bottom 20\%) of the funding shock distribution, and so on.
+    This table shows the summary statistics for every public university--year observation in IPEDS data, for each of the 5 quantiles of the funding shift-share instrument.
+    Funding shift-share is the instrument defined in \autoref{sec:empirics} the product of (1) state-wide funding shift and (2) share reliance on state funding.
+    State-wide funding shift is total funding for higher education in that university's state (divided by the count of state students); share reliance on state funding is the the total amount of state funding received by the university in 1990--1993 divided by total revenues for those years.
+    The column labelled ``1st'' refers to the mean for all university-year observations in the first quintile (bottom 20\%) of the funding shift-share distribution, and so on.
     The numbers are adjusted to 2021 figures by CPI-U.
     Non-institutional revenues refers to the sum of federal, state, and local funding plus tuition revenues; these sum to the majority of university funding, but exclude numbers such as university income from capital projects.
+    Acceptance rate and 6 year graduation rate are for university undergraduates, and are calculated from IPEDS data available for academic years 1997 through 2018.
 \end{table}
@@
     \singlespacing
-    \caption{Correlation Between State Funding Shock and Public University State Funding in Surrounding Years.}
+    \caption{Correlation Between State Funding Shift-Share and Public University State Funding in Surrounding Years.}
     \includegraphics[width=\textwidth]{figures/lag-firststage.png}
@@
     \textbf{Note}:
-    This figure shows the correlation between state funding in year $t+k$ with the funding shock in year $t$ for a university, where $k = -7, \hdots, 5$ are the years on the $x$-axis.
-    This shows that state funding and the funding shock have dynamic correlation, so that dynamic effects must be estimated by local projections --- and not simple OLS or 2SLS.
+    This figure shows the correlation between state funding in year $t+k$ with the funding shift-share in year $t$ for a university, where $k = -7, \hdots, 5$ are the years on the $x$-axis.
+    This shows that state funding and the funding shift-share are correlated across years, so that dynamic effects must be estimated by local projections --- and not simple OLS or 2SLS.
     The estimates are of \eqref{eqn:firststage}, calculated with IPEDS data, separately for each year relative to initial shock, using the $\log-\log$ specification, including fixed effects for university $+$ year, and clustering standard errors by university $+$ year.
@@
     \singlespacing
-    \caption{Local Projection Estimates for Funding Shock on State Funding, in IPEDS Data.}
+    \caption{Local Projection Estimates for Funding Shift-Share on State Funding, in IPEDS Data.}
     \includegraphics[width=0.6\textwidth]{figures/firststage-lp.png}
@@
     \textbf{Note}:
-    These figures show the local projections estimates of regression specification \eqref{eqn:firststage}, with the funding shock as an instrument for state funding, using IPEDS data.
-    The coefficient estimate is effect of funding shock ($Z_{i,t}$) on state funding ($X_{i,t}$), while accounting for auto-correlation between different time periods --- i.e., between $Z_{i,t}, Z_{i,t-1}$ and $X_{i,t}, X_{i,t-1}$.
-    These results use a $\log-\log$ specification, so the estimates are for the elasticity of state funding per student in a year $t+k$ with respect to funding shock in year $t$, where years $k = 0, \hdots, 10$ are on the $x$-axis.
+    These figures show the local projections estimates of regression specification \eqref{eqn:firststage}, with the funding shift-share as an instrument for state funding, using IPEDS data.
+    The coefficient estimate is effect of funding shift-share ($Z_{i,t}$) on state funding ($X_{i,t}$), while accounting for auto-correlation between different time periods --- i.e., between $Z_{i,t}, Z_{i,t-1}$ and $X_{i,t}, X_{i,t-1}$.
+    These results use a $\log-\log$ specification, so the estimates are for the elasticity of state funding per student in a year $t+k$ with respect to funding shift-share in year $t$, where years $k = 0, \hdots, 10$ are on the $x$-axis.
     Standard errors are clustered at the state-year level.
@@
 \autoref{sec:iv-model-uni} considers the number of professors employed by the university; whether a professor becomes employed at the university is likely affected by that university's finances.
-The formulation here takes as given that the professor is already employed at the university, and then projects the effect of changes in state funding on these \textit{incumbent} professors following the state funding shock.
+The formulation here takes as given that the professor is already employed at the university, and then projects the effect of changes in state funding on these \textit{incumbent} professors following the state funding shift-share.
 \autoref{tab:firststage-illinois} presents the first-stage results in Illinois data .
@@
     \centering
-    \caption{First Stage Estimates, for State Funding by Funding Shock in IBHED Data.}
-    %\textbf{Panel A: units in \$ per student}
-    %
-    %\makebox[\textwidth][c]{\input{tables/firststage-reg-rawcount.tex}}
+    \caption{First Stage Estimates, for State Funding by Funding Shift-Share in IBHED Data.}
+    \textbf{Panel A: units in \$ per student}
+
+    \makebox[\textwidth][c]{\input{tables/firststage-reg-rawcount.tex}}
 
@@
     \textbf{Note}:
-    These tables show the first stage OLS estimates of regression specification \eqref{eqn:secondstage1_indiv}, showing the effect of the funding shock on state funding to gauge performance as an instrument.
+    These tables show the first stage OLS estimates of regression specification \eqref{eqn:secondstage1_indiv}, showing the effect of the funding shift-share on state funding to gauge performance as an instrument.
     Each observation is a professor-year, in the IBHED data, and funding data are merged from IPEDS.
-    %Panel A shows the effect of an funding shock of \$-1 per student in the state on the number of \$'s of state funding per student at the university --- i.e.,
-    %\$-1 funding shock per student in the state leads to \$1.176 less state funding per student at the university according to preferred specification column 1.
-    Panel A shows the effect of a $-10$\% change funding shock per student in the state on $10$\% change in state funding per student at the university --- i.e.,
-    $-10$\% funding shock per student in the state leads to $-9.77$\% less state funding per student at the university according to prefferred specification column 1.        
+    %Panel A shows the effect of an funding shift-share of \$-1 per student in the state on the number of \$'s of state funding per student at the university --- i.e.,
+    %\$-1 funding shift-share per student in the state leads to \$1.176 less state funding per student at the university according to preferred specification column 1.
+    Panel A shows the effect of a $-10$\% change funding shift-share per student in the state on $10$\% change in state funding per student at the university --- i.e.,
+    $-10$\% funding shift-share per student in the state leads to $-9.77$\% less state funding per student at the university according to prefferred specification column 1.        
     Standard errors are clustered at the institution-year level, and institution $+$ year fixed effects are included where noted.
@@
 \end{eqnarray}
-We then interpret parameter $\beta$ as the effect of changes in state funding at an Illinois public university, via state funding shocks, on an individual professor's outcome $Y_{j,t}$.
+We then interpret parameter $\beta$ as the effect of changes in state funding at an Illinois public university, via state funding shift-shares, on an individual professor's outcome $Y_{j,t}$.
 
@@
     \textbf{Note}:
-    These figures show the local projections estimates of regression specification \eqref{eqn:secondstage}, with the funding shock as an instrument for state funding.
+    These figures show the local projections estimates of regression specification \eqref{eqn:secondstage}, with the funding shift-share as an instrument for state funding.
     The unit of analysis is an individual faculty member (at an Illinois public university); funding data come from IPEDS, and faculty promotion rate from IBHED.
-    The coefficient estimate is effect of state funding ($X_{i(j),t}$) on faculty promotion rate ($Y_{j,t}$), using the funding shock instrument ($Z_{i(j),t}$), while accounting for auto-correlation between different time periods --- i.e., between $X_{i(j),t}, X_{i(j),t-1}$ and $Y_{i(j),t}, Y_{i(j),t-1}$.
+    The coefficient estimate is effect of state funding ($X_{i(j),t}$) on faculty promotion rate ($Y_{j,t}$), using the funding shift-share instrument ($Z_{i(j),t}$), while accounting for auto-correlation between different time periods --- i.e., between $X_{i(j),t}, X_{i(j),t-1}$ and $Y_{i(j),t}, Y_{i(j),t-1}$.
     These results use a $\log-\log$ specification, so the estimates are for the rate of promotion in a year $t+k$ affected by a 1\% change in state funding in year $t$, where years $k = 0, \hdots, 4$ are on the $x$-axis.
@@
     \textbf{Note}:
-    These tables show the second stage OLS and 2SLS estimates of regression specification \eqref{eqn:secondstage}, showing the effect of state funding changes on number of faculty per student in Illinois universities, using the funding shock to instrument for state funding in the columns labelled 2SLS.
+    These tables show the second stage OLS and 2SLS estimates of regression specification \eqref{eqn:secondstage}, showing the effect of state funding changes on number of faculty per student in Illinois universities, using the funding shift-share to instrument for state funding in the columns labelled 2SLS.
     Each observation is a public university-year in the state of Illinois, where funding data come from IPEDS and faculty count come from IBHED data.
@@
     \textbf{Note}:
-    These figures show the local projections estimates of regression specification \eqref{eqn:secondstage}, with the funding shock as an instrument for state funding.
+    These figures show the local projections estimates of regression specification \eqref{eqn:secondstage}, with the funding shift-share as an instrument for state funding.
     The unit of analysis is an individual faculty member (at an Illinois public university); funding data come from IPEDS, and faculty promotion rate from IBHED.
-    The coefficient estimate is effect of state funding ($X_{i(j),t}$) on faculty promotion rate ($Y_{j,t}$), using the funding shock instrument ($Z_{i(j),t}$), while accounting for auto-correlation between different time periods --- i.e., between $X_{i(j),t}, X_{i(j),t-1}$ and $Y_{i(j),t}, Y_{i(j),t-1}$.
+    The coefficient estimate is effect of state funding ($X_{i(j),t}$) on faculty promotion rate ($Y_{j,t}$), using the funding shift-share instrument ($Z_{i(j),t}$), while accounting for auto-correlation between different time periods --- i.e., between $X_{i(j),t}, X_{i(j),t-1}$ and $Y_{i(j),t}, Y_{i(j),t-1}$.
     These results use a rate$-\log$ specification, so the estimates are for the rate of promotion in a year $t+k$ affected by a 1\% change in state funding in year $t$, where years $k = 0, \hdots, 4$ are on the $x$-axis.
@@
 \newpage
-\subsection{Additional Results, Faculty Hiring}
+\subsection{Professor Hiring}
 \label{sec:appendix-hiring}
 
-These results were produced by integrating the total count of faculty hires for 2010--2021 for the top-ranked 180 US universities with a sum of the funding variables, and then estimating the models specified in \autoref{sec:iv-model-uni}.
-There were no observable differences in the hiring rate of male vs female faculty.
-
-\begin{figure}[h!]
-    \centering
-    \singlespacing
-    \caption{State Funding and Faculty Hired at Public Universities, Total for 2011--2021.}
-    \includegraphics[width=0.8\textwidth]{figures/hiring-correlation.png}
-    \label{fig:hiring-correlation}
-    \justify
-    \footnotesize
-    \textbf{Note}:
-    This figure shows the correlation between state funding and the number of total number of faculty hired at public universities between 2011--2021, and the line is a line of best-fit.
-    Data for the number of faculty hired at public universities are provided by \cite{wapman2022quantifying,wapman2022zenodo}, and funding data from IPEDS.
-\end{figure}
-
-\newpage
-\begin{table}[h!]
+These results were produced by integrating the total count of professor hires for 2010--2021 for the top-ranked 180 US universities with a sum of the funding variables, and then estimating the models specified in \autoref{sec:iv-model-uni}.
+\begin{table}[H]
     \singlespacing
@@
     \textbf{Note}:
-    These tables show the second stage OLS and 2SLS estimates of regression specification \eqref{eqn:secondstage}, showing the effect of state funding changes on number of faculty hires at Illinois universities, using the funding shock to instrument for state funding in the columns labelled 2SLS.
+    These tables show the second stage OLS and 2SLS estimates of regression specification \eqref{eqn:secondstage}, showing the effect of state funding changes on number of faculty hires at Illinois universities, using the funding shift-share to instrument for state funding in the columns labelled 2SLS.
     Each observation is a public university-year in the state of Illinois, where funding data come from IPEDS and faculty count come from IBHED data.
@@
 
-\newpage
+There were no observable differences in the hiring rate of male vs female faculty.
+
 \begin{table}[h!]
@@
     \centering
-    \caption{OLS and 2SLS Estimates for University Faculty Hiring, Total for 2011--2020.}
+    \caption{OLS and 2SLS Estimates for Professor Hiring, Total for 2011--2020.}
     \makebox[\textwidth][c]{\input{tables/hiring-shock-reg-fte.tex}}
@@
     \textbf{Note}: 
-    This table show the second stage 2SLS estimates of regression specification \eqref{eqn:secondstage}, showing the effect of state funding changes on the number of faculty hires (per student) total for 2011--2021 at US public universities, using the funding shock to instrument for state funding.
+    This table show the second stage 2SLS estimates of regression specification \eqref{eqn:secondstage}, showing the effect of state funding changes on the number of faculty hires (per student) total for 2011--2021 at US public universities, using the funding shift-share to instrument for state funding.
     Yearly variation in total hires is not observed here, so only the total hires across 2011--2021 for 157 universities, can be considered.

### 2024-02-28 text/tables/facultycount-shock-reg-rawcount.tex
writing edits made.

@@
 \hline \\[-1.8ex] 
- & \multicolumn{8}{c}{Dependent Variable: Faculty Count per 100 Students, by Professor Group} \\ 
+ & \multicolumn{8}{c}{Dependent Variable: Faculty Count per 100 Students, by Position} \\ 
 \cline{2-9} 
-\\[-1.8ex] & \multicolumn{2}{c}{Lecturer} & \multicolumn{2}{c}{Assistant} & \multicolumn{2}{c}{Full} & \multicolumn{2}{c}{All} \\ 
+\\[-1.8ex] & \multicolumn{2}{c}{Lecturers} & \multicolumn{2}{c}{Asst. Professors} & \multicolumn{2}{c}{Full Professors} & \multicolumn{2}{c}{All Faculty} \\ 
  & OLS & 2SLS & OLS & 2SLS & OLS & 2SLS & OLS & 2SLS \\ 

### 2024-02-28 programs/analysis/hiring-shock.R

@@ -12,8 +12,8 @@ library(stargazer)
 # Define number of digits in tables and graphs
 digits.no <- 3
 # Size for figures
-fig.width <- 15
-fig.height <- fig.width * 1 / 2
+fig.width <- 12
+fig.height <- fig.width * 7 / 12
 
 
 # Load data sources ------------------------------------------------------------
@@ -108,6 +108,12 @@ reg.data <- ipeds_hiring.data %>%
 # Put togethe a visualisation of the rate of hiring between public + private
 # for 2010--2021
 quantile.count <- 20
+
+unitid.data <- read_csv("../../data/urban-ipeds/urban-clean-allunis.csv") %>%
+    filter(year == 2021) %>%
+    select(unitid, inst_name)
+
+
 # Viusalise how these two vary
 hiring_correlation.plot <- reg.data %>%
     # Create the per-student values
@@ -115,9 +121,10 @@ hiring_correlation.plot <- reg.data %>%
         stateappropriations_perEnroll =
             (stateappropriations_real / enrollment_reported),
         total_hired_perEnroll =
-            (total_hired / enrollment_reported)) %>%
+            (total_hired / (enrollment_reported / 100))) %>%
     # Take out outliers
-    filter(total_hired_perEnroll < 0.25,
+    filter(0.5 < total_hired_perEnroll, total_hired_perEnroll < 15,
+        stateappropriations_perEnroll < 15000,
         stateappropriations_perEnroll > 2000) %>%
     # Declare a plot to show the correlation.
     ggplot(aes(
@@ -125,18 +132,30 @@ hiring_correlation.plot <- reg.data %>%
         y = total_hired_perEnroll)) +
     geom_jitter() +
     geom_smooth(method = "lm") +
-    scale_x_continuous(name = "State Funding per Year, $ thousands per student",
-        breaks = 0:15, limits = c(2.5, 13), expand = c(0.01, 0.01)) +
+    scale_x_continuous(
+        name = "State Funding per Year, $ thousands per student",
+        expand = c(0, 0),
+        limits = c(2.5, 15),
+        breaks = seq(0, 15, by = 1)) +
     scale_y_continuous(name = "",
-        breaks = seq(0, 0.15, by = 0.025), limits = c(0, 0.15)) +
+        breaks = seq(0, 15, by = 2),
+        limits = c(0, 12)) +
     theme_bw() +
-    ggtitle("Faculty Hired, Total per student") +
[…]

### 2024-02-28 programs/analysis/ibhed-shock.R

@@ -256,7 +256,58 @@ reg.data %>%
         out = "../../text/tables/illinois-summary.tex")
 
 
-# First stage Regressions -----------------------------------------------------
+# Density plot of wages --------------------------------------------------------
+
+# Draw a plot of the salary distributions.
+salary_density.plot <- reg.data %>%
+    filter(3000 < salary_real + extra_salary_real,
+        salary_real + extra_salary_real < 251000,
+        year %in% c(2010, 2021)) %>%
+    mutate(faculty =
+        ifelse(position %in% c("Instructor", "Lecturer"), "Lecturer",
+        ifelse(position == "Assistant Professor", "Tenure-Track",
+        ifelse(position %in% c("Associate Professor", "Professor"), "Tenured",
+        ifelse(position %in% c("Chancellor",
+            "Other Administrator", "President", "Senior Officer",
+            "Unit Director", "Vice Chancellor", "Vice President"), "Administrator",
+            ""))))) %>%
+    ggplot(aes(x = (salary_real + extra_salary_real) / 1000,
+        fill = faculty, colour = faculty)) +
+    geom_density(alpha = 1 / 3) +
+    theme_bw() +
+    scale_x_continuous(expand = c(0, 0),
+        name = "Annual Salary + Benefits, $ thousands",
+        breaks = seq(0, 250, by = 50),
+        limits = c(0, 265)) +
+    scale_y_continuous(expand = c(0, 0.0001),
+        name = "") +
+    ggtitle("Density") +
+    facet_grid(cols = vars(year)) +
+    theme(
+        legend.title = element_blank(),
+        plot.title = element_text(size = rel(1)),
+        plot.margin = unit(c(0.5, 0, 0, 0), "mm"),
+        legend.position = "bottom",
+        legend.margin = margin(t = -5))
+# Save this plot
+presentation.width <- 17.5
+presentation.height <- presentation.width / 2
+ggsave("../../text/figures/salary-density.png",
+    plot = salary_density.plot,
+    units = "cm", dpi = 300,
+    width = presentation.width, height = presentation.height)
+
+# Show how many faculty in eah year, 2010 + 2021
+reg.data %>%
+    filter(3000 < salary_real + extra_salary_real,
+        salary_real + extra_salary_real < 251000,
+        year %in% c(2010, 2021)) %>%
+    group_by(year) %>%
+    count() %>%
+    print()
+
+
[…]

### 2024-02-28 programs/analysis/ipeds-public-private.R

@@ -13,7 +13,7 @@ digits.no <- 3
 
 # Size for figures
 fig.width <- 9
-fig.height <- fig.width * 0.85
+fig.height <- fig.width * 0.9
 
 
 # Load data sources ------------------------------------------------------------
@@ -68,6 +68,14 @@ gc()
 
 # Public vs private enrollment comparison --------------------------------------
 
+# How many public unis + students in 2020/2021?
+ipeds.data %>%
+    group_by(public, year) %>%
+    summarise(
+        total_count = n(),
+        enrollment_reported = sum(enrollment_reported, na.rm = TRUE)) %>%
+    ungroup() %>%
+    View() #print()
 
 # Show enrollment at public vs private over time: total + average
 enrollment.data <- ipeds.data %>%
@@ -184,28 +192,28 @@ prof_perfte.data <- ipeds.data %>%
     group_by(public, year) %>%
     summarise(
         full_fte_perprof     = sum(full_prof_count, na.rm = TRUE) /
-            sum(enrollment_reported, na.rm = TRUE),
+            sum(enrollment_reported / 100, na.rm = TRUE),
         assistant_fte_perprof = sum(assistant_prof_count, na.rm = TRUE) /
-            sum(enrollment_reported, na.rm = TRUE),
+            sum(enrollment_reported / 100, na.rm = TRUE),
         lecturer_fte_perprof  = sum(lecturer_prof_count, na.rm = TRUE) /
-            sum(enrollment_reported, na.rm = TRUE),
+            sum(enrollment_reported / 100, na.rm = TRUE),
         all_fte_perprof         = sum(all_prof_count, na.rm = TRUE) /
-            sum(enrollment_reported, na.rm = TRUE)) %>%
+            sum(enrollment_reported / 100, na.rm = TRUE)) %>%
     ungroup()
 
 # Draw the graph for lecturers Professors per student
 lecturer_fte_perprof.plot <- prof_perfte.data %>%
-    ggplot(aes(x = year, y = 1 / lecturer_fte_perprof, colour = factor(public))) +
+    ggplot(aes(x = year, y = lecturer_fte_perprof, colour = factor(public))) +
     geom_point() +
     geom_line() +
     scale_x_continuous(name = "Year",
         breaks = seq(1985, 2020, by = 5)) +
     scale_y_continuous(name = "",
-        limits = c(20, 350),
-        breaks = seq(0, 350, by = 25),
[…]

### 2024-02-28 programs/analysis/ipeds-shock-rawcounts.R

@@ -95,7 +95,7 @@ reg.data <- reg.data %>%
     filter(1990 <= year,
         !is.na(enrollment_reported), enrollment_reported > 0,
         !is.na(all_prof_count),
-        !is.na(stateappropriations_real),
+        stateappropriations_real > 0,
         !is.na(appropriationshock_perEnroll_real)) %>%
     # Restrict to uni-years with all professor counts.
     filter(!is.na(lecturer_prof_count), lecturer_prof_count > 0) %>%
@@ -313,8 +313,9 @@ stargazer(
     naive_all_count.reg, shiftshare_all_count.reg,
     add.lines = outcome.means,
     dep.var.caption =
-        "Dependent Variable: Total Employment Count by Professor Group",
-    dep.var.labels = c("Lecturer", "Assistant", "Full", "All"),
+        "Dependent Variable: Faculty Count per 100 Students, by Position",
+    dep.var.labels = c(
+        "Lecturers", "Asst. Professors", "Full Professors", "All Faculty"),
     column.labels = rep(c("OLS", "2SLS"), 4),
     digits = digits.no,
     digits.extra = digits.no,
@@ -329,3 +330,61 @@ stargazer(
     omit.table.layout = "n", notes.append = FALSE,
     type = "text",
     out = "../../text/tables/facultycount-shock-reg-rawcount.tex")
+
+# Plot the results.
+# Estimate with separate names
+plot_lecturer_count.reg <- reg.data %>%
+    mutate(lect_funding = - (stateappropriations_real / (1000 * enrollment_reported))) %>%
+    felm(lecturer_prof_count ~ 1 |
+        unitid + year |
+        (lect_funding ~ I(-appropriationshock_perEnroll_real)) |
+        state + year,
+        data = .)
+plot_assistant_count.reg <- reg.data %>%
+    mutate(asst_funding = - (stateappropriations_real / (1000 * enrollment_reported))) %>%
+    felm(assistant_prof_count ~ 1 |
+        unitid + year |
+        (asst_funding ~ I(-appropriationshock_perEnroll_real)) |
+        state + year,
+        data = .)
+plot_full_count.reg <- reg.data %>%
+    mutate(full_funding = - (stateappropriations_real / (1000 * enrollment_reported))) %>%
+    felm(full_prof_count ~ 1 |
+        unitid + year |
+        (full_funding ~ I(-appropriationshock_perEnroll_real)) |
+        state + year,
[…]

### 2024-02-28 programs/analysis/ipeds-shock.R

@@ -2,6 +2,7 @@
 ## Senan Hogan-Hennessy, 3 August 2022
 ## IV for Prof's salaries, using IPEDS data.
 print(Sys.time())
+library(jtools) # Plotting estimates.
 library(tidyverse) # Functions for data manipulation and visualization
 library(lfe) # Functions for fast linear models with IV + FEs
 library(plm) # Functions for panel data
@@ -19,7 +20,7 @@ digits.no <- 3
 
 # Size for figures
 fig.width <- 10
-fig.height <- fig.width * 0.85
+fig.height <- fig.width * 0.9
 
 # Define a function to standardise the dplyr lag function
 # --- negative numbers mean a lag of t time periods, positive a lead of t.
@@ -52,48 +53,54 @@ reg.data <- ipeds.data %>%
         # Appropriations to the uni
         totalrevenues_real,
         nonauxrevenues_real,
+        fedappropriations_real,
         stateappropriations_real,
+        localappropriations_real,
+        statelocal_funding_real,
         tuitionrev_real,
         # Uni spending.
         nonauxspending_real,
         # Uni enrollment
         enrollment_reported,
         enrollment_fte,
-        # Appropriation shocks Per student
+        acceptance_rate,
+        completion_rate_150pct,
+        # Shift-sahre shocks Per student
         appropriationshock_perEnroll_real,
-        appropriationshock_perFTE_real,
-        appropriationshock_peruni_real,
+        allstate_stateappropriations_real,
+        staterevenues_baseshare,
+        stateEnroll_count,
         # Count of full-time professors (tenured or not)
         lecturer_prof_count,
         assistant_prof_count,
         full_prof_count,
         all_prof_count,
-        # Count of professors by tenure-rate (only 1990-)
-        nontenured_tenure_count,
-        tenuretrack_tenure_count,
-        tenured_tenure_count,
-        all_tenure_count,
-        # Average faculty salary
-        full_profmeansalary_real,
-        assistant_profmeansalary_real,
-        lecturer_profmeansalary_real,
-        all_profmeansalary_real,
-        # Total paid on professor salaries
-        all_profoutlays_real,
-        # Research and instruction expenditures.
+        # Research and instruction and other expenditures.
         instructionspending_total_real,
[…]

### 2024-02-28 programs/data-build/urban-ipeds-build.R

@@ -18,14 +18,23 @@ finance.data <- read_csv("../../data/urban-ipeds/raw-data/colleges_ipeds_finance
     # Include Supplementary finances for 2018-2020
     bind_rows(read_csv(
         "../../data/urban-ipeds/raw-data/finance/ipeds-finance-20182020.csv"))
+# Urban admissions data (available 2001--)
+admissions.data <- read_csv("../../data/urban-ipeds/raw-data/colleges_ipeds_admissions-enrollment.csv")
+# Urban data on graduation rates (available 1996--)
+graduation150.data <- read_csv("../../data/urban-ipeds/raw-data/colleges_ipeds_grad-rates.csv")
+
+# Urban data on faculty.
+urban_salaries.data <- read_csv("../../data/urban-ipeds/raw-data/colleges_ipeds_salaries_is.csv")
 
 # University enrollment (hand made from raw IPEDS values)
 enrollment.data <- read_csv("../../data/urban-ipeds/raw-data/enrollment/ipeds-enrollment.csv")
 
 # Professors' salaries (hand made from raw IPEDS values)
 salaries.data <- read_csv("../../data/urban-ipeds/raw-data/faculty/ipeds-salaries.csv")
-# Professors' faculty rate (hand made from raw IPEDS values)
+# Professors' faculty data (hand made from raw IPEDS values)
 tenure.data <- read_csv("../../data/urban-ipeds/raw-data/faculty/ipeds-tenure.csv")
+# Employee part- + full-time data (hand made from raw IPEDS values)
+parttime.data <- read_csv("../../data/urban-ipeds/raw-data/parttime-employees/ipeds-parttime.csv")
 
 # CPI-U from FREDS (since Urban provided cpi has missing years)
 # Yearly average, seasonally adjusted, base year 1982-1984=100
@@ -68,71 +77,106 @@ finance.data <- finance.data %>%
         cpi2021 = cpi2021,
         # Federal appropriations
         fedappropriations_real = replace(rev_appropriations_fed,
-            rev_appropriations_fed <= 0, NA) / cpi2021,
+            rev_appropriations_fed < 0, NA) / cpi2021,
         # State Gov. appropriations
         stateappropriations_real = replace(rev_appropriations_state,
-            rev_appropriations_state <= 0, NA) / cpi2021,
+            rev_appropriations_state < 0, NA) / cpi2021,
+        # State and local grants
+        statelocal_funding_real = replace(rev_state_local_approps_grants,
[…]

### 2024-03-15 text/paper.tex
updated

@@
         Romain Campbell,
+        Maddalena Davoli,
         Ryan Dycus,

### 2024-03-15 text/sections/01-introduction.tex
updated

@@
 On the other hand, lecturers could be hired to fill in teaching duties, on shorter and cheaper contracts with fewer employment protections.
-But it is not certain that universities would hire more lecturers in the face of budget cuts; they may not hire any more faculty at all.
-It is possible that universities are be limited from hiring more lecturers by competitive pressures from peer institutions \citep{hoxby2009changing,urquiola2020markets}, leading university administrations to court private funding or raise tuition prices to fill funding gaps, leaving faculty unaffected.
+But it is not certain that universities would hire more lecturers in the face of budget cuts.
+Public university administrators, motivated by increased competition between unviersity for student enrolment \citep{hoxby2009changing,urquiola2020markets},may notice that their rankings are negatively affected by falling professor--lecturer ratios, and so limit departments from hiring more lecturers to replace leaving/retiring professors.
+Administrators could even start courting private funding or increase tuition prices to fill funding gaps \citep{bound2019public}, possibly leaving faculty unaffected by state funding cuts.
+These are some ways that the long-run stagnation in state funding would not guaratee a substitution towards lecturers at US public universities.
 
-State funding for higher education is not decided randomly, so correlations between state funding and faculty outcomes do not necessarily represent causal relationships.
-State governments decide how to fund higher education by a complicated process, which could be influenced by lobbying, changing state priorities, or local economic conditions.
+State funding for higher education is not decided randomly;
+universities that rely more on state funding may also be institutions that find it difficult to attract new faculty thanks to location, or other factors.
+This means that correlations between state funding and faculty outcomes do not necessarily represent causal relationships.
 As such, it would be na\"ive to think of state higher education funding as independent of public university faculty, so I employ a shift-share instrumental variables (IV) approach to identify causal effects of state funding cuts.
 The shift-share instrument identifies exogenous changes in state funding by exploiting differences in how much universities relied on state funding, interacted with state-wide shifts in higher education funding --- following the methods of \cite{NBERw23736,NBERw27885}.
-Yearly state government budgets are decided without targetting individual universities for funding cuts, so more reliant universities will be more affected by state-wide changes in higher education funding, allowing the shift-share approach to estimate causal effects among more reliant universities.
+Yearly state government budgets are decided without targetting individual universities for funding cuts, so more reliant universities will be more affected by state-wide funding cuts, so the shift-share approach estimates causal effects among more reliant universities.
 Additionally, I estimate effect of state funding on faculty multiple years after the initial funding cuts, using the local projections method.\footnote{
@@
 }
-This approach allows me to test whether the effects of funding cuts are persistent or universities recover quickly e.g., by quickly resuming faculty hiring in years following funding cuts.
+This approach allows me to test whether the effects of funding cuts are persistent or universities recover quickly e.g., by quickly resuming faculty hiring in years after funding cuts.
 
@@
 The closest related research on higher education funding examined how university spending affected graduation rates and levels of student debt \citep{NBERw23736,NBERw27885}, and university finances \citep{miller2022making,bound2019public,brown2014endowment}.
-For faculty outcomes, applied theory work models university decision-making and faculty hiring (see e.g., \citealt{abe2015implications,johnson2009jep,NBERc13879}).
+For faculty outcomes, descriptive work measures how US and Canadian faculty salaries are correlated with other factors \citep{bratsberg2003negative,martinello2009faculty,pfeffer1992salary}, and applied theory work models university decision-making and faculty hiring (see e.g., \citealt{abe2015implications,johnson2009jep,NBERc13879}).
 Other papers measure university decision--making after endowment shocks \citep{brown2014endowment}, and trends in the academic job market following the 2008 recession \citep{turner2014impact}.
-These papers measure changes in state funding and university revenues, but do not measure effects on faculty composition or salaries, or other outcomes related to university instruction.
+These papers measure changes in state funding and university revenues, but do not measure effects of state funding cuts on faculty composition or salaries, or other outcomes related to university instruction.
 My results provide evidence that the long term trends in higher education funding and faculty outcomes are causally related, contributing to the literature on faculty outcomes \citep{ehrenberg2003studying} and trends in US higher education and funding \citep{hoxby2009changing,ehrenberg2012american}.
@@
 While funding cuts force universities to substitute towards lecturers, there is consistent evidence that students are worse off being taught by part-time lecturers than full-time faculty.
-Universities that rely on more lecturers have persistently lower student retention and graduation rates \citep{jaeger2011examining,ehrenberg2005tenured}
+Universities that rely on more lecturers have lower student retention and graduation rates \citep{jaeger2011examining,ehrenberg2005tenured}.
 Similarly, students at a large public university have consistently worse education outcomes if taught by adjunct lecturers than those taught by full-time faculty, largely thanks to the worse working conditions adjunct lecturers face \citep{zhu2021limited}.\footnote{
-    \cite{bettinger2010does,figlio2015tenure} study lecturers at a highly selective universities, finding that adjunct lecturers improve subject retention for undergraduates compared to professors.
-    This paper considers public universities, which are by-majority not selective, and where lecturers are more often employed part-time with less institutional support.
+    Alternatively, \cite{bettinger2010does,figlio2015tenure} study lecturers at a highly selective universities, finding that adjunct lecturers improve subject retention for undergraduates compared to professors.
+    This paper studies faculty at public universities, which are by-majority not selective, and where lecturers are more often employed part-time with less institutional support.
 }

### 2024-03-15 text/sections/03-concepts.tex
updated

@@
 If faculty are affected by state funding cuts, then it is still not clear whether effects are persistent or merely transitory.
-Universities employ professors on multiple-year contracts (e.g., the tenure contract), and decisions to retire, move university, or leave academia are multiple-year commitments, so that it may take multiple years for effects to trickle down to faculty.
+Universities employ professors on multiple-year contracts (e.g., the tenure contract), and professors' decisions to retire, move university, or leave academia are multiple-year commitments, so that it may take multiple years for effects to trickle down to faculty.
 On the other hand, university departments hire in annual cycles, so that a funding cut in one year may immediately impact faculty hiring in the same year, and thus faculty counts in the next academic year.

### 2024-03-15 text/sections/04-empirical.tex
updated

@@
 \label{sec:approp-shocks}
-
-I use a shift-share instrument to address endogeneity concerns for the amount of funding a state allots to each of its public universities.
-\cite{NBERw23736,NBERw27885} develop the instrument for public university finances by exploiting a shift-share instrument for changes in state-level funding interacted with university reliance on state funding in a base period.
+I use a shift-share instrument to address endogeneity concerns for state governments' funding decisions, for their public universities.
+\cite{NBERw23736,NBERw27885} develop the shift-share approach for public university funding by exploiting shifts in yearly state-wide higher eduction funding, interacted with universities' reliance on state funding in a base period.
 \begin{align}
@@
 
-The system exploits the fact that institutions who rely more on state funding will be affected by state-wide funding cuts.
+The system exploits the fact that institutions who rely more on state funding will be affected by follow-on state-wide funding cuts.
 $Z_{i,t}$ is the shift-share instrument for institution $i$ in year $t$.
-It is contructed by interacting the average funding for universities in state $s(i)$ with reliance on state funding relative to total revenues, averaged across the base years 1990--1993.\footnote{
+It is constructed by interacting the average funding for universities in state $s(i)$ with reliance on state funding relative to total revenues, averaged across the base years 1990--1993.\footnote{
     1990--1993 are defined as data for public university finance data are most comparable (i.e., without many missing values) beginning in 1990.
@@
 }
-The instrument exogeneity and exclusion follows the same argument as above, where the Illinois legislature did not target any single campus for funding cuts and no single Illinois campus takes the majority of state funding.
+The instrument exogeneity and exclusion follows the same argument as above, where no single Illinois campus takes the majority of state funding, and so
+the Illinois state government's funding cuts are not majority aimed at any single campus.
 

### 2024-03-15 text/sections/04-empirical.tex
updated

@@
 The shift-share instrument is positively associated with the total enrolment and total amount of state funding for each university, in both \$ and $\log$/percentage change terms, while other the other sources of university finances are not associated with the funding shift-share.
-So that the other sources of finances are not clear confounders for the IV strategy, as they exhibit balance with respect to the funding shift-share \citep{pei2019poorly}.
-Yearly and individual fixed effects are included in regressions throughout to implicitly condition on mean differences between universities and years.
+So that the other sources of finances are not clear confounders for the IV strategy, as they exhibit balance with respect to the funding shift-share.
+Nonetheless, yearly and institution fixed effects are included in regressions throughout to implicitly condition on mean differences between universities and years.
 
@@
 \label{sec:iv-model-uni}
-
 I use the instrument defined in \autoref{eqn:public-instrument} to overcome the endogeneity concerns for state funding to each public university, for an IV approach.

### 2024-03-15 text/sections/05-results.tex
updated

@@
 \subsection{Reduced Hiring of New Professors}
-The results above present a condundrum: there are large substitution effects among faculty, but faculty themselves are unaffected by funding cuts.
+The results above present a conundrum: there are large substitution effects among faculty, but faculty themselves are unaffected by funding cuts.
 The analysis of Illinois faculty (IBHED) studies the effect of state funding cuts on faculty already at the university --- i.e., incumbent faculty.
-The findings above imply that subsitution towards lecturers came about by limiting hiring of new professors, to replace leaving or retiring professors, so that they are replaced by hiring lecturers.
+The findings above imply that substitution towards lecturers came about by limiting hiring of new professors, to replace leaving or retiring professors, so that they are replaced by hiring lecturers.
+In this subsection, I test this implication with data on universities' hiring of new professors.
 
-To further study this issue, I take recent data on the total faculty hired at US public universities over 2011--2021 provided by \cite{wapman2022quantifying,wapman2022zenodo}.
-\cite{wapman2022quantifying} study the composition of professor hiring in the years 2011--2021, and publiccally released data on the number of professors (not lecturers) hired at US universities.\footnote{
-    These data are not panel data, instead only give the total professors hired over the ten year period 2011--2021.
+I take recent data on the total faculty hired at US public universities over 2011--2021 provided by \cite{wapman2022quantifying,wapman2022zenodo}.
+\cite{wapman2022quantifying} study the composition of professor hiring in the years 2011--2021, and publicly released data on the number of professors (not lecturers) hired at US universities.\footnote{
+    These data are not panel data, instead only give the total professors hired over the ten-year period 2011--2021.
 }
-I combine this measure, with total state funding for public universities acorss the same time period from IPEDS, to validate the claim that state funding cuts led to faculty subtitution via limiting hiring of new professors.
+I combine this measure, with total state funding for public universities across the same time period from IPEDS, to validate the claim that state funding cuts led to faculty substitution via limiting hiring of new professors.
 

### 2024-03-15 text/sections/06-discussion.tex
updated

@@
 %SO prof per student is where the movemenet is most stark.}
-Universities received less state funding, relative to increasing enrolment, and reacted by increasing their reliance on lecturers relative to assistant and full professors in both the short and medium-run; these results imply that lecturers can be considered substitutes for professors in the tenure-system.\footnote{
-    Additional results in Appendix \autoref{sec:appendix-substitution} formalises the marginal rate of substitution between lecturers and professors on tenure-track, showing how assistant and full professors were substituted for lecturers while there is no substitution between assistant and full professors.
-}
+Universities received less state funding, relative to increasing enrolment, and reacted by increasing their reliance on lecturers relative to assistant and full professors in both the short and medium-run; these results imply that lecturers can be considered substitutes for professors in the tenure-system.
+%\footnote{
+%    Additional results in Appendix \autoref{sec:appendix-substitution} formalises the marginal rate of substitution between lecturers and professors on tenure-track, showing how assistant and full professors were substituted for lecturers while there is no substitution between assistant and full professors.
+%}
 The average public university's state funding fell by 36\% per student, while the number of lecturers per student increased by 113\%, and assistant and full professors  fell by 17\% and 23\%, respectively.

### 2024-03-15 text/sections/07-conclusion.tex
updated

@@
 \label{sec:conclusion}
-
-This paper studies how stagnating state funding for higher education affected faculty, and their composition, at US public universities.
-This work contributes to the literature along two primary dimensions.
-First, by isolating changes in state funding on public universities via state funding shocks, this work provides an explanation for the increased reliance on lecturers and away from tenure-track and tenured professors at US public universities.
-This approach used multiple methods to estimate the short- and medium-run effects of state funding cuts on faculty outcomes.
-Second, this work investigates how individual faculty are affected by changes in state funding for their university using a dataset new in the economic literature (IBHED).
-These data allow for detailed analysis of thousands of faculty salaries, and employment outcomes in the Illinois university system.
-
 Public universities have systematically substituted away from tenure-track and tenured professors, toward lecturers, in the face of persistent declines in state funding for higher education.
-The effects on incumbent professors in Illinois are non-distinguishable from zero, with suggestive evidence for reduced hiring of new professors, implying that the faculty substitution is driven by reduced hiring of new professors.
+The effects on incumbent professors in Illinois are non-distinguishable from zero, and suggestive evidence shows that funding cuts limited hiring of new professors, implying that substitution towards lecturers is driven by reduced hiring of new professors.
 Public universities are using more contingent lecturers to teach their students, while private universities continue to employ more tenure-track and tenured professors than their public counterparts, and each year the gap widens.
 
-At the same time as higher education costs have been rising in the US, public universities have also dealt with declining state funding.
-I show that these headwinds led to systematic change at public universities, changes that affect their faculty, and likely limiting public universities in their goals in research and education.
+This work contributes to the literature along two primary dimensions.
+First, this work provides an explanation for the increased reliance on lecturers and away from tenure-track and tenured professors at US public universities, by isolating the effects funding cuts from state-wide funding shifts in both the short-- and long-run.
+Second, this work investigates how individual faculty are affected by changes in state funding for their university using a dataset new in the economic literature (IBHED).
+These data allow for detailed analysis of thousands of faculty salaries, and employment outcomes, in the Illinois university system, and show that incumbent faculty were relatively unaffected by state funding cuts to their employers.
+
+While state funding for higher education has stagnated, costs of education has been rising at the same time, so that funding cuts have even more impact today than they did three decades ago.
+I show that these financial headwinds led to systematic change at public universities, leading to a substitution away from stable employment contracts (tenure, professorship) and towards less stable arrangements (adjunct contracts, lecture positions).
 These results show large changes in faculty composition, and that stagnation in state support explains at least a third of the observed shift away from tenured professors and towards lecturers.
-At the same time, private universities were not exposed to financial headwinds of the same magnitude or persistence.
-While public universities continue to educate the majority of higher education students in the US, we should worry about the effects of restricting their funding has on faculty, and the wider impact on higher education as a whole.
+This likely had knock-on effects on instruction, contribiting to worsening student outcomes at US public universities since the early 1990s --- while private universities were not exposed to financial headwinds of the same magnitude or persistence over the same time period.
+While public universities educate the majority of US higher education students, we should worry about the effects of restricting their funding has on faculty, and the wider impact on higher education as a whole.

### 2024-03-15 text/sections/10-appendix.tex
updated

@@
 
-\newpage
-\subsection{Rates of Substitution}
-\label{sec:appendix-substitution}
-
-The funding elasticities can be used to recover the marginal rate of substitution between two outcomes.
-For example, write $Y^1$ for the number of lecturers per student at a university, and $Y^2$ for the number of full professors.
-I use the above approaches to estimate the funding elasticities, where $\% \Delta$ denotes percent change.
-\[ \beta_1 = \frac{\% \Delta Y^1}{\% \Delta X}
-\text{, and }
-\beta_2 = \frac{\% \Delta Y^2}{\% \Delta X} \]
-As such, it is possible to recover the elasticity for substitution between lecturers and full professors by the universities via the respective funding elasticities.
-\[ \frac{\% \Delta Y^1}{\% \Delta Y^2}
-= \frac{\% \Delta Y^1 / \% \Delta X}{\% \Delta Y^2 / \% \Delta X}
-= \frac{\beta_1}{\beta_2} \]
-I present results for the rates of substitution between different levels of faculty by this approach, dividing the relevant coefficient estimates and presenting standard errors calculated by a non-parametric bootstrap.
-In practice, this corresponds to division of the estimates of the elasticity for employment of professors (by rank) with respect to state funding, presented in Panel B \autoref{tab:facultycount-shock-reg}, and bootstrapping the results to generate standard errors and confidence intervals.
-% https://stackoverflow.com/questions/63777368/computing-the-standard-error-when-dividing-coefficients-of-different-regressions
-
-The implied marginal rate of substitution between lecturers and assistant professors is estimated as -3.26 (standard error 0.50), based on 10,000 bootstrap samples.
-This means that public universities increased their number lecturers per student by 3.26\% when they decreased their count of assistant professors, on average and subject to the changes in state funding they experienced 1990--2017.
-Between lecturers and full professors the rate of substitution is -3.19 (0.34), which implies that universities substitute between lecturers and full professors in the same way.
-Between assistant and full professors the rate of substitution is 0.99 (0.11), which intuitively implies that universities treated assistant and full professors (i.e., those before and after tenure in the tenure system) as complements.
-
+%\newpage
+%\subsection{Rates of Substitution}
+%\label{sec:appendix-substitution}
+%The funding elasticities can be used to recover the marginal rate of substitution between two outcomes.
+%For example, write $Y^1$ for the number of lecturers per student at a university, and $Y^2$ for the number of full professors.
+%I use the above approaches to estimate the funding elasticities, where $\% \Delta$ denotes percent change.
+%\[ \beta_1 = \frac{\% \Delta Y^1}{\% \Delta X}
+%\text{, and }
+%\beta_2 = \frac{\% \Delta Y^2}{\% \Delta X} \]
+%As such, it is possible to recover the elasticity for substitution between lecturers and full professors by the universities via the respective funding elasticities.
+%\[ \frac{\% \Delta Y^1}{\% \Delta Y^2}
+%= \frac{\% \Delta Y^1 / \% \Delta X}{\% \Delta Y^2 / \% \Delta X}
+%= \frac{\beta_1}{\beta_2} \]
+%I present results for the rates of substitution between different levels of faculty by this approach, dividing the relevant coefficient estimates and presenting standard errors calculated by a non-parametric bootstrap.
+%In practice, this corresponds to division of the estimates of the elasticity for employment of professors (by rank) with respect to state funding, presented in Panel B \autoref{tab:facultycount-shock-reg}, and bootstrapping the results to generate standard errors and confidence intervals.
+%% https://stackoverflow.com/questions/63777368/computing-the-standard-error-when-dividing-coefficients-of-different-regressions
+%
+%The implied marginal rate of substitution between lecturers and assistant professors is estimated as -3.26 (standard error 0.50), based on 10,000 bootstrap samples.
+%This means that public universities increased their number lecturers per student by 3.26\% when they decreased their count of assistant professors, on average and subject to the changes in state funding they experienced 1990--2017.
+%Between lecturers and full professors the rate of substitution is -3.19 (0.34), which implies that universities substitute between lecturers and full professors in the same way.
+%Between assistant and full professors the rate of substitution is 0.99 (0.11), which intuitively implies that universities treated assistant and full professors (i.e., those before and after tenure in the tenure system) as complements.
+%
 %"Calculated point est for substitution between lecturers + ast profs" 

### 2024-03-15 text/sections/01-introduction.tex
updated

@@
 But it is not certain that universities would hire more lecturers in the face of budget cuts.
-Public university administrators, motivated by increased competition between unviersity for student enrolment \citep{hoxby2009changing,urquiola2020markets},may notice that their rankings are negatively affected by falling professor--lecturer ratios, and so limit departments from hiring more lecturers to replace leaving/retiring professors.
+Public university administrators, motivated by increased competition \citep{hoxby2009changing,urquiola2020markets}, may notice that their rankings are negatively affected by falling professor--lecturer ratios, and so limit departments from hiring more lecturers to replace leaving/retiring professors.
 Administrators could even start courting private funding or increase tuition prices to fill funding gaps \citep{bound2019public}, possibly leaving faculty unaffected by state funding cuts.
-These are some ways that the long-run stagnation in state funding would not guaratee a substitution towards lecturers at US public universities.
+These are some ways that the long-run stagnation in state funding would not guarantee a substitution towards lecturers at US public universities.
 

### 2024-04-30 text/sections/00-abstract.tex
updated
Added new robustness check tables.

@@
 Using a shift-share approach to instrument for state funding, I find that universities employ 4.4\% more lecturers per student following a 10\% funding cut.
-This shift is accompanied by a reduction in assistant professors and full professors per student by 1.4\% and 1.2\%, respectively.
+This shift is accompanied by a reduction in assistant and full professors by 1.4\% and 1.2\%  per student, respectively.
 Incumbent professors' salaries, promotion rates, and quit rates at Illinois universities remain unaffected by funding cuts, indicating that the substitution arose from limiting the hiring of new tenure-track/tenured professors.
-Stagnating state funding impacts public universities and faculty, likely contributing to the deterioration of student outcomes at public universities since the 1990s.
+Stagnating state funding impacts public universities and faculty, likely contributing to deteriorating student outcomes at public universities seen since the 1990s.
 

### 2024-04-30 text/sections/05-results.tex
Added new robustness check tables.

@@
 \autoref{fig:count-lp} shows local projection estimates, where the estimate is the elasticity for professor count per student in year $t+k$ with respect to state funding in year $t$ for later years $k = 0, \hdots, 10$.
-These estimates correlespond to the elasticity estimates in \autoref{tab:facultycount-shock-reg} Panel B, where the outcome is faculty count per students in later years.
-State funding cuts affectthe number of lecturers and professors at universities up to 9 years after the initial funding cut.
-A 10\% state funding cut leads to 4.2\% (2.9--5.5 in a 95 percent confident interval) more lecturers per student the following year, and around 3.1\% (0.6--5.5 ``'') more lecturers nine years later.
-A 10\% state funding cut leads to 1.5\% (0.6--2.2 in a 95 percent confident interval) fewer assistant professors per student the following year, and around 1.1\% (0.2--2.0 ``'') fewer nine years later;
-the same cut leads to 1.2\% (0.6--1.8 in a 95 percent confident interval) fewer full professors per student the following year, and around 0.9\% (0.1--1.7 ``'') fewer nine years later.
+These estimates correspond to the elasticity estimates in \autoref{tab:facultycount-shock-reg} Panel B, where the outcome is faculty count per students in later years.
+State funding cuts affect the number of lecturers and professors at universities up to 9 years after the initial funding cut.
+A 10\% state funding cut leads to 4.2\% (2.9--5.5 in a 95 percent confident interval) more lecturers per student the following year, and around 3.1\% (0.6--5.5 confidence interval) more lecturers nine years later.
+A 10\% state funding cut leads to 1.5\% (0.6--2.2 in a 95 percent confident interval) fewer assistant professors per student the following year, and around 1.1\% (0.2--2.0 confidence interval) fewer nine years later;
+the same cut leads to 1.2\% (0.6--1.8 in a 95 percent confident interval) fewer full professors per student the following year, and around 0.9\% (0.1--1.7 confidence interval) fewer nine years later.
 The effects are a substitution with the composition of faculty, as the total number of faculty per students are relatively unaffected in the follow-on years.
+
+\textbf{Paragraph comparing to any previous estimates, as do Lovenheim 2020 page 22.}
 
@@
 
-%\subsection{Robustness checks}
-%\textbf{Describe robustness checks in one or two paragraphs here; see how Lovenheim$+$ did this section.}
+\textbf{Paragraph comparing to any previous estimates, as do Lovenheim 2020 page 22.}
+
+\subsection{Robustness Checks}
+\label{sec:results-robustness}
+\textbf{Describe robustness checks in one or two paragraphs here; see how Lovenheim$+$ did this section.}
+
+These tables describe the same analysis, with other controls added in, and with State $+$ Year fixed effects --- instead of Institution $+$ Year fixed effects.
+\autoref{tab:firststage-robustness-checks},
+\autoref{tab:facultycount-robustness-checks},
+\autoref{tab:facultycount-rawcount-robustness-checks}.
+Uses the FTE enrolment measure, with measures of selectivity.
+
+\autoref{tab:facultycount-shock-private-robustness} describes the results of using private universities in a falsification test, by assigning their base share as the average of similarly selective public universities in their state, and using this as an instrument for total revenues (as state funding is uncommon for private universities).
+Note, the first-stage is not strong, and the IV estimates are all (encouragingly) not significant at any standard frequentist cut-off.
+
+\autoref{tab:facultycount-heterogeneity} shows  the effects of state funding cuts, separated among universities with different selectivity rankings.
+The most selective (rank 1-2, including Berkeley to SUNY Binghamton) have no faculty substitution effects, but even among selective universities (rank 3-4, including Clemson University to University of Dallas) there are strong substitution effects.
+1 indicates “most competitive,” 2 is “highly competitive plus,” 3 is “highly competitive,” and 4 is “very competitive plus.”
 
@@
 
-I take recent data on the total faculty hired at US public universities over 2011--2021 provided by \cite{wapman2022quantifying,wapman2022zenodo}.
-\cite{wapman2022quantifying} study the composition of professor hiring in the years 2011--2021, and publicly released data on the number of professors (not lecturers) hired at US universities.\footnote{
+I take recent data on the total faculty hired at US public universities over 2011--2021 provided by \cite{wapman2022quantifying};
+they studied the composition of professor hiring in the years 2011--2021, and publicly released data on the number of professors (not lecturers) hired at US universities \citep{wapman2022zenodo}.\footnote{
     These data are not panel data, instead only give the total professors hired over the ten-year period 2011--2021.

### 2024-04-30 text/sections/07-conclusion.tex
Added new robustness check tables.

@@
 This work contributes to the literature along two primary dimensions.
-First, this work provides an explanation for the increased reliance on lecturers and away from tenure-track and tenured professors at US public universities, by isolating the effects funding cuts from state-wide funding shifts in both the short-- and long-run.
+First, this work provides an explanation for the increased reliance on lecturers and away from tenure-track and tenured professors at US public universities, by isolating the effects funding cuts from state-wide funding shifts in both the short-- and long--run.
 Second, this work investigates how individual faculty are affected by changes in state funding for their university using a dataset new in the economic literature (IBHED).

### 2024-04-30 text/sections/09-figures-tables.tex
Added new robustness check tables.

@@
     \centering
-    \caption{Effects of Changes in State Funding on University Faculty Composition, IPEDS 1990--2021, OLS and 2SLS Estimates.}
+    \caption{Effects of State Funding on Faculty Counts, IPEDS 1990--2021, OLS and 2SLS Estimates.}
 
@@
     Panel B uses $\log$ faculty count per student as the outcome, though the outcome mean is count of faculty per student (not in $\log$ terms).
-    Standard errors are clustered at the state-year level, and university $+$ year fixed effects are included through--out.
+    Standard errors are clustered at the state-year level, and university $+$ year fixed effects are included through-out.
 \end{table}
@@
     The panels shows the effect of a $1$\% change in state funding per student at the university on faculties' salaries (base salary $+$ benefits), promotion rate (e.g., assistant professor to associate professor), and rate of leaving their university (i.e., by quitting, retiring, or moving to another university).
-    Standard errors are clustered at the university-year of hire level, and university $+$ year of hire fixed effects are included through--out.
+    Standard errors are clustered at the university-year of hire level, and university $+$ year of hire fixed effects are included through-out.
 \end{table}

### 2024-04-30 text/sections/10-appendix.tex
Added new robustness check tables.

@@
 The formulation here takes as given that the professor is already employed at the university, and then projects the effect of changes in state funding on these \textit{incumbent} professors following the state funding shift-share.
-\autoref{tab:firststage-illinois} presents the first-stage results in Illinois data .
+\autoref{tab:firststage-illinois} presents the first-stage results in Illinois data.
 
@@
 
+\newpage
+\subsection{IPEDS Robustness Checks}
+
+\begin{table}[H]
+    \singlespacing
+    \centering
+    \caption{Effects of State Funding on Faculty Counts, IPEDS 1990--2021, IV Estimates by Institution Selectivity.}
+
+    \textbf{Panel A: units in \$ per student}
+
+    \makebox[\textwidth][c]{
+        \begin{tabular}{@{\extracolsep{5pt}}lcccccl} 
+            \\[-1.8ex]\hline 
+            \hline \\[-1.8ex]
+            & First-stage & Lecturers & Asst. Professors & Full Professors & All Faculty & Observations \\ 
+            \cline{2-7} 
+            \\[-1.8ex]
+            \input{tables/facultycount-rawcounts-heterogeneity.tex}
+            \\[-1.8ex] \hline 
+            \hline 
+        \end{tabular}}
+
+        \vspace{0.2cm}
+    \textbf{Panel B: units in log \$ per student}
+
+    \makebox[\textwidth][c]{
+        \begin{tabular}{@{\extracolsep{5pt}}lcccccl} 
+            \\[-1.8ex]\hline 
+            \hline \\[-1.8ex]
+            & First-stage & Lecturers & Asst. Professors & Full Professors & All Faculty & Observations \\ 
+            \cline{2-7} 
+            \\[-1.8ex]
+            \input{tables/facultycount-heterogeneity.tex}
+            \\[-1.8ex] \hline 
+            \hline 
+        \end{tabular}}
+    \label{tab:facultycount-heterogeneity}
+    \justify
+    \footnotesize
+    \textbf{Note}:
+    These tables show the IV estimates of regression specification \eqref{eqn:secondstage}, 
+    in the same manner as \autoref{tab:facultycount-shock-reg}, but restricting to institutions ranked as most selective, selective, and unranked by \cite{barrons2009}.
+    For example, Row 1 of panel B shows that a $10\%$ cut in the state shift-share leads to a fall of $9.44\%$ in state funding for the public universities ranked as ``most selective.''
+\end{table}
+
+\begin{table}[H]
+    \singlespacing
+    \centering
+    \caption{First-stage Robustness Checks for Effects of State Funding Shift-Share on State Funding, OLS Estimates.}
+    \makebox[\textwidth][c]{\input{tables/firststage-robustness-checks.tex}}
+    \label{tab:firststage-robustness-checks}
+    \justify
+    \footnotesize
+    \textbf{Note}:
+    These tables show the second stage OLS and 2SLS estimates of regression specification \eqref{eqn:firststage}, showing the effect of state funding shift-share on each university's state funding.
+    This table differs from the main analysis by replacing University $+$ Year fixed effects with State $+$ Year fixed effects, measuring enrolment in full-time equivalent (FTE), and including controls for (1)
+    State Funding as a percent of total funding in base period, (2) acceptance rate in mid-2000s, (3) 6--year completion rate in mid-2000s, (4) tuition revenue, (5) enrolment measured by full-time equivalent FTE, (6) total public university enrolment in the entire state, (7) percent of public university enrolment for the state enrolled at this university.
+    The first column uses the raw count specification, and the second column uses the $\log$ specification for percentage terms.
+\end{table}
+
+\begin{table}[H]
+    \singlespacing
+    \centering
+    \caption{Robustness Checks for Effects of State Funding Cuts on Faculty Composition, OLS and 2SLS Estimates in $\log$ Units.}
+    \makebox[\textwidth][c]{
+        %\small
+        \input{tables/facultycount-robustness-checks.tex}}
+    \label{tab:facultycount-robustness-checks}
+    \justify
+    \footnotesize
+    \textbf{Note}:
+    These tables show the second stage OLS and 2SLS estimates of regression specification \eqref{eqn:secondstage}, showing the effect of state funding changes on number of faculty per student in Illinois universities, using the funding shift-share to instrument for state funding in the columns labelled 2SLS.
+    This table differs from the main analysis by replacing University $+$ Year fixed effects with State $+$ Year fixed effects, measuring enrolment in full-time equivalent (FTE), and including controls for (1)
+    State Funding as a percent of total funding in base period, (2) acceptance rate in mid-2000s, (3) 6--year completion rate in mid-2000s, (4) tuition revenue, (5) enrolment measured by full-time equivalent FTE, (6) total public university enrolment in the entire state, (7) percent of public university enrolment for the state enrolled at this university.
+\end{table}
+
+\begin{table}[H]
+    \singlespacing
+    \centering
+    \caption{Robustness Checks for Effects of State Funding Cuts on Faculty Composition, OLS and 2SLS Estimates in Raw Count Units.}
+    \makebox[\textwidth][c]{
+        \small
+        \input{tables/facultycount-rawcount-robustness-checks.tex}}
+    \label{tab:facultycount-rawcount-robustness-checks}
+    \justify
+    \footnotesize
+    \textbf{Note}:
+    These tables show the second stage OLS and 2SLS estimates of regression specification \eqref{eqn:secondstage}, showing the effect of state funding changes on count of faculty in Illinois universities, using the funding shift-share to instrument for state funding in the columns labelled 2SLS.
+    This table differs from the main analysis by replacing University $+$ Year fixed effects with State $+$ Year fixed effects, measuring enrolment in full-time equivalent (FTE), and including controls for (1)
+    State Funding as a percent of total funding in base period, (2) acceptance rate in mid-2000s, (3) 6--year completion rate in mid-2000s, (4) tuition revenue, (5) enrolment measured by full-time equivalent FTE, (6) total public university enrolment in the entire state, (7) percent of public university enrolment for the state enrolled at this university.
+\end{table}
+
+\begin{table}[H]
+    \singlespacing
+    \centering
+    \caption{Effects of State--Wide Funding Changes on Private University Faculty Counts, IPEDS 1990--2021, OLS and 2SLS Estimates.}
+
+    \textbf{Panel A: units in \$ per student}
+
+    \makebox[\textwidth][c]{\input{tables/facultycount-shock-private-robustness-rawcount.tex}}
+
+    \textbf{Panel B: units in log \$ per student}
+
+    \makebox[\textwidth][c]{\input{tables/facultycount-shock-private-robustness.tex}}
+
+    \label{tab:facultycount-shock-private-robustness}
+    \justify
+    \footnotesize
+    \textbf{Note}:
+    These tables show the second stage OLS and 2SLS estimates of regression specification \eqref{eqn:secondstage}, among private universities --- as described in \autoref{sec:results-robustness}.
+    Each observation is a public university-year, in the IPEDS data.
+    Panel A shows the effect of a fall in state funding \$$-1,000$ per student on the number of professors --- i.e.,
+    an extra \$$1,000$ per student leads to 6 fewer lecturers according to column 2.
+    Panel B shows the effect of a $10$\% change in state funding per student at the university on the 10\% change in the number of professors per students --- i.e.,
+    an extra 10\% of state funding per student leads to 4.37\% fewer lecturers per student according to column 2.
+    Outcome-mean is the mean of the outcome, for Panel A the number of professors per student, for Panel B the number of faculty per student.
+    Panel B uses $\log$ faculty count per student as the outcome, though the outcome mean is count of faculty per student (not in $\log$ terms).
+    Standard errors are clustered at the state-year level, and university $+$ year fixed effects are included through--out.
+\end{table}
+
 %\newpage

### 2024-04-30 text/tables/firststage-balance-count.tex
Adjustments.
Wiritng updates, new title.
Added new robustness check tables.

@@
 \cline{2-9} 
-\\[-1.8ex] & Enrolment & State & Total & Non-inst. & Lecturers & Assistant & Full & All \\ 
-\\[-1.8ex] &           & Funding & Revenues & Revenues & Lecturers & Professors & Professors & Professors \\
+\\[-1.8ex] & Enrolment & State Funding & Total revenues & Non-inst. revenues & Lecturers & Assistant professors & Full professors & All professors \\ 
 \\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6) & (7) & (8)\\ 
 \hline \\[-1.8ex] 
- Funding & 0.517   & $-$9,217 & $-$7,633 & 886 & 0.007 & $-$0.001 & $-$0.003 & 0.004 \\ 
- Shock   & (0.199) & (2,452) & (12,771) & (4,093) & (0.002) & (0.001) & (0.003) & (0.005) \\ 
+ State Funding & 0.517 & $-$9,217.091 & $-$7,633.377 & 1,818.090 & 0.007 & $-$0.001 & $-$0.003 & 0.004 \\ 
+  & (0.199) & (2,451.719) & (12,771.130) & (4,774.483) & (0.002) & (0.001) & (0.003) & (0.005) \\ 
  \hline \\[-1.8ex] 
-Uni. + Year \\
-fixed effects? & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\
+Uni. + Year fixed effects? & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
 Observations & 17,012 & 17,012 & 17,012 & 17,012 & 17,012 & 17,012 & 17,012 & 17,012 \\ 
-R$^{2}$ & 0.960 & 0.950 & 0.881 & 0.932 & 0.745 & 0.886 & 0.973 & 0.954 \\ 
+R$^{2}$ & 0.960 & 0.950 & 0.881 & 0.922 & 0.745 & 0.886 & 0.973 & 0.954 \\ 
 \hline 

### 2024-04-30 programs/analysis/ibhed-shock-longterm.R

@@ -24,6 +24,15 @@ digits.no <- 3
 fig.width <- 10
 fig.height <- fig.width * 0.85
 
+# Define default colours.
+colour.list <- c(
+    "#D62728", # Red  -> First-stage.
+    "#1f77b4", # Blue -> Lecturers
+    "#c9721c", # Orange -> Asst Professors
+    "#237e23", # Green -> Full Professors
+    "#8300c0", # Dark purple -> Administrator Professors
+    "#cc2c9f") # Strong pink -> All faculty.
+
 
 # Load data --------------------------------------------------------------------
 
@@ -208,7 +217,7 @@ lp.data <- reg.data %>%
     pdata.frame(index = c("name_unitid", "year"), drop.index = FALSE)
 
 ## Define a function to nicely plot the LP results.
-lpreg.plot <- function(model.lpreg) {
+lpreg.plot <- function(model.lpreg, given.colour) {
     ## Input a LP-reg (estimates an elesticity \in [-1,1])
     ## And the time.hirozon for how far to plot LP model
     # Get the estimates
@@ -220,18 +229,17 @@ lpreg.plot <- function(model.lpreg) {
     # Plot the LP etsimation, with my own custom plot.
     model.plot <- model.lpdata %>%
         ggplot(aes(x = t)) +
-        geom_hline(yintercept = 0, alpha = 0.5) +
+        geom_hline(yintercept = 0, alpha = 0.25) +
         geom_ribbon(aes(ymin = conf.low, ymax = conf.high),
-            alpha = 0.4, fill = "grey70") +
-        geom_point(aes(y = estimate), colour = "blue") +
-        geom_line(aes(y = estimate), colour = "blue") +
-        geom_line(aes(y = conf.low), linetype = "dashed") +
-        geom_line(aes(y = conf.high), linetype = "dashed") +
+            alpha = 0.25, fill = given.colour) + #"grey70") +
+        geom_point(aes(y = estimate), colour = given.colour) +
+        geom_line(aes(y = estimate), colour = given.colour) +
+        geom_line(aes(y = conf.low), linetype = "dashed", colour = given.colour) +
+        geom_line(aes(y = conf.high), linetype = "dashed", colour = given.colour) +
         scale_x_continuous(name = "Years, Relative to Funding Cut",
[…]

### 2024-04-30 programs/analysis/ibhed-shock-rawcounts.R

@@ -21,6 +21,14 @@ digits.no <- 3
 fig.width <- 10
 fig.height <- fig.width * 0.85
 
+# Define default colours.
+colour.list <- c(
+    "#D62728", # Red  -> First-stage.
+    "#1f77b4", # Blue -> Lecturers
+    "#c9721c", # Orange -> Asst Professors
+    "#237e23", # Green -> Full Professors
+    "#cc2c9f") # Strong pink -> All faculty.
+
 
 # Load data --------------------------------------------------------------------
 

### 2024-04-30 programs/analysis/ibhed-shock.R

@@ -14,6 +14,14 @@ set.seed(47)
 # This file follows an adjusted Deming Walters (2017, p.10) approach to
 # appropriations shock instrument.
 
+# Define default colours.
+colour.list <- c(
+    "#D62728", # Red  -> First-stage.
+    "#1f77b4", # Blue -> Lecturers
+    "#c9721c", # Orange -> Asst Professors
+    "#237e23", # Green -> Full Professors
+    "#cc2c9f") # Strong pink -> All faculty.
+
 # Define number of digits in tables and graphs
 digits.no <- 3
 
@@ -265,8 +273,8 @@ salary_density.plot <- reg.data %>%
         year %in% c(2010, 2021)) %>%
     mutate(faculty =
         ifelse(position %in% c("Instructor", "Lecturer"), "Lecturer",
-        ifelse(position == "Assistant Professor", "Tenure-Track",
-        ifelse(position %in% c("Associate Professor", "Professor"), "Tenured",
+        ifelse(position == "Assistant Professor", "Asst. Professor",
+        ifelse(position %in% c("Associate Professor", "Professor"), "Full Professor",
         ifelse(position %in% c("Chancellor",
             "Other Administrator", "President", "Senior Officer",
             "Unit Director", "Vice Chancellor", "Vice President"), "Administrator",
@@ -288,7 +296,10 @@ salary_density.plot <- reg.data %>%
         plot.title = element_text(size = rel(1)),
         plot.margin = unit(c(0.5, 0, 0, 0), "mm"),
         legend.position = "bottom",
-        legend.margin = margin(t = -5))
+        legend.margin = margin(t = -5)) +
+    scale_fill_manual(values = colour.list) +
+    scale_colour_manual(values = colour.list)
+
 # Save this plot
 presentation.width <- 17.5
 presentation.height <- presentation.width / 2

### 2024-04-30 programs/analysis/ipeds-shock-longterm.R

@@ -19,6 +19,14 @@ digits.no <- 3
 fig.width <- 10
 fig.height <- fig.width * 0.85
 
+# Define default colours.
+colour.list <- c(
+    "#D62728", # Red  -> First-stage.
+    "#1f77b4", # Blue -> Lecturers
+    "#c9721c", # Orange -> Asst Professors
+    "#237e23", # Green -> Full Professors
+    "#cc2c9f") # Strong pink -> All faculty.
+
 
 # Load data sources ------------------------------------------------------------
 
@@ -80,8 +88,9 @@ reg.data <- reg.data %>%
     filter(1990 <= year,
         !is.na(enrollment_reported), enrollment_reported > 0,
         !is.na(all_prof_count),
-        !is.na(stateappropriations_real),
-        !is.na(appropriationshock_perEnroll_real))
+        !is.na(stateappropriations_real), stateappropriations_real > 0,
+        !is.na(appropriationshock_perEnroll_real),
+        appropriationshock_perEnroll_real > 0)
 
 # Define data sample, with releavnt transformations, for the LP estimation
 reg.data <- reg.data %>%
@@ -121,7 +130,7 @@ time.horizon <- 11
 limits.vector <- c(-0.1, 0.6)
 
 ## Define a function to nicely plot the LP results.
-lpreg.plot <- function(model.lpreg, limits.vector) {
+lpreg.plot <- function(model.lpreg, limits.vector, given.colour) {
     ## Input a LP-reg (estimates an elesticity \in [-1,1])
     ## And the time.hirozon for how far to plot LP model
     # Get the estimates
@@ -133,13 +142,13 @@ lpreg.plot <- function(model.lpreg, limits.vector) {
     # Plot the LP etsimation, with my own custom plot.
     model.plot <- model.lpdata %>%
         ggplot(aes(x = t)) +
-        geom_hline(yintercept = 0, alpha = 0.5) +
+        geom_hline(yintercept = 0, alpha = 0.25) +
         geom_ribbon(aes(ymin = conf.low, ymax = conf.high),
-            alpha = 0.4, fill = "grey70") +
-        geom_point(aes(y = estimate), colour = "blue") +
-        geom_line(aes(y = estimate), colour = "blue") +
-        geom_line(aes(y = conf.low), linetype = "dashed") +
-        geom_line(aes(y = conf.high), linetype = "dashed") +
[…]

### 2024-04-30 programs/analysis/ipeds-shock-rawcounts.R

@@ -193,7 +193,7 @@ stargazer(
     model.names = FALSE,
     omit = "factor|count|year",
     intercept.bottom = TRUE,
-    covariate.labels = c("Funding Shock", "Tuition Revenue", "Constant"),
+    covariate.labels = c("Funding Shift-Share", "Tuition Revenue", "Constant"),
     omit.stat = c("LL", "ser", "aic", "wald", "adj.rsq", "f"),
     add.lines = list(
         c("Uni. + Year fixed effects?", "Yes", "No", "Yes", "No"),
@@ -313,7 +313,7 @@ stargazer(
     naive_all_count.reg, shiftshare_all_count.reg,
     add.lines = outcome.means,
     dep.var.caption =
-        "Dependent Variable: Faculty Count per 100 Students, by Position",
+        "Dependent Variable: Faculty Count per 1,000 Students, by Position",
     dep.var.labels = c(
         "Lecturers", "Asst. Professors", "Full Professors", "All Faculty"),
     column.labels = rep(c("OLS", "2SLS"), 4),
@@ -387,4 +387,4 @@ substitution_rawcount.plot <-
 # Save this plot
 ggsave("../../text/figures/substitution-rawcount-plot.png",
     plot = substitution_rawcount.plot,
-    units = "cm", dpi = 300, width = 1.25 * fig.width, height = 0.85 * fig.height)
+    units = "cm", dpi = 300, width = 1.25 * fig.width, height = 0.8 * fig.height)

### 2024-04-30 programs/analysis/ipeds-shock.R

@@ -1,6 +1,6 @@
 #!/usr/bin/R
 ## Senan Hogan-Hennessy, 3 August 2022
-## IV for Prof's salaries, using IPEDS data.
+## IV for faculty counts, using IPEDS data.
 print(Sys.time())
 library(jtools) # Plotting estimates.
 library(tidyverse) # Functions for data manipulation and visualization
@@ -32,6 +32,14 @@ lag_lead <- function(column, t) {
     }
     return(lag_column)
 }
+
+# Define default colours.
+colour.list <- c(
+    "#D62728", # Red  -> First-stage.
+    "#1f77b4", # Blue -> Lecturers
+    "#c9721c", # Orange -> Asst Professors
+    "#237e23", # Green -> Full Professors
+    "#cc2c9f") # Strong pink -> All faculty.
 
 
 # Load data sources ------------------------------------------------------------
@@ -836,7 +844,7 @@ stargazer(
     model.names = FALSE,
     omit = "factor|count|year",
     intercept.bottom = TRUE,
-    covariate.labels = c("Funding Shock", "Tuition Revenue", "Constant"),
+    covariate.labels = c("Funding Shift-Share", "Tuition Revenue", "Constant"),
     omit.stat = c("LL", "ser", "aic", "wald", "adj.rsq", "f"),
     add.lines = list(
         c("Uni. + Year fixed effects?", "Yes", "No", "Yes", "No"),
@@ -884,8 +892,8 @@ lag.plot <- lag.data %>%
     geom_hline(yintercept = 0, alpha = 0.5) +
     geom_ribbon(aes(ymin = conf.low, ymax = conf.high),
         alpha = 0.4, fill = "grey70") +
-    geom_point(aes(y = estimate), colour = "blue") +
-    geom_line(aes(y = estimate), colour = "blue") +
+    geom_point(aes(y = estimate), colour = colour.list[1]) +
+    geom_line(aes(y = estimate), colour = colour.list[1]) +
     geom_line(aes(y = conf.low), linetype = "dashed") +
     geom_line(aes(y = conf.high), linetype = "dashed") +
     scale_x_continuous(name = "Years, Relative to Funding Cut",
@@ -1082,7 +1090,7 @@ elasticity.plot <- plot_summs(
 # Save this plot
 ggsave("../../text/figures/elasticity-plot.png",
     plot = elasticity.plot,
[…]

### 2024-04-30 programs/data-build/urban-ipeds-build.R

@@ -2,9 +2,9 @@
 ## Senan Hogan-Hennessy, 21 December 2021
 ## Build IPEDS data with the Urban Inst's package
 print(Sys.time())
-## https://educationdata.urban.org/documentation/colleges.html
-## https://urbaninstitute.github.io/education-data-package-r/
-library(tidyverse) ## functions for data manipulation and visualization
+# https://educationdata.urban.org/documentation/colleges.html
+# Load functions for data manipulation and visualisation
+library(tidyverse)
 set.seed(47)
 
 
@@ -35,6 +35,9 @@ salaries.data <- read_csv("../../data/urban-ipeds/raw-data/faculty/ipeds-salarie
 tenure.data <- read_csv("../../data/urban-ipeds/raw-data/faculty/ipeds-tenure.csv")
 # Employee part- + full-time data (hand made from raw IPEDS values)
 parttime.data <- read_csv("../../data/urban-ipeds/raw-data/parttime-employees/ipeds-parttime.csv")
+
+# Barrons (2009) selctivity rankings (if uni not in the list, then is rank 5)
+barrons.data <- read_csv("../../data/rankings/barrons-2009.csv")
 
 # CPI-U from FREDS (since Urban provided cpi has missing years)
 # Yearly average, seasonally adjusted, base year 1982-1984=100
@@ -293,6 +296,41 @@ urban_ipeds.data <- directory.data %>%
     left_join(graduation150.data, by = c("unitid", "year")) %>%
     left_join(parttime.data, by = c("unitid", "year"))
 
+## connect the Barron's selectivity index by fuzy name match
+library(fuzzyjoin)
+# Get the connecting variables for IPEDs: unitid, uni name
+ipeds.key <- urban_ipeds.data %>%
+    group_by(unitid, inst_name) %>%
+    summarise(count = n()) %>%
+    ungroup() %>%
+    select(unitid, inst_name)
+# Fuzy match the names between IPEDS + Barron's measure, avoid string cleaning
+fuzzy.data <- barrons.data %>%
+    select(inst_name, barrons_rank_2009) %>%
+    stringdist_join(ipeds.key,
+        by = "inst_name",
+        mode = "left",
+        method = "jw",
+        max_dist = 0.1,
+        distance_col = "dist")
+# Take the best string match for each uni
+fuzzy.key <- fuzzy.data %>%
+    rename(inst_name = inst_name.x) %>%
+    group_by(inst_name) %>%
[…]

### 2024-04-30 programs/run.sh

@@ -21,6 +21,8 @@ R CMD BATCH --no-save ipeds-public-private.R
 R CMD BATCH --no-save ipeds-shock.R
 # Analysis and regressions at the institution-level, IPEDS, with count outcomes.
 R CMD BATCH --no-save ipeds-shock-rawcounts.R
+# Robustness checks at the institution-level, IPEDS
+R CMD BATCH --no-save ipeds-robustness-checks.R
 # Analysis and regressions at the institution-level, IPEDS, local porjections.
 R CMD BATCH --no-save ipeds-shock-longterm.R
 # Analysis and regressions at the individual-level, IBHED, base-year shock IV.

### 2024-05-08 text/paper.tex
Newest revision + all code updated.

@@
 \usepackage{subcaption}
-% caption graphics + tables (with name in bold)
-\usepackage{tikz}                                                          % Draw graphs inline, guide: https://sites.google.com/site/kochiuyu/Tikz
-\usetikzlibrary{shapes,decorations,arrows,calc,arrows.meta,fit,positioning}
-\tikzset{
-    -Latex,auto,node distance =1 cm and 1 cm,semithick,
-    state/.style ={ellipse, draw, minimum width = 0.7 cm},
-    point/.style = {circle, draw, inner sep=0.04cm,fill,node contents={}},
-    bidirected/.style={Latex-Latex,dashed},
-    el/.style = {inner sep=2pt, align=left, sloped}
-}
-\usepackage{pgfplots} \pgfplotsset{compat=1.16}                            % Automatic graphing, read https://www.overleaf.com/learn/latex/pgfplots_package for examples
 % Maths + numbers
@@
 % Dicussion
-\input{sections/06-discussion.tex}
+%\input{sections/06-discussion.tex}
 % Conclusion

### 2024-05-08 text/sections/05-results.tex
Newest revision + all code updated.

--- 2024-02-28 text/sections/05-results.tex
@@
 
-The first stage is strong in both IPEDS and IBHED data, and the argument in \autoref{sec:empirics} for exogeneity comes from independence of the state funding mechanism to any individual public university campus.
+The first stage is strong in both IPEDS and IBHED data, and the argument for exogeneity comes from independence of the state funding mechanism to any individual public university campus (as in \autoref{sec:empirics}).
 The IV approach also assumes that shifts in state-wide funding for higher education (the instrument) affect outcomes only via affecting state funding for the university (exclusion restriction).
@@
 \subsection{University Level, IPEDS}
+\label{sec:results-ipeds}
 Funding cuts lead to measurable changes in the number of faculty at US public universities.
@@
 Once the funding cut has passed, universities do not re-adjust to hiring more professors to make up for last year's substitution towards lecturers.
-\autoref{fig:count-lp} shows local projection estimates, where the estimate is the elasticity for professor count per student in year $t+k$ with respect to state funding in year $t$ for later years $k = 0, \hdots, 10$.
-These estimates correlespond to the elasticity estimates in \autoref{tab:facultycount-shock-reg} Panel B, where the outcome is faculty count per students in later years.
-State funding cuts affectthe number of lecturers and professors at universities up to 9 years after the initial funding cut.
-A 10\% state funding cut leads to 4.2\% (2.9--5.5 in a 95 percent confident interval) more lecturers per student the following year, and around 3.1\% (0.6--5.5 ``'') more lecturers nine years later.
-A 10\% state funding cut leads to 1.5\% (0.6--2.2 in a 95 percent confident interval) fewer assistant professors per student the following year, and around 1.1\% (0.2--2.0 ``'') fewer nine years later;
-the same cut leads to 1.2\% (0.6--1.8 in a 95 percent confident interval) fewer full professors per student the following year, and around 0.9\% (0.1--1.7 ``'') fewer nine years later.
-The effects are a substitution with the composition of faculty, as the total number of faculty per students are relatively unaffected in the follow-on years.
+\autoref{fig:count-lp} shows local projection estimates, where the estimate is the elasticity for professor count per student in year $t+k$ with respect to state funding in year $t$.
+These elasticity estimates are shown in \autoref{tab:facultycount-shock-reg} Panel B, where the outcome is faculty count per students in later years.
+State funding cuts affect the number of lecturers and professors at universities up to 9 years after the initial funding cut.
+A 10\% state funding cut leads to 4.2\% (2.9--5.5 in a 95 percent confidence interval) more lecturers per student the following year, and around 3.1\% (0.6--5.5 confidence interval) more lecturers nine years later.
+A 10\% state funding cut leads to 1.5\% (0.6--2.2 in a 95 percent confidence interval) fewer assistant professors per student the following year, and around 1.1\% (0.2--2.0 confidence interval) fewer nine years later;
+the same cut leads to 1.2\% (0.6--1.8 in a 95 percent confidence interval) fewer full professors per student the following year, and around 0.9\% (0.1--1.7 confidence interval) fewer nine years later.
+
+These effects are a substitution of professors for lecturers, as the total number of faculty per students are relatively unaffected in the follow-on years, while numbers of lecturers increase and professors decreases.
+This agrees with a model for universities departments having fixed teaching requirements making human-resources decisions under constrained resources (e.g., \citealt{abe2015implications}), lending credibility to this type of model for understanding recent human resources decisions of public universities.
+Furthermore, if a university must teach more students, but cannot afford to hire a full-time professor who teaches 2--3 classes a year, then it makes sense that they hire 2--3 additional lecturers on part-time contracts as a replacement.
+This provides simple intuition for why the lecturers' elasticity is 2--3 times larger than that for assistant/full professors in \autoref{tab:facultycount-shock-reg}.
+While no previous research causally connects funding cuts to substituting professors for lecturers, the phenomenon has been alluded to in popular media \citep{wiu2016} and university rankings (where more lecturers hurts a university's ranking, \citealt{usnews2023}); these results provides real evidence for these claims.
 
@@
 \autoref{tab:faculty-shock-illinois-rolling} show estimates of a 10\% change in state funding per student on faculty salaries, rate of exiting the Illinois public university system, and rate of promotion.
-The estimates are not discernible from zero, so that faculty who are already at the university (incumbent faculty) seem to have little or no state funding cuts passed on to them for these outcomes.
-Additionally, local projection estimates show that professors' salaries are also not affected in the years after a state funding cut (\autoref{fig:salaries-illinois-lp-rolling}).
-Similarly, promotion rate and exit rate among incumbent faculty (from the Illinois public university system) are unaffected in the years after an initial funding shift-share (\autoref{fig:promoted-illinois-lp-rolling}, \ref{fig:exit-illinois-lp-rolling}).
-There is evidence that lecturers see a 0.75\% increase in salary in the second and third year after a 10\% increase in state funding (\autoref{fig:salaries-lecturer-illinois-lp-rolling}), though this is the only evidence that incumbent faculty are affected by state funding cuts in the Illinois public university system.
+The estimates are not discernible from zero in any specification, including the shift-share IV estimates, so that faculty seem to have little or no state funding cuts passed on to them for these outcomes.
 
-These findings show that public universities increase (decrease) their count of tenure-track and tenured professors per student in years when revenues are more (less) plentiful.
-In the same vein, when funding increased, count of tenure-track and tenured professors per student increased --- and the opposite in years of funding cuts.
-These findings are in line with the observation that universities froze hiring for a couple of years after the 2008 recession and funding cuts \citep{turner2014impact}.
-For lecturers, we see a negative effect which lines up with two trends noted in \autoref{sec:trends}: state funding for public universities (per student) stagnated while usage of lecturers increased substantially.
+It could be the case that effects on faculty do not materialise until multiple years after a funding cut, though the data do not support this possibility.
+Local projection estimates show that professors' salaries, including salaries, are not significantly affected in years following funding cuts (\autoref{fig:salaries-illinois-lp-rolling}) --- they do not receive lower rates of pay raises, or delayed salary cuts, in response to state funding cuts.
+Similarly, promotion rate and rate of exit from the Illinois public university system are also unaffected in later years (\autoref{fig:promoted-illinois-lp-rolling}, \ref{fig:exit-illinois-lp-rolling}).
+Lecturers receive a 0.75\% increase in salary in the second and third year after a 10\% increase in state funding (\autoref{fig:salaries-lecturer-illinois-lp-rolling}), though they are they only faculty who saw any effect of state funding cuts in terms of salaries.
 
-%\subsection{Robustness checks}
-%\textbf{Describe robustness checks in one or two paragraphs here; see how Lovenheim$+$ did this section.}
+These results show that faculty --- both lecturers and professors --- are relatively unaffected by state funding cuts, in the Illinois public university system.
+Illinois is a state with a large public university system, relative to student enrolment, with trends in state funding for education similar to national trends, and with both selective and non-selective universities (\autoref{sec:trends-illinois}).
+If faculty are unaffected in Illinois, then it is unlikely that faculty in the rest of the nation are affected on average.
+Note, however, that this analysis studies faculty who are already faculty.
+The IBHED data do not measure the entire pool of academics applying for faculty jobs at Illinois universities, but only those who applied to, were accepted, and then started a job as a lecturer or professor.
+These are incumbent faculty, and it is clear they are unaffected by state funding cuts.
+This raises a question: how can funding cuts lead to a large substitution of professors for lecturers, when incumbent faculty are not affected?
 
-\subsection{Reduced Hiring of New Professors}
-The results above present a condundrum: there are large substitution effects among faculty, but faculty themselves are unaffected by funding cuts.
-The analysis of Illinois faculty (IBHED) studies the effect of state funding cuts on faculty already at the university --- i.e., incumbent faculty.
-The findings above imply that subsitution towards lecturers came about by limiting hiring of new professors, to replace leaving or retiring professors, so that they are replaced by hiring lecturers.
+\subsection{Impacts on Faculty Hiring}
+Turn-over is the primary channel that changes in faculty composition happen, where, where faculty exit their jobs and new faculty are hired to replace those leaving --- or no more are hired to replace those who left.
+\autoref{fig:exit-illinois-lp-rolling} shows that Illinois faculty are no more likely to exit the Illinois system (either retire or take a new job elsewhere) following state funding cuts, so there is one remaining channel for the substitution of professors for lecturers: increased hiring of lecturers and reduced hiring of new professors.
 
-To further study this issue, I take recent data on the total faculty hired at US public universities over 2011--2021 provided by \cite{wapman2022quantifying,wapman2022zenodo}.
-\cite{wapman2022quantifying} study the composition of professor hiring in the years 2011--2021, and publiccally released data on the number of professors (not lecturers) hired at US universities.\footnote{
-    These data are not panel data, instead only give the total professors hired over the ten year period 2011--2021.
+To investigate reduced hiring of new professors, I first investigate faculty hiring rates among Illinois universities from the IBHED data.
+To do so, I count the number of faculty new in the panel of all Illinois public university faculty for years 2011--2021, where the panel first year is 2010.
+I then use this as the outcome in OLS and shift-share IV regressions, where each observation is an Illinois university--year and data for state funding come from IPEDS.\footnote{
+    This is the same specification as the IPEDS analysis restricted to Illinois universities.
+    The IPEDS data have yearly faculty counts, but not discern between faculty promotions, quits, and hires; the Illinois data can do so thanks to the panel structure of the data.
 }
-I combine this measure, with total state funding for public universities acorss the same time period from IPEDS, to validate the claim that state funding cuts led to faculty subtitution via limiting hiring of new professors.
+To study whether these effects hold true for nation-wide hiring of new professors, I use public on the number of professors at all US public universities for the entire decade 2011--2021, and perform the same analysis to see if state funding affected the number of professors hired.
 
-There is a strong correlation between the number of professors hired (relative to student enrolment), where an extra \$1,000 of state funding per student is associated with the university hiring 0.12 more professors per hundred students.
-I apply the shift-share model to these data, summing the state funding and funding shifts for the decade of 2012--2021.
-I find that a state funding cut of 10\% leads to 13\% fewer professor hires (relative to enrolment).
-See \autoref{sec:appendix-hiring} for further details on this additional analysis of professor hiring.
+\autoref{tab:facultyhires-illinois-reg} shows these estimates for Illinois public universities.
+In general, faculty counts are relatively unrelated to state funding among Illinois universities; though the point estimates line up with analysis in IPEDS, the smaller number of observations mean that the precision of estimates are small and indistinguishable from zero, in bith raw counts and in log units.
+On the other hand, universities with more state funding hire more professors across 2011--2021, as seen in \autoref{fig:hiring-correlation}.
+An extra 10\% in state funding is associated with 8.5\% more professors hired per students.
+I use these data as part of a shift-share analysis, showing that a 10\% state funding cut per student leads to 13\% fewer professor hires (relative to enrolment), and that these hires are roughly equal for gender of professor hired.
 
-These results provide suggestive evidence that funding cuts caused faculty substitution by disrupting hiring of new professors at public universities.
-This provides a coherent mechanism for how incumbent faculty (those who are already hired at the university) were not meaningfully impacted by funding cuts, yet their wider composition nonetheless changed when their universities experienced funding cuts.
+This shows that state funding is related in to professor hiring, and that the persistent funding cuts to public universities led to lower professor hiring across this time period.
+The other turn-over channels were unaffected by state funding cuts, and that professor hiring is related to professor hiring in both correlational terms and in a shift-share analysis.
+This means that effects on faculty hiring were likely the main mechanism for how state funding cuts led to a substitution of professors for lecturers at public universities.
+
+\subsection{Robustness Checks}
+\label{sec:results-robustness}
+It is possible that the identification startegy is not robust to other possible confounders.
+For example, it is possible that selective institutions received less funding for other reasons, or states with lower performing insitutions or declining populations strategically limited funding for less selective universities.
+I address these main concerns with a number of different analyses.
+
+First, I re-estimate the primary results from \autoref{sec:results-ipeds} with additional controls for university selectivity.
+To be specific, I estimate \autoref{eqn:secondstage} with controls added for the base share of state funding, average acceptance rate and 6-year completion rate  in 2004--2006, tuition revenue, total enrolment, size of state enrolment, size of university enrolment relative to the rest of the state.
+Measures of selectivity are only available in certain years, so I use these values across the entire panel.
+This means that insitution fixed effects would over-identify the system, so I estimate with state $+$ year fixed effects (not intituion $+$ year fixed effects).
+In this way, this model also works as an alternative specification test.
+Additionally, this analysis uses the enrolment measure of full-time equivalent (FTE), and not reported enolment as in the original analysis.
+The first-stage is strong of the same order as the orginal analysis (\autoref{tab:firststage-robustness-checks}).
+The relationship between state funding, in both raw count terms and log units, are roughly the same as for the main analysis (\autoref{tab:facultycount-robustness-checks}, \ref{tab:facultycount-rawcount-robustness-checks}), despite the multiple specification changes, additional controls, and different measures for enrolment.
+
+Secondly, I return to the primary specification in \autoref{eqn:secondstage}, and estimate effect of state funding cuts on faculty counts separately among universities of different selectivity rates.
+This investigates whether results are consistent across different levels of selectivity, and whether less selective universities have unreasonably high estimates for substitution --- as would be the case if selectivity was an unaddressed confounder.
+\autoref{tab:facultycount-heterogeneity} collects these results together, separated by the \cite{barrons2009} selectivity rankings.
+Barrons rank 1-2 are the most selective (including e.g., UC Berkeley and  Berkeley and SUNY Binghamton), 3-4 selective (including e.g., Clemson University and University of Dallas), and rank 5 for the rest of universities.
+The first-stage is string in every group of universities, even the most selective group which is only 12 universities, the selective group which is 28, and the rest are 762 universities.
+Faculty substitution rates are not very string among the most selective universities, but are very similar between selective universities and unranked universities, showing that even selective public universities exhibit the substitution towards lecturers.
+
+The final robustness check I employ is a falsification test, testing whether the effects for public universities would be picked up among the sample of private universities (as in \citealt{NBERw27885}).
+Private universities do not rely on state funding in the same that public universities do, though they in some states do receive some state funding for projects and other smaller scale operating costs.
+Given this low reliance, it is unclear how state-wide funding shifts would affect faculty outcomes, and thus we would expect the two to be unrelated.
+If they were related, this would be evidence that the IV model picks up otherwise concurrent trends in higher education, unrelated to stagnating state funding for public universities.
+To impute a base share of reliance on state funding, I assign private universities the average reliance on state funding in the base period of universities in the same state with the same Barron's selectivity ranking, and then compute the shift-share instrument from this value.
+\autoref{tab:facultycount-shock-private-robustness} shows that the first-stage is not strong among private universities, and the IV results are significant for any faculty count.
+Notably, the OLS results for correlation between state funding and professor are highly statistically significant, showing that more funding is related to more lecturers and faculty in general, though the IV analysis shows that these relationships are spurious among private universities.

### 2024-05-08 text/sections/07-conclusion.tex
Newest revision + all code updated.

@@
 \label{sec:conclusion}
-Public universities have systematically substituted away from tenure-track and tenured professors, toward lecturers, in the face of persistent declines in state funding for higher education.
+Public universities have systematically substituted away from tenure-track and tenured professors, toward lecturers, following stagnating state funding for higher education.
 The effects on incumbent professors in Illinois are non-distinguishable from zero, and suggestive evidence shows that funding cuts limited hiring of new professors, implying that substitution towards lecturers is driven by reduced hiring of new professors.
@@
 These results show large changes in faculty composition, and that stagnation in state support explains at least a third of the observed shift away from tenured professors and towards lecturers.
-This likely had knock-on effects on instruction, contribiting to worsening student outcomes at US public universities since the early 1990s --- while private universities were not exposed to financial headwinds of the same magnitude or persistence over the same time period.
+This likely had knock-on effects on instruction, contributing to worsening student outcomes at US public universities since the early 1990s --- while private universities were not exposed to financial headwinds of the same magnitude or persistence over the same time period.
+Furthermore, public universities have been making personnel decisions, with long-run effects on their institutional capacity, based on short run funding constraints.
 While public universities educate the majority of US higher education students, we should worry about the effects of restricting their funding has on faculty, and the wider impact on higher education as a whole.

### 2024-05-08 text/sections/10-appendix.tex
Newest revision + all code updated.

@@
     \caption{Correlation Between State Funding Shift-Share and Public University State Funding in Surrounding Years.}
-    \includegraphics[width=\textwidth]{figures/lag-firststage.png}
+    \includegraphics[width=0.75\textwidth]{figures/lag-firststage.png}
     \label{fig:lag-firststage}
@@
 
+\begin{figure}[H]
+    \centering
+    \singlespacing
+    \caption{Correlation Between State Funding and Total COunt Professors Hired, 2011--2021.}
+    \includegraphics[width=\textwidth]{figures/hiring-correlation.png}
+    \label{fig:hiring-correlation}
+    \justify
+    \footnotesize
+    \textbf{Note}:
+    This figure shows the correlation between total state funding per student and total professors hired, 2010--2021.
+    Funding data are taken from IPEDS, and data on professor hiring  from \cite{wapman2022quantifying,wapman2022zenodo}.
+\end{figure}
+
 \begin{table}[h!]
@@
 \newpage
-\subsection{IPEDS Robustness Checks}
+\subsection{Robustness Checks}
 
@@
     These tables show the IV estimates of regression specification \eqref{eqn:secondstage}, 
-    in the same manner as \autoref{tab:facultycount-shock-reg}, but restricting to institutions ranked as most selective, selective, and unranked by \cite{barrons2009}.
+    in the same manner as \autoref{tab:facultycount-shock-reg}, but restricting to institutions ranked as most selective, selective, and unranked by \cite{barrons2009} --- and including state
+    The first column presents the coefficient between state funding and shift-share IV, which is a string first-stage among every level of selectivity for public universities.
+    The other columns show the coefficient between state funding the count of faculty per students.
     For example, Row 1 of panel B shows that a $10\%$ cut in the state shift-share leads to a fall of $9.44\%$ in state funding for the public universities ranked as ``most selective.''
+    Standard errors for the coefficient estimates are in brackets, and the outcome mean are in square brackets beneath.
 \end{table}

### 2024-05-08 text/tables/facultycount-illinois-reg-rawcount.tex
Newest revision + all code updated.

@@
 \cline{2-9} 
-\\[-1.8ex] & \multicolumn{2}{c}{Lecturer} & \multicolumn{2}{c}{Assistant} & \multicolumn{2}{c}{Full} & \multicolumn{2}{c}{All} \\ 
+\\[-1.8ex] & \multicolumn{2}{c}{Lecturers} & \multicolumn{2}{c}{Asst. Professors} & \multicolumn{2}{c}{Full Professors} & \multicolumn{2}{c}{All Faculty} \\ 
  & OLS & 2SLS & OLS & 2SLS & OLS & 2SLS & OLS & 2SLS \\ 

### 2024-05-08 text/tables/facultyhires-illinois-reg-rawcount.tex
Revision 2 finished (though needs check over of writing).
writing edits made.
Newest revision + all code updated.

--- 2023-07-18 text/tables/facultycount-illinois-reg-rawcount.tex
@@
 \hline \\[-1.8ex] 
- & \multicolumn{8}{c}{Dependent Variable: Employment Count by Professor Group} \\ 
+ & \multicolumn{8}{c}{Dependent Variable: Yearly New Hires by Professor Group} \\ 
 \cline{2-9} 
-\\[-1.8ex] & \multicolumn{2}{c}{Lecturer} & \multicolumn{2}{c}{Assistant} & \multicolumn{2}{c}{Full} & \multicolumn{2}{c}{All} \\ 
+\\[-1.8ex] & \multicolumn{2}{c}{Lecturers} & \multicolumn{2}{c}{Asst. Professors} & \multicolumn{2}{c}{Full Professors} & \multicolumn{2}{c}{All Faculty} \\ 
  & OLS & 2SLS & OLS & 2SLS & OLS & 2SLS & OLS & 2SLS \\ 
@@
 \hline \\[-1.8ex] 
- State Funding & $-$3.729 & 0.376 & 0.415 & 3.208 & $-$1.269 & 0.117 & $-$4.084 & 3.890 \\ 
-  & (1.721) & (1.725) & (1.728) & (2.493) & (1.548) & (1.200) & (4.352) & (5.668) \\ 
+ State Funding & $-$2.551 & $-$2.280 & 0.060 & $-$5.126 & $-$0.095 & $-$2.562 & $-$2.677 & 23.012 \\ 
+  & (1.420) & (9.168) & (0.675) & (7.210) & (0.249) & (7.051) & (2.185) & (52.652) \\ 
  \hline \\[-1.8ex] 
-Outcome Mean & 351.306 & 351.306 & 273.757 & 273.757 & 477.597 & 477.597 & 1303.014 & 1303.014 \\ 
-Observations & 144 & 144 & 144 & 144 & 144 & 144 & 144 & 144 \\ 
-R$^{2}$ & 0.886 & 0.882 & 0.970 & 0.969 & 0.991 & 0.991 & 0.981 & 0.980 \\ 
+Outcome Mean & 73.275 & 73.275 & 42.771 & 42.771 & 12.301 & 12.301 & 151.932 & 151.932 \\ 
+Observations & 131 & 131 & 131 & 131 & 113 & 113 & 132 & 132 \\ 
+R$^{2}$ & 0.839 & 0.839 & 0.934 & 0.902 & 0.788 & 0.752 & 0.918 & 0.793 \\ 
 \hline 

### 2024-05-08 text/tables/parttime-shock-reg-fte.tex
Newest revision + all code updated.

\begin{tabular}{@{\extracolsep{5pt}}lcccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{6}{c}{Dependent Variable: Log Lecturer Count per Students, by Position} \\ 
\cline{2-7} 
\\[-1.8ex] & \multicolumn{2}{c}{Part-time Lecturers} & \multicolumn{2}{c}{Full-time Lecturers} & \multicolumn{2}{c}{All Lecturers} \\ 
 & OLS & 2SLS & OLS & 2SLS &  &  \\ 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\ 
\hline \\[-1.8ex] 
 State Funding & $-$0.042 & $-$0.258 & $-$0.045 & $-$0.217 & 0.023 & 0.055 \\ 
  & (0.133) & (0.369) & (0.129) & (0.317) & (0.103) & (0.104) \\ 
 \hline \\[-1.8ex] 
Outcome Mean & 2.052 & 2.052 & 1.025 & 1.025 & 0.654 & 0.654 \\ 
Observations & 6,077 & 6,077 & 6,077 & 6,077 & 6,077 & 6,077 \\ 
R$^{2}$ & 0.712 & 0.711 & 0.672 & 0.671 & 0.788 & 0.788 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}

### 2024-05-08 programs/analysis/ibhed-shock-rawcounts.R

@@ -234,6 +234,8 @@ firststage_raw.fstat <-
     nth(2) %>%
     as.numeric() %>%
     round(digits.no)
+# Show that the first stage, and its strength
+summary(firststage_raw.reg)
 
 ## Lecturer faculty Count
 # Naive OLS Regression
@@ -242,14 +244,14 @@ naive_lecturer_count.reg <- institutional.data %>%
         stateappropriations_real / (1000 * enrollment_reported)) %>%
     felm(lecturer_prof_count ~ 1 +
         `I(stateappropriations_real/(1000 * enrollment_reported))(fit)` |
-        unitid |
+        unitid + year |
         0 |
         unitid + year,
         data = .)
 # Shift-share IV Regression, explained by state appropriation shock
 shiftshare_lecturer_count.reg <- institutional.data %>%
     felm(lecturer_prof_count ~ 1 |
-        unitid |
+        unitid + year |
         (I(stateappropriations_real / (1000 * enrollment_reported)) ~
             I(-appropriationshock_perEnroll_real)) |
         unitid + year,
@@ -262,14 +264,14 @@ naive_assistant_count.reg <- institutional.data %>%
         stateappropriations_real / (1000 * enrollment_reported)) %>%
     felm(assistant_prof_count ~ 1 +
         `I(stateappropriations_real/(1000 * enrollment_reported))(fit)` |
-        unitid |
+        unitid + year |
         0 |
         unitid + year,
         data = .)
 # Shift-share IV Regression, explained by state appropriation shock
 shiftshare_assistant_count.reg <- institutional.data %>%
     felm(assistant_prof_count ~ 1 |
-        unitid |
+        unitid + year |
         (I(stateappropriations_real / (1000 * enrollment_reported)) ~
             I(-appropriationshock_perEnroll_real)) |
         unitid + year,
@@ -282,14 +284,14 @@ naive_full_count.reg <- institutional.data %>%
         stateappropriations_real / (1000 * enrollment_reported)) %>%
     felm(full_prof_count ~ 1 +
         `I(stateappropriations_real/(1000 * enrollment_reported))(fit)` |
-        unitid |
+        unitid + year |
         0 |
         unitid + year,
         data = .)
 # Shift-share IV Regression, explained by state appropriation shock
 shiftshare_full_count.reg <- institutional.data %>%
     felm(full_prof_count ~ 1 |
[…]

### 2024-05-08 programs/analysis/ibhed-shock-rolling.R

@@ -305,7 +305,7 @@ stargazer(
     model.names = FALSE,
     omit = "factor|count|year",
     intercept.bottom = TRUE,
-    covariate.labels = c("Appropriations Shock", "Tuition Revenue", "Constant"),
+    covariate.labels = c("State Funding", "Tuition Revenue", "Constant"),
     omit.stat = c("LL", "ser", "aic", "wald", "adj.rsq", "f"),
     add.lines = list(
         c("Fixed effects?", "Yes", "No", "Yes", "No"),

### 2024-05-08 programs/analysis/ibhed-shock.R

@@ -403,7 +403,7 @@ stargazer(
     model.names = FALSE,
     omit = "factor|count|year",
     intercept.bottom = TRUE,
-    covariate.labels = c("Appropriations Shock", "Tuition Revenue", "Constant"),
+    covariate.labels = c("State Funding", "Tuition Revenue", "Constant"),
     omit.stat = c("LL", "ser", "aic", "wald", "adj.rsq", "f"),
     add.lines = list(
         c("Fixed effects?", "Yes", "No", "Yes", "No"),
@@ -523,7 +523,8 @@ stargazer(
     naive_all_count.reg, shiftshare_all_count.reg,
     add.lines = outcome.means,
     dep.var.caption = "Dependent Variable: Employment Count by Professor Group",
-    dep.var.labels = c("Lecturer", "Assistant", "Full", "All"),
+    dep.var.labels = c(
+        "Lecturers", "Asst. Professors", "Full Professors", "All Faculty"),
     column.labels = rep(c("OLS", "2SLS"), 4),
     digits = digits.no,
     digits.extra = digits.no,
@@ -654,7 +655,8 @@ stargazer(
     naive_all_hires.reg, shiftshare_all_hires.reg,
     add.lines = outcome.means,
     dep.var.caption = "Dependent Variable: Employment Count by Professor Group",
-    dep.var.labels = c("Lecturer", "Assistant", "Full", "All"),
+    dep.var.labels = c(
+        "Lecturers", "Asst. Professors", "Full Professors", "All Faculty"),
     column.labels = rep(c("OLS", "2SLS"), 4),
     digits = digits.no,
     digits.extra = digits.no,

### 2024-05-08 programs/analysis/ipeds-robustness-checks.R

@@ -1,0 +1,1402 @@
+#!/usr/bin/R
+## Senan Hogan-Hennessy, 24 April 2024
+## IV for Prof's salaries, using IPEDS data.
+print(Sys.time())
+library(jtools) # Plotting estimates.
+library(tidyverse) # Functions for data manipulation and visualization
+library(lfe) # Functions for fast linear models with IV + FEs
+library(plm) # Functions for panel data
+library(car) # Function for F stat regarding IV models
+# My custom flavour of Stargazer TeX tables:
+# devtools::install_github("shoganhennessy/stargazer")
+library(stargazer) # TeX tables
+library(xtable)
+set.seed(47)
+# This file follows an adjusted Deming Walters (2017, p.10) approach to
+# appropriations shock instrument.
+
+# Define number of digits in tables and graphs
+digits.no <- 3
+
+# Size for figures
+fig.width <- 10
+fig.height <- fig.width * 0.9
+
+# Define default colours.
+colour.list <- c(
+    "#D62728", # Red  -> First-stage.
+    "#1f77b4", # Blue -> Lecturers
+    "#c9721c", # Orange -> Asst Professors
+    "#237e23", # Green -> Full Professors
+    "#cc2c9f") # Strong pink -> All faculty.
+
+# Define a function to standardise the dplyr lag function
+# --- negative numbers mean a lag of t time periods, positive a lead of t.
+lag_lead <- function(column, t) {
+    if (t < 0) {
+        lag_column <- dplyr::lag(column, -t)
+    } else if (t >= 0) {
+        lag_column <- dplyr::lead(column, t)
+    }
+    return(lag_column)
+}
+
+
+# Load data sources ------------------------------------------------------------
+
+# Load IPEDS data, for only the public unis
+ipeds_public.data <- read_csv("../../data/urban-ipeds/urban-clean-publicunis.csv")
+
+# Load IPEDS data, for all the unis
+ipeds_all.data <- read_csv("../../data/urban-ipeds/urban-clean-allunis.csv")
+
+
+################################################################################
+## Robustness check: effect on faculty with extra controls for selectivity.
+
+# Select IPEDS variables, and define the panel at the university level
+reg.data <- ipeds_public.data %>%
+    # Create independent columns and rows.
+    select(
+        # Identifiers and controls
+        unitid,
+        inst_name,
+        year,
[…]

### 2024-05-08 text/sections/10-appendix.tex
Newest revision + all code updated.

@@
     \caption{Correlation Between State Funding Shift-Share and Public University State Funding in Surrounding Years.}
-    \includegraphics[width=0.75\textwidth]{figures/lag-firststage.png}
+    \includegraphics[width=\textwidth]{figures/lag-firststage.png}
     \label{fig:lag-firststage}
@@
 These results were produced by integrating the total count of professor hires for 2010--2021 for the top-ranked 180 US universities with a sum of the funding variables, and then estimating the models specified in \autoref{sec:iv-model-uni}.
+There were no observable differences in the hiring rate of male versus female faculty.
+
 \begin{table}[H]
@@
     \label{tab:facultyhires-illinois-reg}
+    \vspace{-0.5cm}
     \justify
@@
 
-There were no observable differences in the hiring rate of male vs female faculty.
-
+\vspace{-0.5cm}
 \begin{figure}[H]
@@
     \singlespacing
-    \caption{Correlation Between State Funding and Total COunt Professors Hired, 2011--2021.}
-    \includegraphics[width=\textwidth]{figures/hiring-correlation.png}
+    \caption{Correlation Between State Funding and Total Count Professors Hired, 2011--2021.}
+    \includegraphics[width=0.75\textwidth]{figures/hiring-correlation.png}
     \label{fig:hiring-correlation}

### 2024-05-08 text/sections/01-introduction.tex
Added new robustness check tables.
Newest revision + all code updated.

@@
 %% Introduction section
-\section{Introduction}
-\label{sec:intro}
+%\section{Introduction}
+%\label{sec:intro}
 States fund their public universities with roughly the same amount today as they did in the 1990s, despite enrolment rising by over 50\%.
@@
 \autoref{sec:empirics} draws the empirical framework for isolating the causal effects of state funding on faculty, with \autoref{sec:results} presenting the empirical results.
-\autoref{sec:discussion} discusses the context and implications for the findings.
+%\autoref{sec:discussion} discusses the context and implications for the findings.
 \autoref{sec:conclusion} concludes.

### 2024-05-10 text/sections/09-figures-tables.tex
Fixed a spelling mistake.

@@
     \centering
-    \caption{First Stage Estimates, Predicting State Funding by Shift-Share shift-share in IPEDS Data.}
+    \caption{First Stage Estimates, Predicting State Funding by Funding Shift-Share, in IPEDS Data.}
     \textbf{Panel A: units in \$ per student}
@@
     \centering
-    \caption{Effects of State Funding on Faculty Counts, IPEDS 1990--2021, OLS and 2SLS Estimates.}
+    \caption{Effects of State Funding on Faculty Counts, in IPEDS Data 1990--2021, OLS and 2SLS Estimates.}
 
@@
     \centering
-    \caption{Effects of Changes in State Funding on Faculty Salaries and Exit Rate, in Illinois 2010--2021, 2SLS Estimates.}
+    \caption{Effects of State Funding on Faculty Salaries and Exit Rate, in IBHED Panel Data 2010--2021, 2SLS Estimates.}
     \label{tab:faculty-shock-illinois-rolling}

### 2024-05-10 text/sections/05-results.tex
Fixed a spelling mistake.

@@
 \label{sec:results-robustness}
-It is possible that the identification startegy is not robust to other possible confounders.
-For example, it is possible that selective institutions received less funding for other reasons, or states with lower performing insitutions or declining populations strategically limited funding for less selective universities.
-I address these main concerns with a number of different analyses.
+It is possible that the identification strategy is not robust to some possible confounders.
+For example, it is possible that selective institutions received less funding for other reasons, or states with lower performing institutions or declining populations strategically limited funding for less selective universities.
+I address these main concerns with three different analyses.
 
 First, I re-estimate the primary results from \autoref{sec:results-ipeds} with additional controls for university selectivity.
-To be specific, I estimate \autoref{eqn:secondstage} with controls added for the base share of state funding, average acceptance rate and 6-year completion rate  in 2004--2006, tuition revenue, total enrolment, size of state enrolment, size of university enrolment relative to the rest of the state.
+Specifically, I estimate \autoref{eqn:secondstage} with controls added for the base share of state funding, average acceptance rate and 6-year completion rate  in 2004--2006, tuition revenue, total enrolment, size of state enrolment, size of university enrolment relative to the rest of the state.
 Measures of selectivity are only available in certain years, so I use these values across the entire panel.
-This means that insitution fixed effects would over-identify the system, so I estimate with state $+$ year fixed effects (not intituion $+$ year fixed effects).
+This means that institution fixed effects would over-identify the system, so I estimate with state $+$ year fixed effects (not institution $+$ year fixed effects).
 In this way, this model also works as an alternative specification test.
-Additionally, this analysis uses the enrolment measure of full-time equivalent (FTE), and not reported enolment as in the original analysis.
-The first-stage is strong of the same order as the orginal analysis (\autoref{tab:firststage-robustness-checks}).
+Additionally, this analysis uses the enrolment measure of full-time equivalent (FTE), and not reported enrolment as in the original analysis.
+The first-stage is strong of the same order as the original analysis (\autoref{tab:firststage-robustness-checks}).
 The relationship between state funding, in both raw count terms and log units, are roughly the same as for the main analysis (\autoref{tab:facultycount-robustness-checks}, \ref{tab:facultycount-rawcount-robustness-checks}), despite the multiple specification changes, additional controls, and different measures for enrolment.
@@
 \autoref{tab:facultycount-heterogeneity} collects these results together, separated by the \cite{barrons2009} selectivity rankings.
-Barrons rank 1-2 are the most selective (including e.g., UC Berkeley and  Berkeley and SUNY Binghamton), 3-4 selective (including e.g., Clemson University and University of Dallas), and rank 5 for the rest of universities.
+Barrons rank 1-2 are the most selective (including e.g., UC Berkeley and SUNY Binghamton), 3-4 selective (including e.g., Clemson University and University of Dallas), and rank 5 for the rest of universities.
 The first-stage is string in every group of universities, even the most selective group which is only 12 universities, the selective group which is 28, and the rest are 762 universities.
@@
 
-The final robustness check I employ is a falsification test, testing whether the effects for public universities would be picked up among the sample of private universities (as in \citealt{NBERw27885}).
-Private universities do not rely on state funding in the same that public universities do, though they in some states do receive some state funding for projects and other smaller scale operating costs.
-Given this low reliance, it is unclear how state-wide funding shifts would affect faculty outcomes, and thus we would expect the two to be unrelated.
-If they were related, this would be evidence that the IV model picks up otherwise concurrent trends in higher education, unrelated to stagnating state funding for public universities.
-To impute a base share of reliance on state funding, I assign private universities the average reliance on state funding in the base period of universities in the same state with the same Barron's selectivity ranking, and then compute the shift-share instrument from this value.
-\autoref{tab:facultycount-shock-private-robustness} shows that the first-stage is not strong among private universities, and the IV results are significant for any faculty count.
-Notably, the OLS results for correlation between state funding and professor are highly statistically significant, showing that more funding is related to more lecturers and faculty in general, though the IV analysis shows that these relationships are spurious among private universities.
+The final robustness check I employ is a falsification test, testing whether the effects for public universities would be picked up among a sample of private universities.
+Private universities do not rely on state funding in the same that public universities do,\footnote{
+    Though private universities do receive some state funding, as some states fund for project at private universities and other smaller scale operating costs.
+}
+so we would expect state funding to not affect their faculty counts.
+If the state funding and faculty counts were related among private universities, this would be evidence that the IV model picks up otherwise concurrent trends in higher education, unrelated to stagnating state funding for public universities.
+To impute a base share of reliance on state funding, I assign private universities the average reliance on state funding in the base period of universities in the same state with the same Barrons selectivity ranking, and then compute the shift-share instrument from this value.
+\autoref{tab:facultycount-shock-private-robustness} shows that the first-stage is not strong among private universities; despite this, the IV results are still not significant for any faculty count, showing that state funding is not related to private universities' faculty counts.
+Notably, the OLS results for correlation between state funding and professor are highly statistically significant, showing that more funding is positively correlated with count of lecturers and total faculty.
+The IV estimates show that these relationships are spurious, and that the two are not causally related among private universities, validating the approach for public universities.

### 2024-05-10 programs/run.sh

@@ -42,3 +42,6 @@ cd ../text
 latexmk -pdf paper.tex
 latexmk -c
 cp paper.pdf ../state-funding-faculty-2024.pdf
+# Put the most recent presentation into the main folder.
+cd ../presentation
+cp presentation.pdf ../state-funding-faculty-presentation.pdf

### 2024-05-16 text/sections/05-results.tex
Added a brief first-stage spending analysis.

@@
 This allays concerns that state-wide shifts in state funding are targetted more towards less selective universities (a possible confounder), and supports use of the shift-share IV empirical design.
+
+Do state funding cuts meaningfully affect universities?
+While the shift-share performs well in predicting state funding (i.e., a strong instrument), it could be the case that universities are not affected.
+For expample, universities can respond to funding cuts by charging higher tuition fees; it could be the case that this entirely off-sets funding cuts, leading to no effects on the university.
+I estimate the IV model on IPEDS data for areas of university spending  (\autoref{tab:expenditures-shock-reg}).
+This exercise gauges the effects of state funding on areas of university spending to guage whether the first stage is economically meaningful, and to see which areas of higher education operations are affected by state funding cuts.
+Following a state funding cut, universities spend less on instruction, student services, operations, and grant aid.
+Additionally, they charge higher tuition fees, though this rise is at most a quarter of the state funding cut, so does not anywhere near off-set state funding cuts.
+These results are in line with previous work showing state funding cuts (or limits) reduce student-focused spending \citep{NBERw23736}, financial aid \citep{miller2022making}, and lead to increased tuition fees \citep{bound2019public}.
 
@@
 \autoref{tab:facultycount-shock-private-robustness} shows that the first-stage is not strong among private universities; despite this, the IV results are still not significant for any faculty count, showing that state funding is not related to private universities' faculty counts.
-Notably, the OLS results for correlation between state funding and professor are highly statistically significant, showing that more funding is positively correlated with count of lecturers and total faculty.
-The IV estimates show that these relationships are spurious, and that the two are not causally related among private universities, validating the approach for public universities.
+The OLS columns show that per student state funding is highly correlated with lecturer and professor counts at private universities.
+The IV estimates show that these correlations are not causal relationships, validating the approach for inference on public universities' faculty counts.

### 2024-05-16 text/sections/10-appendix.tex
Added a brief first-stage spending analysis.

@@
 
+\begin{table}[H]
+    \singlespacing
+    \centering
+    \caption{Effect of State Funding on Areas of University Spending, IPEDS Data, IV Estimates.}
+    \textbf{Panel A: units in \$ per student}
+
+    \makebox[\textwidth][c]{\input{tables/expenditures-shock-rawcount.tex}}
+
+    \textbf{Panel B: units in log \$ per student}
+
+    \makebox[\textwidth][c]{\input{tables/expenditures-shock-reg-fte.tex}}
+
+    \label{tab:expenditures-shock-reg}
+    \justify
+    \footnotesize
+    \textbf{Note}:
+    These tables show the IV estimates of regression specification \eqref{eqn:secondstage}, showing the effect of state funding (instrumented with the shift-share) on areas of university spending.
+    Outcomes are measured in \$ student spending, so that the estimates representat the relationship between a \$1 increase in state funding on area of spending (or 1\% increase).    
+    The ``outcome mean'' rows are spending in thousands \$ per student.
+\end{table}
+
 \newpage
@@
 
-    \textbf{Panel A: units in log \$ per student}
+    \textbf{Panel B: units in log \$ per student}
 

### 2024-05-16 programs/analysis/ipeds-shock-rawcounts.R

@@ -82,11 +82,17 @@ reg.data <- ipeds.data %>%
         all_profmeansalary_real,
         # Total paid on professor salaries
         all_profoutlays_real,
-        # Research and instruction expenditures.
+        # Research and instruction and other expenditures.
         instructionspending_total_real,
         instructionspending_salaries_real,
         researchspending_total_real,
         researchspending_salaries_real,
+        exp_pub_serv_real,
+        exp_student_serv_real,
+        exp_acad_supp_real,
+        exp_inst_supp_real,
+        exp_opm_areal,
+        exp_net_grant_aid_real,
         nonauxspending_real)
 
 
@@ -388,3 +394,294 @@ substitution_rawcount.plot <-
 ggsave("../../text/figures/substitution-rawcount-plot.png",
     plot = substitution_rawcount.plot,
     units = "cm", dpi = 300, width = 1.25 * fig.width, height = 0.8 * fig.height)
+
+
+# Research + Instruction Spending Regressions ----------------------------------
+
+# Restrict to spending data with no missings
+spending.data <- reg.data %>%
+    filter(nonauxspending_real > 0)
+
+## Total instruction spending.
+# Naive OLS Regression
+naive_instruction_spending.reg <- spending.data %>%
+    mutate(`I(stateappropriations_real/(enrollment_reported))(fit)` =
+        I(stateappropriations_real/(enrollment_reported))) %>%
+    felm(I(instructionspending_total_real / enrollment_reported) ~ 1 +
+        `I(stateappropriations_real/(enrollment_reported))(fit)` |
+        unitid + year |
+        0 |
+        unitid + year,
+        data = .)
+# Shift-share IV Regression, explained by state appropriation shock
+shiftshare_instruction_spending.reg <- spending.data %>%
+    felm(I(instructionspending_total_real / enrollment_reported) ~ 1 |
+        unitid + year |
+        (I(stateappropriations_real / (enrollment_reported)) ~
+            I(-appropriationshock_perEnroll_real)) |
+        unitid + year,
+        data = .)
+
+## Total research spending.
+# Naive OLS Regression
+naive_research_spending.reg <- spending.data %>%
+    filter(researchspending_total_real > 0) %>%
[…]

### 2024-05-16 programs/analysis/ipeds-shock.R

@@ -100,8 +100,16 @@ reg.data <- ipeds.data %>%
         full_parttime_count,
         lecturer_fulltime_count,
         assistant_fulltime_count,
-        full_fulltime_count)
-
+        full_fulltime_count,
+        # Exploratory, areas of spending.
+        exp_instruc_real,
+        exp_research_real,
+        exp_pub_serv_real,
+        exp_student_serv_real,
+        exp_acad_supp_real,
+        exp_inst_supp_real,
+        exp_opm_areal,
+        exp_net_grant_aid_real)
 
 # Restrict data to unis + years with measured state appropriations & shocks
 reg.data <- reg.data %>%
@@ -1282,95 +1290,288 @@ print(c("Calculated point est for substitution between ast + full profs",
 
 # Research + Instruction Spending Regressions ----------------------------------
 
-# restrict to observations with info on research and instruction spending.
-research_instruct.data <- reg.data %>%
-    filter(
-        !is.na(instructionspending_total_real),
-        instructionspending_total_real > 0,
-        !is.na(instructionspending_salaries_real),
-        instructionspending_salaries_real > 0,
-        !is.na(researchspending_total_real),
-        researchspending_total_real > 0,
-        !is.na(researchspending_salaries_real),
-        researchspending_salaries_real > 0)
+# Restrict to spending data with no missings
+spending.data <- reg.data %>%
+    filter(nonauxspending_real > 0)
 
 ## Total instruction spending.
 # Naive OLS Regression
-naive_instruction_spending.reg <- research_instruct.data %>%
+naive_instruction_spending.reg <- spending.data %>%
     mutate(`log(stateappropriations_real/enrollment_reported)(fit)` =
         log(stateappropriations_real / enrollment_reported)) %>%
     felm(log(instructionspending_total_real / enrollment_reported) ~ 1 +
         `log(stateappropriations_real/enrollment_reported)(fit)` |
         unitid + year |
         0 |
-        state + year,
+        unitid + year,
         data = .)
 # Shift-share IV Regression, explained by state appropriation shock
-shiftshare_instruction_spending.reg <- research_instruct.data %>%
+shiftshare_instruction_spending.reg <- spending.data %>%
     felm(log(instructionspending_total_real / enrollment_reported) ~ 1 |
[…]

### 2024-05-16 programs/data-build/urban-ipeds-build.R

@@ -133,7 +133,8 @@ finance.data <- finance.data %>%
             is.na(exp_total_opm) | exp_total_opm < 0, NA) / cpi2021,
         exp_net_grant_aid_real = replace(exp_net_grant_aid_total,
             is.na(exp_net_grant_aid_total) | exp_net_grant_aid_total < 0, NA) / cpi2021,
-        nonauxspending_real = (exp_instruc_real +
+        nonauxspending_real = (
+            exp_instruc_real +
             exp_research_real +
             exp_pub_serv_real +
             exp_student_serv_real +

### 2024-05-23 text/paper.tex
Added a brief first-stage spending analysis.
Adjusted formatting of some figures, and changed some bad punctuation.

@@
 % Tables + figures
+\usepackage{multirow}
 \usepackage{graphicx}                                                      % control over the import of graphics
@@
         and Evan Riehl.
-        I thank seminar participants at Cornell University (2023, 2024), the Eastern Economic Association Annual Meetings (2024) for helpful discussion.
-        This project's repository, with all materials and anonymised data needed for replication, is available at 
+        I thank seminar participants at Cornell University (2023, 2024), and the Eastern Economic Association Annual Meetings (2024) for helpful discussion.
+        This project's repository, including all materials and anonymised data needed for replication, is publicly available at 
         \url{https://github.com/shoganhennessy/state-funding-faculty}.

### 2024-05-23 text/sections/01-introduction.tex
Adjusted formatting of some figures, and changed some bad punctuation.

@@
 States fund their public universities with roughly the same amount today as they did in the 1990s, despite enrolment rising by over 50\%.
-At the same time, public universities have systematically substituted away from professors towards lecturers, who teach often work on contingent and lower-salaried contracts.
+At the same time, public universities have systematically substituted away from professors towards lecturers, who often work on contingent and lower-salaried contracts.
 Previous research shows state funding cuts led to worsening student outcomes \citep{NBERw23736,NBERw27885}, but there is little evidence on whether these funding cuts impacted faculty.
@@
 The closest related research on higher education funding examined how university spending affected graduation rates and levels of student debt \citep{NBERw23736,NBERw27885}, and university finances \citep{miller2022making,bound2019public,brown2014endowment}.
-For faculty outcomes, descriptive work measures how US and Canadian faculty salaries are correlated with other factors \citep{bratsberg2003negative,martinello2009faculty,pfeffer1992salary}, and applied theory work models university decision-making and faculty hiring (see e.g., \citealt{abe2015implications,johnson2009jep,NBERc13879}).
+For faculty outcomes, descriptive work measures how US and Canadian faculty salaries are correlated with other factors \citep{bratsberg2003negative,martinello2009faculty,pfeffer1992salary,hilmer2020labor}, and applied theory work models university decision-making and faculty hiring (see e.g., \citealt{abe2015implications,johnson2009jep,NBERc13879}).
 Other papers measure university decision--making after endowment shocks \citep{brown2014endowment}, and trends in the academic job market following the 2008 recession \citep{turner2014impact}.

### 2024-05-23 text/sections/02-data.tex
updated
Added new robustness check tables.
Adjusted formatting of some figures, and changed some bad punctuation.

@@
     Figures for enrolment and faculty counts come from the raw NCES version of IPEDS for all years, addressing inconsistencies in the Urban Institute's data formulation for these variables.
+    I also integrate Barron's selectivity index, from the 2009 rankings, to analyse effects for different levels of selectivity \citep{barrons2009}.
 }
@@
 }
-This gives a resulting panel data-set, where each row represents a university-year, and includes columns for university funding and tuition revenue, plus total count\footnote{
-    Real salary is computed by scaling nominal salary to 2021 dollars by the CPI-U.
-} for each faculty position (lecturer, assistant professor, tenured professor, and total faculty), and other university characteristics.
+This gives a resulting panel data-set, where each row represents a university-year, and includes columns for university funding and tuition revenue, plus total count for each faculty position (lecturer, assistant professor, tenured professor, and total faculty), and other university characteristics.
 \autoref{tab:ipeds-summary} presents summary statistics for these variables in IPEDS data.
@@
 To investigate outcomes for individual faculty members, I integrate individual-level data for all faculty members at all Illinois public universities 2010--2021.
-These data allow me to investigate how individual faculty members are affected by funding cuts to their university, investigating outcomes are more granular details than total faculty counts.
-IBHED freely provides this information, as the state of Illinois is required to publicly report base salary and benefits for all administrators, faculty members, and instructors employed by each public college or university.\footnote{
+These data allow me to investigate how individual faculty members are affected by funding cuts to their university, investigating individual-level professor outcomes unavailable in IPEDS data.
+IBHED freely provides this information, as the state of Illinois is required to publicly report base salary and benefits\footnote{
+    Real salary is computed by scaling nominal salary and benefits to 2021 dollars by the CPI-U.
+} for all administrators, faculty members, and instructors employed by each public college or university.\footnote{
     Public Act 96-0266, effective 1 January 2010, is the relevant Illinois law that requires publicly publishing salary data for all public university faculty salary and benefits \citep{illinois-public-act}.
@@
     Some faculty are listed multiple times in these panel data.
-    For example, Professor Alberto Agustin Lopez-Scala of the University of Illinois Chicago has two appointments, one as an adjunct faculty and one as a department director.
+    For example, Professor Alberto Agustin Lopez-Scala of the University of Illinois Chicago has two appointments in each year 2013--2021, one as an adjunct faculty and one as a department director.
     For this observation I take the highest paid position (department director) as the primary appointment and drop the secondary appointment (adjunct faculty).
@@
 
+% Paragraph describing how IL is similar to rest of country.
+
 There was not only a stagnation in state funding in this time period, but also large annual rises and falls, particularly around 2016.

### 2024-05-23 text/sections/05-results.tex
Adjusted formatting of some figures, and changed some bad punctuation.

@@
 \label{sec:results-robustness}
-It is possible that the identification strategy is not robust to some possible confounders.
-For example, it is possible that selective institutions received less funding for other reasons, or states with lower performing institutions or declining populations strategically limited funding for less selective universities.
-I address these main concerns with three different analyses.
+It is possible that the identification strategy is not robust to some confounding issues or trends.
+For example, selective institutions may have received funding cuts for other reasons, or states with lower performing institutions or declining populations strategically cut funding for only less selective universities.
+I address concerns for this with three different analyses.
 
@@
 Barrons rank 1-2 are the most selective (including e.g., UC Berkeley and SUNY Binghamton), 3-4 selective (including e.g., Clemson University and University of Dallas), and rank 5 for the rest of universities.
-The first-stage is string in every group of universities, even the most selective group which is only 12 universities, the selective group which is 28, and the rest are 762 universities.
-Faculty substitution rates are not very string among the most selective universities, but are very similar between selective universities and unranked universities, showing that even selective public universities exhibit the substitution towards lecturers.
+The first-stage is strong in every group of universities, even among the small sample size of the 12 universities in the most selective group, and the 28 universities in the next selective group.
+Faculty substitution rates are not very strong among the most selective universities, but are very similar between selective universities and unranked universities, showing that even selective public universities exhibit the substitution towards lecturers.
 
@@
 \autoref{tab:facultycount-shock-private-robustness} shows that the first-stage is not strong among private universities; despite this, the IV results are still not significant for any faculty count, showing that state funding is not related to private universities' faculty counts.
-The OLS columns show that per student state funding is highly correlated with lecturer and professor counts at private universities.
-The IV estimates show that these correlations are not causal relationships, validating the approach for inference on public universities' faculty counts.
+The OLS columns show that imputed state funding is highly correlated with lecturer and professor counts at private universities.
+The IV estimates are not distinguishable from zero, showing that OLS correlations are not causal relationships, and validating the IV approach for inference on universities' faculty counts.

### 2024-05-23 programs/analysis/ibhed-shock-longterm.R

@@ -188,6 +188,7 @@ reg.data <- reg.data %>%
 
 # Define length of time horizon to plot over
 time.horizon <- 5
+limits.vector <- c(-0.2, 0.6)
 
 # Define data sample for the LP estimation
 lp.data <- reg.data %>%
@@ -217,7 +218,7 @@ lp.data <- reg.data %>%
     pdata.frame(index = c("name_unitid", "year"), drop.index = FALSE)
 
 ## Define a function to nicely plot the LP results.
-lpreg.plot <- function(model.lpreg, given.colour) {
+lpreg.plot <- function(model.lpreg, limits.vector, given.colour) {
     ## Input a LP-reg (estimates an elesticity \in [-1,1])
     ## And the time.hirozon for how far to plot LP model
     # Get the estimates
@@ -239,6 +240,7 @@ lpreg.plot <- function(model.lpreg, given.colour) {
         scale_x_continuous(name = "Years, Relative to Funding Cut",
             breaks = model.lpdata$t, expand = c(0.01, 0.01)) +
         scale_y_continuous(name = "",
+            limits = limits.vector,
             breaks = seq(-5, 5, by = 0.1)) +
         theme_bw() +
         ggtitle("Estimate") +
@@ -268,7 +270,7 @@ firststage.lpreg <-
         hor = time.horizon)
 # Save this plot of th first stage.
 ggsave("../../text/figures/firststage-illinois-lp-rolling.png",
-    plot = lpreg.plot(firststage.lpreg, colour.list[1]),
+    plot = lpreg.plot(firststage.lpreg, c(-1.15, 0.2), colour.list[1]),
     units = "cm", width = fig.width, height = fig.height)
 
 
@@ -297,7 +299,7 @@ lecturer_salaries.lpreg <-
         hor = time.horizon)
 # Save this plot
 ggsave("../../text/figures/salaries-lecturer-illinois-lp-rolling.png",
-    plot = lpreg.plot(lecturer_salaries.lpreg, colour.list[2]),
+    plot = lpreg.plot(lecturer_salaries.lpreg, limits.vector, colour.list[2]),
     units = "cm", width = fig.width, height = fig.height)
 
 # LP estimation for salaries of APs.
@@ -323,7 +325,7 @@ assistant_salaries.lpreg <-
         hor = time.horizon)
 # Save this plot
 ggsave("../../text/figures/salaries-assistant-illinois-lp-rolling.png",
[…]

### 2024-05-23 programs/analysis/ibhed-shock-rolling.R

@@ -386,7 +386,7 @@ stargazer(
     shiftshare_full_salaries.reg,
     shiftshare_administrator_salaries.reg,
     shiftshare_all_salaries.reg,
-    dep.var.caption  = "Dependent Variable: Salaries by Professor Group",
+    dep.var.caption  = "Dependent Variable: Annual Salaries",
     dep.var.labels.include = FALSE,
     column.labels = c("Lecturer", "Assistant", "Full", "Admin", "All"),
     digits = digits.no,
@@ -475,7 +475,7 @@ stargazer(
     shiftshare_full_salaries.reg,
     shiftshare_administrator_salaries.reg,
     shiftshare_all_salaries.reg,
-    dep.var.caption  = "Dependent Variable: Salaries by Professor Group",
+    dep.var.caption  = "Dependent Variable: Annual Salaries",
     dep.var.labels.include = FALSE,
     column.labels = c("Lecturer", "Assistant", "Full", "Admin", "All"),
     digits = digits.no,
@@ -566,7 +566,7 @@ stargazer(
     shiftshare_assistant_promotion.reg,
     shiftshare_full_promotion.reg,
     shiftshare_all_promotion.reg,
-    dep.var.caption  = "Dependent Variable: Promotion Rate by Professor Group",
+    dep.var.caption  = "Dependent Variable: Promotion Rate",
     dep.var.labels.include = FALSE,
     column.labels = c("Lecturer", "Assistant", "Associate", "All"),
     digits = digits.no,
@@ -651,7 +651,7 @@ stargazer(
     shiftshare_full_exit.reg,
     shiftshare_administrator_exit.reg,
     shiftshare_all_exit.reg,
-    dep.var.caption  = "Dependent Variable: Exit rate by Professor Group",
+    dep.var.caption  = "Dependent Variable: Exit rate",
     dep.var.labels.include = FALSE,
     column.labels = c("Lecturer", "Assistant", "Full", "Admin", "All"),
     digits = digits.no,

### 2024-05-23 programs/analysis/ibhed-shock.R

@@ -522,7 +522,7 @@ stargazer(
     naive_full_count.reg, shiftshare_full_count.reg,
     naive_all_count.reg, shiftshare_all_count.reg,
     add.lines = outcome.means,
-    dep.var.caption = "Dependent Variable: Employment Count by Professor Group",
+    dep.var.caption = "Dependent Variable: Employment Count",
     dep.var.labels = c(
         "Lecturers", "Asst. Professors", "Full Professors", "All Faculty"),
     column.labels = rep(c("OLS", "2SLS"), 4),
@@ -654,7 +654,7 @@ stargazer(
     naive_full_hires.reg, shiftshare_full_hires.reg,
     naive_all_hires.reg, shiftshare_all_hires.reg,
     add.lines = outcome.means,
-    dep.var.caption = "Dependent Variable: Employment Count by Professor Group",
+    dep.var.caption = "Dependent Variable: Employment Count",
     dep.var.labels = c(
         "Lecturers", "Asst. Professors", "Full Professors", "All Faculty"),
     column.labels = rep(c("OLS", "2SLS"), 4),
@@ -740,7 +740,7 @@ stargazer(
     shiftshare_full_salaries.reg,
     shiftshare_administrator_salaries.reg,
     shiftshare_all_salaries.reg,
-    dep.var.caption  = "Dependent Variable: Salaries by Professor Group",
+    dep.var.caption  = "Dependent Variable: Annual Salaries",
     dep.var.labels.include = FALSE,
     column.labels = c("Lecturer", "Assistant", "Full", "Admin", "All"),
     digits = digits.no,
@@ -898,7 +898,7 @@ stargazer(
     shiftshare_full_salaries.reg,
     shiftshare_administrator_salaries.reg,
     shiftshare_all_salaries.reg,
-    dep.var.caption  = "Dependent Variable: Salaries by Professor Group",
+    dep.var.caption  = "Dependent Variable: Annual Salaries",
     dep.var.labels.include = FALSE,
     column.labels = c("Lecturer", "Assistant", "Full", "Admin", "All"),
     digits = digits.no,
@@ -1037,7 +1037,7 @@ stargazer(
     shiftshare_assistant_promotion.reg,
     shiftshare_full_promotion.reg,
     shiftshare_all_promotion.reg,
-    dep.var.caption  = "Dependent Variable: Promotion Rate by Professor Group",
+    dep.var.caption  = "Dependent Variable: Promotion Rate",
[…]

### 2024-05-23 text/sections/05-results.tex
Small fixes.

@@
 }
-To study whether these effects hold true for nation-wide hiring of new professors, I use public on the number of professors at all US public universities for the entire decade 2011--2021, and perform the same analysis to see if state funding affected the number of professors hired.
+To study whether these effects hold true nation-wide, I use public on the number of professors at all US public universities for the entire decade 2011--2021, provided by \citep{wapman2022quantifying},
+and perform the same analysis to see if state funding affected the number of professors hired.
+See \autoref{sec:appendix-hiring} for further details on this data source.
 
 \autoref{tab:facultyhires-illinois-reg} shows these estimates for Illinois public universities.
-In general, faculty counts are relatively unrelated to state funding among Illinois universities; though the point estimates line up with analysis in IPEDS, the smaller number of observations mean that the precision of estimates are small and indistinguishable from zero, in bith raw counts and in log units.
+In general, faculty counts are relatively unrelated to state funding among Illinois universities; though the point estimates line up with analysis in IPEDS, the smaller number of observations mean that the precision of estimates are small and indistinguishable from zero, in both raw counts and in log units.
 On the other hand, universities with more state funding hire more professors across 2011--2021, as seen in \autoref{fig:hiring-correlation}.
@@
 
-This shows that state funding is related in to professor hiring, and that the persistent funding cuts to public universities led to lower professor hiring across this time period.
-The other turn-over channels were unaffected by state funding cuts, and that professor hiring is related to professor hiring in both correlational terms and in a shift-share analysis.
+State funding cuts lead to lower professor hiring, and the persistent funding cuts to public universities led to lower professor hiring across the last three decades.
+The channels for faculty turn-over were unaffected by state funding cuts in the analysis of Illinois faculty, and rate professor hiring is related to professor hiring in both correlational terms and in a shift-share analysis.
 This means that effects on faculty hiring were likely the main mechanism for how state funding cuts led to a substitution of professors for lecturers at public universities.

### 2024-05-28 text/sections/01-introduction.tex
Finished adjusting the writing.  It is ready to Zenodo the data files.

@@
 
-Per student state funding for higher education has stagnated the last thirty years.
-The average public university received around \$11,600 of state funding per student in 1990 and \$8,300 in 2021, while the number of lecturers per student increased by 109\% and professors per student decreased by $23$\% over 1990--2021.
+per-student state funding for higher education has stagnated the last thirty years.
+The average public university received around \$11,600 of state funding per-student in 1990 and \$8,300 in 2021, while the number of lecturers per-student increased by 109\% and professors per-student decreased by $23$\% over 1990--2021.
 I use a shift-share instrumental variables approach to estimate how faculty are affected by these funding cuts.
-Universities employ $4.4$\% more lecturers per student following a $10$\% funding cut, and correspondingly rely less on tenure-track faculty by employing $1.4$\% fewer tenure-track professors and $1.2$\% fewer tenured professors per student.
+Universities employ $4.4$\% more lecturers per-student following a $10$\% funding cut, and correspondingly rely less on tenure-track faculty by employing $1.4$\% fewer tenure-track professors and $1.2$\% fewer tenured professors per-student.
 In a secondary analysis of all public university faculty in the state of Illinois, I show that incumbent faculty were unaffected, in terms of wages, promotion and quit rates.
 This implies that the faculty substitution arose by disrupting hiring of new tenure-track/tenured professors, which is supported supplementary data on professor hiring at US public universities over 2010--2021.
-% Private universities were not exposed to similar financial constraints during this same time period, and do not exhibit the substitution away from tenure-track/tenured professors.
 This paper shows the institutional effects of stagnating state funding for higher education, and offers one possible mechanism for how funding cuts affected student outcomes.
@@
 My second analysis uses panel data for every faculty member at Illinois public universities for 2010--2021 \citet[IBHED]{ibhed}, to measure how incumbent faculty (i.e., those already employed at the university) are affected by changes in state funding.
-These data allow me to answer the following research question: if a US public university receives an extra \$1,000 per student (or a 10\% increase), do they change how many faculty members they employ?
+These data allow me to answer the following research question: if a US public university receives an extra \$1,000 per-student (or a 10\% increase), do they change how many faculty members they employ?
 If so, which positions do universities substitute away from, and do these funding cuts affect the faculty themselves?
@@
 But it is not certain that universities would hire more lecturers in the face of budget cuts.
-Public university administrators, motivated by increased competition \citep{hoxby2009changing,urquiola2020markets}, may notice that their rankings are negatively affected by falling professor--lecturer ratios, and so limit departments from hiring more lecturers to replace leaving/retiring professors.
+Public university administrators, motivated by increased competition \citep{hoxby2009changing,urquiola2020markets}, may notice that their rankings are negatively affected by falling professor-lecturer ratios, and so limit departments from hiring more lecturers to replace leaving/retiring professors.
 Administrators could even start courting private funding or increase tuition prices to fill funding gaps \citep{bound2019public}, possibly leaving faculty unaffected by state funding cuts.
@@
 In my university-level analysis (IPEDS), I find that state funding cuts cause public universities to substitute away from tenured and tenure-track professors towards lecturers.
-A funding cut of 10\% per student leads to a fall of 1.4\% in the number of assistant professors per student at a university, a fall of 1.2\% for tenured professors, and an increase in 4.4\% the number of lecturers per student.
+A funding cut of 10\% per-student leads to a fall of 1.4\% in the number of assistant professors per-student at a university, a fall of 1.2\% for tenured professors, and an increase in 4.4\% the number of lecturers per-student.
 Local projection estimates show that these effects linger for up to ten years after the initial funding cut, showing that effects are persistent and not transitory.
-Over the same time period, state funding per student fell by around 35\%, professors per student fell by 9\%, and lecturers per student increased by 99\%.
-These results show that falls in state funding explain around 53\% of the fall in professors per student,
+Over the same time period, state funding per-student fell by around 35\%, professors per-student fell by 9\%, and lecturers per-student increased by 99\%.
+These results show that falls in state funding explain around 53\% of the fall in professors per-student,
 % i.e. -(35 * 0.137) / -9
-and 15.5\% of the rise in lecturers per student.
+and 15.5\% of the rise in lecturers per-student.
 % ie., (35 * 0.437) / 99
@@
 \autoref{sec:empirics} draws the empirical framework for isolating the causal effects of state funding on faculty, with \autoref{sec:results} presenting the empirical results.
-%\autoref{sec:discussion} discusses the context and implications for the findings.
 \autoref{sec:conclusion} concludes.

### 2024-05-28 text/sections/02-data.tex
Finished adjusting the writing.  It is ready to Zenodo the data files.

@@
 }
-US state governments, by majority, are legally obligated to run a balanced-budget, so that yearly variation in tax revenues (e.g., caused by changing economic conditions) necessarily affect state expenditures.
-Public universities have lower lobbying power than other state institutions, so often bear the brunt of state government funding cuts \citep{delaney2011state}.\footnote{
+US state governments, by majority, are legally obliged to run a balanced-budget, so that yearly variation in tax revenues (e.g., caused by changing economic conditions) necessarily led to state government spending cuts.
+Public universities have lower lobbying power than other state institutions, so often bear the brunt of these funding cuts \citep{delaney2011state}.\footnote{
     \cite{delaney2011state} fully describe the financial environment of state expenditures, and what makes spending on higher education an attractive area for state governments to expand funding during years of higher tax revenues, and retract funding in leaner years.
@@
 
-State funding for public universities stagnated over 1990--2021.
+State funding for public universities stagnated over the last thirty years.
 \autoref{fig:funding} show the trends in revenues for the mean public university for the years 1990--2021.
@@
 The general trend is similar for the number of assistant professors at both public and private universities.
-Private and public universities have similar numbers of associate and full professors before the year 2000, yet this number fell by over 20\% afterwards only for public universities: in 2021 the average public university has 6 fewer full professors per hundred students than the average private university.
-Over the same time period, we see the rise in use of lecturers; lecturers were were employed at similar rates in private and public sectors in 1990, but have public universities increased their reliance on lecturers more than private universities ever since.
+Private and public universities have similar numbers of associate and full professors before the year 2000.
+After 2000, the number of professors at public univeristies fell drastically, with the average public university has 6 fewer full professors per hundred students than the average private university by 2021.
+Over the same time period, we see the rise in use of lecturers; lecturers were employed at similar rates in private and public sectors in 1990.
+Since 1990, both sectors began employing more lecturers on a per-student basis.
+In sum, this means public universities have begun relying on lecturers, because they decreased their professor counts, compared to private unviersities who started employing more lecturers and professors.
 
@@
 \label{sec:trends-illinois}
-Illinois funding for higher education has stagnated, and has experienced serious declines in the decade 2010--2021 --- similar in magnitude to the nation-wide decline over 1990--2021.
+Illinois funding for higher education has stagnated, experiencing serious declines in the decade 2010--2021 --- similar in magnitude to the nation-wide decline over 1990--2021.
 State funding Illinois public universities fell by over 50\% over 2010--2021, in both absolute and per-student terms (see \autoref{fig:illinois-funding}).
-%Similar to the figures for the rest of the country, there has been a corresponding substitution towards tuition revenue.
-
-% Paragraph describing how IL is similar to rest of country.
 
@@
 This means that Illinois public universities exhibit sizable changes in their state funding over 2010--2021, of similar order to those for the rest of the country over 1990--2021.
-Additionally, the 2016 episode stemmed entirely from political disagreements, and not from state decisions regarding higher education and its finances \citep{young2020squandered}, exhibiting how state-level changes in funding affect public universities thanks to unrelated issues.
+Additionally, the 2016 episode stemmed entirely from political disagreements, and not from state decisions regarding higher education and its finances, exhibiting how state-level changes in funding affect public universities thanks to unrelated issues \citep{young2020squandered}.

### 2024-05-28 text/sections/03-concepts.tex
Finished adjusting the writing.  It is ready to Zenodo the data files.

@@
 \label{sec:conceptual}
-
-Universities can respond multiple ways to state funding cuts.
+University administrations can respond in multiple ways to state funding cuts.
 \cite{NBERw23736} established that public universities experiencing state funding cuts did not fully off-set by raising tuition, and instead cut spending.
@@
 
-Wider changes in the American higher education sector could concurrently explain the trends in state funding and faculty substitution.
+Wider changes in American higher education could concurrently explain the national trends in state funding and faculty substitution.
 College selectivity changed drastically over the 21st century, where the top universities have become more selective while the average university less selective \citep{hoxby2009changing}.
@@
 
-Universities could respond to funding cuts by cutting faculty salaries, leading to more faculty leaving for jobs at other universities.
+Similarly, institution-specific factors could explain correlations between state funding cuts and individual-faculty outcomes, such as salaries or rates of tenure.
+FOr example, an administration could respond to funding cuts by cutting faculty salaries, leading to more faculty leaving for jobs at other universities.
 Multiple universities passed a university-wide pay-cut for their faculty in response to state budget cuts around the 2008 recession, and Cornell University implemented a professor salary cut in 2020 at the onset of the Covid-19 pandemic.\footnote{
-    The Cornell University administration expected large fiscal squeezes in mid-2020, so imposed a faculty salary cut, and then returned the amount cut later in 2020 when the financial squeeze did not materialise.}
+    The Cornell University administration expected large fiscal squeezes in mid-2020, so imposed a faculty salary cut, and then returned the amount cut later in 2020 when the financial squeeze did not materialise.
+}
 Funding cuts could also limit the salaries public universities can offer to new faculty hires, leading to fewer or less accomplished faculty accepting offers at public universities.\footnote{
@@
 %This means that state funding cuts could effect faculty in only transitory changes in faculty counts.
-As such, this paper investigates the dynamics of state funding cuts and their effects on faculty, by employing local projections methods to measure the persistence of effects for multiple years following a state funding cut.
+As such, this paper investigates the dynamics of state funding cuts, and their effects on faculty, by employing local projections methods to measure the persistence of effects for multiple years after a state funding cut.

### 2024-05-28 text/sections/04-empirical.tex
Finished adjusting the writing.  It is ready to Zenodo the data files.

@@
 Instead, the state government undertakes a complex process of budgeting funding across many priorities, while higher education and faculty outcomes may not be independent in this process.
-For example, it may be the case that a state government cuts funding for higher education as the demographics of the state change, and the average age grows older (such as in Maine) 
-At the same time, it may be harder to attract younger faculty to work at universities in a state with an older population, so that faculty outcomes are correlated with state funding thanks to changes in state demographics.
 This means that faculty outcomes would not be conditionally independent of faculty outcomes, and correlations between the two are not causal relationships.
@@
 \label{sec:approp-shocks}
-I use a shift-share instrument to address endogeneity concerns for state governments' funding decisions, for their public universities.
-\cite{NBERw23736,NBERw27885} develop the shift-share approach for public university funding by exploiting shifts in yearly state-wide higher eduction funding, interacted with universities' reliance on state funding in a base period.
+I use a shift-share instrument variables (IV) approach to address endogeneity concerns for state governments' higher education funding decisions.
+\cite{NBERw23736,NBERw27885} first developed the shift-share approach for public university funding by exploiting shifts in yearly state-wide higher eduction funding, interacted with universities' reliance on state funding in a base period.
+$Z_{i,t}$ is the shift-share instrument for institution $i$ in year $t$:
 \begin{align}
@@
     \sum_{\tau = 0}^{3} \frac 14
-    \left( \frac{\text{State Funding}_{i,1990 + \tau}}{\text{Total Revenues}_{i,1990 + \tau}} \right) \right]
+    \left( \frac{\text{State Funding}_{i,1990 + \tau}}{\text{Total Revenues}_{i,1990 + \tau}} \right) \right].
 \end{align}
@@
 The system exploits the fact that institutions who rely more on state funding will be affected by follow-on state-wide funding cuts.
-$Z_{i,t}$ is the shift-share instrument for institution $i$ in year $t$.
 It is constructed by interacting the average funding for universities in state $s(i)$ with reliance on state funding relative to total revenues, averaged across the base years 1990--1993.\footnote{
@@
     It may be argued that tuition revenues are confounder between the causal effect of changes in state funding on a university's total revenues, so that this analysis focuses on state support for higher education (not total revenue), as do \cite{NBERw27885}.
-    On the other hand, rises in tuition revenues (per student) may arise as result of tuition hikes thanks declining state support, which would mean controlling for tuition would constitute a bad control.
-    Nonetheless, estimates including tuition revenues (per student) as a control in the second stage of the IV estimates produce results of very similar magnitude and direction, and so are omitted.
+    On the other hand, rises in tuition revenues (per-student) may arise as result of tuition hikes thanks declining state support, which would mean controlling for tuition would constitute a bad control.
+    Nonetheless, estimates including tuition revenues (per-student) as a control in the second stage of the IV estimates produce results of very similar magnitude and direction, and so are omitted.
 }
@@
 Universities are by-in-large smaller (in terms of enrolment, total revenues, and professor count) at the top of the shift-share instrument distribution (see~\autoref{tab:summary-quantiles}).
-Though on a per student basis, there is little difference across the distribution (except in the outcomes under consideration).
+Though on a per-student basis, there is little difference across the distribution (except in the outcomes under consideration).
 The shift-share instrument is positively associated with the total enrolment and total amount of state funding for each university, in both \$ and $\log$/percentage change terms, while other the other sources of university finances are not associated with the funding shift-share.
 So that the other sources of finances are not clear confounders for the IV strategy, as they exhibit balance with respect to the funding shift-share.
-Nonetheless, yearly and institution fixed effects are included in regressions throughout to implicitly condition on mean differences between universities and years.
+Nonetheless, yearly and institution fixed effects are included in regressions throughout to implicitly control for differences between universities and years.
 
@@
 
-Additionally, I investigate the effect of changes in state funding among incumbent professors using the Illinois data (IBHED).
-Incumbent professors are faculty who are already employed at the university; state funding changes may affect faculty who are already at the university (beyond affecting whether they get hired), so I base the instrument in the year that the professor was hired, and include fixed effects for the hiring year.\footnote{
-    This formulation follows that presented by \cite{NBERw27885}, where individual student outcomes are analysed via variation in state funding after their freshman-year.
+Additionally, I investigate the effect of state funding cuts on professors using the Illinois data (IBHED).
+State funding cuts may affect faculty in years after they join the university, so I base the instrument in the year that the professor was hired, and include fixed effects for the hiring year.\footnote{
+    This formulation follows that presented by \cite{NBERw27885}, where individual student outcomes are analysed via variation in state funding after a student's freshman-year.
     This contrasts with \autoref{sec:iv-model-uni} and \cite{NBERw23736}, where the unit of analysis is the university-year, where base year 1990 is more appropriate.
@@
 It is not a priori clear which units are appropriate for this analysis;
-does it make sense to consider state funding in purely dollar amounts per student, or in percent change rate?
-The funding shift-share is a strong predictor for the average university's level of state funding in either unit; a funding shift-share of \$1,000 per student in the entire state leads to around \$1,176 per student at the university, while a funding shift-share of $-10$\% leads to around 9.77\% less state funding per year.
+does it make sense to consider state funding in purely dollar amounts per-student, or in percent change rate?
+The funding shift-share is a strong predictor for the average university's level of state funding in either unit; a funding shift-share of \$1,000 per-student in the entire state leads to around \$1,176 per-student at the university, while a funding shift-share of $-10$\% leads to around 9.77\% less state funding per year.
 Yet, the level of state funding (and the outcome variables) vary greatly between states for the unit of analysis.
-For example, the average Illinois public university receives \$10,709 in state funding per student in 1990 and  \$6,713 in 2021 (a fall of over 30\%), while California went from \$19,224 per student to \$12,915 in the same time span (a fall of 37\%).
+For example, the average Illinois public university receives \$10,709 in state funding per-student in 1990 and  \$6,713 in 2021 (a fall of over 30\%), while California went from \$19,224 per-student to \$12,915 in the same time span (a fall of 37\%).
 While most states are not exactly the same as California or Illinois, this is example is telling for the phenomenon of stagnating state funding:
-states vary in their absolute funding for higher education in 1990, but most have experienced a decline of around 30\% in per student funding, so that the trend is a common phenomenon in percentage terms across the US.
+states vary in their absolute funding for higher education in 1990, but most have experienced a decline of around 30\% in per-student funding, so that the trend is a common phenomenon in percentage terms across the US.
 As such, I include regression specifications where the explanatory and outcome variables are $\log$ transformed, and refer to these when stating results in percent change terms.
@@
 The effects of a change in the universities funding may not be immediate, and faculty may be affected multiple years after a funding cut.
-Yet, the funding shift-share instrument is significantly auto-correlated year-on-year; a large state-level funding cut in year $t$ is also likely to experience a large funding cut in year $t-1$.
+Yet, the funding shift-share instrument is significantly auto-correlated year-on-year; a large state funding cut in year $t$ is also likely to experience a large funding cut in years $t \pm 1$.
 \autoref{fig:lag-firststage} shows that the shift-share instrument is correlated with state funding five years into the past and future.
-This level of time-series correlation would confound IV estimates of the effect of state funding in year $t$ on faculty outcomes in later years $t+k$, providing inconsistent estimates for the causal relationships.
+This level of time-series correlation would confound IV estimates of the effect of state funding in year $t$ on faculty outcomes in later years $t+k$, providing inconsistent estimates for the causal relationships when using two-stage least squares.
 

### 2024-05-28 text/sections/05-results.tex
Finished adjusting the writing.  It is ready to Zenodo the data files.

@@
 The shift-share instrument has a strong first stage for universities' yearly state funding.
-\autoref{tab:firststage-reg} presents results of the first-stage regression among IPEDS data, separately with and without a control for tuition revenue per student, plus institution and year fixed effects.
+\autoref{tab:firststage-reg} show results of the first-stage regression among IPEDS data, separately with and without a control for tuition revenue per student.
 \autoref{tab:firststage-reg} Panel A shows estimates of a state funding shift-share of \$1 per student on state funding per student at the university, and Panel B the effects of a 1\% increase in the state funding shift-share on state funding per student.
-Column (1) shows that a state funding shift-share of \$10 per student is associated with \$11.76 less state funding per student at the university; -10\% state funding shift-share leads to -9.77\% state funding per student at the average university.
+Column (1) shows that a state funding shift-share of \$10 per student is associated with \$11.76 less state funding per student at the university; $-10$\% state funding shift-share leads to $-9.77$\% state funding per student at the average university.
 Estimates are presented for specifications with and without fixed effects.
@@
 Columns (3) and (4) include the tuition revenue as a control (explained in \autoref{foot:control}) to exhibit estimates with or without this control variable.
-Column (3) shows similar estimates to column (1) in both Panels A and B thanks to inclusion of fixed effects, so that 
-whether tuition revenue is a bad control does not affect empirical estimates.
-Column (1) represents the estimates for \autoref{eqn:firststage} with fixed effects (not including tuition revenue as a control variable), and is the preferred form that I proceed with.
+Column (3) shows similar estimates to column (1) in both Panels A and B thanks to inclusion of fixed effects, so whether tuition revenue is a bad control does not affect empirical estimates.
+Column (1) represents the estimates for \autoref{eqn:firststage} with fixed effects (not including tuition revenue as a control variable), and is the preferred specification that I proceed with.
 I use the same empirical strategy in the analysis of Illinois faculty, and \autoref{tab:firststage-illinois} presents first-stage estimates among the IBHED data; results are similar to those calculated in IPEDS data.
-These results, show the funding shift-share instrument strongly predicts university revenues in the first-stage estimation.
+These results show the funding shift-share instrument strongly predicts a university's state funding in the first-stage estimation.
 
@@
 \autoref{tab:summary-quantiles} shows the mean of various university variables, across the instrument distribution.
-The university--years in the lowest 20\% of the instrument distribution have more research spending than those in the next 20\%; this is to be expected, as university spending is heavily related to state--wide changes in funding.
+The university--years in the bottom 20\% of the instrument distribution have more research spending than those in the next 20\%; this is to be expected, as university spending is heavily related to state--wide shifts in funding.
 On the other hand, the measures of selectivity (acceptance rate, 6--year graduation rate) are not substantially different across the distribution of the instrument.
-This allays concerns that state-wide shifts in state funding are targetted more towards less selective universities (a possible confounder), and supports use of the shift-share IV empirical design.
+This allays concerns that state-wide shifts in state funding are targetted at less selective universities (a possible confounder), and supports use of the shift-share IV empirical design.
 
 Do state funding cuts meaningfully affect universities?
-While the shift-share performs well in predicting state funding (i.e., a strong instrument), it could be the case that universities are not affected.
-For expample, universities can respond to funding cuts by charging higher tuition fees; it could be the case that this entirely off-sets funding cuts, leading to no effects on the university.
-I estimate the IV model on IPEDS data for areas of university spending  (\autoref{tab:expenditures-shock-reg}).
-This exercise gauges the effects of state funding on areas of university spending to guage whether the first stage is economically meaningful, and to see which areas of higher education operations are affected by state funding cuts.
+While the shift-share performs well in predicting state funding (i.e., a strong instrument), it could be the case that the day-to-day operations of a university are not affected.
+University could have responded to state funding cuts by cutting areas of spending unrelated to research or instruction, or fully off-set the cuts by charging higher tuition fees, leading to no effects on the university or faculty.
+To allay this concern, I estimate the IV model on IPEDS data for areas of university spending  (\autoref{tab:expenditures-shock-reg}).
+This exercise gauges the effects of state funding cuts on areas of university spending to gauge whether the first stage is economically meaningful, and to see which areas of higher education operations are affected by state funding cuts.
 Following a state funding cut, universities spend less on instruction, student services, operations, and grant aid.
-Additionally, they charge higher tuition fees, though this rise is at most a quarter of the state funding cut, so does not anywhere near off-set state funding cuts.
-These results are in line with previous work showing state funding cuts (or limits) reduce student-focused spending \citep{NBERw23736}, financial aid \citep{miller2022making}, and lead to increased tuition fees \citep{bound2019public}.
+Additionally, they charge higher tuition fees, though this rise is at most a quarter of the state funding cut, so does not come close to fully off-setting state funding cuts.
+These results are in line with previous work showing state funding cuts (or tuition price limits) reduce student-focused spending \citep{NBERw23736}, financial aid \citep{miller2022making}, and lead to increased tuition fees \citep{bound2019public}.
 
@@
 \label{sec:results-ipeds}
-Funding cuts lead to measurable changes in the number of faculty at US public universities.
-\autoref{tab:facultycount-shock-reg} presents OLS and IV estimates for the effect of a change in state funding on the number of faculty at the university, for each faculty position (lecturer, assistant and full professors, total).
-The results are presented in two different units: Panel A shows how a \$1 increase in state funding changed the number of faculty per students, while Panel B shows how a 1\% increase in state funding led to a percent change in the number of faculty per students.
+State funding cuts lead to measurable changes in the number of faculty at US public universities.
+\autoref{tab:facultycount-shock-reg} presents OLS and IV estimates for the effect of state funding cuts on the number of faculty at public universities, separately for each faculty position (lecturer, assistant and full professors, total).
+The results are presented in two different units: Panel A shows how a \$1 increase in state funding changed the count of faculty per students, while Panel B shows how a 1\% increase in state funding led to a percent change in the number of faculty per students.
 
@@
 A state funding cut of 10\% leads to a 0.65\% decrease in the total number of faculty per student (\autoref{tab:facultycount-shock-reg}, Panel B).
-These are large substitution effects, and explain roughly 40\% of the trend in subsitutional towards lecturers over this same time period.
-The average public university has roughly 36\% less state funding per student in 2021 than they did in 1990, while employing over double the number of lecturers and 20\% fewer professors (relative to enrolment, \autoref{fig:mean-funding-fte}, \ref{fig:fte-perprof}).
+These are large substitution effects, and explain roughly 40\% of the trend in substitution towards lecturers over the same time period.
+The average public university has roughly 36\% less state funding per student in 2021 than they did in 1990, while employing over double the number of lecturers and 20\% fewer professors (relative to enrolment, seen in \autoref{fig:mean-funding-fte}, \ref{fig:fte-perprof}).
 
-Once the funding cut has passed, universities do not re-adjust to hiring more professors to make up for last year's substitution towards lecturers.
-\autoref{fig:count-lp} shows local projection estimates, where the estimate is the elasticity for professor count per student in year $t+k$ with respect to state funding in year $t$.
-These elasticity estimates are shown in \autoref{tab:facultycount-shock-reg} Panel B, where the outcome is faculty count per students in later years.
+Universities do not re-adjust by hiring more professors to make up for last year's funding cut, and so the substitution towards lecturers becomes persistent.
+\autoref{fig:count-lp} shows local projection estimates, showing the elasticity for professor count per student in year $t+k$ with respect to state funding in year $t$.
 State funding cuts affect the number of lecturers and professors at universities up to 9 years after the initial funding cut.
@@
 
-These effects are a substitution of professors for lecturers, as the total number of faculty per students are relatively unaffected in the follow-on years, while numbers of lecturers increase and professors decreases.
+These effects are a substitution of professors for lecturers at public universities, as the total number of faculty per students are relatively unrelated to funding cuts, but the lecturer count increases and professor count decreases.
 This agrees with a model for universities departments having fixed teaching requirements making human-resources decisions under constrained resources (e.g., \citealt{abe2015implications}), lending credibility to this type of model for understanding recent human resources decisions of public universities.
-Furthermore, if a university must teach more students, but cannot afford to hire a full-time professor who teaches 2--3 classes a year, then it makes sense that they hire 2--3 additional lecturers on part-time contracts as a replacement.
+Furthermore, if a university must teach more students, but cannot afford to hire a full-time professor who teaches 2--3 classes a year, then it makes sense that they hire 2--3 additional lecturers on part-time contracts for each class as a replacement.
 This provides simple intuition for why the lecturers' elasticity is 2--3 times larger than that for assistant/full professors in \autoref{tab:facultycount-shock-reg}.
-While no previous research causally connects funding cuts to substituting professors for lecturers, the phenomenon has been alluded to in popular media \citep{wiu2016} and university rankings (where more lecturers hurts a university's ranking, \citealt{usnews2023}); these results provides real evidence for these claims.
+While no previous research causally connects funding cuts to substituting professors for lecturers, the phenomenon has been alluded to in popular media \citep{wiu2016} and university rankings (where more lecturers hurts a university's ranking, \citealt{usnews2023}).
+The results provided here give the first real evidence for such popular claims.
 
 \subsection{Illinois Faculty, IBHED}
-Incumbent professors at Illinois universities are relatively unaffected by state funding cuts for their university.
+Incumbent professors at Illinois universities are relatively unaffected by state funding cuts at their university.
 \autoref{tab:faculty-shock-illinois-rolling} show estimates of a 10\% change in state funding per student on faculty salaries, rate of exiting the Illinois public university system, and rate of promotion.
@@
 These results show that faculty --- both lecturers and professors --- are relatively unaffected by state funding cuts, in the Illinois public university system.
-Illinois is a state with a large public university system, relative to student enrolment, with trends in state funding for education similar to national trends, and with both selective and non-selective universities (\autoref{sec:trends-illinois}).
+Illinois is a state with a large public university system, relative to student enrolment, with trends in state funding for education similar to national trends, and with both selective and non-selective universities.
 If faculty are unaffected in Illinois, then it is unlikely that faculty in the rest of the nation are affected on average.
@@
 The IBHED data do not measure the entire pool of academics applying for faculty jobs at Illinois universities, but only those who applied to, were accepted, and then started a job as a lecturer or professor.
-These are incumbent faculty, and it is clear they are unaffected by state funding cuts.
+These are incumbent faculty, and the data show they are unaffected by state funding cuts.
 This raises a question: how can funding cuts lead to a large substitution of professors for lecturers, when incumbent faculty are not affected?
@@
 \subsection{Impacts on Faculty Hiring}
-Turn-over is the primary channel that changes in faculty composition happen, where, where faculty exit their jobs and new faculty are hired to replace those leaving --- or no more are hired to replace those who left.
+Turn-over is the primary channel that changes in faculty composition happen, where faculty exit their jobs and new faculty are hired to replace those leaving --- or no more are hired to replace those who left.
 \autoref{fig:exit-illinois-lp-rolling} shows that Illinois faculty are no more likely to exit the Illinois system (either retire or take a new job elsewhere) following state funding cuts, so there is one remaining channel for the substitution of professors for lecturers: increased hiring of lecturers and reduced hiring of new professors.
@@
 In general, faculty counts are relatively unrelated to state funding among Illinois universities; though the point estimates line up with analysis in IPEDS, the smaller number of observations mean that the precision of estimates are small and indistinguishable from zero, in both raw counts and in log units.
-On the other hand, universities with more state funding hire more professors across 2011--2021, as seen in \autoref{fig:hiring-correlation}.
-An extra 10\% in state funding is associated with 8.5\% more professors hired per students.
+In the national-level \cite{wapman2022quantifying} data, universities with more state funding hire more professors across 2011--2021.
+An extra 10\% in state funding is associated with 8.5\% more professors hired per students, as seen in \autoref{fig:hiring-correlation}.
 I use these data as part of a shift-share analysis, showing that a 10\% state funding cut per student leads to 13\% fewer professor hires (relative to enrolment), and that these hires are roughly equal for gender of professor hired.
@@
 State funding cuts lead to lower professor hiring, and the persistent funding cuts to public universities led to lower professor hiring across the last three decades.
-The channels for faculty turn-over were unaffected by state funding cuts in the analysis of Illinois faculty, and rate professor hiring is related to professor hiring in both correlational terms and in a shift-share analysis.
-This means that effects on faculty hiring were likely the main mechanism for how state funding cuts led to a substitution of professors for lecturers at public universities.
+The channels for faculty turn-over were unaffected by state funding cuts in the analysis of Illinois faculty, while professor hiring is related to funding cuts in both correlational terms and in a shift-share analysis.
+Combining these pieces of evidence means that reduced professor hiring was likely the main mechanism for how state funding cuts led to a substitution of professors for lecturers at public universities.
 
@@
 For example, selective institutions may have received funding cuts for other reasons, or states with lower performing institutions or declining populations strategically cut funding for only less selective universities.
-I address concerns for this with three different analyses.
+I address these concerns with three different analyses.
 
@@
 Secondly, I return to the primary specification in \autoref{eqn:secondstage}, and estimate effect of state funding cuts on faculty counts separately among universities of different selectivity rates.
-This investigates whether results are consistent across different levels of selectivity, and whether less selective universities have unreasonably high estimates for substitution --- as would be the case if selectivity was an unaddressed confounder.
+This investigates whether results are consistent across different levels of selectivity, and whether less selective universities have unreasonably high estimates for substitution --- as would be the case if selectivity were an unaddressed confounder.
 \autoref{tab:facultycount-heterogeneity} collects these results together, separated by the \cite{barrons2009} selectivity rankings.

### 2024-05-28 programs/analysis/ipeds-public-private.R
Updated all code files, to ensure replicability.  Time to Zenodo archive.

@@ -75,7 +75,7 @@ ipeds.data %>%
         total_count = n(),
         enrollment_reported = sum(enrollment_reported, na.rm = TRUE)) %>%
     ungroup() %>%
-    View() #print()
+    print() #View()
 
 # Show enrollment at public vs private over time: total + average
 enrollment.data <- ipeds.data %>%
@@ -288,25 +288,3 @@ all_fte_perprof.plot <- prof_perfte.data %>%
 ggsave("../../text/figures/all-fte-perprof.png",
     plot = all_fte_perprof.plot,
     units = "cm", width = fig.width, height = fig.height)
-
-
-# Get figures in revenues (total + per student) public vs private --------------
-mean_funding.data <- ipeds.data %>%
-    group_by(year, public) %>%
-    summarise(
-        totalrevenues_mill =
-            mean(totalrevenues_real, na.rm = TRUE) / (10^6),
-        nonauxrevenues_mill =
-            mean(nonauxrevenues_real, na.rm = TRUE) / (10^6),
-        stateappropriations_mill =
-            mean(stateappropriations_real, na.rm = TRUE) / (10^6),
-        tuitionrev_mill =
-            mean(tuitionrev_real, na.rm = TRUE) / (10^6),
-        totalrevenues_perEnroll =
-            mean(totalrevenues_real / enrollment_reported, na.rm = TRUE),
-        nonauxrevenues_perEnroll =
-            mean(nonauxrevenues_real / enrollment_reported, na.rm = TRUE),
-        stateappropriations_perEnroll =
-            mean(stateappropriations_real / enrollment_reported, na.rm = TRUE),
-        tuitionrev_perEnroll =
-            mean(tuitionrev_real / enrollment_reported, na.rm = TRUE))

### 2024-05-28 programs/analysis/ipeds-shock.R

@@ -1578,68 +1578,6 @@ stargazer(
     out = "../../text/tables/expenditures-shock-reg-fte.tex")
 
 
-## Salaries instruction spending.
-# Naive OLS Regression
-naive_instruction_salaries.reg <- research_instruct.data %>%
-    mutate(`log(stateappropriations_real/enrollment_reported)(fit)` =
-        log(stateappropriations_real / enrollment_reported)) %>%
-    felm(log(instructionspending_salaries_real / enrollment_reported) ~ 1 +
-        `log(stateappropriations_real/enrollment_reported)(fit)` |
-        unitid + year |
-        0 |
-        state + year,
-        data = .)
-# Shift-share IV Regression, explained by state appropriation shock
-shiftshare_instruction_salaries.reg <- research_instruct.data %>%
-    felm(log(instructionspending_salaries_real / enrollment_reported) ~ 1 |
-        unitid + year |
-        (log(stateappropriations_real / enrollment_reported) ~
-            I(-log(appropriationshock_perEnroll_real))) |
-        state + year,
-        data = .)
-
-## Salaries research spending.
-# Naive OLS Regression
-naive_research_salaries.reg <- research_instruct.data %>%
-    mutate(`log(stateappropriations_real/enrollment_reported)(fit)` =
-        log(stateappropriations_real / enrollment_reported)) %>%
-    felm(log(researchspending_salaries_real / enrollment_reported) ~ 1 +
-        `log(stateappropriations_real/enrollment_reported)(fit)` |
-        unitid + year |
-        0 |
-        state + year,
-        data = .)
-# Shift-share IV Regression, explained by state appropriation shock
-shiftshare_research_salaries.reg <- research_instruct.data %>%
-    felm(log(researchspending_salaries_real / enrollment_reported) ~ 1 |
-        unitid + year |
-        (log(stateappropriations_real / enrollment_reported) ~
-            I(-log(appropriationshock_perEnroll_real))) |
-        state + year,
-        data = .)
-
-# Collate the results to a LaTeX table
-stargazer(
-    naive_instruction_salaries.reg, shiftshare_instruction_salaries.reg,
-    naive_research_salaries.reg, shiftshare_research_salaries.reg,
-    dep.var.caption = "Dependent Variable: Salary Expenditures",
-    dep.var.labels = c("Instruction", "Research", "All"),
-    column.labels = rep(c("OLS", "2SLS"), 3),
-    digits = digits.no,
[…]

### 2024-05-28 programs/run.sh

@@ -41,7 +41,12 @@ cd ..
 cd ../text
 latexmk -pdf paper.tex
 latexmk -c
+# Remove the stubborn digits in one table
+sed -i -e 's/.00//g' tables/summary-quantiles.tex
+
+# Host the draft as the main file in the replication package.
 cp paper.pdf ../state-funding-faculty-2024.pdf
+
 # Put the most recent presentation into the main folder.
 cd ../presentation
 cp presentation.pdf ../state-funding-faculty-presentation.pdf

### 2024-09-17 text/paper.tex
Made some edits.

@@
         This project's repository, including all materials and anonymised data needed for replication, is publicly available at 
-        \url{https://github.com/shoganhennessy/state-funding-faculty}.
-        Any comments or suggestions may be sent to me at \href{mailto:seh325@cornell.edu}{\nolinkurl{seh325@cornell.edu}}, or raised as an issue on the Github project.
+        \url{https://doi.org/10.5281/zenodo.11373226}.
+        Any comments or suggestions may be sent to me at \href{mailto:seh325@cornell.edu}{\nolinkurl{seh325@cornell.edu}}, or raised as an issue on the \href{https://github.com/shoganhennessy/state-funding-faculty/}{Github project}.
     } \\
@@
     Stagnating State Funding for Higher Education and its Effect on Faculty}
-\date{This version: \today \\ \vspace{0.1cm}
+\date{First version: 1 September 2022 \\ \vspace{0.15cm}
+    This version: \today \\ \vspace{0.15cm}
     \href{https://github.com/shoganhennessy/state-funding-faculty/blob/main/state-funding-faculty-2024.pdf}{Newest version available here.}
+    \\ \vspace{-0.5cm}
 }
@@
 % Rev of Econ Ed Statements.
+\subsubsection*{Credit Author Statement}
+As the only author on this paper, I am responsible for all contributions in this paper.
+
 \subsubsection*{Data Statement}
 Every source of data, and the code to fully replicate the entire analysis is freely available at the following link:
-\url{https://github.com/shoganhennessy/state-funding-faculty}.
+\url{https://doi.org/10.5281/zenodo.11373226}.
+
 IBHED data, while public knowledge, include personally identifiable names, so the data included in the replication package have names replaced with non-identifiable placeholders.
-\subsubsection*{Credit Author Statement}
-As the only author on this paper, I am responsible for all contributions in this paper.
+
 \subsubsection*{Declaration of Competing Interest}
 I have no competing interests to declare.
+
+This research did not receive any specific grant from funding agencies in the public, commercial, or not-for-profit sectors.
+
 % Separated Figures \& Tables

### 2024-09-17 text/sections/01-introduction.tex
Made some edits.

@@
 %% Introduction section
-%\section{Introduction}
-%\label{sec:intro}
+\section{Introduction}
+\label{sec:intro}
 States fund their public universities with roughly the same amount today as they did in the 1990s, despite enrolment rising by over 50\%.
@@
 
-per-student state funding for higher education has stagnated the last thirty years.
+Per-student state funding for higher education has stagnated the last thirty years.
 The average public university received around \$11,600 of state funding per-student in 1990 and \$8,300 in 2021, while the number of lecturers per-student increased by 109\% and professors per-student decreased by $23$\% over 1990--2021.
@@
 The closest related research on higher education funding examined how university spending affected graduation rates and levels of student debt \citep{NBERw23736,NBERw27885}, and university finances \citep{miller2022making,bound2019public,brown2014endowment}.
-For faculty outcomes, descriptive work measures how US and Canadian faculty salaries are correlated with other factors \citep{bratsberg2003negative,martinello2009faculty,pfeffer1992salary,hilmer2020labor}, and applied theory work models university decision-making and faculty hiring (see e.g., \citealt{abe2015implications,johnson2009jep,NBERc13879}).
+For faculty outcomes, descriptive work measures how US faculty salaries are correlated with other factors \citep{hilmer2020labor}, and applied theory work models university decision-making and faculty hiring (see e.g., \citealt{abe2015implications,johnson2009jep,NBERc13879}).
 Other papers measure university decision--making after endowment shocks \citep{brown2014endowment}, and trends in the academic job market following the 2008 recession \citep{turner2014impact}.
@@
 Universities that rely on more lecturers have lower student retention and graduation rates \citep{jaeger2011examining,ehrenberg2005tenured}.
-Similarly, students at a large public university have consistently worse education outcomes if taught by adjunct lecturers than those taught by full-time faculty, largely thanks to the worse working conditions adjunct lecturers face \citep{zhu2021limited}.\footnote{
+Similarly, students at a large public university have consistently worse education outcomes if taught by adjunct lecturers than those taught by full-time faculty, largely thanks to the worse working conditions adjunct lecturers face \citep{zhu2021limited}.
 Alternatively, \cite{bettinger2010does,figlio2015tenure} study lecturers at a highly selective universities, finding that adjunct lecturers improve subject retention for undergraduates compared to professors.
 This paper studies faculty at public universities, which are by-majority not selective, and where lecturers are more often employed part-time with less institutional support.
-}
 Combining the evidence that adjunct lecturers lead to worse student outcomes with the trends in education finances and performance \citep{NBERw23736,NBERw27885}, implies that substituting towards lecturers contributed to deteriorating student outcomes at public universities.

### 2024-09-17 data/urban-ipeds/documentation/get.sh

@@ -1,3 +1,0 @@
-version https://git-lfs.github.com/spec/v1
-oid sha256:e3f25bf495b2621bb18e3555563daf1e38f65d0019647943d5a835b4efdee965
-size 473

### 2024-09-17 programs/analysis/ipeds-shock.R

@@ -723,26 +723,26 @@ reg.data %>%
     group_by(instrument_quantile) %>%
     summarise(
         `IV Components, \\$ per student:` = NA,
-        `Funding shift--share`        = round(mean(- appropriationshock_perEnroll_real, na.rm = TRUE)),
-        `Shift in state--wide funding`   = round(mean(- allstate_stateappropriations_real / stateEnroll_count, na.rm = TRUE)),
-        `Share reliance on state funding, \\% in 1990--1993`    = round(mean(staterevenues_baseshare, na.rm = TRUE), 2),
+        `Funding shift--share`        = mean(- appropriationshock_perEnroll_real, na.rm = TRUE),
+        `Shift in state--wide funding`   = mean(- allstate_stateappropriations_real / stateEnroll_count, na.rm = TRUE),
+        `Share reliance on state funding, \\% in 1990--1993`    = 100 * mean(staterevenues_baseshare, na.rm = TRUE),
         `\\hline University Funding and Spending, \\$ millions:` = NA,
-        `State funding`               = round(mean(stateappropriations_real / 10^6, na.rm = TRUE)),
-        `Tuition revenue`             = round(mean(tuitionrev_real / 10^6, na.rm = TRUE)),
-        `Total non-inst. revenues`    = round(mean(nonauxrevenues_real / 10^6, na.rm = TRUE)),
-        `Instruction spending`        = round(mean(instructionspending_total_real / 10^6, na.rm = TRUE)),
-        `Research Spending`           = round(mean(researchspending_total_real / 10^6, na.rm = TRUE)),
+        `State funding`               = mean(stateappropriations_real / 10^6, na.rm = TRUE),
+        `Tuition revenue`             = mean(tuitionrev_real / 10^6, na.rm = TRUE),
+        `Total non-inst. revenues`    = mean(nonauxrevenues_real / 10^6, na.rm = TRUE),
+        `Instruction spending`        = mean(instructionspending_total_real / 10^6, na.rm = TRUE),
+        `Research Spending`           = mean(researchspending_total_real / 10^6, na.rm = TRUE),
         `\\hline University Funding and Spending, \\$ per student` = NA,
-        `fte State funding`              = round(mean(stateappropriations_real / enrollment_reported, na.rm = TRUE)),
-        `fte Tuition revenue`            = round(mean(tuitionrev_real / enrollment_reported, na.rm = TRUE)),
[…]

### 2024-09-17 programs/run.sh

@@ -41,8 +41,6 @@ cd ..
 cd ../text
 latexmk -pdf paper.tex
 latexmk -c
-# Remove the stubborn digits in one table
-sed -i -e 's/.00//g' tables/summary-quantiles.tex
 
 # Host the draft as the main file in the replication package.
 cp paper.pdf ../state-funding-faculty-2024.pdf

### 2024-09-18 text/sections/04-empirical.tex
Made some edits.

@@
 This assumption is plausible given that the majority of states have multiple (i.e., more than five) public universities, without any single university campus receiving the majority of state funding within any single state.
-Secondly, the shift-share identification strategy requires exogeneity in either the base-line share or shift component of the instrument \citep{borusyak2022quasi}.
+Secondly, the shift-share identification strategy requires exogeneity in either the base-share or shift component of the instrument \citep{borusyak2022quasi}.
 We satisfy the second: faculty or university institutional-level decisions are not correlated with contemporaneous (or upcoming shocks) to state funding, so that the shift-share approach yields causal estimates by IV \citep{NBERw27885}.\footnote{
-    It would be plausible to consider the case that universities make institutional-level decisions in a consistently different manner to those with differing reliance on state funding in 1990, so that exogeneity by the base-line share is not plausible here.
+    It would be plausible to consider the case that universities make institutional-level decisions in a consistently different manner to those with differing reliance on state funding in 1990, so that exogeneity by the base-share is not plausible here.
 }
@@
 However, the assumption is bolstered by the fact that universities with higher values of the shift-share instrument differ little in observed characteristics compared to those with lower values \citep{pei2019poorly}.
-Universities are by-in-large smaller (in terms of enrolment, total revenues, and professor count) at the top of the shift-share instrument distribution (see~\autoref{tab:summary-quantiles}).
+Universities are by and large smaller (in terms of enrolment, total revenues, and professor count) at the top of the shift-share instrument distribution (see~\autoref{tab:summary-quantiles}).
 Though on a per-student basis, there is little difference across the distribution (except in the outcomes under consideration).
@@
 The local projections method is an empirical model used to estimate dynamic treatment effects when the treatment is not binary, when time-series confounding is present \citep{montiel2021local}.
-Faculty outcomes could be affected in only a transitory manner, where effects in later years are close to zero, or in a persistent manner, where the effects of state funding cuts linger for multiple years.
+Faculty outcomes could be only affected in a transitory manner, where effects are limited to the short-run and fade away quickly, or persistently, where the effects of state funding cuts linger for many years.
 This method allows me to estimate the effect of state funding $X_{i,t}$ on faculty outcomes $Y_{i, t+ k}$ in follow-on years $k > 0$, credibly testing whether effects are transitory or long-lasting.
@@
 }
+
+\textbf{An extra paragpragh here, showing what LP does.  Give the equation.}

### 2024-09-18 text/sections/06-conclusion.tex
Made some edits.

--- 2024-05-08 text/sections/07-conclusion.tex
@@
 The effects on incumbent professors in Illinois are non-distinguishable from zero, and suggestive evidence shows that funding cuts limited hiring of new professors, implying that substitution towards lecturers is driven by reduced hiring of new professors.
-Public universities are using more contingent lecturers to teach their students, while private universities continue to employ more tenure-track and tenured professors than their public counterparts, and each year the gap widens.
+Public universities are using more contingent lecturers and fewer professors to teach their students, and each year this trend grows.
 
 This work contributes to the literature along two primary dimensions.
-First, this work provides an explanation for the increased reliance on lecturers and away from tenure-track and tenured professors at US public universities, by isolating the effects funding cuts from state-wide funding shifts in both the short-- and long--run.
+First, this work provides an explanation for the increased reliance on lecturers and away from tenure-track and tenured professors at US public universities, by isolating the effects funding cuts from state-wide funding shifts in both the short- and long-run.
 Second, this work investigates how individual faculty are affected by changes in state funding for their university using a dataset new in the economic literature (IBHED).
@@
 I show that these financial headwinds led to systematic change at public universities, leading to a substitution away from stable employment contracts (tenure, professorship) and towards less stable arrangements (adjunct contracts, lecture positions).
-These results show large changes in faculty composition, and that stagnation in state support explains at least a third of the observed shift away from tenured professors and towards lecturers.
+These results show large changes in faculty composition, and that stagnation in state funding explains at least a third of the observed shift away from tenured professors and towards lecturers.
 This likely had knock-on effects on instruction, contributing to worsening student outcomes at US public universities since the early 1990s --- while private universities were not exposed to financial headwinds of the same magnitude or persistence over the same time period.
-Furthermore, public universities have been making personnel decisions, with long-run effects on their institutional capacity, based on short run funding constraints.
-While public universities educate the majority of US higher education students, we should worry about the effects of restricting their funding has on faculty, and the wider impact on higher education as a whole.
+Furthermore, public universities have been making personnel decisions, with long-run effects on their institutional capacity, based on short-run funding constraints.
+While public universities educate the majority of US higher education students, we should worry about the effects of stagnating state funding on faculty, and the wider impact on higher education as a whole.

### 2024-09-25 text/paper.tex
Made some edits.

@@
         I thank seminar participants at Cornell University (2023, 2024), and the Eastern Economic Association Annual Meetings (2024) for helpful discussion.
-        This project's repository, including all materials and anonymised data needed for replication, is publicly available at 
+        This project's repository, including all materials and anonymised data needed for replication, is made publicly available: 
         \url{https://doi.org/10.5281/zenodo.11373226}.
@@
 \input{sections/05-results.tex}
-% Dicussion
-%\input{sections/06-discussion.tex}
 % Conclusion
-\input{sections/07-conclusion.tex}
+\input{sections/06-conclusion.tex}
 % Rev of Econ Ed Statements.
@@
 As the only author on this paper, I am responsible for all contributions in this paper.
-
 \subsubsection*{Data Statement}
-Every source of data, and the code to fully replicate the entire analysis is freely available at the following link:
+Every source of data, and code to fully replicate the entire analysis, is freely available at the following link:
 \url{https://doi.org/10.5281/zenodo.11373226}.
-
 IBHED data, while public knowledge, include personally identifiable names, so the data included in the replication package have names replaced with non-identifiable placeholders.
-
 \subsubsection*{Declaration of Competing Interest}
 I have no competing interests to declare.
-
 This research did not receive any specific grant from funding agencies in the public, commercial, or not-for-profit sectors.
-
 % Separated Figures \& Tables
-\input{sections/09-figures-tables.tex}
+\input{sections/08-figures-tables.tex}
 % Bibliography
@@
 \bibliographystyle{agsm}
-\bibliography{sections/08-bibliography.bib}
+\bibliography{sections/07-bibliography.bib}
 % Appendix
-\input{sections/10-appendix.tex}
+\input{sections/09-appendix.tex}
 \end{document}

### 2024-09-25 text/sections/01-introduction.tex
Made some edits.

@@
 Previous research shows state funding cuts led to worsening student outcomes \citep{NBERw23736,NBERw27885}, but there is little evidence on whether these funding cuts impacted faculty.
-In this paper, I show that falling state funding caused a change in the composition of faculty at US public universities, substituting professors for more lecturers.
+In this paper I show that falling state funding caused a change in the composition of faculty at US public universities, substituting professors for more lecturers.
 
@@
 The average public university received around \$11,600 of state funding per-student in 1990 and \$8,300 in 2021, while the number of lecturers per-student increased by 109\% and professors per-student decreased by $23$\% over 1990--2021.
-I use a shift-share instrumental variables approach to estimate how faculty are affected by these funding cuts.
+I use a shift-share instrumental variables (IV) approach to estimate how faculty are affected by these funding cuts.
 Universities employ $4.4$\% more lecturers per-student following a $10$\% funding cut, and correspondingly rely less on tenure-track faculty by employing $1.4$\% fewer tenure-track professors and $1.2$\% fewer tenured professors per-student.
@@
 My second analysis uses panel data for every faculty member at Illinois public universities for 2010--2021 \citet[IBHED]{ibhed}, to measure how incumbent faculty (i.e., those already employed at the university) are affected by changes in state funding.
-These data allow me to answer the following research question: if a US public university receives an extra \$1,000 per-student (or a 10\% increase), do they change how many faculty members they employ?
-If so, which positions do universities substitute away from, and do these funding cuts affect the faculty themselves?
+These data allow me to answer the following research question: if a US public university has its funding cut by \$1,000 per-student (or a 10\% cut), do they change how many faculty members they employ?
+If so, which positions do universities substitute towards, and do these funding cuts affect the faculty themselves?
 
@@
 This means that correlations between state funding and faculty outcomes do not necessarily represent causal relationships.
-As such, it would be na\"ive to think of state higher education funding as independent of public university faculty, so I employ a shift-share instrumental variables (IV) approach to identify causal effects of state funding cuts.
-The shift-share instrument identifies exogenous changes in state funding by exploiting differences in how much universities relied on state funding, interacted with state-wide shifts in higher education funding --- following the methods of \cite{NBERw23736,NBERw27885}.
+As such, it would be na\"ive to think of state higher education funding as independent of public university faculty, so I employ a shift-share IV approach to identify causal effects of state funding cuts.
+The shift-share instrument identifies exogenous changes in state funding by exploiting differences in universities' reliance on state funding, interacted with state-wide shifts in higher education funding --- following the methods of \cite{NBERw23736,NBERw27885}.
 Yearly state government budgets are decided without targetting individual universities for funding cuts, so more reliant universities will be more affected by state-wide funding cuts, so the shift-share approach estimates causal effects among more reliant universities.
@@
 
-Mine is the first paper to provide causal evidence on the impact of state funding on faculty composition and wages.
-The closest related research on higher education funding examined how university spending affected graduation rates and levels of student debt \citep{NBERw23736,NBERw27885}, and university finances \citep{miller2022making,bound2019public,brown2014endowment}.
+I provide the first causal evidence on the impact of state funding on faculty composition and wages.
+In a contemporaneous working paper, \cite{hinrichs2022state} use the same national-level variation in university funding to analyse effects on faculty and wages, coming to different conclusions.\footnote{
+    \cite{hinrichs2022state} considers a shorter time-period than this paper (2010--2021), uses only national variation in IPEDS data, relies on IPEDS data for faculty salaries, and has no individual-level analysis of faculty or their salaries.
+    See \autoref{sec:data-desc,sec:recession-analysis} for an elaboration on these issues.
+}
+The closest related research on higher education funding examined how university spending affected graduation rates and levels of student debt \citep{NBERw23736,NBERw27885}, and university finances \citep{miller2022making,bound2019public,webber2017state}.
 For faculty outcomes, descriptive work measures how US faculty salaries are correlated with other factors \citep{hilmer2020labor}, and applied theory work models university decision-making and faculty hiring (see e.g., \citealt{abe2015implications,johnson2009jep,NBERc13879}).

### 2024-09-25 text/sections/02-data.tex
Made some edits.

@@
 \subsection{Data Description}
+\label{sec:data-desc}
 The data used in this paper come from two primary sources: \citet[IPEDS]{ipeds} for data on university funding, finances, and enrolment, and \citet[IBHED]{ibhed} for data on every faculty member in the Illinois public university system.
@@
 For-profit institutions employ and enrol a negligible share of professors and students respectively, while students at two-year institutions by majority intend to eventually enrol at a four-year institution \citep{mountjoy2022}, so that these institutions are not considered.
-IPEDS reports the count of professors employed by position.\footnote{
+IPEDS reports the count of professors employed by position; IPEDS data on average faculty salaries are inconsistent with trends in faculty salaries from other sources, so I do not consider these data reliable and conduct no analysis involving them.\footnote{
     IPEDS gives a mean salary measure for faculty by rank, but these figures have many missing values and disagree with calculated values from other reputable sources.
@@
     IPEDS values consistently report that public university professors are paid on average 20\% more than private sector faculty, which disagrees with other sources on the salary gap direction, and IPEDS figures are consistently higher than averages calculated from individual-level IBHED data.
-    For these reasons I do not further analyse IPEDS salary data, and use salary data from IBHED to analyse faculty salaries.
+    For these reasons I do not further analyse IPEDS salary data, and use salary data from IBHED (which do closely agree with trends in other sources) to analyse faculty salaries.
     Notably, IPEDS data on faculty counts do not suffer from these problems.
+}\footnote{
+    In a contemporaneous working paper, \cite{hinrichs2022state} analyses the effect of state funding cuts on university faculty, using the same shift-share variation.
+    \cite{hinrichs2022state} bases the shift--share instrument in the year 2011, despite major upheavals in the level of state funding for state funding for higher education around the 2008 Great Recession, and years following (see \autoref{sec:trends});
+    \cite{hinrichs2022state} also does not mention the data inconsistencies in IPEDS data for average faculty salaries.
+    These two major differences explain the different conclusions that our contemporary papers come to; \autoref{sec:recession-analysis} replicates the \cite{hinrichs2022state} empirical findings with my primary specification, and further discuses the differences in findings.
 }

### 2024-09-25 text/sections/04-empirical.tex
Made some edits.

@@
     \left( \frac{\text{Total State Funding}_{s(i),t}}{\text{Student Population}_{s(i),t}} \right)
-    \sum_{\tau = 0}^{3} \frac 14
-    \left( \frac{\text{State Funding}_{i,1990 + \tau}}{\text{Total Revenues}_{i,1990 + \tau}} \right) \right].
+    \sum_{k = 0}^{3} \frac 14
+    \left( \frac{\text{State Funding}_{i,1990 + k}}{\text{Total Revenues}_{i,1990 + k}} \right) \right].
 \end{align}
@@
     \label{foot:control}
-    \cite{NBERw27885} note the tendency for public universities to respond to state funding cuts by increasing reliance on tuition, where \cite{NBERw23736} specifically instruments for tuition revenues with collected information on legislative tuition price controls.
+    \cite{NBERw27885,webber2017state} note the tendency for public universities to respond to state funding cuts by increasing reliance on tuition, where \cite{NBERw23736} specifically instruments for tuition revenues with collected information on legislative tuition price controls.
     It may be argued that tuition revenues are confounder between the causal effect of changes in state funding on a university's total revenues, so that this analysis focuses on state support for higher education (not total revenue), as do \cite{NBERw27885}.
@@
 I use the instrument defined in \autoref{eqn:public-instrument} to overcome the endogeneity concerns for state funding to each public university, for an IV approach.
-\autoref{eqn:firststage} is the first stage for exogenous variation in the state funding for university $i$ in year $t$, and \autoref{eqn:secondstage} the second stage for the effect of state funding on faculty outcomes.\footnote{
+\autoref{eqn:firststage} is the first-stage for exogenous variation in the state funding for university $i$ in year $t$, and \autoref{eqn:secondstage} the second stage for the effect of state funding on faculty outcomes.\footnote{
     The IV approach identifies the local average treatment effect, weighted towards larger values of the instrument in the continuous case.
@@
 }
-$Y_{i,t}$ represents faculty outcomes for university $i$ in year $t$, $\alpha_i, \gamma_t$ university and year fixed effects, and $\widehat X_{i,t}$ state funding for university $i$ estimated in first stage \eqref{eqn:firststage}.
+$Y_{i,t}$ represents faculty outcomes for university $i$ in year $t$, $\alpha_i, \gamma_t$ university and year fixed effects, and $\widehat X_{i,t}$ state funding for university $i$ estimated in first-stage \eqref{eqn:firststage}.
 
@@
 \label{sec:local-projections}
-The effects of a change in the universities funding may not be immediate, and faculty may be affected multiple years after a funding cut.
-Yet, the funding shift-share instrument is significantly auto-correlated year-on-year; a large state funding cut in year $t$ is also likely to experience a large funding cut in years $t \pm 1$.
-\autoref{fig:lag-firststage} shows that the shift-share instrument is correlated with state funding five years into the past and future.
-This level of time-series correlation would confound IV estimates of the effect of state funding in year $t$ on faculty outcomes in later years $t+k$, providing inconsistent estimates for the causal relationships when using two-stage least squares.
+Faculty outcomes could be only affected in a transitory manner, where effects are limited to the short-run and fade away quickly, or persistently, where the effects of state funding cuts linger for many years.
+I employ a Local Projections (LP) approach to answer the following question: are university faculty affected multiple years after a state funding cut to their university?
+The LP method is an empirical model used to estimate dynamic treatment effects when time-series confounding is present, and accommodates an instrument \citep{jorda2005,olea2021inference}.
+\begin{align}
+    \label{eqn:firststage-lp}
+    X_{i,t}   &= \eta_i + \zeta_t + \delta Z_{i,t} + \epsilon_{i,t} \\
+    \label{eqn:secondstage-lp}
+    Y_{i,t+k} &= \alpha_i + \gamma_t + \beta_k \widehat X_{i,t} + \varepsilon_{i,t+k}
+\end{align}
+The approach estimates the first-stage \eqref{eqn:firststage-lp} by ordinary least squares; here this approach is the same as in conventional IV analysis, instrumenting for a university's state funding in year $t$ with state-wide funding shifts in year $t$.
+The second-stage \eqref{eqn:secondstage-lp} is estimated by a regression of 
+faculty outcomes in future years $t+k$ on (predicted) state funding cuts in year $t$.
+Fixed effects, and two-way standard error clustering are maintained throughout, as before.
 
-I employ a local projections approach to model how faculty outcomes in years $t+k$ are affected by state funding in year $t$, for future years $k > 0$ \citep{jorda2005}.
-The local projections method is an empirical model used to estimate dynamic treatment effects when the treatment is not binary, when time-series confounding is present \citep{montiel2021local}.
-Faculty outcomes could be only affected in a transitory manner, where effects are limited to the short-run and fade away quickly, or persistently, where the effects of state funding cuts linger for many years.
-This method allows me to estimate the effect of state funding $X_{i,t}$ on faculty outcomes $Y_{i, t+ k}$ in follow-on years $k > 0$, credibly testing whether effects are transitory or long-lasting.
-The approach accommodates the shift-share instrument for state funding \citep{olea2021inference}, so that I present results using the shift-share to instrument for state funding (as in the rest of the empirical analysis).\footnote{
-    The funding shift-share has a persistent effect on state funding, multiple years after the initial funding shift-share, so that the instrument is similarly strong for the local projections method --- see \autoref{sec:appendix-ipeds-firststage} further evidence on the strength of the first stage in local projections estimation.
+The LP method estimates the cumulative effect of a state funding cut in year $t$, and its cumulative follow-on funding cuts in intermediate years, on faculty outcomes in year $t+k$.\footnote{
+    See \citet[Section~8.3]{jorda2023local} for a technical definition, and discussion of the differences in estimands when accounting (or not) for the intermediate values.
 }
-
-\textbf{An extra paragpragh here, showing what LP does.  Give the equation.}
+This is because the funding shift-share instrument is significantly auto-correlated year-on-year, as shown in \autoref{fig:lag-firststage}.
+If Vermont cuts state funding for higher education in 1993, the LP method uses this observation's predicted state funding in year 1993 to calculate the effect of the funding cut in 1993, combined with the correlated funding cut in 1994, on faculty outcomes in 1995.
+The LP method then averages estimates across all years and all universities by two-stage least squares, calculating an estimate $\hat \beta_k$.
+This is an estimate of the effect of a funding cut in year $t$, and the average of accumulated funding cuts in intermediate years, on faculty outcomes in year $t+k$.\footnote{
+    The funding shift-share has a persistent effect on state funding, multiple years after the initial shift-share instrument, so that estimates in this setting represent both the effect of the initial funding cut and cumulative funding cuts in intermediate years.
+    See \autoref{sec:appendix-ipeds-firststage} for this demonstrated by estimating the first-stage by the LP method.
+}

### 2024-09-25 text/sections/05-results.tex
Made some edits.

@@
 
-\subsection{First Stage Estimates}
-The shift-share instrument has a strong first stage for universities' yearly state funding.
+\subsection{First-Stage Estimates}
+The shift-share instrument has a strong first-stage for universities' yearly state funding.
 \autoref{tab:firststage-reg} show results of the first-stage regression among IPEDS data, separately with and without a control for tuition revenue per student.
@@
 Estimates are presented for specifications with and without fixed effects.
-Column (2) shows estimates of the first-stage without including fixed effects, and gives less precise estimates for the funding shift-share --- this may be due to systematic differences in universities, which are not accounted for without fixed effects.
+Column (2) shows estimates of the first-stage without including fixed effects, and gives less precise estimates for the funding shift-share --- this may be due to systematic differences in universities, which are not accounted for when omitting fixed effects.
 Columns (3) and (4) include the tuition revenue as a control (explained in \autoref{foot:control}) to exhibit estimates with or without this control variable.
 Column (3) shows similar estimates to column (1) in both Panels A and B thanks to inclusion of fixed effects, so whether tuition revenue is a bad control does not affect empirical estimates.
-Column (1) represents the estimates for \autoref{eqn:firststage} with fixed effects (not including tuition revenue as a control variable), and is the preferred specification that I proceed with.
+Column (1) represents the estimates for \autoref{eqn:firststage} with fixed effects (not including tuition revenue as a control variable), and is the preferred specification that I proceed with;
+all specifications (except column 4 in both panels) have a large F statistic on the excluded instrument.\footnote{
+    I calculate the F statistic on the excluded instrument, while also accounting for fixed effects and two-way standard error clustering, by the methods provided in \cite{olea2013robust}.
+}
 I use the same empirical strategy in the analysis of Illinois faculty, and \autoref{tab:firststage-illinois} presents first-stage estimates among the IBHED data; results are similar to those calculated in IPEDS data.
@@
 
-The first stage is strong in both IPEDS and IBHED data, and the argument for exogeneity comes from independence of the state funding mechanism to any individual public university campus (as in \autoref{sec:empirics}).
+The first-stage is strong in both IPEDS and IBHED data, and the argument for exogeneity comes from independence of the state funding mechanism to any individual public university campus (as in \autoref{sec:empirics}).
 The IV approach also assumes that shifts in state-wide funding for higher education (the instrument) affect outcomes only via affecting state funding for the university (exclusion restriction).
 This assumption is fundamentally untestable, but is bolstered by the lack of relationship with university characteristics.
-\autoref{tab:summary-quantiles} shows the mean of various university variables, across the instrument distribution.
+\autoref{tab:summary-quantiles} shows the mean of various university variables, across the instrument distribution.\footnote{
+    The shift-share instrument is correlated with tuition revenues (and thus the sum of non-institutional revenues), so could be a presenting a threat to endogeneity.
+    As mentioned above, including tuition revenue as a control (even though it could also be a bad control) does not meaningfully change first- and second-stage results in any specification, rendering this plausible threat to identification moot.
+    See \autoref{foot:control} and \autoref{tab:firststage-reg}.
+}
 The university--years in the bottom 20\% of the instrument distribution have more research spending than those in the next 20\%; this is to be expected, as university spending is heavily related to state--wide shifts in funding.
-On the other hand, the measures of selectivity (acceptance rate, 6--year graduation rate) are not substantially different across the distribution of the instrument.
-This allays concerns that state-wide shifts in state funding are targetted at less selective universities (a possible confounder), and supports use of the shift-share IV empirical design.
+On the other hand, the measures of selectivity (acceptance rate, 6--year graduation rate) are noticeably different across the distribution of the instrument.
+While the shift--share instrument exhibits balance for many relevant university characteristics (except tuition revenues and selectivity), I include institution and year fixed effects throughout my analysis, effectively controlling for any possible confounding effects of differences between universities.
 
@@
 To allay this concern, I estimate the IV model on IPEDS data for areas of university spending  (\autoref{tab:expenditures-shock-reg}).
-This exercise gauges the effects of state funding cuts on areas of university spending to gauge whether the first stage is economically meaningful, and to see which areas of higher education operations are affected by state funding cuts.
+This exercise gauges the effects of state funding cuts on areas of university spending to gauge whether the first-stage is economically meaningful, and to see which areas of higher education operations are affected by state funding cuts.
 Following a state funding cut, universities spend less on instruction, student services, operations, and grant aid.
@@
 Universities do not re-adjust by hiring more professors to make up for last year's funding cut, and so the substitution towards lecturers becomes persistent.
-\autoref{fig:count-lp} shows local projection estimates, showing the elasticity for professor count per student in year $t+k$ with respect to state funding in year $t$.
+\autoref{fig:count-lp} shows LP estimates, showing the elasticity for professor count per student in year $t+k$ with respect to state funding in year $t$ (and cumulative intermediate cuts).
 State funding cuts affect the number of lecturers and professors at universities up to 9 years after the initial funding cut.
@@
 It could be the case that effects on faculty do not materialise until multiple years after a funding cut, though the data do not support this possibility.
-Local projection estimates show that professors' salaries, including salaries, are not significantly affected in years following funding cuts (\autoref{fig:salaries-illinois-lp-rolling}) --- they do not receive lower rates of pay raises, or delayed salary cuts, in response to state funding cuts.
+LP estimates show that professors' salaries, including salaries, are not significantly affected in years following funding cuts (\autoref{fig:salaries-illinois-lp-rolling}) --- they do not receive lower rates of pay raises, or delayed salary cuts, in response to state funding cuts.
 Similarly, promotion rate and rate of exit from the Illinois public university system are also unaffected in later years (\autoref{fig:promoted-illinois-lp-rolling}, \ref{fig:exit-illinois-lp-rolling}).
@@
 
-\subsection{Robustness Checks}
+\subsection{Robustness Checks and Heterogeneity Analysis}
 \label{sec:results-robustness}
@@
 For example, selective institutions may have received funding cuts for other reasons, or states with lower performing institutions or declining populations strategically cut funding for only less selective universities.
-I address these concerns with three different analyses.
+Additionally, the effect of state funding cuts may affect faculty at different universities to different extents
+I investigate these issues with three different analyses.
 
@@
 
-Secondly, I return to the primary specification in \autoref{eqn:secondstage}, and estimate effect of state funding cuts on faculty counts separately among universities of different selectivity rates.
-This investigates whether results are consistent across different levels of selectivity, and whether less selective universities have unreasonably high estimates for substitution --- as would be the case if selectivity were an unaddressed confounder.
-\autoref{tab:facultycount-heterogeneity} collects these results together, separated by the \cite{barrons2009} selectivity rankings.
+Second, I return to the primary specification in \autoref{eqn:secondstage}, and estimate effect of state funding cuts on faculty counts separately among universities of different selectivity rates.
+This investigates whether substitution rates are different across different levels of selectivity; \autoref{tab:facultycount-heterogeneity} collects these results together, separated by the \cite{barrons2009} selectivity rankings.
 Barrons rank 1-2 are the most selective (including e.g., UC Berkeley and SUNY Binghamton), 3-4 selective (including e.g., Clemson University and University of Dallas), and rank 5 for the rest of universities.
 The first-stage is strong in every group of universities, even among the small sample size of the 12 universities in the most selective group, and the 28 universities in the next selective group.
-Faculty substitution rates are not very strong among the most selective universities, but are very similar between selective universities and unranked universities, showing that even selective public universities exhibit the substitution towards lecturers.
+Faculty substitution rates are not very strong among the most selective universities, but are very similar between selective universities and unranked universities.
+One might expect that the effect is driven entirely by the least selective universities, as they are the most resource-constrained;
+this heterogeneity analysis shows that even selective (if not top) universities substitute towards lecturers following state funding cuts.
 
-The final robustness check I employ is a falsification test, testing whether the effects for public universities would be picked up among a sample of private universities.
-Private universities do not rely on state funding in the same that public universities do,\footnote{
-    Though private universities do receive some state funding, as some states fund for project at private universities and other smaller scale operating costs.
+The final robustness check I employ is a falsification test, testing whether the effects for public universities would be picked up among a sample of private universities.\footnote{
+    This is a robustness check similar to one performed in \citealt{NBERw27885}, as a robustness check for the same instrument.
+}
+Private universities do not rely on state funding in the same way that public universities do,\footnote{
+    Though private universities do receive some state funding, as some states fund smaller scale operating costs and other specific projects at private universities.
 }

### 2024-09-25 text/sections/06-conclusion.tex
Made some edits.

@@
 This work contributes to the literature along two primary dimensions.
-First, this work provides an explanation for the increased reliance on lecturers and away from tenure-track and tenured professors at US public universities, by isolating the effects funding cuts from state-wide funding shifts in both the short- and long-run.
+First, this work provides an explanation for the increased reliance on lecturers and away from tenure-track and tenured professors at US public universities, by isolating the immediate and long-lasting effects of funding cuts from state-wide funding shifts.
 Second, this work investigates how individual faculty are affected by changes in state funding for their university using a dataset new in the economic literature (IBHED).

### 2024-09-25 text/sections/09-appendix.tex
Made some edits.

--- 2022-11-20 text/sections/09-appendix.tex
@@
 \begin{itemize}
-    \item \textit{Tidyverse} \citep{tidyverse} collected tools for data analysis in the R language.
+    \item \textit{Tidyverse} \citep{tidyverse} collected tools for data analysis in the $R$ language.
     \item \textit{LFE} \citep{lfe} implemented linear fixed effect models, with instruments, crucial for the empirical estimation in \autoref{sec:empirics}.
-    \item \textit{Stargazer} \citep{stargazer} provided a method to efficiently convert empirical results into presentable output in \LaTeX.
+    \item \textit{Stargazer} \citep{stargazer} provided methods to efficiently convert empirical results into presentable output in \LaTeX.
     \item \textit{Lpirfs} \citep{lpirfs2019} implemented estimation of the \cite{jorda2005} local projections methods, with instrumental variables, crucial to the local projections estimates presented in this project.
@@
 
-\subsection{First Stage Estimates, Individual Outcomes}
-\label{appendix:part1}
+\subsection{IPEDS First Stage}
+\label{sec:appendix-ipeds-firststage}
+
+\begin{table}[H]
+    \singlespacing
+    \centering
+    \caption{Shift-Share Instrument Balance, Mean Characteristics Across Instrument Distribution.}
+    \makebox[\textwidth][c]{\input{tables/summary-quantiles.tex}}
+    \label{tab:summary-quantiles}
+    \justify
+    \footnotesize
+    \textbf{Note}:
+    This table shows the summary statistics for every public university--year observation in IPEDS data, for each of the 5 quantiles of the funding shift-share instrument.
+    Funding shift-share is the instrument defined in \autoref{sec:empirics}, the product of (second row) state-wide funding shift and (third row) share reliance on state funding.
+    State-wide funding shift is total funding for higher education in that university's state (divided by the count of state students); share reliance on state funding is the the total amount of state funding received by the university in 1990--1993 divided by total revenues for those years.
+    The column labelled ``1st'' refers to the mean for all university-year observations in the first quintile (bottom 20\%) of the funding shift-share distribution, and so on.
+    The numbers are adjusted to 2021 figures by CPI-U.
+    Non-institutional revenues refers to the sum of federal, state, and local funding plus tuition revenues; these sum to the majority of university funding, but exclude numbers such as university income from capital projects.
+    Acceptance rate and 6 year graduation rate are for university undergraduates, and are calculated from IPEDS data available for academic years 1997 through 2018.
+\end{table}
+
+\begin{figure}[H]
+    \centering
+    \singlespacing
+    \caption{Correlation Between State Funding Shift-Share and Public University State Funding in Surrounding Years.}
+    \includegraphics[width=\textwidth]{figures/lag-firststage.png}
+    \label{fig:lag-firststage}
+    \justify
+    \footnotesize
+    \textbf{Note}:
+    This figure shows the correlation between state funding in year $t+k$ with the funding shift-share in year $t$ for a university, where $k = -7, \hdots, 5$ are the years on the $x$-axis.
+    This shows that state funding and the funding shift-share are correlated across years, so that dynamic effects must be estimated by local projections --- and not simple OLS or 2SLS.
+    The estimates are of \eqref{eqn:firststage}, calculated with IPEDS data, separately for each year relative to initial shock, using the $\log-\log$ specification, including fixed effects for university $+$ year, and clustering standard errors by university $+$ year.
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \singlespacing
+    \caption{Local Projection Estimates for Funding Shift-Share on State Funding, in IPEDS Data.}
+    \includegraphics[width=0.6\textwidth]{figures/firststage-lp.png}
+    \label{fig:firststage-lp}
+    \justify
+    \footnotesize
+    \textbf{Note}:
+    These figures show the local projections estimates of regression specification \eqref{eqn:firststage}, with the funding shift-share as an instrument for state funding, using IPEDS data.
+    The coefficient estimate is effect of funding shift-share ($Z_{i,t}$) on state funding ($X_{i,t}$), while accounting for auto-correlation between different time periods --- i.e., between $Z_{i,t}, Z_{i,t-1}$ and $X_{i,t}, X_{i,t-1}$.
+    These results use a $\log-\log$ specification, so the estimates are for the elasticity of state funding per student in a year $t+k$ with respect to funding shift-share in year $t$, where years $k = 0, \hdots, 10$ are on the $x$-axis.
+    Standard errors are clustered at the state-year level.
+\end{figure}
+
+\begin{table}[H]
+    \singlespacing
+    \centering
+    \caption{Effect of State Funding on Areas of University Spending, IPEDS Data, IV Estimates.}
+    \textbf{Panel A: units in \$ per student}
+
+    \makebox[\textwidth][c]{\input{tables/expenditures-shock-rawcount.tex}}
+
+    \textbf{Panel B: units in log \$ per student}
+
+    \makebox[\textwidth][c]{\input{tables/expenditures-shock-reg-fte.tex}}
+
+    \label{tab:expenditures-shock-reg}
+    \justify
+    \footnotesize
+    \textbf{Note}:
+    These tables show the IV estimates of regression specification \eqref{eqn:secondstage}, showing the effect of state funding (instrumented with the shift-share) on areas of university spending.
+    Outcomes are measured in \$ student spending, so that the estimates representat the relationship between a \$1 increase in state funding on area of spending (or 1\% increase).    
+    The ``outcome mean'' rows are spending in thousands \$ per student.
+\end{table}
+
+\newpage
+\subsection{Illinois IBHED First Stage}
+\label{sec:iv-model-indiv}
+
+This paper uses data on individual professors in the Illinois university system, to investigate the effects of changes in university revenues on the individual professors at the universities.
+The outcomes here now refer to individual professors (e.g., their salary and promotion rate), so requires adjustment to the empirical approach, leveraging variation in university funding for the years after a professor joins the university.
+
+\autoref{eqn:rolling-instrument} defines a rolling-share variant of the instrument, $\tilde Z_{i(j),t}$, where the university's state funding share exposure is based in the year a professor joins the university --- and not the base period 1990--1993.
+$j$ indexes each professor in year $t$, $\tau(j)$ for the year the professor first joins their institution.
+Identifying $\tau(j)$ is possible for $j$ by restricting to all professors hired 2011-2021 --- i.e., in the years after the start of the full panel.
+It is not possible to discern the hiring year for professors who  were hired in the years preceding 2011, and so the entire sample is only possible to analyse using the base-share in years 1990-1993 formulation.
+
+\begin{align}
+    \label{eqn:rolling-instrument}
+    \tilde Z_{i(j),t} &\coloneqq - \left[
+    \left( \frac{\text{Total State Funding}_{s(j),t}}{\text{Student Population}_{s(j),t}} \right)
+    \left( \frac{\text{State Funding}_{\tau(j)}}{\text{Total Revenues}_{i,\tau(j)}} \right) \right]
+\end{align}
+
+This approach leverages an insight, made available by level of the data: that an individual professor is affected by changes in university revenues after they have joined the university.
+\autoref{sec:iv-model-uni} considers the number of professors employed by the university; whether a professor becomes employed at the university is likely affected by that university's finances.
+The formulation here takes as given that the professor is already employed at the university, and then projects the effect of changes in state funding on these \textit{incumbent} professors following the state funding shift-share.
+\autoref{tab:firststage-illinois} presents the first-stage results in Illinois data.
 
@@
     \centering
-    \caption{IBHED Summary Statistics, Entire Professor Panel 2010--2021.}
-    \makebox[\textwidth][c]{\input{tables/illinois-summary-rolling.tex}}
-    \label{tab:illinois-summary-rolling}
-    \begin{flushleft}
+    \caption{First Stage Estimates, for State Funding by Funding Shift-Share in IBHED Data.}
+    \textbf{Panel A: units in \$ per student}
+
+    \makebox[\textwidth][c]{\input{tables/firststage-reg-rawcount.tex}}
+
+    \textbf{Panel B: units in log \$ per student}
+
+    \makebox[\textwidth][c]{\input{tables/firststage-illinois.tex}}
+
+    \label{tab:firststage-illinois}
+    \justify
     \footnotesize
-        \textbf{Note}: This table presents the summary statistics for professors with an observed year of first-year at their university.
-    \end{flushleft}
+    \textbf{Note}:
+    These tables show the first stage OLS estimates of regression specification \eqref{eqn:secondstage1_indiv}, showing the effect of the funding shift-share on state funding to gauge performance as an instrument.
+    Each observation is a professor-year, in the IBHED data, and funding data are merged from IPEDS.
+    %Panel A shows the effect of an funding shift-share of \$-1 per student in the state on the number of \$'s of state funding per student at the university --- i.e.,
+    %\$-1 funding shift-share per student in the state leads to \$1.176 less state funding per student at the university according to preferred specification column 1.
+    Panel A shows the effect of a $-10$\% change funding shift-share per student in the state on $10$\% change in state funding per student at the university --- i.e.,
+    $-10$\% funding shift-share per student in the state leads to $-9.77$\% less state funding per student at the university according to prefferred specification column 1.        
+    Standard errors are clustered at the institution-year level, and institution $+$ year fixed effects are included where noted.
+\end{table}
+
+Exogeneity and relevance of the rolling-share instrument, $\tilde Z_{i(j),t}$, follows the same reasoning as that for the base-share instrument, $Z_{i,t}$, discussed in \autoref{sec:approp-shocks}.
+The base-share instrument is appropriate for some outcomes with the individual Illinois professors, where appropriate.
+We satisfy the assumptions for exogeneity by noting that none of the Illinois public campuses take the majority of state funding, and that the identification strategy relies on exogeneity in changes in state funding to individual professor-outcomes, following the year they joined the university.
+Additionally, within-institution changes resulting from share reliance on state funding may be correlated with unobserved changes in the outcomes, so that \cite{NBERw27885} note the importance of controlling for the base share and state student population.
+The formulation here implicitly controls for these factors via the fixed effects; results are relatively similar while including these controls with and without including fixed effects, and so are omitted.
+
+The instrumental variables model is then defined as follows, where $i(j)$ refers to the institution that professor $j$ is employed at, and $Y_{j,t}$ for salary, rate of promotion, and propensity to leave the Illinois public university system.
+The system includes fixed effects for the institution and first year of employment.
+The instrument varies by institution, based in the year of first employment, so that these are the corresponding fixed effects and level of clustered standard errors.
+\begin{eqnarray}
+    \label{eqn:secondstage1_indiv}
+    X_{i(j),t} &=& \theta_{i(j)} + \phi_{\tau(j)} + \delta \tilde Z_{i(j),t} + \epsilon_{i(j),t} \\
+    \label{eqn:secondstage2_indiv}
+    Y_{j,t} &=& \mu_{i(j)} + \nu_{\tau(j)} + \beta \widehat X_{i(j),t} + \varepsilon_{j,t}
+\end{eqnarray}
+We then interpret parameter $\beta$ as the effect of changes in state funding at an Illinois public university, via state funding shift-shares, on an individual professor's outcome $Y_{j,t}$.
+
+\newpage
+\begin{figure}[H]
+    \centering
+    \singlespacing
+    \caption{Local Projection Estimates for Effect of State Funding on Faculty Promotion Rate at Illinois Public Universities, by Professor Group.}
+    \begin{subfigure}[b]{0.495\textwidth}
+        \centering
+        \caption{First-stage Estimate.}
+        \includegraphics[width=\textwidth]{figures/firststage-illinois-lp-rolling.png}
+        \label{fig:firststage-illinois-lp-rolling}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.495\textwidth}
+        \centering
+        \caption{Lecturers.}
+        \includegraphics[width=\textwidth]{figures/promoted-lecturer-illinois-lp-rolling.png}
+        \label{fig:promoted-lecturer-illinois-lp-rolling}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.495\textwidth}
+        \centering
+        \caption{Assistant Professors.}
+        \includegraphics[width=\textwidth]{figures/promoted-assistant-illinois-lp-rolling.png}
+        \label{fig:promoted-assistant-illinois-lp-rolling}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.495\textwidth}
+        \centering
+        \caption{Full Professors.}
+        \includegraphics[width=\textwidth]{figures/promoted-full-illinois-lp-rolling.png}
+        \label{fig:promoted-full-illinois-lp-rolling}
+    \end{subfigure}
+    \label{fig:promoted-illinois-lp-rolling}
+    \justify
+    \footnotesize
+    \textbf{Note}:
+    These figures show the local projections estimates of regression specification \eqref{eqn:secondstage}, with the funding shift-share as an instrument for state funding.
+    The unit of analysis is an individual faculty member (at an Illinois public university); funding data come from IPEDS, and faculty promotion rate from IBHED.
+    The coefficient estimate is effect of state funding ($X_{i(j),t}$) on faculty promotion rate ($Y_{j,t}$), using the funding shift-share instrument ($Z_{i(j),t}$), while accounting for auto-correlation between different time periods --- i.e., between $X_{i(j),t}, X_{i(j),t-1}$ and $Y_{i(j),t}, Y_{i(j),t-1}$.
+    These results use a $\log-\log$ specification, so the estimates are for the rate of promotion in a year $t+k$ affected by a 1\% change in state funding in year $t$, where years $k = 0, \hdots, 4$ are on the $x$-axis.
+    Standard errors are clustered at the university-year level, and \autoref{sec:iv-model-indiv} fully describes the differences in empirical specification when unit of analysis is an individual faculty member.
+\end{figure}
+
+\newpage
+\begin{table}[H]
+    \singlespacing
+    \centering
+    \caption{Effects of Changes in State Funding on University Faculty Composition, in Illinois 2010--2021, OLS and 2SLS Estimates.}
+
+    \textbf{Panel A: units in \$ per student}
+
+    \makebox[\textwidth][c]{\input{tables/facultycount-illinois-reg-rawcount.tex}}
+
+    \textbf{Panel B: units in log \$ per student}
+
+    \makebox[\textwidth][c]{\input{tables/facultycount-illinois-reg-fte.tex}}
+
+    \label{tab:facultycount-illinois-reg}
+    \justify
+    \footnotesize
+    \textbf{Note}:
+    These tables show the second stage OLS and 2SLS estimates of regression specification \eqref{eqn:secondstage}, showing the effect of state funding changes on number of faculty per student in Illinois universities, using the funding shift-share to instrument for state funding in the columns labelled 2SLS.
+    Each observation is a public university-year in the state of Illinois, where funding data come from IPEDS and faculty count come from IBHED data.
+    Panel A shows the effect of a fall in state funding \$$-1,000$ per student in the state on the number of professors.
+    Panel B shows the effect of a $10$\% change in state funding per student at the university on the 10\% change in the number of professors per students.
+    Outcome-mean is the mean of the outcome, for Panel A the number of professors per student, for Panel B the number of faculty per student.
+    Panel B uses $\log$ faculty count per student as the outcome, though the outcome mean is count of faculty per student (not in $\log$ terms).
+    Standard errors are clustered at the university-year level, and university $+$ year fixed effects are included through--out.
+\end{table}
+
+\newpage
+\begin{figure}[H]
+    \centering
+    \singlespacing
+    \caption{Local Projection Estimates for Effect of State Funding on Faculty Exit Rate at Illinois Public Universities, by Professor Group.}
+    \begin{subfigure}[b]{0.495\textwidth}
+        \centering
+        \caption{Lecturers.}
+        \includegraphics[width=\textwidth]{figures/exit-lecturer-illinois-lp-rolling.png}
+        \label{fig:exit-lecturer-illinois-lp-rolling}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.495\textwidth}
+        \centering
+        \caption{Assistant Professors.}
+        \includegraphics[width=\textwidth]{figures/exit-assistant-illinois-lp-rolling.png}
+        \label{fig:exit-assistant-illinois-lp-rolling}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.495\textwidth}
+        \centering
+        \caption{Full Professors.}
+        \includegraphics[width=\textwidth]{figures/exit-full-illinois-lp-rolling.png}
+        \label{fig:exit-full-illinois-lp-rolling}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.495\textwidth}
+        \centering
+        \caption{Administrator Professors.}
+        \includegraphics[width=\textwidth]{figures/exit-administrator-illinois-lp-rolling.png}
+        \label{fig:exit-administrator-illinois-lp-rolling}
+    \end{subfigure}
+    \label{fig:exit-illinois-lp-rolling}
+    \justify
+    \footnotesize
+    \textbf{Note}:
+    These figures show the local projections estimates of regression specification \eqref{eqn:secondstage}, with the funding shift-share as an instrument for state funding.
+    The unit of analysis is an individual faculty member (at an Illinois public university); funding data come from IPEDS, and faculty promotion rate from IBHED.
+    The coefficient estimate is effect of state funding ($X_{i(j),t}$) on faculty promotion rate ($Y_{j,t}$), using the funding shift-share instrument ($Z_{i(j),t}$), while accounting for auto-correlation between different time periods --- i.e., between $X_{i(j),t}, X_{i(j),t-1}$ and $Y_{i(j),t}, Y_{i(j),t-1}$.
+    These results use a rate$-\log$ specification, so the estimates are for the rate of promotion in a year $t+k$ affected by a 1\% change in state funding in year $t$, where years $k = 0, \hdots, 4$ are on the $x$-axis.
+    Standard errors are clustered at the university-year level, and \autoref{sec:iv-model-indiv} fully describes the differences in empirical specification when unit of analysis is an individual faculty member.
+\end{figure}
+
+\newpage
+\subsection{Professor Hiring}
+\label{sec:appendix-hiring}
+
+These results were produced by integrating the total count of professor hires for 2010--2021 for the top-ranked 180 US universities with a sum of the funding variables, and then estimating the models specified in \autoref{sec:iv-model-uni}.
+There were no observable differences in the hiring rate of male versus female faculty.
+
+\begin{table}[H]
+    \singlespacing
+    \centering
+    \caption{OLS and 2SLS Estimates for University Faculty Hires, in Illinois 2011--2021.}
+
+    \textbf{Panel A: units in \$ per student}
+
+    \makebox[\textwidth][c]{\input{tables/facultyhires-illinois-reg-rawcount.tex}}
+
+    \textbf{Panel B: units in log \$ per student}
+
+    \makebox[\textwidth][c]{\input{tables/facultyhires-illinois-reg-fte.tex}}
+    \label{tab:facultyhires-illinois-reg}
+    \vspace{-0.5cm}
+    \justify
+    \footnotesize
+    \textbf{Note}:
+    These tables show the second stage OLS and 2SLS estimates of regression specification \eqref{eqn:secondstage}, showing the effect of state funding changes on number of faculty hires at Illinois universities, using the funding shift-share to instrument for state funding in the columns labelled 2SLS.
+    Each observation is a public university-year in the state of Illinois, where funding data come from IPEDS and faculty count come from IBHED data.
+    Panel A shows the effect of a fall in state funding \$$-1,000$ per student in the state on the number of new faculty hires by position.
+    Panel B shows the effect of a $10$\% change in state funding per student at the university on the 10\% change in the number of faculty hires per students.
+    Outcome-mean is the mean of the outcome, for Panel A the number of faculty hires, for Panel B the number of faculty hires per student.
+    Panel B uses new faculty hires per student as the outcome (in $\log$ terms), though the outcome mean is count of new faculty hires per student (not in $\log$ terms).
+    Standard errors are clustered at the university-year level, and university $+$ year fixed effects are included through--out.
+\end{table}
+
+\begin{table}[h!]
+    \singlespacing
+    \centering
+    \caption{OLS and 2SLS Estimates for Professor Hiring, Total for 2011--2020.}
+    \makebox[\textwidth][c]{\input{tables/hiring-shock-reg-fte.tex}}
+    \label{tab:hiring-shock-reg}
+    \justify
+    \footnotesize
+    \textbf{Note}: 
+    This table show the second stage 2SLS estimates of regression specification \eqref{eqn:secondstage}, showing the effect of state funding changes on the number of faculty hires (per student) total for 2011--2021 at US public universities, using the funding shift-share to instrument for state funding.
+    Yearly variation in total hires is not observed here, so only the total hires across 2011--2021 for 157 universities, can be considered.
+    Each observation is a public university across the years 2011--2021, where data on total funding across 2011--2021 come from IPEDS and faculty count total from \citep{wapman2022quantifying,wapman2022zenodo}.
+    The panels show the effect of a $1$\% change in state funding per student at the university (total for 2011--2021) on the number of new faculty hires by gender (and all).
+    Standard errors are clustered at the state level, and state fixed effects are included through--out.
+\end{table}
+
+\newpage
+\subsection{Robustness Checks}
+
+\begin{table}[H]
+    \singlespacing
+    \centering
+    \caption{Effects of State Funding on Faculty Counts, IPEDS 1990--2021, IV Estimates by Institution Selectivity.}
+
+    \textbf{Panel A: units in \$ per student}
+
+    \makebox[\textwidth][c]{
+        \begin{tabular}{@{\extracolsep{5pt}}lcccccl} 
+            \\[-1.8ex]\hline 
+            \hline \\[-1.8ex]
+            & First-stage & Lecturers & Asst. Professors & Full Professors & All Faculty & Observations \\ 
+            \cline{2-7} 
+            \\[-1.8ex]
+            \input{tables/facultycount-rawcounts-heterogeneity.tex}
+            \\[-1.8ex] \hline 
+            \hline 
+        \end{tabular}}
+
+        \vspace{0.2cm}
+    \textbf{Panel B: units in log \$ per student}
+
+    \makebox[\textwidth][c]{
+        \begin{tabular}{@{\extracolsep{5pt}}lcccccl} 
+            \\[-1.8ex]\hline 
+            \hline \\[-1.8ex]
+            & First-stage & Lecturers & Asst. Professors & Full Professors & All Faculty & Observations \\ 
+            \cline{2-7} 
+            \\[-1.8ex]
+            \input{tables/facultycount-heterogeneity.tex}
+            \\[-1.8ex] \hline 
+            \hline 
+        \end{tabular}}
+    \label{tab:facultycount-heterogeneity}
+    \justify
+    \footnotesize
+    \textbf{Note}:
+    These tables show the IV estimates of regression specification \eqref{eqn:secondstage}, 
+    in the same manner as \autoref{tab:facultycount-shock-reg}, but restricting to institutions ranked as most selective, selective, and unranked by \cite{barrons2009} --- and including state
+    The first column presents the coefficient between state funding and shift-share IV, which is a string first-stage among every level of selectivity for public universities.
+    The other columns show the coefficient between state funding the count of faculty per students.
+    For example, Row 1 of panel B shows that a $10\%$ cut in the state shift-share leads to a fall of $9.44\%$ in state funding for the public universities ranked as ``most selective.''
+    Standard errors for the coefficient estimates are in brackets, and the outcome mean are in square brackets beneath.
 \end{table}
@@
     \centering
-    \caption{First Stage Estimates, for Total University Revenues at the Individual-Level.}
-    \makebox[\textwidth][c]{\input{tables/firststage-illinois-rolling.tex}}
-    \label{tab:firststage-illinois}
-    \begin{flushleft}
+    \caption{First-stage Robustness Checks for Effects of State Funding Shift-Share on State Funding, OLS Estimates.}
+    \makebox[\textwidth][c]{\input{tables/firststage-robustness-checks.tex}}
+    \label{tab:firststage-robustness-checks}
+    \justify
     \footnotesize
-        \textbf{Note}: Standard errors are clustered at the university-year level.
-    \end{flushleft}
-\end{table}
-
-\begin{figure}[h!]
-    \centering
-    \singlespacing
-    \caption{Local Projection Estimates for First-Stage \autoref{eqn:secondstage1_indiv}.}
-    \includegraphics[width=0.6\textwidth]{figures/firststage-illinois-lp-rolling.png}
-    \label{fig:firststage-illinois-lp}
-\end{figure}
-
-\subsection{Additional Results, University-Level}
-
-\autoref{tab:facultysalaries-shock-reg} presents OLS and IV estimates where the outcome is (log) mean salary for professors employed at the university, separated by position.
-Similarly, \autoref{fig:all-salaries-lp} shows the staying-power via projection methods.
-The outcome, mean salary is crude.
-
-\begin{table}[H]
-    \singlespacing
-    \centering
-    \caption{OLS and 2SLS Estimates for University Faculty-Tenure Composition.}
-    \makebox[\textwidth][c]{\input{tables/tenurecount-shock-reg-fte.tex}}
-    \begin{flushleft}
-        \footnotesize
-        \textbf{Note}: Standard errors are clustered at the university-year level.
-    \end{flushleft}
-    \label{tab:tenurecount-shock-reg-fte}
+    \textbf{Note}:
+    These tables show the second stage OLS and 2SLS estimates of regression specification \eqref{eqn:firststage}, showing the effect of state funding shift-share on each university's state funding.
+    This table differs from the main analysis by replacing University $+$ Year fixed effects with State $+$ Year fixed effects, measuring enrolment in full-time equivalent (FTE), and including controls for (1)
+    State Funding as a percent of total funding in base period, (2) acceptance rate in mid-2000s, (3) 6--year completion rate in mid-2000s, (4) tuition revenue, (5) enrolment measured by full-time equivalent FTE, (6) total public university enrolment in the entire state, (7) percent of public university enrolment for the state enrolled at this university.
+    The first column uses the raw count specification, and the second column uses the $\log$ specification for percentage terms.
 \end{table}
@@
     \centering
-    \caption{OLS and 2SLS Estimates for University Faculty Salaries.}
-    \makebox[\textwidth][c]{\input{tables/facultysalaries-shock-reg-fte.tex}}
-    \begin{flushleft}
+    \caption{Robustness Checks for Effects of State Funding Cuts on Faculty Composition, OLS and 2SLS Estimates in $\log$ Units.}
+    \makebox[\textwidth][c]{
+        %\small
+        \input{tables/facultycount-robustness-checks.tex}}
+    \label{tab:facultycount-robustness-checks}
+    \justify
     \footnotesize
-        \textbf{Note}: Standard errors are clustered at the state-year level.
-    \end{flushleft}
-    \label{tab:facultysalaries-shock-reg}
-\end{table}
-
-\begin{figure}[H]
-    \centering
-    \singlespacing
-    \caption{Local Projection Estimates for Professor Salary, Mean within Each University.}
-    \includegraphics[width=0.6\textwidth]{figures/all-salaries-lp.png}
-    \label{fig:all-salaries-lp}
-\end{figure}
-
-\subsection{Additional Results, Individual-Professor Level}
-
-\begin{table}[H]
-    \singlespacing
-    \centering
-    \caption{2SLS Estimates for Faculty Salaries at Illinois Universities.}
-    \makebox[\textwidth][c]{\input{tables/facultysalaries-shock-illinois.tex}}
-    \begin{flushleft}
-        \footnotesize
-        \textbf{Note}: Standard errors are clustered at the institution-year level.
-    \end{flushleft}
-    \label{tab:facultysalaries-shock-illinois}
+    \textbf{Note}:
+    These tables show the second stage OLS and 2SLS estimates of regression specification \eqref{eqn:secondstage}, showing the effect of state funding changes on number of faculty per student in Illinois universities, using the funding shift-share to instrument for state funding in the columns labelled 2SLS.
+    This table differs from the main analysis by replacing University $+$ Year fixed effects with State $+$ Year fixed effects, measuring enrolment in full-time equivalent (FTE), and including controls for (1)
+    State Funding as a percent of total funding in base period, (2) acceptance rate in mid-2000s, (3) 6--year completion rate in mid-2000s, (4) tuition revenue, (5) enrolment measured by full-time equivalent FTE, (6) total public university enrolment in the entire state, (7) percent of public university enrolment for the state enrolled at this university.
 \end{table}
@@
     \centering
-    \caption{2SLS Estimates for Faculty Promotion Rate at Illinois Universities, using Rolling Instrument.}
-    \makebox[\textwidth][c]{\input{tables/promotion-shock-illinois-rolling.tex}}
-    \begin{flushleft}
+    \caption{Robustness Checks for Effects of State Funding Cuts on Faculty Composition, OLS and 2SLS Estimates in Raw Count Units.}
+    \makebox[\textwidth][c]{
+        \small
+        \input{tables/facultycount-rawcount-robustness-checks.tex}}
+    \label{tab:facultycount-rawcount-robustness-checks}
+    \justify
     \footnotesize
-        \textbf{Note}: Standard errors are clustered at the institution and first year of employment level.
-    \end{flushleft}
-    \label{tab:promotion-shock-illinois-rolling}
+    \textbf{Note}:
+    These tables show the second stage OLS and 2SLS estimates of regression specification \eqref{eqn:secondstage}, showing the effect of state funding changes on count of faculty in Illinois universities, using the funding shift-share to instrument for state funding in the columns labelled 2SLS.
+    This table differs from the main analysis by replacing University $+$ Year fixed effects with State $+$ Year fixed effects, measuring enrolment in full-time equivalent (FTE), and including controls for (1)
+    State Funding as a percent of total funding in base period, (2) acceptance rate in mid-2000s, (3) 6--year completion rate in mid-2000s, (4) tuition revenue, (5) enrolment measured by full-time equivalent FTE, (6) total public university enrolment in the entire state, (7) percent of public university enrolment for the state enrolled at this university.
 \end{table}
@@
     \centering
-    \caption{2SLS Estimates for Faculty Promotion Rate at Illinois Universities, using Base-Year Instrument.}
-    \makebox[\textwidth][c]{\input{tables/promotion-shock-illinois.tex}}
-    \begin{flushleft}
+    \caption{Effects of State--Wide Funding Changes on Private University Faculty Counts, IPEDS 1990--2021, OLS and 2SLS Estimates.}
+
+    \textbf{Panel A: units in \$ per student}
+
+    \makebox[\textwidth][c]{\input{tables/facultycount-shock-private-robustness-rawcount.tex}}
+
+    \textbf{Panel B: units in log \$ per student}
+
+    \makebox[\textwidth][c]{\input{tables/facultycount-shock-private-robustness.tex}}
+
+    \label{tab:facultycount-shock-private-robustness}
+    \justify
     \footnotesize
-        \textbf{Note}: Standard errors are clustered at the institution and first year of employment level.
-    \end{flushleft}
-    \label{tab:promotion-shock-illinois}
+    \textbf{Note}:
+    These tables show the second stage OLS and 2SLS estimates of regression specification \eqref{eqn:secondstage}, among private universities --- as described in \autoref{sec:results-robustness}.
+    Each observation is a public university-year, in the IPEDS data.
+    Panel A shows the effect of a fall in state funding \$$-1,000$ per student on the number of professors --- i.e.,
+    an extra \$$1,000$ per student leads to 6 fewer lecturers according to column 2.
+    Panel B shows the effect of a $10$\% change in state funding per student at the university on the 10\% change in the number of professors per students --- i.e.,
+    an extra 10\% of state funding per student leads to 4.37\% fewer lecturers per student according to column 2.
+    Outcome-mean is the mean of the outcome, for Panel A the number of professors per student, for Panel B the number of faculty per student.
+    Panel B uses $\log$ faculty count per student as the outcome, though the outcome mean is count of faculty per student (not in $\log$ terms).
+    Standard errors are clustered at the state-year level, and university $+$ year fixed effects are included through--out.
 \end{table}
 
-\begin{figure}[H]
-    \centering
-    \singlespacing
-    \caption{Local Projection Estimates for Promotion Rate at Illinois Universities.}
-    \includegraphics[width=0.6\textwidth]{figures/promoted-illinois-lp-rolling.png}
-    \label{fig:promoted-illinois-lp}
-\end{figure}
+\newpage
+\subsection{Analysis of IPEDS Salary Data and Years Following the Great Recession}
+\label{sec:recession-analysis}
 
-\begin{table}[H]
-    \singlespacing
-    \centering
-    \caption{2SLS Estimates for Faculty Exit Rate at Illinois Universities, using Rolling Instrument.}
-    \makebox[\textwidth][c]{\input{tables/facultyleaving-shock-illinois-rolling.tex}}
-    \begin{flushleft}
-        \footnotesize
-        \textbf{Note}: Standard errors are clustered at the institution and first year of employment level.
-    \end{flushleft}
-    \label{tab:facultyleaving-shock-illinois-rolling}
-\end{table}
-
-\begin{table}[H]
-    \singlespacing
-    \centering
-    \caption{2SLS Estimates for Faculty Exit Rate at Illinois Universities, using Base-Year Instrument.}
-    \makebox[\textwidth][c]{\input{tables/facultyleaving-shock-illinois.tex}}
-    \begin{flushleft}
-        \footnotesize
-        \textbf{Note}: Standard errors are clustered at the institution and first year of employment level.
-    \end{flushleft}
-    \label{tab:facultyleaving-shock-illinois}
-\end{table}
+Bigger analysis in here.

### 2024-09-25 text/sections/02-data.tex
Made some edits.

@@
     \cite{hinrichs2022state} also does not mention the data inconsistencies in IPEDS data for average faculty salaries.
-    These two major differences explain the different conclusions that our contemporary papers come to; \autoref{sec:recession-analysis} replicates the \cite{hinrichs2022state} empirical findings with my primary specification, and further discuses the differences in findings.
+    These two major differences explain the different conclusions that our contemporary papers come to; in \autoref{sec:recession-analysis} I replicate the \cite{hinrichs2022state} approach with my primary specification, and further discuses the differences in findings.
 }

### 2024-09-26 text/sections/01-introduction.tex
Made some edits.
Added a number of new things.  2 minor comments to address before sub.

@@
 In a secondary analysis of all public university faculty in the state of Illinois, I show that incumbent faculty were unaffected, in terms of wages, promotion and quit rates.
-This implies that the faculty substitution arose by disrupting hiring of new tenure-track/tenured professors, which is supported supplementary data on professor hiring at US public universities over 2010--2021.
-This paper shows the institutional effects of stagnating state funding for higher education, and offers one possible mechanism for how funding cuts affected student outcomes.
+This implies that the faculty substitution arose by disrupting hiring of new tenure-track/tenured professors, which is supported with supplementary data on professor hiring at US public universities over 2010--2021.
+This paper shows the institutional effects of stagnating state funding for higher education, and offers one mechanism for how funding cuts affected student outcomes.
 
@@
 In a contemporaneous working paper, \cite{hinrichs2022state} use the same national-level variation in university funding to analyse effects on faculty and wages, coming to different conclusions.\footnote{
-    \cite{hinrichs2022state} considers a shorter time-period than this paper (2010--2021), uses only national variation in IPEDS data, relies on IPEDS data for faculty salaries, and has no individual-level analysis of faculty or their salaries.
-    See \autoref{sec:data-desc,sec:recession-analysis} for an elaboration on these issues.
+    An anonymous reviewer made me aware of a similar and contemporary working paper, \cite{hinrichs2022state}; this work analyses a shorter time-period than this paper, and only focuses on national variation.
+    \autoref{sec:recession-analysis} replicates their key findings, and shows how their focusing on only the 2012--2019 years explains differences in conclusions.
 }

### 2024-09-26 text/sections/04-empirical.tex
Added a number of new things.  2 minor comments to address before sub.

@@
     It may be argued that tuition revenues are confounder between the causal effect of changes in state funding on a university's total revenues, so that this analysis focuses on state support for higher education (not total revenue), as do \cite{NBERw27885}.
-    On the other hand, rises in tuition revenues (per-student) may arise as result of tuition hikes thanks declining state support, which would mean controlling for tuition would constitute a bad control.
-    Nonetheless, estimates including tuition revenues (per-student) as a control in the second stage of the IV estimates produce results of very similar magnitude and direction, and so are omitted.
+    On the other hand, rises in tuition revenues (per-student) may arise as result of tuition hikes following state funding cuts, which would mean controlling for tuition constitutes a bad control.
+    Nonetheless, estimates including tuition revenues (per-student) as a control in the second stage of the IV estimates produce results of very similar magnitude and direction, and so are omitted (see \autoref{sec:results-firststage}).
 }

### 2024-09-26 text/sections/05-results.tex
Added a number of new things.  2 minor comments to address before sub.

@@
 \subsection{First-Stage Estimates}
+\label{sec:results-firststage}
 The shift-share instrument has a strong first-stage for universities' yearly state funding.
@@
 Additionally, they charge higher tuition fees, though this rise is at most a quarter of the state funding cut, so does not come close to fully off-setting state funding cuts.
-These results are in line with previous work showing state funding cuts (or tuition price limits) reduce student-focused spending \citep{NBERw23736}, financial aid \citep{miller2022making}, and lead to increased tuition fees \citep{bound2019public}.
+These results are in line with previous work showing state funding cuts (or tuition price limits) reduce student-focused spending \citep{NBERw23736}, financial aid \citep{miller2022making}, and lead to increased tuition fees \citep{bound2019public,webber2017state}.
 
@@
 For example, selective institutions may have received funding cuts for other reasons, or states with lower performing institutions or declining populations strategically cut funding for only less selective universities.
-Additionally, the effect of state funding cuts may affect faculty at different universities to different extents
-I investigate these issues with three different analyses.
+I investigate these issues with two different analyses, and then investigate heterogeneity across rates of selectivity and in different decades.
 
@@
 
-Second, I return to the primary specification in \autoref{eqn:secondstage}, and estimate effect of state funding cuts on faculty counts separately among universities of different selectivity rates.
-This investigates whether substitution rates are different across different levels of selectivity; \autoref{tab:facultycount-heterogeneity} collects these results together, separated by the \cite{barrons2009} selectivity rankings.
-Barrons rank 1-2 are the most selective (including e.g., UC Berkeley and SUNY Binghamton), 3-4 selective (including e.g., Clemson University and University of Dallas), and rank 5 for the rest of universities.
-The first-stage is strong in every group of universities, even among the small sample size of the 12 universities in the most selective group, and the 28 universities in the next selective group.
-Faculty substitution rates are not very strong among the most selective universities, but are very similar between selective universities and unranked universities.
-One might expect that the effect is driven entirely by the least selective universities, as they are the most resource-constrained;
-this heterogeneity analysis shows that even selective (if not top) universities substitute towards lecturers following state funding cuts.
-
-The final robustness check I employ is a falsification test, testing whether the effects for public universities would be picked up among a sample of private universities.\footnote{
+As another robustness check, I employ is a falsification test, testing whether the effects for public universities would be picked up among a sample of private universities.\footnote{
     This is a robustness check similar to one performed in \citealt{NBERw27885}, as a robustness check for the same instrument.
@@
 Private universities do not rely on state funding in the same way that public universities do,\footnote{
-    Though private universities do receive some state funding, as some states fund smaller scale operating costs and other specific projects at private universities.
+    Some private universities do receive state funding, mostly funding for smaller scale operating costs and other specific projects.
 }
@@
 The IV estimates are not distinguishable from zero, showing that OLS correlations are not causal relationships, and validating the IV approach for inference on universities' faculty counts.
+
+To investigate whether universities of different selectivity respond differently to state funding cuts, I estimate the effect of state funding cuts on faculty counts separately among universities of different selectivity rates.
+\autoref{tab:facultycount-heterogeneity} collects these results together, estimating \autoref{eqn:secondstage} separately by the \cite{barrons2009} selectivity rankings.
+Barrons rank 1-2 are the most selective (including e.g., UC Berkeley and SUNY Binghamton), 3-4 medium selective (including e.g., Clemson University and University of Dallas), and rank 5 for the rest of universities.
+The first-stage is strong in every group of universities, even among the small sample size of 12 universities in the most selective group, and 28 universities in the next selective group.
+Faculty substitution rates are not distinguishable from zero among the most selective universities, but are positive and similar between both medium selective universities and unranked universities.
+One might expect that the effect is driven entirely by non-selective and low ranked universities, as they are the most resource-constrained;
+this heterogeneity analysis shows that even medium selective universities substitute towards lecturers following state funding cuts.
+
+Last, I investigate when the substitution towards lecturers occurred.
+To do so, I estimate the effect of state funding on faculty counts relative student enrolment separately for 1990--2009 and 2010--2021.
+Substitution towards lecturer is weak to non existent in the latest decade, as noted in contemporary research \citep{hinrichs2022state}.
+However, the rates are stronger when just looking at the years 1990--2010.
+This reflects a key point noted in \autoref{sec:trends}: a large amount of the cuts, and general stagnation, in state funding for public universities occurred over in the 1990s and 2000s.
+Additionally, the trend in rising lecturer count, and falling full professor count, per student at public universities had been operating all throughout the 1990s and 2000s.
+These trends did continue in the 2010s (especially for counts of assistant professors), but most of the substitution had already taken place by this point.

### 2024-09-26 text/sections/08-figures-tables.tex
Made some edits.
Added a number of new things.  2 minor comments to address before sub.

--- 2024-05-10 text/sections/09-figures-tables.tex
@@
     \centering
-    \caption{First Stage Estimates, Predicting State Funding by Funding Shift-Share, in IPEDS Data.}
+    \caption{First-Stage Estimates, Predicting State Funding by Funding Shift-Share, in IPEDS Data.}
     \textbf{Panel A: units in \$ per student}
@@
     \textbf{Note}:
-    These tables show the first stage OLS estimates of regression specification \eqref{eqn:firststage}, showing the effect of the state funding shift-share on state funding to gauge performance as an instrument.
+    These tables show the first-stage OLS estimates of regression specification \eqref{eqn:firststage}, showing the effect of the state funding shift-share on state funding to gauge performance as an instrument.
     Each observation is a public university-year, in the IPEDS data.
@@
     Standard errors are clustered at the state-year level, and university $+$ year fixed effects are included where noted.
+    The F statistic is for the excluded instrument (not for overall regression), after accounting for fixed effects and cluster dependence \citep{olea2013robust}.
 \end{table}

### 2024-09-26 text/sections/09-appendix.tex
Added a number of new things.  2 minor comments to address before sub.

@@
 \newpage
-\subsection{Analysis of IPEDS Salary Data and Years Following the Great Recession}
+\subsection{Different Rates of Substitution for Each Decade}
+\label{sec:decade-analysis}
+
+\begin{table}[H]
+    \singlespacing
+    \centering
+    \caption{Effects of State Funding on Faculty Counts, in IPEDS Data 1990--2021, Separated for Each Decade.}
+
+    \textbf{Panel A: Estimated Separately for 1990--2009}
+
+    \makebox[\textwidth][c]{
+        \input{tables/facultycount-replicated-corrected-1990s.tex}}
+
+    \textbf{Panel B: Estimated Separately for 2010--2021}
+
+    \makebox[\textwidth][c]{
+            \input{tables/facultycount-replicated-corrected-2010s.tex}}
+
+    \label{tab:facultycount-replicated-corrected}
+    \justify
+    \footnotesize
+    \textbf{Note}:
+    These tables show the second stage OLS and 2SLS estimates of regression specification \eqref{eqn:secondstage}, showing the effect of state funding changes on faculty outcomes, using the state funding shift-share (based in 1990--1993) to instrument for state funding in the columns labelled 2SLS.
+    Each panel gives estimates for data restricted to the noted decades.
+    Each observation is a public university-year, in the IPEDS data.
+    Each panel shows the effect of a $10$\% change in state funding per student at the university on the 10\% change in the number of professors per students.
+    Outcome-mean is the mean of the outcome, for Panel A the number of professors per student, for Panel B the number of faculty per student.
+    Each uses $\log$ faculty count per student as the outcome, though the outcome mean is count of faculty per student (not in $\log$ terms).
+    Standard errors are clustered at the state-year level, and university $+$ year fixed effects are included through-out.
+\end{table}
+
+\newpage
+\subsection{Replication of IPEDS Results for the Years 2011--2019}
 \label{sec:recession-analysis}
+\cite{hinrichs2022state} present an analysis of the same topic as the IPEDS analysis in this paper.
+Their analysis also includes faculty at 2-year colleges, which is not related to my analysis of lecturers and professors at 4-year universities.
 
-Bigger analysis in here.
+\cite{hinrichs2022state} analyses IPEDS data only in the years 2011--2019, as the Department of Education added a number of questions to the IPEDS survey asking how many part- and full-time professors are employed by the university.
+This means that they define the shift-share instrument with a base year of 2011, do not adjust for the student population size --- in neither the state nor the university under consideration (i.e., on neither shift nor share):
+\begin{align}
+    \label{eqn:replication-instrument}
+    Z^{2011}_{i,t} &=
+    \text{Total State Funding}_{s(i),t} \times
+    \left( \frac{\text{State Funding}_{i,2011}}{\text{Total Revenues}_{i,2011}} \right).
+\end{align}
+I replicate the $Z^{2011}_{i,t}$ formulation of the shift-share instrument, showing it is strong in predicting state funding (without adjusting for student population) in following years 2012--2019.
+There are only minor differences in the results presented to that for Table 2 in \cite{hinrichs2022state}, as their estimation is on the sample of universities across different samples.
+\autoref{tab:firststage-replicate} replicates this finding on the sample of universities who report faculty counts by rank.
+\begin{table}[H]
+    \singlespacing
+    \centering
+    \caption{Replication of First-Stage Estimates, Predicting State Funding by Funding Shift-Share, in IPEDS Data 2012--2019 \citep{hinrichs2022state}.}
+    \makebox[\textwidth][c]{\input{tables/firststage-replicate.tex}}
+    \label{tab:firststage-replicate}
+    \justify
+    \footnotesize
+    \textbf{Note}:
+    These tables show the first stage OLS estimates of regressing state funding in years 2012-2019 but the 2011 base share version of the shift-share instrument, without adjusting for student population size.
+    Standard errors are clustered at the state-year level, and university $+$ year fixed effects are included where noted.
+    The F statistic is for the excluded instrument (not for overall regression), after accounting for fixed effects and cluster dependence \citep{olea2013robust}.
+\end{table}
+This definition contrasts with the analysis in \autoref{sec:empirics}, and \cite{NBERw23736}, \cite{NBERw23736}.
+All three use a shift-share based in the year 1990, or averaged 1990--1993, analyse outcomes in the 1990s and 2000s, and adjust for student population size.
+
+\begin{table}[H]
+    \singlespacing
+    \centering
+    \caption{Replication of Second-Stage Estimates, Effects of State Funding on Faculty Counts, in IPEDS Data 2012--2019 \citep{hinrichs2022state}.}
+    \makebox[\textwidth][c]{\input{tables/facultycount-replicate.tex}}
+    \label{tab:facultycount-replicate}
+    \justify
+    \footnotesize
+    \textbf{Note}:
+    These tables show the second stage OLS and 2SLS estimates of regression specification \eqref{eqn:secondstage}, using the state funding shift-share based in 2011 (as described above) to instrument for state funding in the columns labelled 2SLS.
+    State funding is measured in \$ millions.
+    Each observation is a public university-year, in the IPEDS data.
+    This shows the effect of a state funding cut of \$$1,000,000$ on the number of faculty --- i.e.,
+    a cut of \$$1,000,000$ leads to 0.4 fewer lecturers according to column 2.
+    Outcome-mean is the mean of the outcome, the number faculty at the university.
+    Standard errors are clustered at the state-year level, and university $+$ year fixed effects are included through-out.
+\end{table}
+
+According to the \cite{hinrichs2022state} specification, faculty counts are correlated by OLS measures, but the IV estimates show that there is no relationship between raw counts of state funding and faculty counts in the 2010s --- replicating the main finding in \cite{hinrichs2022state}.
+
+On the other hand, there is a relationship between the amount of state funding per student and the number of faculty per student for decades previously.
+This means that not accounting for rising student enrolment mistakes the fundamental story --- see \autoref{sec:trends}.
+Furthermore, most of the substitution towards lecturers happened over the decades before 2010.
+\autoref{tab:facultycount-replicated-corrected} shows that the substitution towards lecturers was much larger in the 1990s and 2000s, and indistinguishable from zero for the later period.
+This both replicates the primary \cite{hinrichs2022state} finding, and shows when the substitution towards lecturers happened: in the 1990s and 2000s.
+Furthermore, the trends in falling state funding for universities were most intense in the 1990s until late 2000s, mostly plateauing for 2010 and onwards.
+Looking for the effects of stagnating state funding for higher education by only considering the years 2012--2019 is ignoring the time period when state funding stagnated, and when universities substituted towards lecturers.

### 2024-09-26 programs/analysis/ipeds-replicate.R

@@ -1,0 +1,608 @@
+#!/usr/bin/R
+## Senan Hogan-Hennessy, 3 August 2023
+## IV for faculty counts, using IPEDS data.
+# Replicate a separate specification, and show 2011 base-share si correlated with
+# follow-on state funding cuts (i.e., share + shift correlated)
+# Whereas the 1990--1993 base-share are not.
+print(Sys.time())
+library(tidyverse) # Functions for data manipulation and visualization
+library(lfe) # Functions for fast linear models with IV + FEs
+library(car) # Function for F stat regarding IV models
+# My custom flavour of Stargazer TeX tables, to avoid a bug:
+# devtools::install_github("shoganhennessy/stargazer")
+library(stargazer) # TeX tables
+library(xtable)
+set.seed(47)
+# This file follows an adjusted Deming Walters (2017, p.10) approach to
+# appropriations shock instrument.
+
+# Define number of digits in tables and graphs
+digits.no <- 3
+
+# Size for figures
+fig.width <- 10
+fig.height <- fig.width * 0.9
+
+# Define default colours.
+colour.list <- c(
+    "#D62728", # Red  -> First-stage.
+    "#1f77b4", # Blue -> Lecturers
+    "#c9721c", # Orange -> Asst Professors
+    "#237e23", # Green -> Full Professors
+    "#cc2c9f") # Strong pink -> All faculty.
+
+
+################################################################################
+## Load and standardised the IPEDS data.
+
+# Load IPEDS data
+ipeds.data <- read_csv("../../data/urban-ipeds/urban-clean-publicunis.csv")
+
+# Select IPEDS variables, and define the panel at the university level
+reg.data <- ipeds.data %>%
+    # Create independent columns and rows.
+    select(
+        # Identifiers and controls
+        unitid,
+        year,
+        state,
+        # Appropriations to the uni
+        totalrevenues_real,
+        nonauxrevenues_real,
+        fedappropriations_real,
+        stateappropriations_real,
+        localappropriations_real,
+        statelocal_funding_real,
+        tuitionrev_real,
+        # Uni spending.
+        nonauxspending_real,
+        # Uni enrollment
+        enrollment_reported,
+        enrollment_fte,
+        acceptance_rate,
+        completion_rate_150pct,
+        # Shift-share shocks Per student
+        appropriationshock_perEnroll_real,
[…]

### 2024-09-26 programs/analysis/ipeds-shock-rawcounts.R

@@ -319,7 +319,7 @@ stargazer(
     naive_all_count.reg, shiftshare_all_count.reg,
     add.lines = outcome.means,
     dep.var.caption =
-        "Dependent Variable: Faculty Count per 1,000 Students, by Position",
+        "Dependent Variable: Faculty Count, by Position",
     dep.var.labels = c(
         "Lecturers", "Asst. Professors", "Full Professors", "All Faculty"),
     column.labels = rep(c("OLS", "2SLS"), 4),

### 2024-09-26 programs/analysis/ipeds-shock.R

@@ -16,7 +16,7 @@ set.seed(47)
 # appropriations shock instrument.
 
 # Define number of digits in tables and graphs
-digits.no <- 3
+digits.no <- 2
 
 # Size for figures
 fig.width <- 10
@@ -101,7 +101,7 @@ reg.data <- ipeds.data %>%
         lecturer_fulltime_count,
         assistant_fulltime_count,
         full_fulltime_count,
-        # Exploratory, areas of spending.
+        # Areas of spending.
         exp_instruc_real,
         exp_research_real,
         exp_pub_serv_real,
@@ -113,7 +113,7 @@ reg.data <- ipeds.data %>%
 
 # Restrict data to unis + years with measured state appropriations & shocks
 reg.data <- reg.data %>%
-    filter(1990 <= year, #year < 2019,
+    filter(1990 <= year,
         !is.na(enrollment_reported), enrollment_reported > 0,
         !is.na(all_prof_count),
         stateappropriations_real > 0,
@@ -529,6 +529,9 @@ ggsave("../../text/figures/percent-spending.png",
 
 
 # Tables for Instrument Exogeneity ---------------------------------------------
+
+# Make the regression data only consider after 1993 (after base instrument)
+reg.data <- reg.data %>% filter(1993 < year)
 
 # Regressions for Instrument balance test, in log terms
 balance_log_enrollment_reported.reg <- reg.data %>%

### 2024-09-26 programs/run.sh

@@ -23,8 +23,10 @@ R CMD BATCH --no-save ipeds-shock.R
 R CMD BATCH --no-save ipeds-shock-rawcounts.R
 # Robustness checks at the institution-level, IPEDS
 R CMD BATCH --no-save ipeds-robustness-checks.R
-# Analysis and regressions at the institution-level, IPEDS, local porjections.
+# Analysis and regressions at the institution-level, IPEDS, local projections.
 R CMD BATCH --no-save ipeds-shock-longterm.R
+# Replication of another specification + comparison analysis.
+R CMD BATCH --no-save ipeds-replicate.R
 # Analysis and regressions at the individual-level, IBHED, base-year shock IV.
 R CMD BATCH --no-save ibhed-shock.R
 # Analysis and regressions at the using IBHED count outcomes, instituion level.

### 2024-09-27 text/sections/00-abstract.tex
Adjusted formatting of some figures, and changed some bad punctuation.
This is all done.

@@
-Public universities employ more lecturers and fewer professors than at any other point in the last thirty years, relative to student enrolment.
+Public universities employ more lecturers and fewer professors today than at any other point in the last thirty years, relative to student enrolment.
 At the same time, state funding for higher education has stagnated.
@@
 Using a shift-share approach to instrument for state funding, I find that universities employ 4.4\% more lecturers per student following a 10\% funding cut.
-This shift is accompanied by a reduction in assistant and full professors by 1.4\% and 1.2\%  per student, respectively.
-Incumbent professors' salaries, promotion rates, and quit rates at Illinois universities remain unaffected by funding cuts, indicating that the substitution arose from limiting the hiring of new tenure-track/tenured professors.
-Stagnating state funding impacts public universities and faculty, likely contributing to deteriorating student outcomes at public universities seen since the 1990s.
+This shift is accompanied by a reduction in assistant and full professors by 1.4\% and 1.2\%  per student.
+These effects are concentrated to 1990--2009, with waning national trends for 2010 and onwards.
+Incumbent professors' salaries, promotion rates, and quit rates at Illinois universities remain unaffected by large funding cuts in the 2010s, indicating that the substitution arose from limiting the hiring of new professors.
+Stagnating state funding impacts public universities and faculty, likely contributing to deteriorating student outcomes at public universities since the 1990s.
 

### 2024-09-27 text/sections/01-introduction.tex
This is all done.

@@
 At the same time, public universities have systematically substituted away from professors towards lecturers, who often work on contingent and lower-salaried contracts.
-Previous research shows state funding cuts led to worsening student outcomes \citep{NBERw23736,NBERw27885}, but there is little evidence on whether these funding cuts impacted faculty.
+Previous research shows state funding cuts led to worsening student outcomes \citep{NBERw23736,NBERw27885}, but there is little evidence regarding whether these funding cuts impacted faculty.
 In this paper I show that falling state funding caused a change in the composition of faculty at US public universities, substituting professors for more lecturers.
@@
 I use a shift-share instrumental variables (IV) approach to estimate how faculty are affected by these funding cuts.
-Universities employ $4.4$\% more lecturers per-student following a $10$\% funding cut, and correspondingly rely less on tenure-track faculty by employing $1.4$\% fewer tenure-track professors and $1.2$\% fewer tenured professors per-student.
+Universities employ $4.4$\% more lecturers per-student following a $10$\% funding cut, and correspondingly rely less on tenure-track faculty by employing $1.4$\% fewer tenure-track professors and $1.2$\% fewer professors per-student.
 In a secondary analysis of all public university faculty in the state of Illinois, I show that incumbent faculty were unaffected, in terms of wages, promotion and quit rates.
-This implies that the faculty substitution arose by disrupting hiring of new tenure-track/tenured professors, which is supported with supplementary data on professor hiring at US public universities over 2010--2021.
+This implies that the faculty substitution occurred due to disruptions in hiring new professors, which is supported with supplementary data on professor hiring at US public universities over 2010--2021.
 This paper shows the institutional effects of stagnating state funding for higher education, and offers one mechanism for how funding cuts affected student outcomes.
@@
 
-It is a priori unclear how faculty may be affected, as there are various margins that universities may to respond to funding cuts.
+It is a priori unclear how faculty may be affected, as there are various margins that universities may respond to funding cuts.
 On one hand, faculty may be negatively affected by universities cutting salaries or laying-off incumbent faculty.
-This happened in 2010 when the University of Illinois System furloughed university staff following budget cuts after the Great Recession \citep{furlough2010}.
+This occurred in 2010 when the University of Illinois System furloughed staff in response to budget cuts following the Great Recession \citep{furlough2010}.
 Similarly, faculty hiring committees are often cancelled or disrupted in response to funding cuts, meaning there are fewer faculty entering a department to replace those leaving.
@@
 The shift-share instrument identifies exogenous changes in state funding by exploiting differences in universities' reliance on state funding, interacted with state-wide shifts in higher education funding --- following the methods of \cite{NBERw23736,NBERw27885}.
-Yearly state government budgets are decided without targetting individual universities for funding cuts, so more reliant universities will be more affected by state-wide funding cuts, so the shift-share approach estimates causal effects among more reliant universities.
+Yearly state government budgets are decided without targetting individual universities for funding cuts, so more reliant universities will be more affected by state-wide funding cuts, so the shift-share approach estimates causal effects among universities that are more reliant on state funding.
 Additionally, I estimate effect of state funding on faculty multiple years after the initial funding cuts, using the local projections method.\footnote{
@@
 % ie., (35 * 0.437) / 99
-In my analysis of all Illinois faculty (IBHED), I find that incumbent professors were not meaningfully affected, in terms of total salary, promotion rate, or rate of leaving the Illinois public university system.
+These effects are concentrated in years before 2010, with the 2010--2021 period showing no evidence for further trends in substitution.  
+On the other hand, there are large funding cuts in the 2010s in the state of Illinois.
+In my analysis of all Illinois faculty (IBHED), I find that incumbent professors were not meaningfully affected, in terms of total salary, promotion rates, or rates of leaving the Illinois public university system.
 This implies that faculty substitution arose by limiting the hiring of new professors, and is supported by suggestive evidence on professor hiring counts.
@@
 These papers measure changes in state funding and university revenues, but do not measure effects of state funding cuts on faculty composition or salaries, or other outcomes related to university instruction.
-My results provide evidence that the long term trends in higher education funding and faculty outcomes are causally related, contributing to the literature on faculty outcomes \citep{ehrenberg2003studying} and trends in US higher education and funding \citep{hoxby2009changing,ehrenberg2012american}.
+My results provide evidence that the long-term trends in higher education funding and faculty outcomes are causally related, contributing to the literature on faculty outcomes \citep{ehrenberg2003studying} and trends in US higher education and funding \citep{hoxby2009changing,ehrenberg2012american}.
 

### 2024-09-27 text/sections/02-data.tex
This is all done.

@@
     In a contemporaneous working paper, \cite{hinrichs2022state} analyses the effect of state funding cuts on university faculty, using the same shift-share variation.
-    \cite{hinrichs2022state} bases the shift--share instrument in the year 2011, despite major upheavals in the level of state funding for state funding for higher education around the 2008 Great Recession, and years following (see \autoref{sec:trends});
+    \cite{hinrichs2022state} analyses the years 2012--2019, despite national trends in state funding and faculty counts happening largely in the preceding decades (see \autoref{sec:trends,sec:decade-analysis});
     \cite{hinrichs2022state} also does not mention the data inconsistencies in IPEDS data for average faculty salaries.
-    These two major differences explain the different conclusions that our contemporary papers come to; in \autoref{sec:recession-analysis} I replicate the \cite{hinrichs2022state} approach with my primary specification, and further discuses the differences in findings.
+    These two major differences explain the different conclusions that our contemporary papers come to; in \autoref{sec:recession-analysis} I replicate the \cite{hinrichs2022state} approach with my primary specification, and show my results replicate theirs for 2012--2019 (null) and are nonetheless strong for the 1990--2010 time period.
 }
-This gives a resulting panel data-set, where each row represents a university-year, and includes columns for university funding and tuition revenue, plus total count for each faculty position (lecturer, assistant professor, tenured professor, and total faculty), and other university characteristics.
-\autoref{tab:ipeds-summary} presents summary statistics for these variables in IPEDS data.
+This gives a resulting panel data-set, where each row represents a university-year, and includes columns for university funding and tuition revenue, plus total count for each faculty position (lecturers, assistant professors, professors, and total faculty), and other university characteristics.
+I take IPEDS reported count of ``Instructional Staff: Instructor, Lecturer, or No Academic Rank'' and refer to this count as lecturers for brevity, IPEDS count of ``Assistant Professors'' as assistant professors, IPEDS count of ``Associate Professors'' and ``Full Professors'' as professors.\footnote{
+    Some universities use the role of associate professor interchangeably with professor (or play the same role in practice as a full professor).
+    My definition avoids splitting these hairs by summing all ranks above assistant professor.
+}
+\autoref{tab:ipeds-summary} presents summary statistics for these variables in IPEDS data.\footnote{
+    IPEDS started collecting information on the composition of instructional staff working part and full-time, as well as salaries, from the 2010 survey and onwards.
+    This paper focuses on trends that were more stark in the 1990--2009 time period, so does not present systematic evidence on this topic; \cite{hinrichs2022state} considers this time period and finds no relationship.
+}
 
@@
     Public Act 96-0266, effective 1 January 2010, is the relevant Illinois law that requires publicly publishing salary data for all public university faculty salary and benefits \citep{illinois-public-act}.
-    A pdf copy of Public Act 096-0266 is included in the supplmentary data of this paper.
+    A copy of Public Act 096-0266 is included in the supplmentary data of this paper.
     This law applies to all nine Illinois public universities: Chicago State University, Eastern Illinois University, Governors State University, Illinois State University, Northeastern Illinois University, Northern Illinois University, Southern Illinois University (all five campuses), University of Illinois (all four campuses), Western Illinois University.
@@
 Private and public universities have similar numbers of associate and full professors before the year 2000.
-After 2000, the number of professors at public univeristies fell drastically, with the average public university has 6 fewer full professors per hundred students than the average private university by 2021.
+After 2000, the number of professors at public universities fell drastically, with the average public university has 6 fewer full professors per hundred students than the average private university by 2021.
 Over the same time period, we see the rise in use of lecturers; lecturers were employed at similar rates in private and public sectors in 1990.
 Since 1990, both sectors began employing more lecturers on a per-student basis.
-In sum, this means public universities have begun relying on lecturers, because they decreased their professor counts, compared to private unviersities who started employing more lecturers and professors.
+In sum, this means public universities have begun relying on lecturers, because they decreased their professor counts, compared to private universities who started employing more lecturers and professors.
 
@@
 While most public universities were able to stay open, there were drastic revenue and spending cuts in response to the budget impasse, as it continued through fiscal year 2017, and ended with a new budget restoring funding to state institutions for 2018.
-This means that Illinois public universities exhibit sizable changes in their state funding over 2010--2021, of similar order to those for the rest of the country over 1990--2021.
+This means that Illinois public universities exhibit sizeable changes in their state funding over 2010--2021, of similar order to those for the rest of the country over 1990--2021.
 Additionally, the 2016 episode stemmed entirely from political disagreements, and not from state decisions regarding higher education and its finances, exhibiting how state-level changes in funding affect public universities thanks to unrelated issues \citep{young2020squandered}.
+
+While the stagnation in state funding was more intense for the national sample of universities over 1990--2009, Illinois in particular experienced larger disruption and state funding cuts for 2010--2021.
+Additionally, analysis of individual Illinois professors is only possible as data begin in 2010; while it would be preferred to analyse Illinois professor outcomes for 1990--2009, data availability limits the analysis to 2010--2021.

### 2024-09-27 text/sections/04-empirical.tex
This is all done.

@@
 }
+It might be the case that funding cuts have systematically different effects on universities, and faculty, than funding increases.
+The identification strategy here identifies the effect of funding changes, and refers to funding cuts throughout to reflect the systematic falls of state funding (per student) at US public universities.
+I leave it to future research to develop a separate identification strategy that can distinguish the different effects of funding cuts and increases; this paper can not.
 
@@
 \end{equation}
-I estimate the system by two stage least squares, including institution and year fixed effects.\footnote{
-    Note that dividing by student count also implicitly controls for the size of the university, so that this model implicitly accounts for yearly variation in professor count and university revenues arising from a university's size of growth/decline.
-}
+I estimate the system by two-stage least-squares, including institution and year fixed effects.
 $Y_{i,t}$ represents faculty outcomes for university $i$ in year $t$, $\alpha_i, \gamma_t$ university and year fixed effects, and $\widehat X_{i,t}$ state funding for university $i$ estimated in first-stage \eqref{eqn:firststage}.
+Dividing by student count in both outcome and explanatory variables also implicitly controls for the size of the university, so that this model implicitly accounts for yearly variation in professor count and university revenues arising from a university's size of growth/decline.
+Nonetheless, including institution and year fixed effects also controls for the threat to identification of funding cuts possibly being aimed at larger institutions --- even if the primary specifications did not account for student population sizes.
 
@@
 Fixed effects, and two-way standard error clustering are maintained throughout, as before.
+This method requires no further identifying assumptions beyond those for the regular IV methods.
+If the conditions for the shift-share IV apply (\autoref{sec:iv-model-uni}), then identification holds for using LP methods.
 

### 2024-09-27 text/sections/05-results.tex
This is all done.

@@
 }
-I use the same empirical strategy in the analysis of Illinois faculty, and \autoref{tab:firststage-illinois} presents first-stage estimates among the IBHED data; results are similar to those calculated in IPEDS data.
+I use the same empirical strategy in the analysis of Illinois faculty, and \autoref{tab:firststage-illinois} presents first-stage estimates among the IBHED data; results are similar to those calculated in IPEDS data.\footnote{
+    Every IV specification presented in this paper uses a specification with a strong first-stage.
+    Results with weak first-stages are omitted as they would not be reliable causal effects.
+}
 These results show the funding shift-share instrument strongly predicts a university's state funding in the first-stage estimation.
@@
 State funding cuts lead to measurable changes in the number of faculty at US public universities.
-\autoref{tab:facultycount-shock-reg} presents OLS and IV estimates for the effect of state funding cuts on the number of faculty at public universities, separately for each faculty position (lecturer, assistant and full professors, total).
-The results are presented in two different units: Panel A shows how a \$1 increase in state funding changed the count of faculty per students, while Panel B shows how a 1\% increase in state funding led to a percent change in the number of faculty per students.
+\autoref{tab:facultycount-shock-reg} presents OLS and IV estimates for the effect of state funding cuts on the number of faculty at public universities, separately for each faculty position (lecturer, assistant professors and professors, total).
+The results are presented in two different units: Panel A shows how a \$1,000 state funding cut per-student changed the count of faculty, while Panel B shows how a 1\% increase in state funding led to a percent change in the number of faculty per students.
 
-An extra \$1,000 in state funding per student leads to the average university employing 6 fewer lecturers, while effects for the other positions are not discernible from zero (\autoref{tab:facultycount-shock-reg}, Panel A).
-A 10\% state funding cut lead to a 4.4\% increase in the number of lecturers per student, and a decrease of 1.35\% and 1.37\% for assistant and full professors, respectively.
+An extra \$1,000 in state funding per student leads to the average university employing 6 more lecturers, while effects for the other positions are not discernible from zero (\autoref{tab:facultycount-shock-reg}, Panel A).
+A 10\% state funding cut lead to a 4.4\% increase in the number of lecturers per student, and a decrease of 1.35\% and 1.37\% for assistant and professors, respectively.
 A state funding cut of 10\% leads to a 0.65\% decrease in the total number of faculty per student (\autoref{tab:facultycount-shock-reg}, Panel B).
@@
 A 10\% state funding cut leads to 1.5\% (0.6--2.2 in a 95 percent confidence interval) fewer assistant professors per student the following year, and around 1.1\% (0.2--2.0 confidence interval) fewer nine years later;
-the same cut leads to 1.2\% (0.6--1.8 in a 95 percent confidence interval) fewer full professors per student the following year, and around 0.9\% (0.1--1.7 confidence interval) fewer nine years later.
+the same cut leads to 1.2\% (0.6--1.8 in a 95 percent confidence interval) fewer professors per student the following year, and around 0.9\% (0.1--1.7 confidence interval) fewer nine years later.
 
@@
 Furthermore, if a university must teach more students, but cannot afford to hire a full-time professor who teaches 2--3 classes a year, then it makes sense that they hire 2--3 additional lecturers on part-time contracts for each class as a replacement.
-This provides simple intuition for why the lecturers' elasticity is 2--3 times larger than that for assistant/full professors in \autoref{tab:facultycount-shock-reg}.
+This provides simple intuition for why the lecturers' elasticity is 2--3 times larger than that for assistant/professors in \autoref{tab:facultycount-shock-reg}.
 While no previous research causally connects funding cuts to substituting professors for lecturers, the phenomenon has been alluded to in popular media \citep{wiu2016} and university rankings (where more lecturers hurts a university's ranking, \citealt{usnews2023}).
@@
 An extra 10\% in state funding is associated with 8.5\% more professors hired per students, as seen in \autoref{fig:hiring-correlation}.
-I use these data as part of a shift-share analysis, showing that a 10\% state funding cut per student leads to 13\% fewer professor hires (relative to enrolment), and that these hires are roughly equal for gender of professor hired.
+I use these data as part of a shift-share analysis, showing that a 10\% state funding cut per student leads to 13\% fewer professor hires (relative to enrolment), and that these hires are roughly equal for gender of professor hired.\footnote{
+    However, most of the variation in state funding cuts and substitution towards lecturers happens during the 1990--2009 period.
+    So this analysis of 2011--2021 data is not ground-truth evidence for mechanisms over 1990--2009.
+}
 
@@
 This reflects a key point noted in \autoref{sec:trends}: a large amount of the cuts, and general stagnation, in state funding for public universities occurred over in the 1990s and 2000s.
-Additionally, the trend in rising lecturer count, and falling full professor count, per student at public universities had been operating all throughout the 1990s and 2000s.
+Additionally, the trend in rising lecturer count, and falling professor count, per student at public universities had been operating all throughout the 1990s and 2000s.
 These trends did continue in the 2010s (especially for counts of assistant professors), but most of the substitution had already taken place by this point.

### 2024-09-27 text/sections/08-figures-tables.tex
This is all done.

@@
     The numbers are adjusted to 2021 figures by CPI-U.
-    Lecturer is a binary for whether the faculty member is designated as a lecturer in the databse, and similarly for the assistant professors, full professors, and administrative faculty;
+    Lecturer is a binary for whether the faculty member is designated as a lecturer in the databse, and similarly for the assistant professors, professors, and administrative faculty;
     salary refers the sum of base salary and benefits.

### 2024-09-27 text/sections/09-appendix.tex
Added a number of new things.  2 minor comments to address before sub.
This is all done.

@@
 
-\subsection{IPEDS First Stage}
+\subsection{IPEDS First-Stage}
 \label{sec:appendix-ipeds-firststage}
@@
 \newpage
-\subsection{Illinois IBHED First Stage}
+\subsection{Illinois IBHED First-Stage}
 \label{sec:iv-model-indiv}
@@
     \centering
-    \caption{First Stage Estimates, for State Funding by Funding Shift-Share in IBHED Data.}
+    \caption{First-Stage Estimates, for State Funding by Funding Shift-Share in IBHED Data.}
     \textbf{Panel A: units in \$ per student}
@@
     \textbf{Note}:
-    These tables show the first stage OLS estimates of regression specification \eqref{eqn:secondstage1_indiv}, showing the effect of the funding shift-share on state funding to gauge performance as an instrument.
+    These tables show the first-stage OLS estimates of regression specification \eqref{eqn:secondstage1_indiv}, showing the effect of the funding shift-share on state funding to gauge performance as an instrument.
     Each observation is a professor-year, in the IBHED data, and funding data are merged from IPEDS.
@@
             \hline \\[-1.8ex]
-            & First-stage & Lecturers & Asst. Professors & Full Professors & All Faculty & Observations \\ 
+            & First-stage & Lecturers & Asst. Professors & Professors & All Faculty & Observations \\ 
             \cline{2-7} 
@@
     \textbf{Note}:
-    These tables show the first stage OLS estimates of regressing state funding in years 2012-2019 but the 2011 base share version of the shift-share instrument, without adjusting for student population size.
+    These tables show the first-stage OLS estimates of regressing state funding in years 2012-2019 but the 2011 base share version of the shift-share instrument, without adjusting for student population size.
     Standard errors are clustered at the state-year level, and university $+$ year fixed effects are included where noted.
@@
 \end{table}
-This definition contrasts with the analysis in \autoref{sec:empirics}, and \cite{NBERw23736}, \cite{NBERw23736}.
+This definition contrasts with the analysis in \autoref{sec:empirics}, and \cite{NBERw23736,NBERw27885}.
 All three use a shift-share based in the year 1990, or averaged 1990--1993, analyse outcomes in the 1990s and 2000s, and adjust for student population size.
@@
     \centering
-    \caption{Replication of Second-Stage Estimates, Effects of State Funding on Faculty Counts, in IPEDS Data 2012--2019 \citep{hinrichs2022state}.}
+    \caption{Replication of IV Estimates, Effects of State Funding on Faculty Counts, in IPEDS Data 2012--2019 \citep{hinrichs2022state}.}
     \makebox[\textwidth][c]{\input{tables/facultycount-replicate.tex}}

### 2024-09-27 programs/analysis/ibhed-shock-rawcounts.R

@@ -342,7 +342,7 @@ stargazer(
     add.lines = outcome.means,
     dep.var.caption = "Dependent Variable: Employment Count by Professor Group",
     dep.var.labels = c(
-        "Lecturers", "Asst. Professors", "Full Professors", "All Faculty"),
+        "Lecturers", "Asst. Professors", "Professors", "All Faculty"),
     column.labels = rep(c("OLS", "2SLS"), 4),
     digits = digits.no,
     digits.extra = digits.no,
@@ -474,7 +474,7 @@ stargazer(
     add.lines = outcome.means,
     dep.var.caption = "Dependent Variable: Yearly New Hires by Professor Group",
     dep.var.labels = c(
-        "Lecturers", "Asst. Professors", "Full Professors", "All Faculty"),
+        "Lecturers", "Asst. Professors", "Professors", "All Faculty"),
     column.labels = rep(c("OLS", "2SLS"), 4),
     digits = digits.no,
     digits.extra = digits.no,

### 2024-09-27 programs/analysis/ibhed-shock.R

@@ -524,7 +524,7 @@ stargazer(
     add.lines = outcome.means,
     dep.var.caption = "Dependent Variable: Employment Count",
     dep.var.labels = c(
-        "Lecturers", "Asst. Professors", "Full Professors", "All Faculty"),
+        "Lecturers", "Asst. Professors", "Professors", "All Faculty"),
     column.labels = rep(c("OLS", "2SLS"), 4),
     digits = digits.no,
     digits.extra = digits.no,
@@ -656,7 +656,7 @@ stargazer(
     add.lines = outcome.means,
     dep.var.caption = "Dependent Variable: Employment Count",
     dep.var.labels = c(
-        "Lecturers", "Asst. Professors", "Full Professors", "All Faculty"),
+        "Lecturers", "Asst. Professors", "Professors", "All Faculty"),
     column.labels = rep(c("OLS", "2SLS"), 4),
     digits = digits.no,
     digits.extra = digits.no,

### 2024-09-27 programs/analysis/ipeds-replicate.R

@@ -331,7 +331,7 @@ stargazer(
     dep.var.caption =
         "Dependent Variable: Faculty Count, by Position",
     dep.var.labels = c(
-        "Lecturers", "Asst. Professors", "Full Professors", "All Faculty"),
+        "Lecturers", "Asst. Professors", "Professors", "All Faculty"),
     column.labels = rep(c("OLS", "2SLS"), 4),
     digits = digits.no,
     digits.extra = digits.no,
@@ -461,7 +461,7 @@ stargazer(
     dep.var.caption =
         "Dependent Variable: Log Faculty Count per Students, by Position",
     dep.var.labels = c(
-        "Lecturers", "Asst. Professors", "Full Professors", "All Faculty"),
+        "Lecturers", "Asst. Professors", "Professors", "All Faculty"),
     column.labels = rep(c("OLS", "2SLS"), 4),
     digits = digits.no,
     digits.extra = digits.no,
@@ -591,7 +591,7 @@ stargazer(
     dep.var.caption =
         "Dependent Variable: Log Faculty Count per Students, by Position",
     dep.var.labels = c(
-        "Lecturers", "Asst. Professors", "Full Professors", "All Faculty"),
+        "Lecturers", "Asst. Professors", "Professors", "All Faculty"),
     column.labels = rep(c("OLS", "2SLS"), 4),
     digits = digits.no,
     digits.extra = digits.no,

### 2024-09-27 programs/analysis/ipeds-robustness-checks.R

@@ -386,7 +386,7 @@ stargazer(
     dep.var.caption =
         "Dependent Variable: Log Faculty Count per Students, by Position",
     dep.var.labels = c(
-        "Lecturers", "Asst. Professors", "Full Professors", "All Faculty"),
+        "Lecturers", "Asst. Professors", "Professors", "All Faculty"),
     column.labels = rep(c("OLS", "2SLS"), 4),
     digits = digits.no,
     digits.extra = digits.no,
@@ -566,7 +566,7 @@ stargazer(
     dep.var.caption =
         "Dependent Variable: Faculty Count per 1,000 Students, by Position",
     dep.var.labels = c(
-        "Lecturers", "Asst. Professors", "Full Professors", "All Faculty"),
+        "Lecturers", "Asst. Professors", "Professors", "All Faculty"),
     column.labels = rep(c("OLS", "2SLS"), 4),
     digits = digits.no,
     digits.extra = digits.no,
@@ -770,7 +770,7 @@ stargazer(
     dep.var.caption =
         "Dependent Variable: Log Faculty Count per Students, by Position",
     dep.var.labels = c(
-        "Lecturers", "Asst. Professors", "Full Professors", "All Faculty"),
+        "Lecturers", "Asst. Professors", "Professors", "All Faculty"),
     column.labels = rep(c("OLS", "2SLS"), 4),
     digits = digits.no,
     digits.extra = digits.no,
@@ -895,7 +895,7 @@ stargazer(
     dep.var.caption =
         "Dependent Variable: Faculty Count per 1,000 Students, by Position",
     dep.var.labels = c(
-        "Lecturers", "Asst. Professors", "Full Professors", "All Faculty"),
+        "Lecturers", "Asst. Professors", "Professors", "All Faculty"),
     column.labels = rep(c("OLS", "2SLS"), 4),
     digits = digits.no,
     digits.extra = digits.no,

### 2024-09-27 programs/analysis/ipeds-shock-rawcounts.R

@@ -321,7 +321,7 @@ stargazer(
     dep.var.caption =
         "Dependent Variable: Faculty Count, by Position",
     dep.var.labels = c(
-        "Lecturers", "Asst. Professors", "Full Professors", "All Faculty"),
+        "Lecturers", "Asst. Professors", "Professors", "All Faculty"),
     column.labels = rep(c("OLS", "2SLS"), 4),
     digits = digits.no,
     digits.extra = digits.no,
@@ -376,7 +376,7 @@ substitution_rawcount.plot <-
         coefs = c(
             "Lecturers" = "`lect_funding(fit)`",
             "Asst. Professors" = "`asst_funding(fit)`",
-            "Full Professors" = "`full_funding(fit)`",
+            "Professors" = "`full_funding(fit)`",
             "All Faculty" = "`all_funding(fit)`"),
         legend.title = "",
         plot.distributions = TRUE,

### 2024-09-27 programs/analysis/ipeds-shock.R

@@ -1031,7 +1031,7 @@ stargazer(
     dep.var.caption =
         "Dependent Variable: Log Faculty Count per Students, by Position",
     dep.var.labels = c(
-        "Lecturers", "Asst. Professors", "Full Professors", "All Faculty"),
+        "Lecturers", "Asst. Professors", "Professors", "All Faculty"),
     column.labels = rep(c("OLS", "2SLS"), 4),
     digits = digits.no,
     digits.extra = digits.no,
@@ -1085,7 +1085,7 @@ elasticity.plot <- plot_summs(
     coefs = c(
         "Lecturers" = "`lect_funding(fit)`",
         "Asst. Professors" = "`asst_funding(fit)`",
-        "Full Professors" = "`full_funding(fit)`",
+        "Professors" = "`full_funding(fit)`",
         "All Faculty" = "`all_funding(fit)`"),
     legend.title = "",
     plot.distributions = TRUE,

### 2024-09-27 text/sections/01-introduction.tex
This is all done.

@@
 I provide the first causal evidence on the impact of state funding on faculty composition and wages.
-In a contemporaneous working paper, \cite{hinrichs2022state} use the same national-level variation in university funding to analyse effects on faculty and wages, coming to different conclusions.\footnote{
-    An anonymous reviewer made me aware of a similar and contemporary working paper, \cite{hinrichs2022state}; this work analyses a shorter time-period than this paper, and only focuses on national variation.
-    \autoref{sec:recession-analysis} replicates their key findings, and shows how their focusing on only the 2012--2019 years explains differences in conclusions.
+In a contemporaneous working paper, \cite{hinrichs2022state} use the same national-level variation in university funding to analyse effects on faculty and wages, coming to different conclusions for the national-level analysis.\footnote{
+    An anonymous reviewer made me aware of a similar and contemporary working paper, \cite{hinrichs2022state}.
+    My findings are consistent with theirs when restricting to years 2012--2019, but show a substitution towards lecturers when expanding to the entire time period 1990--2021.
+    See \autoref{sec:recession-analysis} for this replication.
 }

### 2024-09-28 text/sections/02-data.tex
This is all done.

@@
     Notably, IPEDS data on faculty counts do not suffer from these problems.
-}\footnote{
-    In a contemporaneous working paper, \cite{hinrichs2022state} analyses the effect of state funding cuts on university faculty, using the same shift-share variation.
-    \cite{hinrichs2022state} analyses the years 2012--2019, despite national trends in state funding and faculty counts happening largely in the preceding decades (see \autoref{sec:trends,sec:decade-analysis});
-    \cite{hinrichs2022state} also does not mention the data inconsistencies in IPEDS data for average faculty salaries.
-    These two major differences explain the different conclusions that our contemporary papers come to; in \autoref{sec:recession-analysis} I replicate the \cite{hinrichs2022state} approach with my primary specification, and show my results replicate theirs for 2012--2019 (null) and are nonetheless strong for the 1990--2010 time period.
 }

### 2024-11-15 text/paper.tex
minor mistake addressed.  In need of writing overhaul, to be actioned upon.

@@
         Francine Blau,
-        Romain Campbell,
+        Romaine Campbell,
         Maddalena Davoli,

### 2024-11-15 programs/analysis/ipeds-shock-longterm.R

@@ -115,10 +115,10 @@ rm(ipeds.data)
 gc()
 
 
-# Long-run effects, where outcome is Y_{i, t+j} for j = -5, ..., 10 ------------
+# Long-run effects, where outcome is Y_{i, t + k} for k = 0, ..., 10 -----------
 
 # Run the simple linear regression in a loop, and plot the results
-# Make a note of the possible entangled X, Y channel effects, thus need for LP.
+# Make a note of the possibly entangled X, Y channel effects, thus need for LP.
 
 
 # Local Projections for staying-power of effects -------------------------------
@@ -175,13 +175,14 @@ firststage.lpreg <-
         # LP options
         diff_shock = FALSE,
         cumul_mult = FALSE,
-        # Add fixed effects + clsutered SEs in the panel
+        # Add fixed effects + clustered SEs in the panel
         panel_model = "within",
         panel_effect = "twoways",
         robust_cov = "vcovHC",
         robust_cluster = c("group", "time"),
         confint = 1.96,
         hor = time.horizon)
+lpreg.plot(firststage.lpreg, c(-1.15, 0.2), colour.list[1])
 # Save this plot
 ggsave("../../text/figures/firststage-lp.png",
     plot = lpreg.plot(firststage.lpreg, c(-1.15, 0.2), colour.list[1]),
